### 2017-10-11 analysis/fps_sanity_checks.Rmd
add sanity check rmd

---
title: "FPS Sanity Checks"
output: html_notebook
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
```

```{r, message=FALSE, warning=FALSE}
new <- read_csv("../data/new_movie_lengths.csv")
old <- read_csv("../data/old_movie_lengths.csv")
```

```{r}
new %>%
  rename(new_fps = fps, new_length = length) %>%
  inner_join(old %>%
              mutate(old_fps = last.frame / length) %>%
              rename(old_length = length), 
            by = "subid") %>%
  mutate(length_close_enough = abs(new_length - old_length) <= 1) %>%
  mutate(fps_close_enough =  abs(new_fps - old_fps) <= 1) %>%
  kable
```

### 2017-10-11 test_scripts/compare_movie_lengths.py

@@ -1,8 +1,0 @@
-import os
-
-DIR = "/"
-
-if __name__ == "__main__":
-    pass
-
-# compare frames / time / fps between old and new move_lengths csv

### 2017-12-20 scripts/detect_faces.py
change slurm parameters

@@ -19,7 +19,7 @@ def detect_faces(gray_img):
     return faceCascade.detectMultiScale(
         gray_img,
         scaleFactor=1.1,
-        minNeighbors=5,
+        minNeighbors=20,
         minSize=(30, 30)
     )
 
@@ -32,8 +32,7 @@ def rotate_img(image, angle):
     return result
 
 
-def process_img(imgpath, i, arg):
-    # print('batch {0}: {1}%'.format(arg, i/1075.0 * 100))
+def process_img(imgpath):
     print imgpath
     gray_img = cv2.cvtColor(cv2.imread(imgpath), cv2.COLOR_BGR2GRAY)
 
@@ -47,38 +46,38 @@ def process_img(imgpath, i, arg):
     rows = []
 
     if len(faces) == 0:
-        rows.append([group, name, frame, False, None, None, None, None, 0])
-    else:
-        for (x, y, w, h) in faces:
-            rows.append([group, name, frame, True, x, y, w, h, 0])
-
-    if len(faces_rotated) == 0:
         rows.append([group, name, frame, False, None, None, None, None, 180])
     else:
-        for (x, y, w, h) in faces_rotated:
+        for (x, y, w, h) in faces:
             rows.append([group, name, frame, True, x, y, w, h, 180])
+
+    if len(faces_rotated) == 0:
+        rows.append([group, name, frame, False, None, None, None, None, 0])
+    else:
+        for (x, y, w, h) in faces_rotated:
+            rows.append([group, name, frame, True, x, y, w, h, 0])
 
     return rows
 
 
 if __name__ == "__main__":
-    img_index = int(sys.argv[1]) * 1075
+    img_index = int(sys.argv[1]) * 4171
 
     with open(IMAGE_PATHS_FILE, 'rb') as img_paths_f:
         img_paths = list(csv.reader(img_paths_f, delimiter=','))
-        images_to_process = img_paths[img_index-1074:img_index+1]
+        images_to_process = img_paths[img_index-4170:img_index+1]
 
     pool = multiprocessing.Pool()
     results = []
 
-    for i, row in enumerate(images_to_process):
-        results.append(pool.apply_async(process_img, args=(row[0], i, sys.argv[1])))
+    for row in images_to_process:
+        results.append(pool.apply_async(process_img, args=(row[0],)))
 
     pool.close()
 
[…]

### 2017-12-20 scripts/wrap.sh

@@ -1,5 +1,5 @@
 #!/bin/sh
-for i in $(seq 679 $END); 
-do sbatch -p normal,hns -t 40:00 -c 16 --mail-type=FAIL --mail-user=sanchez7@stanford.edu --wrap="python detect_faces.py $i"
+for i in $(seq 175 $END);
+do sbatch -p normal,hns -t 40:00 -c 8 --mail-type=FAIL --mail-user=sanchez7@stanford.edu --wrap="python detect_faces.py $i"
 sleep 1 
 done

### 2017-12-20 analysis/face_detection_analysis.Rmd

---
title: "XS-FACE preliminary analysis"
output: html_notebook
---

Imports
```{r, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
```

Load CSV
```{r}
data <- read_csv('../data/face_detections.csv')
```

Summarize
```{r}
data %>%
  distinct(group, video, frame, is_face, angle) %>% # ignore > 1 face detections
  group_by(group, angle) %>%
  summarise(face_frames=sum(is_face=="True"), total=n()) %>%
  mutate(percentage=face_frames/total * 100) %>%
  arrange(angle, group) %>%
  kable
```

### 2018-01-03 analysis/face_detection_analysis.Rmd
random edits to analysis scripts

@@
 Load CSV
-```{r}
-data <- read_csv('../data/face_detections.csv')
+```{r, message=FALSE, warning=FALSE}
+data <- read_csv('/Users/alessandro/Github/xs-face/data/3.csv')
 ```
@@
 data %>%
+  #filter(w>100 | is.na(w)) %>%
   distinct(group, video, frame, is_face, angle) %>% # ignore > 1 face detections
@@
 ```
+
+```{r}
+b <- data %>%
+  filter(video=="XS_0801", angle==0, is_face=="True") %>%
+  arrange(frame)
+
+```
+
+```{r}
+a <- data %>% filter(angle ==0, is_face=="True") %>% sample_n(250)
+b <- data %>% filter(angle ==0, is_face=="False") %>% sample_n(250)
+c <- rbind(a, b)
+c %>% mutate(path = paste0("/share/PI/mcfrank/frames/", group, "-months/", video, "/image-",frame,".jpg"))
+write_csv(c, "sample.csv")
+```
+
+```{r}
+val <- inner_join(c,gold, by = c("group", "video", "frame", "angle"))
+
+```
+
+```{r}
+func <- function(a, b) {
+  if (a == "True") {
+    if (a == b) return ("TP")
+    else return("FP")
+  }
+  else {
+    if (a == b) return("TN")
+    else return("FN")
+  }
+}
+```
+
+```{r}
+answers <- val %>%
+  rowwise() %>%
+  mutate(ans = func(is_face.x, is_face.y))
+```
+
+```{r}
+answers %>% 
+  filter(ans == "TP") %>%
+  nrow()
+```
+
+```{r}
+answers %>% 
+  filter(group == "8", ans == "FP") %>%
+  nrow()
+```
+
+```{r}
+answers %>% 
+  filter(ans == "TN") %>%
+  nrow()
+```
+
+```{r}
+answers %>% 
+  filter(ans == "FN") %>%
+  nrow()
+```
+
+```{r}
+b
+table(diff(as.integer(distinct(a, frame)$frame)))
+```
+
+```{r}
+diff(as.integer((b %>% filter(w < 100) %>% distinct(frame))$frame)) %>% table
+```
+
+```{r}
+d8 <- data %>% 
+  filter(group==8) %>%
+  distinct(group, video, frame, is_face, angle) %>% 
+  arrange(video, frame)
+b
+d8a <- d8 %>% 
+  filter(angle==180)
+
+d8b <- d8 %>%
+  filter(angle==0)
+
+hist(h$area, xlim=c(10, 40000), binwidth=10)
+```

### 2018-01-03 scripts/download.py

@@ -1,0 +1,10 @@
+import os
+import csv
+
+with open("../sample.csv", 'rb') as img_paths_f:
+        img_paths = list(csv.reader(img_paths_f, delimiter=','))
+
+for path in img_paths[1:]:
+	os.system("mkdir -p ~/xsface/sample_images/%s/%s" % (path[0], path[1]))
+	os.system("scp sanchez7@sherlock.stanford.edu:%s ~/xsface/sample_images/%s/%s" % (path[9], path[0], path[1]))
+

### 2018-01-03 scripts/validate.py

@@ -1,0 +1,38 @@
+import os
+import csv
+import ntpath
+
+BRIA_ANNOTATIONS = "/Users/alessandro/Downloads/sample_images_bria"
+
+if __name__ == "__main__":
+
+    rows = {}
+
+    for root, subdirectory, filenames in os.walk(BRIA_ANNOTATIONS):
+
+        if root[len(root) - 1].isdigit(): # inside video directory
+            for filename in filenames:
+                if filename == ".DS_Store": continue
+                group = ntpath.basename(os.path.dirname(root))
+                video = ntpath.basename(root)
+                frame = filename.split(".")[0].split("-")[1]
+
+                rows[filename.split(".")[0]] = [group, video, frame, "False", 0]
+
+        elif ntpath.basename(root) == "annotations": #inside annotations
+            for filename in filenames:
+                arr = rows[filename.split(".")[0]]
+                arr[3] = "True"
+                rows[filename.split(".")[0]] = arr
+
+    print(len(rows))
+
+    with open("ground_truth.csv", 'wb') as csvfile:
+        wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
+        wr.writerow(['group', 'video', 'frame', 'is_face', 'angle'])
+        for key, val in rows.iteritems():
+            wr.writerow(val)
+
+
+
+

### 2018-01-08 analysis/face_detection_analysis.Rmd
updated csv

@@
 ```{r, message=FALSE, warning=FALSE}
-data <- read_csv('/Users/alessandro/Github/xs-face/data/3.csv')
+data1 <- read_csv('/Users/alessandro/workspace/3.csv')
+data <- read_csv("../data/final_output/mtcnn_face_detections.csv")
+```
+
+```{r}
+setwd("..")
 ```
@@
 ```{r}
-data %>%
+data1 %>%
   #filter(w>100 | is.na(w)) %>%
-  distinct(group, video, frame, is_face, angle) %>% # ignore > 1 face detections
-  group_by(group, angle) %>%
+  distinct(group, video, frame, is_face) %>% # ignore > 1 face detections
+  group_by(group) %>%
   summarise(face_frames=sum(is_face=="True"), total=n()) %>%
   mutate(percentage=face_frames/total * 100) %>%
-  arrange(angle, group) %>%
+  arrange(group) %>%
   kable

### 2018-01-08 analysis/merge_all_data.R
copied other data & main anlayses script from original xsface

@@ -1,0 +1,42 @@
+## main data merge script - prerequisite for all other analyses
+rm(list=ls())
+library(dplyr)
+library(readr)
+library(magrittr)
+library(tidyr)
+source("helper.R")
+
+## load detectors, demo data, merge
+dets <- read_csv("../data/all_detectors.csv") %>%
+  mutate(y = as.numeric(y)) # because of leading zeros, apparently
+demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
+  select(-ra, -assist, -len) ## -dot -dib fields did not exist, deleted bll
+d <- dets %>%
+  left_join(demo.data) %>%
+  group_by(subid) %>%
+  do(add.times(.)) %>% 
+  group_by(subid) %>%
+  do(add.posture(.))
+
+
+# Complete the data frame so that zeros get counted: expands so that each subid
+# includes a zero length row for each posture and orientation.
+# this was annoying.
+ages <- d %>%
+  group_by(subid) %>%
+  summarise(age.grp = mean(age.grp))
+
+complete_combos <- expand(d, nesting(posture, orientation), 
+                          subid) %>% 
+  mutate(dt = 0, face = FALSE) %>%
+  left_join(ages)
+
+d <- bind_rows(d, complete_combos)
+
+write_csv(d, "../data/consolidated_data.csv")
+
+## Create a set of videos that passes the sanity checks in length_sanity_check.Rmd
+d_passing <- filter(d, not(subid %in% c("XS_1205","XS_1211","XS_1225",
+                                        "XS_1628","XS_1639")))
+
+write_csv(d_passing, "../data/consolidated_data_passing.csv")

### 2018-01-08 analysis/helper.R
fixed helper functions to import mtcnn2 input

@@ -1,0 +1,178 @@
+library(stringr)
+
+######## MERGE IN EXACT TIMES FOR FRAMES #########
+add.times <- function(x) {
+  print(x$subid[1])
+  fname <- paste0("../data/frame_times/",
+                 str_sub(as.character(x$subid[1]),start=4,end=8),
+                 ".csv")
+  if (file.exists(fname)) {
+    times <- read_csv(fname)
+#     x <- merge(x,times, by.x = "frame", by.y = "frame", all.x = TRUE, all.y = FALSE)
+    x <- left_join(x, times)
+
+    x$dt <- c(diff(x$time),.032)
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing frames ****"))
+    x$time <- NA
+    x$dt <- NA
+  }
+  
+  return(x)
+}
+
+######## ADD POSTURE CODING #########
+add.posture <- function(x) {
+  fname <- paste("../data/posture/",
+                 x$subid[1],
+                 ".csv",
+                 sep="")
+  
+  if (file.exists(fname)) {
+    print(fname)
+    postures <- read_csv(fname)
+    
+    postures <- regularize.postures(subset(postures,code=="posture"))
+    
+    x$posture <- factor(NA,levels=levels(postures$posture))
+    x$orientation <- factor(NA,levels=levels(postures$orientation))
+    
+    # for each row of p, populate posture and orientation to x
+    for (i in 1:nrow(postures)) {
+      range <- x$time > postures$start[i] & x$time <= postures$end[i]
+      x$posture[range] <- postures$posture[i]
+      x$orientation[range] <- postures$orientation[i]
+    }
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing posture ****"))
+    x$posture <- NA
+    x$orientation <- NA
+  }
+  
+  return(x)
+}
+
+######## GET POSTURE CODING SORTED OUT #########
+regularize.postures <- function (p) {
+  original.postures <- c("p","si","st","l","c","w","NIF")
+  
+  # clean up postures
+  p$type <- factor(p$type,levels=original.postures)
+  p$type <- plyr::revalue(p$type,c("p"="prone",
+                   "si"="sit",
+                   "st"="stand",
+                   "l"="prone",
+                   "c"="carry",
+                   "w"="stand",
+                   "NIF"="NIF"))
[…]

### 2018-01-08 analysis/merge_all_data.R

@@ -7,8 +7,9 @@ library(tidyr)
 source("helper.R")
 
 ## load detectors, demo data, merge
-dets <- read_csv("../data/all_detectors.csv") %>%
+dets <- read_csv("../data/final_output/mtcnn2.csv") %>%
   mutate(y = as.numeric(y)) # because of leading zeros, apparently
+dets$subid <- dets$video # common names
 demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
   select(-ra, -assist, -len) ## -dot -dib fields did not exist, deleted bll
 d <- dets %>%

### 2018-01-08 analysis/mtcnn_evaluation.Rmd
new eval script

---
title: "mtcnn evaluation"
output: html_notebook
---

Load data
```{r}
mtcnn_data <- read_csv("data/final_output/mtcnn2.csv") %>%
  select(-X1)
ground_truth_data <- read_csv("data/ground_truth/ground_truth.csv")
```

Join
```{r}
combined <- mtcnn_data %>%
  rename(mtcnn_is_face = is_face) %>%
  right_join(ground_truth_data, by = c("group", "video", "frame"))
```

Eval function
```{r}
evaluate <- function(a, b) {
  if (a == "True") {
    if (a == b) return ("TP")
    else return("FP")
  }
  else {
    if (a == b) return("TN")
    else return("FN")
  }
}
```

Evaluate
```{r}
results <- combined %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, mtcnn_is_face))
```

Precison / Recall / F-score
```{r}
results %>%
  group_by(result) %>%
  summarise(count = n())
```

### 2018-01-08 analysis/main_analyses.Rmd

---
title: "XSFACE Analyses"
author: "Mike Frank"
date: "July 14, 2016"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---

```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = TRUE)
```

```{r}
rm(list=ls())
library(readr)
library(dplyr)
library(tidyr)
library(langcog)
library(ggplot2)
# library(lubridate)
```

Note that the dataset here is those participants (31) who pass the length sanity checks.

```{r}
d <- read_csv("../data/consolidated_data_passing.csv") %>%
  mutate(posture = factor(posture), 
         orientation = factor(orientation))
demo <- read_csv("../data/demographics/demographics.csv")
#cdis <- read_csv("../data/cdis/cdi_summary.csv") ## not in repo
```

# Postures and Orientations

## Postures

```{r}
ps <- d %>%
  filter(!is.na(posture)) %>%
  group_by(subid, age.grp, posture) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(age.grp, posture) %>%
  multi_boot_standard(col = "prop.time", na.rm=TRUE)
```

```{r}
ggplot(ps, aes(x = factor(age.grp), y = mean, 
               fill = posture, pch = posture)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), 
                 position = position_dodge(width = .9)) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  scale_fill_solarized() + 
  theme_bw()
```

## Orientations

```{r}
os <- d %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, subid, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(age.grp, orientation) %>%
  multi_boot_standard(col = "prop.time")
```

```{r}
ggplot(os, aes(x = factor(age.grp), y = mean, 
               fill = orientation, pch = orientation)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), 
                 position = position_dodge(width = .9)) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  scale_fill_solarized() + 
  theme_bw()
```

## Postures and orientations together

Data don't look interpretable typically when sliced this fine.

```{r}
pos <- d %>%
  filter(!is.na(orientation) & !is.na(posture)) %>%
  group_by(age.grp, subid, orientation, posture) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  group_by(age.grp, subid) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(age.grp, orientation, posture) %>%
  multi_boot_standard(col = "prop.time")
```

```{r}
ggplot(pos, aes(x = factor(age.grp), y = mean, 
               fill = posture)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), 
                 position = position_dodge(width = .9)) +
  facet_grid(.~ orientation) + 
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  scale_fill_solarized() + 
  theme_bw()
```

# Face detections

Using automated face detection results.

## Basic detections descriptives

```{r}
mss <- d %>%
  group_by(age.grp, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

ms <- mss %>%
  multi_boot_standard(col = "face")

ggplot(mss, aes(x = age.grp, y = face)) + 
  geom_jitter(width = .25) + 
  geom_pointrange(data = ms, 
                 aes(x = age.grp, 
                     y = mean, 
                     ymin = ci_lower, 
                     ymax = ci_upper), col = "red") +
  geom_line(data = ms, 
             aes(x = age.grp, 
                     y = mean), col = "red") + 
  scale_x_continuous(breaks=c(8,12,16)) +
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

And with continuous age. This is probably the simplest option.

```{r}
mss <- d %>%
  filter(!is.na(age.at.test)) %>%
  group_by(subid, age.at.test) %>%
  summarise(face = mean(face)) %>%
  left_join(demo) %>%
  mutate(len = as.numeric(lubridate::as.duration(lubridate::ms(len))))

ggplot(mss, aes(x = age.at.test, y = face)) + 
  geom_jitter(width = .25, aes(size = len)) + 
  geom_smooth() + 
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_size_continuous(name = "Length (s)") + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

Boxplot, though it's not as good.

```{r}
ggplot(mss, aes(x = factor(age.grp), y = face)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

Stats. Standard ANOVA (violates distributional assumptions).

```{r}
mod <- lm(face ~ factor(age.grp), data = mss)
summary(mod)
anova(mod)
```

Now a non-parametric test.

```{r}
kruskal.test(face ~ factor(age.grp), data = mss)
```

No real signal that the (predicted) U-shaped pattern is reliable, unfortunately.

## Detections timecourse

Do a 10s moving window average.

```{r}
d <- d %>%
  group_by(subid) %>%
  mutate(face_avg = zoo::rollapply(data = face, 
                                   width = 30*10, # 10s window
                                   FUN = mean, 
                                   align = "right", 
                                   fill = NA, 
                                   na.rm = TRUE))

ggplot(d, aes(x = frame, y = face_avg)) + 
  geom_line() + 
  facet_wrap(~subid) + 
  theme_bw()
```

# Detections by posture and orientation

## Posture

```{r}
mss <- d %>%
  filter(!is.na(posture)) %>%
  group_by(age.grp, posture, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

# ms <- mss %>%
#   multi_boot_standard(col = "face")

ggplot(mss, aes(x = factor(age.grp), y = face, col = posture)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

## Orientation

```{r}
mss <- d %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, orientation, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

# ms <- mss %>%
#   multi_boot_standard(col = "face")

ggplot(mss, aes(x = factor(age.grp), y = face, col = orientation)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

## Both

Not really enough data to say much when binned this finely.

```{r}
mss <- d %>%
  filter(!is.na(posture) & !is.na(orientation)) %>%
  group_by(age.grp, posture, orientation, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

# ms <- mss %>%
#   multi_boot_standard(col = "face")

ggplot(mss, aes(x = factor(age.grp), y = face, col = posture)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  facet_grid(~orientation) + 
  ylim(c(0,.2)) + 
  theme_bw() 
```

# Naming data

Need some helper functions to merge in namings.

```{r}
source("helper.R")
```

```{r}
namings <- d %>%
  group_by(subid) %>%
  do(summarize.naming(.)) %>%
  left_join(demo %>% select(subid, gender, age.at.test, len))
```

First examine: how many namings do we have, generally?

```{r}
naming_rate <- namings %>%
  group_by(subid, age.grp, age.at.test) %>%
  summarise(num = n(), 
            len = len[1]) %>%
  mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60))

ggplot(naming_rate, 
       aes(x = age.at.test, y = rate)) +
  geom_point() + 
  geom_smooth(span = 1) + 
  ylab("Namings per minute") + 
  xlab("Age (months)") + 
  theme_bw() 
```

Break down by novel and familiar.

```{r}
naming_rate_familiarity <- namings %>%
  group_by(subid, familiarity, age.at.test) %>%
  summarise(num = n(), 
            len = len[1]) %>%
  mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60))

ggplot(naming_rate_familiarity, 
       aes(x = age.at.test, y = rate, col = familiarity)) +
  geom_point() + 
  geom_smooth(span = 1) + 
  ylab("Namings per minute") + 
  xlab("Age (months)") + 
  theme_bw() 
```

Item effects just for fun.

```{r}
namings %>%
  group_by(subid, name, familiarity, age.at.test) %>%
  summarise(num = n(), 
            len = len[1]) %>%
  mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60)) %>%
  ggplot(aes(x = age.at.test, y = rate, col = name)) +
  geom_point() + 
  facet_grid(.~familiarity) + 
  geom_smooth(span = 1) + 
  ylab("Namings per minute") + 
  xlab("Age (months)") + 
  theme_bw() 
```

# Namings and face detections

## Are there more faces during naming instances?

## Are there more faces before or after naming instances?

```{r}
#   sb <- ddply(d,~subid, summarize.naming, window=c(-3,0))
#   sa <- ddply(d,~subid, summarize.naming, window=c(0,3))
#   sb$window <- "before"
#   sa$window <- "after"
#   s <- rbind(sb,sa)
#   s$window <- factor(s$window)
#   
# #   d <- subset(d,!is.na(time))
# #   s <- ddply(d,~subid, summarize.naming, window=c(-3,3))
#   
#   s$first.instance <- s$naming.instance == 1
#   s$log.instance <- round(log10(s$naming.instance)*4)/4
#   s$binned.instance <- 10^(round(log10(s$naming.instance)*4)/4)
#   s$face.logical <- s$face != 0
#   
#   ## try plotting individuals
#   mss <- ddply(s,~subid+age.grp+window,summarise,
#                face = na.mean(face))
```

```{r}
  # qplot(age.grp,face,colour=window,
  #       xlab="Age (Months)",ylab="Proportion faces detected in window",
  #       data=mss,geom=c("point")) + 
  #   scale_x_continuous(breaks=c(8,12,16)) +
  #   geom_smooth()
```

## How do face detections change with sequence of naming instances?

```{r}
ms <- namings %>%
  group_by(age.grp, subid, naming.instance, familiarity, name) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp, naming.instance, familiarity) %>%
  multi_boot_standard(col = "face", na.rm = TRUE)
```

```{r}
ggplot(ms, 
       aes(x = naming.instance, y = mean, col = familiarity)) +
  geom_point() + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  geom_smooth(se = FALSE) + 
  facet_grid(~age.grp) + 
  xlim(c(0,20)) + 
  scale_color_solarized() + 
  theme_bw()

```

```{r}

#   ## now calculate differences from base rate
#   n <- ddply(s,~subid+age.grp,summarise,
#              face = na.mean(face))
#   
#   n$br <- ddply(d,~subid+age.grp,summarise,
#               face=na.mean(face))$face
#   
#   n$br.diff <- n$face - n$br
#   
# #   pdf("~/Projects/xsface/writeup/figures/naming_faces_diff.pdf",width=4,height=3)
#   qplot(age.grp,br.diff,position=position_jitter(.2),
#         data=subset(n,age.grp>4),xlab="Age (Months)",ylab="Face detections relative to base rate") + 
#     scale_x_continuous(breaks=c(8,12,16,20)) +
#     geom_abline(slope=0,intercept=0,lty=2) + 
#     stat_summary(fun.data="mean_cl_boot",color="red")
# #   dev.off()
```

# CDIs, namings, and face detections

```{r}
cdi_namings <- left_join(naming_rate, cdis)

qplot(comprehension, facets= .~ age.grp, data = cdi_namings) + 
  theme_bw()

qplot(production, facets= .~ age.grp, data = cdi_namings) + 
  theme_bw()

```

## CDIs and namings

Just out of curiousity, are there any major differences in terms of CDIs by naming rate? (Would help to establish the validity of the naming rate analyses).

```{r}
ggplot(cdi_namings, 
       aes(x = rate, y = comprehension)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_grid(.~age.grp) + 
  xlab("Namings per minute") + 
  ylab("CDI Summed Comprehension Vocabulary") + 
  theme_bw()
```

```{r}
cdis <- read_csv("../data/cdis/cdi_summary.csv")
left_join(naming_rate, cdis) %>%
  ggplot(aes(x = rate, y = production)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) + 
  facet_grid(.~age.grp) + 
  xlab("Namings per minute") + 
  ylab("CDI Summed Production Vocabulary") + 
  theme_bw()
```

### 2018-01-08 analysis/main_analyses.Rmd
made main r analyses work with new annotations, rendered notebook

@@
 title: "XSFACE Analyses"
-author: "Mike Frank"
-date: "July 14, 2016"
+author: "Mike Frank, continued Bria Long"
+date: "July 14, 2016; January 8, 2018"
 output:
@@
 ```{r}
+d$face_num<-as.numeric(d$is_face=="TRUE")
+
 mss <- d %>%
   group_by(age.grp, subid) %>%
-  summarise(face = mean(face)) %>%
+  filter(!is.na(face_num)) %>% # filter out NA values
+  summarise(face = mean(face_num)) %>%
   group_by(age.grp)
@@
   filter(!is.na(age.at.test)) %>%
+  filter(!is.na(face_num)) %>%
   group_by(subid, age.at.test) %>%
-  summarise(face = mean(face)) %>%
+  summarise(face = mean(face_num)) %>%
   left_join(demo) %>%
@@
   group_by(subid) %>%
-  mutate(face_avg = zoo::rollapply(data = face, 
+  filter(!is.na(face_num)) %>%
+  mutate(face_avg = zoo::rollapply(data = face_num, 
                                    width = 30*10, # 10s window
@@
   filter(!is.na(posture)) %>%
+  filter(!is.na(face_num)) %>%
   group_by(age.grp, posture, subid) %>%
-  summarise(face = mean(face)) %>%
+  summarise(face = mean(face_num)) %>%
   group_by(age.grp)
@@
   filter(!is.na(orientation)) %>%
+  filter(!is.na(face_num)) %>%
   group_by(age.grp, orientation, subid) %>%
-  summarise(face = mean(face)) %>%
+  summarise(face = mean(face_num)) %>%
   group_by(age.grp)
@@
 mss <- d %>%
-  filter(!is.na(posture) & !is.na(orientation)) %>%
+  filter(!is.na(posture) & !is.na(orientation) & !is.na(face_num)) %>%
   group_by(age.grp, posture, orientation, subid) %>%
-  summarise(face = mean(face)) %>%
+  summarise(face = mean(face_num)) %>%
   group_by(age.grp)
@@
 
-# Naming data
+<!-- # Naming data -->
 
-Need some helper functions to merge in namings.
+<!-- Need some helper functions to merge in namings.  -->
 
-```{r}
-source("helper.R")
-```
+<!-- ```{r} -->
+<!-- source("helper.R") -->
+<!-- ``` -->
 
-```{r}
-namings <- d %>%
-  group_by(subid) %>%
-  do(summarize.naming(.)) %>%
-  left_join(demo %>% select(subid, gender, age.at.test, len))
-```
+<!-- ```{r} -->
+<!-- namings <- d %>% -->
+<!--   group_by(subid) %>% -->
+<!--   do(summarize.naming(.)) %>% -->
+<!--   left_join(demo %>% select(subid, gender, age.at.test, len)) -->
+<!-- ``` -->
 
-First examine: how many namings do we have, generally?
+<!-- First examine: how many namings do we have, generally?  -->
 
-```{r}
-naming_rate <- namings %>%
-  group_by(subid, age.grp, age.at.test) %>%
-  summarise(num = n(), 
-            len = len[1]) %>%
-  mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60))
+<!-- ```{r} -->
+<!-- naming_rate <- namings %>% -->
+<!--   group_by(subid, age.grp, age.at.test) %>% -->
+<!--   summarise(num = n(),  -->
+<!--             len = len[1]) %>% -->
+<!--   mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60))  -->
 
-ggplot(naming_rate, 
-       aes(x = age.at.test, y = rate)) +
-  geom_point() + 
-  geom_smooth(span = 1) + 
-  ylab("Namings per minute") + 
-  xlab("Age (months)") + 
-  theme_bw() 
-```
+<!-- ggplot(naming_rate,  -->
+<!--        aes(x = age.at.test, y = rate)) + -->
+<!--   geom_point() +  -->
+<!--   geom_smooth(span = 1) +  -->
+<!--   ylab("Namings per minute") +  -->
+<!--   xlab("Age (months)") +  -->
+<!--   theme_bw()  -->
+<!-- ``` -->
 
-Break down by novel and familiar.
+<!-- Break down by novel and familiar.  -->
 
-```{r}
-naming_rate_familiarity <- namings %>%
-  group_by(subid, familiarity, age.at.test) %>%
-  summarise(num = n(), 
-            len = len[1]) %>%
-  mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60))
+<!-- ```{r} -->
+<!-- naming_rate_familiarity <- namings %>% -->
+<!--   group_by(subid, familiarity, age.at.test) %>% -->
+<!--   summarise(num = n(),  -->
+<!--             len = len[1]) %>% -->
+<!--   mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60)) -->
 
-ggplot(naming_rate_familiarity, 
-       aes(x = age.at.test, y = rate, col = familiarity)) +
-  geom_point() + 
-  geom_smooth(span = 1) + 
-  ylab("Namings per minute") + 
-  xlab("Age (months)") + 
-  theme_bw() 
-```
+<!-- ggplot(naming_rate_familiarity,  -->
+<!--        aes(x = age.at.test, y = rate, col = familiarity)) + -->
+<!--   geom_point() +  -->
+<!--   geom_smooth(span = 1) +  -->
+<!--   ylab("Namings per minute") +  -->
+<!--   xlab("Age (months)") +  -->
+<!--   theme_bw()  -->
+<!-- ``` -->
 
-Item effects just for fun.
+<!-- Item effects just for fun.  -->
 
-```{r}
-namings %>%
-  group_by(subid, name, familiarity, age.at.test) %>%
-  summarise(num = n(), 
-            len = len[1]) %>%
-  mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60)) %>%
-  ggplot(aes(x = age.at.test, y = rate, col = name)) +
-  geom_point() + 
-  facet_grid(.~familiarity) + 
-  geom_smooth(span = 1) + 
-  ylab("Namings per minute") + 
-  xlab("Age (months)") + 
-  theme_bw() 
-```
+<!-- ```{r} -->
+<!-- namings %>% -->
+<!--   group_by(subid, name, familiarity, age.at.test) %>% -->
+<!--   summarise(num = n(),  -->
+<!--             len = len[1]) %>% -->
+<!--   mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60)) %>% -->
+<!--   ggplot(aes(x = age.at.test, y = rate, col = name)) + -->
+<!--   geom_point() +  -->
+<!--   facet_grid(.~familiarity) +  -->
+<!--   geom_smooth(span = 1) +  -->
+<!--   ylab("Namings per minute") +  -->
+<!--   xlab("Age (months)") +  -->
+<!--   theme_bw()  -->
+<!-- ``` -->
 
-# Namings and face detections
+<!-- # Namings and face detections -->
 
-## Are there more faces during naming instances?
+<!-- ## Are there more faces during naming instances?  -->
 
-## Are there more faces before or after naming instances?
+<!-- ## Are there more faces before or after naming instances?  -->
 
-```{r}
-#   sb <- ddply(d,~subid, summarize.naming, window=c(-3,0))
-#   sa <- ddply(d,~subid, summarize.naming, window=c(0,3))
-#   sb$window <- "before"
-#   sa$window <- "after"
-#   s <- rbind(sb,sa)
-#   s$window <- factor(s$window)
-#   
-# #   d <- subset(d,!is.na(time))
-# #   s <- ddply(d,~subid, summarize.naming, window=c(-3,3))
-#   
-#   s$first.instance <- s$naming.instance == 1
-#   s$log.instance <- round(log10(s$naming.instance)*4)/4
-#   s$binned.instance <- 10^(round(log10(s$naming.instance)*4)/4)
-#   s$face.logical <- s$face != 0
-#   
-#   ## try plotting individuals
-#   mss <- ddply(s,~subid+age.grp+window,summarise,
-#                face = na.mean(face))
-```
+<!-- ```{r} -->
+<!-- #   sb <- ddply(d,~subid, summarize.naming, window=c(-3,0)) -->
+<!-- #   sa <- ddply(d,~subid, summarize.naming, window=c(0,3)) -->
+<!-- #   sb$window <- "before" -->
+<!-- #   sa$window <- "after" -->
+<!-- #   s <- rbind(sb,sa) -->
+<!-- #   s$window <- factor(s$window) -->
+<!-- #    -->
+<!-- # #   d <- subset(d,!is.na(time)) -->
+<!-- # #   s <- ddply(d,~subid, summarize.naming, window=c(-3,3)) -->
+<!-- #    -->
+<!-- #   s$first.instance <- s$naming.instance == 1 -->
+<!-- #   s$log.instance <- round(log10(s$naming.instance)*4)/4 -->
+<!-- #   s$binned.instance <- 10^(round(log10(s$naming.instance)*4)/4) -->
+<!-- #   s$face.logical <- s$face != 0 -->
+<!-- #    -->
+<!-- #   ## try plotting individuals -->
+<!-- #   mss <- ddply(s,~subid+age.grp+window,summarise, -->
+<!-- #                face = na.mean(face)) -->
+<!-- ```   -->
 
-```{r}
-  # qplot(age.grp,face,colour=window,
-  #       xlab="Age (Months)",ylab="Proportion faces detected in window",
-  #       data=mss,geom=c("point")) + 
-  #   scale_x_continuous(breaks=c(8,12,16)) +
-  #   geom_smooth()
-```
+<!-- ```{r} -->
+<!--   # qplot(age.grp,face,colour=window, -->
+<!--   #       xlab="Age (Months)",ylab="Proportion faces detected in window", -->
+<!--   #       data=mss,geom=c("point")) +  -->
+<!--   #   scale_x_continuous(breaks=c(8,12,16)) + -->
+<!--   #   geom_smooth() -->
+<!-- ``` -->
 
-## How do face detections change with sequence of naming instances?
+<!-- ## How do face detections change with sequence of naming instances? -->
 
-```{r}
-ms <- namings %>%
-  group_by(age.grp, subid, naming.instance, familiarity, name) %>%
-  summarise(face = mean(face)) %>%
-  group_by(age.grp, naming.instance, familiarity) %>%
-  multi_boot_standard(col = "face", na.rm = TRUE)
-```
+<!-- ```{r} -->
+<!-- ms <- namings %>% -->
+<!--   group_by(age.grp, subid, naming.instance, familiarity, name) %>% -->
+<!--   summarise(face = mean(face)) %>% -->
+<!--   group_by(age.grp, naming.instance, familiarity) %>% -->
+<!--   multi_boot_standard(col = "face", na.rm = TRUE) -->
+<!-- ``` -->
 
-```{r}
-ggplot(ms, 
-       aes(x = naming.instance, y = mean, col = familiarity)) +
-  geom_point() + 
-  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + 
-  geom_smooth(se = FALSE) + 
-  facet_grid(~age.grp) + 
-  xlim(c(0,20)) + 
-  scale_color_solarized() + 
-  theme_bw()
+<!-- ```{r} -->
+<!-- ggplot(ms,  -->
+<!--        aes(x = naming.instance, y = mean, col = familiarity)) + -->
+<!--   geom_point() +  -->
+<!--   geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +  -->
+<!--   geom_smooth(se = FALSE) +  -->
+<!--   facet_grid(~age.grp) +  -->
+<!--   xlim(c(0,20)) +  -->
+<!--   scale_color_solarized() +  -->
+<!--   theme_bw() -->
 
-```
+<!-- ``` -->
 
-```{r}
+<!-- ```{r} -->
 
-#   ## now calculate differences from base rate
-#   n <- ddply(s,~subid+age.grp,summarise,
-#              face = na.mean(face))
-#   
-#   n$br <- ddply(d,~subid+age.grp,summarise,
-#               face=na.mean(face))$face
-#   
-#   n$br.diff <- n$face - n$br
-#   
-# #   pdf("~/Projects/xsface/writeup/figures/naming_faces_diff.pdf",width=4,height=3)
-#   qplot(age.grp,br.diff,position=position_jitter(.2),
-#         data=subset(n,age.grp>4),xlab="Age (Months)",ylab="Face detections relative to base rate") + 
-#     scale_x_continuous(breaks=c(8,12,16,20)) +
-#     geom_abline(slope=0,intercept=0,lty=2) + 
-#     stat_summary(fun.data="mean_cl_boot",color="red")
-# #   dev.off()
-```
+<!-- #   ## now calculate differences from base rate -->
+<!-- #   n <- ddply(s,~subid+age.grp,summarise, -->
+<!-- #              face = na.mean(face)) -->
+<!-- #    -->
+<!-- #   n$br <- ddply(d,~subid+age.grp,summarise, -->
+<!-- #               face=na.mean(face))$face -->
+<!-- #    -->
+<!-- #   n$br.diff <- n$face - n$br -->
+<!-- #    -->
+<!-- # #   pdf("~/Projects/xsface/writeup/figures/naming_faces_diff.pdf",width=4,height=3) -->
+<!-- #   qplot(age.grp,br.diff,position=position_jitter(.2), -->
+<!-- #         data=subset(n,age.grp>4),xlab="Age (Months)",ylab="Face detections relative to base rate") +  -->
+<!-- #     scale_x_continuous(breaks=c(8,12,16,20)) + -->
+<!-- #     geom_abline(slope=0,intercept=0,lty=2) +  -->
+<!-- #     stat_summary(fun.data="mean_cl_boot",color="red") -->
+<!-- # #   dev.off() -->
+<!-- ``` -->
 
-# CDIs, namings, and face detections
+<!-- # CDIs, namings, and face detections -->
 
-```{r}
-cdi_namings <- left_join(naming_rate, cdis)
+<!-- ```{r} -->
+<!-- cdi_namings <- left_join(naming_rate, cdis) -->
 
-qplot(comprehension, facets= .~ age.grp, data = cdi_namings) + 
-  theme_bw()
+<!-- qplot(comprehension, facets= .~ age.grp, data = cdi_namings) +  -->
+<!--   theme_bw() -->
 
-qplot(production, facets= .~ age.grp, data = cdi_namings) + 
-  theme_bw()
+<!-- qplot(production, facets= .~ age.grp, data = cdi_namings) +  -->
+<!--   theme_bw() -->
 
-```
+<!-- ``` -->
 
-## CDIs and namings
+<!-- ## CDIs and namings -->
 
-Just out of curiousity, are there any major differences in terms of CDIs by naming rate? (Would help to establish the validity of the naming rate analyses).
+<!-- Just out of curiousity, are there any major differences in terms of CDIs by naming rate? (Would help to establish the validity of the naming rate analyses).  -->
 
-```{r}
-ggplot(cdi_namings, 
-       aes(x = rate, y = comprehension)) + 
-  geom_point() + 
-  geom_smooth(method = "lm", se = FALSE) + 
-  facet_grid(.~age.grp) + 
-  xlab("Namings per minute") + 
-  ylab("CDI Summed Comprehension Vocabulary") + 
-  theme_bw()
-```
+<!-- ```{r} -->
+<!-- ggplot(cdi_namings,  -->
+<!--        aes(x = rate, y = comprehension)) +  -->
+<!--   geom_point() +  -->
+<!--   geom_smooth(method = "lm", se = FALSE) +  -->
+<!--   facet_grid(.~age.grp) +  -->
+<!--   xlab("Namings per minute") +  -->
+<!--   ylab("CDI Summed Comprehension Vocabulary") +  -->
+<!--   theme_bw() -->
+<!-- ``` -->
 
-```{r}
-cdis <- read_csv("../data/cdis/cdi_summary.csv")
-left_join(naming_rate, cdis) %>%
-  ggplot(aes(x = rate, y = production)) + 
-  geom_point() + 
-  geom_smooth(method = "lm", se = FALSE) + 
-  facet_grid(.~age.grp) + 
-  xlab("Namings per minute") + 
-  ylab("CDI Summed Production Vocabulary") + 
-  theme_bw()
-```
+<!-- ```{r} -->
+<!-- cdis <- read_csv("../data/cdis/cdi_summary.csv") -->
+<!-- left_join(naming_rate, cdis) %>% -->
+<!--   ggplot(aes(x = rate, y = production)) +  -->
+<!--   geom_point() +  -->
+<!--   geom_smooth(method = "lm", se = FALSE) +  -->
+<!--   facet_grid(.~age.grp) +  -->
+<!--   xlab("Namings per minute") +  -->
+<!--   ylab("CDI Summed Production Vocabulary") +  -->
+<!--   theme_bw() -->
+<!-- ``` -->

### 2018-01-08 analysis/main_analyses_bll.Rmd
made main r analyses work with new annotations, rendered notebook

--- 2018-01-08 analysis/main_analyses.Rmd
@@
 ```{r}
-d$face_num<-as.numeric(d$is_face=="TRUE")
+d$face_num<-as.numeric(d$is_face=="True")
 
@@
 
-```{r}
+```{r, d}
+
 mss <- d %>%

### 2018-01-08 analysis/merge_all_data.R

@@ -10,6 +10,8 @@ source("helper.R")
 dets <- read_csv("../data/final_output/mtcnn2.csv") %>%
   mutate(y = as.numeric(y)) # because of leading zeros, apparently
 dets$subid <- dets$video # common names
+dets$frame<-as.numeric(dets$frame) # numeric
+
 demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
   select(-ra, -assist, -len) ## -dot -dib fields did not exist, deleted bll
 d <- dets %>%
@@ -18,7 +20,6 @@ d <- dets %>%
   do(add.times(.)) %>% 
   group_by(subid) %>%
   do(add.posture(.))
-
 
 # Complete the data frame so that zeros get counted: expands so that each subid
 # includes a zero length row for each posture and orientation.

### 2018-01-09 analysis-validation/openpose_analysis.Rmd
reorganized analyses & integrated new detector outputs

---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(knitr)
```

```{r}
data <- read_csv("../data/openpose.csv")
```

```{r}
check <- function(nose, l, r) {
  return(!is.na(nose) & !is.na(l) & !is.na(r))
}
```

```{r}
data2 <- data %>%
  #rowwise() %>%
  mutate(is_face = check(Nose, LEye, REye))

#!is.na(Nose) && !is.na(LEye) && !is.na(REye)
```

```{r}
data2 %>%
  #filter(w>100 | is.na(w)) %>%
  #distinct(group, video, frame, is_face) %>% # ignore > 1 face detections
  group_by(group) %>%
  summarise(face_frames=sum(is_face==TRUE), total=n()) %>%
  mutate(percentage=face_frames/total * 100) %>%
  arrange(group) %>%
  kable
```

```
```{r}

```

### 2018-01-09 analysis-main/helper.R

@@ -1,0 +1,178 @@
+library(stringr)
+
+######## MERGE IN EXACT TIMES FOR FRAMES #########
+add.times <- function(x) {
+  print(x$subid[1])
+  fname <- paste0("../data/frame_times/",
+                 str_sub(as.character(x$subid[1]),start=4,end=8),
+                 ".csv")
+  if (file.exists(fname)) {
+    times <- read_csv(fname)
+    
+#     x <- merge(x,times, by.x = "frame", by.y = "frame", all.x = TRUE, all.y = FALSE)
+    x <- left_join(x, times)
+    x$dt <- c(diff(x$time),.032) ## assumes data is sorted by frames! also not sure why .032?
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing frames ****"))
+    x$time <- NA
+    x$dt <- NA
+  }
+  
+  return(x)
+}
+
+######## ADD POSTURE CODING #########
+add.posture <- function(x) {
+  fname <- paste("../data/posture/",
+                 x$subid[1],
+                 ".csv",
+                 sep="")
+  
+  if (file.exists(fname)) {
+    print(fname)
+    postures <- read_csv(fname)
+    
+    postures <- regularize.postures(subset(postures,code=="posture"))
+    
+    x$posture <- factor(NA,levels=levels(postures$posture))
+    x$orientation <- factor(NA,levels=levels(postures$orientation))
+    
+    # for each row of p, populate posture and orientation to x
+    for (i in 1:nrow(postures)) {
+      range <- x$time > postures$start[i] & x$time <= postures$end[i]
+      x$posture[range] <- postures$posture[i]
+      x$orientation[range] <- postures$orientation[i]
+    }
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing posture ****"))
+    x$posture <- NA
+    x$orientation <- NA
+  }
+  
+  return(x)
+}
+
+######## GET POSTURE CODING SORTED OUT #########
+regularize.postures <- function (p) {
+  original.postures <- c("p","si","st","l","c","w","NIF")
+  
+  # clean up postures
+  p$type <- factor(p$type,levels=original.postures)
+  p$type <- plyr::revalue(p$type,c("p"="prone",
+                   "si"="sit",
+                   "st"="stand",
+                   "l"="prone",
+                   "c"="carry",
[…]

### 2018-01-09 analysis-main/merge_all_data.R

@@ -1,0 +1,46 @@
+## main data merge script - prerequisite for all other analyses
+rm(list=ls())
+library(dplyr)
+library(readr)
+library(magrittr)
+library(tidyr)
+source("helper.R")
+
+## load detectors, demo data, merge
+dets <- read_csv("../data/final_output/mtcnn2.csv") %>%
+  mutate(y = as.numeric(y)) # because of leading zeros, apparently
+dets<-arrange(dets,video,frame) # comes out in weird order because of sherlock jobs!
+dets$subid <- dets$video # common names
+dets$frame<-as.numeric(dets$frame) # numeric
+dets$face<-as.logical(dets$is_face) # make logical
+
+demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
+  select(-ra, -assist, -len) ## -dot -dib fields did not exist, deleted bll
+d <- dets %>%
+  left_join(demo.data) %>%
+  group_by(subid) %>%
+  do(add.times(.)) %>% 
+  group_by(subid) %>%
+  do(add.posture(.))
+
+# Complete the data frame so that zeros get counted: expands so that each subid
+# includes a zero length row for each posture and orientation.
+# this was annoying.
+ages <- d %>%
+  group_by(subid) %>%
+  summarise(age.grp = mean(age.grp))
+
+complete_combos <- expand(d, nesting(posture, orientation), 
+                          subid) %>% 
+  mutate(dt = 0, face = FALSE) %>%
+  left_join(ages)
+
+d <- bind_rows(d, complete_combos)
+
+write_csv(d, "../data/consolidated_data.csv")
+
+## Create a set of videos that passes the sanity checks in length_sanity_check.Rmd
+d_passing <- filter(d, not(subid %in% c("XS_1205","XS_1211","XS_1225",
+                                        "XS_1628","XS_1639")))
+
+write_csv(d_passing, "../data/consolidated_data_passing.csv")

### 2018-01-09 analysis/helper.R

@@ -1,178 +1,0 @@
-library(stringr)
-
-######## MERGE IN EXACT TIMES FOR FRAMES #########
-add.times <- function(x) {
-  print(x$subid[1])
-  fname <- paste0("../data/frame_times/",
-                 str_sub(as.character(x$subid[1]),start=4,end=8),
-                 ".csv")
-  if (file.exists(fname)) {
-    times <- read_csv(fname)
-#     x <- merge(x,times, by.x = "frame", by.y = "frame", all.x = TRUE, all.y = FALSE)
-    x <- left_join(x, times)
-
-    x$dt <- c(diff(x$time),.032)
-  } else {
-    print(paste0("**** ", x$subid[1], " is missing frames ****"))
-    x$time <- NA
-    x$dt <- NA
-  }
-  
-  return(x)
-}
-
-######## ADD POSTURE CODING #########
-add.posture <- function(x) {
-  fname <- paste("../data/posture/",
-                 x$subid[1],
-                 ".csv",
-                 sep="")
-  
-  if (file.exists(fname)) {
-    print(fname)
-    postures <- read_csv(fname)
-    
-    postures <- regularize.postures(subset(postures,code=="posture"))
-    
-    x$posture <- factor(NA,levels=levels(postures$posture))
-    x$orientation <- factor(NA,levels=levels(postures$orientation))
-    
-    # for each row of p, populate posture and orientation to x
-    for (i in 1:nrow(postures)) {
-      range <- x$time > postures$start[i] & x$time <= postures$end[i]
-      x$posture[range] <- postures$posture[i]
-      x$orientation[range] <- postures$orientation[i]
-    }
-  } else {
-    print(paste0("**** ", x$subid[1], " is missing posture ****"))
-    x$posture <- NA
-    x$orientation <- NA
-  }
-  
-  return(x)
-}
-
-######## GET POSTURE CODING SORTED OUT #########
-regularize.postures <- function (p) {
-  original.postures <- c("p","si","st","l","c","w","NIF")
-  
-  # clean up postures
-  p$type <- factor(p$type,levels=original.postures)
-  p$type <- plyr::revalue(p$type,c("p"="prone",
-                   "si"="sit",
-                   "st"="stand",
-                   "l"="prone",
-                   "c"="carry",
-                   "w"="stand",
-                   "NIF"="NIF"))
[…]

### 2018-01-09 analysis/merge_all_data.R

@@ -1,44 +1,0 @@
-## main data merge script - prerequisite for all other analyses
-rm(list=ls())
-library(dplyr)
-library(readr)
-library(magrittr)
-library(tidyr)
-source("helper.R")
-
-## load detectors, demo data, merge
-dets <- read_csv("../data/final_output/mtcnn2.csv") %>%
-  mutate(y = as.numeric(y)) # because of leading zeros, apparently
-dets$subid <- dets$video # common names
-dets$frame<-as.numeric(dets$frame) # numeric
-
-demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
-  select(-ra, -assist, -len) ## -dot -dib fields did not exist, deleted bll
-d <- dets %>%
-  left_join(demo.data) %>%
-  group_by(subid) %>%
-  do(add.times(.)) %>% 
-  group_by(subid) %>%
-  do(add.posture(.))
-
-# Complete the data frame so that zeros get counted: expands so that each subid
-# includes a zero length row for each posture and orientation.
-# this was annoying.
-ages <- d %>%
-  group_by(subid) %>%
-  summarise(age.grp = mean(age.grp))
-
-complete_combos <- expand(d, nesting(posture, orientation), 
-                          subid) %>% 
-  mutate(dt = 0, face = FALSE) %>%
-  left_join(ages)
-
-d <- bind_rows(d, complete_combos)
-
-write_csv(d, "../data/consolidated_data.csv")
-
-## Create a set of videos that passes the sanity checks in length_sanity_check.Rmd
-d_passing <- filter(d, not(subid %in% c("XS_1205","XS_1211","XS_1225",
-                                        "XS_1628","XS_1639")))
-
-write_csv(d_passing, "../data/consolidated_data_passing.csv")

### 2018-01-10 analysis-main/main_analyses_bll.Rmd

--- 2018-01-08 analysis/main_analyses.Rmd
@@
 title: "XSFACE Analyses"
-author: "Mike Frank"
-date: "July 14, 2016"
+author: "Mike Frank; continued by Bria Long"
+date: "July 14, 2016; January 8, 2018"
 output:
@@
 knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = TRUE)
+options(error = browser)
 ```
@@
 library(ggplot2)
-# library(lubridate)
+library(lubridate)
 ```
@@
 ```{r}
+
 mss <- d %>%
@@
   group_by(subid) %>%
+  filter(!is.na(face)) %>%
   mutate(face_avg = zoo::rollapply(data = face, 
@@
   filter(!is.na(posture)) %>%
+  filter(!is.na(face)) %>%
   group_by(age.grp, posture, subid) %>%
@@
             len = len[1]) %>%
-  mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60))
+  mutate(rate = num / as.numeric(as.duration(lubridate::hms(len)) / dminutes())) # bll edits here to wrangle dates
 
@@
             len = len[1]) %>%
-  mutate(rate = num / (as.numeric(lubridate::as.duration(lubridate::ms(len)))/60)) %>%
+  mutate(rate = num / as.numeric(as.duration(lubridate::hms(len)) / dminutes())) %>%  # bll edits here to wrangle dates
   ggplot(aes(x = age.at.test, y = rate, col = name)) +
@@
 
-## Are there more faces during naming instances?
-
 ## Are there more faces before or after naming instances?
@@
 ```{r}
-#   sb <- ddply(d,~subid, summarize.naming, window=c(-3,0))
-#   sa <- ddply(d,~subid, summarize.naming, window=c(0,3))
-#   sb$window <- "before"
-#   sa$window <- "after"
-#   s <- rbind(sb,sa)
-#   s$window <- factor(s$window)
-#   
-# #   d <- subset(d,!is.na(time))
-# #   s <- ddply(d,~subid, summarize.naming, window=c(-3,3))
-#   
+library(plyr)
+
+  sb <- ddply(d,~subid, summarize.naming, window=c(-3,0))
+  sa <- ddply(d,~subid, summarize.naming, window=c(0,3))
+  sb$window <- "before"
+  sa$window <- "after"
+  s <- rbind(sb,sa)
+  s$window <- factor(s$window)
+
+  # not used, commented out
   # s$first.instance <- s$naming.instance == 1
@@
   # s$face.logical <- s$face != 0
-#   
-#   ## try plotting individuals
-#   mss <- ddply(s,~subid+age.grp+window,summarise,
-#                face = na.mean(face))
+
+  ## try plotting individuals
+  mss <- ddply(s,~subid+age.grp+window,summarise,
+               face = mean(face, na.rm="TRUE"))
 ```
@@
 ```{r}
-  # qplot(age.grp,face,colour=window,
-  #       xlab="Age (Months)",ylab="Proportion faces detected in window",
-  #       data=mss,geom=c("point")) + 
-  #   scale_x_continuous(breaks=c(8,12,16)) +
-  #   geom_smooth()
+  qplot(age.grp,face,colour=window,
+        xlab="Age (Months)",ylab="Proportion faces detected in window",
+        data=mss,geom=c("point")) +
+    scale_x_continuous(breaks=c(8,12,16)) +
+    geom_smooth()
 ```
 
-## How do face detections change with sequence of naming instances?
+<!-- ## How do face detections change with sequence of naming instances? -->
 
-```{r}
-ms <- namings %>%
-  group_by(age.grp, subid, naming.instance, familiarity, name) %>%
-  summarise(face = mean(face)) %>%
-  group_by(age.grp, naming.instance, familiarity) %>%
-  multi_boot_standard(col = "face", na.rm = TRUE)
-```
+<!-- ```{r} -->
+<!-- ms <- namings %>% -->
+<!--   group_by(age.grp, subid, naming.instance, familiarity, name) %>% -->
+<!--   summarise(face = mean(face)) %>% -->
+<!--   group_by(age.grp, naming.instance, familiarity) %>% -->
+<!--   multi_boot_standard(col = "face", na.rm = TRUE) -->
+<!-- ``` -->
 
-```{r}
-ggplot(ms, 
-       aes(x = naming.instance, y = mean, col = familiarity)) +
-  geom_point() + 
-  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + 
-  geom_smooth(se = FALSE) + 
-  facet_grid(~age.grp) + 
-  xlim(c(0,20)) + 
-  scale_color_solarized() + 
-  theme_bw()
+<!-- ```{r} -->
+<!-- ggplot(ms, -->
+<!--        aes(x = naming.instance, y = mean, col = familiarity)) + -->
+<!--   geom_point() + -->
+<!--   geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + -->
+<!--   geom_smooth(se = FALSE) + -->
+<!--   facet_grid(~age.grp) + -->
+<!--   xlim(c(0,20)) + -->
+<!--   scale_color_solarized() + -->
+<!--   theme_bw() -->
+<!-- ``` -->
 
-```
+## Now calculate face detection differences from base rate
 
@@
 
-#   ## now calculate differences from base rate
-#   n <- ddply(s,~subid+age.grp,summarise,
-#              face = na.mean(face))
-#   
-#   n$br <- ddply(d,~subid+age.grp,summarise,
-#               face=na.mean(face))$face
-#   
-#   n$br.diff <- n$face - n$br
-#   
-# #   pdf("~/Projects/xsface/writeup/figures/naming_faces_diff.pdf",width=4,height=3)
-#   qplot(age.grp,br.diff,position=position_jitter(.2),
-#         data=subset(n,age.grp>4),xlab="Age (Months)",ylab="Face detections relative to base rate") + 
-#     scale_x_continuous(breaks=c(8,12,16,20)) +
-#     geom_abline(slope=0,intercept=0,lty=2) + 
-#     stat_summary(fun.data="mean_cl_boot",color="red")
-# #   dev.off()
+  n <- ddply(s,~subid+age.grp,summarise,
+             face = mean(face, na.rm="TRUE"))
+
+  n$br <- ddply(d,~subid+age.grp,summarise,
+              face=mean(face, na.rm="TRUE"))$face
+
+  n$br.diff <- n$face - n$br
+
+#   pdf("~/Projects/xsface/writeup/figures/naming_faces_diff.pdf",width=4,height=3)
+  qplot(age.grp,br.diff,position=position_jitter(.2),
+        data=subset(n,age.grp>4),xlab="Age (Months)",ylab="Face detections relative to base rate") +
+    scale_x_continuous(breaks=c(8,12,16)) +
+    geom_abline(slope=0,intercept=0,lty=2) +
+    stat_summary(fun.data="mean_cl_boot",color="red")
+#   dev.off()
 ```
 
-# CDIs, namings, and face detections
+<!-- # CDIs, namings, and face detections -->
 
-```{r}
-cdi_namings <- left_join(naming_rate, cdis)
+<!-- ```{r} -->
+<!-- cdi_namings <- left_join(naming_rate, cdis) -->
 
-qplot(comprehension, facets= .~ age.grp, data = cdi_namings) + 
-  theme_bw()
+<!-- qplot(comprehension, facets= .~ age.grp, data = cdi_namings) + -->
+<!--   theme_bw() -->
 
-qplot(production, facets= .~ age.grp, data = cdi_namings) + 
-  theme_bw()
+<!-- qplot(production, facets= .~ age.grp, data = cdi_namings) + -->
+<!--   theme_bw() -->
 
-```
+<!-- ``` -->
 
-## CDIs and namings
+<!-- ## CDIs and namings -->
 
-Just out of curiousity, are there any major differences in terms of CDIs by naming rate? (Would help to establish the validity of the naming rate analyses).
+<!-- Just out of curiousity, are there any major differences in terms of CDIs by naming rate? (Would help to establish the validity of the naming rate analyses). -->
 
-```{r}
-ggplot(cdi_namings, 
-       aes(x = rate, y = comprehension)) + 
-  geom_point() + 
-  geom_smooth(method = "lm", se = FALSE) + 
-  facet_grid(.~age.grp) + 
-  xlab("Namings per minute") + 
-  ylab("CDI Summed Comprehension Vocabulary") + 
-  theme_bw()
-```
+<!-- ```{r} -->
+<!-- ggplot(cdi_namings, -->
+<!--        aes(x = rate, y = comprehension)) + -->
+<!--   geom_point() + -->
+<!--   geom_smooth(method = "lm", se = FALSE) + -->
+<!--   facet_grid(.~age.grp) + -->
+<!--   xlab("Namings per minute") + -->
+<!--   ylab("CDI Summed Comprehension Vocabulary") + -->
+<!--   theme_bw() -->
+<!-- ``` -->
 
-```{r}
-cdis <- read_csv("../data/cdis/cdi_summary.csv")
-left_join(naming_rate, cdis) %>%
-  ggplot(aes(x = rate, y = production)) + 
-  geom_point() + 
-  geom_smooth(method = "lm", se = FALSE) + 
-  facet_grid(.~age.grp) + 
-  xlab("Namings per minute") + 
-  ylab("CDI Summed Production Vocabulary") + 
-  theme_bw()
-```
+<!-- ```{r} -->
+<!-- cdis <- read_csv("../data/cdis/cdi_summary.csv") -->
+<!-- left_join(naming_rate, cdis) %>% -->
+<!--   ggplot(aes(x = rate, y = production)) + -->
+<!--   geom_point() + -->
+<!--   geom_smooth(method = "lm", se = FALSE) + -->
+<!--   facet_grid(.~age.grp) + -->
+<!--   xlab("Namings per minute") + -->
+<!--   ylab("CDI Summed Production Vocabulary") + -->
+<!--   theme_bw() -->
+<!-- ``` -->

### 2018-01-10 analysis-validation/face_detection_analysis.Rmd

--- 2018-01-03 analysis/face_detection_analysis.Rmd
@@
 ```{r, message=FALSE, warning=FALSE}
-data <- read_csv('/Users/alessandro/Github/xs-face/data/3.csv')
+# data1 <- read_csv('/Users/alessandro/workspace/3.csv')
+data <- read_csv("../data/final_output/mtcnn2.csv")
+```
+
+```{r}
+setwd("..")
 ```
@@
   #filter(w>100 | is.na(w)) %>%
-  distinct(group, video, frame, is_face, angle) %>% # ignore > 1 face detections
-  group_by(group, angle) %>%
+  distinct(group, video, frame, is_face) %>% # ignore > 1 face detections
+  group_by(group) %>%
   summarise(face_frames=sum(is_face=="True"), total=n()) %>%
   mutate(percentage=face_frames/total * 100) %>%
-  arrange(angle, group) %>%
+  arrange(group) %>%
   kable
@@
 
-```{r}
-b <- data %>%
-  filter(video=="XS_0801", angle==0, is_face=="True") %>%
-  arrange(frame)
-
-```
+# ```{r} # no "angle" in data
+# b <- data %>%
+#   filter(video=="XS_0801", angle==0, is_face=="True") %>%
+#   arrange(frame) 
+# 
+# ```
 

### 2018-01-10 analysis-validation/frame_validation/check_frame_counts_post_detection.Rmd

---
title: "Check frame counts after mtcnn detection"
output: html_notebook
---

```{r}
library(tidyverse)
library(knitr)
```

Load data
```{r}
ffprobe_frame_counts <- read_csv("data/video_stats/ffprobe_frame_count.csv")
extracted_frame_counts <- read_csv("data/video_stats/extracted_frame_count.csv")
mtcnn_data <- read_csv("data/final_output/mtcnn_face_detections.csv")
# openpose_data 
```

Check ffprobe results against actual extracted frames
```{r}
ffprobe_frame_counts %>%
  all.equal(extracted_frame_counts)
```

Check mtcnn results frame counts against extracted frames
```{r}
mtcnn_data %>%
  distinct(video, frame) %>%
  group_by(video) %>%
  summarise(num_frames = n()) %>%
  all.equal(extracted_frame_counts) 
```

Check openpose results frame counts against extracted frames
```{r}

```

### 2018-01-10 scripts/ground_truth_creation/get_gold_set_images.Rmd
gold set rmd

---
title: "get_gold_set_images"
output: html_document
---

```{r}

```

### 2018-01-10 analysis-validation/length_sanity_check.Rmd
updated demo sheets and sanity check .rmd file for current detectors

---
title: "xsface - file length sanity checking"
author: "Mike Frank"
date: "July 11, 2016"
output:
  html_document:
  toc: true
---

The goal of this document is to document processing steps for the xsface data and ensure that three different datasets match up:

1. source files for movies
2. demographic sheet (containing session lengths)
3. detector outputs

```{r}
rm(list=ls())
library(readr)
library(magrittr)
library(dplyr)
library(lubridate)
```

read in lengths on movies

```{r}
demo <- read_csv("../data/demographics/demographics.csv")
movies <- read_csv("../data/video_stats/new_movie_lengths.csv") 
dets <- read_csv("../data/final_output/mtcnn2.csv")
```

Clean up measurements.

```{r}
demo %<>% 
  rename(recorded_len = len_seconds)

## movie lengths
movies$frameCount=movies$fps*movies$length
movies %<>% 
  rename(last_frame_movie = frameCount, 
         length_movie = length)

# standardize
dets$frame=as.numeric(dets$frame)
dets$subid=dets$video

det.lens <- dets %>%
  group_by(subid) %>%
  summarise(frames_dets = max(frame))
```

Merge.

```{r}
all_data <- left_join(demo, movies) %>%
  left_join(det.lens) %>%
  select(subid, recorded_len, last_frame_movie, length_movie, frames_dets) %>%
  mutate(det_frame_err = abs(last_frame_movie - frames_dets) > 100,
         demo_time_err = abs(recorded_len - length_movie) > 1) %>%
  data.frame
```

Print.

```{r}
knitr::kable(all_data, digits = 2)
```

Explanations of discrepancies:

+ 0825 - unknown, ~500 frames off (15s), spot checks seem to match decently well. 
+ 1205, 1211, and 1225 were post-trimed, hypothesis is that the _QT files are wrappers to source files that trim length to appropriate? 
+ 1628 and 1639 both off by a lot of frames, unknown.

to check in the videos
+ 1229 is reported to be much longer in the demographics file, perhaps this is post-trimming? 
+ 1612 is reported to be shorter than it is, check this. 
+ 1622 is reported to be longer than it is, check this.

### 2018-01-10 writeup/xsface.Rmd
copied writeup files from original xsface

---
title: "Developmental and postural changes in children's visual access to faces"
short-title: "Children's visual access to faces"
output: kmr::apa_manuscript
csl: apa6.csl
bibliography: xsface.bib

document-params: "a4paper,man,apacite,floatsintext,longtable"

bib-tex: "xsface.bib"

author-information:
    - \author{Michael C. Frank, Guido Pusiol, Daniel Yurovsky, Kaia Simmons}

affiliation-information:
    - \affiliation{Department of Psychology, Stanford University}

author-note:
    "Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of the Language and Cognition Lab for help in recruitment, data collection, and annotation. This research was supported by a John Merck Scholars grant to MCF. An earlier version of this work was presented to the Cognitive Science Society in @frank2013. Please address correspondence to Michael C. Frank, Department of Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}."

abstract: 
    "The faces of other people are a critical information source for young children. During early development, children undergo significant postural and locomotor development, changing from lying and sitting infants to toddlers who walk independently. We used a head-mounted camera in conjunction with a face-detection system to explore the eff  ects of these changes on children's visual access to their caregivers' faces during an in-lab play session. In a cross-sectional sample of 8--16 month old children, we found substantial changes in face accessibility based on age and posture. These changes may translate into changes in the accessibility of social information during language learning. We make our corpus available for reanalysis, and discuss strengths and limitations of the headcam method more generally."

keywords:
    "social cognition; face-perception; infancy; locomotion; head-cameras"
---

```{r global_options, include=FALSE}
rm(list=ls())
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
rm(list=ls())
library(readr)
# library(magrittr)
library(dplyr)
library(tidyr)
library(xtable)
# library(lubridate)
```

<!-- % A father offers his young daughter a novel object: a bright yellow feather duster. A few moments after she accepts the toy, he remarks, ``Isn't the zem funny?'' Her father may still be talking about the feather duster, or he may be describing a new object.  To find out she has access to a simple and reliable method: she can look to his face to infer the direction of his attention. -->

The ability to follow social signals like eye-gaze is an important part of early social cognition [@scaife1975;@tomasello2009]

In addition to being an important part of early social development, the ability to follow gaze is a strong predictor of children's early language development. For example, @brooks2005 found that children who followed an experimenter's gaze better before their first birthday had larger vocabularies at 18 months. Similarly, @carpenter1998 found that children's level of joint engagement (as well as the degree to which mothers followed the child's focus of attention in their labeling) predicted vocabulary growth in both language production and comprehension. These studies suggest that children's social environment plays a powerful supportive role in language learning.

But at the same time as children are beginning to learn their first words, their view of the world is changing radically [@adolph2007]. As speechless infants, they are unable to locomote independently. Before their first birthday, they begin crawling; soon after, they begin to walk independently. Infants' visual field is subject to the whims of their caregivers, but caregivers often place them in positions conducive to joint attention. In contrast, toddlers determine their own input to a much greater degree, but as a consequence they spend much of their time in a world primarily populated by knees. These postural and locomotor changes may have a profound effect on what children see.

A recent study suggests the possibility of links between motor milestones, social cognition, and language. @walle2014 noted robust correlations between children's ability to walk and their vocabulary, both receptive and productive. On the basis of an observational study of parent input, they speculated that the emergence of walking may change the ability of the child to access social information (because walking toddlers see more of the social world than crawling infants). Accessing more social information may in turn allow children to discover word meanings more effectively.

Recent methodological developments have the potential to provide data that would allow this hypothesis to be tested. The availability of head-mounted cameras and eye-trackers allows for the measurement of children's naturalistic environment in a way that was not previously possible. @yoshida2008 gave the first demonstration of the radical differences between toddler and adult perspectives on the social world, with toddlers' visual field being dominated by hands and objects much more than that of adults. More recent work has used head-mounted eye-tracking methods to measure young toddlers' fixations [@franchak2011], also finding that children look relatively infrequently at their mothers' faces in naturalistic play.

These methods are now being applied to understand inputs to language acquisition. Work by Yu, Smith, and colleagues suggests that word learning is facilitated when parents and children create moments in which the visual field is dominated by a single object [@smith2011;@yuinpress]. Some data even suggest that young children's restricted viewpoint may be more effective for learning words than the comparable adult perspective [@yurovsky2012]. Together, this body of evidence suggests that measuring infants' perspective---and how it changes in motor development---is a critical part of understanding early language learning.

In the current study we took a developmental approach to understanding the relationship between perspective and access to social information. We recorded head-camera data from a group of infants and children across a broad age range as they played with their caregivers during a brief laboratory visit. We then hand-annotated these data for the child's posture and parents' naming behavior and used face-detection algorithms to measure the frequency of faces in the child's visual field.  The resulting dataset allows us to analyze changes in access to faces according to children's age, posture, and linguistic input.

# Methods

## Participants

```{r echo=FALSE, include=FALSE}
all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
  rename(subid = subID)
inc_subs <- read.csv("../data/demographics/demographics.csv")
all_subs$inc <- all_subs$reason=="included"

inc_subs$len.min <- sapply(strsplit(as.character(inc_subs$len),":"),
       function(x) {
         x <- as.numeric(x)
         x[1]+x[2]/60
       }
)

all_subs %>% 
  group_by(reason) %>%
  summarise(n = n()) %>% 
  kable

# # exclusions by age
# all_subs %>% 
#   group_by(reason, age_group) %>%
#   summarise(n = n()) %>%
#   spread(age_group, n) %>%
#   xtable %>%
#   print(digits = 2)

all_subs %>%
  left_join(inc_subs) %>%
  group_by(age_group) %>%
  summarise(n = sum(inc), 
            percent_included = mean(inc),
            age = mean(age.at.test, na.rm=TRUE), 
            video_length = mean(len.min, na.rm=TRUE), 
            female = mean(gender=="female", na.rm=TRUE)) %>%
  xtable %>%
  print(digits = 2, include.rownames=FALSE)
```

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
Age group & N & \% included & Mean age & Video length (min) & \% female \\ 
  \hline
  8 &  12 & 0.46 & 8.71 & 14.41 & 0.50 \\ 
   12 &  12 & 0.40 & 12.62 & 13.48 & 0.58 \\ 
   16 &  12 & 0.31 & 16.29 & 15.00 & 0.50 \\ 
   \hline
\end{tabular}
\caption{\label{tab:pop} Demographics by age group.}
\end{table}

Our final sample consisted of 36 infants and children, distributed into each of three age groups: 8 months, 12 months, and 16 months. Participants were recruited from the surrounding community via state birth records. Participants had no documented disabilities and were reported to hear at least 80\% English at home. Demographics and exclusion rates are given in Table \ref{tab:pop}.

In order to compile this final sample, we tested a substantially larger group of children, who were excluded for the following reasons: 
`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
`r sum(all_subs$reason=="no CDI")` for missing CDI data,
`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses.

## Head-mounted camera

\begin{figure}
\centering
\includegraphics[width=3in]{images/viewangle.pdf}
\includegraphics[width=2in]{images/headcam_w_fisheye3.jpg}
\caption{\label{fig:headcam} Left: Field of view for three different headcam configurations, with the device we used in the middle. The lowest camera is pictured for comparison, but was not available until after our study was already in progress. Right: The light-weight, low-cost head-mounted camera with fisheye lens.}
\end{figure}

Our head-mounted camera ("headcam") is composed of a small, inexpensive MD80 model camera attached to a soft elastic headband from a camping headlamp. An aftermarket fisheye lens intended for iPhones and other Apple devices is attached to increase view angle. The total cost of each camera is approximately \$60. The camera captures 720x480 pixel images at approx. 25 frames per second, and has battery life of 60--90 minutes. Without the fisheye lens, the viewing angle for the camera is $32^{\circ}$ horizontal by $24^{\circ}$ vertical; with the fisheye, $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left). The device is pictured in Figure \ref{fig:headcam}, right. Instructions for creating the camera are given in detail at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.

The vertical field of view of the camera was considerably smaller than the child's approximate vertical field of view, which---even at 6-7 months---spans around 100--120$^{\circ}$ in the vertical dimension [@mayer1988;@cummings1988]. We were therefore faced with a choice in the orientation of the camera. If we chose a lower or higher orientation, we would be at risk of truncating either the child's own hands and physically proximate objects, or the faces of the adults around the child. Yet if we chose the middle orientation, we would still be at risk of underestimating the proportion of faces viewed by the child. Thus, for the purposes of the current study---measuring visual access to faces---we chose to orient the camera towards the upper part of the visual field. While this orientation decreased our chances of recording the objects being manipulated by the child, it nevertheless allowed us to capture the majority of the faces in the child's visual field.

## Procedure

Before coming to the lab, caregivers completed the MacArthur-Bates Communicative Development Inventories Words and Gestures form [@fenson2007], reporting on the words that they believed that their child understood and produced.

After coming to the lab, families were seated in our waiting room where they signed consent documents and where children were fitted with the headcam. After a short period of play, they were escorted to a playroom in the lab where the free-play session (the focus of the current study) was conducted.

In the waiting room, the experimenter placed the headcam on children's heads after they had time to adjust to the environment. For children who resisted wearing the headcam, the experimenter used distractor techniques (presenting stimulating toys or engaging the children in hand-occupying activities) intended to keep children's focus elsewhere and prevent them from taking off the camera [@yoshida2008]. Once the child was wearing the camera comfortably for a period of time, child and caregiver were escorted to the playroom.

In the playroom, the experimenter presented the child's parent with a box containing three labeled pairs of objects, each consisting of a familiar and a novel object (e.g. a ball and a feather duster, marked as a "zem"). Parents confirmed that the child had not previously seen the novel toys. Parents were instructed to play with the object pairs with the child, one at a time, "as they typically would" and to use the novel labels to refer to the three toys. After giving these instructions, the experimenter left the room for a period of approximately 15 minutes. During this time, a tripod-mounted camera recorded video from a corner of the room and the headcam captured video from the child's perspective.

## Data Processing and Annotation

\begin{figure*}
\includegraphics[width=6in]{images/framesample.pdf}
\caption{\label{fig:frames} Sample frames from the headcam videos for a child from each age group, selected because they featured successful face detections (green squares).}
\end{figure*}

All headcam videos were cropped to exclude the period of entry to the playroom and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software. These sessions yielded a substantial amount of video: a total of
`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a milion frames), with an average video length of
`r round(sum(inc_subs$len.min)/60, digits = 2)` minutes
(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).

### Posture and Orientation Annotation

One major goal of our study was to understand the relationship between children's posture and their access to information from the faces of their caregiver. To investigate this relationship, we created a set of annotations for the child's physical posture (e.g. standing, sitting) and orientation of the caregiver relative to the child (e.g. in front of, behind, close, far away). For each headcam video, a coder used OpenSHAPA software to annotate both orientation and posture [@adolph2012].

Orientation was initially categorized as having the caregiver in front, to the side, or behind the child, and close (defined informally as within arm's reach) or farther away. Because of data sparsity, we consolidated this scheme into three categories: close to the caregiver with the caregiver either in front or on the side, farther from the caregiver again with caregiver either in front or on the side, and a global category of caregiver behind the child. Posture was categorized as being held/carried, lying face-up, sitting, prone (crawling or lying), standing, or other. Data from when the child was out of view of the tripod camera was marked as uncodable and excluded from these annotations.
% A second coder coded XYZ videos; their categorizations were reliable at XYZ.

### Labeling Annotation

We were also interested in the availability of social information proximate to naming events in the caregivers' speech to children. Accordingly, a human coder marked the onset time when the name of any of the six objects in the object set was used. Overall, caregivers produced a median of XYZ labels in a highly skewed distribution across participants (range: XYZ -- XYZ).

## Face Detection

An additional goal of the study was to measure the presence of caregivers' faces in the child's field of view (as approximated by the headcam). To avoid hand-annotating the size and position of faces in every frame of video, we tested two face detection systems. Sample frames from the video with successful detections are given in Figure \ref{fig:frames}.

### Face detection algorithms

The first algorithm was based on freely available computer vision tools [@bradski2008] and is described in depth in our previous work [@frank2012b]. This system had two parts. The first was the application of a set of four Haar-style face detection filters [@viola2004] to each frame of the videos independently. These detectors each provide information about whether a face is present in the frame as well as size and position for any detections. In a second step, these detections are then combined via a hidden Markov model (HMM), trained on hand-annotated data (see Appendix). The HMM model (which performed nearly as well as the more complex and computationally-intensive Conditional Random Field model used in our previous work) attempted to estimate whether a face was truly present in each frame of the videos, using as its input the number of Haar detectors that were active in any given frame.

The second algorithm that we evaluated was a semi-automated adaptive tracker-by-detection (SAATD). The algorithm required manual user input (selecting a single face example per video) for its initialization, but then needed no additional training data. The tracker is based on @kalal2010, which uses patches in the trajectory of an optical-flow based tracker [@lucas1981] to train and update a face detector. The optical flow tracker and the face detector work in parallel.
<!-- % For detection, we keep positive and negative templates of previously detected faces. -->
<!-- If the face detector finds a location in a new frame exhibiting a high similarity to its stored template, the tracker is re-initialised on that location. Otherwise, the tracker uses the optical flow to decide the location of a face in the new frame. -->
<!--  % The face detector is trained using visual features [@lepetit2005} and the detection is performed using a random ferns classifier [@ozuysal2007}. -->
<!-- The primary advantage of the SAATD algorithm is the use of motion for face detection: Following the movement of the pixels that define a face it is possible for the algorithm to adapt to new morphologies (i.e. different face poses). -->
<!-- %One disadvantage, however, is the necessity of manual initialization. -->

<!-- % [1] Z. Kalal, J. Matas, and K. Mikolajczyk. P-N learning: Bootstrapping binary classiﬁers by structural constraints. In 2010 IEEE Conference on Computer Vision and Pattern Recognition (CVPR), pages 49–56. IEEE, June 2010. -->

<!-- % [2] B. D. Lucas and T. Kanade. An iterative image registration technique with an application to stereo vision. In Proceedings of the International Joint Conference on Artiﬁcial Intelligence, pages 674–679, 1981 -->

<!-- % [3] V. Lepetit, P. Lagger, and P. Fua. Randomized trees for Real-Time keypoint recognition. -->
<!-- % In IEEE Computer Society Conference on Computer Vision and Pattern Recognition, volume 2, pages 775–781, Los Alamitos, CA, USA, 2005. IEEE Computer Society. -->

<!-- % [4] M. Ozuysal, P. Fua, and V. Lepetit. Fast keypoint recognition in ten lines of code. In IEEE Conference on Computer Vision and Pattern Recognition, Los Alamitos, CA, USA, June 2007. IEEE. -->

### Detector evaluation

\begin{table}[t]
  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
  \begin{center}
    \begin{tabular}{l|ccc|ccc}
      \hline
       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
      % \null Model & P & R & F & P & R & F  \\
      \hline
      % HMM & .55  & .38 &  .45 & .89 & .74 & .81   \\
      % SAATD & .86 & .78 & .81 & .93 & .76 & .83 \\
    \hline
    \end{tabular}
  \end{center}
\end{table}

To ensure that our evaluation was not biased by the relatively rare appearance of faces in the dataset, we annotated two samples, both a random sample from the data and a sample with a high-density of faces. We selected 1 minute of interaction for each age group, divided evenly across the four dyads at that age. For each dyad, we divided the recorded video into contiguous 1 s segments and selected 16 in accordance with two criteria. First, 8 of these segments were selected by choosing the parts of the videos highest in face detection (*high density sample*). To be fair to both algorithms, half of this was chosen from the segments with the most HMM detections and half were chosen from the segments with the most SAATD detections. The remaining segments were chosen by randomly sampling from segments not yet selected for coding (*random sample*).

Segments were annotated frame-by-frame by a human coder, who marked each frame as containing a face if at least half of the face was in the child's view. Detector output for each of these frames was then compared to this gold standard. A detection was counted as correct if it overlapped a face with half of its total area. The HMM training sample was selected via the same method as this gold standard sample, but used separate set of video segments.

We evaluated each algorithm on its precision (hits / hits + false alarms) and recall (hits / hits + misses), as well as F-score (the harmonic mean of these two measures). Results are reported in Table \ref{tab:model_eval}.

The HMM model obtained a relatively high level of performance for the random subsections, but performed poorly when there was a relatively high density of faces present. In contrast, SAATD performed well on both samples, giving better performance especially in cases where there was partial occlusion. Our goal in using face-detection algorithms was to provide a measurement technique that eliminated tedious and expensive hand-coding and provided acceptable results. We therefore selected the SAATD model and report detections from this algorithm as an estimate of face presence in all further analyses.

# Results

We report results from three different sets of analyses. First, we explore developmental changes in posture and orientation in our dataset. Next, we explore how these changes affect access to faces, as measured using our face-detection algorithm. Finally, we report preliminary results on the accessibility of faces during labeling.

## Changes in Posture and Orientation

<!-- \begin{figure}[t] -->
<!-- \centering -->
<!-- \includegraphics[width=5in]{figures/posture.pdf} -->
<!-- \caption{\label{fig:posture} Proportion time in each posture, plotted by child's age. Error bars show 95\% confidence intervals across participants, computed by non-parametric bootstrap.} -->
<!-- \end{figure} -->

<!-- \begin{figure}[t] -->
<!-- \centering -->
<!-- \includegraphics[width=5in]{figures/orientation.pdf} -->
<!-- \caption{\label{fig:orientation} Proportion time in each orientation, plotted by child's age. Error bars show 95\% confidence intervals across participants, computed by non-parametric bootstrap.} -->
<!-- \end{figure} -->

Our posture coding captured typical developmental milestones (Figure \ref{fig:posture}). Overall, sitting was the most common posture for interactions in the caregiver play session. The youngest infants in our sample mostly sat (with parental assistance), but also lay down and were carried a significant proportion of the time. The 12-month-olds were the only group who spent a large amount of time crawling, and the 16- and 20-month-olds sat and stood in equal parts.

Similarly, our coding of orientation revealed some significant developmental changes (Figure \ref{fig:orientation}). Younger children more frequently had the caregiver behind them, often because the caregiver was supporting the child's sitting posture (for the 4-month-olds especially). In contrast, the 12--20 month olds were able to locomote independently and so were able to spend more time further from the caregiver.

## Access to Faces

<!-- % \begin{figure}[ -->
<!-- % \begin{figure}[t] -->
<!-- % \centering -->
<!-- % \includegraphics[width=5in]{figures/posture.pdf} -->
<!-- % \caption{\label{fig:posture} Proportion face detections, split by age group (left panel), posture (middle panel), and caregiver's orientation (right panel). We omit the lying face-up posture due to data sparsity. Black points show individual participants and are jittered slightly on the horizontal, red lines show means and 95\% confidence intervals.} -->
<!-- % \end{figure}t] -->
<!-- % \centering -->
<!-- % \includegraphics[width=5in]{figures/posture.pdf} -->
<!-- % \caption{\label{fig:posture} Proportion face detections, split by age group (left panel), posture (middle panel), and caregiver's orientation (right panel). We omit the lying face-up posture due to data sparsity. Black points show individual participants and are jittered slightly on the horizontal, red lines show means and 95\% confidence intervals.} -->
<!-- % \end{figure} -->

We next investigated the effects of the child's posture and orientation on the presence and size of the caregiver's face in the visual field. Figure \ref{fig:face_dets} shows the proportion of frames with a positive face detection, plotted by the child's age, posture, and orientation relative to the caregiver.

Overall, there were very large differences in access to faces across age. The 4-month-olds saw almost no faces---their parents were behind them most of the time, supporting them since they could not sit independently. In contrast, the 8-month-olds, who could sit independently, typically sat across from their caregiver and saw many faces in both the sitting and prone postures. The 12-month-olds spent a large amount of time in the prone position (typically crawling after the ball, for example) and saw almost no faces in that posture. The 16- and 20-month-olds saw many faces because they were standing while their parents were sitting, putting their faces at a relatively similar level.

a
## Access to Faces During Labeling

Our final analysis concerned the accessibility of caregivers' faces during labeling events. @franchak2011 found that referential speech was marginally more likely to draw toddlers' attention to mothers' faces. We were similarly interested in whether looking at faces occurred during labeling. Accordingly, we used the labeling annotations for each child to identify the 2s before and after each labeling event. We then computed the proportion face detections within this window across ages.

The overall pattern of face accessibility closely mirrored the base rates shown in Figure \ref{fig:face_dets}. Although this general pattern in itself is important in assessing developmental access to social information, in the current analysis we were interested in whether there was differential access to faces around labeling instances. We thus computed difference scores between the baseline face detection rate and the rate of face detections in labeling windows for each participant. Figure \ref{fig:naming_face} shows the results of this analysis.

Although any conclusion must remain extremely tentative because of the small sample, we nevertheless saw an increase in label-related face access for the 20-month-olds. This difference was robust across a variety of window sizes from 1--6 s. (8-month-olds were more variable but similarly showed some trend towards greater face access during naming.) We cannot yet make inferences about the source of these differences: They could be could be caused by children, caregivers, or a combination of the two. Nevertheless, these results converge with previous work and suggest that, in combination with face detection techniques, the headcam may be a viable method for examining social access during language learning.

<!-- % \begin{figure} -->
<!-- % \begin{center} -->
<!-- %   \includegraphics[width=3in]{figures/naming_faces_diff.pdf} -->
<!-- % \end{center} -->
<!-- % \caption{\label{fig:naming_face} Proportion of faces detected in a 4s window of time centered around labeling events, plotted by age group and whether the word was familiar or novel. Error bars show 95\% confidence intervals. 4-month-olds are omitted due to the limited number of total face detections for this group.} -->
<!-- % % \vspace{-ex} -->
<!-- % \end{figure} -->

# General Discussion

Using a head-mounted camera, we explored the relationship between infants' postural and locomotor development and their visual access to social information. The use of automated annotation tools from computer vision allowed us to measure the prevalence of caregivers' faces in their children's visual field. We found systematic differences in the visual accessibility of faces based on posture, orientation, and age, as well as hints of differences in language-related changes in visual access. While these results remain preliminary given the size of our developmental sample, this work nevertheless provides an important proof-of-concept that computer vision techniques can be used as a measurement method in the developmental context.

The measures developed here have broad applicability to the study of individual and cultural differences. Since the physical circumstances of child rearing vary widely across households and across cultures, there may be important and predictable differences in children's visual experience. As suggested by the correlations between walking and vocabulary development [@walleunderreview], postural development may have substantial downstream consequences for language. For example, shifts in how infants are placed in particular postures by strollers or carriers [@zeedyk2008] or how their motor development is encouraged by parent practices [@bril1986] may lead to differences in social input which in turn affect their language learning. Since our variant of the headcam method is both inexpensive and highly portable, we have been able to deploy it in children's homes with some success; it may thus be a valuable tool for investigating differences in child-rearing practices.

A deep body of work uses children's linguistic input---measured using audio recordings---to understand the learning mechanisms underlying vocabulary acquisition [@huttenlocher1991;@hart1995;@fernald2006]. There have been some important initial successes in using visual input to predict language uptake [@yuinpress]. Nevertheless, we have a long way to go before our knowledge about children's visual input parallels our understanding of their linguistic environment. Coming to such an understanding will require the creation of both corpus resources and automated tools such as those we have begun to develop here.

# View angle

We note that previous studies have shown that children's head movements in the horizontal dimension are approximated by (though are slightly lagged by) their head movements [@yoshida2008]. Our own experience with the current apparatus ratifies these conclusions for the horizontal field but suggests that head movements in the vertical field are less reliable. Hence, these studies may run the risk of underestimating the proportion of faces actually seen by children.

# References

### 2018-01-10 analysis-validation/redo_mtcnn_csv.Rmd
more validation stuff

---
title: "redo_mtcnn_csv"
output: html_document
---

```{r}
data <- read_csv("data/final_output/mtcnn2.csv") %>%
  select(-X1)

# TODO preprocessing with new_frames? match header
# new_frames <- read_csv("data/final_output/new_frames.csv")
```

Filter out messed up videos (contain less frames)
```{r}
data %<>% 
  filter(!(video %in% c("XS_1612", "XS_1614", "XS_1631", "XS_1635")))
```

```{r}
result <- data %>%
  rbind(new_frames) %>%
  arrange(video, frame)
```

Sanity check # of frames
```{r}

```

```{r}
write_csv(result, "data/final_output/mtcnn3.csv")
```

### 2018-01-10 scripts/ground_truth_creation/get_gold_set_images.Rmd
more validation stuff

@@
 ```{r}
+library(tidyverse)
+library(magrittr)
+```
 
+# TODO replace with proper csv result
+```{r}
+data <- read_csv("data/final_output/mtcnn2.csv") %>%
+  select(-X1) %>%
+  distinct(video, frame, .keep_all = TRUE)
 ```
+
+Sample high density faces, remove selected from original data
+```{r}
+high_density_faces_sample <- data %>%
+  group_by(video) %>%
+  filter(is_face == "True") %>%
+  sample_n(7)
+
+data %<>%
+  anti_join(high_density_faces_sample)
+```
+
+Choose random frames from remaining
+```{r}
+random_sample <- data %>%
+  group_by(video) %>%
+  sample_n(7)
+```
+
+Combine and save result
+```{r}
+gold_set_candidates <- high_density_faces_sample %>%
+  rbind(random_sample) %>%
+  mutate(path = paste0(video, "/image-", frame, ".jpg"))
+
+# write_csv(gold_set_candidates, "data/ground_truth/gold_set_candidates.csv")
+```

### 2018-01-10 scripts/ground_truth_creation/download.py

@@ -1,10 +1,10 @@
 import os
 import csv
 
-with open("../sample.csv", 'rb') as img_paths_f:
+with open("../../data/gold_set_candidates.csv", 'rb') as img_paths_f:
         img_paths = list(csv.reader(img_paths_f, delimiter=','))
 
 for path in img_paths[1:]:
-	os.system("mkdir -p ~/xsface/sample_images/%s/%s" % (path[0], path[1]))
-	os.system("scp sanchez7@sherlock.stanford.edu:%s ~/xsface/sample_images/%s/%s" % (path[9], path[0], path[1]))
+	os.system("mkdir -p ~/xsface/sample_images2/%s" % (path[1]))
+	os.system("scp sanchez7@sherlock.stanford.edu:/home/groups/mcfrank/frames4/%s ~/xsface/sample_images2/%s" % (path[9], path[1]))
 

### 2018-01-10 scripts/ground_truth_creation/validate.py

@@ -2,6 +2,8 @@ import os
 import csv
 import ntpath
 
+# TODO change folder name to sample or whatever in ~/xsface
+
 BRIA_ANNOTATIONS = "/Users/alessandro/Downloads/sample_images_bria"
 
 if __name__ == "__main__":
@@ -27,7 +29,7 @@ if __name__ == "__main__":
 
     print(len(rows))
 
-    with open("ground_truth.csv", 'wb') as csvfile:
+    with open("ground_truth2.csv", 'wb') as csvfile:
         wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
         wr.writerow(['group', 'video', 'frame', 'is_face', 'angle'])
         for key, val in rows.iteritems():

### 2018-01-10 scripts/viola_jones/viola_jones_detector.py

@@ -1,0 +1,67 @@
+import os
+import re
+import sys
+import csv
+import ntpath
+import traceback
+import multiprocessing
+sys.path.append('/home/sanchez7/local')
+import cv2
+
+OUTPUT_FILE = 'tmp/face_detection_%s.csv'
+
+
+def detect_faces(gray_img):
+    cascPath = "haarcascade_frontalface_default.xml"
+    faceCascade = cv2.CascadeClassifier(cascPath)
+    return faceCascade.detectMultiScale(
+        gray_img,
+        scaleFactor=1.1,
+        minNeighbors=5,
+        minSize=(30, 30)
+    )
+
+
+def process_img(imgpath, i, arg):
+    # print('batch {0}: {1}%'.format(arg, i/1075.0 * 100))
+    print imgpath
+    gray_img = cv2.cvtColor(cv2.imread(imgpath), cv2.COLOR_BGR2GRAY)
+
+    faces = detect_faces(gray_img)
+
+    name = ntpath.basename(os.path.dirname(imgpath))
+    group = str(int(name.split("_")[1][:2]))
+    frame = re.search('image-(.*).jpg', ntpath.basename(imgpath)).group(1)
+
+    rows = []
+
+    if len(faces) == 0:
+        rows.append([group, name, frame, False, None, None, None, None])
+    else:
+        for (x, y, w, h) in faces:
+            rows.append([group, name, frame, True, x, y, w, h])
+
+
+if __name__ == "__main__":
+    vid_folder = sys.argv[1]
+    vid = ntpath.basename(vid_folder)
+
+    pool = multiprocessing.Pool()
+    results = []
+
+    for image in os.listdir(vid_folder):
+        results.append(pool.apply_async(process_img, args=(os.path.join(vid_folder, image),)))
+
+    pool.close()
+
+    with open(OUTPUT_FILE % vid, 'wb') as csvfile:
+        wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
+        # wr.writerow(['group', 'video', 'frame', 'is_face', 'x', 'y', 'w', 'h'])
+
+        for result in results:
+            try:
+                rows = result.get()
+                for row in rows:
+                    wr.writerow(row)
+            except:
+                traceback.print_exc()

### 2018-01-10 scripts/viola_jones/wrap_viola_jones_detector.sh

@@ -1,0 +1,13 @@
+#!/bin/sh
+# TODO add all frames to frames4 first
+for vid in $PI_HOME/frames4/*; do
+	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
+	    echo "Large video (> 10GB): $vid"
+		sbatch -p hns,normal -c 16 -t 10:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+		    --wrap="python viola_jones_detector.py $vid"
+	else
+		sbatch -p hns,normal -c 8 -t 5:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+		    --wrap="python viola_jones_detector.py $vid"
+	fi
+	sleep 1
+done

### 2018-01-10 scripts/mtcnn_face_detection/wrap2.sh
new frames

@@ -1,0 +1,10 @@
+#!/bin/sh
+
+module load tensorflow
+module load opencv/3.0.0
+
+for vid in $PI_HOME/redo_rotate/*; do
+		sbatch -p hns,normal -c 8 -t 10:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+		    --wrap="python detect_faces.py $vid"
+	sleep 1
+done

### 2018-01-10 scripts/ground_truth_creation/get_gold_set_images.Rmd
viola jones edits and gold set candidates

--- 2018-01-10 scripts/ground_truth_creation/get_gold_set_images.Rmd
@@
 ```{r}
+library(tidyverse)
+library(magrittr)
+```
 
+```{r}
+data <- read_csv("data/final_output/mtcnn3.csv") %>%
+  distinct(video, frame, .keep_all = TRUE)
 ```
+
+Sample high density faces, remove selected from original data
+```{r}
+high_density_faces_sample <- data %>%
+  group_by(video) %>%
+  filter(is_face == "True") %>%
+  sample_n(7)
+
+data %<>%
+  anti_join(high_density_faces_sample)
+```
+
+Choose random frames from remaining
+```{r}
+random_sample <- data %>%
+  group_by(video) %>%
+  sample_n(7)
+```
+
+Combine and save result
+```{r}
+gold_set_candidates <- high_density_faces_sample %>%
+  rbind(random_sample) %>%
+  mutate(path = paste0(video, "/image-", frame, ".jpg")) %>%
+  arrange(video, frame)
+
+write_csv(gold_set_candidates, "data/ground_truth/gold_set_candidates.csv")
+```

### 2018-01-10 scripts/ground_truth_creation/download.py

@@ -1,10 +1,10 @@
 import os
 import csv
 
-with open("../../data/gold_set_candidates.csv", 'rb') as img_paths_f:
+with open("../../data/ground_truth/gold_set_candidates.csv", 'rb') as img_paths_f:
         img_paths = list(csv.reader(img_paths_f, delimiter=','))
 
 for path in img_paths[1:]:
-	os.system("mkdir -p ~/xsface/sample_images2/%s" % (path[1]))
-	os.system("scp sanchez7@sherlock.stanford.edu:/home/groups/mcfrank/frames4/%s ~/xsface/sample_images2/%s" % (path[9], path[1]))
+    os.system("mkdir -p ~/xsface/sample_images3/%s" % (path[1]))
+    os.system("scp sanchez7@sherlock.stanford.edu:/share/PI/mcfrank/frames4/%s ~/xsface/sample_images3/%s" % (path[8], path[1]))
 

### 2018-01-10 scripts/viola_jones/viola_jones_detector.py

@@ -22,7 +22,7 @@ def detect_faces(gray_img):
     )
 
 
-def process_img(imgpath, i, arg):
+def process_img(imgpath):
     # print('batch {0}: {1}%'.format(arg, i/1075.0 * 100))
     print imgpath
     gray_img = cv2.cvtColor(cv2.imread(imgpath), cv2.COLOR_BGR2GRAY)

### 2018-01-10 scripts/viola_jones/wrap_viola_jones_detector.sh

@@ -1,5 +1,4 @@
 #!/bin/sh
-# TODO add all frames to frames4 first
 for vid in $PI_HOME/frames4/*; do
 	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
 	    echo "Large video (> 10GB): $vid"

### 2018-01-10 analysis-validation/mtcnn_evaluation.Rmd
eval mtcnn

---
title: "mtcnn evaluation"
output: html_notebook
---

500 frames were taken across all groups, half of them labeled as containing a face and half with no face.

Load data
```{r, message=FALSE, warning=FALSE}
setwd("..")

data <- read_csv("data/final_output/mtcnn3.csv") %>%
  distinct(video, frame, .keep_all = TRUE)

ground_truth_data <- read_csv("data/ground_truth/ground_truth3.csv") %>%
  select(-angle)
```

Join
```{r}
combined <- data %>%
  rename(detector_is_face = is_face) %>%
  inner_join(ground_truth_data, by = c("group", "video", "frame"))
```

Eval function
```{r}
evaluate <- function(a, b) {
  if (a == "True") {
    if (a == b) return ("TP")
    else return("FN")
  }
  else {
    if (a == b) return("TN")
    else return("FP")
  }
}
```

Evaluate
```{r}
results <- combined %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face))
```

Precison / Recall / F-score
```{r}
score <- results %>%
  group_by(result) %>%
  summarise(count = n())

p <- score[4,2] / (score[4,2] + score[2,2])
r <- score[4,2] / (score[4,2] + score[1,2])
f <- 2 * p * r / (p + r)
```

### 2018-01-10 scripts/ground_truth_creation/validate.py

@@ -4,7 +4,7 @@ import ntpath
 
 # TODO change folder name to sample or whatever in ~/xsface
 
-BRIA_ANNOTATIONS = "/Users/alessandro/Downloads/sample_images_bria"
+BRIA_ANNOTATIONS = "/Users/alessandro/xsface/sample_images3"
 
 if __name__ == "__main__":
 
@@ -15,21 +15,22 @@ if __name__ == "__main__":
         if root[len(root) - 1].isdigit(): # inside video directory
             for filename in filenames:
                 if filename == ".DS_Store": continue
-                group = ntpath.basename(os.path.dirname(root))
                 video = ntpath.basename(root)
+                group = str(int(video.split("_")[1][:2]))
                 frame = filename.split(".")[0].split("-")[1]
 
-                rows[filename.split(".")[0]] = [group, video, frame, "False", 0]
+                rows[video + filename.split(".")[0]] = [group, video, frame, "False", 0]
 
         elif ntpath.basename(root) == "annotations": #inside annotations
+            video = ntpath.basename(os.path.dirname(root))
             for filename in filenames:
-                arr = rows[filename.split(".")[0]]
+                arr = rows[video + filename.split(".")[0]]
                 arr[3] = "True"
-                rows[filename.split(".")[0]] = arr
+                rows[video + filename.split(".")[0]] = arr
 
     print(len(rows))
 
-    with open("ground_truth2.csv", 'wb') as csvfile:
+    with open("ground_truth3.csv", 'wb') as csvfile:
         wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
         wr.writerow(['group', 'video', 'frame', 'is_face', 'angle'])
         for key, val in rows.iteritems():

### 2018-01-10 scripts/viola_jones/viola_jones_detector.py

@@ -41,6 +41,8 @@ def process_img(imgpath):
         for (x, y, w, h) in faces:
             rows.append([group, name, frame, True, x, y, w, h])
 
+    return rows
+
 
 if __name__ == "__main__":
     vid_folder = sys.argv[1]

### 2018-01-10 analysis-main/main_analyses_bll.Rmd
integrated last detectors, reran sanity checks, added face size analyses

@@
 library(readr)
+library(plyr) # need to load this for scripts, always load before
 library(dplyr)
@@
 
-Note that the dataset here is those participants (31) who pass the length sanity checks.
+Note that the dataset here is now the complete data (N=36); all pass length sanity checks.
 
 ```{r}
-d <- read_csv("../data/consolidated_data_passing.csv") %>%
+d <- read_csv("../data/consolidated_data.csv") %>%
   mutate(posture = factor(posture), 
@@
   left_join(demo) %>%
-  mutate(len = as.numeric(lubridate::as.duration(lubridate::ms(len))))
+  # mutate(len = as.numeric(lubridate::as.duration(lubridate::ms(len)))) ## error here, used simpler calculation.
+  mutate(len = as.numeric(len_seconds/60))
 
@@
 ```
+# Face area: descriptives, posture and orientation
+## When detected, how much of the field of view did faces occupy? (% of screen)
+
+```{r}
+screenPixels=720 * 480
+d$faceArea = d$w * d$h /screenPixels
+
+mss <- d %>%
+  group_by(age.grp, subid) %>%
+  summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
+  group_by(age.grp)
+
+ms <- mss %>%
+  multi_boot_standard(col = "faceArea")
+
+ggplot(mss, aes(x = age.grp, y = faceArea)) + 
+  geom_jitter(width = .25) + 
+  geom_pointrange(data = ms, 
+                 aes(x = age.grp, 
+                     y = mean, 
+                     ymin = ci_lower, 
+                     ymax = ci_upper), col = "red") +
+  geom_line(data = ms, 
+             aes(x = age.grp, 
+                     y = mean), col = "red") + 
+  scale_x_continuous(breaks=c(8,12,16)) +
+  ylab("% ScreenArea of Face Detections") + 
+  xlab("Age (months)") + 
+  theme_bw() 
+```
+
+## Face size as a function of continous age
+
+```{r}
+mssFaceArea <- d %>%
+  filter(!is.na(age.at.test)) %>%
+  group_by(subid, age.at.test) %>%
+  summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
+  left_join(demo) %>%
+  mutate(len = as.numeric(len_seconds/60))
+
+ggplot(mssFaceArea, aes(x = age.at.test, y = faceArea)) + 
+  geom_jitter(width = .25, aes(size = len)) + 
+  geom_smooth() + 
+  scale_x_continuous(breaks=c(8,12,16)) +
+  scale_size_continuous(name = "Length (s)") + 
+  ylab("% Area of Face Detections") + 
+  xlab("Age (months)") + 
+  theme_bw() 
+```
+
+## Face size by infant's posture
+
+```{r}
+mss <- d %>%
+  filter(!is.na(posture)) %>%
+  filter(!is.na(face)) %>%
+  group_by(age.grp, posture, subid) %>%
+  summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
+  group_by(age.grp)
+
+# ms <- mss %>%
+#   multi_boot_standard(col = "faceArea")
+
+ggplot(mss, aes(x = factor(age.grp), y = faceArea, col = posture)) + 
+  geom_boxplot() + 
+  ylab("% Area of Face Detections") + 
+  xlab("Age (months)") + 
+  theme_bw() 
+```
+
+## Face Area by Caregiver Orientation
+```{r}
+mss <- d %>%
+  filter(!is.na(orientation)) %>%
+  group_by(age.grp, orientation, subid) %>%
+  summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
+  group_by(age.grp)
+
+# ms <- mss %>%
+#   multi_boot_standard(col = "faceArea")
+
+ggplot(mss, aes(x = factor(age.grp), y = faceArea, col = orientation)) + 
+  geom_boxplot() + 
+  ylab("% Area of Face Detections") + 
+  xlab("Age (months)") + 
+  theme_bw() 
+```
 
@@
   do(summarize.naming(.)) %>%
-  left_join(demo %>% select(subid, gender, age.at.test, len))
+  left_join(demo %>% select(subid, gender, age.at.test, len_seconds))
 ```
@@
   summarise(num = n(),
-            len = len[1]) %>%
-  mutate(rate = num / as.numeric(as.duration(lubridate::hms(len)) / dminutes())) # bll edits here to wrangle dates
+            len = len_seconds[1]) %>%
+  mutate(rate = num / as.numeric(len / 60)) # bll edits here to wrangle dates
 
@@
   summarise(num = n(),
-            len = len[1]) %>%
-  mutate(rate = num / as.numeric(as.duration(lubridate::hms(len)) / dminutes())) %>%  # bll edits here to wrangle dates
+            len = len_seconds[1]) %>%
+  mutate(rate = num / as.numeric(len / 60)) %>%  # bll edits here to wrangle dates
   ggplot(aes(x = age.at.test, y = rate, col = name)) +
@@
 ```{r}
-library(plyr)
+sb <- plyr::ddply(d,~subid, summarize.naming, window=c(-2,0)) ## used 2 second window as in ms draft
+sa <- plyr::ddply(d,~subid, summarize.naming, window=c(0,2))
 
-  sb <- ddply(d,~subid, summarize.naming, window=c(-3,0))
-  sa <- ddply(d,~subid, summarize.naming, window=c(0,3))
 sb$window <- "before"
@@
 
-<!-- ## How do face detections change with sequence of naming instances? -->
+## How do face detections change with sequence of naming instances?
 
-<!-- ```{r} -->
-<!-- ms <- namings %>% -->
-<!--   group_by(age.grp, subid, naming.instance, familiarity, name) %>% -->
-<!--   summarise(face = mean(face)) %>% -->
-<!--   group_by(age.grp, naming.instance, familiarity) %>% -->
-<!--   multi_boot_standard(col = "face", na.rm = TRUE) -->
-<!-- ``` -->
+```{r}
+ms <- namings %>%
+  group_by(age.grp, subid, naming.instance, familiarity, name) %>%
+  summarise(face = mean(face)) %>%
+  group_by(age.grp, naming.instance, familiarity) %>%
+  multi_boot_standard(col = "face", na.rm = TRUE)
+```
 
-<!-- ```{r} -->
-<!-- ggplot(ms, -->
-<!--        aes(x = naming.instance, y = mean, col = familiarity)) + -->
-<!--   geom_point() + -->
-<!--   geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) + -->
-<!--   geom_smooth(se = FALSE) + -->
-<!--   facet_grid(~age.grp) + -->
-<!--   xlim(c(0,20)) + -->
-<!--   scale_color_solarized() + -->
-<!--   theme_bw() -->
-<!-- ``` -->
+```{r}
+ggplot(ms,
+       aes(x = naming.instance, y = mean, col = familiarity)) +
+  geom_point() +
+  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  geom_smooth(se = FALSE) +
+  facet_grid(~age.grp) +
+  xlim(c(0,20)) +
+  scale_color_solarized() +
+  theme_bw()
+```
 

### 2018-01-10 analysis-validation/length_sanity_check.Rmd
integrated last detectors, reran sanity checks, added face size analyses

@@
 movies <- read_csv("../data/video_stats/new_movie_lengths.csv") 
-dets <- read_csv("../data/final_output/mtcnn2.csv")
+dets <- read_csv("../data/final_output/mtcnn3.csv")
 ```

### 2018-01-10 analysis-main/helper.R

@@ -149,6 +149,7 @@ summarize.naming <- function (x, window = c(-2,2)) {
   # for each row of p, populate posture and orientation to x
   namings$posture <- factor(NA, levels=levels(x$posture))
   namings$orientation <- factor(NA, levels=levels(x$orientation))
+
   
   for (i in 1:nrow(namings)) {
     t <- namings$time[i]

### 2018-01-10 analysis-main/merge_all_data.R

@@ -7,7 +7,7 @@ library(tidyr)
 source("helper.R")
 
 ## load detectors, demo data, merge
-dets <- read_csv("../data/final_output/mtcnn2.csv") %>%
+dets <- read_csv("../data/final_output/mtcnn3.csv") %>%
   mutate(y = as.numeric(y)) # because of leading zeros, apparently
 dets<-arrange(dets,video,frame) # comes out in weird order because of sherlock jobs!
 dets$subid <- dets$video # common names
@@ -39,8 +39,8 @@ d <- bind_rows(d, complete_combos)
 
 write_csv(d, "../data/consolidated_data.csv")
 
-## Create a set of videos that passes the sanity checks in length_sanity_check.Rmd
-d_passing <- filter(d, not(subid %in% c("XS_1205","XS_1211","XS_1225",
-                                        "XS_1628","XS_1639")))
-
-write_csv(d_passing, "../data/consolidated_data_passing.csv")
+# ## Create a set of videos that passes the sanity checks in length_sanity_check.Rmd
+# d_passing <- filter(d, not(subid %in% c("XS_1205","XS_1211","XS_1225",
+#                                         "XS_1628","XS_1639")))
+# 
+# write_csv(d_passing, "../data/consolidated_data_passing.csv")

### 2018-01-11 analysis-validation/mtcnn_evaluation.Rmd
validation code

@@
 title: "mtcnn evaluation"
-output: html_notebook
+output: html_document
 ---
@@
 setwd("..")
-
 data <- read_csv("data/final_output/mtcnn3.csv") %>%
@@
 f <- 2 * p * r / (p + r)
+
+print(paste("P:", round(p,3), "R:", round(r,3), "F:", round(f,3)))
 ```

### 2018-01-11 scripts/ground_truth_creation/get_gold_set_images.Rmd
validation code

@@
   filter(is_face == "True") %>%
-  sample_n(7)
+  sample_n(7) %>%
+  mutate(sample_type = "high_density")
 
@@
   group_by(video) %>%
-  sample_n(7)
+  sample_n(7) %>%
+  mutate(sample_type = "random")
 ```

### 2018-01-11 scripts/ground_truth_creation/validate.py

@@ -21,7 +21,7 @@ if __name__ == "__main__":
 
                 rows[video + filename.split(".")[0]] = [group, video, frame, "False", 0]
 
-        elif ntpath.basename(root) == "annotations": #inside annotations
+        elif ntpath.basename(root) == "annotations": # inside annotations
             video = ntpath.basename(os.path.dirname(root))
             for filename in filenames:
                 arr = rows[video + filename.split(".")[0]]

### 2018-01-11 analysis-main/main_analyses_bll.Rmd

@@
 ```
+
 # Face area: descriptives, posture and orientation
-## When detected, how much of the field of view did faces occupy? (% of screen)
+
+When detected, how much of the field of view did faces occupy? (% of screen)
+
+## Descriptves by age
 
@@
 
-## Face size as a function of continous age
+## Face Area as a function of continous age
 
@@
 
-## Face size by infant's posture
+## Face Area by infant's posture
 

### 2018-01-11 writeup-cogsci/cogsci2018.Rmd
reorg for barebones cogsci12018 rmarkdown files

--- 2018-01-10 writeup/xsface.Rmd
@@
 title: "Developmental and postural changes in children's visual access to faces"
-short-title: "Children's visual access to faces"
-output: kmr::apa_manuscript
+bibliography: xsface.bib
 csl: apa6.csl
-bibliography: xsface.bib
+document-params: "10pt, letterpaper"
 
-document-params: "a4paper,man,apacite,floatsintext,longtable"
-
-bib-tex: "xsface.bib"
-
-author-information:
-    - \author{Michael C. Frank, Guido Pusiol, Daniel Yurovsky, Kaia Simmons}
-
-affiliation-information:
-    - \affiliation{Department of Psychology, Stanford University}
-
-author-note:
-    "Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of the Language and Cognition Lab for help in recruitment, data collection, and annotation.
-This research was supported by a John Merck Scholars grant to MCF.
-An earlier version of this work was presented to the Cognitive Science Society in @frank2013.
-Please address correspondence to Michael C.
-Frank, Department of Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}."
+author-information: > 
+    \author{{\large \bf Alessandro Sacnhez} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Stanford University
+    \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
+        \And {\large \bf Ally Kraus} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
+        \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}
 
 abstract: 
-    "The faces of other people are a critical information source for young children.
+     The faces of other people are a critical information source for young children.
 During early development, children undergo significant postural and locomotor development, changing from lying and sitting infants to toddlers who walk independently.
-We used a head-mounted camera in conjunction with a face-detection system to explore the eff  ects of these changes on children's visual access to their caregivers' faces during an in-lab play session.
+We used a head-mounted camera in conjunction with a face-detection system to explore the effects of these changes on children's visual access to their caregivers' faces during an in-lab play session.
 In a cross-sectional sample of 8--16 month old children, we found substantial changes in face accessibility based on age and posture.
 These changes may translate into changes in the accessibility of social information during language learning.
-We make our corpus available for reanalysis, and discuss strengths and limitations of the headcam method more generally."
+We make our corpus available for reanalysis, and discuss strengths and limitations of the headcam method more generally.
 
@@
     "social cognition; face-perception; infancy; locomotion; head-cameras"
+
+output: cogsci2016::cogsci_paper
 ---
@@
 rm(list=ls())
+knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
+                      echo=F, warning=F, cache=F, message=F, sanitize = T)
+```
+
+```{r, libraries}
 library(knitr)
@@
                       cache=TRUE, message = FALSE, sanitize = TRUE)
-rm(list=ls())
 library(readr)
-# library(magrittr)
 library(dplyr)
@@
 library(xtable)
-# library(lubridate)
+library(lubridate)
 ```
-
-<!-- % A father offers his young daughter a novel object: a bright yellow feather duster.
-A few moments after she accepts the toy, he remarks, ``Isn't the zem funny?'' Her father may still be talking about the feather duster, or he may be describing a new object.
-To find out she has access to a simple and reliable method: she can look to his face to infer the direction of his attention.
--->
-
-The ability to follow social signals like eye-gaze is an important part of early social cognition [@scaife1975;@tomasello2009]
-
-In addition to being an important part of early social development, the ability to follow gaze is a strong predictor of children's early language development.
-For example, @brooks2005 found that children who followed an experimenter's gaze better before their first birthday had larger vocabularies at 18 months.
-Similarly, @carpenter1998 found that children's level of joint engagement (as well as the degree to which mothers followed the child's focus of attention in their labeling) predicted vocabulary growth in both language production and comprehension.
-These studies suggest that children's social environment plays a powerful supportive role in language learning.
-
-But at the same time as children are beginning to learn their first words, their view of the world is changing radically [@adolph2007].
-As speechless infants, they are unable to locomote independently.
-Before their first birthday, they begin crawling; soon after, they begin to walk independently.
-Infants' visual field is subject to the whims of their caregivers, but caregivers often place them in positions conducive to joint attention.
-In contrast, toddlers determine their own input to a much greater degree, but as a consequence they spend much of their time in a world primarily populated by knees.
-These postural and locomotor changes may have a profound effect on what children see.
-
-A recent study suggests the possibility of links between motor milestones, social cognition, and language.
-@walle2014 noted robust correlations between children's ability to walk and their vocabulary, both receptive and productive.
-On the basis of an observational study of parent input, they speculated that the emergence of walking may change the ability of the child to access social information (because walking toddlers see more of the social world than crawling infants).
-Accessing more social information may in turn allow children to discover word meanings more effectively.
-
-Recent methodological developments have the potential to provide data that would allow this hypothesis to be tested.
-The availability of head-mounted cameras and eye-trackers allows for the measurement of children's naturalistic environment in a way that was not previously possible.
-@yoshida2008 gave the first demonstration of the radical differences between toddler and adult perspectives on the social world, with toddlers' visual field being dominated by hands and objects much more than that of adults.
-More recent work has used head-mounted eye-tracking methods to measure young toddlers' fixations [@franchak2011], also finding that children look relatively infrequently at their mothers' faces in naturalistic play.
-
-These methods are now being applied to understand inputs to language acquisition.
-Work by Yu, Smith, and colleagues suggests that word learning is facilitated when parents and children create moments in which the visual field is dominated by a single object [@smith2011;@yuinpress].
-Some data even suggest that young children's restricted viewpoint may be more effective for learning words than the comparable adult perspective [@yurovsky2012].
-Together, this body of evidence suggests that measuring infants' perspective---and how it changes in motor development---is a critical part of understanding early language learning.
-
-In the current study we took a developmental approach to understanding the relationship between perspective and access to social information.
-We recorded head-camera data from a group of infants and children across a broad age range as they played with their caregivers during a brief laboratory visit.
-We then hand-annotated these data for the child's posture and parents' naming behavior and used face-detection algorithms to measure the frequency of faces in the child's visual field.
-The resulting dataset allows us to analyze changes in access to faces according to children's age, posture, and linguistic input.
 
@@
 `r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses.
+```
 
-## Head-mounted camera
+# Acknowledgements
 
-\begin{figure}
-\centering
-\includegraphics[width=3in]{images/viewangle.pdf}
-\includegraphics[width=2in]{images/headcam_w_fisheye3.jpg}
-\caption{\label{fig:headcam} Left: Field of view for three different headcam configurations, with the device we used in the middle.
-The lowest camera is pictured for comparison, but was not available until after our study was already in progress.
-Right: The light-weight, low-cost head-mounted camera with fisheye lens.}
-\end{figure}
-
-Our head-mounted camera ("headcam") is composed of a small, inexpensive MD80 model camera attached to a soft elastic headband from a camping headlamp.
-An aftermarket fisheye lens intended for iPhones and other Apple devices is attached to increase view angle.
-The total cost of each camera is approximately \$60.
-The camera captures 720x480 pixel images at approx.
-25 frames per second, and has battery life of 60--90 minutes.
-Without the fisheye lens, the viewing angle for the camera is $32^{\circ}$ horizontal by $24^{\circ}$ vertical; with the fisheye, $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left).
-The device is pictured in Figure \ref{fig:headcam}, right.
-Instructions for creating the camera are given in detail at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
-
-The vertical field of view of the camera was considerably smaller than the child's approximate vertical field of view, which---even at 6-7 months---spans around 100--120$^{\circ}$ in the vertical dimension [@mayer1988;@cummings1988].
-We were therefore faced with a choice in the orientation of the camera.
-If we chose a lower or higher orientation, we would be at risk of truncating either the child's own hands and physically proximate objects, or the faces of the adults around the child.
-Yet if we chose the middle orientation, we would still be at risk of underestimating the proportion of faces viewed by the child.
-Thus, for the purposes of the current study---measuring visual access to faces---we chose to orient the camera towards the upper part of the visual field.
-While this orientation decreased our chances of recording the objects being manipulated by the child, it nevertheless allowed us to capture the majority of the faces in the child's visual field.
-
-## Procedure
-
-Before coming to the lab, caregivers completed the MacArthur-Bates Communicative Development Inventories Words and Gestures form [@fenson2007], reporting on the words that they believed that their child understood and produced.
-
-After coming to the lab, families were seated in our waiting room where they signed consent documents and where children were fitted with the headcam.
-After a short period of play, they were escorted to a playroom in the lab where the free-play session (the focus of the current study) was conducted.
-
-In the waiting room, the experimenter placed the headcam on children's heads after they had time to adjust to the environment.
-For children who resisted wearing the headcam, the experimenter used distractor techniques (presenting stimulating toys or engaging the children in hand-occupying activities) intended to keep children's focus elsewhere and prevent them from taking off the camera [@yoshida2008].
-Once the child was wearing the camera comfortably for a period of time, child and caregiver were escorted to the playroom.
-
-In the playroom, the experimenter presented the child's parent with a box containing three labeled pairs of objects, each consisting of a familiar and a novel object (e.g.
-a ball and a feather duster, marked as a "zem").
-Parents confirmed that the child had not previously seen the novel toys.
-Parents were instructed to play with the object pairs with the child, one at a time, "as they typically would" and to use the novel labels to refer to the three toys.
-After giving these instructions, the experimenter left the room for a period of approximately 15 minutes.
-During this time, a tripod-mounted camera recorded video from a corner of the room and the headcam captured video from the child's perspective.
-
-## Data Processing and Annotation
-
-\begin{figure*}
-\includegraphics[width=6in]{images/framesample.pdf}
-\caption{\label{fig:frames} Sample frames from the headcam videos for a child from each age group, selected because they featured successful face detections (green squares).}
-\end{figure*}
-
-All headcam videos were cropped to exclude the period of entry to the playroom and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software.
-These sessions yielded a substantial amount of video: a total of
-`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a milion frames), with an average video length of
-`r round(sum(inc_subs$len.min)/60, digits = 2)` minutes
-(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).
-
-### Posture and Orientation Annotation
-
-One major goal of our study was to understand the relationship between children's posture and their access to information from the faces of their caregiver.
-To investigate this relationship, we created a set of annotations for the child's physical posture (e.g.
-standing, sitting) and orientation of the caregiver relative to the child (e.g.
-in front of, behind, close, far away).
-For each headcam video, a coder used OpenSHAPA software to annotate both orientation and posture [@adolph2012].
-
-Orientation was initially categorized as having the caregiver in front, to the side, or behind the child, and close (defined informally as within arm's reach) or farther away.
-Because of data sparsity, we consolidated this scheme into three categories: close to the caregiver with the caregiver either in front or on the side, farther from the caregiver again with caregiver either in front or on the side, and a global category of caregiver behind the child.
-Posture was categorized as being held/carried, lying face-up, sitting, prone (crawling or lying), standing, or other.
-Data from when the child was out of view of the tripod camera was marked as uncodable and excluded from these annotations.
-% A second coder coded XYZ videos; their categorizations were reliable at XYZ.
-
-### Labeling Annotation
-
-We were also interested in the availability of social information proximate to naming events in the caregivers' speech to children.
-Accordingly, a human coder marked the onset time when the name of any of the six objects in the object set was used.
-Overall, caregivers produced a median of XYZ labels in a highly skewed distribution across participants (range: XYZ -- XYZ).
-
-## Face Detection
-
-An additional goal of the study was to measure the presence of caregivers' faces in the child's field of view (as approximated by the headcam).
-To avoid hand-annotating the size and position of faces in every frame of video, we tested two face detection systems.
-Sample frames from the video with successful detections are given in Figure \ref{fig:frames}.
-
-### Face detection algorithms
-
-The first algorithm was based on freely available computer vision tools [@bradski2008] and is described in depth in our previous work [@frank2012b].
-This system had two parts.
-The first was the application of a set of four Haar-style face detection filters [@viola2004] to each frame of the videos independently.
-These detectors each provide information about whether a face is present in the frame as well as size and position for any detections.
-In a second step, these detections are then combined via a hidden Markov model (HMM), trained on hand-annotated data (see Appendix).
-The HMM model (which performed nearly as well as the more complex and computationally-intensive Conditional Random Field model used in our previous work) attempted to estimate whether a face was truly present in each frame of the videos, using as its input the number of Haar detectors that were active in any given frame.
-
-The second algorithm that we evaluated was a semi-automated adaptive tracker-by-detection (SAATD).
-The algorithm required manual user input (selecting a single face example per video) for its initialization, but then needed no additional training data.
-The tracker is based on @kalal2010, which uses patches in the trajectory of an optical-flow based tracker [@lucas1981] to train and update a face detector.
-The optical flow tracker and the face detector work in parallel.
-<!-- % For detection, we keep positive and negative templates of previously detected faces. -->
-<!-- If the face detector finds a location in a new frame exhibiting a high similarity to its stored template, the tracker is re-initialised on that location.
-Otherwise, the tracker uses the optical flow to decide the location of a face in the new frame.
--->
-<!--  % The face detector is trained using visual features [@lepetit2005} and the detection is performed using a random ferns classifier [@ozuysal2007}. -->
-<!-- The primary advantage of the SAATD algorithm is the use of motion for face detection: Following the movement of the pixels that define a face it is possible for the algorithm to adapt to new morphologies (i.e.
-different face poses).
--->
-<!-- %One disadvantage, however, is the necessity of manual initialization. -->
-
-<!-- % [1] Z.
-Kalal, J.
-Matas, and K.
-Mikolajczyk.
-P-N learning: Bootstrapping binary classiﬁers by structural constraints.
-In 2010 IEEE Conference on Computer Vision and Pattern Recognition (CVPR), pages 49–56.
-IEEE, June 2010.
--->
-
-<!-- % [2] B.
-D.
-Lucas and T.
-Kanade.
-An iterative image registration technique with an application to stereo vision.
-In Proceedings of the International Joint Conference on Artiﬁcial Intelligence, pages 674–679, 1981 -->
-
-<!-- % [3] V. Lepetit, P. Lagger, and P. Fua. Randomized trees for Real-Time keypoint recognition. -->
-<!-- % In IEEE Computer Society Conference on Computer Vision and Pattern Recognition, volume 2, pages 775–781, Los Alamitos, CA, USA, 2005. IEEE Computer Society. -->
-
-<!-- % [4] M.
-Ozuysal, P.
-Fua, and V.
-Lepetit.
-Fast keypoint recognition in ten lines of code.
-In IEEE Conference on Computer Vision and Pattern Recognition, Los Alamitos, CA, USA, June 2007.
-IEEE.
--->
-
-### Detector evaluation
-
-\begin{table}[t]
-  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
-  \begin{center}
-    \begin{tabular}{l|ccc|ccc}
-      \hline
-       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
-      % \null Model & P & R & F & P & R & F  \\
-      \hline
-      % HMM & .55  & .38 &  .45 & .89 & .74 & .81   \\
-      % SAATD & .86 & .78 & .81 & .93 & .76 & .83 \\
-    \hline
-    \end{tabular}
-  \end{center}
-\end{table}
-
-To ensure that our evaluation was not biased by the relatively rare appearance of faces in the dataset, we annotated two samples, both a random sample from the data and a sample with a high-density of faces.
-We selected 1 minute of interaction for each age group, divided evenly across the four dyads at that age.
-For each dyad, we divided the recorded video into contiguous 1 s segments and selected 16 in accordance with two criteria.
-First, 8 of these segments were selected by choosing the parts of the videos highest in face detection (*high density sample*).
-To be fair to both algorithms, half of this was chosen from the segments with the most HMM detections and half were chosen from the segments with the most SAATD detections.
-The remaining segments were chosen by randomly sampling from segments not yet selected for coding (*random sample*).
-
-Segments were annotated frame-by-frame by a human coder, who marked each frame as containing a face if at least half of the face was in the child's view.
-Detector output for each of these frames was then compared to this gold standard.
-A detection was counted as correct if it overlapped a face with half of its total area.
-The HMM training sample was selected via the same method as this gold standard sample, but used separate set of video segments.
-
-We evaluated each algorithm on its precision (hits / hits + false alarms) and recall (hits / hits + misses), as well as F-score (the harmonic mean of these two measures).
-Results are reported in Table \ref{tab:model_eval}.
-
-The HMM model obtained a relatively high level of performance for the random subsections, but performed poorly when there was a relatively high density of faces present.
-In contrast, SAATD performed well on both samples, giving better performance especially in cases where there was partial occlusion.
-Our goal in using face-detection algorithms was to provide a measurement technique that eliminated tedious and expensive hand-coding and provided acceptable results.
-We therefore selected the SAATD model and report detections from this algorithm as an estimate of face presence in all further analyses.
-
-# Results
-
-We report results from three different sets of analyses.
-First, we explore developmental changes in posture and orientation in our dataset.
-Next, we explore how these changes affect access to faces, as measured using our face-detection algorithm.
-Finally, we report preliminary results on the accessibility of faces during labeling.
-
-## Changes in Posture and Orientation
-
-<!-- \begin{figure}[t] -->
-<!-- \centering -->
-<!-- \includegraphics[width=5in]{figures/posture.pdf} -->
-<!-- \caption{\label{fig:posture} Proportion time in each posture, plotted by child's age. Error bars show 95\% confidence intervals across participants, computed by non-parametric bootstrap.} -->
-<!-- \end{figure} -->
-
-<!-- \begin{figure}[t] -->
-<!-- \centering -->
-<!-- \includegraphics[width=5in]{figures/orientation.pdf} -->
-<!-- \caption{\label{fig:orientation} Proportion time in each orientation, plotted by child's age.
-Error bars show 95\% confidence intervals across participants, computed by non-parametric bootstrap.} -->
-<!-- \end{figure} -->
-
-Our posture coding captured typical developmental milestones (Figure \ref{fig:posture}).
-Overall, sitting was the most common posture for interactions in the caregiver play session.
-The youngest infants in our sample mostly sat (with parental assistance), but also lay down and were carried a significant proportion of the time.
-The 12-month-olds were the only group who spent a large amount of time crawling, and the 16- and 20-month-olds sat and stood in equal parts.
-
-Similarly, our coding of orientation revealed some significant developmental changes (Figure \ref{fig:orientation}).
-Younger children more frequently had the caregiver behind them, often because the caregiver was supporting the child's sitting posture (for the 4-month-olds especially).
-In contrast, the 12--20 month olds were able to locomote independently and so were able to spend more time further from the caregiver.
-
-## Access to Faces
-
-<!-- % \begin{figure}[ -->
-<!-- % \begin{figure}[t] -->
-<!-- % \centering -->
-<!-- % \includegraphics[width=5in]{figures/posture.pdf} -->
-<!-- % \caption{\label{fig:posture} Proportion face detections, split by age group (left panel), posture (middle panel), and caregiver's orientation (right panel).
-We omit the lying face-up posture due to data sparsity.
-Black points show individual participants and are jittered slightly on the horizontal, red lines show means and 95\% confidence intervals.} -->
-<!-- % \end{figure}t] -->
-<!-- % \centering -->
-<!-- % \includegraphics[width=5in]{figures/posture.pdf} -->
-<!-- % \caption{\label{fig:posture} Proportion face detections, split by age group (left panel), posture (middle panel), and caregiver's orientation (right panel).
-We omit the lying face-up posture due to data sparsity.
-Black points show individual participants and are jittered slightly on the horizontal, red lines show means and 95\% confidence intervals.} -->
-<!-- % \end{figure} -->
-
-We next investigated the effects of the child's posture and orientation on the presence and size of the caregiver's face in the visual field.
-Figure \ref{fig:face_dets} shows the proportion of frames with a positive face detection, plotted by the child's age, posture, and orientation relative to the caregiver.
-
-Overall, there were very large differences in access to faces across age.
-The 4-month-olds saw almost no faces---their parents were behind them most of the time, supporting them since they could not sit independently.
-In contrast, the 8-month-olds, who could sit independently, typically sat across from their caregiver and saw many faces in both the sitting and prone postures.
-The 12-month-olds spent a large amount of time in the prone position (typically crawling after the ball, for example) and saw almost no faces in that posture.
-The 16- and 20-month-olds saw many faces because they were standing while their parents were sitting, putting their faces at a relatively similar level.
-
-a
-## Access to Faces During Labeling
-
-Our final analysis concerned the accessibility of caregivers' faces during labeling events.
-@franchak2011 found that referential speech was marginally more likely to draw toddlers' attention to mothers' faces.
-We were similarly interested in whether looking at faces occurred during labeling.
-Accordingly, we used the labeling annotations for each child to identify the 2s before and after each labeling event.
-We then computed the proportion face detections within this window across ages.
-
-The overall pattern of face accessibility closely mirrored the base rates shown in Figure \ref{fig:face_dets}.
-Although this general pattern in itself is important in assessing developmental access to social information, in the current analysis we were interested in whether there was differential access to faces around labeling instances.
-We thus computed difference scores between the baseline face detection rate and the rate of face detections in labeling windows for each participant.
-Figure \ref{fig:naming_face} shows the results of this analysis.
-
-Although any conclusion must remain extremely tentative because of the small sample, we nevertheless saw an increase in label-related face access for the 20-month-olds.
-This difference was robust across a variety of window sizes from 1--6 s.
-(8-month-olds were more variable but similarly showed some trend towards greater face access during naming.) We cannot yet make inferences about the source of these differences: They could be could be caused by children, caregivers, or a combination of the two.
-Nevertheless, these results converge with previous work and suggest that, in combination with face detection techniques, the headcam may be a viable method for examining social access during language learning.
-
-<!-- % \begin{figure} -->
-<!-- % \begin{center} -->
-<!-- %   \includegraphics[width=3in]{figures/naming_faces_diff.pdf} -->
-<!-- % \end{center} -->
-<!-- % \caption{\label{fig:naming_face} Proportion of faces detected in a 4s window of time centered around labeling events, plotted by age group and whether the word was familiar or novel.
-Error bars show 95\% confidence intervals.
-4-month-olds are omitted due to the limited number of total face detections for this group.} -->
-<!-- % % \vspace{-ex} -->
-<!-- % \end{figure} -->
-
-# General Discussion
-
-Using a head-mounted camera, we explored the relationship between infants' postural and locomotor development and their visual access to social information.
-The use of automated annotation tools from computer vision allowed us to measure the prevalence of caregivers' faces in their children's visual field.
-We found systematic differences in the visual accessibility of faces based on posture, orientation, and age, as well as hints of differences in language-related changes in visual access.
-While these results remain preliminary given the size of our developmental sample, this work nevertheless provides an important proof-of-concept that computer vision techniques can be used as a measurement method in the developmental context.
-
-The measures developed here have broad applicability to the study of individual and cultural differences.
-Since the physical circumstances of child rearing vary widely across households and across cultures, there may be important and predictable differences in children's visual experience.
-As suggested by the correlations between walking and vocabulary development [@walleunderreview], postural development may have substantial downstream consequences for language.
-For example, shifts in how infants are placed in particular postures by strollers or carriers [@zeedyk2008] or how their motor development is encouraged by parent practices [@bril1986] may lead to differences in social input which in turn affect their language learning.
-Since our variant of the headcam method is both inexpensive and highly portable, we have been able to deploy it in children's homes with some success; it may thus be a valuable tool for investigating differences in child-rearing practices.
-
-A deep body of work uses children's linguistic input---measured using audio recordings---to understand the learning mechanisms underlying vocabulary acquisition [@huttenlocher1991;@hart1995;@fernald2006].
-There have been some important initial successes in using visual input to predict language uptake [@yuinpress].
-Nevertheless, we have a long way to go before our knowledge about children's visual input parallels our understanding of their linguistic environment.
-Coming to such an understanding will require the creation of both corpus resources and automated tools such as those we have begun to develop here.
-
-# View angle
-
-We note that previous studies have shown that children's head movements in the horizontal dimension are approximated by (though are slightly lagged by) their head movements [@yoshida2008].
-Our own experience with the current apparatus ratifies these conclusions for the horizontal field but suggests that head movements in the vertical field are less reliable.
-Hence, these studies may run the risk of underestimating the proportion of faces actually seen by children.
+"Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of the Language and Cognition Lab for help in recruitment, data collection, and annotation.
+This research was supported by a John Merck Scholars grant to MCF.
+An earlier version of this work was presented to the Cognitive Science Society in @frank2013.
+Please address correspondence to Michael C.
+Frank, Department of Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}."
 
 # References
+
+```{r}
+# References will be generated automatically by Pandoc and included here.
+# The following code is some latex to format the bibliography. Do not remove it.
+```
+
+\setlength{\parindent}{-0.1in} 
+\setlength{\leftskip}{0.125in}
+\noindent

### 2018-01-11 writeup-cogsci/cogsci2018.tex
reorg for barebones cogsci12018 rmarkdown files

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Developmental and postural changes in children's visual access to faces}

\author{{\large \bf Alessandro Sacnhez} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Ally Kraus} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

\begin{document}

\maketitle

\begin{abstract}
The faces of other people are a critical information source for young
children. During early development, children undergo significant
postural and locomotor development, changing from lying and sitting
infants to toddlers who walk independently. We used a head-mounted
camera in conjunction with a face-detection system to explore the
effects of these changes on children's visual access to their
caregivers' faces during an in-lab play session. In a cross-sectional
sample of 8--16 month old children, we found substantial changes in face
accessibility based on age and posture. These changes may translate into
changes in the accessibility of social information during language
learning. We make our corpus available for reanalysis, and discuss
strengths and limitations of the headcam method more generally.

\textbf{Keywords:}
social cognition; face-perception; infancy; locomotion; head-cameras
\end{abstract}

\section{Methods}\label{methods}

\subsection{Participants}\label{participants}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
Age group & N & \% included & Mean age & Video length (min) & \% female \\ 
  \hline
   8 &  12 & 0.46 & 8.71 & 14.41 & 0.50 \\ 
   12 &  12 & 0.40 & 12.62 & 13.48 & 0.58 \\ 
   16 &  12 & 0.31 & 16.29 & 15.00 & 0.50 \\ 
   \hline
\end{tabular}
\caption{\label{tab:pop} Demographics by age group.}
\end{table}

Our final sample consisted of 36 infants and children, distributed into
each of three age groups: 8 months, 12 months, and 16 months.
Participants were recruited from the surrounding community via state
birth records. Participants had no documented disabilities and were
reported to hear at least 80\% English at home. Demographics and
exclusion rates are given in Table \ref{tab:pop}.

In order to compile this final sample, we tested a substantially larger
group of children, who were excluded for the following reasons: 20 for
technical issues related to the headcam, 15 for failing to wear the
headcam, 10 for fewer than 4 minutes of headcam footage, 5 for having
multiple adults present, 5 for missing CDI data, 2 for missing scene
camera footage, 1 for fussiness, and one excluded for sample symmetry.
All inclusion decisions were made independent of the results of
subsequent analyses. ```

\section{Acknowledgements}\label{acknowledgements}

``Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of
the Language and Cognition Lab for help in recruitment, data collection,
and annotation. This research was supported by a John Merck Scholars
grant to MCF. An earlier version of this work was presented to the
Cognitive Science Society in Frank, Simmons, Yurovsky, \& Pusiol (2013).
Please address correspondence to Michael C. Frank, Department of
Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford,
CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}.''

\section{References}\label{references}

\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent

\hypertarget{refs}{}
\hypertarget{ref-frank2013}{}
Frank, M. C., Simmons, K., Yurovsky, D., \& Pusiol, G. (2013).
Developmental and postural changes in childrens visual access to faces.
In \emph{Proceedings of the 35th annual meeting of the cognitive science
society} (pp. 454--459).

\end{document}

### 2018-01-11 writeup-cogsci/cogsci2018.Rmd
rewrote some methods sections, redid demo tables

@@
 author-information: > 
-    \author{{\large \bf Alessandro Sacnhez} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Stanford University
+    \author{{\large \bf Alessandro Sanchez} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Stanford University
     \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
@@
 abstract:
-     The faces of other people are a critical information source for young children.
-During early development, children undergo significant postural and locomotor development, changing from lying and sitting infants to toddlers who walk independently.
-We used a head-mounted camera in conjunction with a face-detection system to explore the effects of these changes on children's visual access to their caregivers' faces during an in-lab play session.
-In a cross-sectional sample of 8--16 month old children, we found substantial changes in face accessibility based on age and posture.
-These changes may translate into changes in the accessibility of social information during language learning.
-We make our corpus available for reanalysis, and discuss strengths and limitations of the headcam method more generally.
 
@@
 # Methods
-
 ## Participants
-
 ```{r echo=FALSE, include=FALSE}
@@
 
-# # exclusions by age
-# all_subs %>% 
-#   group_by(reason, age_group) %>%
-#   summarise(n = n()) %>%
-#   spread(age_group, n) %>%
-#   xtable %>%
-#   print(digits = 2)
-
 all_subs %>%
@@
             video_length = mean(len.min, na.rm=TRUE), 
-            female = mean(gender=="female", na.rm=TRUE)) %>%
+            female = sum(gender=="female", na.rm=TRUE)) %>%
   xtable %>%
-  print(digits = 2, include.rownames=FALSE)
+  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
 ```
-
-\begin{table}[ht]
+Our final sample consisted of 36 infants and children, with 8 participants in three age groups: 8 months (6 females), 12 months (7 females), and 16 months (6 females).
+Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80\% English at home.
+Demographics and exclusion rates are given in Table \ref{tab:pop}.
+\begin{table}[H]
 \centering
@@
   \hline
-Age group & N & \% included & Mean age & Video length (min) & \% female \\ 
+ Group & N & \% incl. & Mean age & Videos length (min) \\ 
   \hline
-   8 &  12 & 0.46 & 8.71 & 14.41 & 0.50 \\ 
-   12 &  12 & 0.40 & 12.62 & 13.48 & 0.58 \\ 
-   16 &  12 & 0.31 & 16.29 & 15.00 & 0.50 \\ 
+   8 mo. &   12 & 0.46 & 8.71 & 14.41 \\ 
+   12 mo. &  12 & 0.40 & 12.62 & 13.48 \\ 
+   16 mo. &  12 & 0.31 & 16.29 & 15.00\\ 
    \hline
@@
 \end{table}
-
-Our final sample consisted of 36 infants and children, distributed into each of three age groups: 8 months, 12 months, and 16 months.
-Participants were recruited from the surrounding community via state birth records.
-Participants had no documented disabilities and were reported to hear at least 80\% English at home.
-Demographics and exclusion rates are given in Table \ref{tab:pop}.
-
-In order to compile this final sample, we tested a substantially larger group of children, who were excluded for the following reasons: 
+To obtain this final sample, we tested `r length(all_subs$reason)`, excluding  `r sum(all_subs$reason!="included")` children for the following reasons:
 `r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
@@
 `r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses.
+
+## Head-mounted camera
+We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband.Videos captured by the headcam were 720x480 pixels with 25 frames per second.Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
+A fisheye lens was attached to the camera to cincrease the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left).
+
+Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's approximate vertical field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988].
+As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirely of the child's upper visual field where the child is likely to see the faces of adults around them.
+This allowed us to maximize our chances of capturing faces that the child would have seen during the play session.
+
+```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for three different headcam configurations, with the device we used in the middle.
+The lowest camera is pictured for comparison, but was not available until after our study was already in progress."}
+img <- png::readPNG("images/viewangle.png")
+grid::grid.raster(img)
 ```
+
+## Procedure
+
+First, all parents signed consent documentsin a waiting room  where children were fitted with the headcam.
+After the child was comfortable in the waiting room and with the experimenter, the experimenter placed the headcam on the child's head.
+If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam [@yoshida2008].
+
+After the child was comfortable with wearing the headcam, the child and caregiver were shown to a playroom for the free-play session--the focus of the current study.
+Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a feather duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previosuly seen the novel toys and were instructed to use the novel labels to refer to the novel toys.
+
+The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.
+
+## Data Processing and Annotation
+
+\begin{figure*}
+\includegraphics[width=6in]{images/framesample.pdf}
+\caption{\label{fig:frames} Sample frames from the headcam videos for a child from each age group, selected because they featured successful face detections (green squares).}
+\end{figure*}
+
+All headcam videos were cropped to exclude the period of entry to the playroom and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software.
+These sessions yielded a substantial amount of video: a total of
+`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a milion frames), with an average video length of
+`r round(sum(inc_subs$len.min)/60, digits = 2)` minutes
+(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).
+
+### Posture and Orientation Annotation
+
+We created a set of custom annotations that described the child's physical posture (e.g.
+standing) and the orientation of the caregiver relative to the child (e.g.
+far away).
+The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing.
+The caregiver's orientation was characterized as being close to the child, farther from the child, and a global category of caregiver behind the child.
+For the first two annotationes (close/far from the child), the caregiver could either be to the the front or to the side of the child.
+All annontations were made using OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera was marked as uncodable and was excluded from these annotations.
+
+## Face Detection
+
+An additional goal of the study was to measure the presence of caregivers' faces in the child's field of view (as approximated by the headcam).
+To avoid hand-annotating the size and position of faces in every frame of video, we tested two face detection systems.
+Sample frames from the video with successful detections are given in Figure \ref{fig:frames}.
+
+### Face detection algorithms
+
+### Detector evaluation
+
+\begin{table}[t]
+  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
+  \begin{center}
+    \begin{tabular}{l|ccc|ccc}
+      \hline
+       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
+      % \null Model & P & R & F & P & R & F  \\
+      \hline
+      % Viola-Jones & .55  & .38 &  .45 & .89 & .74 & .81   \\
+      % MTCNN & .86 & .78 & .81 & .93 & .76 & .83 \\
+      % OpenPose & .86 & .78 & .81 & .93 & .76 & .83 \\
+    \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+# Results
+We report results from three different sets of analyses.
+First, we explore developmental changes in posture and orientation in our dataset.
+Next, we explore how these changes affect access to faces and to hands, as measured using computater vision algorithms.
+Finally, we explore how these changes impact the accessibility of faces and hands during labeling events.
+
+## Changes in Posture and Orientation
+
+## Changes in Access to Faces
+
+## Changes in Access to Hands
 
@@
 
-"Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of the Language and Cognition Lab for help in recruitment, data collection, and annotation.
+Thanks to Kaia Simmons, Kathy Woo, Aditi Maliwal, and other members of the Language and Cognition Lab for help in recruitment, data collection, and annotation.
 This research was supported by a John Merck Scholars grant to MCF.
@@
 Please address correspondence to Michael C.
-Frank, Department of Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}."
+Frank, Department of Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}.
 

### 2018-01-11 writeup-cogsci/cogsci2018.tex
rewrote some methods sections, redid demo tables

@@
 
-\author{{\large \bf Alessandro Sacnhez} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
+\author{{\large \bf Alessandro Sanchez} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
     \And {\large \bf Ally Kraus} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
@@
 \begin{abstract}
-The faces of other people are a critical information source for young
-children. During early development, children undergo significant
-postural and locomotor development, changing from lying and sitting
-infants to toddlers who walk independently. We used a head-mounted
-camera in conjunction with a face-detection system to explore the
-effects of these changes on children's visual access to their
-caregivers' faces during an in-lab play session. In a cross-sectional
-sample of 8--16 month old children, we found substantial changes in face
-accessibility based on age and posture. These changes may translate into
-changes in the accessibility of social information during language
-learning. We make our corpus available for reanalysis, and discuss
-strengths and limitations of the headcam method more generally.
 
@@
 
-\begin{table}[ht]
+Our final sample consisted of 36 infants and children, with 8
+participants in three age groups: 8 months (6 females), 12 months (7
+females), and 16 months (6 females). Participants were recruited from
+the surrounding community via state birth records, had no documented
+disabilities, and were reported to hear at least 80\% English at home.
+Demographics and exclusion rates are given in Table \ref{tab:pop}.
+
+\begin{table}[H]
 \centering
@@
 
-Our final sample consisted of 36 infants and children, distributed into
-each of three age groups: 8 months, 12 months, and 16 months.
-Participants were recruited from the surrounding community via state
-birth records. Participants had no documented disabilities and were
-reported to hear at least 80\% English at home. Demographics and
-exclusion rates are given in Table \ref{tab:pop}.
+To obtain this final sample, we tested 95, excluding 59 children for the
+following reasons: 20 for technical issues related to the headcam, 15
+for failing to wear the headcam, 10 for fewer than 4 minutes of headcam
+footage, 5 for having multiple adults present, 5 for missing CDI data, 2
+for missing scene camera footage, 1 for fussiness, and one excluded for
+sample symmetry. All inclusion decisions were made independent of the
+results of subsequent analyses.
 
-In order to compile this final sample, we tested a substantially larger
-group of children, who were excluded for the following reasons: 20 for
-technical issues related to the headcam, 15 for failing to wear the
-headcam, 10 for fewer than 4 minutes of headcam footage, 5 for having
-multiple adults present, 5 for missing CDI data, 2 for missing scene
-camera footage, 1 for fussiness, and one excluded for sample symmetry.
-All inclusion decisions were made independent of the results of
-subsequent analyses. ```
+\subsection{Head-mounted camera}\label{head-mounted-camera}
+
+We used a small, head-mounted camera (``headcam'') that was constructed
+from a MD80 model camera attached to a soft elastic headband.Videos
+captured by the headcam were 720x480 pixels with 25 frames per
+second.Detailed instructions for creating this headcam can be found at
+\url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
+A fisheye lens was attached to the camera to cincrease the view angle
+from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
+\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
+\ref{fig:headcam}, left).
+
+Even with the fish-eye lens, the vertical field of view of the camera is
+still considerably reduced compared to the child's approximate vertical
+field of view, which spans around 100--120\(^{\circ}\) in the vertical
+dimension by 6-7 months of age (Cummings, Van Hof-Van Duin, Mayer,
+Hansen, \& Fulton, 1988; Mayer, Fulton, \& Cummings, 1988). As we were
+primarily interested in the presence of faces in the child's field of
+view, we chose to orient the camera upwards to capture the entirely of
+the child's upper visual field where the child is likely to see the
+faces of adults around them. This allowed us to maximize our chances of
+capturing faces that the child would have seen during the play session.
+
+\begin{CodeChunk}
+\begin{figure}[H]
+
+{\centering \includegraphics{figs/image-1}
+
+}
+
+\caption[Field of view for three different headcam configurations, with the device we used in the middle]{Field of view for three different headcam configurations, with the device we used in the middle.
+The lowest camera is pictured for comparison, but was not available until after our study was already in progress.}\label{fig:image}
+\end{figure}
+\end{CodeChunk}
+
+\subsection{Procedure}\label{procedure}
+
+First, all parents signed consent documentsin a waiting room where
+children were fitted with the headcam. After the child was comfortable
+in the waiting room and with the experimenter, the experimenter placed
+the headcam on the child's head. If the child was uninterested in
+wearing the headcam or tried to take it off, the experimenter presented
+engaging toys to try to draw the child's focus away from the headcam
+(Yoshida \& Smith, 2008).
+
+After the child was comfortable with wearing the headcam, the child and
+caregiver were shown to a playroom for the free-play session--the focus
+of the current study. Parents were shown a box containing three pairs of
+novel and familiar objects (e.g., a ball and a feather duster, named a
+``zem''), and were instructed to play with the object pairs with their
+child one at a time, ``as they typically would.'' All parents confirmed
+that their child had not previosuly seen the novel toys and were
+instructed to use the novel labels to refer to the novel toys.
+
+The experimenter then left the playroom for approximately 15 minutes,
+during which a tripod-mounted camera in the corner of the room recorded
+the session and the headcam captured video from the child's perspective.
+
+\subsection{Data Processing and
+Annotation}\label{data-processing-and-annotation}
+
+\begin{figure*}
+\includegraphics[width=6in]{images/framesample.pdf}
+\caption{\label{fig:frames} Sample frames from the headcam videos for a child from each age group, selected because they featured successful face detections (green squares).}
+\end{figure*}
+
+All headcam videos were cropped to exclude the period of entry to the
+playroom and were automatically synchronized with the tripod-mounted
+videos using FinalCut Pro Software. These sessions yielded a substantial
+amount of video: a total of 516 minutes (almost a milion frames), with
+an average video length of 8.6 minutes (min = 4.53, max = 19.35).
+
+\subsubsection{Posture and Orientation
+Annotation}\label{posture-and-orientation-annotation}
+
+We created a set of custom annotations that described the child's
+physical posture (e.g.~standing) and the orientation of the caregiver
+relative to the child (e.g.~far away). The child's posture was
+categorized as being held/carried, prone (crawling or lying), sitting,
+or standing. The caregiver's orientation was characterized as being
+close to the child, farther from the child, and a global category of
+caregiver behind the child. For the first two annotationes (close/far
+from the child), the caregiver could either be to the the front or to
+the side of the child. All annontations were made using
+OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
+Millman, 2012), and times when the child was out of view of the tripod
+camera was marked as uncodable and was excluded from these annotations.
+
+\subsection{Face Detection}\label{face-detection}
+
+An additional goal of the study was to measure the presence of
+caregivers' faces in the child's field of view (as approximated by the
+headcam). To avoid hand-annotating the size and position of faces in
+every frame of video, we tested two face detection systems. Sample
+frames from the video with successful detections are given in Figure
+\ref{fig:frames}.
+
+\subsubsection{Face detection
+algorithms}\label{face-detection-algorithms}
+
+\subsubsection{Detector evaluation}\label{detector-evaluation}
+
+\begin{table}[t]
+  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
+  \begin{center}
+    \begin{tabular}{l|ccc|ccc}
+      \hline
+       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
+      % \null Model & P & R & F & P & R & F  \\
+      \hline
+      % Viola-Jones & .55  & .38 &  .45 & .89 & .74 & .81   \\
+      % MTCNN & .86 & .78 & .81 & .93 & .76 & .83 \\
+      % OpenPose & .86 & .78 & .81 & .93 & .76 & .83 \\
+    \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+\section{Results}\label{results}
+
+We report results from three different sets of analyses. First, we
+explore developmental changes in posture and orientation in our dataset.
+Next, we explore how these changes affect access to faces and to hands,
+as measured using computater vision algorithms. Finally, we explore how
+these changes impact the accessibility of faces and hands during
+labeling events.
+
+\subsection{Changes in Posture and
+Orientation}\label{changes-in-posture-and-orientation}
+
+\subsection{Changes in Access to
+Faces}\label{changes-in-access-to-faces}
+
+\subsection{Changes in Access to
+Hands}\label{changes-in-access-to-hands}
 
@@
 
-``Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of
+Thanks to Kaia Simmons, Kathy Woo, Aditi Maliwal, and other members of
 the Language and Cognition Lab for help in recruitment, data collection,
@@
 Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford,
-CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}.''
+CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}.
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-adolph2012}{}
+Adolph, K., Gilmore, R., Freeman, C., Sanderson, P., \& Millman, D.
+(2012). Toward open behavioral science. \emph{Psychological Inquiry},
+\emph{23}(3), 244--247.
+
+\hypertarget{ref-cummings1988}{}
+Cummings, M., Van Hof-Van Duin, J., Mayer, D., Hansen, R., \& Fulton, A.
+(1988). Visual fields of young children. \emph{Behavioural and Brain
+Research}, \emph{29}(1), 7--16.
+
 \hypertarget{ref-frank2013}{}
@@
 
+\hypertarget{ref-mayer1988}{}
+Mayer, D., Fulton, A., \& Cummings, M. (1988). Visual fields of infants
+assessed with a new perimetric technique. \emph{Investigative
+Ophthalmology \& Visual Science}, \emph{29}(3), 452--459.
+
+\hypertarget{ref-yoshida2008}{}
+Yoshida, H., \& Smith, L. (2008). What's in view for toddlers? Using a
+head camera to study visual experience. \emph{Infancy}, \emph{13},
+229--248.
+
 \end{document}

### 2018-01-11 writeup-cogsci/cogsci2018.Rmd
face detection text

@@
 ## Face Detection
+Measuring the availability of caregiver’s faces across development was an important goal of the study.
+We used a total of three face detection systems to explore this hypothesis and to avoid the cost of hand-annotating every frame.
+Results for the evaluation of these three systems are shown in Table X.
 
-An additional goal of the study was to measure the presence of caregivers' faces in the child's field of view (as approximated by the headcam).
-To avoid hand-annotating the size and position of faces in every frame of video, we tested two face detection systems.
-Sample frames from the video with successful detections are given in Figure \ref{fig:frames}.
+### Face Detection Algorithms
+The first face detection system made use of a series of Haar feature-based cascade classifiers (viola) applied to each individual frame.
+This detector provided information about the presence of a face as well as its size and position.
 
-### Face detection algorithms
+The second algorithm was based on the work by () using multi-task cascaded convolutional neural networks (MTCNNs).
+The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
+We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet)
+
+The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc).
+The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
+The candidates are then matched to a single individual and finally assembled into a pose.
+For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose).
 

### 2018-01-11 writeup-cogsci/cogsci2018.Rmd
figures etc.

--- 2018-01-11 writeup-cogsci/cogsci2018.Rmd
@@
 ---
-title: "Developmental and postural changes in children's visual access to faces"
+title: "Postural changes mediate children's visual access to faces and hands"
 bibliography: xsface.bib
@@
 library(lubridate)
+library(langcog)
+library(ggplot2)
 ```
@@
 \includegraphics[width=6in]{images/framesample.pdf}
-\caption{\label{fig:frames} Sample frames from the headcam videos for a child from each age group, selected because they featured successful face detections (green squares).}
+\caption{\label{fig:frames} Example face detections made by MTCNN for the headcam videos from a child in each group  (green squares).}
 \end{figure*}
 
-All headcam videos were cropped to exclude the period of entry to the playroom and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software.
-These sessions yielded a substantial amount of video: a total of
+Headcam videos were trimmed such that they excluded the instruction phase when the experimenter was in the room and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software.
+These sessions yielded videos of
 `r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a milion frames), with an average video length of
@@
 We report results from three different sets of analyses.
-First, we explore developmental changes in posture and orientation in our dataset.
-Next, we explore how these changes affect access to faces and to hands, as measured using computater vision algorithms.
-Finally, we explore how these changes impact the accessibility of faces and hands during labeling events.
+First, we explore how these changes affect access to faces and to hands, as measured using computater vision algorithms.
+Second, we explore how these changes impact the accessibility of faces and hands during labeling events.
 
-## Changes in Posture and Orientation
+```{r}
+d <- read_csv("../data/consolidated_data.csv") %>%
+  mutate(posture = factor(posture), 
+         orientation = factor(orientation))
 
-## Changes in Access to Faces
+facebyAge <- d %>%
+  filter(!is.na(age.at.test)) %>%
+  group_by(subid, age.at.test) %>%
+  summarise(face = mean(face)) %>%
+  left_join(inc_subs) %>%
+  mutate(len = as.numeric(len_seconds/60))
 
-## Changes in Access to Hands
+faceByOrientation <- d %>%
+  filter(!is.na(orientation)) %>%
+  group_by(age.grp, orientation, subid) %>%
+  summarise(face = mean(face)) %>%
+  group_by(age.grp)
+
+facebyPosture <- d %>%
+  filter(!is.na(posture)) %>%
+  group_by(age.grp, posture, subid) %>%
+  summarise(face = mean(face)) %>%
+  group_by(age.grp)
+```
+
+# Changes in Access to Faces
+
+```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
+ggplot(facebyAge, aes(x = age.at.test, y = face)) + 
+  geom_jitter(width = .25, aes(size = len)) + 
+  geom_smooth() + 
+  scale_x_continuous(breaks=c(8,12,16)) +
+  scale_size_continuous(name = "Length (s)") + 
+  ylab("Proportion Face Detections") + 
+  xlab("Age (months)") + 
+  theme_bw() 
+```
+
+```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
+ggplot(facebyPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
+  geom_boxplot() + 
+  ylab("Proportion Face Detections") + 
+  xlab("Age (months)") + 
+  theme_bw()
+```
+
+```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
+ggplot(faceByOrientation, aes(x = factor(age.grp), y = face, col = orientation)) + 
+  geom_boxplot() + 
+  ylab("Proportion Face Detections") + 
+  xlab("Age (months)") + 
+  theme_bw() 
+```
+
+# Changes in Access to Hands
 

### 2018-01-11 writeup-cogsci/cogsci2018.Rmd
few edits to face detection code & model starts

@@
 library(ggplot2)
+library(lme4)
 ```
@@
 ### Face Detection Algorithms
-The first face detection system made use of a series of Haar feature-based cascade classifiers (viola) applied to each individual frame.
+The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, CITEXX) applied to each individual frame.
 This detector provided information about the presence of a face as well as its size and position.
 
-The second algorithm was based on the work by () using multi-task cascaded convolutional neural networks (MTCNNs).
+The second algorithm was based on the work by (CITEXX) using multi-task cascaded convolutional neural networks (MTCNNs).
 The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
@@
 
-The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc).
+The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (CITEXX).
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
@@
 
+## Model
+```{r}
+
+fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), 
+      data = d,  
+      family = "binomial")
+
+kable(summary(mod_covariates_age_v2)$coef, digits = 3)
+
+```
+
 # Changes in Access to Hands

### 2018-01-11 writeup-cogsci/cogsci2018.tex

@@
 
-\title{Developmental and postural changes in children's visual access to faces}
+\title{Postural changes mediate children's visual access to faces and hands}
 
@@
 \includegraphics[width=6in]{images/framesample.pdf}
-\caption{\label{fig:frames} Sample frames from the headcam videos for a child from each age group, selected because they featured successful face detections (green squares).}
+\caption{\label{fig:frames} Example face detections made by MTCNN for the headcam videos from a child in each group  (green squares).}
 \end{figure*}
 
-All headcam videos were cropped to exclude the period of entry to the
-playroom and were automatically synchronized with the tripod-mounted
-videos using FinalCut Pro Software. These sessions yielded a substantial
-amount of video: a total of 516 minutes (almost a milion frames), with
-an average video length of 8.6 minutes (min = 4.53, max = 19.35).
+Headcam videos were trimmed such that they excluded the instruction
+phase when the experimenter was in the room and were automatically
+synchronized with the tripod-mounted videos using FinalCut Pro Software.
+These sessions yielded videos of 516 minutes (almost a milion frames),
+with an average video length of 8.6 minutes (min = 4.53, max = 19.35).
 
@@
 We report results from three different sets of analyses. First, we
-explore developmental changes in posture and orientation in our dataset.
-Next, we explore how these changes affect access to faces and to hands,
-as measured using computater vision algorithms. Finally, we explore how
+explore how these changes affect access to faces and to hands, as
+measured using computater vision algorithms. Second, we explore how
 these changes impact the accessibility of faces and hands during
@@
 
-\subsection{Changes in Posture and
-Orientation}\label{changes-in-posture-and-orientation}
+\section{Changes in Access to Faces}\label{changes-in-access-to-faces}
 
-\subsection{Changes in Access to
-Faces}\label{changes-in-access-to-faces}
+\begin{CodeChunk}
+\begin{figure}[H]
 
-\subsection{Changes in Access to
-Hands}\label{changes-in-access-to-hands}
+{\centering \includegraphics{figs/unnamed-chunk-3-1}
+
+}
+
+\caption[R plot]{R plot}\label{fig:unnamed-chunk-3}
+\end{figure}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure}[H]
+
+{\centering \includegraphics{figs/unnamed-chunk-4-1}
+
+}
+
+\caption[R plot]{R plot}\label{fig:unnamed-chunk-4}
+\end{figure}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure}[H]
+
+{\centering \includegraphics{figs/unnamed-chunk-5-1}
+
+}
+
+\caption[R plot]{R plot}\label{fig:unnamed-chunk-5}
+\end{figure}
+\end{CodeChunk}
+
+\section{Changes in Access to Hands}\label{changes-in-access-to-hands}
 

### 2018-01-12 analysis-validation/mtcnn_evaluation.Rmd
random edits and new gold set candidates

@@
 500 frames were taken across all groups, half of them labeled as containing a face and half with no face.
+```{r}
+library(tidyverse)
+library(magrittr)
+```
 
@@
 
+For openpose only 
+```{r}
+data <- read_csv("data/final_output/openpose_results_truncated.csv") %>%
+  distinct(name, frame, .keep_all = TRUE)
+```
+
+```{r}
+contains_face <- function(n, l, r) {
+  return(n != 0)
+}
+```
+
+```{r}
+data2 <- data %>%
+  mutate(is_face = (LEye_conf != 0 | REye_conf != 0) | Nose_conf != 0) %>%
+  select(group, name, frame, is_face) %>%
+  rename(video = name)
+```
+
 Join
 ```{r}
-combined <- data %>%
+combined <- data2 %>%
   rename(detector_is_face = is_face) %>%
   inner_join(ground_truth_data, by = c("group", "video", "frame"))
+```
+openpose, annoying
+```{r}
+combined %<>% 
+  mutate(detector_is_face = ifelse(detector_is_face == TRUE, "True", "False"))
 ```

### 2018-01-12 analysis-validation/detector_evaluation.Rmd
hand validation

---
title: "detector evaluation"
output: html_document
---

504 frames were taken across all groups, half of them labeled as containing a face and half with no face.
```{r}
library(tidyverse)
library(magrittr)
```

Select detector to test:

MTCNN
```{r}
setwd("xs-face")
data <- read_csv("data/final_output/mtcnn3.csv") %>%
  distinct(video, frame, .keep_all = TRUE)

original_candidates <- read_csv("data/ground_truth/gold_set_candidates2.csv") %>%
  arrange(video, frame)

ground_truth_data <- read_csv("data/ground_truth/ground_truth3.csv") %>%
  select(-angle) %>%
  arrange(video, frame) %>%
  cbind(original_candidates %>%
          select(sample_type))
```

OPENPOSE
```{r}
setwd("xs-face")
data <- read_csv("data/final_output/openpose_results_truncated.csv") %>%
  distinct(name, frame, .keep_all = TRUE)

data %<>%
  mutate(is_face = ifelse(Nose_conf != 0, "True", "False"))%>%
  select(group, name, frame, is_face) %>%
  rename(video = name)

original_candidates <- read_csv("data/ground_truth/gold_set_candidates2.csv") %>%
  arrange(video, frame)

ground_truth_data <- read_csv("data/ground_truth/ground_truth3.csv") %>%
  select(-angle) %>%
  arrange(video, frame) %>%
  cbind(original_candidates %>%
          select(sample_type))
```

OPENPOSE HANDS
```{r}
setwd("..")
data <- read_csv("data/final_output/openpose_results_truncated.csv") %>%
  distinct(name, frame, .keep_all = TRUE)

data %<>%
  mutate(is_face = ifelse(LWrist_conf != 0 | RWrist_conf != 0, "True", "False"))%>%
  select(group, name, frame, is_face) %>%
  rename(video = name)

ground_truth_data <- read_csv("data/ground_truth/ground_truth_hands.csv") %>%
  select(-angle) %>%
  arrange(video, frame)
```

Join data
```{r}
combined <- data %>%
  rename(detector_is_face = is_face) %>%
  inner_join(ground_truth_data, by = c("group", "video", "frame"))
```

Eval function
```{r}
evaluate <- function(a, b) {
  if (a == "True") {
    if (a == b) return ("TP")
    else return("FN")
  }
  else {
    if (a == b) return("TN")
    else return("FP")
  }
}
```

Evaluate
```{r}
results <- combined %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face))
```

Precison / Recall / F-score
```{r}
score <- results %>%
  group_by(sample_type, result) %>%
  summarise(count = n())

# p <- score[4,2] / (score[4,2] + score[2,2])
# r <- score[4,2] / (score[4,2] + score[1,2])
# f <- 2 * p * r / (p + r)
# 
# print(paste("P:", round(p,3), "R:", round(r,3), "F:", round(f,3)))
```

### 2018-01-12 scripts/ground_truth_creation/get_gold_set_hand_images.Rmd
hand validation

---
title: "get_gold_set_hand_images"
output: html_document
---

```{r}
library(tidyverse)
library(magrittr)
```

```{r}
setwd("/Users/alessandro/Github/xs-face")
data <- read_csv("data/final_output/openpose_results_truncated.csv")
```

```{r}
data2 <- data %>% 
  rename(video = name) %>%
  distinct(video, frame, .keep_all = TRUE)
```

Sample high density faces, remove selected from original data
```{r}
high_density_hands_sample <- data2 %>%
  group_by(video) %>%
  filter(LWrist_conf != 0 | RWrist_conf != 0) %>%
  sample_n(7) %>%
  mutate(sample_type = "high_density")

data3 <- data2 %>%
  anti_join(high_density_hands_sample)
```

Choose random frames from remaining
```{r}
random_sample <- data3 %>%
  group_by(video) %>%
  sample_n(7) %>%
  mutate(sample_type = "random")
```

Combine and save result
```{r}
gold_set_candidates <- high_density_hands_sample %>%
  rbind(random_sample) %>%
  mutate(path = paste0(video, "/image-", frame, ".jpg")) %>%
  arrange(video, frame)

write_csv(gold_set_candidates, "data/ground_truth/gold_set_hand_candidates.csv")
```

### 2018-01-12 scripts/ground_truth_creation/download.py

@@ -1,10 +1,10 @@
 import os
 import csv
 
-with open("../../data/ground_truth/gold_set_candidates.csv", 'rb') as img_paths_f:
+with open("../../data/ground_truth/gold_set_hand_candidates.csv", 'rb') as img_paths_f:
         img_paths = list(csv.reader(img_paths_f, delimiter=','))
 
 for path in img_paths[1:]:
-    os.system("mkdir -p ~/xsface/sample_images3/%s" % (path[1]))
-    os.system("scp sanchez7@sherlock.stanford.edu:/share/PI/mcfrank/frames4/%s ~/xsface/sample_images3/%s" % (path[8], path[1]))
+    os.system("mkdir -p ~/xsface/sample_hand_images/%s" % (path[1]))
+    os.system("scp sanchez7@sherlock.stanford.edu:/share/PI/mcfrank/frames4/%s ~/xsface/sample_hand_images/%s" % (path[58], path[1]))
 

### 2018-01-12 scripts/openpose/json_to_csv.py

@@ -6,6 +6,7 @@ import ntpath
 
 CSV_OUTPUT_DIR = os.path.expandvars("$HOME/xs-face/scripts/openpose/tmp")
 
+# TODO 1 index
 # TODO 1229 is weird because title contains "cropped"
 # TODO if no face or hand or whatever, fill in with 0s!!! first check if, and if not, extend with 0s
 

### 2018-01-12 analysis-main/merge_all_data.R
reorg

@@ -12,23 +12,38 @@ dets <- read_csv("../data/final_output/mtcnn3.csv") %>%
   mutate(subid = video) %>%
   mutate(frame = as.numeric(frame)) %>%
   mutate(faceMT = as.logical(is_face)) %>%
+<<<<<<< HEAD
+  distinct(video,frame,.keep_all=TRUE)   %>%
+
+# open pose detectors
+detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") 
+=======
   distinct(video,frame,.keep_all=TRUE) 
 
 # open pose detectors
 detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated.csv") 
+>>>>>>> ef16392816084c123fa7a598eeda72e854cb4677
 detsOpenPose <- detsOpenPose %>%
   mutate(video = name) %>%
   distinct(video,frame,.keep_all=TRUE)  %>%
   mutate(frame = as.numeric(frame)) %>%
   mutate(faceOP = Nose_conf!=0 & REye_conf!=0 | LEye_conf!=0 )  %>%
+<<<<<<< HEAD
+  mutate(handOP = LWrist_conf!=0 | RWrist_conf!=0 )   %>%
+=======
   mutate(handOP = LWrist_conf!=0 | RWrist_conf!=0 )   
+>>>>>>> ef16392816084c123fa7a598eeda72e854cb4677
   
 # viola jones detectors
 detsViola <- read_csv("../data/final_output/viola.csv") 
 detsViola <- detsViola %>%
   distinct(video,frame,.keep_all=TRUE)  %>%
   mutate(faceVJ = as.logical(is_face))  %>%
+<<<<<<< HEAD
+  mutate(frame = as.numeric(frame)) %>%
+=======
   mutate(frame = as.numeric(frame))
+>>>>>>> ef16392816084c123fa7a598eeda72e854cb4677
 
 # merge all three detectors
 alldets=left_join(dets,detsViola[,c("video","frame","faceVJ")]) 
@@ -41,9 +56,27 @@ demo.data <- read_csv("../data/demographics/demographics.csv") %>%
 # rearrange so we make sure it is in the order of the frames
 alldets<-arrange(alldets,video,frame) 
 
+<<<<<<< HEAD
+## OpenPose doesn't get all frames. Let's check out the frames it deleted somehow.
+missingInd=is.na(alldets$faceOP)
[…]

### 2018-01-12 analysis-main/main_analyses_bll.Rmd
hope this works

@@
 options(error = browser)
+
 ```
@@
 library(readr)
-library(plyr) # need to load this for scripts, always load before
+library(langcog)
 library(dplyr)
 library(tidyr)
-library(langcog)
 library(ggplot2)
@@
 ```{r}
-d <- read_csv("../data/consolidated_data.csv") %>%
+d <- read_csv("../data/consolidated_data_3dets.csv") %>%
   mutate(posture = factor(posture), 
@@
 ```{r}
+#d$face=d$faceVJ
+#d$face=d$faceMT
+d$face=d$faceOP
+#d$face=d$handOP
 
@@
 
-# Namings and face detections
+<!-- # Namings and face detections -->
 
@@
 ```{r}
-sb <- plyr::ddply(d,~subid, summarize.naming, window=c(-2,0)) ## used 2 second window as in ms draft
-sa <- plyr::ddply(d,~subid, summarize.naming, window=c(0,2))
+aroundNamings <- d %>%
+  filter(!is.na(posture)) %>%
+  group_by(subid) %>%
+  do(summarize.naming(.))
 
-sb$window <- "before"
-sa$window <- "after"
-s <- rbind(sb,sa)
-s$window <- factor(s$window)
+ggplot(aroundNamings, aes(x = factor(age.grp), y = face, col = posture)) + 
+  geom_boxplot() + 
+  ylab("detections within naming window") + 
+  xlab("Age (months)") + 
+  facet_grid(.~familiarity) +
+  theme_bw()
 
-# not used, commented out
-# s$first.instance <- s$naming.instance == 1
-# s$log.instance <- round(log10(s$naming.instance)*4)/4
-# s$binned.instance <- 10^(round(log10(s$naming.instance)*4)/4)
-# s$face.logical <- s$face != 0
+```{r}
+# mss <- d %>%
+#   filter(!is.na(posture)) %>%
+#   filter(!is.na(face)) %>%
+#   group_by(age.grp, posture, subid) %>%
+#   summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
+#   group_by(age.grp)
+# 
+# 
+# sb_summary <- sb %>%
+#   filter(!is.na(posture)) %>%
+#   filter(!is.na(face)) %>%
+#   group_by(age.grp, posture, subid) %>%
+#   dplyr::summarise(face = count(face)) %>%
+#   group_by(age.grp)
+# 
+# 
+# nameFreq <- sb %>%
+#   filter(!is.na(posture)) %>%
+#   filter(!is.na(face)) %>%
+#   group_by(age.grp, posture, subid) %>%
+#   dplyr::summarise(nameFreq= max(naming.instance, na.rm="TRUE")) %>%
+#   group_by(age.grp)
+# 
+# ggplot(nameFreq, aes(x = factor(age.grp), y = nameFreq, col = posture)) + 
+#   geom_boxplot() + 
+#   ylab("within window") + 
+#   xlab("Age (months)") + 
+#   theme_bw() 
+# 
+# 
+# # ggplot(sb_summary, aes(x = factor(age.grp), y = face, col = posture)) + 
+# #   geom_boxplot() + 
+# #   ylab("within window") + 
+# #   xlab("Age (months)") + 
+# #   theme_bw()
 
-## try plotting individuals
-mss <- ddply(s,~subid+age.grp+window,summarise,
-               face = mean(face, na.rm="TRUE"))
+sb <- sb %>%
+  filter(!is.na(posture))
+
+ggplot(sb, aes(x = factor(age.grp), y = face, col = posture)) + 
+  geom_boxplot() + 
+  ylab("within window") + 
+  xlab("Age (months)") + 
+  facet_grid(.~familiarity) +
+  theme_bw()
+
 ```
@@
 ms <- namings %>%
-  group_by(age.grp, subid, naming.instance, familiarity, name) %>%
-  summarise(face = mean(face)) %>%
-  group_by(age.grp, naming.instance, familiarity) %>%
+  dplyr::group_by(age.grp, subid, naming.instance, familiarity, name) %>%
+  dplyr::summarise(face = mean(face)) %>%
+  dplyr::group_by(age.grp, naming.instance, familiarity) %>%
   multi_boot_standard(col = "face", na.rm = TRUE)
@@
 
-## Now calculate face detection differences from base rate
-
-```{r}
-
-  n <- ddply(s,~subid+age.grp,summarise,
-             face = mean(face, na.rm="TRUE"))
-
-  n$br <- ddply(d,~subid+age.grp,summarise,
-              face=mean(face, na.rm="TRUE"))$face
-
-  n$br.diff <- n$face - n$br
-
-#   pdf("~/Projects/xsface/writeup/figures/naming_faces_diff.pdf",width=4,height=3)
-  qplot(age.grp,br.diff,position=position_jitter(.2),
-        data=subset(n,age.grp>4),xlab="Age (Months)",ylab="Face detections relative to base rate") +
-    scale_x_continuous(breaks=c(8,12,16)) +
-    geom_abline(slope=0,intercept=0,lty=2) +
-    stat_summary(fun.data="mean_cl_boot",color="red")
-#   dev.off()
-```
-
-<!-- # CDIs, namings, and face detections -->
+<!-- ## Now calculate face detection differences from base rate -->
 
 <!-- ```{r} -->
-<!-- cdi_namings <- left_join(naming_rate, cdis) -->
 
-<!-- qplot(comprehension, facets= .~ age.grp, data = cdi_namings) + -->
-<!--   theme_bw() -->
+<!--   n <- ddply(s,~subid+age.grp,summarise, -->
+<!--              face = mean(face, na.rm="TRUE")) -->
 
-<!-- qplot(production, facets= .~ age.grp, data = cdi_namings) + -->
-<!--   theme_bw() -->
+<!--   n$br <- ddply(d,~subid+age.grp,summarise, -->
+<!--               face=mean(face, na.rm="TRUE"))$face -->
 
+<!--   n$br.diff <- n$face - n$br -->
+
+<!-- #   pdf("~/Projects/xsface/writeup/figures/naming_faces_diff.pdf",width=4,height=3) -->
+<!--   qplot(age.grp,br.diff,position=position_jitter(.2), -->
+<!--         data=subset(n,age.grp>4),xlab="Age (Months)",ylab="Face detections relative to base rate") + -->
+<!--     scale_x_continuous(breaks=c(8,12,16)) + -->
+<!--     geom_abline(slope=0,intercept=0,lty=2) + -->
+<!--     stat_summary(fun.data="mean_cl_boot",color="red") -->
+<!-- #   dev.off() -->
 <!-- ``` -->
 
-<!-- ## CDIs and namings -->
+<!-- <!-- # CDIs, namings, and face detections --> -->
 
-<!-- Just out of curiousity, are there any major differences in terms of CDIs by naming rate? (Would help to establish the validity of the naming rate analyses). -->
+<!-- <!-- ```{r} --> -->
+<!-- <!-- cdi_namings <- left_join(naming_rate, cdis) --> -->
 
-<!-- ```{r} -->
-<!-- ggplot(cdi_namings, -->
-<!--        aes(x = rate, y = comprehension)) + -->
-<!--   geom_point() + -->
-<!--   geom_smooth(method = "lm", se = FALSE) + -->
-<!--   facet_grid(.~age.grp) + -->
-<!--   xlab("Namings per minute") + -->
-<!--   ylab("CDI Summed Comprehension Vocabulary") + -->
-<!--   theme_bw() -->
-<!-- ``` -->
+<!-- <!-- qplot(comprehension, facets= .~ age.grp, data = cdi_namings) + --> -->
+<!-- <!--   theme_bw() --> -->
 
-<!-- ```{r} -->
-<!-- cdis <- read_csv("../data/cdis/cdi_summary.csv") -->
-<!-- left_join(naming_rate, cdis) %>% -->
-<!--   ggplot(aes(x = rate, y = production)) + -->
-<!--   geom_point() + -->
-<!--   geom_smooth(method = "lm", se = FALSE) + -->
-<!--   facet_grid(.~age.grp) + -->
-<!--   xlab("Namings per minute") + -->
-<!--   ylab("CDI Summed Production Vocabulary") + -->
-<!--   theme_bw() -->
-<!-- ``` -->
+<!-- <!-- qplot(production, facets= .~ age.grp, data = cdi_namings) + --> -->
+<!-- <!--   theme_bw() --> -->
+
+<!-- <!-- ``` --> -->
+
+<!-- <!-- ## CDIs and namings --> -->
+
+<!-- <!-- Just out of curiousity, are there any major differences in terms of CDIs by naming rate? (Would help to establish the validity of the naming rate analyses). --> -->
+
+<!-- <!-- ```{r} --> -->
+<!-- <!-- ggplot(cdi_namings, --> -->
+<!-- <!--        aes(x = rate, y = comprehension)) + --> -->
+<!-- <!--   geom_point() + --> -->
+<!-- <!--   geom_smooth(method = "lm", se = FALSE) + --> -->
+<!-- <!--   facet_grid(.~age.grp) + --> -->
+<!-- <!--   xlab("Namings per minute") + --> -->
+<!-- <!--   ylab("CDI Summed Comprehension Vocabulary") + --> -->
+<!-- <!--   theme_bw() --> -->
+<!-- <!-- ``` --> -->
+
+<!-- <!-- ```{r} --> -->
+<!-- <!-- cdis <- read_csv("../data/cdis/cdi_summary.csv") --> -->
+<!-- <!-- left_join(naming_rate, cdis) %>% --> -->
+<!-- <!--   ggplot(aes(x = rate, y = production)) + --> -->
+<!-- <!--   geom_point() + --> -->
+<!-- <!--   geom_smooth(method = "lm", se = FALSE) + --> -->
+<!-- <!--   facet_grid(.~age.grp) + --> -->
+<!-- <!--   xlab("Namings per minute") + --> -->
+<!-- <!--   ylab("CDI Summed Production Vocabulary") + --> -->
+<!-- <!--   theme_bw() --> -->
+<!-- <!-- ``` --> -->

### 2018-01-12 analysis-main/merge_all_data.R

@@ -12,38 +12,23 @@ dets <- read_csv("../data/final_output/mtcnn3.csv") %>%
   mutate(subid = video) %>%
   mutate(frame = as.numeric(frame)) %>%
   mutate(faceMT = as.logical(is_face)) %>%
-<<<<<<< HEAD
   distinct(video,frame,.keep_all=TRUE)   %>%
 
 # open pose detectors
 detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") 
-=======
-  distinct(video,frame,.keep_all=TRUE) 
-
-# open pose detectors
-detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated.csv") 
->>>>>>> ef16392816084c123fa7a598eeda72e854cb4677
 detsOpenPose <- detsOpenPose %>%
   mutate(video = name) %>%
   distinct(video,frame,.keep_all=TRUE)  %>%
   mutate(frame = as.numeric(frame)) %>%
   mutate(faceOP = Nose_conf!=0 & REye_conf!=0 | LEye_conf!=0 )  %>%
-<<<<<<< HEAD
   mutate(handOP = LWrist_conf!=0 | RWrist_conf!=0 )   %>%
-=======
-  mutate(handOP = LWrist_conf!=0 | RWrist_conf!=0 )   
->>>>>>> ef16392816084c123fa7a598eeda72e854cb4677
   
 # viola jones detectors
 detsViola <- read_csv("../data/final_output/viola.csv") 
 detsViola <- detsViola %>%
   distinct(video,frame,.keep_all=TRUE)  %>%
   mutate(faceVJ = as.logical(is_face))  %>%
-<<<<<<< HEAD
   mutate(frame = as.numeric(frame)) %>%
-=======
-  mutate(frame = as.numeric(frame))
->>>>>>> ef16392816084c123fa7a598eeda72e854cb4677
 
 # merge all three detectors
 alldets=left_join(dets,detsViola[,c("video","frame","faceVJ")]) 
@@ -56,7 +41,6 @@ demo.data <- read_csv("../data/demographics/demographics.csv") %>%
 # rearrange so we make sure it is in the order of the frames
 alldets<-arrange(alldets,video,frame) 
 
-<<<<<<< HEAD
 ## OpenPose doesn't get all frames. Let's check out the frames it deleted somehow.
 missingInd=is.na(alldets$faceOP)
 assert_that(sum(is.na(alldets$faceOP))==57)
[…]

### 2018-01-13 analysis-main/merge_all_data.R
fixed merge data file

@@ -12,7 +12,7 @@ dets <- read_csv("../data/final_output/mtcnn3.csv") %>%
   mutate(subid = video) %>%
   mutate(frame = as.numeric(frame)) %>%
   mutate(faceMT = as.logical(is_face)) %>%
-  distinct(video,frame,.keep_all=TRUE)   %>%
+  distinct(video,frame,.keep_all=TRUE)   
 
 # open pose detectors
 detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") 
@@ -21,14 +21,14 @@ detsOpenPose <- detsOpenPose %>%
   distinct(video,frame,.keep_all=TRUE)  %>%
   mutate(frame = as.numeric(frame)) %>%
   mutate(faceOP = Nose_conf!=0 & REye_conf!=0 | LEye_conf!=0 )  %>%
-  mutate(handOP = LWrist_conf!=0 | RWrist_conf!=0 )   %>%
+  mutate(handOP = LWrist_conf!=0 | RWrist_conf!=0 )   
   
 # viola jones detectors
 detsViola <- read_csv("../data/final_output/viola.csv") 
 detsViola <- detsViola %>%
   distinct(video,frame,.keep_all=TRUE)  %>%
   mutate(faceVJ = as.logical(is_face))  %>%
-  mutate(frame = as.numeric(frame)) %>%
+  mutate(frame = as.numeric(frame)) 
 
 # merge all three detectors
 alldets=left_join(dets,detsViola[,c("video","frame","faceVJ")]) 
@@ -68,8 +68,9 @@ d <- alldets %>%
 
 ages <- d %>%
   group_by(subid) %>%
-  summarise(age.grp = mean(age.grp))
-  complete_combos <- expand(d, nesting(posture, orientation), 
+  summarise(age.grp = mean(age.grp)) 
+
+complete_combos <- expand(d, nesting(posture, orientation), 
                           subid) %>% 
   mutate(dt = 0, faceMT = FALSE, faceVJ = FALSE, faceOP = FALSE, handOP = FALSE) %>%
   left_join(ages)
@@ -77,5 +78,5 @@ ages <- d %>%
 d <- bind_rows(d, complete_combos)
 
 ## save it out
-write_csv(d, "../data/consolidated_data_3dets.csv")
+write_csv(d, "../data/consolidateddata/consolidated_data_3dets.csv")
 

### 2018-01-13 analysis-main/main_analyses_bll.Rmd
discussion intro integrated

@@
 ```{r}
-d <- read_csv("../data/consolidated_data_3dets.csv") %>%
+d <- read_csv("../data/consolidateddata/consolidated_data_3dets.csv") %>%
   mutate(posture = factor(posture), 
@@
 
-```{r}
-# mss <- d %>%
-#   filter(!is.na(posture)) %>%
-#   filter(!is.na(face)) %>%
-#   group_by(age.grp, posture, subid) %>%
-#   summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
-#   group_by(age.grp)
-# 
-# 
-# sb_summary <- sb %>%
-#   filter(!is.na(posture)) %>%
-#   filter(!is.na(face)) %>%
-#   group_by(age.grp, posture, subid) %>%
-#   dplyr::summarise(face = count(face)) %>%
-#   group_by(age.grp)
-# 
-# 
-# nameFreq <- sb %>%
-#   filter(!is.na(posture)) %>%
-#   filter(!is.na(face)) %>%
-#   group_by(age.grp, posture, subid) %>%
-#   dplyr::summarise(nameFreq= max(naming.instance, na.rm="TRUE")) %>%
-#   group_by(age.grp)
-# 
-# ggplot(nameFreq, aes(x = factor(age.grp), y = nameFreq, col = posture)) + 
-#   geom_boxplot() + 
-#   ylab("within window") + 
-#   xlab("Age (months)") + 
-#   theme_bw() 
-# 
-# 
-# # ggplot(sb_summary, aes(x = factor(age.grp), y = face, col = posture)) + 
-# #   geom_boxplot() + 
-# #   ylab("within window") + 
-# #   xlab("Age (months)") + 
-# #   theme_bw()
-
-sb <- sb %>%
-  filter(!is.na(posture))
-
-ggplot(sb, aes(x = factor(age.grp), y = face, col = posture)) + 
-  geom_boxplot() + 
-  ylab("within window") + 
-  xlab("Age (months)") + 
-  facet_grid(.~familiarity) +
-  theme_bw()
-
 ```
 
-```{r}
-  qplot(age.grp,face,colour=window,
-        xlab="Age (Months)",ylab="Proportion faces detected in window",
-        data=mss,geom=c("point")) +
-    scale_x_continuous(breaks=c(8,12,16)) +
-    geom_smooth()
-```
+<!-- # mss <- d %>% -->
+<!-- #   filter(!is.na(posture)) %>% -->
+<!-- #   filter(!is.na(face)) %>% -->
+<!-- #   group_by(age.grp, posture, subid) %>% -->
+<!-- #   summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>% -->
+<!-- #   group_by(age.grp) -->
+<!-- #  -->
+<!-- #  -->
+<!-- # sb_summary <- sb %>% -->
+<!-- #   filter(!is.na(posture)) %>% -->
+<!-- #   filter(!is.na(face)) %>% -->
+<!-- #   group_by(age.grp, posture, subid) %>% -->
+<!-- #   dplyr::summarise(face = count(face)) %>% -->
+<!-- #   group_by(age.grp) -->
+<!-- #  -->
+<!-- #  -->
+<!-- # nameFreq <- sb %>% -->
+<!-- #   filter(!is.na(posture)) %>% -->
+<!-- #   filter(!is.na(face)) %>% -->
+<!-- #   group_by(age.grp, posture, subid) %>% -->
+<!-- #   dplyr::summarise(nameFreq= max(naming.instance, na.rm="TRUE")) %>% -->
+<!-- #   group_by(age.grp) -->
+<!-- #  -->
+<!-- # ggplot(nameFreq, aes(x = factor(age.grp), y = nameFreq, col = posture)) +  -->
+<!-- #   geom_boxplot() +  -->
+<!-- #   ylab("within window") +  -->
+<!-- #   xlab("Age (months)") +  -->
+<!-- #   theme_bw()  -->
+<!-- #  -->
+<!-- #  -->
+<!-- # # ggplot(sb_summary, aes(x = factor(age.grp), y = face, col = posture)) +  -->
+<!-- # #   geom_boxplot() +  -->
+<!-- # #   ylab("within window") +  -->
+<!-- # #   xlab("Age (months)") +  -->
+<!-- # #   theme_bw()  -->
+
+<!-- sb <- sb %>% -->
+<!--   filter(!is.na(posture))  -->
+
+<!-- ggplot(sb, aes(x = factor(age.grp), y = face, col = posture)) +  -->
+<!--   geom_boxplot() +  -->
+<!--   ylab("within window") +  -->
+<!--   xlab("Age (months)") +  -->
+<!--   facet_grid(.~familiarity) + -->
+<!--   theme_bw()  -->
+
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!--   qplot(age.grp,face,colour=window, -->
+<!--         xlab="Age (Months)",ylab="Proportion faces detected in window", -->
+<!--         data=mss,geom=c("point")) + -->
+<!--     scale_x_continuous(breaks=c(8,12,16)) + -->
+<!--     geom_smooth() -->
+<!-- ``` -->
 

### 2018-01-13 writeup-cogsci-copy/cogsci2018.Rmd
discussion intro integrated

--- 2018-01-11 writeup-cogsci/cogsci2018.Rmd
@@
 ---
-title: "Postural changes mediate children's visual access to faces and hands"
+title: "Postural changes mediate children's visual access to social information"
 bibliography: xsface.bib
@@
 abstract: 
+The aiblity to rpocess social cues (e.g., eye gaze) is a critical component in children/s early language development.
+However, As children reach their first birthday, they begin to locomote themselves, exploring their visual environment in an entirely new way.
+How do these postural changes affect children's access to the social information relevant for word-learning?
 
@@
 library(readr)
-library(plyr) # need to load this for scripts, always load before
 library(dplyr)
@@
 ```
+# Introduction
+Children are remarkably skilled language learners, connecting arbitrary labels (“cup”) with specific visual concepts at a rapid pace through the first two years of life.
+However, children do not learn words in a vacuum but rather in a rich social environment, where social cues provided by speakers (e.g., eye-gaze) provide strong scaffolding for this learning process.
+Further, children’s ability to effectively process these social cues may be a key factor in their early language development.
+In a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies (@carpenter1998).
+More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months (@brooks2005) and throughout the second year of life (@brooks2008).
+
+However, as children are learning their first words, their view of the world is also changing radically [@adolph2007].
+Infant’s motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see.
+During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011), when they have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).
+
+One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities (Iverson, 2010).
+Indeed, these postural changes also impact how children interact with their mothers;  walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”).
+(Karasik et al., 2014).
+More directly, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
+On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access the social information (e.g., faces, gaze) relative to children who are still crawling and sitting.
+In other words, the ability to access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.
+
+Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different than the adult perspective (and not easily predicted by our own adult intuitions; @franchak2011, @yoshida2008; @clerkin2017).
+Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning what objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017].
+Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2017].
+
+Here, we directly ask whether the postural changes that infants’ experience as they reach their first birthday change the availability of the social information relevant for word learning.
+To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and their orientation to their caregiver were also recorded from a third-person perspective and hand-annotated.
+We capitalize on recent improvements in computer vision algorithms (@citemtcnn, @openpose) that allow the automated detection of both faces and hands, and we analyze the frequency of both faces and hands in the child’s visual environment.
 
@@
       % MTCNN & .86 & .78 & .81 & .93 & .76 & .83 \\
-      % OpenPose & .86 & .78 & .81 & .93 & .76 & .83 \\
+      % OpenPose-Faces & .86 & .78 & .81 & .93 & .76 & .83 \\
+      % OpenPose-Wrists & .86 & .78 & .81 & .93 & .76 & .83 \\
+      % OpenPose-Hands & .86 & .78 & .81 & .93 & .76 & .83 \\
     \hline
@@
 ```{r}
-d <- read_csv("../data/consolidated_data.csv") %>%
-  mutate(posture = factor(posture), 
-         orientation = factor(orientation))
+d <- read_csv("../data/consolidateddata/consolidated_data_3dets.csv") %>%
+  mutate(posture = factor(posture), orientation = factor(orientation))
 
-facebyAge <- d %>%
+## make long form version for plotting
+dlong <- d %>%
+  mutate(id = paste(subid,frame)) %>%
+  gather(key = detectorType, value = face, faceMT, faceOP)
+
+facebyAge <- dlong %>%
   filter(!is.na(age.at.test)) %>%
-  group_by(subid, age.at.test) %>%
+  group_by(subid, age.at.test, detectorType) %>%
   summarise(face = mean(face)) %>%
@@
 
-faceByOrientation <- d %>%
+faceByOrientation <- dlong %>%
   filter(!is.na(orientation)) %>%
-  group_by(age.grp, orientation, subid) %>%
+  group_by(age.grp, orientation, subid, detectorType) %>%
   summarise(face = mean(face)) %>%
@@
 
-facebyPosture <- d %>%
+facebyPosture <- dlong %>%
   filter(!is.na(posture)) %>%
-  group_by(age.grp, posture, subid) %>%
+  group_by(age.grp, posture, subid,detectorType) %>%
   summarise(face = mean(face)) %>%
@@
 ```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
-ggplot(facebyAge, aes(x = age.at.test, y = face)) + 
-  geom_jitter(width = .25, aes(size = len)) + 
+
+  ggplot(facebyAge, aes(x = age.at.test, y = face, color = detectorType, size = len)) +
+    geom_jitter(width = .25) + 
     geom_smooth() +
     scale_x_continuous(breaks=c(8,12,16)) +
-  scale_size_continuous(name = "Length (s)") + 
+    scale_size_continuous(range=c(1,4),name = "Video Length (mins)") + 
     ylab("Proportion Face Detections") + 
     xlab("Age (months)") + 
-  theme_bw() 
+    theme_bw() +
+    theme(legend.position="bottom", legend.box = "horizontal")
+
 ```
 
-```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
+```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=6, fig.height=2, fig.cap = "Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)"}
 ggplot(facebyPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
@@
   xlab("Age (months)") + 
+  facet_grid(~detectorType) +
   theme_bw()
+
 ```
 
-```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
+```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=6, fig.height=2, fig.cap = "Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)" }
 ggplot(faceByOrientation, aes(x = factor(age.grp), y = face, col = orientation)) + 
@@
   xlab("Age (months)") + 
+  facet_grid(~detectorType) +
   theme_bw() 
-```
-
-## Model
-```{r}
-
-fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), 
-      data = d,  
-      family = "binomial")
-
-kable(summary(mod_covariates_age_v2)$coef, digits = 3)
+  # theme(legend.position="bottom", legend.box = "horizontal")
 
@@
 
+<!-- ## Model -->
+
+<!-- # ```{r} -->
+<!-- # d$faceNum=as.numeric(d$faceMT) -->
+<!-- # -->
+<!-- # # Fails to converge -->
+<!-- # # fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), -->
+<!-- # #       data = d, -->
+<!-- # #       family = "binomial") -->
+<!-- # -->
+<!-- # # Fails to converge -->
+<!-- # # agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid), -->
+<!-- # #       data = d, -->
+<!-- # #       family = "binomial") -->
+<!-- # -->
+<!-- # fullModelminsInt <- glmer(faceNum ~ scale(age.at.test) + posture + orientation + (1|subid), -->
+<!-- #        data = d, -->
+<!-- #        family = "binomial") -->
+<!-- # -->
+<!-- # ageOnly <- glmer(faceNum ~ scale(age.at.test) + (1|subid), -->
+<!-- #       data = d, -->
+<!-- #       family = "binomial") -->
+<!-- # -->
+<!-- # ``` -->
+
+To formalize these observations, we fit a generalized logistic mixed-effect model to all with the presence/absence of a face on every frame (as detected by the MTCNN3 model) as the dependent variable, and participant's age, orientation, and posture as predictors.
+Interactions between predictors were not included as this maximal model failed to converge.
+A summary of the coefficients of this model can be found in Table {ref}.
+While age remained a significant predictor even when accounting for the effects of infants' posture and caregiver's oirentation, it held relativley less predictive power.
+Overall, these results confirm that infant's access to faces is heavily infleunced by their own posture and their caregivers orientation towards them.
+
+<!-- ```{r xtable, results="asis"} -->
+<!-- tab2 <- xtable::xtable(summary(fullModelminsInt)$coef, digits=3,  -->
+<!--                       caption = "Results from GLMM model prediting the presence/absence of a face across                       the entire dataset.") -->
+
+<!-- print(tab2, type="latex", comment = F, table.placement = "H") -->
+<!-- ``` -->
+
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
+  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
+  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
+  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
+  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
+  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
+  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{Results from GLMM model prediting the presence/absence of a face across                       the entire dataset.} 
+\end{table}
+
 # Changes in Access to Hands
+
+Next, we examined whether we would see similar trends in infant and children's access to hands.
+
+# General Discussion
+We use a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information relevant for word-learning, here operationalized as the presence of the faces and hands of their caregiver.
+We found that children's posture and orientation towards their caregiver changed systematically across age, and that all three of these factors dramatically impacted the proportion of faces and hands that were avaliable in the child's visual field.
+Thus, infants postural development is mediating factor that explains age-related changes in the proportion of faces and hands avaliable in infants' visual field.
+
+This work also deploys novel advancements in computer vision to the study of developmental psychology.
+The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms (Krihevsky et al., 2012), creating a new generation of algorithmic tools that both are substantially better equipped to deal with noisier, more complicated datasets and that can  extract richer and more detailed information.
+Indeed, these videos from the infant perspective provided sustantial challenges (e.g., partially occluded faces) for the classc models of face detedction (e.g., viola jones).
+Further, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to new infant learners.
+
+Thus, we suggest that the combined use of these new tools can be leveraged to understand the changing infant perspecive on the visual world and the implications of these changes for both linguistic, cognitive, and social development.
 

### 2018-01-13 writeup-cogsci-copy/cogsci2018.Rmd
integrate hand detectors etc.

@@
 
-abstract: 
-The aiblity to rpocess social cues (e.g., eye gaze) is a critical component in children/s early language development.
-However, As children reach their first birthday, they begin to locomote themselves, exploring their visual environment in an entirely new way.
-How do these postural changes affect children's access to the social information relevant for word-learning?
+# abstract: 
+# The ability to process social cues (e.g., eye gaze) is a critical component of children's early language and cognitive development.
+However, as children reach their first birthday, they begin to locomote themselves, exploring their visual environment in an entirely new way.
+How do these postural changes affect children\'s access to the social information relevant for word-learning?
+Here, explore this question by using head-mounted cameras to record infants\' (8-16 months of age) egocentric visual perspective and by using computer-vision algorithms to detect the proportion of faces and hands in infants' environments.
+We find that infant\'s postural changes largely mediate infants\' accesses to both faces and hands, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
 
@@
 # Changes in Access to Faces
-
 ```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }

### 2018-01-13 analysis-main/merge_all_data.R

@@ -21,8 +21,19 @@ detsOpenPose <- detsOpenPose %>%
   distinct(video,frame,.keep_all=TRUE)  %>%
   mutate(frame = as.numeric(frame)) %>%
   mutate(faceOP = Nose_conf!=0 & REye_conf!=0 | LEye_conf!=0 )  %>%
-  mutate(handOP = LWrist_conf!=0 | RWrist_conf!=0 )   
-  
+  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
+
+# open pose detectors
+detsOpenPose_Hands <- read_csv("../data/final_output/openpose_results_with_detail_hands.csv") 
+detsOpenPose_HandConf <- detsOpenPose_Hands %>%
+  mutate(video = name) %>%
+  distinct(video,frame,.keep_all=TRUE)  %>%
+  mutate(frame = as.numeric(frame)) %>%
+  select(video,frame,ends_with("conf")) %>%
+  mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
+  mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
+  mutate(handOP = confRight>.5 | confLeft>.5 )
+
 # viola jones detectors
 detsViola <- read_csv("../data/final_output/viola.csv") 
 detsViola <- detsViola %>%
@@ -32,7 +43,8 @@ detsViola <- detsViola %>%
 
 # merge all three detectors
 alldets=left_join(dets,detsViola[,c("video","frame","faceVJ")]) 
-alldets=left_join(alldets,detsOpenPose[,c("video","frame","faceOP","handOP")]) 
+alldets=left_join(alldets,detsOpenPose[,c("video","frame","faceOP","wristOP")]) 
+alldets=left_join(alldets,detsOpenPose_HandConf[,c("video","frame","handOP")]) 
 
 # load demographics and pose / orientation / timing
 demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
@@ -51,6 +63,7 @@ missingOP<-alldets %>%
 # None of the missing frames are faces. Replace OP "NAN" detections with "FALSE"
 alldets$faceOP[missingInd]=FALSE
 alldets$handOP[missingInd]=FALSE
+alldets$wristOP[missingInd]=FALSE
 
 # calls helper functions in helper.r to get times and posture coding integrated
 d <- alldets %>%

### 2018-01-13 writeup-cogsci-copy/cogsci2018.Rmd
latest cogsci writeup files

@@
 library(lme4)
+library(gridExtra)
 ```
@@
 ## Head-mounted camera
-We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband.Videos captured by the headcam were 720x480 pixels with 25 frames per second.Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
+We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband.
+Videos captured by the headcam were 720x480 pixels with 25 frames per second.Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
 A fisheye lens was attached to the camera to cincrease the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left).
@@
 
-First, all parents signed consent documentsin a waiting room  where children were fitted with the headcam.
+First, all parents signed consent documents in a waiting room  where children were fitted with the headcam.
 After the child was comfortable in the waiting room and with the experimenter, the experimenter placed the headcam on the child's head.
@@
 ### Detector evaluation
-
 \begin{table}[t]
@@
 # Results
-We report results from three different sets of analyses.
-First, we explore how these changes affect access to faces and to hands, as measured using computater vision algorithms.
-Second, we explore how these changes impact the accessibility of faces and hands during labeling events.
+First, we report developmental shifts in infants' posture and their orientation to their caregiver.
+Then, we explore how these changes influence children's visual access to faces and to hands.
+Finally, we explore how these changes impact the accessibility of faces and hands during labeling events.
 
-```{r}
+```{r echo=FALSE, include=FALSE}
 d <- read_csv("../data/consolidateddata/consolidated_data_3dets.csv") %>%
@@
 
-## make long form version for plotting
+## make long form version for plotting both face detections on same graph
 dlong <- d %>%
   mutate(id = paste(subid,frame)) %>%
-  gather(key = detectorType, value = face, faceMT, faceOP)
+  gather(key = detectorType, value = face, faceMT, faceOP) %>%
+  mutate(detectorType=as.factor(detectorType))
 
+# for better plotting names
+levels(detectorType) <- c("MTCNN", "OpenPose") 
+```
+
+# Changes in Posture and Orientation
+```{r echo=FALSE, include=FALSE}
+orientation <- d %>%
+  filter(!is.na(orientation)) %>%
+  group_by(age.grp, subid, orientation) %>%
+  summarise(time = sum(dt, na.rm=TRUE)) %>%
+  mutate(prop.time = time/sum(time)) %>%
+  group_by(subid, age.grp, orientation)
+
+posture <- d %>%
+  filter(!is.na(posture)) %>%
+  group_by(age.grp, subid, posture) %>%
+  summarise(time = sum(dt, na.rm=TRUE)) %>%
+  mutate(prop.time = time/sum(time)) %>%
+  group_by(subid, age.grp, posture) 
+```
+
+```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=6, fig.height=3, fig.cap = "R plot" }
+
+p1=ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
+  geom_boxplot() + 
+  ylab("Proportion Time") + 
+  xlab("Age (months)") + 
+  theme_bw()
+
+p2=ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
+  geom_boxplot() + 
+  ylab("Proportion Time") + 
+  xlab("Age (months)") + 
+  theme_bw()
+
+grid.arrange(p1,p2,nrow = 1)
+
+```
+
+# Changes in Access to Faces
+```{r fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=6, fig.height=2, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age.
+Note that the axes are different; hands were detected overall much less frequently in this dataset." }
 facebyAge <- dlong %>%
@@
   left_join(inc_subs) %>%
-  mutate(len = as.numeric(len_seconds/60))
+  mutate(len = as.numeric(len.min))
 
-faceByOrientation <- dlong %>%
-  filter(!is.na(orientation)) %>%
-  group_by(age.grp, orientation, subid, detectorType) %>%
-  summarise(face = mean(face)) %>%
-  group_by(age.grp)
+handsbyAge <- d %>%
+  filter(!is.na(age.at.test)) %>%
+  group_by(subid, age.at.test) %>%
+  summarise(hands = mean(handOP)) %>%
+  left_join(inc_subs) %>%
+  mutate(len = as.numeric(len.min))
 
-facebyPosture <- dlong %>%
-  filter(!is.na(posture)) %>%
-  group_by(age.grp, posture, subid,detectorType) %>%
-  summarise(face = mean(face)) %>%
-  group_by(age.grp)
-```
-
-# Changes in Access to Faces
-```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=4, fig.height=3, fig.cap = "R plot" }
-
-  ggplot(facebyAge, aes(x = age.at.test, y = face, color = detectorType, size = len)) +
+p1<-ggplot(facebyAge, aes(x = age.at.test, y = face, color = detectorType, size = len)) +
     geom_jitter(width = .25) + 
@@
 
+p2<-ggplot(handsbyAge, aes(x = age.at.test, y = hands, size = len)) +
+    geom_jitter(width = .25) + 
+    geom_smooth() +
+    scale_x_continuous(breaks=c(8,12,16)) +
+    scale_size_continuous(range=c(1,4),name = "Video Length (mins)") + 
+    ylab("Proportion Hand Detections") + 
+    xlab("Age (months)") + 
+    theme_bw() +
+    theme(legend.position="bottom", legend.box = "horizontal")
+
+grid.arrange(p1,p2,nrow = 1)
+
 ```
@@
 ```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=6, fig.height=2, fig.cap = "Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)"}
-ggplot(facebyPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
+
+facebyPosture <- dlong %>%
+  filter(!is.na(posture)) %>%
+  group_by(age.grp, posture, subid,detectorType) %>%
+  summarise(face = mean(face)) %>%
+  group_by(age.grp)
+
+handbyPosture <- d %>%
+  filter(!is.na(posture)) %>%
+  group_by(age.grp, posture, subid) %>%
+  summarise(hand = mean(handOP)) %>%
+  group_by(age.grp)
+
+p2=ggplot(handbyPosture, aes(x = factor(age.grp), y = hand, col = posture)) + 
+  geom_boxplot() + 
+  ylab("Proportion Hand Detections") + 
+  xlab("Age (months)") + 
+  ggtitle("Hand detections") +
+  theme_bw()
+
+p1=ggplot(facebyPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
   geom_boxplot() + 
@@
   facet_grid(~detectorType) +
-  theme_bw()
+  theme_bw() +
+  theme(strip.background = element_blank(), strip.text.y = element_blank())
+
+grid.arrange(p1,p2,nrow = 1, widths=2:1)
 
@@
 ```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=6, fig.height=2, fig.cap = "Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)" }
+faceByOrientation <- dlong %>%
+  filter(!is.na(orientation)) %>%
+  group_by(age.grp, orientation, subid, detectorType) %>%
+  summarise(face = mean(face)) %>%
+  group_by(age.grp)
+
 ggplot(faceByOrientation, aes(x = factor(age.grp), y = face, col = orientation)) + 

### 2018-01-13 writeup-cogsci-copy/cogsci2018.tex
latest cogsci writeup files

--- 2018-01-11 writeup-cogsci/cogsci2018.tex
@@
 
-\title{Postural changes mediate children's visual access to faces and hands}
+\title{Postural changes mediate children's visual access to social information}
 
@@
 \end{abstract}
+
+\section{Introduction}\label{introduction}
+
+Children are remarkably skilled language learners, connecting arbitrary
+labels (``cup'') with specific visual concepts at a rapid pace through
+the first two years of life. However, children do not learn words in a
+vacuum but rather in a rich social environment, where social cues
+provided by speakers (e.g., eye-gaze) provide strong scaffolding for
+this learning process. Further, children's ability to effectively
+process these social cues may be a key factor in their early language
+development. In a longitudinal study, children's level of joint
+engagement with their mother was found to predict both their receptive
+and productive vocabularies (Carpenter, Nagell, \& Tomasello (1998)).
+More recently, 10 month-olds who follow an adult's gaze in an
+experimental context have larger vocabularies at 18 months (Brooks \&
+Meltzoff (2005)) and throughout the second year of life
+(({\textbf{???}})).
+
+However, as children are learning their first words, their view of the
+world is also changing radically (Adolph \& Berger, 2007). Infant's
+motor abilities improve dramatically near the end of the first year of
+life, allowing them to locomote independently and to determine what they
+see. During spontaneous play in a laboratory playroom, toddlers are more
+likely to look at the floor while crawling than while walking (Franchak
+et al., 2011), when they have full visual access to their environment
+and the people in it (Kretch, Franchak, and Adolph, 2014).
+
+One possibility is that these motor improvements have strong
+developmental cascades, impacting children's emerging social, cognitive,
+and linguistic abilities (Iverson, 2010). Indeed, these postural changes
+also impact how children interact with their mothers; walking infants
+make different kinds of object-related bids for attention from their
+mothers than crawling infants, and tend to hear more action directed
+statements (e.g., ``open it''). (Karasik et al., 2014). More directly,
+in an observational study, Walle \& Campos (2014) found that children
+who were able to walk had both higher receptive and productive
+vocabularies. On their account, children's ability to stand and
+independently locomote may fundamentally change their ability to access
+the social information (e.g., faces, gaze) relative to children who are
+still crawling and sitting. In other words, the ability to access more
+detailed social information may allow infants to learn words quicker and
+more efficiently, facilitating language growth.
+
+Recent work has begun to use egocentric, head-mounted cameras to
+document the visual experiences of infants and children --- which even
+for walking children are radically different than the adult perspective
+(and not easily predicted by our own adult intuitions; Franchak, Kretch,
+Soska, \& Adolph (2011), Yoshida \& Smith (2008); ({\textbf{???}})).
+Children's views tend to be restricted and to be dominated by objects
+and hands much more than that of adults (Yoshida \& Smith, 2008), and
+both computational and empirical work suggest that this restricted
+viewpoint may be more effective for learning what objects and their
+labels than the comparable adult perspective (D. Yurovsky, Smith, \& Yu,
+in press, ({\textbf{???}})). Further, recent work also suggests dramatic
+changes in the child's perspective over the first two years of life, as
+views transition from primarily containing close up view of faces to
+capturing views of hands paired with the objects they are acting on
+({\textbf{???}}).
+
+Here, we directly ask whether the postural changes that infants'
+experience as they reach their first birthday change the availability of
+the social information relevant for word learning. To do so, we recorded
+the visual experience of a group of infants and children using
+head-mounted cameras during a brief laboratory free-play session;
+children's posture and their orientation to their caregiver were also
+recorded from a third-person perspective and hand-annotated. We
+capitalize on recent improvements in computer vision algorithms
+(({\textbf{???}}), ({\textbf{???}})) that allow the automated detection
+of both faces and hands, and we analyze the frequency of both faces and
+hands in the child's visual environment.
 
@@
 We used a small, head-mounted camera (``headcam'') that was constructed
-from a MD80 model camera attached to a soft elastic headband.Videos
+from a MD80 model camera attached to a soft elastic headband. Videos
 captured by the headcam were 720x480 pixels with 25 frames per
@@
 
-First, all parents signed consent documentsin a waiting room where
+First, all parents signed consent documents in a waiting room where
 children were fitted with the headcam. After the child was comfortable
@@
 
-An additional goal of the study was to measure the presence of
-caregivers' faces in the child's field of view (as approximated by the
-headcam). To avoid hand-annotating the size and position of faces in
-every frame of video, we tested two face detection systems. Sample
-frames from the video with successful detections are given in Figure
-\ref{fig:frames}.
+Measuring the availability of caregiver's faces across development was
+an important goal of the study. We used a total of three face detection
+systems to explore this hypothesis and to avoid the cost of
+hand-annotating every frame. Results for the evaluation of these three
+systems are shown in Table X.
 
-\subsubsection{Face detection
-algorithms}\label{face-detection-algorithms}
+\subsubsection{Face Detection
+Algorithms}\label{face-detection-algorithms}
+
+The first face detection system made use of a series of Haar
+feature-based cascade classifiers (Viola-Jones, CITEXX) applied to each
+individual frame. This detector provided information about the presence
+of a face as well as its size and position.
+
+The second algorithm was based on the work by (CITEXX) using multi-task
+cascaded convolutional neural networks (MTCNNs). The system was built
+using a novel cascaded CNN-based framework for joint detection and
+alignment, built to perform well in real-world environments where
+varying illuminations and occlusions are present. We used a Tensorflow
+implementation of this algorithm provided by
+(\url{https://github.com/davidsandberg/facenet})
+
+The third algorithm was a CNN-based pose detector that provided the
+locations of 18 body parts (ears, nose, wrists, etc) (CITEXX). The
+system uses a CNN for initial anatomical detection and subsequently
+applies part affinity fields (PAFs) for part association, producing a
+series of body part candidates. The candidates are then matched to a
+single individual and finally assembled into a pose. For the purposes of
+our investigation we only made use of the body parts relevant to the
+face (ears, eyes, nose).
 
@@
 
-We report results from three different sets of analyses. First, we
-explore how these changes affect access to faces and to hands, as
-measured using computater vision algorithms. Second, we explore how
-these changes impact the accessibility of faces and hands during
-labeling events.
+First, we report developmental shifts in infants' posture and their
+orientation to their caregiver. Then, we explore how these changes
+influence children's visual access to faces and to hands. Finally, we
+explore how these changes impact the accessibility of faces and hands
+during labeling events.
 
-\section{Changes in Access to Faces}\label{changes-in-access-to-faces}
-
-\begin{CodeChunk}
-\begin{figure}[H]
-
-{\centering \includegraphics{figs/unnamed-chunk-3-1}
-
-}
-
-\caption[R plot]{R plot}\label{fig:unnamed-chunk-3}
-\end{figure}
-\end{CodeChunk}
+\section{Changes in Posture and
+Orientation}\label{changes-in-posture-and-orientation}
 
@@
 
+\section{Changes in Access to Faces}\label{changes-in-access-to-faces}
+
 \begin{CodeChunk}
-\begin{figure}[H]
+\begin{figure*}[h]
 
@@
 
-\caption[R plot]{R plot}\label{fig:unnamed-chunk-5}
-\end{figure}
+\caption[Proportion face (left) and hand (right) detections as a function of participant's age]{Proportion face (left) and hand (right) detections as a function of participant's age.
+Note that the axes are different; hands were detected overall much less frequently in this dataset.}\label{fig:unnamed-chunk-5}
+\end{figure*}
 \end{CodeChunk}
 
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/faceByPostureFig-1}
+
+}
+
+\caption[Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)]{Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)}\label{fig:faceByPostureFig}
+\end{figure*}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/faceByOrientFig-1}
+
+}
+
+\caption[Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)]{Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)}\label{fig:faceByOrientFig}
+\end{figure*}
+\end{CodeChunk}
+
+To formalize these observations, we fit a generalized logistic
+mixed-effect model to all with the presence/absence of a face on every
+frame (as detected by the MTCNN3 model) as the dependent variable, and
+participant's age, orientation, and posture as predictors. Interactions
+between predictors were not included as this maximal model failed to
+converge. A summary of the coefficients of this model can be found in
+Table \{ref\}. While age remained a significant predictor even when
+accounting for the effects of infants' posture and caregiver's
+oirentation, it held relativley less predictive power. Overall, these
+results confirm that infant's access to faces is heavily infleunced by
+their own posture and their caregivers orientation towards them.
+
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
+  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
+  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
+  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
+  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
+  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
+  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{Results from GLMM model prediting the presence/absence of a face across                       the entire dataset.} 
+\end{table}
+
 \section{Changes in Access to Hands}\label{changes-in-access-to-hands}
+
+Next, we examined whether we would see similar trends in infant and
+children's access to hands.
+
+\section{General Discussion}\label{general-discussion}
+
+We use a head-mounted camera to explore how children's postural and
+locomotive development directly impacts their access to social
+information relevant for word-learning, here operationalized as the
+presence of the faces and hands of their caregiver. We found that
+children's posture and orientation towards their caregiver changed
+systematically across age, and that all three of these factors
+dramatically impacted the proportion of faces and hands that were
+avaliable in the child's visual field. Thus, infants postural
+development is mediating factor that explains age-related changes in the
+proportion of faces and hands avaliable in infants' visual field.
+
+This work also deploys novel advancements in computer vision to the
+study of developmental psychology. The field of object detection and
+recognition has advanced dramatically in the past five years since the
+re-birth of deep learning algorithms (Krihevsky et al., 2012), creating
+a new generation of algorithmic tools that both are substantially better
+equipped to deal with noisier, more complicated datasets and that can
+extract richer and more detailed information. Indeed, these videos from
+the infant perspective provided sustantial challenges (e.g., partially
+occluded faces) for the classc models of face detedction (e.g., viola
+jones). Further, as the headcam technologies employed here were
+inexpensive (\textasciitilde{}\$60 a camera) and the computer vision
+algorithms freely avaliable, this method is a promising avenue for
+quanitfying the visual and social information avaliable to new infant
+learners.
+
+Thus, we suggest that the combined use of these new tools can be
+leveraged to understand the changing infant perspecive on the visual
+world and the implications of these changes for both linguistic,
+cognitive, and social development.
 
@@
 \hypertarget{refs}{}
+\hypertarget{ref-adolph2007}{}
+Adolph, K., \& Berger, S. (2007). Motor development. In \emph{Handbook
+of child psychology}. Wiley Online Library.
+
 \hypertarget{ref-adolph2012}{}
@@
 
+\hypertarget{ref-brooks2005}{}
+Brooks, R., \& Meltzoff, A. (2005). The development of gaze following
+and its relation to language. \emph{Developmental Science}, \emph{8}(6),
+535--543.
+
+\hypertarget{ref-carpenter1998}{}
+Carpenter, M., Nagell, K., \& Tomasello, M. (1998). Social cognition,
+joint attention, and communicative competence from 9 to 15 months of
+age. \emph{Monographs of the Society for Research in Child Development},
+\emph{63}(4).
+
 \hypertarget{ref-cummings1988}{}
@@
 Research}, \emph{29}(1), 7--16.
+
+\hypertarget{ref-franchak2011}{}
+Franchak, J., Kretch, K., Soska, K., \& Adolph, K. (2011). Head-mounted
+eye tracking: A new method to describe infant looking. \emph{Child
+Development}.
 
@@
 
+\hypertarget{ref-walle2014}{}
+Walle, E. A., \& Campos, J. J. (2014). Infant language development is
+related to the acquisition of walking. \emph{Developmental Psychology},
+\emph{50}(2), 336.
+
 \hypertarget{ref-yoshida2008}{}
@@
 
+\hypertarget{ref-yurovsky2012}{}
+Yurovsky, D., Smith, L., \& Yu, C. (in press). Statistical word learning
+at scale: The baby's view is better. \emph{Developmental Science}.
+
 \end{document}

### 2018-01-13 writeup-cogsci-copy/extra.Rmd
latest cogsci writeup files

---
title: "Untitled"
author: "Bria Long"
date: "1/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}

Children (N = `r length(unique(joint$session_id))`, *M* = `r round(mean(d$age[d$trial_num==2]),2)` years, range 4-10 years) participated in an iPad-based drawing game where they were prompted with a verbal cue to draw one of sixteen familiar objects (e.g., “Can you draw a cup?”). Children drew each object category for 30 seconds, after which they were prompted to either make another drawing or to stop drawing altogether. Afterwards, a group of naive adults (N = `r length(unique(r$workerid))`) guessed the identity of each drawn object (`r length(joint$imNameShort)` drawings).

Number of drawings: `r length(joint$imNameShort)`

Number of drawers: `r length(unique(joint$session_id))`

Average age of drawers: `r round(mean(d$age[d$trial_num==2]),2)`

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2018-01-13 analysis-main/merge_all_data.R
integrate wrist detections so no na.s, change filename output

@@ -83,13 +83,12 @@ ages <- d %>%
   group_by(subid) %>%
   summarise(age.grp = mean(age.grp)) 
 
-complete_combos <- expand(d, nesting(posture, orientation), 
-                          subid) %>% 
-  mutate(dt = 0, faceMT = FALSE, faceVJ = FALSE, faceOP = FALSE, handOP = FALSE) %>%
+complete_combos <- expand(d, nesting(posture, orientation), subid) %>% 
+  mutate(dt = 0, faceMT = FALSE, faceVJ = FALSE, faceOP = FALSE, handOP = FALSE, wristOP = FALSE) %>%
   left_join(ages)
 
 d <- bind_rows(d, complete_combos)
 
 ## save it out
-write_csv(d, "../data/consolidateddata/consolidated_data_3dets.csv")
+write_csv(d, "../data/consolidateddata/consolidated_data_4dets.csv")
 

### 2018-01-13 writeup-cogsci-copy/cogsci2018.Rmd
All detectors in ggplot2 facets

@@
 ```{r echo=FALSE, include=FALSE}
-d <- read_csv("../data/consolidateddata/consolidated_data_3dets.csv") %>%
+d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
   mutate(posture = factor(posture), orientation = factor(orientation))
@@
   mutate(id = paste(subid,frame)) %>%
-  gather(key = detectorType, value = face, faceMT, faceOP) %>%
+  gather(key = detectorType, value = detection, faceMT, faceOP, handOP, wristOP) %>%
   mutate(detectorType=as.factor(detectorType))
@@
 # for better plotting names
-levels(detectorType) <- c("MTCNN", "OpenPose") 
+#levels(dlong$detectorType) 
+#[1] "faceMT"  "faceOP"  "handOP"  "wristOP"
+levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","OpenPose-Hands", "OpenPose-Wrists") 
 ```
@@
 
-```{r fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=6, fig.height=3, fig.cap = "R plot" }
+```{r fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
 
@@
 # Changes in Access to Faces
-```{r fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=6, fig.height=2, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age.
+```{r fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age.
 Note that the axes are different; hands were detected overall much less frequently in this dataset." }
-facebyAge <- dlong %>%
+detectionsByAge <- dlong %>%
   filter(!is.na(age.at.test)) %>%
   group_by(subid, age.at.test, detectorType) %>%
-  summarise(face = mean(face)) %>%
+  summarise(detections = mean(detection)) %>%
   left_join(inc_subs) %>%
@@
 
-handsbyAge <- d %>%
-  filter(!is.na(age.at.test)) %>%
-  group_by(subid, age.at.test) %>%
-  summarise(hands = mean(handOP)) %>%
-  left_join(inc_subs) %>%
-  mutate(len = as.numeric(len.min))
-
-p1<-ggplot(facebyAge, aes(x = age.at.test, y = face, color = detectorType, size = len)) +
+ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
     geom_jitter(width = .25) +
@@
     scale_x_continuous(breaks=c(8,12,16)) +
-    scale_size_continuous(range=c(1,4),name = "Video Length (mins)") + 
-    ylab("Proportion Face Detections") + 
+    scale_size_continuous(range=c(.2,2),name = "Video Length (mins)") +
+    ylab("Proportion Detections") +
     xlab("Age (months)") +
-    theme_bw() +
-    theme(legend.position="bottom", legend.box = "horizontal")
+    scale_color_manual(values=c("#1F618D", "#7FB3D5", "#C70039", "#FF5733")) +
+    facet_grid(~detectorType, scales = "free") +
+    theme_bw()
+```
 
-p2<-ggplot(handsbyAge, aes(x = age.at.test, y = hands, size = len)) +
-    geom_jitter(width = .25) + 
-    geom_smooth() +
-    scale_x_continuous(breaks=c(8,12,16)) +
-    scale_size_continuous(range=c(1,4),name = "Video Length (mins)") + 
-    ylab("Proportion Hand Detections") + 
+```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=7, fig.height=3, fig.cap = "Proportion of face detections as a function of infant's posture, using the MTCNN3 face detections and OpenPose face detections (left two panels) and the proportion of hand detections as a function of infant's posture, using OpenPose hand detections and the OpenPose wrist detections.
+"}
+
+detectionsByPosture <- dlong %>%
+  filter(!is.na(posture)) %>%
+  group_by(age.grp, posture, subid,detectorType) %>%
+  summarise(face = mean(detection)) %>%
+  group_by(age.grp)
+
+ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
+  geom_boxplot() + 
+  ylab("Proportion  Detections") + 
   xlab("Age (months)") + 
-    theme_bw() +
-    theme(legend.position="bottom", legend.box = "horizontal")
-
-grid.arrange(p1,p2,nrow = 1)
+  facet_grid(~detectorType) +
+  theme_bw()
 
@@
 
-```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=6, fig.height=2, fig.cap = "Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)"}
-
-facebyPosture <- dlong %>%
-  filter(!is.na(posture)) %>%
-  group_by(age.grp, posture, subid,detectorType) %>%
-  summarise(face = mean(face)) %>%
+```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, num.cols.cap=2, fig.cap = "Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)" }
+detectionsByPosture <- dlong %>%
+  filter(!is.na(orientation)) %>%
+  group_by(age.grp, orientation, subid,detectorType) %>%
+  summarise(detections = mean(detection)) %>%
   group_by(age.grp)
 
-handbyPosture <- d %>%
-  filter(!is.na(posture)) %>%
-  group_by(age.grp, posture, subid) %>%
-  summarise(hand = mean(handOP)) %>%
-  group_by(age.grp)
-
-p2=ggplot(handbyPosture, aes(x = factor(age.grp), y = hand, col = posture)) + 
+ggplot(detectionsByPosture, aes(x = factor(age.grp), y = detections, col = orientation)) + 
   geom_boxplot() + 
-  ylab("Proportion Hand Detections") + 
-  xlab("Age (months)") + 
-  ggtitle("Hand detections") +
-  theme_bw()
-
-p1=ggplot(facebyPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
-  geom_boxplot() + 
-  ylab("Proportion Face Detections") + 
-  xlab("Age (months)") + 
-  facet_grid(~detectorType) +
-  theme_bw() +
-  theme(strip.background = element_blank(), strip.text.y = element_blank())
-
-grid.arrange(p1,p2,nrow = 1, widths=2:1)
-
-```
-
-```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=6, fig.height=2, fig.cap = "Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)" }
-faceByOrientation <- dlong %>%
-  filter(!is.na(orientation)) %>%
-  group_by(age.grp, orientation, subid, detectorType) %>%
-  summarise(face = mean(face)) %>%
-  group_by(age.grp)
-
-ggplot(faceByOrientation, aes(x = factor(age.grp), y = face, col = orientation)) + 
-  geom_boxplot() + 
-  ylab("Proportion Face Detections") + 
+  ylab("Proportion  Detections") + 
   xlab("Age (months)") + 
@@
   theme_bw() 
-  # theme(legend.position="bottom", legend.box = "horizontal")
-
 ```
@@
 
-# Changes in Access to Hands
-
-Next, we examined whether we would see similar trends in infant and children's access to hands.
-
 # General Discussion

### 2018-01-13 writeup-cogsci-copy/cogsci2018.Rmd
motivations for wrist detectors, placeholders in detection stuff

@@
 Measuring the availability of caregiver’s faces across development was an important goal of the study.
-We used a total of three face detection systems to explore this hypothesis and to avoid the cost of hand-annotating every frame.
-Results for the evaluation of these three systems are shown in Table X.
+We thus used a total of three face detection systems to explore this hypothesis and to avoid the cost of hand-annotating every frame.
+We first measure the performance of the most commonly-used and widely avlaialbe face detection algorithmsm (Viola-Jones).
+We use this as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is is notoriously bad at dealing with occluded faces (@citeAnthony2016).
+We next capitalized on recent improvements in computer vision, testing the performance of two other algorithms that extract face information in distinct ways.
+Rsults for the evaluation of these three systems are shown in Table XX.
+
+The second of these algorithms (OpenPose) also captures information about the "skeleton" of people in a given frame, outputting information about the entire set of body parts that were detected.
+This allowed us to extract hand information as well as wrist information.
+We analyze wrist detections as a good proxy for the presence of a hand.
+We do so first because we did not capture the entire visual field experienced by the infants/children.
+Thus evne though the video may only show the presence of a wrist, the caregiver's hand may be within the child's field of view.
+Second, we do so because hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
+For example,if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy).
 
@@
 The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
-We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet)
+We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet).
+Like Viola-Jones, this detector provided ifnormation about the presence of a face as well as it's size and position.
 
@@
 For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose).
+We operationalized face detections as ....., we operationalized hand detedctions as .....
+we operationalized wrist detections as.....
 
 ### Detector evaluation
+
+Evaluating face detectors (text here about how gold standard was selected)
+Evaluating hand/wrist detectors (text here about how gold standard was selected)
+
 \begin{table}[t]
@@
 ```
+We noted characteristic changes in infants posture and orientation across this developmental time range.
+The proportion of time infant's spent sitting decreased with age, and the proportion of time infant's spent standing increased with age.
+Both 8-month-olds and 12-month-olds psent equivalent amounts of time either lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time either sitting or standing (see Figure \ref{fig:posture}).
+We also observed characteristic changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:orientation}).
 
-```{r fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
+```{r posture fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
 
@@
 # Changes in Access to Faces
+
+Next, we examined the impact of these changes
+
 ```{r fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age.
@@
 
-```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=7, fig.height=3, fig.cap = "Proportion of face detections as a function of infant's posture, using the MTCNN3 face detections and OpenPose face detections (left two panels) and the proportion of hand detections as a function of infant's posture, using OpenPose hand detections and the OpenPose wrist detections.
-"}
+```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=7, fig.height=3, fig.cap = "Proportion of face detections as a function of infant's orientation, using the MTCNN3 face detections and OpenPose face detections (left two panels) and the proportion of hand detections as a function of infant's orientation, using OpenPose hand detections and the OpenPose wrist detections."}
 
@@
 
-```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, num.cols.cap=2, fig.cap = "Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)" }
+```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, num.cols.cap=2, fig.cap = "sing the MTCNN3 face detections and OpenPose face detections (left two panels) and the proportion of hand detections as a function of infant's posture, using OpenPose hand detections and the OpenPose wrist detections" }
 detectionsByPosture <- dlong %>%
@@
 
-<!-- ## Model -->
+## Model
 
-<!-- # ```{r} -->
-<!-- # d$faceNum=as.numeric(d$faceMT) -->
-<!-- # -->
-<!-- # # Fails to converge -->
-<!-- # # fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), -->
-<!-- # #       data = d, -->
-<!-- # #       family = "binomial") -->
-<!-- # -->
-<!-- # # Fails to converge -->
-<!-- # # agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid), -->
-<!-- # #       data = d, -->
-<!-- # #       family = "binomial") -->
-<!-- # -->
-<!-- # fullModelminsInt <- glmer(faceNum ~ scale(age.at.test) + posture + orientation + (1|subid), -->
-<!-- #        data = d, -->
-<!-- #        family = "binomial") -->
-<!-- # -->
-<!-- # ageOnly <- glmer(faceNum ~ scale(age.at.test) + (1|subid), -->
-<!-- #       data = d, -->
-<!-- #       family = "binomial") -->
-<!-- # -->
-<!-- # ``` -->
+```{r echo=FALSE, include=FALSE}
+d$faceNum=as.numeric(d$faceMT)
+d$handNum=as.numeric(d$handOP)
+
+# Fails to converge
+# fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid),
+#       data = d,
+#       family = "binomial")
+
+# Fails to converge
+# agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid),
+#       data = d,
+#       family = "binomial")
+
+fullModelminsInt_Faces <- glmer(faceNum ~ scale(age.at.test) + posture + orientation + (1|subid),
+       data = d,
+       family = "binomial")
+
+fullModelminsInt_Handd <- glmer(handNum ~ scale(age.at.test) + posture + orientation + (1|subid),
+       data = d,
+       family = "binomial")
+
+ageOnly <- glmer(faceNum ~ scale(age.at.test) + (1|subid),
+      data = d,
+      family = "binomial")
+
+```
 
@@
 
-<!-- ```{r xtable, results="asis"} -->
-<!-- tab2 <- xtable::xtable(summary(fullModelminsInt)$coef, digits=3,  -->
-<!--                       caption = "Results from GLMM model prediting the presence/absence of a face across                       the entire dataset.") -->
+```{r xtable, results="asis"}
+tab2 <- xtable::xtable(summary(fullModelminsInt_Faces)$coef, digits=3,
+                      caption = "Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across                       the entire dataset.")
 
-<!-- print(tab2, type="latex", comment = F, table.placement = "H") -->
-<!-- ``` -->
+print(tab2, type="latex", comment = F, table.placement = "H")
+```
+
+```{r xtable, results="asis"}
+tab2 <- xtable::xtable(summary(fullModelminsInt_Hands)$coef, digits=3,
+                      caption = "Results from GLMM model prediting the presence/absence of a hands (OpenPose-Hands) across                       the entire dataset.")
+
+print(tab2, type="latex", comment = F, table.placement = "H")
+```
 

### 2018-01-13 writeup-cogsci-copy/cogsci2018.Rmd
merged detection figures, wrote more results text

@@
 
-```{r posture fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
+```{r posture, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
 
@@
 
-# Changes in Access to Faces
+# Changes in Access to Faces & Hands
 
-Next, we examined the impact of these changes
+ We first examined the proportion of face and hand detections across age; a full summary can be seen in Figure \ref{fig:detsbyAge}.
+We observed  a slight u-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds may experience less faces than 8 or 12-month-olds.
+Next, we observed that hand detections increased with the age of the participant, though they were relatively infrequent in this dataset.
+As expected, we found that wrist detections we relatively more frequent and also appeared to increase somewhat with age, suggesting that caregivers hands may be just outside of the field of view of our camera or that they were often occluded by toys.
 
-```{r fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age.
+Next, we turned to our hypothesis of interest, examinig the impact of postural and locomotive developments on children's visual access to faces and hands.
+Overall, we ofund that children's posture was a major factor both in how many faces they saw during the play session.
+Infants who were sitting saw mor faces than infants who were lying down or being carried, while infants who were standing saw the most faces.
+
+```{r detByAge, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age.
 Note that the axes are different; hands were detected overall much less frequently in this dataset." }
@@
 
-```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=7, fig.height=3, fig.cap = "Proportion of face detections as a function of infant's orientation, using the MTCNN3 face detections and OpenPose face detections (left two panels) and the proportion of hand detections as a function of infant's orientation, using OpenPose hand detections and the OpenPose wrist detections."}
+```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=7, fig.height=6, fig.cap = "Proportion of face/hand detections as a function of infant's posture (top panels) and orientation (bottom panels), using the MTCNN3 face detections and OpenPose face detections (left two panels) and using OpenPose hand detections and the OpenPose wrist detections (right two panels)."}
 
@@
 
-ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
+p1=ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
   geom_boxplot() + 
@@
 
-```
-
-```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, num.cols.cap=2, fig.cap = "sing the MTCNN3 face detections and OpenPose face detections (left two panels) and the proportion of hand detections as a function of infant's posture, using OpenPose hand detections and the OpenPose wrist detections" }
-detectionsByPosture <- dlong %>%
+detectionsByOrientation <- dlong %>%
   filter(!is.na(orientation)) %>%
@@
 
-ggplot(detectionsByPosture, aes(x = factor(age.grp), y = detections, col = orientation)) + 
+p2=ggplot(detectionsByOrientation, aes(x = factor(age.grp), y = detections, col = orientation)) + 
   geom_boxplot() + 
@@
   theme_bw()
+
+grid.arrange(p1,p2, nrow=2)
+
+```
+
+```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, num.cols.cap=2, fig.cap = "sing the MTCNN3 face detections and OpenPose face detections (left two panels) and the proportion of hand detections as a function of infant's posture, using OpenPose hand detections and the OpenPose wrist detections" }
+
 ```
@@
 
-```{r echo=FALSE, include=FALSE}
-d$faceNum=as.numeric(d$faceMT)
-d$handNum=as.numeric(d$handOP)
+<!-- ```{r echo=FALSE, include=FALSE} -->
+<!-- d$faceNum=as.numeric(d$faceMT) -->
+<!-- d$handNum=as.numeric(d$handOP) -->
 
-# Fails to converge
-# fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid),
-#       data = d,
-#       family = "binomial")
+<!-- # Fails to converge -->
+<!-- # fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), -->
+<!-- #       data = d, -->
+<!-- #       family = "binomial") -->
 
-# Fails to converge
-# agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid),
-#       data = d,
-#       family = "binomial")
+<!-- # Fails to converge -->
+<!-- # agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid), -->
+<!-- #       data = d, -->
+<!-- #       family = "binomial") -->
 
-fullModelminsInt_Faces <- glmer(faceNum ~ scale(age.at.test) + posture + orientation + (1|subid),
-       data = d,
-       family = "binomial")
+<!-- fullModelminsInt_Faces <- glmer(faceNum ~ scale(age.at.test) + posture + orientation + (1|subid), -->
+<!--        data = d, -->
+<!--        family = "binomial") -->
 
-fullModelminsInt_Handd <- glmer(handNum ~ scale(age.at.test) + posture + orientation + (1|subid),
-       data = d,
-       family = "binomial")
+<!-- fullModelminsInt_Handd <- glmer(handNum ~ scale(age.at.test) + posture + orientation + (1|subid), -->
+<!--        data = d, -->
+<!--        family = "binomial") -->
 
-ageOnly <- glmer(faceNum ~ scale(age.at.test) + (1|subid),
-      data = d,
-      family = "binomial")
+<!-- ageOnly <- glmer(faceNum ~ scale(age.at.test) + (1|subid), -->
+<!--       data = d, -->
+<!--       family = "binomial") -->
 
-```
+<!-- ``` -->
 
@@
 
-```{r xtable, results="asis"}
-tab2 <- xtable::xtable(summary(fullModelminsInt_Faces)$coef, digits=3,
-                      caption = "Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across                       the entire dataset.")
+<!-- ```{r xtable, results="asis"} -->
+<!-- tab2 <- xtable::xtable(summary(fullModelminsInt_Faces)$coef, digits=3, -->
+<!--                       caption = "Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.") -->
 
-print(tab2, type="latex", comment = F, table.placement = "H")
-```
+<!-- print(tab2, type="latex", comment = F, table.placement = "H") -->
+<!-- ``` -->
 
-```{r xtable, results="asis"}
-tab2 <- xtable::xtable(summary(fullModelminsInt_Hands)$coef, digits=3,
-                      caption = "Results from GLMM model prediting the presence/absence of a hands (OpenPose-Hands) across                       the entire dataset.")
+<!-- ```{r xtable, results="asis"} -->
+<!-- tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3, -->
+<!--                       caption = "Results from GLMM model prediting the presence/absence of a hands (OpenPose-Hands) across the entire dataset.") -->
 
-print(tab2, type="latex", comment = F, table.placement = "H")
-```
+<!-- print(tab3, type="latex", comment = F, table.placement = "H") -->
+<!-- ``` -->
 
@@
 \caption{Results from GLMM model prediting the presence/absence of a face across  the entire dataset.} 
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -8.775 & 0.236 & -37.259 & 0.000 \\ 
+  Age & 0.879 & 0.166 & 5.292 & 0.000 \\ 
+  Posture-Prone  & 1.215 & 0.110 & 11.082 & 0.000 \\ 
+  Posture-Sit & 1.943 & 0.108 & 18.000 & 0.000 \\ 
+  Posture-Stand  & 2.175 & 0.107 & 20.260 & 0.000 \\ 
+  Orient.-Close  & 0.608 & 0.059 & 10.371 & 0.000 \\ 
+  Orient.-Far & 2.244 & 0.060 & 37.222 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.} 
 \end{table}

### 2018-01-13 writeup-cogsci-copy/cogsci2018.Rmd
some rearranging of figures / texdt, not ifnished...

@@
 
-# abstract: 
-# The ability to process social cues (e.g., eye gaze) is a critical component of children's early language and cognitive development.
+abstract: 
+  "The ability to process social cues (e.g., eye gaze) is a critical component of children's early language and cognitive development.
 However, as children reach their first birthday, they begin to locomote themselves, exploring their visual environment in an entirely new way.
-How do these postural changes affect children\'s access to the social information relevant for word-learning?
-Here, explore this question by using head-mounted cameras to record infants\' (8-16 months of age) egocentric visual perspective and by using computer-vision algorithms to detect the proportion of faces and hands in infants' environments.
-We find that infant\'s postural changes largely mediate infants\' accesses to both faces and hands, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
+How do these postural changes affect children's access to the social information relevant for word-learning?
+Here, explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and by using computer-vision algorithms to detect the proportion of faces and hands in infants' environments.
+We find that infant's postural changes largely mediate infants' accesses to both faces and hands, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities."
 
@@
 \end{table}
-To obtain this final sample, we tested `r length(all_subs$reason)`, excluding  `r sum(all_subs$reason!="included")` children for the following reasons:
+To obtain this final sample, we tested `r length(all_subs$reason)` children, excluding  `r sum(all_subs$reason!="included")` children for the following reasons:
 `r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
@@
 ## Head-mounted camera
+```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for three different headcam configurations, with the device we used in the middle.
+The lowest camera is pictured for comparison, but was not available until after our study was already in progress."}
+img <- png::readPNG("images/viewangle.png")
+grid::grid.raster(img)
+```
+
 We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband.
@@
 This allowed us to maximize our chances of capturing faces that the child would have seen during the play session.
-
-```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for three different headcam configurations, with the device we used in the middle.
-The lowest camera is pictured for comparison, but was not available until after our study was already in progress."}
-img <- png::readPNG("images/viewangle.png")
-grid::grid.raster(img)
-```
 
@@
 
-\begin{table}[t]
-  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
-  \begin{center}
-    \begin{tabular}{l|ccc|ccc}
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
   \hline
-       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
-      % \null Model & P & R & F & P & R & F  \\
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
   \hline
-      % Viola-Jones & .55  & .38 &  .45 & .89 & .74 & .81   \\
-      % MTCNN & .86 & .78 & .81 & .93 & .76 & .83 \\
-      % OpenPose-Faces & .86 & .78 & .81 & .93 & .76 & .83 \\
-      % OpenPose-Wrists & .86 & .78 & .81 & .93 & .76 & .83 \\
-      % OpenPose-Hands & .86 & .78 & .81 & .93 & .76 & .83 \\
+(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
+  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
+  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
+  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
+  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
+  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
+  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
    \hline
 \end{tabular}
-  \end{center}
+\caption{TABLE PLACEHOLDER} 
 \end{table}
@@
 
-# Changes in Posture and Orientation
+## Changes in Posture and Orientation
+
 ```{r echo=FALSE, include=FALSE}
@@
 ```
+
+```{r orientation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
+
+ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
+  geom_boxplot() + 
+  ylab("Proportion Time") + 
+  xlab("Age (months)") + 
+  theme_bw()
+```
 We noted characteristic changes in infants posture and orientation across this developmental time range.
@@
 
-```{r posture, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
-
-p1=ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
+```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
+ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
   geom_boxplot() + 
@@
   theme_bw()
-
-p2=ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
-  geom_boxplot() + 
-  ylab("Proportion Time") + 
-  xlab("Age (months)") + 
-  theme_bw()
-
-grid.arrange(p1,p2,nrow = 1)
-
 ```
@@
 Overall, we ofund that children's posture was a major factor both in how many faces they saw during the play session.
-Infants who were sitting saw mor faces than infants who were lying down or being carried, while infants who were standing saw the most faces.
+Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces.
+We observed a similar pattern with respect to hands: children who were sitting or standing tended to see more hands/wrists than children who were lying down or held.
 
-```{r detByAge, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age.
+We next examined how the child's orientation relative to their caregiver impacted their visual access to faces and hands.
+Overlal, we saw that children who were far away from their caregiver were more likely to see faces and hands than children who were close to their caregiver; this was true within all age groups and for all detectors.
+
+```{r detByAge, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=2, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age.
 Note that the axes are different; hands were detected overall much less frequently in this dataset." }
@@
 grid.arrange(p1,p2, nrow=2)
-
-```
-
-```{r faceByOrientFig, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, num.cols.cap=2, fig.cap = "sing the MTCNN3 face detections and OpenPose face detections (left two panels) and the proportion of hand detections as a function of infant's posture, using OpenPose hand detections and the OpenPose wrist detections" }
 

### 2018-01-14 analysis-validation/detector_evaluation.Rmd
merge all data changes

@@
 ```{r}
-setwd("..")
-data <- read_csv("data/final_output/openpose_results_truncated.csv") %>%
+setwd("xs-face")
+a <- read_csv("data/final_output/openpose_results_truncated.csv") %>%
   distinct(name, frame, .keep_all = TRUE)
 
-data %<>%
-  mutate(is_face = ifelse(LWrist_conf != 0 | RWrist_conf != 0, "True", "False"))%>%
+data <- a %>%
+  mutate(is_face = ifelse(LWrist_conf != 0 | RWrist_conf != 0, "True", "False")) %>%
   select(group, name, frame, is_face) %>%
@@
   select(-angle) %>%
-  arrange(video, frame)
+  arrange(video, frame) %>%
+  cbind(read_csv("data/ground_truth/gold_set_hand_candidates.csv") %>%
+          arrange(video, frame) %>%
+          select(sample_type))
+```
+
+OPENPOSE HAND DETAIL
+```{r}
+a <- read_csv("/Users/alessandro/workspace/openpose_results_with_detail.csv") %>%
+  distinct(name, frame, .keep_all = TRUE)
+
+data <- a %>%
+ mutate(video = name) %>%
+ distinct(video,frame,.keep_all=TRUE)  %>%
+ #mutate(frame = as.numeric(frame)) %>%
+ select(group, video,frame,ends_with("conf")) %>%
+ mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
+ mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
+ mutate(handOP = confRight>.2 | confLeft>.2 )
+  #mutate(is_face = ifelse(LWrist_conf != 0 | RWrist_conf != 0, "True", "False"))%>%
+  #select(group, name, frame, is_face) %>%
+  # rename(video = name) %>%
+  # select(group, video,frame,ends_with("conf")) %>%
+  # mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
+  # mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
+  # mutate(handOP = confRight>0 | confLeft>0 )
+
+data %<>%
+  mutate(is_face = ifelse(handOP, "True", "False"))
+
+ground_truth_data <- read_csv("data/ground_truth/ground_truth_hands.csv") %>%
+  select(-angle) %>%
+  arrange(video, frame) %>%
+  cbind(read_csv("data/ground_truth/gold_set_hand_candidates.csv") %>%
+          arrange(video, frame) %>%
+          select(sample_type))
 ```
@@
 ```{r}
-score <- results %>%
-  group_by(sample_type, result) %>%
-  summarise(count = n())
+results %>%
+  count(result)
 

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
merge all data changes

@@
 Evaluating face detectors (text here about how gold standard was selected)
+
+\begin{table}[t]
+  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
+  \begin{center}
+    \begin{tabular}{l|ccc|ccc}
+      \hline
+       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
+      % \null Model & P & R & F & P & R & F  \\
+      \hline
+      % MTCNN & .84  & 1 &  .91 & .95 & .75 & .85   \\
+      % OpenPose & .99 & .87 & .93 & .71 & .92 & .8 \\
+    \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
 Evaluating hand/wrist detectors (text here about how gold standard was selected)

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
edits to paper

--- 2018-01-13 writeup-cogsci-copy/cogsci2018.Rmd
@@
 Children are remarkably skilled language learners, connecting arbitrary labels (“cup”) with specific visual concepts at a rapid pace through the first two years of life.
-However, children do not learn words in a vacuum but rather in a rich social environment, where social cues provided by speakers (e.g., eye-gaze) provide strong scaffolding for this learning process.
-Further, children’s ability to effectively process these social cues may be a key factor in their early language development.
-In a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies (@carpenter1998).
+However, children do not learn words in a vacuum but in a rich social environment, where social cues provided by speakers (e.g., eye-gaze) provide strong scaffolding for this learning process.
+Indeed, children’s ability to effectively process these social cues may be a key factor in their early language development.
+For example,in a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies (@carpenter1998).
 More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months (@brooks2005) and throughout the second year of life (@brooks2008).
@@
 Infant’s motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see.
-During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011), when they have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).
+These motor changes have drastic consequences for what children see: crawling and walking infants have radically different views of the world.
+During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011); walking vs.
+crawling infants tend to have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).
 
@@
 ```
+
 Our final sample consisted of 36 infants and children, with 8 participants in three age groups: 8 months (6 females), 12 months (7 females), and 16 months (6 females).
@@
 ### Posture and Orientation Annotation
-
 We created a set of custom annotations that described the child's physical posture (e.g.
@@
 We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet).
-Like Viola-Jones, this detector provided ifnormation about the presence of a face as well as it's size and position.
+Like Viola-Jones, this detector provided information about the presence of a face as well as it's size and position.
 
@@
 For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose).
-We operationalized face detections as ....., we operationalized hand detedctions as .....
-we operationalized wrist detections as.....
+We operationalized face detections as any frames containing a nose and either the left or right eye; we operationalized hand detections as any frames containing any of the 21 points detected on either the left or the right hand, and we operationalized wrist detections any frames containing either the left or the right wrist.
 
 ### Detector evaluation
-
-Evaluating face detectors (text here about how gold standard was selected)
-Evaluating hand/wrist detectors (text here about how gold standard was selected)
+In order to evaluate the performance of these detectors, we constructed a "gold set" of frames
 
@@
     geom_jitter(width = .25) +
-    geom_smooth() +
+    geom_smooth(size=1) +
     scale_x_continuous(breaks=c(8,12,16)) +
-    scale_size_continuous(range=c(.2,2),name = "Video Length (mins)") +
+    scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
     ylab("Proportion Detections") +
@@
     scale_color_manual(values=c("#1F618D", "#7FB3D5", "#C70039", "#FF5733")) +
-    facet_grid(~detectorType, scales = "free") +
-    theme_bw()
+    facet_grid(~detectorType, scales = "free_y") 
+    # theme_bw()
 ```

### 2018-01-14 analysis-main/merge_all_data.R
change confidence for hand detectors to 0

@@ -32,7 +32,7 @@ detsOpenPose_HandConf <- detsOpenPose_Hands %>%
   select(video,frame,ends_with("conf")) %>%
   mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
   mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
-  mutate(handOP = confRight>.5 | confLeft>.5 )
+  mutate(handOP = confRight>0 | confLeft>0 )
 
 # viola jones detectors
 detsViola <- read_csv("../data/final_output/viola.csv") 

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
got rid of merge conflict inserts

@@
 In order to evaluate the performance of these detectors, we constructed a "gold set" of frames
+Evaluating face detectors (text here about how gold standard was selected)
+
+\begin{table}[t]
+  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
+  \begin{center}
+    \begin{tabular}{l|ccc|ccc}
+      \hline
+       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
+      % \null Model & P & R & F & P & R & F  \\
+      \hline
+      % MTCNN & .84  & 1 &  .91 & .95 & .75 & .85   \\
+      % OpenPose & .99 & .87 & .93 & .71 & .92 & .8 \\
+    \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
+Evaluating hand/wrist detectors (text here about how gold standard was selected)
 

### 2018-01-14 writeup-cogsci-copy/cogsci2018.tex
got rid of merge conflict inserts

@@
 \begin{abstract}
+The ability to process social cues (e.g., eye gaze) is a critical
+component of children's early language and cognitive development.
+However, as children reach their first birthday, they begin to locomote
+themselves, exploring their visual environment in an entirely new way.
+How do these postural changes affect children's access to the social
+information relevant for word-learning? Here, explore this question by
+using head-mounted cameras to record infants' (8-16 months of age)
+egocentric visual perspective and by using computer-vision algorithms to
+detect the proportion of faces and hands in infants' environments. We
+find that infant's postural changes largely mediate infants' accesses to
+both faces and hands, suggesting that these postural and locomotive
+developments facilitate infants' emerging linguistic and social
+capacities.
 
@@
 the first two years of life. However, children do not learn words in a
-vacuum but rather in a rich social environment, where social cues
-provided by speakers (e.g., eye-gaze) provide strong scaffolding for
-this learning process. Further, children's ability to effectively
-process these social cues may be a key factor in their early language
-development. In a longitudinal study, children's level of joint
-engagement with their mother was found to predict both their receptive
-and productive vocabularies (Carpenter, Nagell, \& Tomasello (1998)).
-More recently, 10 month-olds who follow an adult's gaze in an
-experimental context have larger vocabularies at 18 months (Brooks \&
-Meltzoff (2005)) and throughout the second year of life
-(({\textbf{???}})).
+vacuum but in a rich social environment, where social cues provided by
+speakers (e.g., eye-gaze) provide strong scaffolding for this learning
+process. Indeed, children's ability to effectively process these social
+cues may be a key factor in their early language development. For
+example,in a longitudinal study, children's level of joint engagement
+with their mother was found to predict both their receptive and
+productive vocabularies (Carpenter, Nagell, \& Tomasello (1998)). More
+recently, 10 month-olds who follow an adult's gaze in an experimental
+context have larger vocabularies at 18 months (Brooks \& Meltzoff
+(2005)) and throughout the second year of life (({\textbf{???}})).
 
@@
 life, allowing them to locomote independently and to determine what they
-see. During spontaneous play in a laboratory playroom, toddlers are more
-likely to look at the floor while crawling than while walking (Franchak
-et al., 2011), when they have full visual access to their environment
-and the people in it (Kretch, Franchak, and Adolph, 2014).
+see. These motor changes have drastic consequences for what children
+see: crawling and walking infants have radically different views of the
+world. During spontaneous play in a laboratory playroom, toddlers are
+more likely to look at the floor while crawling than while walking
+(Franchak et al., 2011); walking vs.~crawling infants tend to have full
+visual access to their environment and the people in it (Kretch,
+Franchak, and Adolph, 2014).
 
@@
 
-To obtain this final sample, we tested 95, excluding 59 children for the
-following reasons: 20 for technical issues related to the headcam, 15
-for failing to wear the headcam, 10 for fewer than 4 minutes of headcam
-footage, 5 for having multiple adults present, 5 for missing CDI data, 2
-for missing scene camera footage, 1 for fussiness, and one excluded for
-sample symmetry. All inclusion decisions were made independent of the
-results of subsequent analyses.
+To obtain this final sample, we tested 95 children, excluding 59
+children for the following reasons: 20 for technical issues related to
+the headcam, 15 for failing to wear the headcam, 10 for fewer than 4
+minutes of headcam footage, 5 for having multiple adults present, 5 for
+missing CDI data, 2 for missing scene camera footage, 1 for fussiness,
+and one excluded for sample symmetry. All inclusion decisions were made
+independent of the results of subsequent analyses.
 
 \subsection{Head-mounted camera}\label{head-mounted-camera}
+
+\begin{CodeChunk}
+\begin{figure}[H]
+
+{\centering \includegraphics{figs/image-1}
+
+}
+
+\caption[Field of view for three different headcam configurations, with the device we used in the middle]{Field of view for three different headcam configurations, with the device we used in the middle.
+The lowest camera is pictured for comparison, but was not available until after our study was already in progress.}\label{fig:image}
+\end{figure}
+\end{CodeChunk}
 
@@
 capturing faces that the child would have seen during the play session.
-
-\begin{CodeChunk}
-\begin{figure}[H]
-
-{\centering \includegraphics{figs/image-1}
-
-}
-
-\caption[Field of view for three different headcam configurations, with the device we used in the middle]{Field of view for three different headcam configurations, with the device we used in the middle.
-The lowest camera is pictured for comparison, but was not available until after our study was already in progress.}\label{fig:image}
-\end{figure}
-\end{CodeChunk}
 
@@
 Measuring the availability of caregiver's faces across development was
-an important goal of the study. We used a total of three face detection
-systems to explore this hypothesis and to avoid the cost of
-hand-annotating every frame. Results for the evaluation of these three
-systems are shown in Table X.
+an important goal of the study. We thus used a total of three face
+detection systems to explore this hypothesis and to avoid the cost of
+hand-annotating every frame. We first measure the performance of the
+most commonly-used and widely avlaialbe face detection algorithmsm
+(Viola-Jones). We use this as a benchmark for performance, as while it
+can achieve impressive accuracy in some situations, it is is notoriously
+bad at dealing with occluded faces (({\textbf{???}})). We next
+capitalized on recent improvements in computer vision, testing the
+performance of two other algorithms that extract face information in
+distinct ways. Rsults for the evaluation of these three systems are
+shown in Table XX.
+
+The second of these algorithms (OpenPose) also captures information
+about the ``skeleton'' of people in a given frame, outputting
+information about the entire set of body parts that were detected. This
+allowed us to extract hand information as well as wrist information. We
+analyze wrist detections as a good proxy for the presence of a hand. We
+do so first because we did not capture the entire visual field
+experienced by the infants/children. Thus evne though the video may only
+show the presence of a wrist, the caregiver's hand may be within the
+child's field of view. Second, we do so because hands are often occluded
+by objects when caregivers are interacting with children, yet still
+visually accessible by the child and part of their joint interaction.
+For example,if a caregiver was holding a toy and presenting it to the
+infant, their wrists may be visible but their hand may not necessarily
+be (as it is occluded by the toy).
 
@@
 implementation of this algorithm provided by
-(\url{https://github.com/davidsandberg/facenet})
+(\url{https://github.com/davidsandberg/facenet}). Like Viola-Jones, this
+detector provided information about the presence of a face as well as
+it's size and position.
 
@@
 our investigation we only made use of the body parts relevant to the
-face (ears, eyes, nose).
+face (ears, eyes, nose). We operationalized face detections as any
+frames containing a nose and either the left or right eye; we
+operationalized hand detections as any frames containing any of the 21
+points detected on either the left or the right hand, and we
+operationalized wrist detections any frames containing either the left
+or the right wrist.
 
 \subsubsection{Detector evaluation}\label{detector-evaluation}
+
+In order to evaluate the performance of these detectors, we constructed
+a ``gold set'' of frames Evaluating face detectors (text here about how
+gold standard was selected)
 
@@
       \hline
-      % Viola-Jones & .55  & .38 &  .45 & .89 & .74 & .81   \\
-      % MTCNN & .86 & .78 & .81 & .93 & .76 & .83 \\
-      % OpenPose-Faces & .86 & .78 & .81 & .93 & .76 & .83 \\
-      % OpenPose-Wrists & .86 & .78 & .81 & .93 & .76 & .83 \\
-      % OpenPose-Hands & .86 & .78 & .81 & .93 & .76 & .83 \\
+      % MTCNN & .84  & 1 &  .91 & .95 & .75 & .85   \\
+      % OpenPose & .99 & .87 & .93 & .71 & .92 & .8 \\
     \hline
@@
   \end{center}
+\end{table}
+
+Evaluating hand/wrist detectors (text here about how gold standard was
+selected)
+
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
+  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
+  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
+  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
+  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
+  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
+  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{TABLE PLACEHOLDER} 
 \end{table}
@@
 
-\section{Changes in Posture and
+\subsection{Changes in Posture and
 Orientation}\label{changes-in-posture-and-orientation}
@@
 \begin{CodeChunk}
-\begin{figure}[H]
+\begin{figure}[h]
 
-{\centering \includegraphics{figs/unnamed-chunk-4-1}
+{\centering \includegraphics{figs/orientation-1}
 
@@
 
-\caption[R plot]{R plot}\label{fig:unnamed-chunk-4}
+\caption[Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)]{Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel).}\label{fig:orientation}
 \end{figure}
@@
 
-\section{Changes in Access to Faces}\label{changes-in-access-to-faces}
+We noted characteristic changes in infants posture and orientation
+across this developmental time range. The proportion of time infant's
+spent sitting decreased with age, and the proportion of time infant's
+spent standing increased with age. Both 8-month-olds and 12-month-olds
+psent equivalent amounts of time either lying/crawling, which was
+markedly decreased in the 16-month-olds, who spent most of their time
+either sitting or standing (see Figure \ref{fig:posture}). We also
+observed characteristic changes in children's orientation relative to
+their caregivers: the 8-month-olds spent more time with their caregiver
+behind them supporting their sitting positions (see Figure
+\ref{fig:orientation}).
+
+\begin{CodeChunk}
+\begin{figure}[h]
+
+{\centering \includegraphics{figs/posture-1}
+
+}
+
+\caption[Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)]{Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel).}\label{fig:posture}
+\end{figure}
+\end{CodeChunk}
+
+\section{Changes in Access to Faces \&
+Hands}\label{changes-in-access-to-faces-hands}
+
+We first examined the proportion of face and hand detections across age;
+a full summary can be seen in Figure \ref{fig:detsbyAge}. We observed a
+slight u-shaped function both when analyzing the output of the MTCNN and
+OpenPose detectors, such that 12-month-olds may experience less faces
+than 8 or 12-month-olds. Next, we observed that hand detections
+increased with the age of the participant, though they were relatively
+infrequent in this dataset. As expected, we found that wrist detections
+we relatively more frequent and also appeared to increase somewhat with
+age, suggesting that caregivers hands may be just outside of the field
+of view of our camera or that they were often occluded by toys.
+
+Next, we turned to our hypothesis of interest, examinig the impact of
+postural and locomotive developments on children's visual access to
+faces and hands. Overall, we ofund that children's posture was a major
+factor both in how many faces they saw during the play session. Infants
+who were sitting saw more faces than infants who were lying down or
+being carried, while infants who were standing saw the most faces. We
+observed a similar pattern with respect to hands: children who were
+sitting or standing tended to see more hands/wrists than children who
+were lying down or held.
+
+We next examined how the child's orientation relative to their caregiver
+impacted their visual access to faces and hands. Overlal, we saw that
+children who were far away from their caregiver were more likely to see
+faces and hands than children who were close to their caregiver; this
+was true within all age groups and for all detectors.
 
@@
 
-{\centering \includegraphics{figs/unnamed-chunk-5-1}
+{\centering \includegraphics{figs/detByAge-1}
 
@@
 \caption[Proportion face (left) and hand (right) detections as a function of participant's age]{Proportion face (left) and hand (right) detections as a function of participant's age.
-Note that the axes are different; hands were detected overall much less frequently in this dataset.}\label{fig:unnamed-chunk-5}
+Note that the axes are different; hands were detected overall much less frequently in this dataset.}\label{fig:detByAge}
 \end{figure*}
@@
 
-\caption[Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)]{Proportion of face detections as a function of infant's posture, using the MTCNN3 detections (left panel) and OpenPose detections (right panel)}\label{fig:faceByPostureFig}
+\caption[Proportion of face/hand detections as a function of infant's posture (top panels) and orientation (bottom panels), using the MTCNN3 face detections and OpenPose face detections (left two panels) and using OpenPose hand detections and the OpenPose wrist detections (right two panels)]{Proportion of face/hand detections as a function of infant's posture (top panels) and orientation (bottom panels), using the MTCNN3 face detections and OpenPose face detections (left two panels) and using OpenPose hand detections and the OpenPose wrist detections (right two panels).}\label{fig:faceByPostureFig}
 \end{figure*}
@@
 
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/faceByOrientFig-1}
-
-}
-
-\caption[Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)]{Proportion of face detections as a function of caregiver's orientation,using the MTCNN3 detections (left panel) and OpenPose detections (right panel)}\label{fig:faceByOrientFig}
-\end{figure*}
-\end{CodeChunk}
+\subsection{Model}\label{model}
 
@@
 
-\section{Changes in Access to Hands}\label{changes-in-access-to-hands}
-
-Next, we examined whether we would see similar trends in infant and
-children's access to hands.
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -8.775 & 0.236 & -37.259 & 0.000 \\ 
+  Age & 0.879 & 0.166 & 5.292 & 0.000 \\ 
+  Posture-Prone  & 1.215 & 0.110 & 11.082 & 0.000 \\ 
+  Posture-Sit & 1.943 & 0.108 & 18.000 & 0.000 \\ 
+  Posture-Stand  & 2.175 & 0.107 & 20.260 & 0.000 \\ 
+  Orient.-Close  & 0.608 & 0.059 & 10.371 & 0.000 \\ 
+  Orient.-Far & 2.244 & 0.060 & 37.222 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.} 
+\end{table}
 

### 2018-01-14 analysis-main/merge_all_data.R
change assert to 45 with 1 indexing

@@ -55,7 +55,7 @@ alldets<-arrange(alldets,video,frame)
 
 ## OpenPose doesn't get all frames. Let's check out the frames it deleted somehow.
 missingInd=is.na(alldets$faceOP)
-assert_that(sum(is.na(alldets$faceOP))==57)
+assert_that(sum(is.na(alldets$faceOP))==45)
 
 missingOP<-alldets %>% 
   filter(missingInd) 

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
edits to text and citations

@@
 abstract: 
-  "The ability to process social cues (e.g., eye gaze) is a critical component of children's early language and cognitive development.
-However, as children reach their first birthday, they begin to locomote themselves, exploring their visual environment in an entirely new way.
-How do these postural changes affect children's access to the social information relevant for word-learning?
-Here, explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and by using computer-vision algorithms to detect the proportion of faces and hands in infants' environments.
-We find that infant's postural changes largely mediate infants' accesses to both faces and hands, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities."
+  "The ability to process social cues--including eye gaze--is a critical component of children's early language and cognitive development.
+However, as children reach their first birthday, they begin to locomote themselves, walking and exploring their visual environment in an entirely new way.
+How do these postural and locomotive changes affect children's access to the social information relevant for word-learning?
+Here, explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and then using modern computer-vision algorithms to detect the proportion of faces and hands in infants' environments.
+We find that infant's postural changes largely mediate infants' acess to both faces and hands, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
+Broadly, we suggest that the combined use of head-mounted cameras and the application of novel deep learning algorithms is a promising avenue for understanding the statistics of infants' visual and linguistic experience.
+"
 
 keywords:
-    "social cognition; face-perception; infancy; locomotion; head-cameras"
+    "social cognition; face-perception; infancy; locomotion; head-cameras; deep learning"
 
@@
 These motor changes have drastic consequences for what children see: crawling and walking infants have radically different views of the world.
-During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011); walking vs.
-crawling infants tend to have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).
+During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; walking vs.
+crawling infants tend to have full visual access to their environment and the people in it [@kretch2014].
 
 One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities (Iverson, 2010).
-Indeed, these postural changes also impact how children interact with their mothers;  walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”).
-(Karasik et al., 2014).
+Indeed, these postural changes also impact how children interact with their mothers;  walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].
 More directly, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
@@
 Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning what objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017].
-Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2017].
+Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
 
@@
 To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and their orientation to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We capitalize on recent improvements in computer vision algorithms (@citemtcnn, @openpose) that allow the automated detection of both faces and hands, and we analyze the frequency of both faces and hands in the child’s visual environment.
+We capitalize on recent improvements in computer vision algorithms (@zhang2016, @cao2017realtime) that allow the automated detection of both faces and hands, and we analyze the frequency of both faces and hands in the child’s visual environment.
 
@@
 We first measure the performance of the most commonly-used and widely avlaialbe face detection algorithmsm (Viola-Jones).
-We use this as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is is notoriously bad at dealing with occluded faces (@citeAnthony2016).
+We use this as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is is notoriously bad at dealing with occluded faces (@scheirer2014perceptual).
 We next capitalized on recent improvements in computer vision, testing the performance of two other algorithms that extract face information in distinct ways.
@@
 ### Face Detection Algorithms
-The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, CITEXX) applied to each individual frame.
+The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, @viola2004robust) applied to each individual frame.
 This detector provided information about the presence of a face as well as its size and position.
 
-The second algorithm was based on the work by (CITEXX) using multi-task cascaded convolutional neural networks (MTCNNs).
+The second algorithm was based on the work by @zhang2016 using multi-task cascaded convolutional neural networks (MTCNNs).
 The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
@@
 
-The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (CITEXX).
+The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
@@
 
-<!-- ```{r xtable, results="asis"} -->
-<!-- tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3, -->
-<!--                       caption = "Results from GLMM model prediting the presence/absence of a hands (OpenPose-Hands) across the entire dataset.") -->
-
-<!-- print(tab3, type="latex", comment = F, table.placement = "H") -->
-<!-- ``` -->
+# ```{r xtable, results="asis"}
+# tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3,
+#                       caption = "Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.")
+# 
+# print(tab3, type="latex", comment = F, table.placement = "H")
+# ```
 

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
new citation, brief edits to GD

@@
 
-# ```{r xtable, results="asis"}
-# tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3,
-#                       caption = "Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.")
-# 
-# print(tab3, type="latex", comment = F, table.placement = "H")
-# ```
+<!-- ```{r xtable, results="asis"} -->
+<!--  tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3, -->
+<!--                        caption = "Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.") -->
+
+<!--  print(tab3, type="latex", comment = F, table.placement = "H") -->
+<!-- ``` -->
 
@@
 This work also deploys novel advancements in computer vision to the study of developmental psychology.
-The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms (Krihevsky et al., 2012), creating a new generation of algorithmic tools that both are substantially better equipped to deal with noisier, more complicated datasets and that can  extract richer and more detailed information.
-Indeed, these videos from the infant perspective provided sustantial challenges (e.g., partially occluded faces) for the classc models of face detedction (e.g., viola jones).
-Further, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to new infant learners.
+The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms @krizhevsky2012imagenet, creating a new generation of algorithmic tools.
+These tools are both substantially better equipped to deal with noisier, more complicated datasets and that can extract richer and more detailed information.
+Videos from the infant perspective provided sustantial challenges (e.g., partially occluded faces) for the classc models of face detedction (e.g., ViolaJones, @viola2004robust).
+urther, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to infant learners.
 

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
Prettify figures

@@
 
-```{r detByAge, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=2, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age.
-Note that the axes are different; hands were detected overall much less frequently in this dataset." }
+```{r detByAge, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=6, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age."  }
 detectionsByAge <- dlong %>%
@@
 
-ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
+p1=ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
     geom_jitter(width = .25) +
@@
     scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
-    ylab("Proportion Detections") +
-    xlab("Age (months)") +
+    labs(y = "Prop. detected", x = element_blank()) +
+    theme_bw() + 
+    theme(legend.position="none") +
     scale_color_manual(values=c("#1F618D", "#7FB3D5", "#C70039", "#FF5733")) +
     facet_grid(~detectorType, scales = "free_y")
-    # theme_bw()
-```
-
-```{r faceByPostureFig, fig.env="figure*", fig.pos = "h", fig.align = "center", num.cols.cap=2, fig.width=7, fig.height=6, fig.cap = "Proportion of face/hand detections as a function of infant's posture (top panels) and orientation (bottom panels), using the MTCNN3 face detections and OpenPose face detections (left two panels) and using OpenPose hand detections and the OpenPose wrist detections (right two panels)."}
 
@@
 
-p1=ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
-  geom_boxplot() + 
-  ylab("Proportion  Detections") + 
-  xlab("Age (months)") + 
-  facet_grid(~detectorType) +
-  theme_bw()
+p2=ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
+  geom_boxplot(outlier.size = 0.5) + 
+  labs(y = "Prop. detected", x = element_blank()) +
+  facet_grid(~detectorType, scales="free_y") +
+  theme_bw() +
+  theme(legend.position="top", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
 
@@
 
-p2=ggplot(detectionsByOrientation, aes(x = factor(age.grp), y = detections, col = orientation)) + 
-  geom_boxplot() + 
-  ylab("Proportion  Detections") + 
-  xlab("Age (months)") + 
-  facet_grid(~detectorType) +
-  theme_bw()
+p3=ggplot(detectionsByOrientation, aes(x = factor(age.grp), y = detections, col = orientation)) + 
+  geom_boxplot(outlier.size = 0.5) + 
+  labs(y = "Prop. detected", x = element_blank()) +
+  facet_grid(~detectorType, scales="free_y") +
+  theme_bw() +
+  theme(legend.position="top", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
+  labs(y = "Prop. detected", x = element_blank())
 
-grid.arrange(p1,p2, nrow=2)
+grid.arrange(p1, p2, p3, nrow=3)
 
@@
 
-To formalize these observations, we fit a generalized logistic mixed-effect model to all with the presence/absence of a face on every frame (as detected by the MTCNN3 model) as the dependent variable, and participant's age, orientation, and posture as predictors.
+For our anlaysis, we fit models to the two models that the had the highest F-scores (i.e., highest performance on detecting faces and hands, respectively).
+To formalize these observations, we fit a generalized logistic mixed-effect model to all with the presence/absence of a face on every frame as the dependent variable, and participant's age, orientation, and posture as predictors.
 Interactions between predictors were not included as this maximal model failed to converge.
-A summary of the coefficients of this model can be found in Table {ref}.
+A summary of the coefficients of this model can be found in Table XX.
 While age remained a significant predictor even when accounting for the effects of infants' posture and caregiver's oirentation, it held relativley less predictive power.
@@
 \end{tabular}
-\caption{Results from GLMM model prediting the presence/absence of a face across  the entire dataset.} 
+\caption{Results from GLMM model prediting the presence/absence of a face across the entire dataset (MTCNN detectors).} 
 \end{table}

### 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
citation brackets redone

@@
 Indeed, children’s ability to effectively process these social cues may be a key factor in their early language development.
-For example,in a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies (@carpenter1998).
-More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months (@brooks2005) and throughout the second year of life (@brooks2008).
+For example,in a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies [@carpenter1998].
+More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months [@brooks2005] and throughout the second year of life ]@brooks2008].
 
@@
 
-One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities (Iverson, 2010).
+One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities [@iverson2010].
 Indeed, these postural changes also impact how children interact with their mothers;  walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].
@@
 
-Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different than the adult perspective (and not easily predicted by our own adult intuitions; @franchak2011, @yoshida2008; @clerkin2017).
-Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning what objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017].
+Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different than the adult perspective (and not easily predicted by our own adult intuitions; [@clerkin2017;@franchak2011, @yoshida2008] ).
+Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning what objects and their labels than the comparable adult perspective [ @bambach2017;@yurovsky2012].
 Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
@@
 To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and their orientation to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We capitalize on recent improvements in computer vision algorithms (@zhang2016, @cao2017realtime) that allow the automated detection of both faces and hands, and we analyze the frequency of both faces and hands in the child’s visual environment.
+We capitalize on recent improvements in computer vision algorithms [@zhang2016, @cao2017realtime] that allow the automated detection of both faces and hands, and we analyze the frequency of both faces and hands in the child’s visual environment.
 
@@
 
-\begin{table}[t]
+\begin{table}[H]
   \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}

### 2018-01-15 writeup-cogsci-copy/cogsci2018.Rmd
face detector results

--- 2018-01-14 writeup-cogsci-copy/cogsci2018.Rmd
@@
 
-\begin{table}[t]
-  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
-  \begin{center}
-    \begin{tabular}{l|ccc|ccc}
-      \hline
-       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
-      % \null Model & P & R & F & P & R & F  \\
-      \hline
-      % MTCNN & .84  & 1 &  .91 & .95 & .75 & .85   \\
-      % OpenPose & .99 & .87 & .93 & .71 & .92 & .8 \\
-    \hline
-    \end{tabular}
-  \end{center}
-\end{table}
+```{r, message=FALSE, warning=FALSE, results='asis'}
+data_mtcnn <- read_csv("../data/final_output/mtcnn3.csv") %>%
+  distinct(video, frame, .keep_all = TRUE)
+
+data_openpose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") %>%
+  distinct(name, frame, .keep_all = TRUE) %>%
+  mutate(frame = sprintf("%05d", as.numeric(frame) + 1)) %>%
+  mutate(is_face = ifelse(Nose_conf != 0, "True", "False")) %>%
+  select(group, name, frame, is_face) %>%
+  rename(video = name)
+
+original_candidates <- read_csv("../data/ground_truth/gold_set_candidates2.csv") %>%
+  arrange(video, frame)
+
+ground_truth_data <- read_csv("../data/ground_truth/ground_truth3.csv") %>%
+  select(-angle) %>%
+  arrange(video, frame) %>%
+  cbind(original_candidates %>%
+          select(sample_type))
+
+evaluate <- function(a, b) {
+  if (a == "True") {
+    if (a == b) return ("TP")
+    else return("FN")
+  }
+  else {
+    if (a == b) return("TN")
+    else return("FP")
+  }
+}
+
+mtcnn_results <- data_mtcnn %>%
+  rename(detector_is_face = is_face) %>%
+  inner_join(ground_truth_data, by = c("group", "video", "frame")) %>%
+  rowwise() %>%
+  mutate(result = evaluate(is_face, detector_is_face))
+
+openpose_results <- data_openpose %>%
+  rename(detector_is_face = is_face) %>%
+  inner_join(ground_truth_data, by = c("group", "video", "frame")) %>%
+  rowwise() %>%
+  mutate(result = evaluate(is_face, detector_is_face))
+
+get_prf <- function(df) {
+  tp <- sum(df$result == "TP")
+  fp <- sum(df$result == "FP")
+  fn <- sum(df$result == "FN")
+
+  p <- tp / (tp + fp)
+  r <- tp / (tp + fn)
+  f <- 2 * p * r / (p + r)
+  return(c(p, r, f))
+}
+
+render_table <- function(df, detector) {
+  hd <- df %>%
+    filter(sample_type == "high_density")
+
+  prf_vec <- get_prf(hd)
+
+  xtable::xtable(tibble(P = prf_vec[1], R = prf_vec[2], F = prf_vec[3]), caption=paste(detector, "High Density Sample")) %>%
+    print(type="latex", comment = F, table.placement = "H")
+
+  r <- df %>%
+    filter(sample_type == "random")
+
+  prf_vec <- get_prf(r)
+
+  xtable::xtable(tibble(P = prf_vec[1], R = prf_vec[2], F = prf_vec[3]), caption=paste(detector, "Random Sample")) %>%
+    print(type="latex", comment = F, table.placement = "H")
+}
+
+render_table(mtcnn_results, "MTCNN")
+render_table(openpose_results, "OpenPose")
+
+```
+
+```{r xtable, eval=FALSE, include=FALSE, results="asis"}
+  tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3,
+                        caption = "Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.")
+
+  print(tab3, type="latex", comment = F, table.placement = "H")
+```
 

### 2018-01-15 writeup-cogsci-copy/cogsci2018.tex
face detector results

@@
 \begin{abstract}
-The ability to process social cues (e.g., eye gaze) is a critical
+The ability to process social cues--including eye gaze--is a critical
 component of children's early language and cognitive development.
 However, as children reach their first birthday, they begin to locomote
-themselves, exploring their visual environment in an entirely new way.
-How do these postural changes affect children's access to the social
-information relevant for word-learning? Here, explore this question by
-using head-mounted cameras to record infants' (8-16 months of age)
-egocentric visual perspective and by using computer-vision algorithms to
-detect the proportion of faces and hands in infants' environments. We
-find that infant's postural changes largely mediate infants' accesses to
-both faces and hands, suggesting that these postural and locomotive
-developments facilitate infants' emerging linguistic and social
-capacities.
+themselves, walking and exploring their visual environment in an
+entirely new way. How do these postural and locomotive changes affect
+children's access to the social information relevant for word-learning?
+Here, explore this question by using head-mounted cameras to record
+infants' (8-16 months of age) egocentric visual perspective and then
+using modern computer-vision algorithms to detect the proportion of
+faces and hands in infants' environments. We find that infant's postural
+changes largely mediate infants' acess to both faces and hands,
+suggesting that these postural and locomotive developments facilitate
+infants' emerging linguistic and social capacities. Broadly, we suggest
+that the combined use of head-mounted cameras and the application of
+novel deep learning algorithms is a promising avenue for understanding
+the statistics of infants' visual and linguistic experience.
 
 \textbf{Keywords:}
-social cognition; face-perception; infancy; locomotion; head-cameras
+social cognition; face-perception; infancy; locomotion; head-cameras;
+deep learning
 \end{abstract}
@@
 recently, 10 month-olds who follow an adult's gaze in an experimental
-context have larger vocabularies at 18 months (Brooks \& Meltzoff
-(2005)) and throughout the second year of life (({\textbf{???}})).
+context have larger vocabularies at 18 months (R. Brooks \& Meltzoff
+(2005)) and throughout the second year of life (Rechele Brooks \&
+Meltzoff (2008)).
 
 However, as children are learning their first words, their view of the
-world is also changing radically (Adolph \& Berger, 2007). Infant's
+world is also changing radically (K. Adolph \& Berger, 2007). Infant's
 motor abilities improve dramatically near the end of the first year of
@@
 world. During spontaneous play in a laboratory playroom, toddlers are
-more likely to look at the floor while crawling than while walking
-(Franchak et al., 2011); walking vs.~crawling infants tend to have full
-visual access to their environment and the people in it (Kretch,
-Franchak, and Adolph, 2014).
+more likely to look at the floor while crawling than while walking (J.
+Franchak, Kretch, Soska, \& Adolph, 2011); walking vs.~crawling infants
+tend to have full visual access to their environment and the people in
+it (K. S. Kretch, Franchak, \& Adolph, 2014).
 
@@
 mothers than crawling infants, and tend to hear more action directed
-statements (e.g., ``open it''). (Karasik et al., 2014). More directly,
-in an observational study, Walle \& Campos (2014) found that children
-who were able to walk had both higher receptive and productive
-vocabularies. On their account, children's ability to stand and
-independently locomote may fundamentally change their ability to access
-the social information (e.g., faces, gaze) relative to children who are
-still crawling and sitting. In other words, the ability to access more
-detailed social information may allow infants to learn words quicker and
-more efficiently, facilitating language growth.
+statements (e.g., ``open it'') (Karasik, Tamis-LeMonda, \& Adolph,
+2014). More directly, in an observational study, Walle \& Campos (2014)
+found that children who were able to walk had both higher receptive and
+productive vocabularies. On their account, children's ability to stand
+and independently locomote may fundamentally change their ability to
+access the social information (e.g., faces, gaze) relative to children
+who are still crawling and sitting. In other words, the ability to
+access more detailed social information may allow infants to learn words
+quicker and more efficiently, facilitating language growth.
 
@@
 for walking children are radically different than the adult perspective
-(and not easily predicted by our own adult intuitions; Franchak, Kretch,
-Soska, \& Adolph (2011), Yoshida \& Smith (2008); ({\textbf{???}})).
-Children's views tend to be restricted and to be dominated by objects
-and hands much more than that of adults (Yoshida \& Smith, 2008), and
-both computational and empirical work suggest that this restricted
-viewpoint may be more effective for learning what objects and their
-labels than the comparable adult perspective (D. Yurovsky, Smith, \& Yu,
-in press, ({\textbf{???}})). Further, recent work also suggests dramatic
-changes in the child's perspective over the first two years of life, as
-views transition from primarily containing close up view of faces to
-capturing views of hands paired with the objects they are acting on
-({\textbf{???}}).
+(and not easily predicted by our own adult intuitions; J. Franchak et
+al. (2011), Yoshida \& Smith (2008); Clerkin, Hart, Rehg, Yu, \& Smith
+(2017)). Children's views tend to be restricted and to be dominated by
+objects and hands much more than that of adults (Yoshida \& Smith,
+2008), and both computational and empirical work suggest that this
+restricted viewpoint may be more effective for learning what objects and
+their labels than the comparable adult perspective (D. Yurovsky, Smith,
+\& Yu, in press, Bambach, Crandall, Smith, \& Yu (2017)). Further,
+recent work also suggests dramatic changes in the child's perspective
+over the first two years of life, as views transition from primarily
+containing close up view of faces to capturing views of hands paired
+with the objects they are acting on (Fausey, Jayaraman, \& Smith, 2016).
 
@@
 recorded from a third-person perspective and hand-annotated. We
-capitalize on recent improvements in computer vision algorithms
-(({\textbf{???}}), ({\textbf{???}})) that allow the automated detection
-of both faces and hands, and we analyze the frequency of both faces and
-hands in the child's visual environment.
+capitalize on recent improvements in computer vision algorithms (K.
+Zhang, Zhang, Li, \& Qiao (2016), Cao, Simon, Wei, \& Sheikh (2017))
+that allow the automated detection of both faces and hands, and we
+analyze the frequency of both faces and hands in the child's visual
+environment.
 
@@
 the side of the child. All annontations were made using
-OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
+OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman, Sanderson, \&
 Millman, 2012), and times when the child was out of view of the tripod
@@
 can achieve impressive accuracy in some situations, it is is notoriously
-bad at dealing with occluded faces (({\textbf{???}})). We next
-capitalized on recent improvements in computer vision, testing the
-performance of two other algorithms that extract face information in
-distinct ways. Rsults for the evaluation of these three systems are
-shown in Table XX.
+bad at dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox
+(2014)). We next capitalized on recent improvements in computer vision,
+testing the performance of two other algorithms that extract face
+information in distinct ways. Rsults for the evaluation of these three
+systems are shown in Table XX.
 
@@
 The first face detection system made use of a series of Haar
-feature-based cascade classifiers (Viola-Jones, CITEXX) applied to each
-individual frame. This detector provided information about the presence
-of a face as well as its size and position.
+feature-based cascade classifiers (Viola-Jones, Viola \& Jones (2004))
+applied to each individual frame. This detector provided information
+about the presence of a face as well as its size and position.
 
-The second algorithm was based on the work by (CITEXX) using multi-task
-cascaded convolutional neural networks (MTCNNs). The system was built
-using a novel cascaded CNN-based framework for joint detection and
-alignment, built to perform well in real-world environments where
-varying illuminations and occlusions are present. We used a Tensorflow
-implementation of this algorithm provided by
+The second algorithm was based on the work by K. Zhang et al. (2016)
+using multi-task cascaded convolutional neural networks (MTCNNs). The
+system was built using a novel cascaded CNN-based framework for joint
+detection and alignment, built to perform well in real-world
+environments where varying illuminations and occlusions are present. We
+used a Tensorflow implementation of this algorithm provided by
 (\url{https://github.com/davidsandberg/facenet}). Like Viola-Jones, this
@@
 The third algorithm was a CNN-based pose detector that provided the
-locations of 18 body parts (ears, nose, wrists, etc) (CITEXX). The
+locations of 18 body parts (ears, nose, wrists, etc) (Cao et al. (2017),
+Simon, Joo, Matthews, \& Sheikh (2017), Wei, Ramakrishna, Kanade, \&
+Sheikh (2016)) avaliable at
+\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
 system uses a CNN for initial anatomical detection and subsequently
@@
 
-\begin{table}[t]
-  \caption{Model performance on gold standard generalization training set dataset. P, R, and F denote precision, recall, and F-score for each of the two samples. \label{tab:model_eval}}
-  \begin{center}
-    \begin{tabular}{l|ccc|ccc}
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrr}
   \hline
-       &  \multicolumn{3}{c|}{High-density} &  \multicolumn{3}{c}{Random} \\
-      % \null Model & P & R & F & P & R & F  \\
+ & P & R & F \\ 
   \hline
-      % MTCNN & .84  & 1 &  .91 & .95 & .75 & .85   \\
-      % OpenPose & .99 & .87 & .93 & .71 & .92 & .8 \\
+1 & 0.84 & 1.00 & 0.91 \\ 
    \hline
 \end{tabular}
-  \end{center}
+\caption{MTCNN High Density Sample} 
+\end{table}\begin{table}[H]
+\centering
+\begin{tabular}{rrrr}
+  \hline
+ & P & R & F \\ 
+  \hline
+1 & 0.95 & 0.75 & 0.84 \\ 
+   \hline
+\end{tabular}
+\caption{MTCNN Random Sample} 
+\end{table}\begin{table}[H]
+\centering
+\begin{tabular}{rrrr}
+  \hline
+ & P & R & F \\ 
+  \hline
+1 & 0.99 & 0.87 & 0.93 \\ 
+   \hline
+\end{tabular}
+\caption{OpenPose High Density Sample} 
+\end{table}\begin{table}[H]
+\centering
+\begin{tabular}{rrrr}
+  \hline
+ & P & R & F \\ 
+  \hline
+1 & 0.71 & 0.92 & 0.80 \\ 
+   \hline
+\end{tabular}
+\caption{OpenPose Random Sample} 
 \end{table}
@@
 
-\caption[Proportion face (left) and hand (right) detections as a function of participant's age]{Proportion face (left) and hand (right) detections as a function of participant's age.
-Note that the axes are different; hands were detected overall much less frequently in this dataset.}\label{fig:detByAge}
-\end{figure*}
-\end{CodeChunk}
-
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/faceByPostureFig-1}
-
-}
-
-\caption[Proportion of face/hand detections as a function of infant's posture (top panels) and orientation (bottom panels), using the MTCNN3 face detections and OpenPose face detections (left two panels) and using OpenPose hand detections and the OpenPose wrist detections (right two panels)]{Proportion of face/hand detections as a function of infant's posture (top panels) and orientation (bottom panels), using the MTCNN3 face detections and OpenPose face detections (left two panels) and using OpenPose hand detections and the OpenPose wrist detections (right two panels).}\label{fig:faceByPostureFig}
+\caption[Proportion face (left) and hand (right) detections as a function of participant's age]{Proportion face (left) and hand (right) detections as a function of participant's age.}\label{fig:detByAge}
 \end{figure*}
@@
 
-To formalize these observations, we fit a generalized logistic
-mixed-effect model to all with the presence/absence of a face on every
-frame (as detected by the MTCNN3 model) as the dependent variable, and
-participant's age, orientation, and posture as predictors. Interactions
-between predictors were not included as this maximal model failed to
-converge. A summary of the coefficients of this model can be found in
-Table \{ref\}. While age remained a significant predictor even when
-accounting for the effects of infants' posture and caregiver's
-oirentation, it held relativley less predictive power. Overall, these
-results confirm that infant's access to faces is heavily infleunced by
-their own posture and their caregivers orientation towards them.
+For our anlaysis, we fit models to the two models that the had the
+highest F-scores (i.e., highest performance on detecting faces and
+hands, respectively). To formalize these observations, we fit a
+generalized logistic mixed-effect model to all with the presence/absence
+of a face on every frame as the dependent variable, and participant's
+age, orientation, and posture as predictors. Interactions between
+predictors were not included as this maximal model failed to converge. A
+summary of the coefficients of this model can be found in Table XX.
+While age remained a significant predictor even when accounting for the
+effects of infants' posture and caregiver's oirentation, it held
+relativley less predictive power. Overall, these results confirm that
+infant's access to faces is heavily infleunced by their own posture and
+their caregivers orientation towards them.
 
@@
 recognition has advanced dramatically in the past five years since the
-re-birth of deep learning algorithms (Krihevsky et al., 2012), creating
-a new generation of algorithmic tools that both are substantially better
-equipped to deal with noisier, more complicated datasets and that can
-extract richer and more detailed information. Indeed, these videos from
-the infant perspective provided sustantial challenges (e.g., partially
-occluded faces) for the classc models of face detedction (e.g., viola
-jones). Further, as the headcam technologies employed here were
-inexpensive (\textasciitilde{}\$60 a camera) and the computer vision
-algorithms freely avaliable, this method is a promising avenue for
-quanitfying the visual and social information avaliable to new infant
-learners.
+re-birth of deep learning algorithms Krizhevsky, Sutskever, \& Hinton
+(2012), creating a new generation of algorithmic tools. These tools are
+both substantially better equipped to deal with noisier, more
+complicated datasets and that can extract richer and more detailed
+information. Videos from the infant perspective provided sustantial
+challenges (e.g., partially occluded faces) for the classc models of
+face detedction (e.g., ViolaJones, Viola \& Jones (2004)). urther, as
+the headcam technologies employed here were inexpensive
+(\textasciitilde{}\$60 a camera) and the computer vision algorithms
+freely avaliable, this method is a promising avenue for quanitfying the
+visual and social information avaliable to infant learners.
 
@@
 
+\hypertarget{ref-bambach2017}{}
+Bambach, S., Crandall, D. J., Smith, L. B., \& Yu, C. (2017). An
+egocentric perspective on active vision and visual object learning in
+toddlers. In \emph{Proceedings of the seventh joint ieee conference on
+development and learning and on epigenetic robotics}.
+
 \hypertarget{ref-brooks2005}{}
@@
 535--543.
+
+\hypertarget{ref-brooks2008}{}
+Brooks, R., \& Meltzoff, A. N. (2008). Infant gaze following and
+pointing predict accelerated vocabulary growth through two years of age:
+A longitudinal, growth curve modeling study. \emph{Journal of Child
+Language}, \emph{35}(1), 207--220.
+
+\hypertarget{ref-cao2017realtime}{}
+Cao, Z., Simon, T., Wei, S.-E., \& Sheikh, Y. (2017). Realtime
+multi-person 2D pose estimation using part affinity fields. In
+\emph{CVPR}.
 
@@
 
+\hypertarget{ref-clerkin2017}{}
+Clerkin, E. M., Hart, E., Rehg, J. M., Yu, C., \& Smith, L. B. (2017).
+Real-world visual statistics and infants' first-learned object names.
+\emph{Phil. Trans. R. Soc. B}, \emph{372}(1711), 20160055.
+
 \hypertarget{ref-cummings1988}{}
@@
 Research}, \emph{29}(1), 7--16.
+
+\hypertarget{ref-fausey2016}{}
+Fausey, C. M., Jayaraman, S., \& Smith, L. B. (2016). From faces to
+hands: Changing visual input in the first two years. \emph{Cognition},
+\emph{152}, 101--107.
 
@@
 
+\hypertarget{ref-karasik2014}{}
+Karasik, L. B., Tamis-LeMonda, C. S., \& Adolph, K. E. (2014). Crawling
+and walking infants elicit different verbal responses from mothers.
+\emph{Developmental Science}, \emph{17}(3), 388--395.
+
+\hypertarget{ref-kretch2014}{}
+Kretch, K. S., Franchak, J. M., \& Adolph, K. E. (2014). Crawling and
+walking infants see the world differently. \emph{Child Development},
+\emph{85}(4), 1503--1518.
+
+\hypertarget{ref-krizhevsky2012imagenet}{}
+Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012). Imagenet
+classification with deep convolutional neural networks. In
+\emph{Advances in neural information processing systems} (pp.
+1097--1105).
+
 \hypertarget{ref-mayer1988}{}
@@
 
+\hypertarget{ref-scheirer2014perceptual}{}
+Scheirer, W. J., Anthony, S. E., Nakayama, K., \& Cox, D. D. (2014).
+Perceptual annotation: Measuring human vision to improve computer
+vision. \emph{IEEE Transactions on Pattern Analysis and Machine
+Intelligence}, \emph{36}(8), 1679--1686.
+
+\hypertarget{ref-simon2017hand}{}
+Simon, T., Joo, H., Matthews, I., \& Sheikh, Y. (2017). Hand keypoint
+detection in single images using multiview bootstrapping. In
+\emph{CVPR}.
+
+\hypertarget{ref-viola2004robust}{}
+Viola, P., \& Jones, M. J. (2004). Robust real-time face detection.
+\emph{International Journal of Computer Vision}, \emph{57}(2), 137--154.
+
 \hypertarget{ref-walle2014}{}
@@
 \emph{50}(2), 336.
+
+\hypertarget{ref-wei2016cpm}{}
+Wei, S.-E., Ramakrishna, V., Kanade, T., \& Sheikh, Y. (2016).
+Convolutional pose machines. In \emph{CVPR}.
 
@@
 
+\hypertarget{ref-zhang2016}{}
+Zhang, K., Zhang, Z., Li, Z., \& Qiao, Y. (2016). Joint face detection
+and alignment using multitask cascaded convolutional networks.
+\emph{IEEE Signal Processing Letters}, \emph{23}(10), 1499--1503.
+
 \end{document}

### 2018-01-15 writeup-cogsci-copy/cogsci2018.Rmd
wrist and hand

@@
 
-original_candidates <- read_csv("../data/ground_truth/gold_set_candidates2.csv") %>%
-  arrange(video, frame)
-
 ground_truth_data <- read_csv("../data/ground_truth/ground_truth3.csv") %>%
@@
   arrange(video, frame) %>%
-  cbind(original_candidates %>%
+  cbind(read_csv("../data/ground_truth/gold_set_candidates2.csv") %>%
+          arrange(video, frame) %>%
+          select(sample_type))
+
+data_openpose_wrists <- read_csv("../data/final_output/openpose_results_truncated_2.csv") %>%
+  distinct(name, frame, .keep_all = TRUE) %>%
+  mutate(is_hand = ifelse(LWrist_conf != 0 | RWrist_conf != 0, "True", "False")) %>%
+  select(group, name, frame, is_hand) %>%
+  rename(video = name)
+
+data_openpose_hands <- read_csv("../data/final_output/openpose_results_with_detail_hands.csv") %>%
+  distinct(name, frame, .keep_all = TRUE) %>%
+  mutate(video = name) %>%
+  distinct(video,frame,.keep_all=TRUE)  %>%
+  select(group, video,frame,ends_with("conf")) %>%
+  mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
+  mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
+  mutate(is_hand = ifelse(confRight>=.5 | confLeft>=.5, "True", "False"))
+
+ground_truth_data_hands <- read_csv("../data/ground_truth/ground_truth_hands.csv") %>%
+  rename(is_hand = is_face) %>%
+  select(-angle) %>%
+  arrange(video, frame) %>%
+  cbind(read_csv("../data/ground_truth/gold_set_hand_candidates.csv") %>%
+          arrange(video, frame) %>%
           select(sample_type))
@@
   mutate(result = evaluate(is_face, detector_is_face))
+
+openpose_wrist_results <- data_openpose_wrists %>%
+  rename(detector_is_hand = is_hand) %>%
+  inner_join(ground_truth_data_hands, by = c("group", "video", "frame")) %>%
+  rowwise() %>%
+  mutate(result = evaluate(is_hand, detector_is_hand))
+
+openpose_hand_results <- data_openpose_hands %>%
+  rename(detector_is_hand = is_hand) %>%
+  inner_join(ground_truth_data_hands, by = c("group", "video", "frame")) %>%
+  rowwise() %>%
+  mutate(result = evaluate(is_hand, detector_is_hand))
 
@@
 render_table(openpose_results, "OpenPose")
+render_table(openpose_wrist_results, "OpenPose Wrist")
+render_table(openpose_hand_results, "OpenPose Hand")
 

### 2018-01-15 writeup-cogsci-copy/cogsci2018.tex
wrist and hand

@@
 with their mother was found to predict both their receptive and
-productive vocabularies (Carpenter, Nagell, \& Tomasello (1998)). More
+productive vocabularies (Carpenter, Nagell, \& Tomasello, 1998). More
 recently, 10 month-olds who follow an adult's gaze in an experimental
-context have larger vocabularies at 18 months (R. Brooks \& Meltzoff
-(2005)) and throughout the second year of life (Rechele Brooks \&
-Meltzoff (2008)).
+context have larger vocabularies at 18 months (R. Brooks \& Meltzoff,
+2005) and throughout the second year of life {]}Rechele Brooks \&
+Meltzoff (2008){]}.
 
@@
 for walking children are radically different than the adult perspective
-(and not easily predicted by our own adult intuitions; J. Franchak et
-al. (2011), Yoshida \& Smith (2008); Clerkin, Hart, Rehg, Yu, \& Smith
-(2017)). Children's views tend to be restricted and to be dominated by
+(and not easily predicted by our own adult intuitions; (Clerkin, Hart,
+Rehg, Yu, \& Smith, 2017; J. Franchak et al., 2011, Yoshida \& Smith
+(2008)) ). Children's views tend to be restricted and to be dominated by
 objects and hands much more than that of adults (Yoshida \& Smith,
@@
 restricted viewpoint may be more effective for learning what objects and
-their labels than the comparable adult perspective (D. Yurovsky, Smith,
-\& Yu, in press, Bambach, Crandall, Smith, \& Yu (2017)). Further,
+their labels than the comparable adult perspective (Bambach, Crandall,
+Smith, \& Yu, 2017; D. Yurovsky, Smith, \& Yu, in press). Further,
 recent work also suggests dramatic changes in the child's perspective
@@
 capitalize on recent improvements in computer vision algorithms (K.
-Zhang, Zhang, Li, \& Qiao (2016), Cao, Simon, Wei, \& Sheikh (2017))
-that allow the automated detection of both faces and hands, and we
-analyze the frequency of both faces and hands in the child's visual
-environment.
+Zhang, Zhang, Li, \& Qiao, 2016, Cao, Simon, Wei, \& Sheikh (2017)) that
+allow the automated detection of both faces and hands, and we analyze
+the frequency of both faces and hands in the child's visual environment.
 
@@
 \caption{OpenPose Random Sample} 
+\end{table}\begin{table}[H]
+\centering
+\begin{tabular}{rrrr}
+  \hline
+ & P & R & F \\ 
+  \hline
+1 & 0.65 & 0.85 & 0.74 \\ 
+   \hline
+\end{tabular}
+\caption{OpenPose Wrist High Density Sample} 
+\end{table}\begin{table}[H]
+\centering
+\begin{tabular}{rrrr}
+  \hline
+ & P & R & F \\ 
+  \hline
+1 & 0.32 & 0.13 & 0.19 \\ 
+   \hline
+\end{tabular}
+\caption{OpenPose Wrist Random Sample} 
+\end{table}\begin{table}[H]
+\centering
+\begin{tabular}{rrrr}
+  \hline
+ & P & R & F \\ 
+  \hline
+1 & 1.00 & 0.21 & 0.35 \\ 
+   \hline
+\end{tabular}
+\caption{OpenPose Hand High Density Sample} 
+\end{table}\begin{table}[H]
+\centering
+\begin{tabular}{rrrr}
+  \hline
+ & P & R & F \\ 
+  \hline
+1 & 1.00 & 0.04 & 0.09 \\ 
+   \hline
+\end{tabular}
+\caption{OpenPose Hand Random Sample} 
 \end{table}
@@
 
+\hypertarget{ref-iverson2010}{}
+Iverson, J. M. (2010). Developing language in a developing body: The
+relationship between motor development and language development.
+\emph{Journal of Child Language}, \emph{37}(2), 229--261.
+
 \hypertarget{ref-karasik2014}{}

### 2018-01-15 writeup-cogsci-copy/cogsci2018.Rmd
>1 keypioint

@@
   mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
-  mutate(is_hand = ifelse(confRight>=.5 | confLeft>=.5, "True", "False"))
+  mutate(is_hand = ifelse(confRight>=.05 | confLeft>=.05, "True", "False"))
 

### 2018-01-15 analysis-main/helper.R
added age.at.test to namings helper

@@ -162,7 +162,7 @@ summarize.naming <- function (x, window = c(-2,2)) {
   }
   
   namings$age.grp <- x$age.grp[1]
-
+  namings$age.at.test <- x$age.at.test[1]
   return(namings)
 }
 

### 2018-01-15 analysis-main/main_analyses_bll.Rmd
integrate hand detectors etc.
slight updates; ref new data file

@@
 ```{r}
-d <- read_csv("../data/consolidateddata/consolidated_data_3dets.csv") %>%
+d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
   mutate(posture = factor(posture), 
@@
 #d$face=d$faceVJ
-#d$face=d$faceMT
-d$face=d$faceOP
+d$face=d$faceMT
+#d$face=d$faceOP
 #d$face=d$handOP

### 2018-01-15 writeup-cogsci-copy/cogsci2018_faceOnly.Rmd
Face only cogsci paper

--- 2018-01-15 writeup-cogsci-copy/cogsci2018.Rmd
@@
 How do these postural and locomotive changes affect children's access to the social information relevant for word-learning?
-Here, explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and then using modern computer-vision algorithms to detect the proportion of faces and hands in infants' environments.
-We find that infant's postural changes largely mediate infants' acess to both faces and hands, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
+Here, explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and then using modern computer-vision algorithms to detect the proportion of faces in infants' environments.
+We find that infant's postural changes largely mediate infants' acess to faces, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
 Broadly, we suggest that the combined use of head-mounted cameras and the application of novel deep learning algorithms is a promising avenue for understanding the statistics of infants' visual and linguistic experience.
@@
 To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and their orientation to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We capitalize on recent improvements in computer vision algorithms [@zhang2016, @cao2017realtime] that allow the automated detection of both faces and hands, and we analyze the frequency of both faces and hands in the child’s visual environment.
+We capitalize on recent improvements in computer vision algorithms [@zhang2016, @cao2017realtime] that allow the automated detection of both faces and hands, and we analyze the frequency of faces in the child’s visual environment.
 
@@
 
-The second of these algorithms (OpenPose) also captures information about the "skeleton" of people in a given frame, outputting information about the entire set of body parts that were detected.
+<!-- The second of these algorithms (OpenPose) also captures information about the "skeleton" of people in a given frame, outputting information about the entire set of body parts that were detected.
 This allowed us to extract hand information as well as wrist information.
@@
 For example,if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy).
+-->
 
@@
 For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose).
-We operationalized face detections as any frames containing a nose and either the left or right eye; we operationalized hand detections as any frames containing any of the 21 points detected on either the left or the right hand, and we operationalized wrist detections any frames containing either the left or the right wrist.
+We operationalized face detections as any frames containing a nose, as any half of a face necessarily contains a nose.
+<!-- we operationalized hand detections as any frames containing any of the 21 points detected on either the left or the right hand, and we operationalized wrist detections any frames containing either the left or the right wrist.
+-->
 
@@
 
-render_table(mtcnn_results, "MTCNN")
-render_table(openpose_results, "OpenPose")
-render_table(openpose_wrist_results, "OpenPose Wrist")
-render_table(openpose_hand_results, "OpenPose Hand")
+# render_table(mtcnn_results, "MTCNN")
+# render_table(openpose_results, "OpenPose")
 
@@
 
-```{r xtable, eval=FALSE, include=FALSE, results="asis"}
-  tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3,
-                        caption = "Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.")
-
-  print(tab3, type="latex", comment = F, table.placement = "H")
-```
-
-Evaluating hand/wrist detectors (text here about how gold standard was selected)
-
 \begin{table}[H]
 \centering
-\begin{tabular}{rrrrr}
+\begin{tabular}{rrrr}
   \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  & P & R & F \\ 
 \hline
-(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
-  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
-  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
-  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
-  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
-  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
-  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
+OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\ 
+OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\
+MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\ 
+MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\ 
    \hline
 \end{tabular}
-\caption{TABLE PLACEHOLDER} 
+\caption{Face detector performance} 
 \end{table}
@@
 First, we report developmental shifts in infants' posture and their orientation to their caregiver.
-Then, we explore how these changes influence children's visual access to faces and to hands.
-Finally, we explore how these changes impact the accessibility of faces and hands during labeling events.
+Then, we explore how these changes influence children's visual access to faces across this developmental time range.
+Finally, we explore how these changes impact the accessibility of faces during labeling events.
 
@@
   mutate(id = paste(subid,frame)) %>%
-  gather(key = detectorType, value = detection, faceMT, faceOP, handOP, wristOP) %>%
+  gather(key = detectorType, value = detection, faceMT, faceOP) %>%
   mutate(detectorType=as.factor(detectorType))
@@
 #[1] "faceMT"  "faceOP"  "handOP"  "wristOP"
-levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","OpenPose-Hands", "OpenPose-Wrists") 
+levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces") 
 ```
@@
 
-# Changes in Access to Faces & Hands
+# Changes in Access to Faces
 
-We first examined the proportion of face and hand detections across age; a full summary can be seen in Figure \ref{fig:detsbyAge}.
-We observed  a slight u-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds may experience less faces than 8 or 12-month-olds.
-Next, we observed that hand detections increased with the age of the participant, though they were relatively infrequent in this dataset.
-As expected, we found that wrist detections we relatively more frequent and also appeared to increase somewhat with age, suggesting that caregivers hands may be just outside of the field of view of our camera or that they were often occluded by toys.
+We first examined the proportion of face detections across age; a full summary can be seen in Figure \ref{fig:detsbyAge}.
+We observed  a slight u-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slighly fewer faces faces than 8 or 16-month-olds.
 
-Next, we turned to our hypothesis of interest, examinig the impact of postural and locomotive developments on children's visual access to faces and hands.
+Next, we turned to our hypothesis of interest, examinig the impact of postural and locomotive developments on children's visual access to faces.
 Overall, we ofund that children's posture was a major factor both in how many faces they saw during the play session.
 Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces.
-We observed a similar pattern with respect to hands: children who were sitting or standing tended to see more hands/wrists than children who were lying down or held.
-
-We next examined how the child's orientation relative to their caregiver impacted their visual access to faces and hands.
-Overlal, we saw that children who were far away from their caregiver were more likely to see faces and hands than children who were close to their caregiver; this was true within all age groups and for all detectors.
+We next examined how the child's orientation relative to their caregiver impacted their visual access to faces.
+Overall, we saw that children who were far away from their caregiver were more likely to see faces  than children who were close to their caregiver; this was true within all age groups and for all detectors.
 
@@
 
-For our anlaysis, we fit models to the two models that the had the highest F-scores (i.e., highest performance on detecting faces and hands, respectively).
 To formalize these observations, we fit a generalized logistic mixed-effect model to all with the presence/absence of a face on every frame as the dependent variable, and participant's age, orientation, and posture as predictors.
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-(Intercept) & -8.775 & 0.236 & -37.259 & 0.000 \\ 
-  Age & 0.879 & 0.166 & 5.292 & 0.000 \\ 
-  Posture-Prone  & 1.215 & 0.110 & 11.082 & 0.000 \\ 
-  Posture-Sit & 1.943 & 0.108 & 18.000 & 0.000 \\ 
-  Posture-Stand  & 2.175 & 0.107 & 20.260 & 0.000 \\ 
-  Orient.-Close  & 0.608 & 0.059 & 10.371 & 0.000 \\ 
-  Orient.-Far & 2.244 & 0.060 & 37.222 & 0.000 \\ 
-   \hline
-\end{tabular}
-\caption{Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.} 
-\end{table}
+## Access to Faces During Labeling Events
+Finally, we analyzed how face detections changed during object labeling events as a function of infant's posture and orientation.
+Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labelling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
+We again found that infant's posture the degree to which they saw their caregivers face during a labeling event; infants who were sitting or standing were more likely to have access to their caregivers face.
+
+```{r echo=FALSE, include=FALSE }
+source("helper.R")
+d$face=d$faceMT
+aroundNamings <- d %>%
+  group_by(subid) %>%
+  do(summarize.naming(.))
+
+aroundNamings <- aroundNamings %>%
+  filter(!is.na(posture)) %>%
+  filter(!is.na(orientation))
+
+# aroundNamingsModel <- lmer(face ~ scale(age.at.test) + posture + orientation + (1|subid), 
+#          data = aroundNamings)
+
+```
+
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections as a function of infants's posture"  }
+ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
+  geom_boxplot() + 
+  ylab("Proportion of face detections within naming window") + 
+  xlab("Posture") + 
+  theme_bw() 
+```
 
 # General Discussion
-We use a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information relevant for word-learning, here operationalized as the presence of the faces and hands of their caregiver.
-We found that children's posture and orientation towards their caregiver changed systematically across age, and that all three of these factors dramatically impacted the proportion of faces and hands that were avaliable in the child's visual field.
-Thus, infants postural development is mediating factor that explains age-related changes in the proportion of faces and hands avaliable in infants' visual field.
+We use a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information relevant for word-learning, here operationalized as the presence of the faces of their caregiver.
+We found that children's posture and orientation towards their caregiver changed systematically across age, and that all three of these factors dramatically impacted the proportion of faces that were avaliable in the child's visual field.
+Thus, infants postural development is mediating factor that explains age-related changes in the proportion of faces avaliable in infants' visual field.
 
@@
 Videos from the infant perspective provided sustantial challenges (e.g., partially occluded faces) for the classc models of face detedction (e.g., ViolaJones, @viola2004robust).
-urther, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to infant learners.
+Further, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to infant learners.
 

### 2018-01-15 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
renaming directories

--- 2018-01-15 writeup-cogsci-copy/cogsci2018.tex
@@
 using modern computer-vision algorithms to detect the proportion of
-faces and hands in infants' environments. We find that infant's postural
-changes largely mediate infants' acess to both faces and hands,
-suggesting that these postural and locomotive developments facilitate
-infants' emerging linguistic and social capacities. Broadly, we suggest
-that the combined use of head-mounted cameras and the application of
-novel deep learning algorithms is a promising avenue for understanding
-the statistics of infants' visual and linguistic experience.
+faces in infants' environments. We find that infant's postural changes
+largely mediate infants' acess to faces, suggesting that these postural
+and locomotive developments facilitate infants' emerging linguistic and
+social capacities. Broadly, we suggest that the combined use of
+head-mounted cameras and the application of novel deep learning
+algorithms is a promising avenue for understanding the statistics of
+infants' visual and linguistic experience.
 
@@
 allow the automated detection of both faces and hands, and we analyze
-the frequency of both faces and hands in the child's visual environment.
+the frequency of faces in the child's visual environment.
 
@@
 
-The second of these algorithms (OpenPose) also captures information
-about the ``skeleton'' of people in a given frame, outputting
-information about the entire set of body parts that were detected. This
-allowed us to extract hand information as well as wrist information. We
-analyze wrist detections as a good proxy for the presence of a hand. We
-do so first because we did not capture the entire visual field
-experienced by the infants/children. Thus evne though the video may only
-show the presence of a wrist, the caregiver's hand may be within the
-child's field of view. Second, we do so because hands are often occluded
-by objects when caregivers are interacting with children, yet still
-visually accessible by the child and part of their joint interaction.
-For example,if a caregiver was holding a toy and presenting it to the
-infant, their wrists may be visible but their hand may not necessarily
-be (as it is occluded by the toy).
-
 \subsubsection{Face Detection
@@
 face (ears, eyes, nose). We operationalized face detections as any
-frames containing a nose and either the left or right eye; we
-operationalized hand detections as any frames containing any of the 21
-points detected on either the left or the right hand, and we
-operationalized wrist detections any frames containing either the left
-or the right wrist.
+frames containing a nose, as any half of a face necessarily contains a
+nose.\\
 
@@
 \hline
-1 & 0.84 & 1.00 & 0.91 \\ 
+OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\ 
+OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\
+MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\ 
+MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\ 
    \hline
 \end{tabular}
-\caption{MTCNN High Density Sample} 
-\end{table}\begin{table}[H]
-\centering
-\begin{tabular}{rrrr}
-  \hline
- & P & R & F \\ 
-  \hline
-1 & 0.95 & 0.75 & 0.84 \\ 
-   \hline
-\end{tabular}
-\caption{MTCNN Random Sample} 
-\end{table}\begin{table}[H]
-\centering
-\begin{tabular}{rrrr}
-  \hline
- & P & R & F \\ 
-  \hline
-1 & 0.99 & 0.87 & 0.93 \\ 
-   \hline
-\end{tabular}
-\caption{OpenPose High Density Sample} 
-\end{table}\begin{table}[H]
-\centering
-\begin{tabular}{rrrr}
-  \hline
- & P & R & F \\ 
-  \hline
-1 & 0.71 & 0.92 & 0.80 \\ 
-   \hline
-\end{tabular}
-\caption{OpenPose Random Sample} 
-\end{table}\begin{table}[H]
-\centering
-\begin{tabular}{rrrr}
-  \hline
- & P & R & F \\ 
-  \hline
-1 & 0.65 & 0.85 & 0.74 \\ 
-   \hline
-\end{tabular}
-\caption{OpenPose Wrist High Density Sample} 
-\end{table}\begin{table}[H]
-\centering
-\begin{tabular}{rrrr}
-  \hline
- & P & R & F \\ 
-  \hline
-1 & 0.32 & 0.13 & 0.19 \\ 
-   \hline
-\end{tabular}
-\caption{OpenPose Wrist Random Sample} 
-\end{table}\begin{table}[H]
-\centering
-\begin{tabular}{rrrr}
-  \hline
- & P & R & F \\ 
-  \hline
-1 & 1.00 & 0.21 & 0.35 \\ 
-   \hline
-\end{tabular}
-\caption{OpenPose Hand High Density Sample} 
-\end{table}\begin{table}[H]
-\centering
-\begin{tabular}{rrrr}
-  \hline
- & P & R & F \\ 
-  \hline
-1 & 1.00 & 0.04 & 0.09 \\ 
-   \hline
-\end{tabular}
-\caption{OpenPose Hand Random Sample} 
-\end{table}
-
-Evaluating hand/wrist detectors (text here about how gold standard was
-selected)
-
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
-  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
-  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
-  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
-  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
-  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
-  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
-   \hline
-\end{tabular}
-\caption{TABLE PLACEHOLDER} 
+\caption{Face detector performance} 
 \end{table}
@@
 orientation to their caregiver. Then, we explore how these changes
-influence children's visual access to faces and to hands. Finally, we
-explore how these changes impact the accessibility of faces and hands
-during labeling events.
+influence children's visual access to faces across this developmental
+time range. Finally, we explore how these changes impact the
+accessibility of faces during labeling events.
 
@@
 
-\section{Changes in Access to Faces \&
-Hands}\label{changes-in-access-to-faces-hands}
+\section{Changes in Access to Faces}\label{changes-in-access-to-faces}
 
-We first examined the proportion of face and hand detections across age;
-a full summary can be seen in Figure \ref{fig:detsbyAge}. We observed a
-slight u-shaped function both when analyzing the output of the MTCNN and
-OpenPose detectors, such that 12-month-olds may experience less faces
-than 8 or 12-month-olds. Next, we observed that hand detections
-increased with the age of the participant, though they were relatively
-infrequent in this dataset. As expected, we found that wrist detections
-we relatively more frequent and also appeared to increase somewhat with
-age, suggesting that caregivers hands may be just outside of the field
-of view of our camera or that they were often occluded by toys.
+We first examined the proportion of face detections across age; a full
+summary can be seen in Figure \ref{fig:detsbyAge}. We observed a slight
+u-shaped function both when analyzing the output of the MTCNN and
+OpenPose detectors, such that 12-month-olds appeared to experience
+slighly fewer faces faces than 8 or 16-month-olds.
 
@@
 postural and locomotive developments on children's visual access to
-faces and hands. Overall, we ofund that children's posture was a major
-factor both in how many faces they saw during the play session. Infants
-who were sitting saw more faces than infants who were lying down or
-being carried, while infants who were standing saw the most faces. We
-observed a similar pattern with respect to hands: children who were
-sitting or standing tended to see more hands/wrists than children who
-were lying down or held.
-
-We next examined how the child's orientation relative to their caregiver
-impacted their visual access to faces and hands. Overlal, we saw that
-children who were far away from their caregiver were more likely to see
-faces and hands than children who were close to their caregiver; this
-was true within all age groups and for all detectors.
+faces. Overall, we ofund that children's posture was a major factor both
+in how many faces they saw during the play session. Infants who were
+sitting saw more faces than infants who were lying down or being
+carried, while infants who were standing saw the most faces. We next
+examined how the child's orientation relative to their caregiver
+impacted their visual access to faces. Overall, we saw that children who
+were far away from their caregiver were more likely to see faces than
+children who were close to their caregiver; this was true within all age
+groups and for all detectors.
 
@@
 
-For our anlaysis, we fit models to the two models that the had the
-highest F-scores (i.e., highest performance on detecting faces and
-hands, respectively). To formalize these observations, we fit a
-generalized logistic mixed-effect model to all with the presence/absence
-of a face on every frame as the dependent variable, and participant's
-age, orientation, and posture as predictors. Interactions between
-predictors were not included as this maximal model failed to converge. A
-summary of the coefficients of this model can be found in Table XX.
-While age remained a significant predictor even when accounting for the
-effects of infants' posture and caregiver's oirentation, it held
-relativley less predictive power. Overall, these results confirm that
-infant's access to faces is heavily infleunced by their own posture and
-their caregivers orientation towards them.
+To formalize these observations, we fit a generalized logistic
+mixed-effect model to all with the presence/absence of a face on every
+frame as the dependent variable, and participant's age, orientation, and
+posture as predictors. Interactions between predictors were not included
+as this maximal model failed to converge. A summary of the coefficients
+of this model can be found in Table XX. While age remained a significant
+predictor even when accounting for the effects of infants' posture and
+caregiver's oirentation, it held relativley less predictive power.
+Overall, these results confirm that infant's access to faces is heavily
+infleunced by their own posture and their caregivers orientation towards
+them.
 
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-(Intercept) & -8.775 & 0.236 & -37.259 & 0.000 \\ 
-  Age & 0.879 & 0.166 & 5.292 & 0.000 \\ 
-  Posture-Prone  & 1.215 & 0.110 & 11.082 & 0.000 \\ 
-  Posture-Sit & 1.943 & 0.108 & 18.000 & 0.000 \\ 
-  Posture-Stand  & 2.175 & 0.107 & 20.260 & 0.000 \\ 
-  Orient.-Close  & 0.608 & 0.059 & 10.371 & 0.000 \\ 
-  Orient.-Far & 2.244 & 0.060 & 37.222 & 0.000 \\ 
-   \hline
-\end{tabular}
-\caption{Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.} 
-\end{table}
+\subsection{Access to Faces During Labeling
+Events}\label{access-to-faces-during-labeling-events}
+
+Finally, we analyzed how face detections changed during object labeling
+events as a function of infant's posture and orientation. Specfically,
+we analyzed a four-second window around each labeling event (e.g.,
+``Look at the {[}zem{]}!''); these labelling events were hand-annotated
+and automatically synchronized with the frame-by-frame face detections.
+We again found that infant's posture the degree to which they saw their
+caregivers face during a labeling event; infants who were sitting or
+standing were more likely to have access to their caregivers face.
+
+\begin{CodeChunk}
+\begin{figure}[H]
+
+{\centering \includegraphics{figs/detByNaming-1}
+
+}
+
+\caption[Proportion face detections as a function of infants's posture]{Proportion face detections as a function of infants's posture}\label{fig:detByNaming}
+\end{figure}
+\end{CodeChunk}
 
@@
 information relevant for word-learning, here operationalized as the
-presence of the faces and hands of their caregiver. We found that
-children's posture and orientation towards their caregiver changed
-systematically across age, and that all three of these factors
-dramatically impacted the proportion of faces and hands that were
-avaliable in the child's visual field. Thus, infants postural
-development is mediating factor that explains age-related changes in the
-proportion of faces and hands avaliable in infants' visual field.
+presence of the faces of their caregiver. We found that children's
+posture and orientation towards their caregiver changed systematically
+across age, and that all three of these factors dramatically impacted
+the proportion of faces that were avaliable in the child's visual field.
+Thus, infants postural development is mediating factor that explains
+age-related changes in the proportion of faces avaliable in infants'
+visual field.
 
@@
 challenges (e.g., partially occluded faces) for the classc models of
-face detedction (e.g., ViolaJones, Viola \& Jones (2004)). urther, as
+face detedction (e.g., ViolaJones, Viola \& Jones (2004)). Further, as
 the headcam technologies employed here were inexpensive

### 2018-01-15 writeup-cogsci-jan14/helper.R

@@ -1,0 +1,179 @@
+library(stringr)
+
+######## MERGE IN EXACT TIMES FOR FRAMES #########
+add.times <- function(x) {
+  print(x$subid[1])
+  fname <- paste0("../data/frame_times/",
+                 str_sub(as.character(x$subid[1]),start=4,end=8),
+                 ".csv")
+  if (file.exists(fname)) {
+    times <- read_csv(fname)
+    
+#     x <- merge(x,times, by.x = "frame", by.y = "frame", all.x = TRUE, all.y = FALSE)
+    x <- left_join(x, times)
+    x$dt <- c(diff(x$time),.032) ## assumes data is sorted by frames! also not sure why .032?
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing frames ****"))
+    x$time <- NA
+    x$dt <- NA
+  }
+  
+  return(x)
+}
+
+######## ADD POSTURE CODING #########
+add.posture <- function(x) {
+  fname <- paste("../data/posture/",
+                 x$subid[1],
+                 ".csv",
+                 sep="")
+  
+  if (file.exists(fname)) {
+    print(fname)
+    postures <- read_csv(fname)
+    
+    postures <- regularize.postures(subset(postures,code=="posture"))
+    
+    x$posture <- factor(NA,levels=levels(postures$posture))
+    x$orientation <- factor(NA,levels=levels(postures$orientation))
+    
+    # for each row of p, populate posture and orientation to x
+    for (i in 1:nrow(postures)) {
+      range <- x$time > postures$start[i] & x$time <= postures$end[i]
+      x$posture[range] <- postures$posture[i]
+      x$orientation[range] <- postures$orientation[i]
+    }
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing posture ****"))
+    x$posture <- NA
+    x$orientation <- NA
+  }
+  
+  return(x)
+}
+
+######## GET POSTURE CODING SORTED OUT #########
+regularize.postures <- function (p) {
+  original.postures <- c("p","si","st","l","c","w","NIF")
+  
+  # clean up postures
+  p$type <- factor(p$type,levels=original.postures)
+  p$type <- plyr::revalue(p$type,c("p"="prone",
+                   "si"="sit",
+                   "st"="stand",
+                   "l"="prone",
+                   "c"="carry",
[…]

### 2018-01-15 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
eval text

--- 2018-01-15 writeup-cogsci-copy/cogsci2018_faceOnly.Rmd
@@
 author-information: > 
-    \author{{\large \bf Alessandro Sanchez} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Stanford University
+    \author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University
     \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
@@
 ### Detector evaluation
-In order to evaluate the performance of these detectors, we constructed a "gold set" of frames
-Evaluating face detectors (text here about how gold standard was selected)
+In order to evaluate the performance of these detectors, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (as reported by the detectors) and a random sample from the remaining frames.
+This was done so as to not bias our evaluation by the relatively rare appearance of faces in the dataset.
+A face was present in a frame if at least half of the face was showing.
+Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated and are reported in Table 2.
+Both detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
+Due to high performance of these two detectors we have included the results of both in the analyses to follow.
 

### 2018-01-15 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
eval text

@@
 
-\author{{\large \bf Alessandro Sanchez} \\ \texttt{author1@university.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
+\author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
     \And {\large \bf Ally Kraus} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
@@
 In order to evaluate the performance of these detectors, we constructed
-a ``gold set'' of frames Evaluating face detectors (text here about how
-gold standard was selected)
+a ``gold set'' of frames by hand labeling both a sample of frames with a
+high density of faces (as reported by the detectors) and a random sample
+from the remaining frames. This was done so as to not bias our
+evaluation by the relatively rare appearance of faces in the dataset. A
+face was present in a frame if at least half of the face was showing.
+Precision (hits / hits + false alarms), recall (hits / hits + misses),
+and F-score (harmonic mean of previous measures) were calculated and are
+reported in Table 2. Both detectors performed relatively well on the
+gold set, with MTCNN outperforming OpenPose on the random sample but
+trailing behind in the high density sample. Due to high performance of
+these two detectors we have included the results of both in the analyses
+to follow.
 

### 2018-01-15 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
typos / edits

@@
 How do these postural and locomotive changes affect children's access to the social information relevant for word-learning?
-Here, explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and then using modern computer-vision algorithms to detect the proportion of faces in infants' environments.
-We find that infant's postural changes largely mediate infants' acess to faces, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
+Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and modern computer vision algorithms to detect the proportion of faces in infants' environments.
+We find that infant's postural changes largely mediate infants' access to faces, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
 Broadly, we suggest that the combined use of head-mounted cameras and the application of novel deep learning algorithms is a promising avenue for understanding the statistics of infants' visual and linguistic experience.
@@
 
-Here, we directly ask whether the postural changes that infants’ experience as they reach their first birthday change the availability of the social information relevant for word learning.
+Here, we directly ask whether the postural changes that infants experience as they reach their first birthday change the availability of the social information relevant for word learning.
 To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and their orientation to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We capitalize on recent improvements in computer vision algorithms [@zhang2016, @cao2017realtime] that allow the automated detection of both faces and hands, and we analyze the frequency of faces in the child’s visual environment.
+We capitalize on recent improvements in face detection algorithms [@zhang2016, @cao2017realtime] and analyze the frequences of faces in the child’s visual environment.
 
@@
 ## Head-mounted camera
-```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for three different headcam configurations, with the device we used in the middle.
+```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for three different headcam configurations, with the device we used in the middle.
 The lowest camera is pictured for comparison, but was not available until after our study was already in progress."}
@@
 Videos captured by the headcam were 720x480 pixels with 25 frames per second.Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
-A fisheye lens was attached to the camera to cincrease the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left).
+A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left).
 
 Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's approximate vertical field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988].
-As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirely of the child's upper visual field where the child is likely to see the faces of adults around them.
+As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see the faces of adults around them.
 This allowed us to maximize our chances of capturing faces that the child would have seen during the play session.
@@
 The caregiver's orientation was characterized as being close to the child, farther from the child, and a global category of caregiver behind the child.
-For the first two annotationes (close/far from the child), the caregiver could either be to the the front or to the side of the child.
-All annontations were made using OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera was marked as uncodable and was excluded from these annotations.
+For the first two annotations (close/far from the child), the caregiver could either be to the the front or to the side of the child.
+All annotations were made using OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera was marked as uncodable and was excluded from these annotations.
 
@@
 Measuring the availability of caregiver’s faces across development was an important goal of the study.
-We thus used a total of three face detection systems to explore this hypothesis and to avoid the cost of hand-annotating every frame.
-We first measure the performance of the most commonly-used and widely avlaialbe face detection algorithmsm (Viola-Jones).
-We use this as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is is notoriously bad at dealing with occluded faces (@scheirer2014perceptual).
-We next capitalized on recent improvements in computer vision, testing the performance of two other algorithms that extract face information in distinct ways.
-Rsults for the evaluation of these three systems are shown in Table XX.
+We thus used a total of two face detection systems to explore this hypothesis and to avoid the cost of hand-annotating every frame.
+We capitalized on recent improvements in computer vision, testing the performance of two algorithms that extract face information in distinct ways.
+Results for the evaluation of these three systems are shown in Table 2.
 
@@
 ### Face Detection Algorithms
-The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, @viola2004robust) applied to each individual frame.
+The first algorithm was based on the work by @zhang2016 using multi-task cascaded convolutional neural networks (MTCNNs).
+The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
+We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet).
 This detector provided information about the presence of a face as well as its size and position.
 
-The second algorithm was based on the work by @zhang2016 using multi-task cascaded convolutional neural networks (MTCNNs).
-The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
-We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet).
-Like Viola-Jones, this detector provided information about the presence of a face as well as it's size and position.
-
-The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The second algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
@@
 ```
-We noted characteristic changes in infants posture and orientation across this developmental time range.
-The proportion of time infant's spent sitting decreased with age, and the proportion of time infant's spent standing increased with age.
-Both 8-month-olds and 12-month-olds psent equivalent amounts of time either lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time either sitting or standing (see Figure \ref{fig:posture}).
+We noted characteristic changes in infants' posture and orientation across this developmental time range.
+The proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with age.
+Both 8-month-olds and 12-month-olds spent equivalent amounts of time either lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time either sitting or standing (see Figure \ref{fig:posture}).
 We also observed characteristic changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:orientation}).
@@
 We first examined the proportion of face detections across age; a full summary can be seen in Figure \ref{fig:detsbyAge}.
-We observed  a slight u-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slighly fewer faces faces than 8 or 16-month-olds.
+We observed  a slight U-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slighly fewer faces faces than 8 or 16-month-olds.
 
-Next, we turned to our hypothesis of interest, examinig the impact of postural and locomotive developments on children's visual access to faces.
-Overall, we ofund that children's posture was a major factor both in how many faces they saw during the play session.
+Next, we turned to our hypothesis of interest, examining the impact of postural and locomotive developments on children's visual access to faces.
+Overall, we found that children's posture was a major factor both in how many faces they saw during the play session.
 Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces.
@@
 Interactions between predictors were not included as this maximal model failed to converge.
-A summary of the coefficients of this model can be found in Table XX.
-While age remained a significant predictor even when accounting for the effects of infants' posture and caregiver's oirentation, it held relativley less predictive power.
+A summary of the coefficients of this model can be found in Table 3.
+While age remained a significant predictor even when accounting for the effects of infants' posture and caregivers' orientation, it held relatively less predictive power.
 Overall, these results confirm that infant's access to faces is heavily infleunced by their own posture and their caregivers orientation towards them.
@@
 Finally, we analyzed how face detections changed during object labeling events as a function of infant's posture and orientation.
-Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labelling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
-We again found that infant's posture the degree to which they saw their caregivers face during a labeling event; infants who were sitting or standing were more likely to have access to their caregivers face.
+Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
+We again found that infant's posture impacted the degree to which they saw their caregivers face during a labeling event; infants who were sitting or standing were more likely to have access to their caregiver's face.
 

### 2018-01-15 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
typos / edits

@@
 children's access to the social information relevant for word-learning?
-Here, explore this question by using head-mounted cameras to record
-infants' (8-16 months of age) egocentric visual perspective and then
-using modern computer-vision algorithms to detect the proportion of
-faces in infants' environments. We find that infant's postural changes
-largely mediate infants' acess to faces, suggesting that these postural
-and locomotive developments facilitate infants' emerging linguistic and
-social capacities. Broadly, we suggest that the combined use of
-head-mounted cameras and the application of novel deep learning
-algorithms is a promising avenue for understanding the statistics of
-infants' visual and linguistic experience.
+Here, we explore this question by using head-mounted cameras to record
+infants' (8-16 months of age) egocentric visual perspective and modern
+computer vision algorithms to detect the proportion of faces in infants'
+environments. We find that infant's postural changes largely mediate
+infants' access to faces, suggesting that these postural and locomotive
+developments facilitate infants' emerging linguistic and social
+capacities. Broadly, we suggest that the combined use of head-mounted
+cameras and the application of novel deep learning algorithms is a
+promising avenue for understanding the statistics of infants' visual and
+linguistic experience.
 
@@
 
-Here, we directly ask whether the postural changes that infants'
+Here, we directly ask whether the postural changes that infants
 experience as they reach their first birthday change the availability of
@@
 recorded from a third-person perspective and hand-annotated. We
-capitalize on recent improvements in computer vision algorithms (K.
-Zhang, Zhang, Li, \& Qiao, 2016, Cao, Simon, Wei, \& Sheikh (2017)) that
-allow the automated detection of both faces and hands, and we analyze
-the frequency of faces in the child's visual environment.
+capitalize on recent improvements in face detection algorithms (K.
+Zhang, Zhang, Li, \& Qiao, 2016, Cao, Simon, Wei, \& Sheikh (2017)) and
+analyze the frequences of faces in the child's visual environment.
 
@@
 
-{\centering \includegraphics{figs/image-1}
+{\centering \includegraphics{figs/headcam-1}
 
@@
 \caption[Field of view for three different headcam configurations, with the device we used in the middle]{Field of view for three different headcam configurations, with the device we used in the middle.
-The lowest camera is pictured for comparison, but was not available until after our study was already in progress.}\label{fig:image}
+The lowest camera is pictured for comparison, but was not available until after our study was already in progress.}\label{fig:headcam}
 \end{figure}
@@
 \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
-A fisheye lens was attached to the camera to cincrease the view angle
+A fisheye lens was attached to the camera to increase the view angle
 from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
@@
 primarily interested in the presence of faces in the child's field of
-view, we chose to orient the camera upwards to capture the entirely of
+view, we chose to orient the camera upwards to capture the entirety of
 the child's upper visual field where the child is likely to see the
@@
 close to the child, farther from the child, and a global category of
-caregiver behind the child. For the first two annotationes (close/far
+caregiver behind the child. For the first two annotations (close/far
 from the child), the caregiver could either be to the the front or to
-the side of the child. All annontations were made using
-OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman, Sanderson, \&
-Millman, 2012), and times when the child was out of view of the tripod
-camera was marked as uncodable and was excluded from these annotations.
+the side of the child. All annotations were made using OpenSHAPA/Datavyu
+software (K. Adolph, Gilmore, Freeman, Sanderson, \& Millman, 2012), and
+times when the child was out of view of the tripod camera was marked as
+uncodable and was excluded from these annotations.
 
@@
 Measuring the availability of caregiver's faces across development was
-an important goal of the study. We thus used a total of three face
+an important goal of the study. We thus used a total of two face
 detection systems to explore this hypothesis and to avoid the cost of
-hand-annotating every frame. We first measure the performance of the
-most commonly-used and widely avlaialbe face detection algorithmsm
-(Viola-Jones). We use this as a benchmark for performance, as while it
-can achieve impressive accuracy in some situations, it is is notoriously
-bad at dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox
-(2014)). We next capitalized on recent improvements in computer vision,
-testing the performance of two other algorithms that extract face
-information in distinct ways. Rsults for the evaluation of these three
-systems are shown in Table XX.
+hand-annotating every frame. We capitalized on recent improvements in
+computer vision, testing the performance of two algorithms that extract
+face information in distinct ways. Results for the evaluation of these
+three systems are shown in Table 2.
 
@@
 
-The first face detection system made use of a series of Haar
-feature-based cascade classifiers (Viola-Jones, Viola \& Jones (2004))
-applied to each individual frame. This detector provided information
-about the presence of a face as well as its size and position.
-
-The second algorithm was based on the work by K. Zhang et al. (2016)
+The first algorithm was based on the work by K. Zhang et al. (2016)
 using multi-task cascaded convolutional neural networks (MTCNNs). The
@@
 
-The third algorithm was a CNN-based pose detector that provided the
+The second algorithm was a CNN-based pose detector that provided the
 locations of 18 body parts (ears, nose, wrists, etc) (Cao et al. (2017),
@@
 
-We noted characteristic changes in infants posture and orientation
-across this developmental time range. The proportion of time infant's
-spent sitting decreased with age, and the proportion of time infant's
+We noted characteristic changes in infants' posture and orientation
+across this developmental time range. The proportion of time infants
+spent sitting decreased with age, and the proportion of time infants
 spent standing increased with age. Both 8-month-olds and 12-month-olds
-psent equivalent amounts of time either lying/crawling, which was
+spent equivalent amounts of time either lying/crawling, which was
 markedly decreased in the 16-month-olds, who spent most of their time
@@
 summary can be seen in Figure \ref{fig:detsbyAge}. We observed a slight
-u-shaped function both when analyzing the output of the MTCNN and
+U-shaped function both when analyzing the output of the MTCNN and
 OpenPose detectors, such that 12-month-olds appeared to experience
@@
 
-Next, we turned to our hypothesis of interest, examinig the impact of
+Next, we turned to our hypothesis of interest, examining the impact of
 postural and locomotive developments on children's visual access to
-faces. Overall, we ofund that children's posture was a major factor both
+faces. Overall, we found that children's posture was a major factor both
 in how many faces they saw during the play session. Infants who were
@@
 as this maximal model failed to converge. A summary of the coefficients
-of this model can be found in Table XX. While age remained a significant
+of this model can be found in Table 3. While age remained a significant
 predictor even when accounting for the effects of infants' posture and
-caregiver's oirentation, it held relativley less predictive power.
+caregivers' orientation, it held relatively less predictive power.
 Overall, these results confirm that infant's access to faces is heavily
@@
 we analyzed a four-second window around each labeling event (e.g.,
-``Look at the {[}zem{]}!''); these labelling events were hand-annotated
+``Look at the {[}zem{]}!''); these labeling events were hand-annotated
 and automatically synchronized with the frame-by-frame face detections.
-We again found that infant's posture the degree to which they saw their
-caregivers face during a labeling event; infants who were sitting or
-standing were more likely to have access to their caregivers face.
+We again found that infant's posture impacted the degree to which they
+saw their caregivers face during a labeling event; infants who were
+sitting or standing were more likely to have access to their caregiver's
+face.
 
@@
 
-\hypertarget{ref-scheirer2014perceptual}{}
-Scheirer, W. J., Anthony, S. E., Nakayama, K., \& Cox, D. D. (2014).
-Perceptual annotation: Measuring human vision to improve computer
-vision. \emph{IEEE Transactions on Pattern Analysis and Machine
-Intelligence}, \emph{36}(8), 1679--1686.
-
 \hypertarget{ref-simon2017hand}{}

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
more edits

@@
 Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and modern computer vision algorithms to detect the proportion of faces in infants' environments.
-We find that infant's postural changes largely mediate infants' access to faces, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
-Broadly, we suggest that the combined use of head-mounted cameras and the application of novel deep learning algorithms is a promising avenue for understanding the statistics of infants' visual and linguistic experience.
-"
+We find that infants' postural changes largely mediate infants' access to faces, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
+Broadly, we suggest that the combined use of head-mounted cameras and the application of novel deep learning algorithms is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
 
@@
 Indeed, children’s ability to effectively process these social cues may be a key factor in their early language development.
-For example,in a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies [@carpenter1998].
-More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months [@brooks2005] and throughout the second year of life ]@brooks2008].
+For example, in a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies [@carpenter1998].
+More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months [@brooks2005] and throughout the second year of life [@brooks2008].
 
 However, as children are learning their first words, their view of the world is also changing radically [@adolph2007].
-Infant’s motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see.
+Infants' motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see.
 These motor changes have drastic consequences for what children see: crawling and walking infants have radically different views of the world.
@@
 
-Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different than the adult perspective (and not easily predicted by our own adult intuitions; [@clerkin2017;@franchak2011, @yoshida2008] ).
-Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning what objects and their labels than the comparable adult perspective [ @bambach2017;@yurovsky2012].
-Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
+Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different from the adult perspective (and not easily predicted by our own adult intuitions [@clerkin2017;@franchak2011, @yoshida2008]).
+Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [ @bambach2017;@yurovsky2012].
+Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing a close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
 
 Here, we directly ask whether the postural changes that infants experience as they reach their first birthday change the availability of the social information relevant for word learning.
-To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and their orientation to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We capitalize on recent improvements in face detection algorithms [@zhang2016, @cao2017realtime] and analyze the frequences of faces in the child’s visual environment.
+To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
+We capitalize on recent improvements in face detection algorithms [@zhang2016, @cao2017realtime] to analyze the frequences of faces in the child’s visual environment.
 
@@
 
-After the child was comfortable with wearing the headcam, the child and caregiver were shown to a playroom for the free-play session--the focus of the current study.
-Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a feather duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previosuly seen the novel toys and were instructed to use the novel labels to refer to the novel toys.
+After the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session--the focus of the current study.
+Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a feather duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the novel toys.
 
@@
 The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing.
-The caregiver's orientation was characterized as being close to the child, farther from the child, and a global category of caregiver behind the child.
+The caregiver's orientation was characterized as being close to the child, far from the child, and a global category of caregiver behind the child.
 For the first two annotations (close/far from the child), the caregiver could either be to the the front or to the side of the child.
-All annotations were made using OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera was marked as uncodable and was excluded from these annotations.
+All annotations were made using OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.
 
 ## Face Detection
-Measuring the availability of caregiver’s faces across development was an important goal of the study.
-We thus used a total of two face detection systems to explore this hypothesis and to avoid the cost of hand-annotating every frame.
-We capitalized on recent improvements in computer vision, testing the performance of two algorithms that extract face information in distinct ways.
-Results for the evaluation of these three systems are shown in Table 2.
+Measuring the availability of caregiver’s faces across development was an important goal of this study.
+We used a total of two face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame.
+We capitalized on recent improvements in computer vision algorithms, testing the performance of two state-of-the-art face detectors that used novel methods to extract face information.
 
@@
 
-The second algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The second algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
@@
 Both detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
-Due to high performance of these two detectors we have included the results of both in the analyses to follow.
+Due to the high performance of these two detectors we have included the results of both in the analyses to follow.
 
@@
 # Results
-First, we report developmental shifts in infants' posture and their orientation to their caregiver.
+First, we report developmental shifts in infants' posture and their orientation relative to their caregiver.
 Then, we explore how these changes influence children's visual access to faces across this developmental time range.
@@
 
-```{r orientation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
+```{r orientation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each orientation relative to their caregiver." }
 
@@
 
-```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)." }
+```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each posture." }
 ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
@@
 
-We first examined the proportion of face detections across age; a full summary can be seen in Figure \ref{fig:detsbyAge}.
+We first examined the proportion of face detections across age; a full summary can be seen in Figure \ref{fig:detByAge}.
 We observed  a slight U-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slighly fewer faces faces than 8 or 16-month-olds.
@@
 While age remained a significant predictor even when accounting for the effects of infants' posture and caregivers' orientation, it held relatively less predictive power.
-Overall, these results confirm that infant's access to faces is heavily infleunced by their own posture and their caregivers orientation towards them.
+Overall, these results confirm that infants' access to faces is heavily infleunced by their own posture and their caregiver's orientation towards them.
 
@@
 \end{tabular}
-\caption{Results from GLMM model prediting the presence/absence of a face across the entire dataset (MTCNN detectors).} 
+\caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).} 
 \end{table}
@@
 Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
-We again found that infant's posture impacted the degree to which they saw their caregivers face during a labeling event; infants who were sitting or standing were more likely to have access to their caregiver's face.
+We again found that infant's posture impacted the degree to which they saw their caregiver's face during a labeling event; infants who were sitting or standing were more likely to have access to their caregiver's face.
 
@@
 We found that children's posture and orientation towards their caregiver changed systematically across age, and that all three of these factors dramatically impacted the proportion of faces that were avaliable in the child's visual field.
-Thus, infants postural development is mediating factor that explains age-related changes in the proportion of faces avaliable in infants' visual field.
+Thus, infants' postural development is a mediating factor that explains age-related changes in the proportion of faces avaliable in infants' visual field.
 
@@
 The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms @krizhevsky2012imagenet, creating a new generation of algorithmic tools.
-These tools are both substantially better equipped to deal with noisier, more complicated datasets and that can extract richer and more detailed information.
-Videos from the infant perspective provided sustantial challenges (e.g., partially occluded faces) for the classc models of face detedction (e.g., ViolaJones, @viola2004robust).
+These tools are substantially better equipped to deal with noisier, more complicated datasets and can extract richer and more detailed information.
+Videos from the infant perspective provided substantial challenges (e.g., partially occluded faces) for the classic models of face detection (e.g., ViolaJones, @viola2004robust).
 Further, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to infant learners.

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
more edits

@@
 computer vision algorithms to detect the proportion of faces in infants'
-environments. We find that infant's postural changes largely mediate
+environments. We find that infants' postural changes largely mediate
 infants' access to faces, suggesting that these postural and locomotive
@@
 cues may be a key factor in their early language development. For
-example,in a longitudinal study, children's level of joint engagement
+example, in a longitudinal study, children's level of joint engagement
 with their mother was found to predict both their receptive and
@@
 context have larger vocabularies at 18 months (R. Brooks \& Meltzoff,
-2005) and throughout the second year of life {]}Rechele Brooks \&
-Meltzoff (2008){]}.
+2005) and throughout the second year of life (Rechele Brooks \&
+Meltzoff, 2008).
 
@@
 for walking children are radically different than the adult perspective
-(and not easily predicted by our own adult intuitions; (Clerkin, Hart,
+(and not easily predicted by our own adult intuitions (Clerkin, Hart,
 Rehg, Yu, \& Smith, 2017; J. Franchak et al., 2011, Yoshida \& Smith
-(2008)) ). Children's views tend to be restricted and to be dominated by
+(2008))). Children's views tend to be restricted and to be dominated by
 objects and hands much more than that of adults (Yoshida \& Smith,
 2008), and both computational and empirical work suggest that this
-restricted viewpoint may be more effective for learning what objects and
+restricted viewpoint may be more effective for learning objects and
 their labels than the comparable adult perspective (Bambach, Crandall,
@@
 over the first two years of life, as views transition from primarily
-containing close up view of faces to capturing views of hands paired
+containing a close up view of faces to capturing views of hands paired
 with the objects they are acting on (Fausey, Jayaraman, \& Smith, 2016).
@@
 head-mounted cameras during a brief laboratory free-play session;
-children's posture and their orientation to their caregiver were also
+children's posture and orientation relative to their caregiver were also
 recorded from a third-person perspective and hand-annotated. We
 capitalize on recent improvements in face detection algorithms (K.
-Zhang, Zhang, Li, \& Qiao, 2016, Cao, Simon, Wei, \& Sheikh (2017)) and
+Zhang, Zhang, Li, \& Qiao, 2016, Cao, Simon, Wei, \& Sheikh (2017)) to
 analyze the frequences of faces in the child's visual environment.
@@
 
-After the child was comfortable with wearing the headcam, the child and
+After the child was comfortable wearing the headcam, the child and
 caregiver were shown to a playroom for the free-play session--the focus
@@
 child one at a time, ``as they typically would.'' All parents confirmed
-that their child had not previosuly seen the novel toys and were
+that their child had not previously seen the novel toys and were
 instructed to use the novel labels to refer to the novel toys.
@@
 or standing. The caregiver's orientation was characterized as being
-close to the child, farther from the child, and a global category of
+close to the child, far from the child, and a global category of
 caregiver behind the child. For the first two annotations (close/far
@@
 software (K. Adolph, Gilmore, Freeman, Sanderson, \& Millman, 2012), and
-times when the child was out of view of the tripod camera was marked as
-uncodable and was excluded from these annotations.
+times when the child was out of view of the tripod camera were marked as
+uncodable and were excluded from these annotations.
 
@@
 Measuring the availability of caregiver's faces across development was
-an important goal of the study. We thus used a total of two face
-detection systems to explore this hypothesis and to avoid the cost of
-hand-annotating every frame. We capitalized on recent improvements in
-computer vision, testing the performance of two algorithms that extract
-face information in distinct ways. Results for the evaluation of these
-three systems are shown in Table 2.
+an important goal of this study. We used a total of two face detection
+systems to explore infants' changing access to social information and to
+avoid the cost of hand-annotating every frame. We capitalized on recent
+improvements in computer vision algorithms, testing the performance of
+two state-of-the-art face detectors that used novel methods to extract
+face information.
 
@@
 used a Tensorflow implementation of this algorithm provided by
-(\url{https://github.com/davidsandberg/facenet}). Like Viola-Jones, this
-detector provided information about the presence of a face as well as
-it's size and position.
+(\url{https://github.com/davidsandberg/facenet}). This detector provided
+information about the presence of a face as well as its size and
+position.
 
 The second algorithm was a CNN-based pose detector that provided the
-locations of 18 body parts (ears, nose, wrists, etc) (Cao et al. (2017),
-Simon, Joo, Matthews, \& Sheikh (2017), Wei, Ramakrishna, Kanade, \&
-Sheikh (2016)) avaliable at
+locations of 18 body parts (ears, nose, wrists, etc.) (Cao et al.
+(2017), Simon, Joo, Matthews, \& Sheikh (2017), Wei, Ramakrishna,
+Kanade, \& Sheikh (2016)) avaliable at
 \url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
@@
 gold set, with MTCNN outperforming OpenPose on the random sample but
-trailing behind in the high density sample. Due to high performance of
-these two detectors we have included the results of both in the analyses
-to follow.
+trailing behind in the high density sample. Due to the high performance
+of these two detectors we've included the results of both in the
+analyses to follow.
 
@@
 We first examined the proportion of face detections across age; a full
-summary can be seen in Figure \ref{fig:detsbyAge}. We observed a slight
+summary can be seen in Figure \ref{fig:detByAge}. We observed a slight
 U-shaped function both when analyzing the output of the MTCNN and
@@
 the proportion of faces that were avaliable in the child's visual field.
-Thus, infants postural development is mediating factor that explains
+Thus, infants' postural development is a mediating factor that explains
 age-related changes in the proportion of faces avaliable in infants'
@@
 (2012), creating a new generation of algorithmic tools. These tools are
-both substantially better equipped to deal with noisier, more
-complicated datasets and that can extract richer and more detailed
-information. Videos from the infant perspective provided sustantial
-challenges (e.g., partially occluded faces) for the classc models of
-face detedction (e.g., ViolaJones, Viola \& Jones (2004)). Further, as
-the headcam technologies employed here were inexpensive
-(\textasciitilde{}\$60 a camera) and the computer vision algorithms
-freely avaliable, this method is a promising avenue for quanitfying the
-visual and social information avaliable to infant learners.
+substantially better equipped to deal with noisier, more complicated
+datasets and can extract richer and more detailed information. Videos
+from the infant perspective provided substantial challenges (e.g.,
+partially occluded faces) for the classic models of face detection
+(e.g., ViolaJones, Viola \& Jones (2004)). Further, as the headcam
+technologies employed here were inexpensive (\textasciitilde{}\$60 a
+camera) and the computer vision algorithms freely avaliable, this method
+is a promising avenue for quanitfying the visual and social information
+avaliable to infant learners.
 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
detector evaluation in tidyr

@@
 How do these postural and locomotive changes affect children's access to the social information relevant for word-learning?
-Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and modern computer vision algorithms to detect the proportion of faces in infants' environments.
-We find that infants' postural changes largely mediate infants' access to faces, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
+Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use state-of-the-art computer vision algorithms to detect the proportion of faces in infants' environments.
+We find that infants' postural changes largely mediate infants' access to faces, hinting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
 Broadly, we suggest that the combined use of head-mounted cameras and the application of novel deep learning algorithms is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
@@
 
-However, as children are learning their first words, their view of the world is also changing radically [@adolph2007].
-Infants' motor abilities improve dramatically near the end of the first year of life,  allowing them to locomote independently and to determine what they see.
-These motor changes have drastic consequences for what children see: crawling and walking infants have radically different views of the world.
-During spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; walking vs.
-crawling infants tend to have full visual access to their environment and the people in it [@kretch2014].
+However, as children are learning their first words, their view of the world is also radically changing [@adolph2007].
+Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently.
+These motor changes have drastic consequences for what children see; crawling and walking infants have radically different views of the world.
+For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
 
 One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities [@iverson2010].
-Indeed, these postural changes also impact how children interact with their mothers;  walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].
-More directly, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
-On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access the social information (e.g., faces, gaze) relative to children who are still crawling and sitting.
+Indeed, these postural changes also impact how children interact with their mothers;  walking (vs.
+crawling) infants make different kinds of object-related bids for attention from their mothers and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].
+Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
+On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access social information (e.g., faces, gaze) relative to children who are still crawling and sitting.
 In other words, the ability to access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.
 
-Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different from the adult perspective (and not easily predicted by our own adult intuitions [@clerkin2017;@franchak2011, @yoshida2008]).
-Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [ @bambach2017;@yurovsky2012].
+Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011; @yoshida2008].
+Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2012].
 Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing a close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
 
-Here, we directly ask whether the postural changes that infants experience as they reach their first birthday change the availability of the social information relevant for word learning.
+Here, we directly ask whether the postural and locomotive developments change the availability of the social information relevant for word learning.
 To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We capitalize on recent improvements in face detection algorithms [@zhang2016, @cao2017realtime] to analyze the frequences of faces in the child’s visual environment.
+We capitalize on recent improvements in face detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces in the child’s visual environment, both overall and relative to naming events by their caregivers.
 
@@
 ```{r, message=FALSE, warning=FALSE, results='asis'}
-data_mtcnn <- read_csv("../data/final_output/mtcnn3.csv") %>%
-  distinct(video, frame, .keep_all = TRUE)
+### Needs to be consolidated so that we're aren't rereading all of these files here...
+# data_vj<- read_csv("../data/final_output/viola.csv") %>%
+#   distinct(video, frame, .keep_all = TRUE)
+# 
+# data_mtcnn <- read_csv("../data/final_output/mtcnn3.csv") %>%
+#   distinct(video, frame, .keep_all = TRUE)
+# 
+# data_openpose <- read_csv("../data/final_output/openpose_results_truncated_2.csv")%>%
+#   distinct(name, frame, .keep_all = TRUE) %>%
+#   mutate(frame = sprintf("%05d", as.numeric(frame) + 1)) %>%
+#   mutate(is_face = ifelse(Nose_conf != 0, "True", "False")) %>%
+#   select(group, name, frame, is_face) %>%
+#   rename(video = name)
 
-data_openpose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") %>%
-  distinct(name, frame, .keep_all = TRUE) %>%
-  mutate(frame = sprintf("%05d", as.numeric(frame) + 1)) %>%
-  mutate(is_face = ifelse(Nose_conf != 0, "True", "False")) %>%
-  select(group, name, frame, is_face) %>%
-  rename(video = name)
+d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
+  mutate(posture = factor(posture), orientation = factor(orientation))
 
+## make long form version for plotting both face detections on same graph
+dlong <- d %>%
+  mutate(id = paste(subid,frame)) %>%
+  gather(key = detectorType, value = detection, faceMT, faceOP, faceVJ) %>%
+  mutate(detectorType=as.factor(detectorType))
+
+## couple of helper functions
+
+evaluate <- function(a, b) {
+  if (a == TRUE) {
+    if (a == b) return ("TP") # was face, said face
+    else return("FN") # was face, missed face
+  }
+  else {
+    if (a == b) return("TN") # was not face, said not face
+    else return("FP") # was not face, said face
+  }
+}
+
+## okay, big long tidyr chunk
 ground_truth_data <- read_csv("../data/ground_truth/ground_truth3.csv") %>%
@@
 
-data_openpose_wrists <- read_csv("../data/final_output/openpose_results_truncated_2.csv") %>%
-  distinct(name, frame, .keep_all = TRUE) %>%
-  mutate(is_hand = ifelse(LWrist_conf != 0 | RWrist_conf != 0, "True", "False")) %>%
-  select(group, name, frame, is_hand) %>%
-  rename(video = name)
-
-data_openpose_hands <- read_csv("../data/final_output/openpose_results_with_detail_hands.csv") %>%
-  distinct(name, frame, .keep_all = TRUE) %>%
-  mutate(video = name) %>%
-  distinct(video,frame,.keep_all=TRUE)  %>%
-  select(group, video,frame,ends_with("conf")) %>%
-  mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
-  mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
-  mutate(is_hand = ifelse(confRight>=.05 | confLeft>=.05, "True", "False"))
-
-ground_truth_data_hands <- read_csv("../data/ground_truth/ground_truth_hands.csv") %>%
-  rename(is_hand = is_face) %>%
-  select(-angle) %>%
-  arrange(video, frame) %>%
-  cbind(read_csv("../data/ground_truth/gold_set_hand_candidates.csv") %>%
-          arrange(video, frame) %>%
-          select(sample_type))
-
-evaluate <- function(a, b) {
-  if (a == "True") {
-    if (a == b) return ("TP")
-    else return("FN")
-  }
-  else {
-    if (a == b) return("TN")
-    else return("FP")
-  }
-}
-
-mtcnn_results <- data_mtcnn %>%
-  rename(detector_is_face = is_face) %>%
-  inner_join(ground_truth_data, by = c("group", "video", "frame")) %>%
+performance <- ground_truth_data  %>%
+  mutate(frame=as.numeric(frame), subid=video) %>%
+  left_join(dlong)   %>%
+  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
   rowwise() %>%
-  mutate(result = evaluate(is_face, detector_is_face))
-
-openpose_results <- data_openpose %>%
-  rename(detector_is_face = is_face) %>%
-  inner_join(ground_truth_data, by = c("group", "video", "frame")) %>%
-  rowwise() %>%
-  mutate(result = evaluate(is_face, detector_is_face))
-
-openpose_wrist_results <- data_openpose_wrists %>%
-  rename(detector_is_hand = is_hand) %>%
-  inner_join(ground_truth_data_hands, by = c("group", "video", "frame")) %>%
-  rowwise() %>%
-  mutate(result = evaluate(is_hand, detector_is_hand))
-
-openpose_hand_results <- data_openpose_hands %>%
-  rename(detector_is_hand = is_hand) %>%
-  inner_join(ground_truth_data_hands, by = c("group", "video", "frame")) %>%
-  rowwise() %>%
-  mutate(result = evaluate(is_hand, detector_is_hand))
-
-get_prf <- function(df) {
-  tp <- sum(df$result == "TP")
-  fp <- sum(df$result == "FP")
-  fn <- sum(df$result == "FN")
-
-  p <- tp / (tp + fp)
-  r <- tp / (tp + fn)
-  f <- 2 * p * r / (p + r)
-  return(c(p, r, f))
-}
-
-render_table <- function(df, detector) {
-  hd <- df %>%
-    filter(sample_type == "high_density")
-
-  prf_vec <- get_prf(hd)
-
-  xtable::xtable(tibble(P = prf_vec[1], R = prf_vec[2], F = prf_vec[3]), caption=paste(detector, "High Density Sample")) %>%
-    print(type="latex", comment = F, table.placement = "H")
-
-  r <- df %>%
-    filter(sample_type == "random")
-
-  prf_vec <- get_prf(r)
-
-  xtable::xtable(tibble(P = prf_vec[1], R = prf_vec[2], F = prf_vec[3]), caption=paste(detector, "Random Sample")) %>%
-    print(type="latex", comment = F, table.placement = "H")
-}
-
-# render_table(mtcnn_results, "MTCNN")
-# render_table(openpose_results, "OpenPose")
-
+  mutate(result = evaluate(is_face, detector_is_face)) %>%
+  group_by(detectorType,sample_type) %>%
+  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
+  group_by(detectorType,sample_type) %>%
+  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
 ```
 
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrr}
-  \hline
-  & P & R & F \\ 
-\hline
-OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\ 
-OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\
-MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\ 
-MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\ 
-   \hline
-\end{tabular}
-\caption{Face detector performance} 
-\end{table}
+```{r xtable, results="asis"}
+ tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance")
+```
+
+<!-- \begin{table}[H] -->
+<!-- \centering -->
+<!-- \begin{tabular}{rrrr} -->
+<!--   \hline -->
+<!--   & P & R & F \\  -->
+<!-- \hline -->
+<!-- OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\  -->
+<!-- OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\ -->
+<!-- MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\  -->
+<!-- MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\  -->
+<!--    \hline -->
+<!-- \end{tabular} -->
+<!-- \caption{Face detector performance}  -->
+<!-- \end{table} -->
 
@@
 Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces.
-We next examined how the child's orientation relative to their caregiver impacted their visual access to faces.
-Overall, we saw that children who were far away from their caregiver were more likely to see faces than children who were close to their caregiver; this was true within all age groups and for all detectors.
+We also examined how the child's orientation relative to their caregiver impacted their visual access to faces: children who were far away from their caregiver were more likely to see faces than children who were close to their caregiver; this was true within all age groups and for both face  detectors.
 
@@
 
-## Access to Faces During Labeling Events
-Finally, we analyzed how face detections changed during object labeling events as a function of infant's posture and orientation.
-Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
-We again found that infant's posture impacted the degree to which they saw their caregiver's face during a labeling event; infants who were sitting or standing were more likely to have access to their caregiver's face.
-
 ```{r echo=FALSE, include=FALSE }
@@
 ```
-
 ```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections as a function of infants's posture"  }
@@
 ```
+## Access to Faces During Labeling Events
+Finally, we analyzed how face detections changed during object labeling events as a function of infant's posture and orientation.
+Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
+We again found that infant's posture impacted the degree to which they saw their caregiver's face during a labeling event; infants who were sitting or standing were more likely to have access to their caregiver's face.
 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
+ viola

--- 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
 Measuring the availability of caregiver’s faces across development was an important goal of this study.
-We used a total of two face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame.
-We capitalized on recent improvements in computer vision algorithms, testing the performance of two state-of-the-art face detectors that used novel methods to extract face information.
+We used a total of three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame.
+We first measured the performance of the most commonly-used and widely available face detection algorithms (Viola-Jones).
+We used this as a benchmark for performance, and while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces (@scheirer2014perceptual).
+We next capitalized on recent improvements in computer vision, testing the performance of two state-of-the-art face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information.
 
@@
 ### Face Detection Algorithms
-The first algorithm was based on the work by @zhang2016 using multi-task cascaded convolutional neural networks (MTCNNs).
+The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, @viola2004robust) applied to each individual frame.
+This detector provided information about the presence of a face as well as its size and position.
+
+The second algorithm was based on the work by @zhang2016 using multi-task cascaded convolutional neural networks (MTCNNs).
 The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
 We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet).
-This detector provided information about the presence of a face as well as its size and position.
+Like Viola-Jones, this detector provided information about the presence of a face as well as its size and position.
 
-The second algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
@@
 Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated and are reported in Table 2.
-Both detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
+Both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
 Due to the high performance of these two detectors we have included the results of both in the analyses to follow.
+The ViolaJones detector did not perform as well on either gold set, so the results associated with this detector were not included.
 
@@
   rename(video = name)
+
+data_viola <- read_csv("../data/final_output/viola.csv") %>%
+  distinct(video, frame, .keep_all = TRUE)
 
@@
 }
+
+viola_results <- data_viola %>%
+  rename(detector_is_face = is_face) %>%
+  inner_join(ground_truth_data, by = c("group", "video", "frame")) %>%
+  rowwise() %>%
+  mutate(result = evaluate(is_face, detector_is_face))
 
@@
 
+# render_table(viola_results, "Viola")
 # render_table(mtcnn_results, "MTCNN")
@@
 \hline
+ViolaJones High Density Sample & 0.99 & 0.50 & 0.66 \\
+ViolaJones Random Sample & 0.39 & 0.29 & 0.33 \\
 OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\ 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
+ viola

@@
 However, as children are learning their first words, their view of the
-world is also changing radically (K. Adolph \& Berger, 2007). Infant's
+world is also changing radically (K. Adolph \& Berger, 2007). Infants'
 motor abilities improve dramatically near the end of the first year of
@@
 document the visual experiences of infants and children --- which even
-for walking children are radically different than the adult perspective
+for walking children are radically different from the adult perspective
 (and not easily predicted by our own adult intuitions (Clerkin, Hart,
@@
 Measuring the availability of caregiver's faces across development was
-an important goal of this study. We used a total of two face detection
+an important goal of this study. We used a total of three face detection
 systems to explore infants' changing access to social information and to
-avoid the cost of hand-annotating every frame. We capitalized on recent
-improvements in computer vision algorithms, testing the performance of
-two state-of-the-art face detectors that used novel methods to extract
-face information.
+avoid the cost of hand-annotating every frame. We first measured the
+performance of the most commonly-used and widely available face
+detection algorithms (Viola-Jones). We used this as a benchmark for
+performance, and while it can achieve impressive accuracy in some
+situations, it is notoriously bad at dealing with occluded faces
+(Scheirer, Anthony, Nakayama, \& Cox (2014)). We next capitalized on
+recent improvements in computer vision, testing the performance of two
+state-of-the-art face detectors that both made use of Convolutional
+Neural Networks (CNNs) to extract face information.
 
@@
 
-The first algorithm was based on the work by K. Zhang et al. (2016)
+The first face detection system made use of a series of Haar
+feature-based cascade classifiers (Viola-Jones, Viola \& Jones (2004))
+applied to each individual frame. This detector provided information
+about the presence of a face as well as its size and position.
+
+The second algorithm was based on the work by K. Zhang et al. (2016)
 using multi-task cascaded convolutional neural networks (MTCNNs). The
@@
 used a Tensorflow implementation of this algorithm provided by
-(\url{https://github.com/davidsandberg/facenet}). This detector provided
-information about the presence of a face as well as its size and
-position.
+(\url{https://github.com/davidsandberg/facenet}). Like Viola-Jones, this
+detector provided information about the presence of a face as well as
+its size and position.
 
-The second algorithm was a CNN-based pose detector that provided the
+The third algorithm was a CNN-based pose detector that provided the
 locations of 18 body parts (ears, nose, wrists, etc.) (Cao et al.
@@
 and F-score (harmonic mean of previous measures) were calculated and are
-reported in Table 2. Both detectors performed relatively well on the
-gold set, with MTCNN outperforming OpenPose on the random sample but
-trailing behind in the high density sample. Due to the high performance
-of these two detectors we've included the results of both in the
-analyses to follow.
+reported in Table 2. Both OpenPose and MTCNN detectors performed
+relatively well on the gold set, with MTCNN outperforming OpenPose on
+the random sample but trailing behind in the high density sample. Due to
+the high performance of these two detectors we have included the results
+of both in the analyses to follow. The ViolaJones detector did not
+perform as well on either gold set, so the results associated with this
+detector were not included.
 
@@
 First, we report developmental shifts in infants' posture and their
-orientation to their caregiver. Then, we explore how these changes
-influence children's visual access to faces across this developmental
-time range. Finally, we explore how these changes impact the
-accessibility of faces during labeling events.
+orientation relative to their caregiver. Then, we explore how these
+changes influence children's visual access to faces across this
+developmental time range. Finally, we explore how these changes impact
+the accessibility of faces during labeling events.
 
@@
 
-\caption[Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)]{Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel).}\label{fig:orientation}
+\caption[Proportion time that infants in each age group spent in each orientation relative to their caregiver]{Proportion time that infants in each age group spent in each orientation relative to their caregiver.}\label{fig:orientation}
 \end{figure}
@@
 
-\caption[Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel)]{Proportion time that infants in each age group spent in each posture (left panel) and orientation relative to their caregiver (right  panel).}\label{fig:posture}
+\caption[Proportion time that infants in each age group spent in each posture]{Proportion time that infants in each age group spent in each posture.}\label{fig:posture}
 \end{figure}
@@
 caregivers' orientation, it held relatively less predictive power.
-Overall, these results confirm that infant's access to faces is heavily
-infleunced by their own posture and their caregivers orientation towards
-them.
+Overall, these results confirm that infants' access to faces is heavily
+infleunced by their own posture and their caregiver's orientation
+towards them.
 
@@
 We again found that infant's posture impacted the degree to which they
-saw their caregivers face during a labeling event; infants who were
+saw their caregiver's face during a labeling event; infants who were
 sitting or standing were more likely to have access to their caregiver's
@@
 
+\hypertarget{ref-scheirer2014perceptual}{}
+Scheirer, W. J., Anthony, S. E., Nakayama, K., \& Cox, D. D. (2014).
+Perceptual annotation: Measuring human vision to improve computer
+vision. \emph{IEEE Transactions on Pattern Analysis and Machine
+Intelligence}, \emph{36}(8), 1679--1686.
+
 \hypertarget{ref-simon2017hand}{}

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
cleanup and get rid of VJ for plotting

--- 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
 ```{r, message=FALSE, warning=FALSE, results='asis'}
-### Needs to be consolidated so that we're aren't rereading all of these files here...
-# data_vj<- read_csv("../data/final_output/viola.csv") %>%
-#   distinct(video, frame, .keep_all = TRUE)
-# 
-# data_mtcnn <- read_csv("../data/final_output/mtcnn3.csv") %>%
-#   distinct(video, frame, .keep_all = TRUE)
-# 
-# data_openpose <- read_csv("../data/final_output/openpose_results_truncated_2.csv")%>%
-#   distinct(name, frame, .keep_all = TRUE) %>%
-#   mutate(frame = sprintf("%05d", as.numeric(frame) + 1)) %>%
-#   mutate(is_face = ifelse(Nose_conf != 0, "True", "False")) %>%
-#   select(group, name, frame, is_face) %>%
-#   rename(video = name)
-
 d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
@@
 ## couple of helper functions
-
 evaluate <- function(a, b) {
@@
 
+## evaluate detectors
 performance <- ground_truth_data  %>%
@@
 ```{r xtable, results="asis"}
- tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance")
+ tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance", type="latex", comment = F, table.placement = "H")
 ```
 
-<!-- \begin{table}[H] -->
-<!-- \centering -->
-<!-- \begin{tabular}{rrrr} -->
-<!--   \hline -->
-<!--   & P & R & F \\  -->
-<!-- \hline -->
-<!-- OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\  -->
-<!-- OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\ -->
-<!-- MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\  -->
-<!-- MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\  -->
-<!--    \hline -->
-<!-- \end{tabular} -->
-<!-- \caption{Face detector performance}  -->
-<!-- \end{table} -->
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllrrr}
+  \hline
+ & detectorType & sample\_type & p & r & f \\ 
+  \hline
+1 & MTCNN & high\_density & 0.841 & 1.000 & 0.914 \\ 
+  2 & MTCNN & random & 0.947 & 0.750 & 0.837 \\ 
+  3 & OpenPose & high\_density & 0.995 & 0.863 & 0.924 \\ 
+  4 & OpenPose & random & 0.710 & 0.917 & 0.800 \\ 
+  5 & ViolaJones & high\_density & 0.991 & 0.495 & 0.660 \\ 
+  6 & ViolaJones & random & 0.389 & 0.292 & 0.333 \\ 
+   \hline
+\end{tabular}
+\caption{Detector performance} 
+\end{table}
 
@@
 ```{r echo=FALSE, include=FALSE}
-d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
-  mutate(posture = factor(posture), orientation = factor(orientation))
-
-## make long form version for plotting both face detections on same graph
-dlong <- d %>%
-  mutate(id = paste(subid,frame)) %>%
-  gather(key = detectorType, value = detection, faceMT, faceOP) %>%
-  mutate(detectorType=as.factor(detectorType))
+# get rid of violajones
+dlong <- dlong %>%
+  filter(detectorType!="faceVJ")
 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
cleanup & couple renames

@@
 
-## couple of helper functions
+## helper function
 evaluate <- function(a, b) {
@@
 
-## okay, big long tidyr chunk
+## read in ground truth detectors
 ground_truth_data <- read_csv("../data/ground_truth/ground_truth3.csv") %>%
@@
 
-```{r xtable, results="asis"}
- tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance", type="latex", comment = F, table.placement = "H")
-```
+# ```{r xtable, results="asis"}
+#  tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance", type="latex", comment = F, table.placement = "H")
+# ```
 
@@
 #[1] "faceMT"  "faceOP"  "handOP"  "wristOP"
-  levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces") 
+  levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces") 
 ```

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd

--- 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
 Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated and are reported in Table 2.
-Both detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
+Both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
 Due to the high performance of these two detectors we have included the results of both in the analyses to follow.
+The ViolaJones detector did not perform as well on either gold set, so the results associated with this detector were not included.
 
 ```{r, message=FALSE, warning=FALSE, results='asis'}
+### Needs to be consolidated so that we're aren't rereading all of these files here...
+# data_vj<- read_csv("../data/final_output/viola.csv") %>%
+#   distinct(video, frame, .keep_all = TRUE)
+# 
+# data_mtcnn <- read_csv("../data/final_output/mtcnn3.csv") %>%
+#   distinct(video, frame, .keep_all = TRUE)
+# 
+# data_openpose <- read_csv("../data/final_output/openpose_results_truncated_2.csv")%>%
+#   distinct(name, frame, .keep_all = TRUE) %>%
+#   mutate(frame = sprintf("%05d", as.numeric(frame) + 1)) %>%
+#   mutate(is_face = ifelse(Nose_conf != 0, "True", "False")) %>%
+#   select(group, name, frame, is_face) %>%
+#   rename(video = name)
+
 d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
@@
 ## couple of helper functions
+
 evaluate <- function(a, b) {
@@
 
-## evaluate detectors
 performance <- ground_truth_data  %>%
@@
 ```{r xtable, results="asis"}
- tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance", type="latex", comment = F, table.placement = "H")
+ tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance")
 ```
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{rllrrr}
-  \hline
- & detectorType & sample\_type & p & r & f \\ 
-  \hline
-1 & MTCNN & high\_density & 0.841 & 1.000 & 0.914 \\ 
-  2 & MTCNN & random & 0.947 & 0.750 & 0.837 \\ 
-  3 & OpenPose & high\_density & 0.995 & 0.863 & 0.924 \\ 
-  4 & OpenPose & random & 0.710 & 0.917 & 0.800 \\ 
-  5 & ViolaJones & high\_density & 0.991 & 0.495 & 0.660 \\ 
-  6 & ViolaJones & random & 0.389 & 0.292 & 0.333 \\ 
-   \hline
-\end{tabular}
-\caption{Detector performance} 
-\end{table}
+<!-- \begin{table}[H] -->
+<!-- \centering -->
+<!-- \begin{tabular}{rrrr} -->
+<!--   \hline -->
+<!--   & P & R & F \\  -->
+<!-- \hline -->
+<!-- ViolaJones High Density Sample & 0.99 & 0.50 & 0.66 \\ -->
+<!-- ViolaJones Random Sample & 0.39 & 0.29 & 0.33 \\ -->
+<!-- OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\  -->
+<!-- OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\ -->
+<!-- MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\  -->
+<!-- MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\  -->
+<!--    \hline -->
+<!-- \end{tabular} -->
+<!-- \caption{Face detector performance}  -->
+<!-- \end{table} -->
 
@@
 ```{r echo=FALSE, include=FALSE}
-# get rid of violajones
-dlong <- dlong %>%
-  filter(detectorType!="faceVJ")
+d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
+  mutate(posture = factor(posture), orientation = factor(orientation))
+
+## make long form version for plotting both face detections on same graph
+dlong <- d %>%
+  mutate(id = paste(subid,frame)) %>%
+  gather(key = detectorType, value = detection, faceMT, faceOP) %>%
+  mutate(detectorType=as.factor(detectorType))
 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.tex

@@
 Here, we explore this question by using head-mounted cameras to record
-infants' (8-16 months of age) egocentric visual perspective and modern
-computer vision algorithms to detect the proportion of faces in infants'
-environments. We find that infants' postural changes largely mediate
-infants' access to faces, suggesting that these postural and locomotive
-developments facilitate infants' emerging linguistic and social
-capacities. Broadly, we suggest that the combined use of head-mounted
-cameras and the application of novel deep learning algorithms is a
-promising avenue for understanding the statistics of infants' visual and
-linguistic experience.
+infants' (8-16 months of age) egocentric visual perspective and use
+state-of-the-art computer vision algorithms to detect the proportion of
+faces in infants' environments. We find that infants' postural changes
+largely mediate infants' access to faces, hinting that these postural
+and locomotive developments facilitate infants' emerging linguistic and
+social capacities. Broadly, we suggest that the combined use of
+head-mounted cameras and the application of novel deep learning
+algorithms is a promising avenue for understanding the statistics of
+infants' visual and linguistic experience.
 
@@
 However, as children are learning their first words, their view of the
-world is also changing radically (K. Adolph \& Berger, 2007). Infants'
+world is also radically changing (K. Adolph \& Berger, 2007). Infants'
 motor abilities improve dramatically near the end of the first year of
-life, allowing them to locomote independently and to determine what they
-see. These motor changes have drastic consequences for what children
-see: crawling and walking infants have radically different views of the
-world. During spontaneous play in a laboratory playroom, toddlers are
-more likely to look at the floor while crawling than while walking (J.
-Franchak, Kretch, Soska, \& Adolph, 2011); walking vs.~crawling infants
-tend to have full visual access to their environment and the people in
-it (K. S. Kretch, Franchak, \& Adolph, 2014).
+life, allowing them to locomote independently. These motor changes have
+drastic consequences for what children see; crawling and walking infants
+have radically different views of the world. For example, during
+spontaneous play in a laboratory playroom, toddlers are more likely to
+look at the floor while crawling than while walking (J. Franchak,
+Kretch, Soska, \& Adolph, 2011); in general, walking infants tend to
+have full visual access to their environment and the people in it, while
+crawling infants do not (K. S. Kretch, Franchak, \& Adolph, 2014).
 
@@
 and linguistic abilities (Iverson, 2010). Indeed, these postural changes
-also impact how children interact with their mothers; walking infants
-make different kinds of object-related bids for attention from their
-mothers than crawling infants, and tend to hear more action directed
+also impact how children interact with their mothers; walking
+(vs.~crawling) infants make different kinds of object-related bids for
+attention from their mothers and tend to hear more action directed
 statements (e.g., ``open it'') (Karasik, Tamis-LeMonda, \& Adolph,
-2014). More directly, in an observational study, Walle \& Campos (2014)
-found that children who were able to walk had both higher receptive and
+2014). Further, in an observational study, Walle \& Campos (2014) found
+that children who were able to walk had both higher receptive and
 productive vocabularies. On their account, children's ability to stand
 and independently locomote may fundamentally change their ability to
-access the social information (e.g., faces, gaze) relative to children
-who are still crawling and sitting. In other words, the ability to
-access more detailed social information may allow infants to learn words
+access social information (e.g., faces, gaze) relative to children who
+are still crawling and sitting. In other words, the ability to access
+more detailed social information may allow infants to learn words
 quicker and more efficiently, facilitating language growth.
@@
 for walking children are radically different from the adult perspective
-(and not easily predicted by our own adult intuitions (Clerkin, Hart,
-Rehg, Yu, \& Smith, 2017; J. Franchak et al., 2011, Yoshida \& Smith
-(2008))). Children's views tend to be restricted and to be dominated by
+(and not easily predicted by our own adult intuitions) (Clerkin, Hart,
+Rehg, Yu, \& Smith, 2017; J. Franchak et al., 2011; Yoshida \& Smith,
+2008). Children's views tend to be restricted and to be dominated by
 objects and hands much more than that of adults (Yoshida \& Smith,
@@
 
-Here, we directly ask whether the postural changes that infants
-experience as they reach their first birthday change the availability of
-the social information relevant for word learning. To do so, we recorded
-the visual experience of a group of infants and children using
-head-mounted cameras during a brief laboratory free-play session;
-children's posture and orientation relative to their caregiver were also
-recorded from a third-person perspective and hand-annotated. We
-capitalize on recent improvements in face detection algorithms (K.
-Zhang, Zhang, Li, \& Qiao, 2016, Cao, Simon, Wei, \& Sheikh (2017)) to
-analyze the frequences of faces in the child's visual environment.
+Here, we directly ask whether the postural and locomotive developments
+change the availability of the social information relevant for word
+learning. To do so, we recorded the visual experience of a group of
+infants and children using head-mounted cameras during a brief
+laboratory free-play session; children's posture and orientation
+relative to their caregiver were also recorded from a third-person
+perspective and hand-annotated. We capitalize on recent improvements in
+face detection algorithms (Cao, Simon, Wei, \& Sheikh, 2017; K. Zhang,
+Zhang, Li, \& Qiao, 2016) to analyze the frequencies of faces in the
+child's visual environment, both overall and relative to naming events
+by their caregivers.
 
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrr}
-  \hline
-  & P & R & F \\ 
-\hline
-ViolaJones High Density Sample & 0.99 & 0.50 & 0.66 \\
-ViolaJones Random Sample & 0.39 & 0.29 & 0.33 \\
-OpenPose High Density Sample & 0.99 & 0.87 & 0.93 \\ 
-OpenPose Random Sample & 0.71 & 0.92 & 0.80 \\
-MTCNN High Density Sample & 0.84 & 1.00 & 0.91 \\ 
-MTCNN Random Sample & 0.95 & 0.75 & 0.84 \\ 
-   \hline
-\end{tabular}
-\caption{Face detector performance} 
-\end{table}
-
 \section{Results}\label{results}
@@
 sitting saw more faces than infants who were lying down or being
-carried, while infants who were standing saw the most faces. We next
+carried, while infants who were standing saw the most faces. We also
 examined how the child's orientation relative to their caregiver
-impacted their visual access to faces. Overall, we saw that children who
-were far away from their caregiver were more likely to see faces than
-children who were close to their caregiver; this was true within all age
-groups and for all detectors.
+impacted their visual access to faces: children who were far away from
+their caregiver were more likely to see faces than children who were
+close to their caregiver; this was true within all age groups and for
+both face detectors.
 
@@
 
+\begin{CodeChunk}
+\begin{figure}[H]
+
+{\centering \includegraphics{figs/detByNaming-1}
+
+}
+
+\caption[Proportion face detections as a function of infants's posture]{Proportion face detections as a function of infants's posture}\label{fig:detByNaming}
+\end{figure}
+\end{CodeChunk}
+
 \subsection{Access to Faces During Labeling
@@
 face.
-
-\begin{CodeChunk}
-\begin{figure}[H]
-
-{\centering \includegraphics{figs/detByNaming-1}
-
-}
-
-\caption[Proportion face detections as a function of infants's posture]{Proportion face detections as a function of infants's posture}\label{fig:detByNaming}
-\end{figure}
-\end{CodeChunk}
 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
comment out table code

--- 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
   summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
-```
 
@@
 # ```
+
+```
 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd

@@
 
-## helper function
 evaluate <- function(a, b) {
@@
 
-## read in ground truth detectors
 ground_truth_data <- read_csv("../data/ground_truth/ground_truth3.csv") %>%
@@
   summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
-
-# ```{r xtable, results="asis"}
-#  tab2 <- xtable::xtable(performance, digits=3, caption = "Detector performance", type="latex", comment = F, table.placement = "H")
-# ```
 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.tex

@@
 
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllrrr}
+  \hline
+ & detectorType & sample\_type & p & r & f \\ 
+  \hline
+1 & MTCNN & high\_density & 0.841 & 1.000 & 0.914 \\ 
+  2 & MTCNN & random & 0.947 & 0.750 & 0.837 \\ 
+  3 & OpenPose & high\_density & 0.995 & 0.863 & 0.924 \\ 
+  4 & OpenPose & random & 0.710 & 0.917 & 0.800 \\ 
+  5 & ViolaJones & high\_density & 0.991 & 0.495 & 0.660 \\ 
+  6 & ViolaJones & random & 0.389 & 0.292 & 0.333 \\ 
+   \hline
+\end{tabular}
+\caption{Detector performance} 
+\end{table}
+
 \section{Results}\label{results}

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
resolve conflicts

--- 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
  & detectorType & sample\_type & p & r & f \\ 
+ & Algorithm & Sample\ Type & P & R & F \\ 
   \hline
@@
 ```
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections as a function of infants's posture"  }
+
 ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
@@
   ylab("Proportion of face detections within naming window") + 
+  ylab("Prop. face detections in naming window") + 
   xlab("Posture") + 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
wording
resolve conflicts

@@
   \hline
- & detectorType & sample\_type & p & r & f \\ 
  & Algorithm & Sample\ Type & P & R & F \\ 
@@
 
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming window as a function of infants's posture"  }
 ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
   geom_boxplot() + 
-  ylab("Proportion of face detections within naming window") + 
   ylab("Prop. face detections in naming window") + 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd

--- 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
 Due to the high performance of these two detectors we have included the results of both in the analyses to follow.
-The ViolaJones detector did not perform as well on either gold set, so the results associated with this detector were not included.
+The ViolaJones detector did not perform well on either gold set, so the results from this detector were not included.
 
@@
   \hline
- & detectorType & sample\_type & p & r & f \\ 
+ & Algorithm & Sample\ Type & P & R & F \\ 
   \hline
@@
 ```
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections as a function of infants's posture"  }
+
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming window as a function of infants's posture"  }
 ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
   geom_boxplot() + 
-  ylab("Proportion of face detections within naming window") + 
+  ylab("Prop. face detections in naming window") + 
   xlab("Posture") + 

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
got rid of a few extraneous things

@@
 ## Face Detection
-Measuring the availability of caregiver’s faces across development was an important goal of this study.
 We used a total of three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame.
@@
 
-## Model
-
 <!-- ```{r echo=FALSE, include=FALSE} -->

### 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
new paragraph

--- 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
 Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated and are reported in Table 2.
+
 Both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
@@
   \hline
- & Algorithm & Sample\ Type & P & R & F \\ 
+ & detectorType & sample\_type & p & r & f \\ 
   \hline
@@
 ```
-
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming window as a function of infants's posture"  }
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections as a function of infants's posture"  }
 ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
   geom_boxplot() + 
-  ylab("Prop. face detections in naming window") + 
+  ylab("Proportion of face detections within naming window") + 
   xlab("Posture") + 

### 2018-01-17 scripts/openpose/csv_header.py
json csv updates

@@ -4,7 +4,7 @@ import csv
 TOTAL_POSE_POINTS = 18
 TOTAL_FACE_POINTS = 70
 TOTAL_HAND_POINTS = 21
-CSV_OUTPUT_DIR = os.path.expandvars("$HOME/xs-face/scripts/openpose/tmp2")
+CSV_OUTPUT_DIR = os.path.expandvars("$HOME/xs-face/scripts/openpose/tmp")
 # https://github.com/CMU-Perceptual-Computing-Lab/openpose/blob/master/src/openpose/pose/poseParameters.cpp
 
 def chunker(seq, size):

### 2018-01-17 scripts/openpose/json_to_csv.py

@@ -6,9 +6,6 @@ import ntpath
 
 CSV_OUTPUT_DIR = os.path.expandvars("$HOME/xs-face/scripts/openpose/tmp")
 
-# TODO 1229 is weird because title contains "cropped"
-# TODO if no face or hand or whatever, fill in with 0s!!! first check if, and if not, extend with 0s
-
 if __name__ == "__main__":
     video_full_path = sys.argv[1]
     video_filename = ntpath.basename(video_full_path)
@@ -21,7 +18,10 @@ if __name__ == "__main__":
         wr = csv.writer(f, quoting=csv.QUOTE_ALL)
 
         for filename in os.listdir(video_full_path):
-            frame = filename.split("_")[3][7:]
+            if "XS_1229" in filename:
+                frame = filename.split("_")[4][7:]
+            else:
+                frame = filename.split("_")[3][7:]
 
             with open(os.path.join(video_full_path, filename), "rb") as json_file:
                 data = json.load(json_file)
@@ -29,11 +29,28 @@ if __name__ == "__main__":
 
             if people:
                 for person in people:
-                    row = [group, subject, frame] + \
-                           person["pose_keypoints"] + \
-                           person["face_keypoints"] + \
-                           person["hand_left_keypoints"] + \
-                           person["hand_right_keypoints"]
+                    row = [group, subject, frame]
+
+                    if person["pose_keypoints"]:
+                        row +=  person["pose_keypoints"]
+                    else:
+                        row += [0 for _ in xrange(18*3)]
+
+                    if person["face_keypoints"]:
+                        row += person["face_keypoints"]
+                    else:
+                        row += [0 for _ in xrange(70*3)]
+
+                    if person["hand_left_keypoints"]:
+                        row += person["hand_left_keypoints"]
+                    else:
+                        row += [0 for _ in xrange(21*3)]
+
+                    if person["hand_right_keypoints"]:
+                        row += person["hand_right_keypoints"]
+                    else:
+                        row += [0 for _ in xrange(21*3)]
+
                     wr.writerow(row)
             else:
                 wr.writerow([group, subject, frame] + [0 for _ in xrange(390)])

### 2018-01-17 scripts/openpose/wrap_json_to_csv.sh

@@ -1,6 +1,6 @@
 #!/bin/sh
 
-for vid in $PI_HOME/openpose_output4/*; do
+for vid in $PI_HOME/openpose_output6/*; do
 	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
 	    echo "Large video (> 10GB): $vid"
 	    time=3:00:00

### 2018-01-17 writeup-old/xsface.tex

--- 2018-01-10 writeup/xsface.Rmd
@@
----
-title: "Developmental and postural changes in children's visual access to faces"
-short-title: "Children's visual access to faces"
-output: kmr::apa_manuscript
-csl: apa6.csl
-bibliography: xsface.bib
+% Template for APA submission with R Markdown
 
-document-params: "a4paper,man,apacite,floatsintext,longtable"
+% Stuff changed from PLOS Template
+\documentclass[a4paper,man,apacite,floatsintext,longtable]{apa6}
+\usepackage{apacite}
 
-bib-tex: "xsface.bib"
+% amsmath package, useful for mathematical formulas
+\usepackage{amsmath}
+% amssymb package, useful for mathematical symbols
+\usepackage{amssymb}
 
-author-information:
-    - \author{Michael C. Frank, Guido Pusiol, Daniel Yurovsky, Kaia Simmons}
+% hyperref package, useful for hyperlinks
+\usepackage{hyperref}
 
-affiliation-information:
-    - \affiliation{Department of Psychology, Stanford University}
+% graphicx package, useful for including eps and pdf graphics
+% include graphics with the command \includegraphics
+\usepackage{graphicx}
 
-author-note:
-    "Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of the Language and Cognition Lab for help in recruitment, data collection, and annotation.
-This research was supported by a John Merck Scholars grant to MCF.
-An earlier version of this work was presented to the Cognitive Science Society in @frank2013.
-Please address correspondence to Michael C.
-Frank, Department of Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}."
+% Sweave(-like)
+\usepackage{fancyvrb}
+\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
+\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
+\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
+\newenvironment{Schunk}{}{}
+\DefineVerbatimEnvironment{Code}{Verbatim}{}
+\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
+\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
+\newenvironment{CodeChunk}{}{}
 
-abstract: 
-    "The faces of other people are a critical information source for young children.
-During early development, children undergo significant postural and locomotor development, changing from lying and sitting infants to toddlers who walk independently.
-We used a head-mounted camera in conjunction with a face-detection system to explore the eff  ects of these changes on children's visual access to their caregivers' faces during an in-lab play session.
-In a cross-sectional sample of 8--16 month old children, we found substantial changes in face accessibility based on age and posture.
-These changes may translate into changes in the accessibility of social information during language learning.
-We make our corpus available for reanalysis, and discuss strengths and limitations of the headcam method more generally."
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
 
-keywords:
-    "social cognition; face-perception; infancy; locomotion; head-cameras"
----
+\usepackage{color}
 
-```{r global_options, include=FALSE}
-rm(list=ls())
-library(knitr)
-knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
-                      fig.path='figs/', echo=FALSE, warning = FALSE, 
-                      cache=TRUE, message = FALSE, sanitize = TRUE)
-rm(list=ls())
-library(readr)
-# library(magrittr)
-library(dplyr)
-library(tidyr)
-library(xtable)
-# library(lubridate)
-```
+% Use doublespacing - comment out for single spacing
+%\usepackage{setspace}
+%\doublespacing
 
-<!-- % A father offers his young daughter a novel object: a bright yellow feather duster.
-A few moments after she accepts the toy, he remarks, ``Isn't the zem funny?'' Her father may still be talking about the feather duster, or he may be describing a new object.
-To find out she has access to a simple and reliable method: she can look to his face to infer the direction of his attention.
--->
+% Text layout
+\topmargin 0.0cm
+\oddsidemargin 0.5cm
+\evensidemargin 0.5cm
+\textwidth 16cm
+\textheight 21cm
 
-The ability to follow social signals like eye-gaze is an important part of early social cognition [@scaife1975;@tomasello2009]
+% Bold the 'Figure #' in the caption and separate it with a period
+% Captions will be left justified
+\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
 
-In addition to being an important part of early social development, the ability to follow gaze is a strong predictor of children's early language development.
-For example, @brooks2005 found that children who followed an experimenter's gaze better before their first birthday had larger vocabularies at 18 months.
-Similarly, @carpenter1998 found that children's level of joint engagement (as well as the degree to which mothers followed the child's focus of attention in their labeling) predicted vocabulary growth in both language production and comprehension.
-These studies suggest that children's social environment plays a powerful supportive role in language learning.
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
 
-But at the same time as children are beginning to learn their first words, their view of the world is changing radically [@adolph2007].
-As speechless infants, they are unable to locomote independently.
-Before their first birthday, they begin crawling; soon after, they begin to walk independently.
-Infants' visual field is subject to the whims of their caregivers, but caregivers often place them in positions conducive to joint attention.
-In contrast, toddlers determine their own input to a much greater degree, but as a consequence they spend much of their time in a world primarily populated by knees.
-These postural and locomotor changes may have a profound effect on what children see.
+% Leave date blank
+\date{}
 
-A recent study suggests the possibility of links between motor milestones, social cognition, and language.
-@walle2014 noted robust correlations between children's ability to walk and their vocabulary, both receptive and productive.
-On the basis of an observational study of parent input, they speculated that the emergence of walking may change the ability of the child to access social information (because walking toddlers see more of the social world than crawling infants).
-Accessing more social information may in turn allow children to discover word meanings more effectively.
+%\pagestyle{myheadings}
+%% ** EDIT HERE **
 
-Recent methodological developments have the potential to provide data that would allow this hypothesis to be tested.
-The availability of head-mounted cameras and eye-trackers allows for the measurement of children's naturalistic environment in a way that was not previously possible.
-@yoshida2008 gave the first demonstration of the radical differences between toddler and adult perspectives on the social world, with toddlers' visual field being dominated by hands and objects much more than that of adults.
-More recent work has used head-mounted eye-tracking methods to measure young toddlers' fixations [@franchak2011], also finding that children look relatively infrequently at their mothers' faces in naturalistic play.
+%% ** EDIT HERE **
+%% PLEASE INCLUDE ALL MACROS BELOW
 
-These methods are now being applied to understand inputs to language acquisition.
-Work by Yu, Smith, and colleagues suggests that word learning is facilitated when parents and children create moments in which the visual field is dominated by a single object [@smith2011;@yuinpress].
-Some data even suggest that young children's restricted viewpoint may be more effective for learning words than the comparable adult perspective [@yurovsky2012].
-Together, this body of evidence suggests that measuring infants' perspective---and how it changes in motor development---is a critical part of understanding early language learning.
+%% END MACROS SECTION
 
-In the current study we took a developmental approach to understanding the relationship between perspective and access to social information.
-We recorded head-camera data from a group of infants and children across a broad age range as they played with their caregivers during a brief laboratory visit.
-We then hand-annotated these data for the child's posture and parents' naming behavior and used face-detection algorithms to measure the frequency of faces in the child's visual field.
-The resulting dataset allows us to analyze changes in access to faces according to children's age, posture, and linguistic input.
+% ALL OF THE TITLE PAGE INFORMATION IS SPECIFIED IN THE YAML
+\title{\textbf{Developmental and postural changes in children's visual access to faces}}
+\shorttitle{Children's visual access to faces}
 
-# Methods
+\author{Michael C. Frank, Guido Pusiol, Daniel Yurovsky, Kaia Simmons}
 
-## Participants
+\affiliation{Department of Psychology, Stanford University}
 
-```{r echo=FALSE, include=FALSE}
-all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
-  rename(subid = subID)
-inc_subs <- read.csv("../data/demographics/demographics.csv")
-all_subs$inc <- all_subs$reason=="included"
+\authornote{Thanks to Ally Kraus, Kathy Woo, Aditi Maliwal, and other members of the
+Language and Cognition Lab for help in recruitment, data collection, and
+annotation. This research was supported by a John Merck Scholars grant
+to MCF. An earlier version of this work was presented to the Cognitive
+Science Society in Michael C Frank, Simmons, Yurovsky, \& Pusiol (2013).
+Please address correspondence to Michael C. Frank, Department of
+Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford,
+CA, 94305, tel: (650) 724-4003, email:
+\href{mailto:exttt\%7Bmcfrank@stanford.edu\%7D}{\nolinkurl{exttt\{mcfrank@stanford.edu\}}}.}
+\abstract{The faces of other people are a critical information source for young
+children. During early development, children undergo significant
+postural and locomotor development, changing from lying and sitting
+infants to toddlers who walk independently. We used a head-mounted
+camera in conjunction with a face-detection system to explore the eff
+ects of these changes on children's visual access to their caregivers'
+faces during an in-lab play session. In a cross-sectional sample of
+8--16 month old children, we found substantial changes in face
+accessibility based on age and posture. These changes may translate into
+changes in the accessibility of social information during language
+learning. We make our corpus available for reanalysis, and discuss
+strengths and limitations of the headcam method more generally.}
+\keywords{social cognition; face-perception; infancy; locomotion; head-cameras}
 
-inc_subs$len.min <- sapply(strsplit(as.character(inc_subs$len),":"),
-       function(x) {
-         x <- as.numeric(x)
-         x[1]+x[2]/60
-       }
-)
+\begin{document}
+\maketitle
 
-all_subs %>% 
-  group_by(reason) %>%
-  summarise(n = n()) %>% 
-  kable
+The ability to follow social signals like eye-gaze is an important part
+of early social cognition (Scaife \& Bruner, 1975; M. Tomasello, 2009)
 
-# # exclusions by age
-# all_subs %>% 
-#   group_by(reason, age_group) %>%
-#   summarise(n = n()) %>%
-#   spread(age_group, n) %>%
-#   xtable %>%
-#   print(digits = 2)
+In addition to being an important part of early social development, the
+ability to follow gaze is a strong predictor of children's early
+language development. For example, Brooks \& Meltzoff (2005) found that
+children who followed an experimenter's gaze better before their first
+birthday had larger vocabularies at 18 months. Similarly, Carpenter,
+Nagell, \& Tomasello (1998) found that children's level of joint
+engagement (as well as the degree to which mothers followed the child's
+focus of attention in their labeling) predicted vocabulary growth in
+both language production and comprehension. These studies suggest that
+children's social environment plays a powerful supportive role in
+language learning.
 
-all_subs %>%
-  left_join(inc_subs) %>%
-  group_by(age_group) %>%
-  summarise(n = sum(inc), 
-            percent_included = mean(inc),
-            age = mean(age.at.test, na.rm=TRUE), 
-            video_length = mean(len.min, na.rm=TRUE), 
-            female = mean(gender=="female", na.rm=TRUE)) %>%
-  xtable %>%
-  print(digits = 2, include.rownames=FALSE)
-```
+But at the same time as children are beginning to learn their first
+words, their view of the world is changing radically (Adolph \& Berger,
+2007). As speechless infants, they are unable to locomote independently.
+Before their first birthday, they begin crawling; soon after, they begin
+to walk independently. Infants' visual field is subject to the whims of
+their caregivers, but caregivers often place them in positions conducive
+to joint attention. In contrast, toddlers determine their own input to a
+much greater degree, but as a consequence they spend much of their time
+in a world primarily populated by knees. These postural and locomotor
+changes may have a profound effect on what children see.
+
+A recent study suggests the possibility of links between motor
+milestones, social cognition, and language. Walle \& Campos (2014) noted
+robust correlations between children's ability to walk and their
+vocabulary, both receptive and productive. On the basis of an
+observational study of parent input, they speculated that the emergence
+of walking may change the ability of the child to access social
+information (because walking toddlers see more of the social world than
+crawling infants). Accessing more social information may in turn allow
+children to discover word meanings more effectively.
+
+Recent methodological developments have the potential to provide data
+that would allow this hypothesis to be tested. The availability of
+head-mounted cameras and eye-trackers allows for the measurement of
+children's naturalistic environment in a way that was not previously
+possible. Yoshida \& Smith (2008) gave the first demonstration of the
+radical differences between toddler and adult perspectives on the social
+world, with toddlers' visual field being dominated by hands and objects
+much more than that of adults. More recent work has used head-mounted
+eye-tracking methods to measure young toddlers' fixations (Franchak,
+Kretch, Soska, \& Adolph, 2011), also finding that children look
+relatively infrequently at their mothers' faces in naturalistic play.
+
+These methods are now being applied to understand inputs to language
+acquisition. Work by Yu, Smith, and colleagues suggests that word
+learning is facilitated when parents and children create moments in
+which the visual field is dominated by a single object (L. B. Smith, Yu,
+\& Pereira, 2011; Yu \& Smith, in press). Some data even suggest that
+young children's restricted viewpoint may be more effective for learning
+words than the comparable adult perspective (D. Yurovsky, Smith, \& Yu,
+in press). Together, this body of evidence suggests that measuring
+infants' perspective---and how it changes in motor development---is a
+critical part of understanding early language learning.
+
+In the current study we took a developmental approach to understanding
+the relationship between perspective and access to social information.
+We recorded head-camera data from a group of infants and children across
+a broad age range as they played with their caregivers during a brief
+laboratory visit. We then hand-annotated these data for the child's
+posture and parents' naming behavior and used face-detection algorithms
+to measure the frequency of faces in the child's visual field. The
+resulting dataset allows us to analyze changes in access to faces
+according to children's age, posture, and linguistic input.
+
+\section{Methods}\label{methods}
+
+\subsection{Participants}\label{participants}
 
@@
 
-Our final sample consisted of 36 infants and children, distributed into each of three age groups: 8 months, 12 months, and 16 months.
-Participants were recruited from the surrounding community via state birth records.
-Participants had no documented disabilities and were reported to hear at least 80\% English at home.
-Demographics and exclusion rates are given in Table \ref{tab:pop}.
+Our final sample consisted of 36 infants and children, distributed into
+each of three age groups: 8 months, 12 months, and 16 months.
+Participants were recruited from the surrounding community via state
+birth records. Participants had no documented disabilities and were
+reported to hear at least 80\% English at home. Demographics and
+exclusion rates are given in Table \ref{tab:pop}.
 
-In order to compile this final sample, we tested a substantially larger group of children, who were excluded for the following reasons: 
-`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
-`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
-`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
-`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
-`r sum(all_subs$reason=="no CDI")` for missing CDI data,
-`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
-`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses.
+In order to compile this final sample, we tested a substantially larger
+group of children, who were excluded for the following reasons: 20 for
+technical issues related to the headcam, 15 for failing to wear the
+headcam, 10 for fewer than 4 minutes of headcam footage, 5 for having
+multiple adults present, 5 for missing CDI data, 2 for missing scene
+camera footage, 1 for fussiness, and one excluded for sample symmetry.
+All inclusion decisions were made independent of the results of
+subsequent analyses.
 
-## Head-mounted camera
+\subsection{Head-mounted camera}\label{head-mounted-camera}
 
@@
 
-Our head-mounted camera ("headcam") is composed of a small, inexpensive MD80 model camera attached to a soft elastic headband from a camping headlamp.
-An aftermarket fisheye lens intended for iPhones and other Apple devices is attached to increase view angle.
-The total cost of each camera is approximately \$60.
-The camera captures 720x480 pixel images at approx.
-25 frames per second, and has battery life of 60--90 minutes.
-Without the fisheye lens, the viewing angle for the camera is $32^{\circ}$ horizontal by $24^{\circ}$ vertical; with the fisheye, $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left).
-The device is pictured in Figure \ref{fig:headcam}, right.
-Instructions for creating the camera are given in detail at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
+Our head-mounted camera (``headcam'') is composed of a small,
+inexpensive MD80 model camera attached to a soft elastic headband from a
+camping headlamp. An aftermarket fisheye lens intended for iPhones and
+other Apple devices is attached to increase view angle. The total cost
+of each camera is approximately \$60. The camera captures 720x480 pixel
+images at approx. 25 frames per second, and has battery life of 60--90
+minutes. Without the fisheye lens, the viewing angle for the camera is
+\(32^{\circ}\) horizontal by \(24^{\circ}\) vertical; with the fisheye,
+\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
+\ref{fig:headcam}, left). The device is pictured in Figure
+\ref{fig:headcam}, right. Instructions for creating the camera are given
+in detail at
+\url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
 
-The vertical field of view of the camera was considerably smaller than the child's approximate vertical field of view, which---even at 6-7 months---spans around 100--120$^{\circ}$ in the vertical dimension [@mayer1988;@cummings1988].
-We were therefore faced with a choice in the orientation of the camera.
-If we chose a lower or higher orientation, we would be at risk of truncating either the child's own hands and physically proximate objects, or the faces of the adults around the child.
-Yet if we chose the middle orientation, we would still be at risk of underestimating the proportion of faces viewed by the child.
-Thus, for the purposes of the current study---measuring visual access to faces---we chose to orient the camera towards the upper part of the visual field.
-While this orientation decreased our chances of recording the objects being manipulated by the child, it nevertheless allowed us to capture the majority of the faces in the child's visual field.
+The vertical field of view of the camera was considerably smaller than
+the child's approximate vertical field of view, which---even at 6-7
+months---spans around 100--120\(^{\circ}\) in the vertical dimension
+(Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton, 1988; Mayer,
+Fulton, \& Cummings, 1988). We were therefore faced with a choice in the
+orientation of the camera. If we chose a lower or higher orientation, we
+would be at risk of truncating either the child's own hands and
+physically proximate objects, or the faces of the adults around the
+child. Yet if we chose the middle orientation, we would still be at risk
+of underestimating the proportion of faces viewed by the child. Thus,
+for the purposes of the current study---measuring visual access to
+faces---we chose to orient the camera towards the upper part of the
+visual field. While this orientation decreased our chances of recording
+the objects being manipulated by the child, it nevertheless allowed us
+to capture the majority of the faces in the child's visual field.
 
-## Procedure
+\subsection{Procedure}\label{procedure}
 
-Before coming to the lab, caregivers completed the MacArthur-Bates Communicative Development Inventories Words and Gestures form [@fenson2007], reporting on the words that they believed that their child understood and produced.
+Before coming to the lab, caregivers completed the MacArthur-Bates
+Communicative Development Inventories Words and Gestures form (Fenson et
+al., 2007), reporting on the words that they believed that their child
+understood and produced.
 
-After coming to the lab, families were seated in our waiting room where they signed consent documents and where children were fitted with the headcam.
-After a short period of play, they were escorted to a playroom in the lab where the free-play session (the focus of the current study) was conducted.
+After coming to the lab, families were seated in our waiting room where
+they signed consent documents and where children were fitted with the
+headcam. After a short period of play, they were escorted to a playroom
+in the lab where the free-play session (the focus of the current study)
+was conducted.
 
-In the waiting room, the experimenter placed the headcam on children's heads after they had time to adjust to the environment.
-For children who resisted wearing the headcam, the experimenter used distractor techniques (presenting stimulating toys or engaging the children in hand-occupying activities) intended to keep children's focus elsewhere and prevent them from taking off the camera [@yoshida2008].
-Once the child was wearing the camera comfortably for a period of time, child and caregiver were escorted to the playroom.
+In the waiting room, the experimenter placed the headcam on children's
+heads after they had time to adjust to the environment. For children who
+resisted wearing the headcam, the experimenter used distractor
+techniques (presenting stimulating toys or engaging the children in
+hand-occupying activities) intended to keep children's focus elsewhere
+and prevent them from taking off the camera (Yoshida \& Smith, 2008).
+Once the child was wearing the camera comfortably for a period of time,
+child and caregiver were escorted to the playroom.
 
-In the playroom, the experimenter presented the child's parent with a box containing three labeled pairs of objects, each consisting of a familiar and a novel object (e.g.
-a ball and a feather duster, marked as a "zem").
-Parents confirmed that the child had not previously seen the novel toys.
-Parents were instructed to play with the object pairs with the child, one at a time, "as they typically would" and to use the novel labels to refer to the three toys.
-After giving these instructions, the experimenter left the room for a period of approximately 15 minutes.
-During this time, a tripod-mounted camera recorded video from a corner of the room and the headcam captured video from the child's perspective.
+In the playroom, the experimenter presented the child's parent with a
+box containing three labeled pairs of objects, each consisting of a
+familiar and a novel object (e.g.~a ball and a feather duster, marked as
+a ``zem''). Parents confirmed that the child had not previously seen the
+novel toys. Parents were instructed to play with the object pairs with
+the child, one at a time, ``as they typically would'' and to use the
+novel labels to refer to the three toys. After giving these
+instructions, the experimenter left the room for a period of
+approximately 15 minutes. During this time, a tripod-mounted camera
+recorded video from a corner of the room and the headcam captured video
+from the child's perspective.
 
-## Data Processing and Annotation
+\subsection{Data Processing and
+Annotation}\label{data-processing-and-annotation}
 
@@
 
-All headcam videos were cropped to exclude the period of entry to the playroom and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software.
-These sessions yielded a substantial amount of video: a total of
-`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a milion frames), with an average video length of
-`r round(sum(inc_subs$len.min)/60, digits = 2)` minutes
-(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).
+All headcam videos were cropped to exclude the period of entry to the
+playroom and were automatically synchronized with the tripod-mounted
+videos using FinalCut Pro Software. These sessions yielded a substantial
+amount of video: a total of 515 minutes (almost a milion frames), with
+an average video length of 8.58 minutes (min = 4.53, max = 19.35).
 
-### Posture and Orientation Annotation
+\subsubsection{Posture and Orientation
+Annotation}\label{posture-and-orientation-annotation}
 
-One major goal of our study was to understand the relationship between children's posture and their access to information from the faces of their caregiver.
-To investigate this relationship, we created a set of annotations for the child's physical posture (e.g.
-standing, sitting) and orientation of the caregiver relative to the child (e.g.
-in front of, behind, close, far away).
-For each headcam video, a coder used OpenSHAPA software to annotate both orientation and posture [@adolph2012].
+One major goal of our study was to understand the relationship between
+children's posture and their access to information from the faces of
+their caregiver. To investigate this relationship, we created a set of
+annotations for the child's physical posture (e.g.~standing, sitting)
+and orientation of the caregiver relative to the child (e.g.~in front
+of, behind, close, far away). For each headcam video, a coder used
+OpenSHAPA software to annotate both orientation and posture (Adolph,
+Gilmore, Freeman, Sanderson, \& Millman, 2012).
 
-Orientation was initially categorized as having the caregiver in front, to the side, or behind the child, and close (defined informally as within arm's reach) or farther away.
-Because of data sparsity, we consolidated this scheme into three categories: close to the caregiver with the caregiver either in front or on the side, farther from the caregiver again with caregiver either in front or on the side, and a global category of caregiver behind the child.
-Posture was categorized as being held/carried, lying face-up, sitting, prone (crawling or lying), standing, or other.
-Data from when the child was out of view of the tripod camera was marked as uncodable and excluded from these annotations.
-% A second coder coded XYZ videos; their categorizations were reliable at XYZ.
+Orientation was initially categorized as having the caregiver in front,
+to the side, or behind the child, and close (defined informally as
+within arm's reach) or farther away. Because of data sparsity, we
+consolidated this scheme into three categories: close to the caregiver
+with the caregiver either in front or on the side, farther from the
+caregiver again with caregiver either in front or on the side, and a
+global category of caregiver behind the child. Posture was categorized
+as being held/carried, lying face-up, sitting, prone (crawling or
+lying), standing, or other. Data from when the child was out of view of
+the tripod camera was marked as uncodable and excluded from these
+annotations. \% A second coder coded XYZ videos; their categorizations
+were reliable at XYZ.
 
-### Labeling Annotation
+\subsubsection{Labeling Annotation}\label{labeling-annotation}
 
-We were also interested in the availability of social information proximate to naming events in the caregivers' speech to children.
-Accordingly, a human coder marked the onset time when the name of any of the six objects in the object set was used.
-Overall, caregivers produced a median of XYZ labels in a highly skewed distribution across participants (range: XYZ -- XYZ).
+We were also interested in the availability of social information
+proximate to naming events in the caregivers' speech to children.
+Accordingly, a human coder marked the onset time when the name of any of
+the six objects in the object set was used. Overall, caregivers produced
+a median of XYZ labels in a highly skewed distribution across
+participants (range: XYZ -- XYZ).
 
-## Face Detection
+\subsection{Face Detection}\label{face-detection}
 
-An additional goal of the study was to measure the presence of caregivers' faces in the child's field of view (as approximated by the headcam).
-To avoid hand-annotating the size and position of faces in every frame of video, we tested two face detection systems.
-Sample frames from the video with successful detections are given in Figure \ref{fig:frames}.
+An additional goal of the study was to measure the presence of
+caregivers' faces in the child's field of view (as approximated by the
+headcam). To avoid hand-annotating the size and position of faces in
+every frame of video, we tested two face detection systems. Sample
+frames from the video with successful detections are given in Figure
+\ref{fig:frames}.
 
-### Face detection algorithms
+\subsubsection{Face detection
+algorithms}\label{face-detection-algorithms}
 
-The first algorithm was based on freely available computer vision tools [@bradski2008] and is described in depth in our previous work [@frank2012b].
-This system had two parts.
-The first was the application of a set of four Haar-style face detection filters [@viola2004] to each frame of the videos independently.
-These detectors each provide information about whether a face is present in the frame as well as size and position for any detections.
-In a second step, these detections are then combined via a hidden Markov model (HMM), trained on hand-annotated data (see Appendix).
-The HMM model (which performed nearly as well as the more complex and computationally-intensive Conditional Random Field model used in our previous work) attempted to estimate whether a face was truly present in each frame of the videos, using as its input the number of Haar detectors that were active in any given frame.
+The first algorithm was based on freely available computer vision tools
+(Bradski \& Kaehler, 2008) and is described in depth in our previous
+work (M. C. Frank, 2012). This system had two parts. The first was the
+application of a set of four Haar-style face detection filters (Viola \&
+Jones, 2004) to each frame of the videos independently. These detectors
+each provide information about whether a face is present in the frame as
+well as size and position for any detections. In a second step, these
+detections are then combined via a hidden Markov model (HMM), trained on
+hand-annotated data (see Appendix). The HMM model (which performed
+nearly as well as the more complex and computationally-intensive
+Conditional Random Field model used in our previous work) attempted to
+estimate whether a face was truly present in each frame of the videos,
+using as its input the number of Haar detectors that were active in any
+given frame.
 
-The second algorithm that we evaluated was a semi-automated adaptive tracker-by-detection (SAATD).
-The algorithm required manual user input (selecting a single face example per video) for its initialization, but then needed no additional training data.
-The tracker is based on @kalal2010, which uses patches in the trajectory of an optical-flow based tracker [@lucas1981] to train and update a face detector.
-The optical flow tracker and the face detector work in parallel.
-<!-- % For detection, we keep positive and negative templates of previously detected faces. -->
-<!-- If the face detector finds a location in a new frame exhibiting a high similarity to its stored template, the tracker is re-initialised on that location.
-Otherwise, the tracker uses the optical flow to decide the location of a face in the new frame.
--->
-<!--  % The face detector is trained using visual features [@lepetit2005} and the detection is performed using a random ferns classifier [@ozuysal2007}. -->
-<!-- The primary advantage of the SAATD algorithm is the use of motion for face detection: Following the movement of the pixels that define a face it is possible for the algorithm to adapt to new morphologies (i.e.
-different face poses).
--->
-<!-- %One disadvantage, however, is the necessity of manual initialization. -->
+The second algorithm that we evaluated was a semi-automated adaptive
+tracker-by-detection (SAATD). The algorithm required manual user input
+(selecting a single face example per video) for its initialization, but
+then needed no additional training data. The tracker is based on Kalal,
+Mikolajczyk, \& Matas (2010), which uses patches in the trajectory of an
+optical-flow based tracker (Lucas \& Kanade, 1981) to train and update a
+face detector. The optical flow tracker and the face detector work in
+parallel.
 
-<!-- % [1] Z.
-Kalal, J.
-Matas, and K.
-Mikolajczyk.
-P-N learning: Bootstrapping binary classiﬁers by structural constraints.
-In 2010 IEEE Conference on Computer Vision and Pattern Recognition (CVPR), pages 49–56.
-IEEE, June 2010.
--->
-
-<!-- % [2] B.
-D.
-Lucas and T.
-Kanade.
-An iterative image registration technique with an application to stereo vision.
-In Proceedings of the International Joint Conference on Artiﬁcial Intelligence, pages 674–679, 1981 -->
-
-<!-- % [3] V. Lepetit, P. Lagger, and P. Fua. Randomized trees for Real-Time keypoint recognition. -->
-<!-- % In IEEE Computer Society Conference on Computer Vision and Pattern Recognition, volume 2, pages 775–781, Los Alamitos, CA, USA, 2005. IEEE Computer Society. -->
-
-<!-- % [4] M.
-Ozuysal, P.
-Fua, and V.
-Lepetit.
-Fast keypoint recognition in ten lines of code.
-In IEEE Conference on Computer Vision and Pattern Recognition, Los Alamitos, CA, USA, June 2007.
-IEEE.
--->
-
-### Detector evaluation
+\subsubsection{Detector evaluation}\label{detector-evaluation}
 
@@
 
-To ensure that our evaluation was not biased by the relatively rare appearance of faces in the dataset, we annotated two samples, both a random sample from the data and a sample with a high-density of faces.
-We selected 1 minute of interaction for each age group, divided evenly across the four dyads at that age.
-For each dyad, we divided the recorded video into contiguous 1 s segments and selected 16 in accordance with two criteria.
-First, 8 of these segments were selected by choosing the parts of the videos highest in face detection (*high density sample*).
-To be fair to both algorithms, half of this was chosen from the segments with the most HMM detections and half were chosen from the segments with the most SAATD detections.
-The remaining segments were chosen by randomly sampling from segments not yet selected for coding (*random sample*).
+To ensure that our evaluation was not biased by the relatively rare
+appearance of faces in the dataset, we annotated two samples, both a
+random sample from the data and a sample with a high-density of faces.
+We selected 1 minute of interaction for each age group, divided evenly
+across the four dyads at that age. For each dyad, we divided the
+recorded video into contiguous 1 s segments and selected 16 in
+accordance with two criteria. First, 8 of these segments were selected
+by choosing the parts of the videos highest in face detection
+(\emph{high density sample}). To be fair to both algorithms, half of
+this was chosen from the segments with the most HMM detections and half
+were chosen from the segments with the most SAATD detections. The
+remaining segments were chosen by randomly sampling from segments not
+yet selected for coding (\emph{random sample}).
 
-Segments were annotated frame-by-frame by a human coder, who marked each frame as containing a face if at least half of the face was in the child's view.
-Detector output for each of these frames was then compared to this gold standard.
-A detection was counted as correct if it overlapped a face with half of its total area.
-The HMM training sample was selected via the same method as this gold standard sample, but used separate set of video segments.
+Segments were annotated frame-by-frame by a human coder, who marked each
+frame as containing a face if at least half of the face was in the
+child's view. Detector output for each of these frames was then compared
+to this gold standard. A detection was counted as correct if it
+overlapped a face with half of its total area. The HMM training sample
+was selected via the same method as this gold standard sample, but used
+separate set of video segments.
 
-We evaluated each algorithm on its precision (hits / hits + false alarms) and recall (hits / hits + misses), as well as F-score (the harmonic mean of these two measures).
-Results are reported in Table \ref{tab:model_eval}.
+We evaluated each algorithm on its precision (hits / hits + false
+alarms) and recall (hits / hits + misses), as well as F-score (the
+harmonic mean of these two measures). Results are reported in Table
+\ref{tab:model_eval}.
 
-The HMM model obtained a relatively high level of performance for the random subsections, but performed poorly when there was a relatively high density of faces present.
-In contrast, SAATD performed well on both samples, giving better performance especially in cases where there was partial occlusion.
-Our goal in using face-detection algorithms was to provide a measurement technique that eliminated tedious and expensive hand-coding and provided acceptable results.
-We therefore selected the SAATD model and report detections from this algorithm as an estimate of face presence in all further analyses.
+The HMM model obtained a relatively high level of performance for the
+random subsections, but performed poorly when there was a relatively
+high density of faces present. In contrast, SAATD performed well on both
+samples, giving better performance especially in cases where there was
+partial occlusion. Our goal in using face-detection algorithms was to
+provide a measurement technique that eliminated tedious and expensive
+hand-coding and provided acceptable results. We therefore selected the
+SAATD model and report detections from this algorithm as an estimate of
+face presence in all further analyses.
 
-# Results
+\section{Results}\label{results}
 
-We report results from three different sets of analyses.
-First, we explore developmental changes in posture and orientation in our dataset.
-Next, we explore how these changes affect access to faces, as measured using our face-detection algorithm.
-Finally, we report preliminary results on the accessibility of faces during labeling.
+We report results from three different sets of analyses. First, we
+explore developmental changes in posture and orientation in our dataset.
+Next, we explore how these changes affect access to faces, as measured
+using our face-detection algorithm. Finally, we report preliminary
+results on the accessibility of faces during labeling.
 
-## Changes in Posture and Orientation
+\subsection{Changes in Posture and
+Orientation}\label{changes-in-posture-and-orientation}
 
-<!-- \begin{figure}[t] -->
-<!-- \centering -->
-<!-- \includegraphics[width=5in]{figures/posture.pdf} -->
-<!-- \caption{\label{fig:posture} Proportion time in each posture, plotted by child's age. Error bars show 95\% confidence intervals across participants, computed by non-parametric bootstrap.} -->
-<!-- \end{figure} -->
+Our posture coding captured typical developmental milestones (Figure
+\ref{fig:posture}). Overall, sitting was the most common posture for
+interactions in the caregiver play session. The youngest infants in our
+sample mostly sat (with parental assistance), but also lay down and were
+carried a significant proportion of the time. The 12-month-olds were the
+only group who spent a large amount of time crawling, and the 16- and
+20-month-olds sat and stood in equal parts.
 
-<!-- \begin{figure}[t] -->
-<!-- \centering -->
-<!-- \includegraphics[width=5in]{figures/orientation.pdf} -->
-<!-- \caption{\label{fig:orientation} Proportion time in each orientation, plotted by child's age.
-Error bars show 95\% confidence intervals across participants, computed by non-parametric bootstrap.} -->
-<!-- \end{figure} -->
+Similarly, our coding of orientation revealed some significant
+developmental changes (Figure \ref{fig:orientation}). Younger children
+more frequently had the caregiver behind them, often because the
+caregiver was supporting the child's sitting posture (for the
+4-month-olds especially). In contrast, the 12--20 month olds were able
+to locomote independently and so were able to spend more time further
+from the caregiver.
 
-Our posture coding captured typical developmental milestones (Figure \ref{fig:posture}).
-Overall, sitting was the most common posture for interactions in the caregiver play session.
-The youngest infants in our sample mostly sat (with parental assistance), but also lay down and were carried a significant proportion of the time.
-The 12-month-olds were the only group who spent a large amount of time crawling, and the 16- and 20-month-olds sat and stood in equal parts.
+\subsection{Access to Faces}\label{access-to-faces}
 
-Similarly, our coding of orientation revealed some significant developmental changes (Figure \ref{fig:orientation}).
-Younger children more frequently had the caregiver behind them, often because the caregiver was supporting the child's sitting posture (for the 4-month-olds especially).
-In contrast, the 12--20 month olds were able to locomote independently and so were able to spend more time further from the caregiver.
+We next investigated the effects of the child's posture and orientation
+on the presence and size of the caregiver's face in the visual field.
+Figure \ref{fig:face_dets} shows the proportion of frames with a
+positive face detection, plotted by the child's age, posture, and
+orientation relative to the caregiver.
 
-## Access to Faces
+Overall, there were very large differences in access to faces across
+age. The 4-month-olds saw almost no faces---their parents were behind
+them most of the time, supporting them since they could not sit
+independently. In contrast, the 8-month-olds, who could sit
+independently, typically sat across from their caregiver and saw many
+faces in both the sitting and prone postures. The 12-month-olds spent a
+large amount of time in the prone position (typically crawling after the
+ball, for example) and saw almost no faces in that posture. The 16- and
+20-month-olds saw many faces because they were standing while their
+parents were sitting, putting their faces at a relatively similar level.
 
-<!-- % \begin{figure}[ -->
-<!-- % \begin{figure}[t] -->
-<!-- % \centering -->
-<!-- % \includegraphics[width=5in]{figures/posture.pdf} -->
-<!-- % \caption{\label{fig:posture} Proportion face detections, split by age group (left panel), posture (middle panel), and caregiver's orientation (right panel).
-We omit the lying face-up posture due to data sparsity.
-Black points show individual participants and are jittered slightly on the horizontal, red lines show means and 95\% confidence intervals.} -->
-<!-- % \end{figure}t] -->
-<!-- % \centering -->
-<!-- % \includegraphics[width=5in]{figures/posture.pdf} -->
-<!-- % \caption{\label{fig:posture} Proportion face detections, split by age group (left panel), posture (middle panel), and caregiver's orientation (right panel).
-We omit the lying face-up posture due to data sparsity.
-Black points show individual participants and are jittered slightly on the horizontal, red lines show means and 95\% confidence intervals.} -->
-<!-- % \end{figure} -->
+Across ages, the carrying and prone postures resulted in the smallest
+number of faces seen, while standing and sitting resulted in far more.
+These postures both presented opportunities for seeing faces in large
+part because parents were sitting or lying on the floor with children.
+Although far fewer faces were seen when the caregiver was behind the
+child,\footnote{Since orientation was coded via body posture, faces seen
+  while the caregiver was behind the child were due to children looking
+  over their shoulder.} both the close and far positions resulted in
+approximately equal proportions of face detections.
 
-We next investigated the effects of the child's posture and orientation on the presence and size of the caregiver's face in the visual field.
-Figure \ref{fig:face_dets} shows the proportion of frames with a positive face detection, plotted by the child's age, posture, and orientation relative to the caregiver.
+\subsection{Access to Faces During
+Labeling}\label{access-to-faces-during-labeling}
 
-Overall, there were very large differences in access to faces across age.
-The 4-month-olds saw almost no faces---their parents were behind them most of the time, supporting them since they could not sit independently.
-In contrast, the 8-month-olds, who could sit independently, typically sat across from their caregiver and saw many faces in both the sitting and prone postures.
-The 12-month-olds spent a large amount of time in the prone position (typically crawling after the ball, for example) and saw almost no faces in that posture.
-The 16- and 20-month-olds saw many faces because they were standing while their parents were sitting, putting their faces at a relatively similar level.
+Our final analysis concerned the accessibility of caregivers' faces
+during labeling events. Franchak et al. (2011) found that referential
+speech was marginally more likely to draw toddlers' attention to
+mothers' faces. We were similarly interested in whether looking at faces
+occurred during labeling. Accordingly, we used the labeling annotations
+for each child to identify the 2s before and after each labeling event.
+We then computed the proportion face detections within this window
+across ages.
 
-a
-## Access to Faces During Labeling
-
-Our final analysis concerned the accessibility of caregivers' faces during labeling events.
-@franchak2011 found that referential speech was marginally more likely to draw toddlers' attention to mothers' faces.
-We were similarly interested in whether looking at faces occurred during labeling.
-Accordingly, we used the labeling annotations for each child to identify the 2s before and after each labeling event.
-We then computed the proportion face detections within this window across ages.
-
-The overall pattern of face accessibility closely mirrored the base rates shown in Figure \ref{fig:face_dets}.
-Although this general pattern in itself is important in assessing developmental access to social information, in the current analysis we were interested in whether there was differential access to faces around labeling instances.
-We thus computed difference scores between the baseline face detection rate and the rate of face detections in labeling windows for each participant.
+The overall pattern of face accessibility closely mirrored the base
+rates shown in Figure \ref{fig:face_dets}. Although this general pattern
+in itself is important in assessing developmental access to social
+information, in the current analysis we were interested in whether there
+was differential access to faces around labeling instances. We thus
+computed difference scores between the baseline face detection rate and
+the rate of face detections in labeling windows for each participant.
 Figure \ref{fig:naming_face} shows the results of this analysis.
 
-Although any conclusion must remain extremely tentative because of the small sample, we nevertheless saw an increase in label-related face access for the 20-month-olds.
-This difference was robust across a variety of window sizes from 1--6 s.
-(8-month-olds were more variable but similarly showed some trend towards greater face access during naming.) We cannot yet make inferences about the source of these differences: They could be could be caused by children, caregivers, or a combination of the two.
-Nevertheless, these results converge with previous work and suggest that, in combination with face detection techniques, the headcam may be a viable method for examining social access during language learning.
+Although any conclusion must remain extremely tentative because of the
+small sample, we nevertheless saw an increase in label-related face
+access for the 20-month-olds. This difference was robust across a
+variety of window sizes from 1--6 s. (8-month-olds were more variable
+but similarly showed some trend towards greater face access during
+naming.) We cannot yet make inferences about the source of these
+differences: They could be could be caused by children, caregivers, or a
+combination of the two. Nevertheless, these results converge with
+previous work and suggest that, in combination with face detection
+techniques, the headcam may be a viable method for examining social
+access during language learning.
 
-<!-- % \begin{figure} -->
-<!-- % \begin{center} -->
-<!-- %   \includegraphics[width=3in]{figures/naming_faces_diff.pdf} -->
-<!-- % \end{center} -->
-<!-- % \caption{\label{fig:naming_face} Proportion of faces detected in a 4s window of time centered around labeling events, plotted by age group and whether the word was familiar or novel.
-Error bars show 95\% confidence intervals.
-4-month-olds are omitted due to the limited number of total face detections for this group.} -->
-<!-- % % \vspace{-ex} -->
-<!-- % \end{figure} -->
+\section{General Discussion}\label{general-discussion}
 
-# General Discussion
+Using a head-mounted camera, we explored the relationship between
+infants' postural and locomotor development and their visual access to
+social information. The use of automated annotation tools from computer
+vision allowed us to measure the prevalence of caregivers' faces in
+their children's visual field. We found systematic differences in the
+visual accessibility of faces based on posture, orientation, and age, as
+well as hints of differences in language-related changes in visual
+access. While these results remain preliminary given the size of our
+developmental sample, this work nevertheless provides an important
+proof-of-concept that computer vision techniques can be used as a
+measurement method in the developmental context.
 
-Using a head-mounted camera, we explored the relationship between infants' postural and locomotor development and their visual access to social information.
-The use of automated annotation tools from computer vision allowed us to measure the prevalence of caregivers' faces in their children's visual field.
-We found systematic differences in the visual accessibility of faces based on posture, orientation, and age, as well as hints of differences in language-related changes in visual access.
-While these results remain preliminary given the size of our developmental sample, this work nevertheless provides an important proof-of-concept that computer vision techniques can be used as a measurement method in the developmental context.
+The measures developed here have broad applicability to the study of
+individual and cultural differences. Since the physical circumstances of
+child rearing vary widely across households and across cultures, there
+may be important and predictable differences in children's visual
+experience. As suggested by the correlations between walking and
+vocabulary development ({\textbf{???}}), postural development may have
+substantial downstream consequences for language. For example, shifts in
+how infants are placed in particular postures by strollers or carriers
+(Zeedyk, 2008) or how their motor development is encouraged by parent
+practices (Bril \& Sabatier, 1986) may lead to differences in social
+input which in turn affect their language learning. Since our variant of
+the headcam method is both inexpensive and highly portable, we have been
+able to deploy it in children's homes with some success; it may thus be
+a valuable tool for investigating differences in child-rearing
+practices.
 
-The measures developed here have broad applicability to the study of individual and cultural differences.
-Since the physical circumstances of child rearing vary widely across households and across cultures, there may be important and predictable differences in children's visual experience.
-As suggested by the correlations between walking and vocabulary development [@walleunderreview], postural development may have substantial downstream consequences for language.
-For example, shifts in how infants are placed in particular postures by strollers or carriers [@zeedyk2008] or how their motor development is encouraged by parent practices [@bril1986] may lead to differences in social input which in turn affect their language learning.
-Since our variant of the headcam method is both inexpensive and highly portable, we have been able to deploy it in children's homes with some success; it may thus be a valuable tool for investigating differences in child-rearing practices.
+A deep body of work uses children's linguistic input---measured using
+audio recordings---to understand the learning mechanisms underlying
+vocabulary acquisition (Fernald, Perfors, \& Marchman, 2006; Hart \&
+Risley, 1995; Huttenlocher, Haight, Bryk, Seltzer, \& Lyons, 1991).
+There have been some important initial successes in using visual input
+to predict language uptake (Yu \& Smith, in press). Nevertheless, we
+have a long way to go before our knowledge about children's visual input
+parallels our understanding of their linguistic environment. Coming to
+such an understanding will require the creation of both corpus resources
+and automated tools such as those we have begun to develop here.
 
-A deep body of work uses children's linguistic input---measured using audio recordings---to understand the learning mechanisms underlying vocabulary acquisition [@huttenlocher1991;@hart1995;@fernald2006].
-There have been some important initial successes in using visual input to predict language uptake [@yuinpress].
-Nevertheless, we have a long way to go before our knowledge about children's visual input parallels our understanding of their linguistic environment.
-Coming to such an understanding will require the creation of both corpus resources and automated tools such as those we have begun to develop here.
+\section{View angle}\label{view-angle}
 
-# View angle
+We note that previous studies have shown that children's head movements
+in the horizontal dimension are approximated by (though are slightly
+lagged by) their head movements (Yoshida \& Smith, 2008). Our own
+experience with the current apparatus ratifies these conclusions for the
+horizontal field but suggests that head movements in the vertical field
+are less reliable. Hence, these studies may run the risk of
+underestimating the proportion of faces actually seen by children.
 
-We note that previous studies have shown that children's head movements in the horizontal dimension are approximated by (though are slightly lagged by) their head movements [@yoshida2008].
-Our own experience with the current apparatus ratifies these conclusions for the horizontal field but suggests that head movements in the vertical field are less reliable.
-Hence, these studies may run the risk of underestimating the proportion of faces actually seen by children.
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 
-# References
+\hypertarget{refs}{}
+\hypertarget{ref-adolph2007}{}
+Adolph, K., \& Berger, S. (2007). Motor development. In \emph{Handbook
+of child psychology}. Wiley Online Library.
+
+\hypertarget{ref-adolph2012}{}
+Adolph, K., Gilmore, R., Freeman, C., Sanderson, P., \& Millman, D.
+(2012). Toward open behavioral science. \emph{Psychological Inquiry},
+\emph{23}(3), 244--247.
+
+\hypertarget{ref-bradski2008}{}
+Bradski, G., \& Kaehler, A. (2008). \emph{Learning OpenCV: Computer
+vision with the OpenCV library}. O'Reilly Media.
+
+\hypertarget{ref-bril1986}{}
+Bril, B., \& Sabatier, C. (1986). The cultural context of motor
+development: Postural manipulations in the daily life of Bambara babies
+(Mali). \emph{International Journal of Behavioral Development},
+\emph{9}(4), 439--453.
+
+\hypertarget{ref-brooks2005}{}
+Brooks, R., \& Meltzoff, A. (2005). The development of gaze following
+and its relation to language. \emph{Developmental Science}, \emph{8}(6),
+535--543.
+
+\hypertarget{ref-carpenter1998}{}
+Carpenter, M., Nagell, K., \& Tomasello, M. (1998). Social cognition,
+joint attention, and communicative competence from 9 to 15 months of
+age. \emph{Monographs of the Society for Research in Child Development},
+\emph{63}(4).
+
+\hypertarget{ref-cummings1988}{}
+Cummings, M., Van Hof-Van Duin, J., Mayer, D., Hansen, R., \& Fulton, A.
+(1988). Visual fields of young children. \emph{Behavioural and Brain
+Research}, \emph{29}(1), 7--16.
+
+\hypertarget{ref-fenson2007}{}
+Fenson, L., Bates, E., Dale, P. S., Marchman, V. A., Reznick, J. S., \&
+Thal, D. J. (2007). \emph{MacArthur-bates communicative development
+inventories: Users manual}. Brookes Publishing Company.
+
+\hypertarget{ref-fernald2006}{}
+Fernald, A., Perfors, A., \& Marchman, V. (2006). Picking up speed in
+understanding: Speech processing efficiency and vocabulary growth across
+the 2nd year. \emph{Developmental Psychology}, \emph{42}(1), 98--116.
+
+\hypertarget{ref-franchak2011}{}
+Franchak, J., Kretch, K., Soska, K., \& Adolph, K. (2011). Head-mounted
+eye tracking: A new method to describe infant looking. \emph{Child
+Development}.
+
+\hypertarget{ref-frank2012b}{}
+Frank, M. C. (2012). Measuring children's visual access to social
+information using face detection. In \emph{Proceedings of the 33nd
+Annual Meeting of the Cognitive Science Society}.
+
+\hypertarget{ref-frank2013}{}
+Frank, M. C., Simmons, K., Yurovsky, D., \& Pusiol, G. (2013).
+Developmental and postural changes in childrens visual access to faces.
+In \emph{Proceedings of the 35th annual meeting of the cognitive science
+society} (pp. 454--459).
+
+\hypertarget{ref-hart1995}{}
+Hart, B., \& Risley, T. (1995). \emph{Meaningful differences in the
+everyday experience of young American children}. Baltimore, MD: Brookes
+Publishing Company.
+
+\hypertarget{ref-huttenlocher1991}{}
+Huttenlocher, J., Haight, W., Bryk, A., Seltzer, M., \& Lyons, T.
+(1991). Early vocabulary growth: Relation to language input and gender.
+\emph{Developmental Psychology}, \emph{27}(2), 236--248.
+
+\hypertarget{ref-kalal2010}{}
+Kalal, Z., Mikolajczyk, K., \& Matas, J. (2010). Forward-backward error:
+Automatic detection of tracking failures. In \emph{20th International
+Conference on Pattern Recognition} (pp. 2756--2759). IEEE.
+
+\hypertarget{ref-lucas1981}{}
+Lucas, B., \& Kanade, T. (1981). An iterative image registration
+technique with an application to stereo vision. In \emph{Proceedings of
+the 7th International Joint Conference on Artificial intelligence}.
+
+\hypertarget{ref-mayer1988}{}
+Mayer, D., Fulton, A., \& Cummings, M. (1988). Visual fields of infants
+assessed with a new perimetric technique. \emph{Investigative
+Ophthalmology \& Visual Science}, \emph{29}(3), 452--459.
+
+\hypertarget{ref-scaife1975}{}
+Scaife, M., \& Bruner, J. (1975). The capacity for joint visual
+attention in the infant. \emph{Nature}, \emph{253}, 265--266.
+
+\hypertarget{ref-smith2011}{}
+Smith, L. B., Yu, C., \& Pereira, A. F. (2011). Not your mother's view:
+The dynamics of toddler visual experience. \emph{Developmental Science},
+\emph{14}, 9--17.
+
+\hypertarget{ref-tomasello2009}{}
+Tomasello, M. (2009). \emph{The cultural origins of human cognition}.
+Harvard University Press.
+
+\hypertarget{ref-viola2004}{}
+Viola, P., \& Jones, D. H. (2004). Robust real-time face detection.
+\emph{International Journal of Computer Vision}.
+
+\hypertarget{ref-walle2014}{}
+Walle, E. A., \& Campos, J. J. (2014). Infant language development is
+related to the acquisition of walking. \emph{Developmental Psychology},
+\emph{50}(2), 336.
+
+\hypertarget{ref-yoshida2008}{}
+Yoshida, H., \& Smith, L. (2008). What's in view for toddlers? Using a
+head camera to study visual experience. \emph{Infancy}, \emph{13},
+229--248.
+
+\hypertarget{ref-yuinpress}{}
+Yu, C., \& Smith, L. B. (in press). Embodied attention and word learning
+by toddlers. \emph{Cognition}.
+
+\hypertarget{ref-yurovsky2012}{}
+Yurovsky, D., Smith, L., \& Yu, C. (in press). Statistical word learning
+at scale: The baby's view is better. \emph{Developmental Science}.
+
+\hypertarget{ref-zeedyk2008}{}
+Zeedyk, M. (2008). \emph{What's life in a baby buggy like?: The impact
+of buggy orientation on parent-infant interaction and infant stress}.
+London, UK: National Literacy Trust.
+
+\bibliography{xsface.bib}
+
+\end{document}

### 2018-01-17 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
quick edits from first pass of abdellah's comments

--- 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
     \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
-        \And {\large \bf Ally Kraus} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
+        \And {\large \bf Ally Kraus} \\ \texttt{XX@XX} \\ Department of Psychology \\ Stanford University
         \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}
@@
 Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use state-of-the-art computer vision algorithms to detect the proportion of faces in infants' environments.
-We find that infants' postural changes largely mediate infants' access to faces, hinting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
+We find that infants' posture and orientation to their caregiver largely mediate infants' access to faces, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
 Broadly, we suggest that the combined use of head-mounted cameras and the application of novel deep learning algorithms is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
@@
 On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access social information (e.g., faces, gaze) relative to children who are still crawling and sitting.
-In other words, the ability to access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.
+In other words, the ability to walk and move around independently and subsequently access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.
 
-Recent work has begun to use egocentric, head-mounted cameras to document the visual experiences of infants and children — which even for walking children are radically different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011; @yoshida2008].
+Recent technological developments allow for testing of this hypothesis by documenting the experiences of infants and children from their own perspective.
+By using head-mounted cameras, researchers have begun to document visual experiences of infants and children — which even for walking children are radically different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011; @yoshida2008].
 Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2012].
@@
 
-Here, we directly ask whether the postural and locomotive developments change the availability of the social information relevant for word learning.
+Here, we directly ask whether  postural and locomotive developments change the availability of the social information relevant for word learning.
 To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
@@
 For the first two annotations (close/far from the child), the caregiver could either be to the the front or to the side of the child.
-All annotations were made using OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.
+All annotations were made by a trained coder using the OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.
 
@@
 We first measured the performance of the most commonly-used and widely available face detection algorithms (Viola-Jones).
-We used this as a benchmark for performance, and while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces (@scheirer2014perceptual).
+We used this as a benchmark for performance, and while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
 We next capitalized on recent improvements in computer vision, testing the performance of two state-of-the-art face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information.
+The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of agent's bodyparts.
 
@@
 ### Face Detection Algorithms
-The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, @viola2004robust) applied to each individual frame.
+The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, [@viola2004robust]) applied to each individual frame.
 This detector provided information about the presence of a face as well as its size and position.
@@
 
-The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) (@cao2017realtime, @simon2017hand, @wei2016cpm) avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm] avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
@@
   \hline
-1 & MTCNN & high\_density & 0.841 & 1.000 & 0.914 \\ 
-  2 & MTCNN & random & 0.947 & 0.750 & 0.837 \\ 
-  3 & OpenPose & high\_density & 0.995 & 0.863 & 0.924 \\ 
-  4 & OpenPose & random & 0.710 & 0.917 & 0.800 \\ 
-  5 & ViolaJones & high\_density & 0.991 & 0.495 & 0.660 \\ 
-  6 & ViolaJones & random & 0.389 & 0.292 & 0.333 \\ 
+1 & MTCNN & High\ density & 0.841 & 1.000 & 0.914 \\ 
+  2 & MTCNN & Random & 0.947 & 0.750 & 0.837 \\ 
+  3 & OpenPose & High\ density & 0.995 & 0.863 & 0.924 \\ 
+  4 & OpenPose & Random & 0.710 & 0.917 & 0.800 \\ 
+  5 & ViolaJones & High\ Density & 0.991 & 0.495 & 0.660 \\ 
+  6 & ViolaJones & Random & 0.389 & 0.292 & 0.333 \\ 
    \hline
 \end{tabular}
-\caption{Detector performance} 
+\caption{Detector performance.} 
 \end{table}
@@
 
-Next, we turned to our hypothesis of interest, examining the impact of postural and locomotive developments on children's visual access to faces.
-Overall, we found that children's posture was a major factor both in how many faces they saw during the play session.
+However, we found that any age related effects were much smaller compared to the impact of postural and locomotive changes on children's visual access to faces.
+Children's posture was a major factor both in how many faces they saw during the play session.
 Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces.
@@
 
-<!-- ```{r echo=FALSE, include=FALSE} -->
-<!-- d$faceNum=as.numeric(d$faceMT) -->
-<!-- d$handNum=as.numeric(d$handOP) -->
+```{r echo=FALSE, include=FALSE}
+d$faceNum_MT=as.numeric(d$faceMT)
+d$faceNum_OP=as.numeric(d$faceMT)
 
-<!-- # Fails to converge -->
-<!-- # fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), -->
-<!-- #       data = d, -->
-<!-- #       family = "binomial") -->
+# Fails to converge
+# fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid),
+#       data = d,
+#       family = "binomial")
 
-<!-- # Fails to converge -->
-<!-- # agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid), -->
-<!-- #       data = d, -->
-<!-- #       family = "binomial") -->
+# Fails to converge
+# agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid),
+#       data = d,
+#       family = "binomial")
 
-<!-- fullModelminsInt_Faces <- glmer(faceNum ~ scale(age.at.test) + posture + orientation + (1|subid), -->
-<!--        data = d, -->
-<!--        family = "binomial") -->
+fullModelminsInt_Faces_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + posture + orientation + (1|subid),
+       data = d,
+       family = "binomial")
 
-<!-- fullModelminsInt_Handd <- glmer(handNum ~ scale(age.at.test) + posture + orientation + (1|subid), -->
-<!--        data = d, -->
-<!--        family = "binomial") -->
+fullModelminsInt_Faces_OP <- glmer(faceNum_OP ~ scale(age.at.test) + posture + orientation + (1|subid),
+       data = d,
+       family = "binomial")
 
-<!-- ageOnly <- glmer(faceNum ~ scale(age.at.test) + (1|subid), -->
-<!--       data = d, -->
-<!--       family = "binomial") -->
-
-<!-- ``` -->
+```
 
@@
 Interactions between predictors were not included as this maximal model failed to converge.
-A summary of the coefficients of this model can be found in Table 3.
-While age remained a significant predictor even when accounting for the effects of infants' posture and caregivers' orientation, it held relatively less predictive power.
-Overall, these results confirm that infants' access to faces is heavily infleunced by their own posture and their caregiver's orientation towards them.
+A summary of the coefficients of this model can be found in Table 3 for the MTCNN detections, though we found the same pattern of results for the OpenPose detections.
+While age remained a significant predictor even when accounting for the effects of infants' posture and  orientation, it accounted for the least amount of variance.
+Overall, these results suggest that, instead, infants' access to faces is heavily infleunced by their own posture and theirorientation towards their caregiver.
 
-<!-- ```{r xtable, results="asis"} -->
-<!-- tab2 <- xtable::xtable(summary(fullModelminsInt_Faces)$coef, digits=3, -->
-<!--                       caption = "Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.") -->
+```{r xtable, results="asis"}
+tab2 <- xtable::xtable(summary(fullModelminsInt_Faces_MTCNN)$coef, digits=3,
+                      caption = "Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.")
 
-<!-- print(tab2, type="latex", comment = F, table.placement = "H") -->
-<!-- ``` -->
+print(tab2, type="latex", comment = F, table.placement = "H")
+```
 
-<!-- ```{r xtable, results="asis"} -->
-<!--  tab3 <- xtable::xtable(summary(fullModelminsInt_Handd)$coef, digits=3, -->
-<!--                        caption = "Results from GLMM model prediting the presence/absence of hands (OpenPose-Hands) across the entire dataset.") -->
+```{r xtable, results="asis"}
+ tab3 <- xtable::xtable(summary(fullModelminsInt_Faces_OP)$coef, digits=3,
+                       caption = "Results from GLMM model prediting the presence/absence of faces (OpenPose-Faces) across the entire dataset.")
 
-<!--  print(tab3, type="latex", comment = F, table.placement = "H") -->
-<!-- ``` -->
+ print(tab3, type="latex", comment = F, table.placement = "H")
+```
 
@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming window as a function of infants's posture"  }
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ("Look, a Zem"; +/- 2 seconds around each utternace) as a function of infants's posture.
+Each point represents a naming instance for each child in the dataset."  }
 ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
   geom_boxplot() + 
-  ylab("Prop. face detections in naming window") + 
+  ylab("Prop. face detections") + 
   xlab("Posture") + 

### 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
both detectors -- tables (need formatting), some edits to figure captions

@@
 
-```{r xtable, results="asis"}
+```{r MTCNNModel, results="asis"}
 tab2 <- xtable::xtable(summary(fullModelminsInt_Faces_MTCNN)$coef, digits=3,
@@
 
-```{r xtable, results="asis"}
+```{r OPModel, results="asis"}
  tab3 <- xtable::xtable(summary(fullModelminsInt_Faces_OP)$coef, digits=3,
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
-  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
-  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
-  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
-  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
-  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
-  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
-   \hline
-\end{tabular}
-\caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).} 
-\end{table}
+<!-- \begin{table}[H] -->
+<!-- \centering -->
+<!-- \begin{tabular}{rrrrr} -->
+<!--   \hline -->
+<!--  & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\  -->
+<!--   \hline -->
+<!-- (Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\  -->
+<!--   Age & 0.170 & 0.061 & 2.766 & 0.006 \\  -->
+<!--   Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\  -->
+<!--   Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\  -->
+<!--   Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\  -->
+<!--   Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\  -->
+<!--   Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\  -->
+<!--    \hline -->
+<!-- \end{tabular} -->
+<!-- \caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).}  -->
+<!-- \end{table} -->
 
@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ("Look, a Zem"; +/- 2 seconds around each utternace) as a function of infants's posture.
-Each point represents a naming instance for each child in the dataset."  }
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utternace) as a function of infants's posture.
+Each point represents a naming instance for each child in the dataset."}
 ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 

### 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
new paragraph
both detectors -- tables (need formatting), some edits to figure captions

--- 2018-01-16 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
@@
 \author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
-    \And {\large \bf Ally Kraus} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
+    \And {\large \bf Ally Kraus} \\ \texttt{XX@XX} \\ Department of Psychology \\ Stanford University
     \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}
@@
 state-of-the-art computer vision algorithms to detect the proportion of
-faces in infants' environments. We find that infants' postural changes
-largely mediate infants' access to faces, hinting that these postural
-and locomotive developments facilitate infants' emerging linguistic and
-social capacities. Broadly, we suggest that the combined use of
-head-mounted cameras and the application of novel deep learning
-algorithms is a promising avenue for understanding the statistics of
-infants' visual and linguistic experience.
+faces in infants' environments. We find that infants' posture and
+orientation to their caregiver largely mediate infants' access to faces,
+suggesting that these postural and locomotive developments facilitate
+infants' emerging linguistic and social capacities. Broadly, we suggest
+that the combined use of head-mounted cameras and the application of
+novel deep learning algorithms is a promising avenue for understanding
+the statistics of infants' visual and linguistic experience.
 
@@
 access social information (e.g., faces, gaze) relative to children who
-are still crawling and sitting. In other words, the ability to access
-more detailed social information may allow infants to learn words
-quicker and more efficiently, facilitating language growth.
+are still crawling and sitting. In other words, the ability to walk and
+move around independently and subsequently access more detailed social
+information may allow infants to learn words quicker and more
+efficiently, facilitating language growth.
 
-Recent work has begun to use egocentric, head-mounted cameras to
-document the visual experiences of infants and children --- which even
-for walking children are radically different from the adult perspective
-(and not easily predicted by our own adult intuitions) (Clerkin, Hart,
-Rehg, Yu, \& Smith, 2017; J. Franchak et al., 2011; Yoshida \& Smith,
-2008). Children's views tend to be restricted and to be dominated by
-objects and hands much more than that of adults (Yoshida \& Smith,
-2008), and both computational and empirical work suggest that this
-restricted viewpoint may be more effective for learning objects and
-their labels than the comparable adult perspective (Bambach, Crandall,
-Smith, \& Yu, 2017; D. Yurovsky, Smith, \& Yu, in press). Further,
-recent work also suggests dramatic changes in the child's perspective
-over the first two years of life, as views transition from primarily
-containing a close up view of faces to capturing views of hands paired
-with the objects they are acting on (Fausey, Jayaraman, \& Smith, 2016).
+Recent technological developments allow for testing of this hypothesis
+by documenting the experiences of infants and children from their own
+perspective. By using head-mounted cameras, researchers have begun to
+document visual experiences of infants and children --- which even for
+walking children are radically different from the adult perspective (and
+not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
+Yu, \& Smith, 2017; J. Franchak et al., 2011; Yoshida \& Smith, 2008).
+Children's views tend to be restricted and to be dominated by objects
+and hands much more than that of adults (Yoshida \& Smith, 2008), and
+both computational and empirical work suggest that this restricted
+viewpoint may be more effective for learning objects and their labels
+than the comparable adult perspective (Bambach, Crandall, Smith, \& Yu,
+2017; D. Yurovsky, Smith, \& Yu, in press). Further, recent work also
+suggests dramatic changes in the child's perspective over the first two
+years of life, as views transition from primarily containing a close up
+view of faces to capturing views of hands paired with the objects they
+are acting on (Fausey, Jayaraman, \& Smith, 2016).
 
-Here, we directly ask whether the postural and locomotive developments
+Here, we directly ask whether postural and locomotive developments
 change the availability of the social information relevant for word
@@
 from the child), the caregiver could either be to the the front or to
-the side of the child. All annotations were made using OpenSHAPA/Datavyu
-software (K. Adolph, Gilmore, Freeman, Sanderson, \& Millman, 2012), and
-times when the child was out of view of the tripod camera were marked as
-uncodable and were excluded from these annotations.
+the side of the child. All annotations were made by a trained coder
+using the OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman,
+Sanderson, \& Millman, 2012), and times when the child was out of view
+of the tripod camera were marked as uncodable and were excluded from
+these annotations.
 
@@
 
-Measuring the availability of caregiver's faces across development was
-an important goal of this study. We used a total of three face detection
-systems to explore infants' changing access to social information and to
-avoid the cost of hand-annotating every frame. We first measured the
-performance of the most commonly-used and widely available face
-detection algorithms (Viola-Jones). We used this as a benchmark for
-performance, and while it can achieve impressive accuracy in some
-situations, it is notoriously bad at dealing with occluded faces
-(Scheirer, Anthony, Nakayama, \& Cox (2014)). We next capitalized on
-recent improvements in computer vision, testing the performance of two
-state-of-the-art face detectors that both made use of Convolutional
-Neural Networks (CNNs) to extract face information.
+We used a total of three face detection systems to explore infants'
+changing access to social information and to avoid the cost of
+hand-annotating every frame. We first measured the performance of the
+most commonly-used and widely available face detection algorithms
+(Viola-Jones). We used this as a benchmark for performance, and while it
+can achieve impressive accuracy in some situations, it is notoriously
+bad at dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox,
+2014). We next capitalized on recent improvements in computer vision,
+testing the performance of two state-of-the-art face detectors that both
+made use of Convolutional Neural Networks (CNNs) to extract face
+information. The first algorithm was specifically optimized for face
+detection, and the second algorithm was optimized to extract information
+about the position of agent's bodyparts.
 
@@
 The first face detection system made use of a series of Haar
-feature-based cascade classifiers (Viola-Jones, Viola \& Jones (2004))
+feature-based cascade classifiers (Viola-Jones, (Viola \& Jones, 2004))
 applied to each individual frame. This detector provided information
@@
 The third algorithm was a CNN-based pose detector that provided the
-locations of 18 body parts (ears, nose, wrists, etc.) (Cao et al.
-(2017), Simon, Joo, Matthews, \& Sheikh (2017), Wei, Ramakrishna,
-Kanade, \& Sheikh (2016)) avaliable at
+locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose
+(Cao et al., 2017; Simon, Joo, Matthews, \& Sheikh, 2017; Wei,
+Ramakrishna, Kanade, \& Sheikh, 2016) avaliable at
 \url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
@@
 and F-score (harmonic mean of previous measures) were calculated and are
-reported in Table 2. Both OpenPose and MTCNN detectors performed
-relatively well on the gold set, with MTCNN outperforming OpenPose on
-the random sample but trailing behind in the high density sample. Due to
-the high performance of these two detectors we have included the results
-of both in the analyses to follow. The ViolaJones detector did not
-perform as well on either gold set, so the results associated with this
-detector were not included.
+reported in Table 2.
+
+Both OpenPose and MTCNN detectors performed relatively well on the gold
+set, with MTCNN outperforming OpenPose on the random sample but trailing
+behind in the high density sample. Due to the high performance of these
+two detectors we have included the results of both in the analyses to
+follow. The ViolaJones detector did not perform well on either gold set,
+so the results from this detector were not included.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllrrr}
+  \hline
+ & Algorithm & Sample\ Type & P & R & F \\ 
+  \hline
+1 & MTCNN & High\ density & 0.841 & 1.000 & 0.914 \\ 
+  2 & MTCNN & Random & 0.947 & 0.750 & 0.837 \\ 
+  3 & OpenPose & High\ density & 0.995 & 0.863 & 0.924 \\ 
+  4 & OpenPose & Random & 0.710 & 0.917 & 0.800 \\ 
+  5 & ViolaJones & High\ Density & 0.991 & 0.495 & 0.660 \\ 
+  6 & ViolaJones & Random & 0.389 & 0.292 & 0.333 \\ 
+   \hline
+\end{tabular}
+\caption{Detector performance.} 
+\end{table}
 
@@
 
-Next, we turned to our hypothesis of interest, examining the impact of
-postural and locomotive developments on children's visual access to
-faces. Overall, we found that children's posture was a major factor both
-in how many faces they saw during the play session. Infants who were
+However, we found that any age related effects were much smaller
+compared to the impact of postural and locomotive changes on children's
+visual access to faces. Children's posture was a major factor both in
+how many faces they saw during the play session. Infants who were
 sitting saw more faces than infants who were lying down or being
@@
 
-\subsection{Model}\label{model}
-
 To formalize these observations, we fit a generalized logistic
@@
 as this maximal model failed to converge. A summary of the coefficients
-of this model can be found in Table 3. While age remained a significant
-predictor even when accounting for the effects of infants' posture and
-caregivers' orientation, it held relatively less predictive power.
-Overall, these results confirm that infants' access to faces is heavily
-infleunced by their own posture and their caregiver's orientation
-towards them.
+of this model can be found in Table 3 for the MTCNN detections, though
+we found the same pattern of results for the OpenPose detections. While
+age remained a significant predictor even when accounting for the
+effects of infants' posture and orientation, it accounted for the least
+amount of variance. Overall, these results suggest that, instead,
+infants' access to faces is heavily infleunced by their own posture and
+theirorientation towards their caregiver.
 
@@
 (Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
-  Age & 0.170 & 0.061 & 2.766 & 0.006 \\ 
-  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
-  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
-  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
-  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\ 
-  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\ 
+  scale(age.at.test) & 0.170 & 0.061 & 2.766 & 0.006 \\ 
+  postureprone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
+  posturesit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
+  posturestand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
+  orientationclose & 1.549 & 0.021 & 72.757 & 0.000 \\ 
+  orientationfar & 2.258 & 0.022 & 101.138 & 0.000 \\ 
    \hline
 \end{tabular}
-\caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).} 
+\caption{Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.} 
+\end{table}
+
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
+  scale(age.at.test) & 0.170 & 0.061 & 2.766 & 0.006 \\ 
+  postureprone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
+  posturesit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
+  posturestand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
+  orientationclose & 1.549 & 0.021 & 72.757 & 0.000 \\ 
+  orientationfar & 2.258 & 0.022 & 101.138 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{Results from GLMM model prediting the presence/absence of faces (OpenPose-Faces) across the entire dataset.} 
 \end{table}
@@
 
-\caption[Proportion face detections as a function of infants's posture]{Proportion face detections as a function of infants's posture}\label{fig:detByNaming}
+\caption[Proportion face detections around a naming instance ('Look, a Zem']{Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utternace) as a function of infants's posture.
+Each point represents a naming instance for each child in the dataset.}\label{fig:detByNaming}
 \end{figure}

### 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
typos

@@
 While age remained a significant predictor even when accounting for the effects of infants' posture and  orientation, it accounted for the least amount of variance.
-Overall, these results suggest that, instead, infants' access to faces is heavily infleunced by their own posture and theirorientation towards their caregiver.
+Overall, these results suggest that, instead, infants' access to faces is heavily infleunced by their own posture and their orientation towards their caregiver.
 
@@
 tab2 <- xtable::xtable(summary(fullModelminsInt_Faces_MTCNN)$coef, digits=3,
-                      caption = "Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.")
+                      caption = "Results from GLMM model predicting the presence/absence of a face (MTCNN-Faces) across the entire dataset.")
 
@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utternace) as a function of infants's posture.
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
 Each point represents a naming instance for each child in the dataset."}

### 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
typos

@@
   \hline
-(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\ 
-  scale(age.at.test) & 0.170 & 0.061 & 2.766 & 0.006 \\ 
-  postureprone & -0.513 & 0.036 & -14.045 & 0.000 \\ 
-  posturesit & 0.648 & 0.034 & 18.853 & 0.000 \\ 
-  posturestand & 1.005 & 0.035 & 28.402 & 0.000 \\ 
-  orientationclose & 1.549 & 0.021 & 72.757 & 0.000 \\ 
-  orientationfar & 2.258 & 0.022 & 101.138 & 0.000 \\ 
+(Intercept) & -4.847 & 0.064 & -75.303 & 0.000 \\ 
+  scale(age.at.test) & 0.170 & 0.075 & 2.273 & 0.023 \\ 
+  postureprone & -0.513 & 0.042 & -12.326 & 0.000 \\ 
+  posturesit & 0.648 & 0.039 & 16.422 & 0.000 \\ 
+  posturestand & 1.005 & 0.041 & 24.721 & 0.000 \\ 
+  orientationclose & 1.549 & 0.022 & 71.940 & 0.000 \\ 
+  orientationfar & 2.258 & 0.022 & 100.357 & 0.000 \\ 
    \hline

### 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
table 2 rewrite
various edits, and optional namings plot with CI intervals

--- 2018-01-17 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
   if (a == TRUE) {
-    if (a == b) return ("TP") # was face, said face
+    if (a == b) return ("TP") # was face, detected face
     else return("FN") # was face, missed face
@@
   else {
-    if (a == b) return("TN") # was not face, said not face
+    if (a == b) return("TN") # was not face, detected face
     else return("FP") # was not face, said face
@@
 
-```{r orientation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each orientation relative to their caregiver." }
-
-ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
+```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each posture." }
+ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
   geom_boxplot() + 
@@
 
-```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each posture." }
-ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
+```{r orientation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each orientation relative to their caregiver." }
+
+ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
   geom_boxplot() + 
@@
 # Changes in Access to Faces 
+```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and the OpenPose model (right panel) as a function of child's age.
+Larger dots indicate children who had longer playsessions and thus for whom there was more data"}
+detectionsByAge <- dlong %>%
+  filter(!is.na(age.at.test)) %>%
+  group_by(subid, age.at.test, detectorType) %>%
+  summarise(detections = mean(detection)) %>%
+  left_join(inc_subs) %>%
+  mutate(len = as.numeric(len.min))
+
+ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
+    geom_jitter(width = .25) +
+    geom_smooth(size=1) +
+    scale_x_continuous(breaks=c(8,12,16)) +
+    scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
+    labs(y = "Prop. detected", x = element_blank()) +
+    theme_bw() + 
+    theme(legend.position="none") +
+    scale_color_manual(values=c("#1F618D", "#7FB3D5", "#C70039", "#FF5733")) +
+    facet_grid(~detectorType, scales = "free_y") 
+```
 
@@
 
-```{r detByAge, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=6, fig.cap = "Proportion face (left) and hand (right) detections as a function of participant's age."  }
-detectionsByAge <- dlong %>%
-  filter(!is.na(age.at.test)) %>%
-  group_by(subid, age.at.test, detectorType) %>%
-  summarise(detections = mean(detection)) %>%
-  left_join(inc_subs) %>%
-  mutate(len = as.numeric(len.min))
-
-p1=ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
-    geom_jitter(width = .25) +
-    geom_smooth(size=1) +
-    scale_x_continuous(breaks=c(8,12,16)) +
-    scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
-    labs(y = "Prop. detected", x = element_blank()) +
-    theme_bw() + 
-    theme(legend.position="none") +
-    scale_color_manual(values=c("#1F618D", "#7FB3D5", "#C70039", "#FF5733")) +
-    facet_grid(~detectorType, scales = "free_y")
+```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=4, fig.cap = "Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
 
@@
   summarise(face = mean(detection)) %>%
-  group_by(age.grp)
+  group_by(age.grp) %%
 
-p2=ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
+p1=ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
   geom_boxplot(outlier.size = 0.5) + 
-  labs(y = "Prop. detected", x = element_blank()) +
+  labs(y = "Prop. faces detected", x = element_blank()) +
   facet_grid(~detectorType, scales="free_y") +
   theme_bw() +
-  theme(legend.position="top", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
 
@@
 
-p3=ggplot(detectionsByOrientation, aes(x = factor(age.grp), y = detections, col = orientation)) + 
+p2=ggplot(detectionsByOrientation, aes(x = factor(age.grp), y = detections, col = orientation)) + 
   geom_boxplot(outlier.size = 0.5) + 
-  labs(y = "Prop. detected", x = element_blank()) +
+  labs(y = "Prop. faces detected", x = element_blank()) +
   facet_grid(~detectorType, scales="free_y") +
   theme_bw() +
-  theme(legend.position="top", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
-  labs(y = "Prop. detected", x = element_blank())
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
+  labs(y = "Prop. faces detected", x = element_blank())
 
-grid.arrange(p1, p2, p3, nrow=3)
+grid.arrange(p1, p1, nrow=2)
 
@@
 
-```{r echo=FALSE, include=FALSE}
-d$faceNum_MT=as.numeric(d$faceMT)
-d$faceNum_OP=as.numeric(d$faceMT)
+<!-- # ```{r echo=FALSE, include=FALSE} -->
+<!-- d$faceNum_MT=as.numeric(d$faceMT) -->
+<!-- d$faceNum_OP=as.numeric(d$faceOP) -->
 
-# Fails to converge
-# fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid),
-#       data = d,
-#       family = "binomial")
+<!-- # Fails to converge -->
+<!-- # fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), -->
+<!-- #       data = d, -->
+<!-- #       family = "binomial") -->
 
-# Fails to converge
-# agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid),
-#       data = d,
-#       family = "binomial")
+<!-- # Fails to converge -->
+<!-- # agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid), -->
+<!-- #       data = d, -->
+<!-- #       family = "binomial") -->
 
-fullModelminsInt_Faces_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + posture + orientation + (1|subid),
-       data = d,
-       family = "binomial")
+<!-- fullModelminsInt_Faces_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + posture + orientation + (1|subid), -->
+<!--        data = d, -->
+<!--        family = "binomial") -->
 
-fullModelminsInt_Faces_OP <- glmer(faceNum_OP ~ scale(age.at.test) + posture + orientation + (1|subid),
-       data = d,
-       family = "binomial")
+<!-- fullModelminsInt_Faces_OP <- glmer(faceNum_OP ~ scale(age.at.test) + posture + orientation + (1|subid), -->
+<!--        data = d, -->
+<!--        family = "binomial") -->
 
-```
+<!-- ``` -->
 
@@
 While age remained a significant predictor even when accounting for the effects of infants' posture and  orientation, it accounted for the least amount of variance.
-Overall, these results suggest that, instead, infants' access to faces is heavily infleunced by their own posture and theirorientation towards their caregiver.
+Overall, these results suggest that, instead, infants' access to faces is heavily infleunced by their own posture and their orientation towards their caregiver.
 
-```{r xtable, results="asis"}
-tab2 <- xtable::xtable(summary(fullModelminsInt_Faces_MTCNN)$coef, digits=3,
-                      caption = "Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.")
+<!-- ```{r MTCNNModel, results="asis"} -->
+<!-- tab2 <- xtable::xtable(summary(fullModelminsInt_Faces_MTCNN)$coef, digits=3, -->
+<!--                       caption = "Results from GLMM model predicting the presence/absence of a face (MTCNN-Faces) across the entire dataset.") -->
 
-print(tab2, type="latex", comment = F, table.placement = "H")
-```
+<!-- print(tab2, type="latex", comment = F, table.placement = "H") -->
+<!-- ``` -->
 
-```{r xtable, results="asis"}
- tab3 <- xtable::xtable(summary(fullModelminsInt_Faces_OP)$coef, digits=3,
-                       caption = "Results from GLMM model prediting the presence/absence of faces (OpenPose-Faces) across the entire dataset.")
+<!-- ```{r OPModel, results="asis"} -->
+<!--  tab3 <- xtable::xtable(summary(fullModelminsInt_Faces_OP)$coef, digits=3, -->
+<!--                        caption = "Results from GLMM model prediting the presence/absence of faces (OpenPose-Faces) across the entire dataset.") -->
 
- print(tab3, type="latex", comment = F, table.placement = "H")
-```
+<!--  print(tab3, type="latex", comment = F, table.placement = "H") -->
+<!-- ``` -->
 
@@
 source("helper.R")
-d$face=d$faceMT
+# d$face=d$faceMT # just choosing one here, but both look the same
+d$face=d$faceOP
+
+# for boxplot
 aroundNamings <- d %>%
   group_by(subid) %>%
-  do(summarize.naming(.))
+  do(summarize.naming(.))   %>%
+  filter(!is.na(posture))
 
-aroundNamings <- aroundNamings %>%
+# for pointrange with CI - could include instead?
+aroundNamings_subAvg <- d %>%
+  group_by(subid) %>%
+  do(summarize.naming(.)) %>%
+  group_by(subid, posture) %>%
+  summarize(propFacesDetected = mean(face)) %>%
+  filter(!is.na(propFacesDetected)) %>%
   filter(!is.na(posture)) %>%
-  filter(!is.na(orientation))
-
-# aroundNamingsModel <- lmer(face ~ scale(age.at.test) + posture + orientation + (1|subid), 
-#          data = aroundNamings)
+  group_by(posture) %>%
+  multi_boot_standard(col = "propFacesDetected")
 
@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ("Look, a Zem"; +/- 2 seconds around each utternace) as a function of infants's posture.
-Each point represents a naming instance for each child in the dataset."  }
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
+Each point represents a naming instance for each child in the dataset."}
+#boxplot
 ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
   geom_boxplot() + 
-  ylab("Prop. face detections") + 
+  ylab("Prop. faces detedcted") + 
   xlab("Posture") + 
   theme_bw()
+
+# ggplot(aroundNamings_subAvg, aes(x = posture, y = mean, col = posture)) +
+#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+#   ylab("Prop. faces detected") +
+#   xlab("Posture") +
+#   theme_bw()
 ```

### 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.tex

@@
 \end{tabular}
-\caption{Detector performance.} 
+\caption{Detector performance on both high density samples (where proportion of faces detected was high) and random samples (where frames were randomly selected).
+P, R, and F denote precision, recall, and F-score, respectively.}
 \end{table}
@@
 infants' access to faces is heavily infleunced by their own posture and
-theirorientation towards their caregiver.
+their orientation towards their caregiver.
 
@@
 \end{tabular}
-\caption{Results from GLMM model prediting the presence/absence of a face (MTCNN-Faces) across the entire dataset.} 
+\caption{Results from GLMM model predicting the presence/absence of a face (MTCNN-Faces) across the entire dataset.} 
 \end{table}
@@
 
-\caption[Proportion face detections around a naming instance ('Look, a Zem']{Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utternace) as a function of infants's posture.
+\caption[Proportion face detections around a naming instance ('Look, a Zem']{Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
 Each point represents a naming instance for each child in the dataset.}\label{fig:detByNaming}

### 2018-01-18 writeup-cogsci-jan14/helper.R
cleanup / organization

@@ -1,179 +1,0 @@
-library(stringr)
-
-######## MERGE IN EXACT TIMES FOR FRAMES #########
-add.times <- function(x) {
-  print(x$subid[1])
-  fname <- paste0("../data/frame_times/",
-                 str_sub(as.character(x$subid[1]),start=4,end=8),
-                 ".csv")
-  if (file.exists(fname)) {
-    times <- read_csv(fname)
-    
-#     x <- merge(x,times, by.x = "frame", by.y = "frame", all.x = TRUE, all.y = FALSE)
-    x <- left_join(x, times)
-    x$dt <- c(diff(x$time),.032) ## assumes data is sorted by frames! also not sure why .032?
-  } else {
-    print(paste0("**** ", x$subid[1], " is missing frames ****"))
-    x$time <- NA
-    x$dt <- NA
-  }
-  
-  return(x)
-}
-
-######## ADD POSTURE CODING #########
-add.posture <- function(x) {
-  fname <- paste("../data/posture/",
-                 x$subid[1],
-                 ".csv",
-                 sep="")
-  
-  if (file.exists(fname)) {
-    print(fname)
-    postures <- read_csv(fname)
-    
-    postures <- regularize.postures(subset(postures,code=="posture"))
-    
-    x$posture <- factor(NA,levels=levels(postures$posture))
-    x$orientation <- factor(NA,levels=levels(postures$orientation))
-    
-    # for each row of p, populate posture and orientation to x
-    for (i in 1:nrow(postures)) {
-      range <- x$time > postures$start[i] & x$time <= postures$end[i]
-      x$posture[range] <- postures$posture[i]
-      x$orientation[range] <- postures$orientation[i]
-    }
-  } else {
-    print(paste0("**** ", x$subid[1], " is missing posture ****"))
-    x$posture <- NA
-    x$orientation <- NA
-  }
-  
-  return(x)
-}
-
-######## GET POSTURE CODING SORTED OUT #########
-regularize.postures <- function (p) {
-  original.postures <- c("p","si","st","l","c","w","NIF")
-  
-  # clean up postures
-  p$type <- factor(p$type,levels=original.postures)
-  p$type <- plyr::revalue(p$type,c("p"="prone",
-                   "si"="sit",
-                   "st"="stand",
-                   "l"="prone",
-                   "c"="carry",
[…]

### 2018-01-22 writeup-cogsci/cogsci2018_faceOnly.Rmd
typo

--- 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
 
-Our final sample consisted of 36 infants and children, with 8 participants in three age groups: 8 months (6 females), 12 months (7 females), and 16 months (6 females).
+Our final sample consisted of 36 infants and children, with 12 participants in three age groups: 8 months (6 females), 12 months (7 females), and 16 months (6 females).
 Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80\% English at home.

### 2018-01-23 writeup-cogsci/cogsci2018_faceOnly.Rmd
add devtools session info call

--- 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.Rmd
@@
 ```{r orientation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each orientation relative to their caregiver." }
-
 ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
@@
 \noindent
+
+```{r session_info, include=FALSE, echo=FALSE}
+devtools::session_info()
+```

### 2018-01-23 writeup-cogsci/cogsci2018_faceOnly.Rmd
fix syntax error in rmd

@@
   summarise(face = mean(detection)) %>%
-  group_by(age.grp) %%
+  group_by(age.grp)
 

### 2018-01-23 writeup-cogsci/cogsci2018_faceOnly.tex
fix syntax error in rmd

--- 2018-01-18 writeup-cogsci-jan14/cogsci2018_faceOnly.tex
@@
 
-Our final sample consisted of 36 infants and children, with 8
+Our final sample consisted of 36 infants and children, with 12
 participants in three age groups: 8 months (6 females), 12 months (7
@@
 
-{\centering \includegraphics{figs/orientation-1}
+{\centering \includegraphics{figs/posture-1}
 
@@
 
-\caption[Proportion time that infants in each age group spent in each orientation relative to their caregiver]{Proportion time that infants in each age group spent in each orientation relative to their caregiver.}\label{fig:orientation}
+\caption[Proportion time that infants in each age group spent in each posture]{Proportion time that infants in each age group spent in each posture.}\label{fig:posture}
 \end{figure}
@@
 
-{\centering \includegraphics{figs/posture-1}
+{\centering \includegraphics{figs/orientation-1}
 
@@
 
-\caption[Proportion time that infants in each age group spent in each posture]{Proportion time that infants in each age group spent in each posture.}\label{fig:posture}
+\caption[Proportion time that infants in each age group spent in each orientation relative to their caregiver]{Proportion time that infants in each age group spent in each orientation relative to their caregiver.}\label{fig:orientation}
 \end{figure}
@@
 \section{Changes in Access to Faces}\label{changes-in-access-to-faces}
+
+\begin{CodeChunk}
+\begin{figure}[h]
+
+{\centering \includegraphics{figs/detByAge-1}
+
+}
+
+\caption[Proportion of faces detected by the MTCNN model (left) and the OpenPose model (right panel) as a function of child's age]{Proportion of faces detected by the MTCNN model (left) and the OpenPose model (right panel) as a function of child's age.
+Larger dots indicate children who had longer playsessions and thus for whom there was more data}\label{fig:detByAge}
+\end{figure}
+\end{CodeChunk}
 
@@
 
-{\centering \includegraphics{figs/detByAge-1}
+{\centering \includegraphics{figs/detByPosOrient-1}
 
@@
 
-\caption[Proportion face (left) and hand (right) detections as a function of participant's age]{Proportion face (left) and hand (right) detections as a function of participant's age.}\label{fig:detByAge}
+\caption[Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant]{Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant.}\label{fig:detByPosOrient}
 \end{figure*}
@@
   \hline
-(Intercept) & -4.847 & 0.064 & -75.303 & 0.000 \\ 
-  scale(age.at.test) & 0.170 & 0.075 & 2.273 & 0.023 \\ 
-  postureprone & -0.513 & 0.042 & -12.326 & 0.000 \\ 
-  posturesit & 0.648 & 0.039 & 16.422 & 0.000 \\ 
-  posturestand & 1.005 & 0.041 & 24.721 & 0.000 \\ 
-  orientationclose & 1.549 & 0.022 & 71.940 & 0.000 \\ 
-  orientationfar & 2.258 & 0.022 & 100.357 & 0.000 \\ 
+(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\
+  Age & 0.170 & 0.061 & 2.766 & 0.006 \\
+  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\
+  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\
+  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\
+  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\
+  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\
    \hline
 \end{tabular}
-\caption{Results from GLMM model predicting the presence/absence of a face (MTCNN-Faces) across the entire dataset.} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-(Intercept) & -4.847 & 0.064 & -75.303 & 0.000 \\ 
-  scale(age.at.test) & 0.170 & 0.075 & 2.273 & 0.023 \\ 
-  postureprone & -0.513 & 0.042 & -12.326 & 0.000 \\ 
-  posturesit & 0.648 & 0.039 & 16.422 & 0.000 \\ 
-  posturestand & 1.005 & 0.041 & 24.721 & 0.000 \\ 
-  orientationclose & 1.549 & 0.022 & 71.940 & 0.000 \\ 
-  orientationfar & 2.258 & 0.022 & 100.357 & 0.000 \\ 
-   \hline
-\end{tabular}
-\caption{Results from GLMM model prediting the presence/absence of faces (OpenPose-Faces) across the entire dataset.} 
+\caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).}
 \end{table}

### 2018-01-23 writeup-cogsci/cogsci2018_faceOnly.Rmd
fix figure 6

--- 2018-01-23 writeup-cogsci/cogsci2018_faceOnly.Rmd
@@
 
-# for boxplot
-aroundNamings <- d %>%
-  group_by(subid) %>%
-  do(summarize.naming(.))   %>%
-  filter(!is.na(posture))
-
 # for pointrange with CI - could include instead?
@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
-Each point represents a naming instance for each child in the dataset."}
-#boxplot
-ggplot(aroundNamings, aes(x = factor(posture), y = face, color=posture)) + 
-  geom_boxplot() + 
-  ylab("Prop. faces detedcted") + 
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.""}
+
+ggplot(aroundNamings_subAvg, aes(x = posture, y = mean, col = posture)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. faces detected") +
   xlab("Posture") +
   theme_bw()
-
-# ggplot(aroundNamings_subAvg, aes(x = posture, y = mean, col = posture)) +
-#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
-#   ylab("Prop. faces detected") +
-#   xlab("Posture") +
-#   theme_bw()
 ```

### 2018-01-23 writeup-cogsci/cogsci2018_faceOnly.Rmd
plots fixed but still squished

@@
 
-grid.arrange(p1, p1, nrow=2)
+grid.arrange(p1, p2, nrow=2)
 
@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.""}
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture."}
 

### 2018-01-23 writeup-cogsci/cogsci2018_faceOnly.tex
plots fixed but still squished

@@
 
-\caption[Proportion face detections around a naming instance ('Look, a Zem']{Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
-Each point represents a naming instance for each child in the dataset.}\label{fig:detByNaming}
+\caption[Proportion face detections around a naming instance ('Look, a Zem']{Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.}\label{fig:detByNaming}
 \end{figure}

### 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
quick changes

@@
 ---
-title: "Postural changes mediate children's visual access to social information"
+title: "Postural developments mediate children's visual access to social information"
 bibliography: xsface.bib
@@
     \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
-        \And {\large \bf Ally Kraus} \\ \texttt{XX@XX} \\ Department of Psychology \\ Stanford University
+    \And {\large \bf Ally Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University
     \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}
@@
 Recent technological developments allow for testing of this hypothesis by documenting the experiences of infants and children from their own perspective.
-By using head-mounted cameras, researchers have begun to document visual experiences of infants and children — which even for walking children are radically different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011; @yoshida2008].
+By using head-mounted cameras, researchers have begun to document visual experiences of infants and children — which even for walking children are radically different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008].
 Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2012].
@@
 We capitalize on recent improvements in face detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces in the child’s visual environment, both overall and relative to naming events by their caregivers.
+We hypothesized that there would be differences in access to social information based on children's posture: that as infant's learned how to crawl, they would see relatively fewer faces as they would be looking primarily at the ground, and that as infant's learned how to walk and locomote independently, that they would experience a much greater portion of faces in their environment.
 
@@
 
-<!-- The second of these algorithms (OpenPose) also captures information about the "skeleton" of people in a given frame, outputting information about the entire set of body parts that were detected.
-This allowed us to extract hand information as well as wrist information.
+The second of these algorithms (OpenPose) also captures information about the "skeleton" of people in a given frame, outputting information about the entire set of body parts that were detected.
 We analyze wrist detections as a good proxy for the presence of a hand.
 We do so first because we did not capture the entire visual field experienced by the infants/children.
-Thus evne though the video may only show the presence of a wrist, the caregiver's hand may be within the child's field of view.
+Thus, even though the video may only show the presence of a wrist, the caregiver's hand may be within the child's field of view.
 Second, we do so because hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
-For example,if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy).
--->
+For example, if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy).
 
@@
 We operationalized face detections as any frames containing a nose, as any half of a face necessarily contains a nose.
-<!-- we operationalized hand detections as any frames containing any of the 21 points detected on either the left or the right hand, and we operationalized wrist detections any frames containing either the left or the right wrist.
--->
+We operationalized wrist detections as any frames containing either the left or the right wrist.
 
@@
   mutate(id = paste(subid,frame)) %>%
-  gather(key = detectorType, value = detection, faceMT, faceOP, faceVJ) %>%
+  gather(key = detectorType, value = detection, faceMT, faceOP, faceVJ, wristOP) %>%
   mutate(detectorType=as.factor(detectorType))
@@
 # for better plotting names
-#levels(dlong$detectorType) 
+levels(dlong$detectorType) 
 #[1] "faceMT"  "faceOP"  "handOP"  "wristOP"

### 2018-01-24 analysis-validation/detector_evaluation.Rmd
wrist eval

@@
   cbind(original_candidates %>%
+          select(sample_type))
+```
+
+OPENPOSE WRISTS
+```{r}
+data <- read_csv("data/final_output/openpose_results_truncated_2.csv") %>%
+  distinct(name, frame, .keep_all = TRUE)
+
+data %<>%
+  mutate(is_face = ifelse(LWrist_conf != 0 | RWrist_conf != 0, "True", "False"))%>%
+  select(group, name, frame, is_face) %>%
+  rename(video = name)
+
+ground_truth_data <- read_csv("data/ground_truth/ground_truth_wrists.csv") %>%
+  arrange(video, frame) %>%
+  cbind(read_csv("data/ground_truth/gold_set_candidates2.csv") %>%
+          arrange(video, frame) %>%
           select(sample_type))

### 2018-01-24 scripts/ground_truth_creation/validate.py

@@ -4,7 +4,7 @@ import ntpath
 
 # TODO change folder name to sample or whatever in ~/xsface
 
-BRIA_ANNOTATIONS = "/Users/alessandro/xsface/sample_hand_images"
+BRIA_ANNOTATIONS = "/Users/alessandro/Downloads/sample_hand_images_bll"
 
 if __name__ == "__main__":
 
@@ -19,7 +19,7 @@ if __name__ == "__main__":
                 group = str(int(video.split("_")[1][:2]))
                 frame = filename.split(".")[0].split("-")[1]
 
-                rows[video + filename.split(".")[0]] = [group, video, frame, "False", 0]
+                rows[video + filename.split(".")[0]] = [group, video, frame, "False"]
 
         elif ntpath.basename(root) == "annotations": # inside annotations
             video = ntpath.basename(os.path.dirname(root))
@@ -30,9 +30,9 @@ if __name__ == "__main__":
 
     print(len(rows))
 
-    with open("ground_truth_hands.csv", 'wb') as csvfile:
+    with open("ground_truth_wrists.csv", 'wb') as csvfile:
         wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
-        wr.writerow(['group', 'video', 'frame', 'is_face', 'angle'])
+        wr.writerow(['group', 'video', 'frame', 'is_face'])
         for key, val in rows.iteritems():
             wr.writerow(val)
 

### 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
plot arranging, etc

@@
 library(ggplot2)
+library(ggthemes)
+library(egg) # plot arranging
 library(lme4)
@@
 
-## evaluate detectors
+## evaluate face detectors
 performance <- ground_truth_data  %>%
@@
 #[1] "faceMT"  "faceOP"  "handOP"  "wristOP"
-  levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces") 
+  levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces","OpenPose-Wrists") 
 ```
@@
   xlab("Age (months)") + 
-  theme_bw()
+  theme_few()
 ```
@@
 detectionsByAge <- dlong %>%
+  filter(detectorType!='MTCNN-Faces') %>%
   filter(!is.na(age.at.test)) %>%
@@
     labs(y = "Prop. detected", x = element_blank()) +
-    theme_bw() + 
+    theme_few() + 
     theme(legend.position="none") +
-    scale_color_manual(values=c("#1F618D", "#7FB3D5", "#C70039", "#FF5733")) +
     facet_grid(~detectorType, scales = "free_y") 
@@
 
-detectionsByPosture <- dlong %>%
+dlongPlot <-dlong %>%
+    filter(detectorType!="OpenPose-Faces")
+
+detectionsByPosture <- dlongPlot %>%
   filter(!is.na(posture)) %>%
@@
   geom_boxplot(outlier.size = 0.5) + 
-  labs(y = "Prop. faces detected", x = element_blank()) +
+  labs(y = "Prop. detected", x = element_blank()) +
   facet_grid(~detectorType, scales="free_y") +
-  theme_bw() +
+  theme_few() +
   theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
 
-detectionsByOrientation <- dlong %>%
+detectionsByOrientation <- dlongPlot %>%
   filter(!is.na(orientation)) %>%
@@
   geom_boxplot(outlier.size = 0.5) + 
-  labs(y = "Prop. faces detected", x = element_blank()) +
+  labs(y = "Prop. detected", x = element_blank()) +
   facet_grid(~detectorType, scales="free_y") +
-  theme_bw() +
+  theme_few() +
   theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
@@
 
-grid.arrange(p1, p2, nrow=2)
+ggarrange(p1, p2)
 
@@
 
-# for pointrange with CI - could include instead?
-aroundNamings_subAvg <- d %>%
+aroundNamings_subAvg_Posture <- d %>%
   group_by(subid) %>%
@@
 
+aroundNamings_subAvg_Orient <- d %>%
+  group_by(subid) %>%
+  do(summarize.naming(.)) %>%
+  group_by(subid, orientation) %>%
+  summarize(propFacesDetected = mean(face)) %>%
+  filter(!is.na(propFacesDetected)) %>%
+  filter(!is.na(orientation)) %>%
+  group_by(orientation) %>%
+  multi_boot_standard(col = "propFacesDetected")
+
 ```
@@
 
-ggplot(aroundNamings_subAvg, aes(x = posture, y = mean, col = posture)) +
+p1=ggplot(aroundNamings_subAvg_Posture, aes(x = posture, y = mean, col = posture)) +
   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
@@
   xlab("Posture") +
-  theme_bw()
+  theme_few() + 
+  theme(legend.position="none")
+
+p2=ggplot(aroundNamings_subAvg_Orient, aes(x = orientation, y = mean, col = orientation)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. faces detected") +
+  xlab("Orientation") +
+  theme_few() +
+  theme(legend.position="none")
+
+ggarrange(p1,p2)
 ```

### 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
new 1000 ground truth

@@
 
-ground_truth_data <- read_csv("../data/ground_truth/ground_truth3.csv") %>%
+ground_truth_data <- read_csv("../data/ground_truth/ground_truth4.csv") %>%
   select(-angle) %>%

### 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
extra text about detectors

@@
 ### Detector evaluation
-In order to evaluate the performance of these detectors, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (as reported by the detectors) and a random sample from the remaining frames.
-This was done so as to not bias our evaluation by the relatively rare appearance of faces in the dataset.
-A face was present in a frame if at least half of the face was showing.
-Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated and are reported in Table 2.
-
-Both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
-Due to the high performance of these two detectors we have included the results of both in the analyses to follow.
-The ViolaJones detector did not perform well on either gold set, so the results from this detector were not included.
 
@@
 
-ground_truth_data <- read_csv("../data/ground_truth/ground_truth4.csv") %>%
   select(-angle) %>%

### 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
adding detector text again?

@@
 ### Detector evaluation
+To evaluate detector performance, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (half reported by the MTCNN, half by OpenPose) and a random sample from the remaining frames; each sample was comprised of frames sampled from each participant (i.e., video).
+This was done so as to not bias our evaluation by the relatively rare appearance of faces in the dataset.
+Faces were classifed as present in a frame if at least half of the face was showing.
+Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated for all detectors and are reported in Table 2.
+
+Both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
+We found a clear tradeoff between recall and precision between the two models.
+That is, MTCNN almost always detected a face if it was present, but made more false alarms; conversely, OpenPose made less false alarms but overall detected fewer faces.
+ViolaJones performed quite poorly, missing many of the faces in the randomly sampled frames.
 
@@
 
+ground_truth_data <- read_csv("../data/ground_truth/ground_truth4.csv") %>%
   select(-angle) %>%

### 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
many new references and rewritten first paragraph

@@
 # Introduction
-Children are remarkably skilled language learners, connecting arbitrary labels (“cup”) with specific visual concepts at a rapid pace through the first two years of life.
-However, children do not learn words in a vacuum but in a rich social environment, where social cues provided by speakers (e.g., eye-gaze) provide strong scaffolding for this learning process.
-Indeed, children’s ability to effectively process these social cues may be a key factor in their early language development.
-For example, in a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies [@carpenter1998].
-More recently, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months [@brooks2005] and throughout the second year of life [@brooks2008].
 
-However, as children are learning their first words, their view of the world is also radically changing [@adolph2007].
+Children are deeply engaged in learning from others around them  [@csibra2009natural; @meltzoff2007like] and attend to the social information in their environment from their earliest days.
+Even newborns tend to prefer to look at faces with direct vs.
+averted gaze [@farroni2002eye] and follow overt gaze shifts by 6 months of age [@gredeback2008microstructure, @gredeback2010development].
+And when infants free-view videos,infants tend to look mostly at faces at the expense of all of the other complex natural scenes -- though older infants also start to look more towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
+
+However, as children are learning from others around them, their view of the world is also radically changing [@adolph2007].
 Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently.
@@
 
-Here, we directly ask whether postural and locomotive developments change the availability of the social information relevant for word learning.
+Here, we directly ask whether postural and locomotive developments change the availability of the social information--the presence of faces and hands.
 To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
@@
 
-<!-- # ```{r echo=FALSE, include=FALSE} -->
-<!-- d$faceNum_MT=as.numeric(d$faceMT) -->
-<!-- d$faceNum_OP=as.numeric(d$faceOP) -->
+```{r echo=FALSE, include=FALSE}
+d$faceNum_MT=as.numeric(d$faceMT)
+d$faceNum_OP=as.numeric(d$faceOP)
 
-<!-- # Fails to converge -->
-<!-- # fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid), -->
-<!-- #       data = d, -->
-<!-- #       family = "binomial") -->
+# Fails to converge
+# fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid),
+#       data = d,
+#       family = "binomial")
 
-<!-- # Fails to converge -->
-<!-- # agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid), -->
-<!-- #       data = d, -->
-<!-- #       family = "binomial") -->
+# Fails to converge
+# agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid),
+#       data = d,
+#       family = "binomial")
 
-<!-- fullModelminsInt_Faces_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + posture + orientation + (1|subid), -->
-<!--        data = d, -->
-<!--        family = "binomial") -->
+fullModelminsInt_Faces_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + posture + orientation + (1|subid),
+       data = d,
+       family = "binomial")
 
-<!-- fullModelminsInt_Faces_OP <- glmer(faceNum_OP ~ scale(age.at.test) + posture + orientation + (1|subid), -->
-<!--        data = d, -->
-<!--        family = "binomial") -->
+ageOnly_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + (1|subid),
+       data = d,
+       family = "binomial")
 
-<!-- ``` -->
+fullModelminsInt_Faces_OP <- glmer(faceNum_OP ~ scale(age.at.test) + posture + orientation + (1|subid),
+       data = d,
+       family = "binomial")
 
-To formalize these observations, we fit a generalized logistic mixed-effect model to all with the presence/absence of a face on every frame as the dependent variable, and participant's age, orientation, and posture as predictors.
+```
+
+To formalize these observations, we fit a generalized logistic mixed-effect model to all with the presence/absence of a face on every frame (as detected by MTCNN), with participant's age, orientation, and posture as predictors.
 Interactions between predictors were not included as this maximal model failed to converge.
-A summary of the coefficients of this model can be found in Table 3 for the MTCNN detections, though we found the same pattern of results for the OpenPose detections.
+A summary of the coefficients of this model can be found in Table 3.
 While age remained a significant predictor even when accounting for the effects of infants' posture and  orientation, it accounted for the least amount of variance.
 Overall, these results suggest that, instead, infants' access to faces is heavily infleunced by their own posture and their orientation towards their caregiver.
+We found the same pattern of results for the OpenPose detections, omitted here for brevity.
 

### 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
intro edits

@@
 averted gaze [@farroni2002eye] and follow overt gaze shifts by 6 months of age [@gredeback2008microstructure, @gredeback2010development].
-And when infants free-view videos,infants tend to look mostly at faces at the expense of all of the other complex natural scenes -- though older infants also start to look more towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
+And when infants free-view videos,they tend to look mostly at faces at the expense of all of complex information--though older infants also start to looktowards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
 

### 2018-01-24 data/ground_truth/merge_ground_truth.Rmd
new detector stuff

---
title: "merge_ground_truth"
output: html_document
---

```{r}
ground_truth_mtcnn <- read_csv("intermediate/ground_truth_mtcnn.csv")
ground_truth_openpose <- read_csv("intermediate/ground_truth_openpose.csv")
candidates_mtcnn <- read_csv("intermediate/candidates_mtcnn.csv")
candidates_openpose <- read_csv("intermediate/candidates_openpose.csv")
```

```{r}
df1 <- ground_truth_mtcnn %>%
  inner_join(candidates_mtcnn %>%
               select(group, video, frame, sample_type), by = c("group", "video", "frame")) %>%
  mutate(detector_sample = "mtcnn")

df2 <- ground_truth_openpose %>%
  inner_join(candidates_openpose %>%
               select(group, video, frame, sample_type), by = c("group", "video", "frame")) %>%
  mutate(detector_sample = "openpose")

write_csv(rbind(df1, df2), "ground_truth_faces.csv")
```

### 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
new detector stuff

--- 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
@@
 ### Detector evaluation
-To evaluate detector performance, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (half reported by the MTCNN, half by OpenPose) and a random sample from the remaining frames; each sample was comprised of frames sampled from each participant (i.e., video).
+In order to evaluate the performance of these detectors, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (as reported by the detectors) and a random sample from the remaining frames.
 This was done so as to not bias our evaluation by the relatively rare appearance of faces in the dataset.
-Faces were classifed as present in a frame if at least half of the face was showing.
-Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated for all detectors and are reported in Table 2.
+A face was present in a frame if at least half of the face was showing.
+Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated and are reported in Table 2.
 
 Both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
-We found a clear tradeoff between recall and precision between the two models.
-That is, MTCNN almost always detected a face if it was present, but made more false alarms; conversely, OpenPose made less false alarms but overall detected fewer faces.
-ViolaJones performed quite poorly, missing many of the faces in the randomly sampled frames.
+Due to the high performance of these two detectors we have included the results of both in the analyses to follow.
+The ViolaJones detector did not perform well on either gold set, so the results from this detector were not included.
 
 ```{r, message=FALSE, warning=FALSE, results='asis'}
+setwd("data")
 d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
@@
 
-ground_truth_data <- read_csv("../data/ground_truth/ground_truth4.csv") %>%
-  select(-angle) %>%
-  arrange(video, frame) %>%
-  cbind(read_csv("../data/ground_truth/gold_set_candidates2.csv") %>%
-          arrange(video, frame) %>%
-          select(sample_type))
+ground_truth_data <- read_csv("data/ground_truth/ground_truth_faces.csv")
 
@@
   mutate(result = evaluate(is_face, detector_is_face)) %>%
-  group_by(detectorType,sample_type) %>%
+  group_by(detectorType, sample_type) %>%
   summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
-  group_by(detectorType,sample_type) %>%
+  group_by(detectorType, sample_type) %>%
   summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))

### 2018-01-24 analysis-main/merge_all_data.R

@@ -20,7 +20,7 @@ detsOpenPose <- detsOpenPose %>%
   mutate(video = name) %>%
   distinct(video,frame,.keep_all=TRUE)  %>%
   mutate(frame = as.numeric(frame) + 1) %>%
-  mutate(faceOP = Nose_conf!=0 & REye_conf!=0 | LEye_conf!=0 )  %>%
+  mutate(faceOP = Nose_conf!=0)  %>%
   mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
 
 # open pose detectors
@@ -28,7 +28,7 @@ detsOpenPose_Hands <- read_csv("../data/final_output/openpose_results_with_detai
 detsOpenPose_HandConf <- detsOpenPose_Hands %>%
   mutate(video = name) %>%
   distinct(video,frame,.keep_all=TRUE)  %>%
-  mutate(frame = as.numeric(frame)) %>%
+  mutate(frame = as.numeric(frame) + 1) %>%
   select(video,frame,ends_with("conf")) %>%
   mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
   mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
@@ -90,5 +90,5 @@ complete_combos <- expand(d, nesting(posture, orientation), subid) %>%
 d <- bind_rows(d, complete_combos)
 
 ## save it out
-write_csv(d, "../data/consolidateddata/consolidated_data_4dets.csv")
+write_csv(d, "../data/consolidateddata/consolidated_data_4dets2.csv")
 

### 2018-01-24 scripts/ground_truth_creation/validate.py

@@ -1,6 +1,7 @@
 import os
 import csv
 import ntpath
+import xml.etree.ElementTree as ET
 
 # TODO change folder name to sample or whatever in ~/xsface
 
@@ -24,9 +25,11 @@ if __name__ == "__main__":
         elif ntpath.basename(root) == "annotations": # inside annotations
             video = ntpath.basename(os.path.dirname(root))
             for filename in filenames:
-                arr = rows[video + filename.split(".")[0]]
-                arr[3] = "True"
-                rows[video + filename.split(".")[0]] = arr
+                xmlroot = ET.parse(os.path.join(root, filename)).getroot()
+                if xmlroot.findall("object"):
+                    arr = rows[video + filename.split(".")[0]]
+                    arr[3] = "True"
+                    rows[video + filename.split(".")[0]] = arr
 
     print(len(rows))
 

### 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd

--- 2018-01-24 writeup-cogsci/cogsci2018_faceOnly.Rmd
@@
 ```{r, message=FALSE, warning=FALSE, results='asis'}
+setwd("data")
 d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%

### 2018-01-25 writeup-cogsci/cogsci2018_faceOnly.Rmd
redoing last night's work on .rmd

@@
 
-Children are deeply engaged in learning from others around them  [@csibra2009natural; @meltzoff2007like] and attend to the social information in their environment from their earliest days.
+Children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like] and attend to the social information in their environment from their earliest days.
 Even newborns tend to prefer to look at faces with direct vs.
-averted gaze [@farroni2002eye] and follow overt gaze shifts by 6 months of age [@gredeback2008microstructure, @gredeback2010development].
-And when infants free-view videos,they tend to look mostly at faces at the expense of all of complex information--though older infants also start to looktowards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
+averted gaze [@farroni2002eye] and follow overt gaze shifts by 6 months of age [@gredeback2008microstructure; @gredeback2010development].
+And when infants free-view videos, they tend to look mostly at faces at the expense of other visual information--though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
 
@@
 On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access social information (e.g., faces, gaze) relative to children who are still crawling and sitting.
-In other words, the ability to walk and move around independently and subsequently access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.
+In other words, the ability to walk and move around independently and subsequently access more detailed social information may accelerate infants’ ability to learn the people in their environments.
 
@@
 
-Here, we directly ask whether postural and locomotive developments change the availability of the social information--the presence of faces and hands.
-To do so, we recorded the visual experience of a group of infants and children using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We capitalize on recent improvements in face detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces in the child’s visual environment, both overall and relative to naming events by their caregivers.
-We hypothesized that there would be differences in access to social information based on children's posture: that as infant's learned how to crawl, they would see relatively fewer faces as they would be looking primarily at the ground, and that as infant's learned how to walk and locomote independently, that they would experience a much greater portion of faces in their environment.
+Here, we directly ask whether postural and locomotive developments change the availability of social information--the presence of faces and hands.
+To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
+We then capitalize on recent improvements in face detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and wrists/hands in the child’s visual environment, both overall and relative to naming events by their caregivers.
+We hypothesized that there would be differential access to social information based on children's postural developments: that as infant's learned how to crawl, they would see relatively fewer faces as they would be looking primarily at the ground, and that as infant's learned how to walk and locomote independently, that they will experience a much greater portion of social information in their environment.
 
@@
 ```{r, message=FALSE, warning=FALSE, results='asis'}
-setwd("data")
-d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
+
+d <- read_csv("../data/consolidateddata/consolidated_data_4detectors.csv") %>%
   mutate(posture = factor(posture), orientation = factor(orientation))
@@
   mutate(detectorType=as.factor(detectorType))
+
+# for better plotting names
+levels(dlong$detectorType) 
+#[1] "faceMT"  "faceOP"  "handOP"  "wristOP"
+levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces","OpenPose-Wrists")
 
@@
 
-ground_truth_data <- read_csv("data/ground_truth/ground_truth_faces.csv")
+ground_truth_faces <- read_csv("data/ground_truth/ground_truth_faces.csv")
+ground_truth_wrists <- read_csv("data/ground_truth/ground_truth_wrists.csv")
 
 ## evaluate face detectors
-performance <- ground_truth_data  %>%
+performanceFaces <- ground_truth_faces  %>%
   mutate(frame=as.numeric(frame), subid=video) %>%
@@
   summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
+  xtable %>%
+  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
+
+## evaluate wrist detectors
+performanceWrists <- ground_truth_wrists  %>%
+  mutate(frame=as.numeric(frame), subid=video) %>%
+  left_join(dlong)   %>%
+  filter(detectorType=="wristOP")
+  mutate(detector_is_wrist = detection, is_wrist = as.logical(is_face)) %>%
+  rowwise() %>%
+  mutate(result = evaluate(is_wrist, detector_is_wrist)) %>%
+  group_by(detectorType, sample_type) %>%
+  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
+  group_by(detectorType, sample_type) %>%
+  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) %>%
+  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
 
@@
 
-# for better plotting names
-levels(dlong$detectorType) 
-#[1] "faceMT"  "faceOP"  "handOP"  "wristOP"
-  levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces","OpenPose-Wrists") 
 ```
@@
 
+## Check for singularity
+fm1=fullModelminsInt_Faces_MTCNN
+diag.vals <- getME(fm1,"theta")[getME(fm1,"lower") == 0]
+any(diag.vals < 1e-6) # FALSE
+
+source(system.file("utils", "allFit.R", package="lme4"))
+fm1.all <- allFit(fm1)
+ss <- summary(fm1.all)
+
 ageOnly_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + (1|subid),
@@
        family = "binomial")
+
+fullModelminsInt_Faces_OP <- glm(faceNum_OP ~ scale(age.at.test) + posture + orientation,
+       data = d,
+       family = "binomial")
+
+## try plain logistic regression
+## think about classification techniques -- ?
 
@@
 source("helper.R")
-# d$face=d$faceMT # just choosing one here, but both look the same
-d$face=d$faceOP
+d$face=d$faceMT # just choosing one here, but both look the same
 
@@
 This work also deploys novel advancements in computer vision to the study of developmental psychology.
-The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms @krizhevsky2012imagenet, creating a new generation of algorithmic tools.
+The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms [@krizhevsky2012imagenet], creating a new generation of algorithmic tools.
 These tools are substantially better equipped to deal with noisier, more complicated datasets and can extract richer and more detailed information.
-Videos from the infant perspective provided substantial challenges (e.g., partially occluded faces) for the classic models of face detection (e.g., ViolaJones, @viola2004robust).
+Videos from the infant perspective provided substantial challenges (e.g., partially occluded faces) for the classic models of face detection (e.g., ViolaJones) [@viola2004robust].
 Further, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to infant learners.

### 2018-01-25 writeup-cogsci/cogsci2018_faceOnly.Rmd
figure redo again

@@
 
-ground_truth_faces <- read_csv("data/ground_truth/ground_truth_faces.csv")
-ground_truth_wrists <- read_csv("data/ground_truth/ground_truth_wrists.csv")
+ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv")
+ground_truth_wrists <- read_csv("../data/ground_truth/ground_truth_wrists.csv")
 
@@
   left_join(dlong)   %>%
-  filter(detectorType=="wristOP")
+  filter(detectorType=="wristOP") %>%
   mutate(detector_is_wrist = detection, is_wrist = as.logical(is_face)) %>%
@@
   summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) %>%
+  xtable %>%
   print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
@@
   \hline
- & Algorithm & Sample\ Type & P & R & F \\ 
+   Algorithm & Sample\ Type & P & R & F \\ 
   \hline
-1 & MTCNN & High\ density & 0.841 & 1.000 & 0.914 \\ 
-  2 & MTCNN & Random & 0.947 & 0.750 & 0.837 \\ 
-  3 & OpenPose & High\ density & 0.995 & 0.863 & 0.924 \\ 
-  4 & OpenPose & Random & 0.710 & 0.917 & 0.800 \\ 
-  5 & ViolaJones & High\ Density & 0.991 & 0.495 & 0.660 \\ 
-  6 & ViolaJones & Random & 0.389 & 0.292 & 0.333 \\ 
+   MTCNN & High\ density & 0.841 & 1.000 & 0.914 \\ 
+   MTCNN & Random & 0.947 & 0.750 & 0.837 \\ 
+  OpenPose & High\ density & 0.995 & 0.863 & 0.924 \\ 
+   OpenPose & Random & 0.710 & 0.917 & 0.800 \\ 
+   ViolaJones & High\ Density & 0.991 & 0.495 & 0.660 \\ 
+   ViolaJones & Random & 0.389 & 0.292 & 0.333 \\ 
    \hline
@@
 dlong <- dlong %>%
+  filter(detectorType!="faceVJ") %>%
   filter(detectorType!="faceVJ")
-
 ```
@@
 
-```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each posture." }
-ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
+```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3.5, fig.height=4, fig.cap = "Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver." }
+p1=ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
   geom_boxplot() + 
   ylab("Proportion Time") + 
-  xlab("Age (months)") + 
-  theme_few()
+    xlab("") + 
+  theme_few() +
+  scale_x_discrete( labels=c("8 m.", "12 m.", "16 m."))
+
+p2=ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
+  geom_boxplot() + 
+  ylab("Proportion Time") + 
+  xlab("") + 
+  theme_few() +
+  scale_x_discrete( labels=c("8 m.", "12 m.", "16 m.")) +
+  theme(axis.ticks = element_blank())
+
+ggarrange(p1,p2)
 ```
@@
 
-```{r orientation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion time that infants in each age group spent in each orientation relative to their caregiver." }
-ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
-  geom_boxplot() + 
-  ylab("Proportion Time") + 
-  xlab("Age (months)") + 
-  theme_few()
-```
-
 # Changes in Access to Faces 
-```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and the OpenPose model (right panel) as a function of child's age.
+```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right panel) as a function of child's age.
 Larger dots indicate children who had longer playsessions and thus for whom there was more data"}
 detectionsByAge <- dlong %>%
-  filter(detectorType!='MTCNN-Faces') %>%
   filter(!is.na(age.at.test)) %>%

### 2018-01-25 writeup-cogsci/cogsci2018_faceOnly.Rmd
detector evaluation table again

@@
 
-```{r, message=FALSE, warning=FALSE, results='asis'}
-
+```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
+### READ IN DETECTIONS
 d <- read_csv("../data/consolidateddata/consolidated_data_4detectors.csv") %>%
@@
 
-evaluate <- function(a, b) {
-  if (a == TRUE) {
-    if (a == b) return ("TP") # was face, detected face
-    else return("FN") # was face, missed face
-  }
-  else {
-    if (a == b) return("TN") # was not face, detected face
-    else return("FP") # was not face, said face
-  }
-}
+### READ IN GROUND TRUTH
+ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv") %>%
+  mutate(sample_type=as.factor(sample_type))
 
-ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv")
-ground_truth_wrists <- read_csv("../data/ground_truth/ground_truth_wrists.csv")
+ground_truth_wrists <- read_csv("../data/ground_truth/ground_truth_wrists.csv") %>%
+  mutate(sample_type=as.factor(sample_type))
 
-## evaluate face detectors
+# Renaming for table output
+levels(ground_truth_faces$sample_type)=c("High density", "Random")
+levels(ground_truth_wrists$sample_type)=c("High density", "Random")
+
+## EVALUATE FACE DETECTORS
 performanceFaces <- ground_truth_faces  %>%
@@
   left_join(dlong)   %>%
+  filter(detectorType!="OpenPose-Wrists") %>%
   mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
@@
   group_by(detectorType, sample_type) %>%
-  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
+  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) %>%
   xtable %>%
@@
 
-## evaluate wrist detectors
+## EVALUATE WRIST DETECTOR
 performanceWrists <- ground_truth_wrists  %>%
@@
   left_join(dlong)   %>%
-  filter(detectorType=="wristOP") %>%
+  filter(detectorType=="OpenPose-Wrists") %>%
   mutate(detector_is_wrist = detection, is_wrist = as.logical(is_face)) %>%
@@
 
+## get rid of ViolaJones detections from data structure for future analyses
+dlong <- dlong %>%
+  filter(detectorType!="faceVJ")
+
 ```
@@
   \hline
-   MTCNN & High\ density & 0.841 & 1.000 & 0.914 \\ 
-   MTCNN & Random & 0.947 & 0.750 & 0.837 \\ 
-  OpenPose & High\ density & 0.995 & 0.863 & 0.924 \\ 
-   OpenPose & Random & 0.710 & 0.917 & 0.800 \\ 
-   ViolaJones & High\ Density & 0.991 & 0.495 & 0.660 \\ 
-   ViolaJones & Random & 0.389 & 0.292 & 0.333 \\ 
+  MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
+  MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
+  OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
+  OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
+  ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
+  ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
+  OpenPose-Wrists & High density & 0.64 & 1.00 & 0.78 \\ 
+  OpenPose-Wrists & Random & 0.76 & 0.26 & 0.39 \\ 
    \hline
@@
 Finally, we explore how these changes impact the accessibility of faces during labeling events.
-
-```{r echo=FALSE, include=FALSE}
-# get rid of violajones
-dlong <- dlong %>%
-  filter(detectorType!="faceVJ") %>%
-  filter(detectorType!="faceVJ")
-```
 
@@
   theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
-  labs(y = "Prop. faces detected", x = element_blank())
 
@@
   ylab("Prop. faces detected") +
-  xlab("Posture") +
+  xlab("") +
   theme_few() + 
@@
   ylab("Prop. faces detected") +
-  xlab("Orientation") +
+  xlab("") +
   theme_few() +

### 2018-01-25 writeup-cogsci/cogsci2018_faceOnly.Rmd
glm/glmer model details

@@
 
+# Function to evaluate detectors
+evaluate <- function(a, b) {
+  if (a == TRUE) {
+    if (a == b) return ("TP") # was face/wrist, detected face/wrist
+    else return("FN") # was face/wrist, missed face/wrist
+  }
+  else {
+    if (a == b) return("TN") # was not face/wrist, detected face/wrist
+    else return("FP") # was not face/wrist, said face/wrist
+  }
+}
+
 ## EVALUATE FACE DETECTORS
@@
 detectionsByAge <- dlong %>%
+  filter(detectorType!="OpenPose-Faces") %>% # for brevity
+  filter(detectorType!="ViolaJones-Faces") %>% # low performance
   filter(!is.na(age.at.test)) %>%
@@
     theme(legend.position="none") +
-    facet_grid(~detectorType, scales = "free_y") 
+    facet_grid(~detectorType) 
 ```
@@
 dlongPlot <-dlong %>%
-    filter(detectorType!="OpenPose-Faces")
+    filter(detectorType!="OpenPose-Faces") %>%
+    filter(detectorType!="ViolaJones-Faces")
 
@@
   theme_few() +
-  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
 
@@
 ```{r echo=FALSE, include=FALSE}
-d$faceNum_MT=as.numeric(d$faceMT)
-d$faceNum_OP=as.numeric(d$faceOP)
+
+## First try a normal logistic regressino, grouping by subject / posture / orientation
+# wrangle data
+dModelGLM <- d %>%
+  group_by(subid, posture, orientation) %>%
+  summarize(countDetected=sum(faceMT==TRUE), countTotal = sum(faceMT==FALSE), age.at.test=age.at.test[1])
+y=cbind(dModelGLM$countDetected, dModelGLM$countTotal) # successes / failures for modeling proportions
+dModelGLM$age.at.test=scale(dModelGLM$age.at.test) # scale and center continous predictor
+# run model
+normalGLM<-glm(y ~ dModelGLM$age.at.test + dModelGLM$posture + dModelGLM$orientation,family=binomial(logit))
+# output results
+normalGLM <- normalGLM %>%
+  xtable %>%
+  print(type="latex", comment = F, table.placement = "H")
+
+# Then fit a mixed effet model
+fullModel <- glmer(faceMT ~ scale(age.at.test)+(posture)+(orientation) + (1|subid),
+       data = d,
+       family = "binomial")
 
@@
 fullModelminsInt_Faces_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + posture + orientation + (1|subid),
+       data = d,
+       family = "binomial")
+
+ <- glm(faceNum_MT ~ scale(age.at.test) + posture + orientation,
        data = d,
@@
 fullModelminsInt_Faces_OP <- glmer(faceNum_OP ~ scale(age.at.test) + posture + orientation + (1|subid),
-       data = d,
-       family = "binomial")
-
-fullModelminsInt_Faces_OP <- glm(faceNum_OP ~ scale(age.at.test) + posture + orientation,
        data = d,

### 2018-01-25 writeup-cogsci/cogsci2018_faceOnly.Rmd
fixed plots and glm details

@@
   group_by(detectorType, sample_type) %>%
-  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) %>%
-  xtable %>%
-  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
+  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
+  # xtable %>%
+  # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
 
@@
 ```{r echo=FALSE, include=FALSE}
+# First try a normal logistic regressino, grouping by subject / posture / orientation
+#wrangle data
+dGLM_faces <- d %>%
+  group_by(subid, posture, orientation) %>%
+  summarize(countDetected_Faces=sum(faceMT==TRUE), countTotal_Frames = sum(faceMT==FALSE), age.at.test=age.at.test[1])
 
-## First try a normal logistic regressino, grouping by subject / posture / orientation
-# wrangle data
-dModelGLM <- d %>%
+dGLM_wrists <- d %>%
   group_by(subid, posture, orientation) %>%
-  summarize(countDetected=sum(faceMT==TRUE), countTotal = sum(faceMT==FALSE), age.at.test=age.at.test[1])
-y=cbind(dModelGLM$countDetected, dModelGLM$countTotal) # successes / failures for modeling proportions
-dModelGLM$age.at.test=scale(dModelGLM$age.at.test) # scale and center continous predictor
-# run model
-normalGLM<-glm(y ~ dModelGLM$age.at.test + dModelGLM$posture + dModelGLM$orientation,family=binomial(logit))
-# output results
-normalGLM <- normalGLM %>%
-  xtable %>%
-  print(type="latex", comment = F, table.placement = "H")
+  summarize(countDetected_Wrists=sum(wristOP==TRUE), countTotal_Frames = sum(wristOP==FALSE),   age.at.test=age.at.test[1])
 
-# Then fit a mixed effet model
-fullModel <- glmer(faceMT ~ scale(age.at.test)+(posture)+(orientation) + (1|subid),
-       data = d,
-       family = "binomial")
+# scale cont predictors
+dGLM_faces$age.at.test=scale(dGLM_faces$age.at.test)
+dGLM_wrists$age.at.test=scale(dGLM_wrists$age.at.test)
 
-# Fails to converge
-# fullModel <- glmer(is_face ~ scale(age.at.test)*(posture) *(orientation) + (1|subid),
-#       data = d,
-#       family = "binomial")
+# successes / failures data structure for glm
+y_faces=cbind(dGLM_faces$countDetected_Faces, dGLM_faces$countTotal_Frames)  
+y_wrists=cbind(dGLM_wrists$countDetected_Wrists, dGLM_wrists$countTotal_Frames)
 
-# Fails to converge
-# agexOrientation <- glmer(faceNum ~ scale(age.at.test) * orientation + (1|subid),
-#       data = d,
-#       family = "binomial")
+# run models
+glm_faces<-glm(y_faces ~ dGLM_faces$age.at.test * dGLM_faces$posture * dGLM_faces$orientation,family=binomial(logit))
+xtable(summary(glm_faces))
 
-fullModelminsInt_Faces_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + posture + orientation + (1|subid),
-       data = d,
-       family = "binomial")
+glm_wrists<-glm(y_wrists ~ dGLM_wrists$age.at.test * dGLM_wrists$posture * dGLM_wrists$orientation,family=binomial(logit))
 
- <- glm(faceNum_MT ~ scale(age.at.test) + posture + orientation,
-       data = d,
-       family = "binomial")
+## OUTPUT GLM RESULTS
+cleanedNames = c("Intercept","Age","Prone","Sit","Stand","Close","Far", "Age:Prone","Age:Sit","Age:Stand","Age:Close",
+          "Age:Far","Prone:Close","Sit:Close","Stand:Close","Prone:Far","Sit:Far","Stand:Far","Age:Prone:Close",
+          "Age:Sit:Close","Age:Prone:Far","Age:Sit:Far")
+faceGLMOut=xtable(summary(glm_faces))
+rownames(faceGLMOut)<-cleanedNames
 
-## Check for singularity
-fm1=fullModelminsInt_Faces_MTCNN
-diag.vals <- getME(fm1,"theta")[getME(fm1,"lower") == 0]
-any(diag.vals < 1e-6) # FALSE
-
-source(system.file("utils", "allFit.R", package="lme4"))
-fm1.all <- allFit(fm1)
-ss <- summary(fm1.all)
-
-ageOnly_MTCNN <- glmer(faceNum_MT ~ scale(age.at.test) + (1|subid),
-       data = d,
-       family = "binomial")
-
-fullModelminsInt_Faces_OP <- glmer(faceNum_OP ~ scale(age.at.test) + posture + orientation + (1|subid),
-       data = d,
-       family = "binomial")
-
-## try plain logistic regression
-## think about classification techniques -- ?
+wristGLMOut=xtable(summary(glm_wrists))
+rownames(wristGLMOut)<-cleanedNames
 
 ```
+
+```{r results="asis"}
+print(faceGLMOut, type="latex", comment = F, table.placement = "H")
+print(wristGLMOut, type="latex", comment = F, table.placement = "H")
+```
+
+<!-- # # output results -->
+<!-- normalGLM <- normalGLM %>% -->
+<!--   xtable %>% -->
+<!--   print(type="latex", comment = F, table.placement = "H") -->
+
+<!-- # ## Then try a run mixed effet model -->
+<!-- # ## Interactions fail to converge, as does the one below -->
+<!-- # fullModelminsInt_Faces_MTCNN <- glmer(faceMT ~ scale(age.at.test)+(posture)+(orientation) + (1|subid), -->
+<!-- #        data = d, -->
+<!-- #        family = "binomial") -->
+
+<!-- ## Reading https://rdrr.io/cran/lme4/man/convergence.html -->
+<!-- ## Check for singularity - not an issue -->
+<!-- # fm1=fullModelminsInt_Faces_MTCNN -->
+<!-- # diag.vals <- getME(fm1,"theta")[getME(fm1,"lower") == 0] -->
+<!-- # any(diag.vals < 1e-6) # FALSE -->
+
+<!-- ## But has issues with every optimizer... -->
+<!-- #source(system.file("utils", "allFit.R", package="lme4")) -->
+<!-- #fm1.all <- allFit(fm1) -->
+<!-- #ss <- summary(fm1.all) -->
+
+<!-- #No convergence issues with open pose detectors -->
+<!-- fullModelminsInt_Faces_OP <- glmer(faceOP ~ scale(age.at.test) + posture + orientation + (1|subid), -->
+<!--        data = d, -->
+<!--        family = "binomial") -->
+
+<!-- fullModelminsInt_Wrist_OP <- glmer(wristOP ~ scale(age.at.test) + posture + orientation + (1|subid), -->
+<!--        data = d, -->
+<!--        family = "binomial") -->
+<!-- ``` -->
 
@@
 <!-- ``` -->
-
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\
-  \hline
-(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\
-  Age & 0.170 & 0.061 & 2.766 & 0.006 \\
-  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\
-  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\
-  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\
-  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\
-  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\
-   \hline
-\end{tabular}
-\caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).}
-\end{table}
 
@@
   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
-  ylab("Prop. faces detected") +
+  ylab("Prop. faces") +
   xlab("") +
+  ylim(0,.3) +
   theme_few() + 
@@
   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
-  ylab("Prop. faces detected") +
+  ylab("Prop. faces") +
   xlab("") +
+  ylim(0,.3) +
   theme_few() +
@@
 
-```{r session_info, include=FALSE, echo=FALSE}
-devtools::session_info()
-```
+# ```{r session_info, include=FALSE, echo=FALSE}
+# devtools::session_info()
+# ```

### 2018-01-25 writeup-cogsci/cogsci2018_faceOnly.Rmd

@@
 
+```
+
+```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
 ### READ IN GROUND TRUTH
@@
 glm_faces<-glm(y_faces ~ dGLM_faces$age.at.test * dGLM_faces$posture * dGLM_faces$orientation,family=binomial(logit))
-xtable(summary(glm_faces))
+
+glm_faces_noInt<-glm(y_faces ~ dGLM_faces$age.at.test + dGLM_faces$posture + dGLM_faces$orientation,family=binomial(logit))
 
@@
 
+glm_wrists_noInt<-glm(y_wrists ~ dGLM_wrists$age.at.test + dGLM_wrists$posture + dGLM_wrists$orientation,family=binomial(logit))
+
 ## OUTPUT GLM RESULTS
+cleanedNames_NoInt = c("Intercept","Age","Prone","Sit","Stand","Close","Far")
 cleanedNames = c("Intercept","Age","Prone","Sit","Stand","Close","Far", "Age:Prone","Age:Sit","Age:Stand","Age:Close",
@@
           "Age:Sit:Close","Age:Prone:Far","Age:Sit:Far")
+
 faceGLMOut=xtable(summary(glm_faces))
@@
 
+wristGLMOut_noInt=xtable(summary(glm_wrists_noInt))
+rownames(wristGLMOut_noInt)<-cleanedNames_NoInt
+
+faceGLMOut_noInt=xtable(summary(glm_faces_noInt))
+rownames(faceGLMOut_noInt)<-cleanedNames_NoInt
+
 ```
@@
 ```{r results="asis"}
-print(faceGLMOut, type="latex", comment = F, table.placement = "H")
-print(wristGLMOut, type="latex", comment = F, table.placement = "H")
+print(faceGLMOut_noInt, type="latex", comment = F, table.placement = "H")
 ```
 
-<!-- # # output results -->
-<!-- normalGLM <- normalGLM %>% -->
-<!--   xtable %>% -->
-<!--   print(type="latex", comment = F, table.placement = "H") -->
+```{r results="asis"}
+print(wristGLMOut_noInt, type="latex", comment = F, table.placement = "H")
+```
 
-<!-- # ## Then try a run mixed effet model -->
-<!-- # ## Interactions fail to converge, as does the one below -->
-<!-- # fullModelminsInt_Faces_MTCNN <- glmer(faceMT ~ scale(age.at.test)+(posture)+(orientation) + (1|subid), -->
-<!-- #        data = d, -->
-<!-- #        family = "binomial") -->
+```{r echo=FALSE, include=FALSE}
+## Mixed effect model tries - many convergence issues
 
-<!-- ## Reading https://rdrr.io/cran/lme4/man/convergence.html -->
-<!-- ## Check for singularity - not an issue -->
-<!-- # fm1=fullModelminsInt_Faces_MTCNN -->
-<!-- # diag.vals <- getME(fm1,"theta")[getME(fm1,"lower") == 0] -->
-<!-- # any(diag.vals < 1e-6) # FALSE -->
+ # Interactions fail to converge, as does the one below 
+ # fullModelminsInt_Faces_MTCNN <- glmer(faceMT ~ scale(age.at.test)+(posture)+(orientation) + (1|subid), 
+ #        data = d, 
+ #        family = "binomial")
 
-<!-- ## But has issues with every optimizer... -->
-<!-- #source(system.file("utils", "allFit.R", package="lme4")) -->
-<!-- #fm1.all <- allFit(fm1) -->
-<!-- #ss <- summary(fm1.all) -->
+## Reading https://rdrr.io/cran/lme4/man/convergence.html 
+## Check for singularity - not an issue 
+# fm1=fullModelminsInt_Faces_MTCNN 
+# diag.vals <- getME(fm1,"theta")[getME(fm1,"lower") == 0] 
+# any(diag.vals < 1e-6) # FALSE
 
-<!-- #No convergence issues with open pose detectors -->
-<!-- fullModelminsInt_Faces_OP <- glmer(faceOP ~ scale(age.at.test) + posture + orientation + (1|subid), -->
-<!--        data = d, -->
-<!--        family = "binomial") -->
+## But has issues with every optimizer... -->
+#source(system.file("utils", "allFit.R", package="lme4")) 
+#fm1.all <- allFit(fm1)
+#ss <- summary(fm1.all)
 
-<!-- fullModelminsInt_Wrist_OP <- glmer(wristOP ~ scale(age.at.test) + posture + orientation + (1|subid), -->
-<!--        data = d, -->
-<!--        family = "binomial") -->
-<!-- ``` -->
+#No convergence issues with open pose detectors for faces, but for wrists
+#fullModelminsInt_Faces_OP <- glmer(faceOP ~ scale(age.at.test) + posture + orientation + (1|subid), 
+#      data = d, 
+#        family = "binomial") 
+```
 
@@
 We found the same pattern of results for the OpenPose detections, omitted here for brevity.
-
-<!-- ```{r MTCNNModel, results="asis"} -->
-<!-- tab2 <- xtable::xtable(summary(fullModelminsInt_Faces_MTCNN)$coef, digits=3, -->
-<!--                       caption = "Results from GLMM model predicting the presence/absence of a face (MTCNN-Faces) across the entire dataset.") -->
-
-<!-- print(tab2, type="latex", comment = F, table.placement = "H") -->
-<!-- ``` -->
-
-<!-- ```{r OPModel, results="asis"} -->
-<!--  tab3 <- xtable::xtable(summary(fullModelminsInt_Faces_OP)$coef, digits=3, -->
-<!--                        caption = "Results from GLMM model prediting the presence/absence of faces (OpenPose-Faces) across the entire dataset.") -->
-
-<!--  print(tab3, type="latex", comment = F, table.placement = "H") -->
-<!-- ``` -->
 
@@
 
-# ```{r session_info, include=FALSE, echo=FALSE}
-# devtools::session_info()
-# ```
+ ```{r session_info, include=FALSE, echo=FALSE}
+ devtools::session_info()
+ ```

### 2018-01-26 writeup-cogsci/cogsci2018_faceOnly.tex
test feather:

@@
 
-\title{Postural changes mediate children's visual access to social information}
+\title{Postural developments mediate children's visual access to social
+information}
 
-\author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
-    \And {\large \bf Ally Kraus} \\ \texttt{XX@XX} \\ Department of Psychology \\ Stanford University
-    \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}
+\author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Ally Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University \And {\large \bf Michael C.
+Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}
 
@@
 
-Children are remarkably skilled language learners, connecting arbitrary
-labels (``cup'') with specific visual concepts at a rapid pace through
-the first two years of life. However, children do not learn words in a
-vacuum but in a rich social environment, where social cues provided by
-speakers (e.g., eye-gaze) provide strong scaffolding for this learning
-process. Indeed, children's ability to effectively process these social
-cues may be a key factor in their early language development. For
-example, in a longitudinal study, children's level of joint engagement
-with their mother was found to predict both their receptive and
-productive vocabularies (Carpenter, Nagell, \& Tomasello, 1998). More
-recently, 10 month-olds who follow an adult's gaze in an experimental
-context have larger vocabularies at 18 months (R. Brooks \& Meltzoff,
-2005) and throughout the second year of life (Rechele Brooks \&
-Meltzoff, 2008).
+Children are deeply engaged in learning from others ({\textbf{???}};
+Csibra \& Gergely, 2009) and attend to the social information in their
+environment from their earliest days. Even newborns tend to prefer to
+look at faces with direct vs.~averted gaze ({\textbf{???}}) and follow
+overt gaze shifts by 6 months of age (Gredebäck, Fikke, \& Melinder,
+2010; Gredebäck, Theuring, Hauf, \& Kenward, 2008). And when infants
+free-view videos, they tend to look mostly at faces at the expense of
+other visual information--though older infants start to look towards
+people's hands and the actions they are performing (Frank, Amso, \&
+Johnson, 2014; Frank, Vul, \& Saxe, 2012).
 
-However, as children are learning their first words, their view of the
-world is also radically changing (K. Adolph \& Berger, 2007). Infants'
-motor abilities improve dramatically near the end of the first year of
-life, allowing them to locomote independently. These motor changes have
-drastic consequences for what children see; crawling and walking infants
-have radically different views of the world. For example, during
-spontaneous play in a laboratory playroom, toddlers are more likely to
-look at the floor while crawling than while walking (J. Franchak,
-Kretch, Soska, \& Adolph, 2011); in general, walking infants tend to
-have full visual access to their environment and the people in it, while
-crawling infants do not (K. S. Kretch, Franchak, \& Adolph, 2014).
+However, as children are learning from others around them, their view of
+the world is also radically changing (K. Adolph \& Berger, 2007).
+Infants' motor abilities improve dramatically near the end of the first
+year of life, allowing them to locomote independently. These motor
+changes have drastic consequences for what children see; crawling and
+walking infants have radically different views of the world. For
+example, during spontaneous play in a laboratory playroom, toddlers are
+more likely to look at the floor while crawling than while walking (J.
+Franchak, Kretch, Soska, \& Adolph, 2011); in general, walking infants
+tend to have full visual access to their environment and the people in
+it, while crawling infants do not (K. S. Kretch, Franchak, \& Adolph,
+2014).
 
@@
 move around independently and subsequently access more detailed social
-information may allow infants to learn words quicker and more
-efficiently, facilitating language growth.
+information may accelerate infants' ability to learn the people in their
+environments.
 
@@
 Here, we directly ask whether postural and locomotive developments
-change the availability of the social information relevant for word
-learning. To do so, we recorded the visual experience of a group of
-infants and children using head-mounted cameras during a brief
-laboratory free-play session; children's posture and orientation
-relative to their caregiver were also recorded from a third-person
-perspective and hand-annotated. We capitalize on recent improvements in
-face detection algorithms (Cao, Simon, Wei, \& Sheikh, 2017; K. Zhang,
-Zhang, Li, \& Qiao, 2016) to analyze the frequencies of faces in the
-child's visual environment, both overall and relative to naming events
-by their caregivers.
+change the availability of social information--the presence of faces and
+hands. To do so, we recorded the visual experience of a group of infants
+in three age ranges (8,12, and 16 months) using head-mounted cameras
+during a brief laboratory free-play session; children's posture and
+orientation relative to their caregiver were also recorded from a
+third-person perspective and hand-annotated. We then capitalize on
+recent improvements in face detection algorithms (Cao, Simon, Wei, \&
+Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
+frequencies of faces and wrists/hands in the child's visual environment,
+both overall and relative to naming events by their caregivers. We
+hypothesized that there would be differential access to social
+information based on children's postural developments: that as infant's
+learned how to crawl, they would see relatively fewer faces as they
+would be looking primarily at the ground, and that as infant's learned
+how to walk and locomote independently, that they will experience a much
+greater portion of social information in their environment.
 
@@
 
+The second of these algorithms (OpenPose) also captures information
+about the ``skeleton'' of people in a given frame, outputting
+information about the entire set of body parts that were detected. We
+analyze wrist detections as a good proxy for the presence of a hand. We
+do so first because we did not capture the entire visual field
+experienced by the infants/children. Thus, even though the video may
+only show the presence of a wrist, the caregiver's hand may be within
+the child's field of view. Second, we do so because hands are often
+occluded by objects when caregivers are interacting with children, yet
+still visually accessible by the child and part of their joint
+interaction. For example, if a caregiver was holding a toy and
+presenting it to the infant, their wrists may be visible but their hand
+may not necessarily be (as it is occluded by the toy).
+
 \subsubsection{Face Detection
@@
 frames containing a nose, as any half of a face necessarily contains a
-nose.\\
+nose. We operationalized wrist detections as any frames containing
+either the left or the right wrist.
 
@@
 
-In order to evaluate the performance of these detectors, we constructed
-a ``gold set'' of frames by hand labeling both a sample of frames with a
-high density of faces (as reported by the detectors) and a random sample
-from the remaining frames. This was done so as to not bias our
-evaluation by the relatively rare appearance of faces in the dataset. A
-face was present in a frame if at least half of the face was showing.
-Precision (hits / hits + false alarms), recall (hits / hits + misses),
-and F-score (harmonic mean of previous measures) were calculated and are
-reported in Table 2.
+To evaluate detector performance, we constructed a ``gold set'' of
+frames by hand labeling both a sample of frames with a high density of
+faces (half reported by the MTCNN, half by OpenPose) and a random sample
+from the remaining frames; each sample was comprised of frames sampled
+from each participant (i.e., video). This was done so as to not bias our
+evaluation by the relatively rare appearance of faces in the dataset.
+Faces were classifed as present in a frame if at least half of the face
+was showing. Precision (hits / hits + false alarms), recall (hits / hits
++ misses), and F-score (harmonic mean of previous measures) were
+calculated for all detectors and are reported in Table 2.
 
@@
 set, with MTCNN outperforming OpenPose on the random sample but trailing
-behind in the high density sample. Due to the high performance of these
-two detectors we have included the results of both in the analyses to
-follow. The ViolaJones detector did not perform well on either gold set,
-so the results from this detector were not included.
+behind in the high density sample. We found a clear tradeoff between
+recall and precision between the two models. That is, MTCNN almost
+always detected a face if it was present, but made more false alarms;
+conversely, OpenPose made less false alarms but overall detected fewer
+faces. ViolaJones performed quite poorly, missing many of the faces in
+the randomly sampled frames.
 
@@
   \hline
- & Algorithm & Sample\ Type & P & R & F \\ 
+   Algorithm & Sample\ Type & P & R & F \\ 
   \hline
-1 & MTCNN & High\ density & 0.841 & 1.000 & 0.914 \\ 
-  2 & MTCNN & Random & 0.947 & 0.750 & 0.837 \\ 
-  3 & OpenPose & High\ density & 0.995 & 0.863 & 0.924 \\ 
-  4 & OpenPose & Random & 0.710 & 0.917 & 0.800 \\ 
-  5 & ViolaJones & High\ Density & 0.991 & 0.495 & 0.660 \\ 
-  6 & ViolaJones & Random & 0.389 & 0.292 & 0.333 \\ 
+  MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
+  MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
+  OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
+  OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
+  ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
+  ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
+  OpenPose-Wrists & High density & 0.64 & 1.00 & 0.78 \\ 
+  OpenPose-Wrists & Random & 0.76 & 0.26 & 0.39 \\ 
    \hline
@@
 
-\caption[Proportion time that infants in each age group spent in each posture]{Proportion time that infants in each age group spent in each posture.}\label{fig:posture}
+\caption[Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver]{Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver.}\label{fig:posture}
 \end{figure}
@@
 
-\begin{CodeChunk}
-\begin{figure}[h]
-
-{\centering \includegraphics{figs/orientation-1}
-
-}
-
-\caption[Proportion time that infants in each age group spent in each orientation relative to their caregiver]{Proportion time that infants in each age group spent in each orientation relative to their caregiver.}\label{fig:orientation}
-\end{figure}
-\end{CodeChunk}
-
 \section{Changes in Access to Faces}\label{changes-in-access-to-faces}
@@
 
-\caption[Proportion of faces detected by the MTCNN model (left) and the OpenPose model (right panel) as a function of child's age]{Proportion of faces detected by the MTCNN model (left) and the OpenPose model (right panel) as a function of child's age.
+\caption[Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right panel) as a function of child's age]{Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right panel) as a function of child's age.
 Larger dots indicate children who had longer playsessions and thus for whom there was more data}\label{fig:detByAge}
@@
 
-To formalize these observations, we fit a generalized logistic
-mixed-effect model to all with the presence/absence of a face on every
-frame as the dependent variable, and participant's age, orientation, and
-posture as predictors. Interactions between predictors were not included
-as this maximal model failed to converge. A summary of the coefficients
-of this model can be found in Table 3 for the MTCNN detections, though
-we found the same pattern of results for the OpenPose detections. While
-age remained a significant predictor even when accounting for the
-effects of infants' posture and orientation, it accounted for the least
-amount of variance. Overall, these results suggest that, instead,
-infants' access to faces is heavily infleunced by their own posture and
-their orientation towards their caregiver.
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+Intercept & -5.2469 & 0.0584 & -89.88 & 0.0000 \\ 
+  Age & 0.0847 & 0.0041 & 20.89 & 0.0000 \\ 
+  Prone & 0.2015 & 0.0564 & 3.57 & 0.0004 \\ 
+  Sit & 1.4053 & 0.0541 & 25.98 & 0.0000 \\ 
+  Stand & 1.4272 & 0.0542 & 26.33 & 0.0000 \\ 
+  Close & 1.8239 & 0.0230 & 79.17 & 0.0000 \\ 
+  Far & 2.5479 & 0.0239 & 106.42 & 0.0000 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
@@
   \hline
-(Intercept) & -4.846 & 0.077 & -62.682 & 0.000 \\
-  Age & 0.170 & 0.061 & 2.766 & 0.006 \\
-  Posture-Prone & -0.513 & 0.036 & -14.045 & 0.000 \\
-  Posture-Sit & 0.648 & 0.034 & 18.853 & 0.000 \\
-  Posture-Stand & 1.005 & 0.035 & 28.402 & 0.000 \\
-  Orient.-Close & 1.549 & 0.021 & 72.757 & 0.000 \\
-  Orient.-Far & 2.258 & 0.022 & 101.138 & 0.000 \\
+Intercept & -5.0818 & 0.0776 & -65.46 & 0.0000 \\ 
+  Age & 0.0564 & 0.0050 & 11.35 & 0.0000 \\ 
+  Prone & 0.9499 & 0.0774 & 12.27 & 0.0000 \\ 
+  Sit & 1.7282 & 0.0758 & 22.79 & 0.0000 \\ 
+  Stand & 1.6618 & 0.0760 & 21.88 & 0.0000 \\ 
+  Close & 0.7472 & 0.0188 & 39.81 & 0.0000 \\ 
+  Far & 1.6357 & 0.0200 & 81.61 & 0.0000 \\ 
    \hline
 \end{tabular}
-\caption{Results from GLMM model predicting the presence/absence of a face across the entire dataset (MTCNN detectors).}
 \end{table}
+
+To formalize these observations, we fit a generalized logistic
+mixed-effect model to all with the presence/absence of a face on every
+frame (as detected by MTCNN), with participant's age, orientation, and
+posture as predictors. Interactions between predictors were not included
+as this maximal model failed to converge. A summary of the coefficients
+of this model can be found in Table 3. While age remained a significant
+predictor even when accounting for the effects of infants' posture and
+orientation, it accounted for the least amount of variance. Overall,
+these results suggest that, instead, infants' access to faces is heavily
+infleunced by their own posture and their orientation towards their
+caregiver. We found the same pattern of results for the OpenPose
+detections, omitted here for brevity.
 
@@
 recognition has advanced dramatically in the past five years since the
-re-birth of deep learning algorithms Krizhevsky, Sutskever, \& Hinton
-(2012), creating a new generation of algorithmic tools. These tools are
+re-birth of deep learning algorithms (Krizhevsky, Sutskever, \& Hinton,
+2012), creating a new generation of algorithmic tools. These tools are
 substantially better equipped to deal with noisier, more complicated
@@
 partially occluded faces) for the classic models of face detection
-(e.g., ViolaJones, Viola \& Jones (2004)). Further, as the headcam
+(e.g., ViolaJones) (Viola \& Jones, 2004). Further, as the headcam
 technologies employed here were inexpensive (\textasciitilde{}\$60 a
@@
 
-\hypertarget{ref-brooks2005}{}
-Brooks, R., \& Meltzoff, A. (2005). The development of gaze following
-and its relation to language. \emph{Developmental Science}, \emph{8}(6),
-535--543.
-
-\hypertarget{ref-brooks2008}{}
-Brooks, R., \& Meltzoff, A. N. (2008). Infant gaze following and
-pointing predict accelerated vocabulary growth through two years of age:
-A longitudinal, growth curve modeling study. \emph{Journal of Child
-Language}, \emph{35}(1), 207--220.
-
 \hypertarget{ref-cao2017realtime}{}
@@
 
-\hypertarget{ref-carpenter1998}{}
-Carpenter, M., Nagell, K., \& Tomasello, M. (1998). Social cognition,
-joint attention, and communicative competence from 9 to 15 months of
-age. \emph{Monographs of the Society for Research in Child Development},
-\emph{63}(4).
-
 \hypertarget{ref-clerkin2017}{}
@@
 \emph{Phil. Trans. R. Soc. B}, \emph{372}(1711), 20160055.
+
+\hypertarget{ref-csibra2009natural}{}
+Csibra, G., \& Gergely, G. (2009). Natural pedagogy. \emph{Trends in
+Cognitive Sciences}, \emph{13}(4), 148--153.
 
@@
 
+\hypertarget{ref-frank2014visual}{}
+Frank, M. C., Amso, D., \& Johnson, S. P. (2014). Visual search and
+attention to faces during early infancy. \emph{Journal of Experimental
+Child Psychology}, \emph{118}, 13--26.
+
 \hypertarget{ref-frank2013}{}
@@
 society} (pp. 454--459).
+
+\hypertarget{ref-frank2012measuring}{}
+Frank, M. C., Vul, E., \& Saxe, R. (2012). Measuring the development of
+social attention using free-viewing. \emph{Infancy}, \emph{17}(4),
+355--375.
+
+\hypertarget{ref-gredeback2010development}{}
+Gredebäck, G., Fikke, L., \& Melinder, A. (2010). The development of
+joint visual attention: A longitudinal study of gaze following during
+interactions with mothers and strangers. \emph{Developmental Science},
+\emph{13}(6), 839--848.
+
+\hypertarget{ref-gredeback2008microstructure}{}
+Gredebäck, G., Theuring, C., Hauf, P., \& Kenward, B. (2008). The
+microstructure of infants' gaze as they view adult shifts in overt
+attention. \emph{Infancy}, \emph{13}(5), 533--543.
 

### 2018-01-26 writeup-cogsci/cogsci2018_faceOnly.Rmd
add feather call

@@
 library(gridExtra)
+library(feather)
 ```
@@
 ### READ IN DETECTIONS
-d <- read_csv("../data/consolidateddata/consolidated_data_4detectors.csv") %>%
+getwd()
+d <- read_feather("../data/consolidated_data_4detectors.feather") %>%
   mutate(posture = factor(posture), orientation = factor(orientation))

### 2018-01-26 writeup-cogsci/helper.R
edit naming function to take other detectors

@@ -156,7 +156,7 @@ summarize.naming <- function (x, window = c(-2,2)) {
     range <- c(max(c(0,t + window[1])),
                min(c(t+window[2],max(x$time,na.rm=T))))
     
-    namings$face[i] <- mean(x$face[x$time > range[1] & x$time < range[2]], na.rm=TRUE)
+    namings$detections[i] <- mean(x$detections[x$time > range[1] & x$time < range[2]], na.rm=TRUE)
     namings$posture[i] <- x$posture[(x$time > namings$time[i])][1]
     namings$orientation[i] <- x$orientation[(x$time > namings$time[i])][1]  
   }

### 2018-01-26 writeup-cogsci/cogsci2018_faceOnly.Rmd
detector output writing, cleanup glm analyses

--- 2018-01-25 writeup-cogsci/cogsci2018_faceOnly.Rmd
@@
 ### Detector evaluation
-To evaluate detector performance, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (half reported by the MTCNN, half by OpenPose) and a random sample from the remaining frames; each sample was comprised of frames sampled from each participant (i.e., video).
+To evaluate face detector performance, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (half reported by the MTCNN, half by OpenPose) and a random sample from the remaining frames; each sample was comprised of frames sampled from each participant (i.e., video).
 This was done so as to not bias our evaluation by the relatively rare appearance of faces in the dataset.
@@
 Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated for all detectors and are reported in Table 2.
-
-Both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
-We found a clear tradeoff between recall and precision between the two models.
-That is, MTCNN almost always detected a face if it was present, but made more false alarms; conversely, OpenPose made less false alarms but overall detected fewer faces.
-ViolaJones performed quite poorly, missing many of the faces in the randomly sampled frames.
+For wrist detections, the "gold set" was constructed in the same way, except that frames with a high density of wrists were taken by wrist detections made by OpenPose.
 
@@
 ```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
-### READ IN GROUND TRUTH
+### READ IN GROUND TRUTH and load function
 ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv") %>%
@@
   rowwise() %>%
-  mutate(result = evaluate(is_face, detector_is_face)) %>%
+  mutate(result = evaluate(is_face, detector_is_face))
+
+perfFacesbySampleType <- performanceFaces %>%
   group_by(detectorType, sample_type) %>%
@@
   # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
+
+perfFacesOverall <- performanceFaces %>%
+  group_by(detectorType) %>%
+  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
+  group_by(detectorType) %>%
+  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
 
@@
   rowwise() %>%
-  mutate(result = evaluate(is_wrist, detector_is_wrist)) %>%
+  mutate(result = evaluate(is_wrist, detector_is_wrist))
+
+perfWristsbySampleType<- performanceWrists %>%
   group_by(detectorType, sample_type) %>%
@@
 
+perfWristsOverall<- performanceWrists %>%
+  group_by(detectorType) %>%
+  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
+  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
+
 ## get rid of ViolaJones detections from data structure for future analyses
@@
 ```
+
+For face detectionm, both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
+We found a clear tradeoff between recall and precision between the two models.
+That is, MTCNN almost always detected a face if it was present, but made more false alarms; conversely, OpenPose made less false alarms but overall detected fewer faces.
+ViolaJones performed quite poorly, missing many of the faces in the randomly sampled frames.
+Overall, MTCNN had a slightly higher f-score (f = `r round(perfFacesOverall$f[1],2)`) than OpenPose (f = `r round(perfFacesOverall$f[2],2)`); for brevity, we thus report analyses only using these detector outputs (though we found the same pattern of results; anlayses code avaliable at https://github.com/amsan7/xs-face).
+
+For wrist detection, OpenPose performed moderately well (f = `r round(perfWristsOverall$f,2)`) with relatively high precision, but low recall on the randomly sampled set of frames (see Table 2).
+We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of wrists overall in the datatset.
 
@@
 ```{r echo=FALSE, include=FALSE}
-# First try a normal logistic regressino, grouping by subject / posture / orientation
+#  Try a normal logistic regressino, grouping by subject / posture / orientation -- model convergance errors for glmer.
+
 #wrangle data
@@
   group_by(subid, posture, orientation) %>%
-  summarize(countDetected_Faces=sum(faceMT==TRUE), countTotal_Frames = sum(faceMT==FALSE), age.at.test=age.at.test[1])
+  summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1])
 
@@
   group_by(subid, posture, orientation) %>%
-  summarize(countDetected_Wrists=sum(wristOP==TRUE), countTotal_Frames = sum(wristOP==FALSE),   age.at.test=age.at.test[1])
+  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE),   age.at.test=age.at.test[1])
 
@@
 # successes / failures data structure for glm
-y_faces=cbind(dGLM_faces$countDetected_Faces, dGLM_faces$countTotal_Frames)  
-y_wrists=cbind(dGLM_wrists$countDetected_Wrists, dGLM_wrists$countTotal_Frames)
+y_faces=cbind(dGLM_faces$countFaces, dGLM_faces$countNotFaces)  
+y_wrists=cbind(dGLM_wrists$countWrists, dGLM_wrists$countNotWrists)
 
@@
 
-faceGLMOut=xtable(summary(glm_faces))
-rownames(faceGLMOut)<-cleanedNames
+faceGLMOut=(summary(glm_faces))
+# rownames(faceGLMOut)<-cleanedNames
 
-wristGLMOut=xtable(summary(glm_wrists))
-rownames(wristGLMOut)<-cleanedNames
+wristGLMOut=(summary(glm_wrists))
+# rownames(wristGLMOut)<-cleanedNames
 
-wristGLMOut_noInt=xtable(summary(glm_wrists_noInt))
+wristGLMOut_noInt=xtable(summary(glm_wrists_noInt), caption="Model coefficients from a generalized linear model predicting the proportion of wrists seen by infants.")
 rownames(wristGLMOut_noInt)<-cleanedNames_NoInt
 
-faceGLMOut_noInt=xtable(summary(glm_faces_noInt))
+faceGLMOut_noInt=xtable(summary(glm_faces_noInt), caption="Model coefficients from a generalized linear model predicting the proportion of wrists seen by infants.")
 rownames(faceGLMOut_noInt)<-cleanedNames_NoInt
@@
 
-To formalize these observations, we fit a generalized logistic mixed-effect model to all with the presence/absence of a face on every frame (as detected by MTCNN), with participant's age, orientation, and posture as predictors.
-Interactions between predictors were not included as this maximal model failed to converge.
-A summary of the coefficients of this model can be found in Table 3.
-While age remained a significant predictor even when accounting for the effects of infants' posture and  orientation, it accounted for the least amount of variance.
-Overall, these results suggest that, instead, infants' access to faces is heavily infleunced by their own posture and their orientation towards their caregiver.
+To formalize these observations, we fit a generalized linear model to the proportion faces seen each infant in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
+A summary of the coefficients of a model with only main effects (and no interactions) model can be found in Table 3.
+When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p =`r round(faceGLMOut$coef[2,4],2)`).
 We found the same pattern of results for the OpenPose detections, omitted here for brevity.
+With respect to wrist/hand detections,  we also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also reduced any main effect of age on the proportion of hands detected ((b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)` , p =`r round(wristGLMOut$coef[2,4],2)`).
+Thus, these results suggest that infants' access to faces is heavily infleunced by their postural and locomotive developments.
 
@@
 source("helper.R")
-d$face=d$faceMT # just choosing one here, but both look the same
 
-aroundNamings_subAvg_Posture <- d %>%
+aroundNamings_Posture_Faces <- d %>%
   group_by(subid) %>%
+  mutate(detections = faceMT) %>%
   do(summarize.naming(.)) %>%
   group_by(subid, posture) %>%
-  summarize(propFacesDetected = mean(face)) %>%
+  summarize(propFacesDetected = mean(detections)) %>%
   filter(!is.na(propFacesDetected)) %>%
@@
 
-aroundNamings_subAvg_Orient <- d %>%
+aroundNamings_Orient_Faces <- d %>%
   group_by(subid) %>%
+  mutate(detections = faceMT) %>%
   do(summarize.naming(.)) %>%
   group_by(subid, orientation) %>%
-  summarize(propFacesDetected = mean(face)) %>%
+  summarize(propFacesDetected = mean(detections)) %>%
   filter(!is.na(propFacesDetected)) %>%
@@
 
+aroundNamings_Posture_Wrists <- d %>%
+  group_by(subid) %>%
+  mutate(detections = wristOP) %>%
+  do(summarize.naming(.)) %>%
+  group_by(subid, posture) %>%
+  summarize(propWristsDetected = mean(detections)) %>%
+  filter(!is.na(propWristsDetected)) %>%
+  filter(!is.na(posture)) %>%
+  group_by(posture) %>%
+  multi_boot_standard(col = "propWristsDetected")
+
+aroundNamings_Orient_Wrists <- d %>%
+  group_by(subid) %>%
+  mutate(detections = wristOP) %>%
+  do(summarize.naming(.)) %>%
+  group_by(subid, orientation) %>%
+  summarize(propWristsDetected = mean(detections)) %>%
+  filter(!is.na(propWristsDetected)) %>%
+  filter(!is.na(orientation)) %>%
+  group_by(orientation) %>%
+  multi_boot_standard(col = "propWristsDetected")
+
 ```
@@
 
-p1=ggplot(aroundNamings_subAvg_Posture, aes(x = posture, y = mean, col = posture)) +
+p1=ggplot(aroundNamings_Posture_Faces, aes(x = posture, y = mean, col = posture)) +
   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
@@
 
-p2=ggplot(aroundNamings_subAvg_Orient, aes(x = orientation, y = mean, col = orientation)) +
+p2=ggplot(aroundNamings_Orient_Faces, aes(x = orientation, y = mean, col = orientation)) +
   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
@@
 
-ggarrange(p1,p2)
+p3=ggplot(aroundNamings_Posture_Wrists, aes(x = posture, y = mean, col = posture)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. wrists") +
+  xlab("") +
+  ylim(0,.3) +
+  theme_few() +
+  theme(legend.position="none")
+
+p4=ggplot(aroundNamings_Orient_Wrists, aes(x = orientation, y = mean, col = orientation)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. wrists") +
+  xlab("") +
+  ylim(0,.3) +
+  theme_few() +
+  theme(legend.position="none")
+
+ggarrange(p1,p2, p3, p4)
 ```
-## Access to Faces During Labeling Events
-Finally, we analyzed how face detections changed during object labeling events as a function of infant's posture and orientation.
+## Access to Faces and Wrists During Labeling Events
+Finally, we analyzed how face and wrists detections changed during object labeling events as a function of infant's posture and orientation.
 Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
-We again found that infant's posture impacted the degree to which they saw their caregiver's face during a labeling event; infants who were sitting or standing were more likely to have access to their caregiver's face.
+We again found that infant's posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information.
 
 # General Discussion
-We use a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information relevant for word-learning, here operationalized as the presence of the faces of their caregiver.
-We found that children's posture and orientation towards their caregiver changed systematically across age, and that all three of these factors dramatically impacted the proportion of faces that were avaliable in the child's visual field.
-Thus, infants' postural development is a mediating factor that explains age-related changes in the proportion of faces avaliable in infants' visual field.
+We use a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
+We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and wrists that were avaliable in the child's visual field.
+Thus, infants' postural and locomotive developments are mediating factors that explain some of the age-related changes in the proportion of faces and hands avaliable in infants' visual field.
 
+Broadly, this work suggests that motoric developments mediate how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
 This work also deploys novel advancements in computer vision to the study of developmental psychology.

### 2018-01-27 scripts/mtcnn_face_detection/align/align_dataset_mtcnn.py
deleted stuff

@@ -1,160 +1,0 @@
-"""Performs face alignment and stores face thumbnails in the output directory."""
-# MIT License
-# 
-# Copyright (c) 2016 David Sandberg
-# 
-# Permission is hereby granted, free of charge, to any person obtaining a copy
-# of this software and associated documentation files (the "Software"), to deal
-# in the Software without restriction, including without limitation the rights
-# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
-# copies of the Software, and to permit persons to whom the Software is
-# furnished to do so, subject to the following conditions:
-# 
-# The above copyright notice and this permission notice shall be included in all
-# copies or substantial portions of the Software.
-# 
-# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
-# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
-# FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
-# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
-# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
-# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
-# SOFTWARE.
-
-from __future__ import absolute_import
-from __future__ import division
-from __future__ import print_function
-
-import os
-import random
-import sys
-from time import sleep
-
-import argparse
-import facenet
-import numpy as np
-import tensorflow as tf
-from scipy import misc
-
-
-def main(args):
-    sleep(random.random())
-    output_dir = os.path.expanduser(args.output_dir)
-    if not os.path.exists(output_dir):
-        os.makedirs(output_dir)
-    # Store some git revision info in a text file in the log directory
-    src_path,_ = os.path.split(os.path.realpath(__file__))
-    facenet.store_revision_info(src_path, output_dir, ' '.join(sys.argv))
-    dataset = facenet.get_dataset(args.input_dir)
-    
-    print('Creating networks and loading parameters')
-    
-    with tf.Graph().as_default():
[…]

### 2018-01-27 scripts/mtcnn_face_detection/wrap2.sh

@@ -1,10 +1,0 @@
-#!/bin/sh
-
-module load tensorflow
-module load opencv/3.0.0
-
-for vid in $PI_HOME/redo_rotate/*; do
-		sbatch -p hns,normal -c 8 -t 10:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
-		    --wrap="python detect_faces.py $vid"
-	sleep 1
-done

### 2018-01-27 scripts/mtcnn_face_detection/wrap_mtcnn_detect_faces.sh

@@ -3,7 +3,7 @@
 module load tensorflow
 module load opencv/3.0.0
 
-for vid in $PI_HOME/frames4/*; do
+for vid in $PI_HOME/frames_rotated/*; do
 	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
 	    echo "Large video (> 10GB): $vid"
 		sbatch -p hns,normal -c 16 -t 20:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \

### 2018-01-27 scripts/openpose/wrap_openpose.sh

@@ -1,14 +1,12 @@
 #!/bin/sh
 
-# TODO add 64GB memory to large videos
-
 module load system
 module load singularity/2.4
 module load cuda/8.0.61
 module load caffe2/0.8.1
 module load cudnn/5.1
 
-mkdir -p $PI_HOME/openpose_output6
+mkdir -p $PI_HOME/openpose_json_output
 
 for vid in $SCRATCH/globus/*/*; do
 	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then

### 2018-01-27 scripts/viola_jones/wrap_viola_jones_detector.sh

@@ -1,5 +1,5 @@
 #!/bin/sh
-for vid in $PI_HOME/frames4/*; do
+for vid in $PI_HOME/frames_rotated/*; do
 	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
 	    echo "Large video (> 10GB): $vid"
 		sbatch -p hns,normal -c 16 -t 10:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \

### 2018-01-27 data/ground_truth/merge_ground_truth.Rmd

@@
 ```
+
+```{r}
+ground_truth_wrists <- read_csv("intermediate/ground_truth_wrists.csv")
+ground_truth_hands <- read_csv("intermediate/ground_truth_hands.csv")
+candidates_hands <- read_csv("intermediate/candidates_hands.csv")
+
+df3 <- ground_truth_wrists %>%
+  inner_join(candidates_hands %>%
+               select(group, video, frame, sample_type), by = c("group", "video", "frame")) %>%
+  mutate(detector_sample = "openpose")
+
+df4 <- ground_truth_hands %>%
+  inner_join(candidates_hands %>%
+               select(group, video, frame, sample_type), by = c("group", "video", "frame")) %>%
+  mutate(detector_sample = "openpose")
+
+write_csv(df3, "ground_truth_wrists.csv")
+write_csv(df4, "ground_truth_hands.csv")
+```

### 2018-01-27 writeup-cogsci/z-old/xsface-firstWriteup.Rmd

--- 2018-01-10 writeup/xsface.Rmd
@@
 During early development, children undergo significant postural and locomotor development, changing from lying and sitting infants to toddlers who walk independently.
-We used a head-mounted camera in conjunction with a face-detection system to explore the eff  ects of these changes on children's visual access to their caregivers' faces during an in-lab play session.
+We used a head-mounted camera in conjunction with a face-detection system to explore the effects of these changes on children's visual access to their caregivers' faces during an in-lab play session.
 In a cross-sectional sample of 8--16 month old children, we found substantial changes in face accessibility based on age and posture.
@@
 library(xtable)
-# library(lubridate)
+library(lubridate)
 ```

### 2018-01-27 writeup-cogsci/cogsci2018_faceOnly.Rmd

@@
 
-To formalize these observations, we fit a generalized linear model to the proportion faces seen each infant in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
+To formalize these observations, we fit a generalized linear model to the proportion faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
 A summary of the coefficients of a model with only main effects (and no interactions) model can be found in Table 3.
 When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p =`r round(faceGLMOut$coef[2,4],2)`).
-We found the same pattern of results for the OpenPose detections, omitted here for brevity.
-With respect to wrist/hand detections,  we also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also reduced any main effect of age on the proportion of hands detected ((b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)` , p =`r round(wristGLMOut$coef[2,4],2)`).
-Thus, these results suggest that infants' access to faces is heavily infleunced by their postural and locomotive developments.
+We found the same pattern of results for the OpenPose face detections, omitted here for brevity.
+We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also reduced any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)` , p =`r round(wristGLMOut$coef[2,4],2)`).
+Thus, these results suggest that infants access to social information is heavily infleunced by their postural and locomotive development.
 
@@
 Thus, infants' postural and locomotive developments are mediating factors that explain some of the age-related changes in the proportion of faces and hands avaliable in infants' visual field.
+Broadly, this work suggests that motoric developments mediate how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
 
-Broadly, this work suggests that motoric developments mediate how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
 This work also deploys novel advancements in computer vision to the study of developmental psychology.

### 2018-01-27 writeup-cogsci/images/frames_with_overlayed_coords/mtcnn_correct/draw_rectangles.py
drawn images

@@ -29,4 +29,4 @@ if __name__ == "__main__":
 
 		cv2.rectangle(image, (x, y), (x+w, y+h), (0, 255, 0), 2)
 
-		cv2.imwrite(ntpath.basename(k).split(".")[0] + "_with_box.jpg")
+		cv2.imwrite(ntpath.basename(k).split(".")[0] + "_with_box.jpg", image)

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
new banner

@@
 \begin{figure*}
-\includegraphics[width=6in]{images/framesample.pdf}
-\caption{\label{fig:frames} Example face detections made by MTCNN for the headcam videos from a child in each group  (green squares).}
+\includegraphics[width=6in]{images/detector_samples_banner.pdf}
+\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
+The last column features a false positive from OpenPose and a false negative from MTCNN.}
 \end{figure*}

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
new banner

@@
 
-To evaluate detector performance, we constructed a ``gold set'' of
+To evaluate face detector performance, we constructed a ``gold set'' of
 frames by hand labeling both a sample of frames with a high density of
@@
 + misses), and F-score (harmonic mean of previous measures) were
-calculated for all detectors and are reported in Table 2.
+calculated for all detectors and are reported in Table 2. For wrist
+detections, the ``gold set'' was constructed in the same way, except
+that frames with a high density of wrists were taken by wrist detections
+made by OpenPose.
 
-Both OpenPose and MTCNN detectors performed relatively well on the gold
-set, with MTCNN outperforming OpenPose on the random sample but trailing
-behind in the high density sample. We found a clear tradeoff between
-recall and precision between the two models. That is, MTCNN almost
-always detected a face if it was present, but made more false alarms;
-conversely, OpenPose made less false alarms but overall detected fewer
-faces. ViolaJones performed quite poorly, missing many of the faces in
-the randomly sampled frames.
+For face detectionm, both OpenPose and MTCNN detectors performed
+relatively well on the gold set, with MTCNN outperforming OpenPose on
+the random sample but trailing behind in the high density sample. We
+found a clear tradeoff between recall and precision between the two
+models. That is, MTCNN almost always detected a face if it was present,
+but made more false alarms; conversely, OpenPose made less false alarms
+but overall detected fewer faces. ViolaJones performed quite poorly,
+missing many of the faces in the randomly sampled frames. Overall, MTCNN
+had a slightly higher f-score (f = 0.89) than OpenPose (f = 0.83); for
+brevity, we thus report analyses only using these detector outputs
+(though we found the same pattern of results; anlayses code avaliable at
+\url{https://github.com/amsan7/xs-face}).
+
+For wrist detection, OpenPose performed moderately well (f = 0.72) with
+relatively high precision, but low recall on the randomly sampled set of
+frames (see Table 2). We thus analyze wrist detections, with the caveat
+that we are likely underestimating the proportion of wrists overall in
+the datatset.
 
@@
 
-To formalize these observations, we fit a generalized logistic
-mixed-effect model to all with the presence/absence of a face on every
-frame (as detected by MTCNN), with participant's age, orientation, and
-posture as predictors. Interactions between predictors were not included
-as this maximal model failed to converge. A summary of the coefficients
-of this model can be found in Table 3. While age remained a significant
-predictor even when accounting for the effects of infants' posture and
-orientation, it accounted for the least amount of variance. Overall,
-these results suggest that, instead, infants' access to faces is heavily
-infleunced by their own posture and their orientation towards their
-caregiver. We found the same pattern of results for the OpenPose
-detections, omitted here for brevity.
+To formalize these observations, we fit a generalized linear model to
+the proportion faces infants saw in each posture and orientation (as
+detected by MTCNN), with participant's age, orientation, and posture as
+independent variables. A summary of the coefficients of a model with
+only main effects (and no interactions) model can be found in Table 3.
+When we did include interaction terms between age, posture, and
+orientation, age no longer remained a significant predictor (b = -15.84,
+SE = 113.98, z = -0.14 , p =0.89). We found the same pattern of results
+for the OpenPose face detections, omitted here for brevity. We also
+found the same pattern with respect to infant's visual access to hands
+(see Table 4); including interaction terms also reduced any main effect
+of age on the proportion of hands detected (b = -15.7, SE = 113.98, z =
+-0.14 , p =0.89). Thus, these results suggest that infants access to
+social information is heavily infleunced by their postural and
+locomotive development.
 
@@
 
-\subsection{Access to Faces During Labeling
-Events}\label{access-to-faces-during-labeling-events}
+\subsection{Access to Faces and Wrists During Labeling
+Events}\label{access-to-faces-and-wrists-during-labeling-events}
 
-Finally, we analyzed how face detections changed during object labeling
-events as a function of infant's posture and orientation. Specfically,
-we analyzed a four-second window around each labeling event (e.g.,
-``Look at the {[}zem{]}!''); these labeling events were hand-annotated
-and automatically synchronized with the frame-by-frame face detections.
-We again found that infant's posture impacted the degree to which they
-saw their caregiver's face during a labeling event; infants who were
-sitting or standing were more likely to have access to their caregiver's
-face.
+Finally, we analyzed how face and wrists detections changed during
+object labeling events as a function of infant's posture and
+orientation. Specfically, we analyzed a four-second window around each
+labeling event (e.g., ``Look at the {[}zem{]}!''); these labeling events
+were hand-annotated and automatically synchronized with the
+frame-by-frame face detections. We again found that infant's posture and
+orientations impacted the degree to which they saw their caregiver's
+face and wrist during a labeling event; infants who were sitting or
+standing were more likely to have access to this social information.
 
@@
 locomotive development directly impacts their access to social
-information relevant for word-learning, here operationalized as the
-presence of the faces of their caregiver. We found that children's
-posture and orientation towards their caregiver changed systematically
-across age, and that all three of these factors dramatically impacted
-the proportion of faces that were avaliable in the child's visual field.
-Thus, infants' postural development is a mediating factor that explains
-age-related changes in the proportion of faces avaliable in infants'
-visual field.
+information, here operationalized as the presence of the faces and
+wrists of their caregiver. We found that children's posture and
+orientation towards their caregiver changed systematically across age,
+and that both of these factors dramatically impacted the proportion of
+faces and wrists that were avaliable in the child's visual field. Thus,
+infants' postural and locomotive developments are mediating factors that
+explain some of the age-related changes in the proportion of faces and
+hands avaliable in infants' visual field. Broadly, this work suggests
+that motoric developments mediate how infants experience their visual
+world and the social information in it: infants that are sitting and
+standing have a different view of their world, the people in it, and the
+actions that are being performed.
 

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
add bibtex / figure refs

@@
 Both 8-month-olds and 12-month-olds spent equivalent amounts of time either lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time either sitting or standing (see Figure \ref{fig:posture}).
-We also observed characteristic changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:orientation}).
+We also observed characteristic changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:posture}).
 

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
add bibtex / figure refs

@@
 
-Children are deeply engaged in learning from others ({\textbf{???}};
-Csibra \& Gergely, 2009) and attend to the social information in their
+Children are deeply engaged in learning from others (Csibra \& Gergely,
+2009; Meltzoff, 2007) and attend to the social information in their
 environment from their earliest days. Even newborns tend to prefer to
-look at faces with direct vs.~averted gaze ({\textbf{???}}) and follow
-overt gaze shifts by 6 months of age (Gredebäck, Fikke, \& Melinder,
-2010; Gredebäck, Theuring, Hauf, \& Kenward, 2008). And when infants
-free-view videos, they tend to look mostly at faces at the expense of
-other visual information--though older infants start to look towards
-people's hands and the actions they are performing (Frank, Amso, \&
-Johnson, 2014; Frank, Vul, \& Saxe, 2012).
+look at faces with direct vs.~averted gaze (Farroni, Csibra, Simion, \&
+Johnson, 2002) and follow overt gaze shifts by 6 months of age
+(Gredebäck, Fikke, \& Melinder, 2010; Gredebäck, Theuring, Hauf, \&
+Kenward, 2008). And when infants free-view videos, they tend to look
+mostly at faces at the expense of other visual information--though older
+infants start to look towards people's hands and the actions they are
+performing (Frank, Amso, \& Johnson, 2014; Frank, Vul, \& Saxe, 2012).
 
@@
 behind them supporting their sitting positions (see Figure
-\ref{fig:orientation}).
+\ref{fig:posture}).
 
@@
 
+\hypertarget{ref-farroni2002eye}{}
+Farroni, T., Csibra, G., Simion, F., \& Johnson, M. H. (2002). Eye
+contact detection in humans from birth. \emph{Proceedings of the
+National Academy of Sciences}, \emph{99}(14), 9602--9605.
+
 \hypertarget{ref-fausey2016}{}
@@
 
+\hypertarget{ref-meltzoff2007like}{}
+Meltzoff, A. N. (2007). ``Like me'': A foundation for social cognition.
+\emph{Developmental Science}, \emph{10}(1), 126--134.
+
 \hypertarget{ref-scheirer2014perceptual}{}

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
paper updates

@@
 How do these postural and locomotive changes affect children's access to the social information relevant for word-learning?
-Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use state-of-the-art computer vision algorithms to detect the proportion of faces in infants' environments.
+Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to detect the proportion of faces in infants' environments.
 We find that infants' posture and orientation to their caregiver largely mediate infants' access to faces, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
-Broadly, we suggest that the combined use of head-mounted cameras and the application of novel deep learning algorithms is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
+Broadly, we suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
 
@@
 `r sum(all_subs$reason=="multiple people")` for having multiple adults present,
-`r sum(all_subs$reason=="no CDI")` for missing CDI data,
+`r sum(all_subs$reason=="no CDI")` for missing Communicative Development Inventory (CDI) data,
 `r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
-`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses.
+`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry.
+All inclusion decisions were made independent of the results of subsequent analyses.
+Some of these data were also used in @frank2013.
 
 ## Head-mounted camera
-```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for three different headcam configurations, with the device we used in the middle.
-The lowest camera is pictured for comparison, but was not available until after our study was already in progress."}
-img <- png::readPNG("images/viewangle.png")
+```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for two different headcam configurations.
+We used the second device in our experiments."}
+img <- png::readPNG("images/viewangle_cut.png")
 grid::grid.raster(img)
@@
 We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband.
-Videos captured by the headcam were 720x480 pixels with 25 frames per second.Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
+Videos captured by the headcam were 720x480 pixels with 25 frames per second.
+Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
 A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left).
@@
 As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see the faces of adults around them.
-This allowed us to maximize our chances of capturing faces that the child would have seen during the play session.
+This choice allowed us to maximize our chances of capturing faces that the child would have seen during the play session.
 
@@
 After the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session--the focus of the current study.
-Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a feather duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the novel toys.
+Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the novel toys.
 
@@
 The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing.
-The caregiver's orientation was characterized as being close to the child, far from the child, and a global category of caregiver behind the child.
+The caregiver's orientation was characterized as being close to the child, far from the child, and behind the child (whether close or far).
 For the first two annotations (close/far from the child), the caregiver could either be to the the front or to the side of the child.
@@
 ## Face Detection
-We used a total of three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame.
-We first measured the performance of the most commonly-used and widely available face detection algorithms (Viola-Jones).
+We used three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame.
+We first measured the performance of the most commonly-used and widely available face detection algorithm, Viola-Jones.
 We used this as a benchmark for performance, and while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
-We next capitalized on recent improvements in computer vision, testing the performance of two state-of-the-art face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information.
+We next capitalized on recent improvements in computer vision, testing the performance of two face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information.
 The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of agent's bodyparts.
@@
 ### Face Detection Algorithms
-The first face detection system made use of a series of Haar feature-based cascade classifiers (Viola-Jones, [@viola2004robust]) applied to each individual frame.
+The first face detection system made use of a series of Haar feature-based cascade classifiers [@viola2004robust] applied to each individual frame.
 This detector provided information about the presence of a face as well as its size and position.
@@
 ### Detector evaluation
-To evaluate face detector performance, we constructed a "gold set" of frames by hand labeling both a sample of frames with a high density of faces (half reported by the MTCNN, half by OpenPose) and a random sample from the remaining frames; each sample was comprised of frames sampled from each participant (i.e., video).
-This was done so as to not bias our evaluation by the relatively rare appearance of faces in the dataset.
+To evaluate face detector performance, we needed to construct a "gold set" of labeled frames that did not carry any biases that might be due to the relatively rare appearance of faces in the dataset.
+We thus hand-labeled two samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
+Each sample was comprised of an equal number of frames taken from each participant (i.e., video).
 Faces were classifed as present in a frame if at least half of the face was showing.
-Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of previous measures) were calculated for all detectors and are reported in Table 2.
+Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 2.
 For wrist detections, the "gold set" was constructed in the same way, except that frames with a high density of wrists were taken by wrist detections made by OpenPose.
@@
 
-For face detectionm, both OpenPose and MTCNN detectors performed relatively well on the gold set, with MTCNN outperforming OpenPose on the random sample but trailing behind in the high density sample.
-We found a clear tradeoff between recall and precision between the two models.
-That is, MTCNN almost always detected a face if it was present, but made more false alarms; conversely, OpenPose made less false alarms but overall detected fewer faces.
+For face detection, MTCNN outperformed OpenPose when looking at the composite F-score (`r round(perfFacesOverall$f[1],2)` vs.
+`r round(perfFacesOverall$f[2],2)`); thus the face presence analyses to follow all result from MTCNN detections.
+We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN had the edge with the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
 ViolaJones performed quite poorly, missing many of the faces in the randomly sampled frames.
-Overall, MTCNN had a slightly higher f-score (f = `r round(perfFacesOverall$f[1],2)`) than OpenPose (f = `r round(perfFacesOverall$f[2],2)`); for brevity, we thus report analyses only using these detector outputs (though we found the same pattern of results; anlayses code avaliable at https://github.com/amsan7/xs-face).
 
@@
   \hline
-  MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
+  MTCNN-Faces & High density & 0.89 & 0.92 & \textbf{0.90} \\ 
   MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
   OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
-  OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
+  OpenPose-Faces & Random & 0.72 & 0.80 & \textbf{0.76} \\ 
   ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
@@
 
-To formalize these observations, we fit a generalized linear model to the proportion faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
+To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
 A summary of the coefficients of a model with only main effects (and no interactions) model can be found in Table 3.
@@
 # General Discussion
-We use a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
+We used a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
 We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and wrists that were avaliable in the child's visual field.

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
paper updates

@@
 infants' (8-16 months of age) egocentric visual perspective and use
-state-of-the-art computer vision algorithms to detect the proportion of
-faces in infants' environments. We find that infants' posture and
-orientation to their caregiver largely mediate infants' access to faces,
-suggesting that these postural and locomotive developments facilitate
-infants' emerging linguistic and social capacities. Broadly, we suggest
-that the combined use of head-mounted cameras and the application of
-novel deep learning algorithms is a promising avenue for understanding
-the statistics of infants' visual and linguistic experience.
+computer vision algorithms to detect the proportion of faces in infants'
+environments. We find that infants' posture and orientation to their
+caregiver largely mediate infants' access to faces, suggesting that
+these postural and locomotive developments facilitate infants' emerging
+linguistic and social capacities. Broadly, we suggest that the combined
+use of head-mounted cameras and the application of new computer vision
+techniques is a promising avenue for understanding the statistics of
+infants' visual and linguistic experience.
 
@@
 minutes of headcam footage, 5 for having multiple adults present, 5 for
-missing CDI data, 2 for missing scene camera footage, 1 for fussiness,
-and one excluded for sample symmetry. All inclusion decisions were made
-independent of the results of subsequent analyses.
+missing Communicative Development Inventory (CDI) data, 2 for missing
+scene camera footage, 1 for fussiness, and one excluded for sample
+symmetry. All inclusion decisions were made independent of the results
+of subsequent analyses. Some of these data were also used in Frank,
+Simmons, Yurovsky, \& Pusiol (2013).
 
@@
 
-\caption[Field of view for three different headcam configurations, with the device we used in the middle]{Field of view for three different headcam configurations, with the device we used in the middle.
-The lowest camera is pictured for comparison, but was not available until after our study was already in progress.}\label{fig:headcam}
+\caption[Field of view for three different headcam configurations, with the device we used in the middle]{Field of view for three different headcam configurations, with the device we used in the middle.}\label{fig:headcam}
 \end{figure}
@@
 the child's upper visual field where the child is likely to see the
-faces of adults around them. This allowed us to maximize our chances of
-capturing faces that the child would have seen during the play session.
+faces of adults around them. This choice allowed us to maximize our
+chances of capturing faces that the child would have seen during the
+play session.
 
@@
 of the current study. Parents were shown a box containing three pairs of
-novel and familiar objects (e.g., a ball and a feather duster, named a
-``zem''), and were instructed to play with the object pairs with their
+novel and familiar objects (e.g., a ball and a microfiber duster, named
+a ``zem''), and were instructed to play with the object pairs with their
 child one at a time, ``as they typically would.'' All parents confirmed
@@
 
-We used a total of three face detection systems to explore infants'
-changing access to social information and to avoid the cost of
-hand-annotating every frame. We first measured the performance of the
-most commonly-used and widely available face detection algorithms
-(Viola-Jones). We used this as a benchmark for performance, and while it
-can achieve impressive accuracy in some situations, it is notoriously
-bad at dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox,
-2014). We next capitalized on recent improvements in computer vision,
-testing the performance of two state-of-the-art face detectors that both
-made use of Convolutional Neural Networks (CNNs) to extract face
-information. The first algorithm was specifically optimized for face
-detection, and the second algorithm was optimized to extract information
-about the position of agent's bodyparts.
+We used three face detection systems to explore infants' changing access
+to social information and to avoid the cost of hand-annotating every
+frame. We first measured the performance of the most commonly-used and
+widely available face detection algorithm, Viola-Jones. We used this as
+a benchmark for performance, and while it can achieve impressive
+accuracy in some situations, it is notoriously bad at dealing with
+occluded faces (Scheirer, Anthony, Nakayama, \& Cox, 2014). We next
+capitalized on recent improvements in computer vision, testing the
+performance of two state-of-the-art face detectors that both made use of
+Convolutional Neural Networks (CNNs) to extract face information. The
+first algorithm was specifically optimized for face detection, and the
+second algorithm was optimized to extract information about the position
+of agent's bodyparts.
 
@@
 The first face detection system made use of a series of Haar
-feature-based cascade classifiers (Viola-Jones, (Viola \& Jones, 2004))
-applied to each individual frame. This detector provided information
-about the presence of a face as well as its size and position.
+feature-based cascade classifiers (Viola \& Jones, 2004) applied to each
+individual frame. This detector provided information about the presence
+of a face as well as its size and position.
 
@@
 
-To evaluate face detector performance, we constructed a ``gold set'' of
-frames by hand labeling both a sample of frames with a high density of
-faces (half reported by the MTCNN, half by OpenPose) and a random sample
-from the remaining frames; each sample was comprised of frames sampled
-from each participant (i.e., video). This was done so as to not bias our
-evaluation by the relatively rare appearance of faces in the dataset.
-Faces were classifed as present in a frame if at least half of the face
-was showing. Precision (hits / hits + false alarms), recall (hits / hits
-+ misses), and F-score (harmonic mean of previous measures) were
-calculated for all detectors and are reported in Table 2. For wrist
-detections, the ``gold set'' was constructed in the same way, except
-that frames with a high density of wrists were taken by wrist detections
-made by OpenPose.
+To evaluate face detector performance, we needed to construct a ``gold
+set'' of labeled frames that did not carry any biases that might be due
+to the relatively rare appearance of faces in the dataset. We thus
+hand-labeled two samples: a sample containing a high density of faces
+(half reported by MTCNN, half by OpenPose) and a random sample from the
+remaining frames. Each sample was comprised of an equal number of frames
+taken from each participant (i.e., video). Faces were classifed as
+present in a frame if at least half of the face was showing. Precision
+(hits / hits + false alarms), recall (hits / hits + misses), and F-score
+(harmonic mean of precision and recall) were calculated for all
+detectors and are reported in Table 2. For wrist detections, the ``gold
+set'' was constructed in the same way, except that frames with a high
+density of wrists were taken by wrist detections made by OpenPose.
 
-For face detectionm, both OpenPose and MTCNN detectors performed
-relatively well on the gold set, with MTCNN outperforming OpenPose on
-the random sample but trailing behind in the high density sample. We
-found a clear tradeoff between recall and precision between the two
-models. That is, MTCNN almost always detected a face if it was present,
-but made more false alarms; conversely, OpenPose made less false alarms
-but overall detected fewer faces. ViolaJones performed quite poorly,
-missing many of the faces in the randomly sampled frames. Overall, MTCNN
-had a slightly higher f-score (f = 0.89) than OpenPose (f = 0.83); for
-brevity, we thus report analyses only using these detector outputs
-(though we found the same pattern of results; anlayses code avaliable at
-\url{https://github.com/amsan7/xs-face}).
+For face detection, MTCNN outperformed OpenPose when looking at the
+composite F-score (0.89 vs.~0.83); thus the face presence analyses to
+follow all result from MTCNN detections. We found that although MTCNN
+and OpenPose perform comparably with the random sample, MTCNN had the
+edge with the high density sample (specifically looking at precision),
+suggesting that OpenPose generated more false positives than MTCNN.
+ViolaJones performed quite poorly, missing many of the faces in the
+randomly sampled frames.
 
@@
 grant to MCF. An earlier version of this work was presented to the
-Cognitive Science Society in Frank, Simmons, Yurovsky, \& Pusiol (2013).
-Please address correspondence to Michael C. Frank, Department of
-Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford,
-CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}.
+Cognitive Science Society in Frank et al. (2013). Please address
+correspondence to Michael C. Frank, Department of Psychology, Stanford
+University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel:
+(650) 724-4003, email: \texttt{mcfrank@stanford.edu}.
 

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
more edits

@@
 ```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right panel) as a function of child's age.
-Larger dots indicate children who had longer playsessions and thus for whom there was more data"}
+Larger dots indicate children who had longer play sessions and thus for whom there was more data"}
 detectionsByAge <- dlong %>%
@@
 ```{r results="asis"}
-print(faceGLMOut_noInt, type="latex", comment = F, table.placement = "H")
+#print(faceGLMOut_noInt, type="latex", comment = F, table.placement = "H")
 ```
 
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+Intercept & -5.2469 & 0.0584 & -89.88 & 0.0000 \\ 
+  Age & 0.0847 & 0.0041 & 20.89 & 0.0000 \\ 
+  Prone & 0.2015 & 0.0564 & 3.57 & 0.0004 \\ 
+  Sit & 1.4053 & 0.0541 & 25.98 & 0.0000 \\ 
+  Stand & 1.4272 & 0.0542 & 26.33 & 0.0000 \\ 
+  Close & 1.8239 & 0.0230 & 79.17 & 0.0000 \\ 
+  Far & 2.5479 & 0.0239 & 106.42 & 0.0000 \\ 
+   \hline
+\end{tabular}
+\caption{Model coefficients from a generalized linear model predicting the proportion of faces seen by infants.} 
+\end{table}
+
 ```{r results="asis"}
-print(wristGLMOut_noInt, type="latex", comment = F, table.placement = "H")
+#print(wristGLMOut_noInt, type="latex", comment = F, table.placement = "H")
 ```
+
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+Intercept & -5.0818 & 0.0776 & -65.46 & 0.0000 \\ 
+  Age & 0.0564 & 0.0050 & 11.35 & 0.0000 \\ 
+  Prone & 0.9499 & 0.0774 & 12.27 & 0.0000 \\ 
+  Sit & 1.7282 & 0.0758 & 22.79 & 0.0000 \\ 
+  Stand & 1.6618 & 0.0760 & 21.88 & 0.0000 \\ 
+  Close & 0.7472 & 0.0188 & 39.81 & 0.0000 \\ 
+  Far & 1.6357 & 0.0200 & 81.61 & 0.0000 \\ 
+   \hline
+\end{tabular}
+\caption{Model coefficients from a generalized linear model predicting the proportion of wrists seen by infants.} 
+\end{table}
 
@@
 Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
-We again found that infant's posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information.
+We again found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information.
 

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
more edits

@@
 
-\caption[Field of view for three different headcam configurations, with the device we used in the middle]{Field of view for three different headcam configurations, with the device we used in the middle.}\label{fig:headcam}
+\caption[Field of view for two different headcam configurations]{Field of view for two different headcam configurations. We used the second device in our experiments.}\label{fig:headcam}
 \end{figure}
@@
 from a MD80 model camera attached to a soft elastic headband. Videos
-captured by the headcam were 720x480 pixels with 25 frames per
-second.Detailed instructions for creating this headcam can be found at
+captured by the headcam were 720x480 pixels with 25 frames per second.
+Detailed instructions for creating this headcam can be found at
 \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
@@
 or standing. The caregiver's orientation was characterized as being
-close to the child, far from the child, and a global category of
-caregiver behind the child. For the first two annotations (close/far
-from the child), the caregiver could either be to the the front or to
-the side of the child. All annotations were made by a trained coder
-using the OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman,
-Sanderson, \& Millman, 2012), and times when the child was out of view
-of the tripod camera were marked as uncodable and were excluded from
-these annotations.
+close to the child, far from the child, and behind the child (whether
+close or far). For the first two annotations (close/far from the child),
+the caregiver could either be to the the front or to the side of the
+child. All annotations were made by a trained coder using the
+OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman, Sanderson, \&
+Millman, 2012), and times when the child was out of view of the tripod
+camera were marked as uncodable and were excluded from these
+annotations.
 
@@
 capitalized on recent improvements in computer vision, testing the
-performance of two state-of-the-art face detectors that both made use of
-Convolutional Neural Networks (CNNs) to extract face information. The
-first algorithm was specifically optimized for face detection, and the
-second algorithm was optimized to extract information about the position
-of agent's bodyparts.
+performance of two face detectors that both made use of Convolutional
+Neural Networks (CNNs) to extract face information. The first algorithm
+was specifically optimized for face detection, and the second algorithm
+was optimized to extract information about the position of agent's
+bodyparts.
 
@@
 \end{tabular}
+\caption{Model coefficients from a generalized linear model predicting the proportion of wrists seen by infants.} 
 \end{table}
@@
 To formalize these observations, we fit a generalized linear model to
-the proportion faces infants saw in each posture and orientation (as
+the proportion of faces infants saw in each posture and orientation (as
 detected by MTCNN), with participant's age, orientation, and posture as
@@
 
-We use a head-mounted camera to explore how children's postural and
+We used a head-mounted camera to explore how children's postural and
 locomotive development directly impacts their access to social

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
more edits

@@
 How do these postural and locomotive changes affect children's access to the social information relevant for word-learning?
-Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to detect the proportion of faces in infants' environments.
-We find that infants' posture and orientation to their caregiver largely mediate infants' access to faces, suggesting that these postural and locomotive developments facilitate infants' emerging linguistic and social capacities.
+Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to detect the proportion of faces and hands in infants' environments.
+We find that infants' posture and orientation to their caregiver largely mediate infants' access to socially informative features, suggesting that postural and locomotive development play a significant role in mediating children's emerging linguistic and social capacities.
 Broadly, we suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
@@
 On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access social information (e.g., faces, gaze) relative to children who are still crawling and sitting.
-In other words, the ability to walk and move around independently and subsequently access more detailed social information may accelerate infants’ ability to learn the people in their environments.
+In other words, the ability to walk and move around independently and subsequently access more detailed social information may accelerate infants’ ability to learn from the people in their environments.
 
@@
 To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We then capitalize on recent improvements in face detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and wrists/hands in the child’s visual environment, both overall and relative to naming events by their caregivers.
-We hypothesized that there would be differential access to social information based on children's postural developments: that as infant's learned how to crawl, they would see relatively fewer faces as they would be looking primarily at the ground, and that as infant's learned how to walk and locomote independently, that they will experience a much greater portion of social information in their environment.
+We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (operationalized by the presence of wrists) in the child’s visual environment, both overall and relative to naming events by their caregivers.
+We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, and walking toddlers would have access to a richer visual landscape that would contain a larger portion of the social information in the environment.
 
@@
 ## Head-mounted camera
-```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for two different headcam configurations.
-We used the second device in our experiments."}
+```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for two different headcam configurations (we used the one below in our current study)."}
 img <- png::readPNG("images/viewangle_cut.png")
@@
 The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing.
-The caregiver's orientation was characterized as being close to the child, far from the child, and behind the child (whether close or far).
+The caregiver's orientation was characterized as being close to the child, far from the child, and behind the child (independent of distance).
 For the first two annotations (close/far from the child), the caregiver could either be to the the front or to the side of the child.
@@
 
-## Face Detection
+## Face and Wrist Detection
 We used three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame.
-We first measured the performance of the most commonly-used and widely available face detection algorithm, Viola-Jones.
+The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones.
 We used this as a benchmark for performance, and while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
 We next capitalized on recent improvements in computer vision, testing the performance of two face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information.
-The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of agent's bodyparts.
+The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts belonging to an agent.
 
-The second of these algorithms (OpenPose) also captures information about the "skeleton" of people in a given frame, outputting information about the entire set of body parts that were detected.
-We analyze wrist detections as a good proxy for the presence of a hand.
+This second of the CNN-based algorithms - the pose detector - provided us with the location of an agent's wrists, which we used as a good proxy for the presence of a hand.
 We do so first because we did not capture the entire visual field experienced by the infants/children.
@@
 
-### Face Detection Algorithms
+### Algorithms
 The first face detection system made use of a series of Haar feature-based cascade classifiers [@viola2004robust] applied to each individual frame.
@@
 The candidates are then matched to a single individual and finally assembled into a pose.
-For the purposes of our investigation we only made use of the body parts relevant to the face (ears, eyes, nose).
+For the purposes of our investigation we only made use of the body parts relevant to the face and hands (nose and wrists).
 We operationalized face detections as any frames containing a nose, as any half of a face necessarily contains a nose.
@@
 ### Detector evaluation
-To evaluate face detector performance, we needed to construct a "gold set" of labeled frames that did not carry any biases that might be due to the relatively rare appearance of faces in the dataset.
+To evaluate face detector performance, we needed to construct an unbiased "gold set" of labeled frames that made sure to account for the relatively rare appearance of faces in the dataset.
 We thus hand-labeled two samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
@@
 Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 2.
-For wrist detections, the "gold set" was constructed in the same way, except that frames with a high density of wrists were taken by wrist detections made by OpenPose.
+For wrist detections, the "gold set" was constructed in the same manner, with the exception that frames with a high density of wrists were taken by wrist detections made by OpenPose.
 
@@
 `r round(perfFacesOverall$f[2],2)`); thus the face presence analyses to follow all result from MTCNN detections.
-We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN had the edge with the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
-ViolaJones performed quite poorly, missing many of the faces in the randomly sampled frames.
+We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN performs better on the high density sample evaluation (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
+ViolaJones performed quite poorly relative to the other detectors, finding much less faces in the random sample.
 
@@
 \caption{Detector performance on both high density samples (where proportion of faces detected was high) and random samples (where frames were randomly selected).
-P, R, and F denote precision, recall, and F-score, respectively.}
+P, R, and F denote precision, recall, and F-score, respectively.
+Scores in bold are the highest F-scores for each sample type.}
 \end{table}
@@
 First, we report developmental shifts in infants' posture and their orientation relative to their caregiver.
-Then, we explore how these changes influence children's visual access to faces across this developmental time range.
-Finally, we explore how these changes impact the accessibility of faces during labeling events.
+Then, we explore how these changes influence children's visual access to faces and hands across this developmental time range.
+Finally, we explore how these changes impact the accessibility of faces and hands during labeling events.
 
@@
 
-# Changes in Access to Faces 
-```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right panel) as a function of child's age.
-Larger dots indicate children who had longer play sessions and thus for whom there was more data"}
+# Changes in Access to Faces and Hands
+```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age.
+Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
 detectionsByAge <- dlong %>%
@@
 
-We first examined the proportion of face detections across age; a full summary can be seen in Figure \ref{fig:detByAge}.
-We observed a slight U-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slighly fewer faces faces than 8 or 16-month-olds.
+We first examined the proportion of face and wrist detections across age; a full summary can be seen in Figure \ref{fig:detByAge}.
+When looking at face presence, we observed a slight U-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slighly fewer faces faces than 8 or 16-month-olds.
 
@@
 We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also reduced any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)` , p =`r round(wristGLMOut$coef[2,4],2)`).
-Thus, these results suggest that infants access to social information is heavily infleunced by their postural and locomotive development.
+Thus, these results suggest that infants' access to social information is heavily influenced by their postural and locomotive development.
 

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
more edits

@@
 infants' (8-16 months of age) egocentric visual perspective and use
-computer vision algorithms to detect the proportion of faces in infants'
-environments. We find that infants' posture and orientation to their
-caregiver largely mediate infants' access to faces, suggesting that
-these postural and locomotive developments facilitate infants' emerging
+computer vision algorithms to detect the proportion of faces and hands
+in infants' environments. We find that infants' posture and orientation
+to their caregiver largely mediate infants' access to socially
+informative features, suggesting that postural and locomotive
+development play a significant role in mediating children's emerging
 linguistic and social capacities. Broadly, we suggest that the combined
@@
 move around independently and subsequently access more detailed social
-information may accelerate infants' ability to learn the people in their
-environments.
+information may accelerate infants' ability to learn from the people in
+their environments.
 
@@
 third-person perspective and hand-annotated. We then capitalize on
-recent improvements in face detection algorithms (Cao, Simon, Wei, \&
-Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
-frequencies of faces and wrists/hands in the child's visual environment,
-both overall and relative to naming events by their caregivers. We
-hypothesized that there would be differential access to social
-information based on children's postural developments: that as infant's
-learned how to crawl, they would see relatively fewer faces as they
-would be looking primarily at the ground, and that as infant's learned
-how to walk and locomote independently, that they will experience a much
-greater portion of social information in their environment.
+recent improvements in face and pose detection algorithms (Cao, Simon,
+Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
+frequencies of faces and hands (operationalized by the presence of
+wrists) in the child's visual environment, both overall and relative to
+naming events by their caregivers. We hypothesized that there would be
+differential access to social information based on children's postural
+developments: crawling infants would see fewer faces because they would
+primarily be looking at the ground, and walking toddlers would have
+access to a richer visual landscape that would contain a larger portion
+of the social information in the environment.
 
@@
 
-\caption[Field of view for two different headcam configurations]{Field of view for two different headcam configurations. We used the second device in our experiments.}\label{fig:headcam}
+\caption[Field of view for two different headcam configurations (we used the one below in our current study)]{Field of view for two different headcam configurations (we used the one below in our current study).}\label{fig:headcam}
 \end{figure}
@@
 or standing. The caregiver's orientation was characterized as being
-close to the child, far from the child, and behind the child (whether
-close or far). For the first two annotations (close/far from the child),
-the caregiver could either be to the the front or to the side of the
-child. All annotations were made by a trained coder using the
-OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman, Sanderson, \&
-Millman, 2012), and times when the child was out of view of the tripod
-camera were marked as uncodable and were excluded from these
+close to the child, far from the child, and behind the child
+(independent of distance). For the first two annotations (close/far from
+the child), the caregiver could either be to the the front or to the
+side of the child. All annotations were made by a trained coder using
+the OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman, Sanderson,
+\& Millman, 2012), and times when the child was out of view of the
+tripod camera were marked as uncodable and were excluded from these
 annotations.
 
-\subsection{Face Detection}\label{face-detection}
+\subsection{Face and Wrist Detection}\label{face-and-wrist-detection}
 
@@
 to social information and to avoid the cost of hand-annotating every
-frame. We first measured the performance of the most commonly-used and
-widely available face detection algorithm, Viola-Jones. We used this as
-a benchmark for performance, and while it can achieve impressive
-accuracy in some situations, it is notoriously bad at dealing with
-occluded faces (Scheirer, Anthony, Nakayama, \& Cox, 2014). We next
-capitalized on recent improvements in computer vision, testing the
-performance of two face detectors that both made use of Convolutional
-Neural Networks (CNNs) to extract face information. The first algorithm
-was specifically optimized for face detection, and the second algorithm
-was optimized to extract information about the position of agent's
-bodyparts.
+frame. The first of these is the most commonly-used and widely available
+face detection algorithm: Viola-Jones. We used this as a benchmark for
+performance, and while it can achieve impressive accuracy in some
+situations, it is notoriously bad at dealing with occluded faces
+(Scheirer, Anthony, Nakayama, \& Cox, 2014). We next capitalized on
+recent improvements in computer vision, testing the performance of two
+face detectors that both made use of Convolutional Neural Networks
+(CNNs) to extract face information. The first algorithm was specifically
+optimized for face detection, and the second algorithm was optimized to
+extract information about the position of 18 different body parts
+belonging to an agent.
 
-The second of these algorithms (OpenPose) also captures information
-about the ``skeleton'' of people in a given frame, outputting
-information about the entire set of body parts that were detected. We
-analyze wrist detections as a good proxy for the presence of a hand. We
-do so first because we did not capture the entire visual field
-experienced by the infants/children. Thus, even though the video may
-only show the presence of a wrist, the caregiver's hand may be within
-the child's field of view. Second, we do so because hands are often
-occluded by objects when caregivers are interacting with children, yet
-still visually accessible by the child and part of their joint
-interaction. For example, if a caregiver was holding a toy and
+This second of the CNN-based algorithms - the pose detector - provided
+us with the location of an agent's wrists, which we used as a good proxy
+for the presence of a hand. We do so first because we did not capture
+the entire visual field experienced by the infants/children. Thus, even
+though the video may only show the presence of a wrist, the caregiver's
+hand may be within the child's field of view. Second, we do so because
+hands are often occluded by objects when caregivers are interacting with
+children, yet still visually accessible by the child and part of their
+joint interaction. For example, if a caregiver was holding a toy and
 presenting it to the infant, their wrists may be visible but their hand
@@
 
-\subsubsection{Face Detection
-Algorithms}\label{face-detection-algorithms}
+\subsubsection{Algorithms}\label{algorithms}
 
@@
 our investigation we only made use of the body parts relevant to the
-face (ears, eyes, nose). We operationalized face detections as any
-frames containing a nose, as any half of a face necessarily contains a
-nose. We operationalized wrist detections as any frames containing
+face and hands (nose and wrists). We operationalized face detections as
+any frames containing a nose, as any half of a face necessarily contains
+a nose. We operationalized wrist detections as any frames containing
 either the left or the right wrist.
@@
 
-To evaluate face detector performance, we needed to construct a ``gold
-set'' of labeled frames that did not carry any biases that might be due
-to the relatively rare appearance of faces in the dataset. We thus
+To evaluate face detector performance, we needed to construct an
+unbiased ``gold set'' of labeled frames that made sure to account for
+the relatively rare appearance of faces in the dataset. We thus
 hand-labeled two samples: a sample containing a high density of faces
@@
 detectors and are reported in Table 2. For wrist detections, the ``gold
-set'' was constructed in the same way, except that frames with a high
-density of wrists were taken by wrist detections made by OpenPose.
+set'' was constructed in the same manner, with the exception that frames
+with a high density of wrists were taken by wrist detections made by
+OpenPose.
 
@@
 follow all result from MTCNN detections. We found that although MTCNN
-and OpenPose perform comparably with the random sample, MTCNN had the
-edge with the high density sample (specifically looking at precision),
-suggesting that OpenPose generated more false positives than MTCNN.
-ViolaJones performed quite poorly, missing many of the faces in the
-randomly sampled frames.
+and OpenPose perform comparably with the random sample, MTCNN performs
+better on the high density sample evaluation (specifically looking at
+precision), suggesting that OpenPose generated more false positives than
+MTCNN. ViolaJones performed quite poorly relative to the other
+detectors, finding much less faces in the random sample.
 
@@
 orientation relative to their caregiver. Then, we explore how these
-changes influence children's visual access to faces across this
-developmental time range. Finally, we explore how these changes impact
-the accessibility of faces during labeling events.
+changes influence children's visual access to faces and hands across
+this developmental time range. Finally, we explore how these changes
+impact the accessibility of faces and hands during labeling events.
 
@@
 
-\section{Changes in Access to Faces}\label{changes-in-access-to-faces}
+\section{Changes in Access to Faces and
+Hands}\label{changes-in-access-to-faces-and-hands}
 
@@
 
-\caption[Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right panel) as a function of child's age]{Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right panel) as a function of child's age.
-Larger dots indicate children who had longer playsessions and thus for whom there was more data}\label{fig:detByAge}
+\caption[Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age]{Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age.
+Larger dots indicate children who had longer play sessions and thus for whom there was more data.}\label{fig:detByAge}
 \end{figure}
@@
 
-We first examined the proportion of face detections across age; a full
-summary can be seen in Figure \ref{fig:detByAge}. We observed a slight
-U-shaped function both when analyzing the output of the MTCNN and
-OpenPose detectors, such that 12-month-olds appeared to experience
-slighly fewer faces faces than 8 or 16-month-olds.
+We first examined the proportion of face and wrist detections across
+age; a full summary can be seen in Figure \ref{fig:detByAge}. When
+looking at face presence, we observed a slight U-shaped function both
+when analyzing the output of the MTCNN and OpenPose detectors, such that
+12-month-olds appeared to experience slighly fewer faces faces than 8 or
+16-month-olds.
 
@@
 \end{tabular}
-\caption{Model coefficients from a generalized linear model predicting the proportion of wrists seen by infants.} 
+\caption{Model coefficients from a generalized linear model predicting the proportion of faces seen by infants.} 
 \end{table}
@@
 of age on the proportion of hands detected (b = -15.7, SE = 113.98, z =
--0.14 , p =0.89). Thus, these results suggest that infants access to
-social information is heavily infleunced by their postural and
+-0.14 , p =0.89). Thus, these results suggest that infants' access to
+social information is heavily influenced by their postural and
 locomotive development.
@@
 were hand-annotated and automatically synchronized with the
-frame-by-frame face detections. We again found that infant's posture and
+frame-by-frame face detections. We again found that infants' posture and
 orientations impacted the degree to which they saw their caregiver's

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly_briaStashed.Rmd
stashing changes in sep file"

--- 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
@@
 # run models
+## faces
 glm_faces<-glm(y_faces ~ dGLM_faces$age.at.test * dGLM_faces$posture * dGLM_faces$orientation,family=binomial(logit))
@@
 
+## wrists
 glm_wrists<-glm(y_wrists ~ dGLM_wrists$age.at.test * dGLM_wrists$posture * dGLM_wrists$orientation,family=binomial(logit))
@@
 We found the same pattern of results for the OpenPose face detections, omitted here for brevity.
-We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also reduced any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)` , p =`r round(wristGLMOut$coef[2,4],2)`).
+We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also reduced any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)` , p = å`r round(wristGLMOut$coef[2,4],2)`).
 Thus, these results suggest that infants access to social information is heavily infleunced by their postural and locomotive development.

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
abstract update

@@
 Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to detect the proportion of faces and hands in infants' environments.
-We find that infants' posture and orientation to their caregiver largely mediate infants' access to socially informative features, suggesting that postural and locomotive development play a significant role in mediating children's emerging linguistic and social capacities.
+We find that infants' posture and orientation to their caregiver largely mediate infants' access to socially informative features, suggesting that postural and locomotive developments play a significant role in the emergence of children's linguistic and social capacities.
 Broadly, we suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience."

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
detector wording

@@
 To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (operationalized by the presence of wrists) in the child’s visual environment, both overall and relative to naming events by their caregivers.
+We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrist presence as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
 We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, and walking toddlers would have access to a richer visual landscape that would contain a larger portion of the social information in the environment.
@@
   \hline
- Group & N & \% incl. & Mean age & Videos length (min) \\ 
+ Group & N & \% incl. & Avg age & Avg video length (min) \\ 
   \hline
@@
 
-## Face and Wrist Detection
+## Face and Hand Detection
 We used three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame.
@@
 
-This second of the CNN-based algorithms - the pose detector - provided us with the location of an agent's wrists, which we used as a good proxy for the presence of a hand.
-We do so first because we did not capture the entire visual field experienced by the infants/children.
+This second of the CNN-based algorithms--the pose detector--provided us (along with other parts) with the location of an agent's wrists, which we used as a good proxy for the presence of a hand.
+We did so first because we did not capture the entire visual field experienced by the infants/children.
 Thus, even though the video may only show the presence of a wrist, the caregiver's hand may be within the child's field of view.
-Second, we do so because hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
+Second, we did so because hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
 For example, if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy).
+Similarly, we used the nose (one of the 18 body parts detected) to operationalize face presence, as any half of a face necessarily contains a nose.
 
@@
 For the purposes of our investigation we only made use of the body parts relevant to the face and hands (nose and wrists).
-We operationalized face detections as any frames containing a nose, as any half of a face necessarily contains a nose.
-We operationalized wrist detections as any frames containing either the left or the right wrist.
+We operationalized face detections as any frames containing a nose, and hand detections as any frames containing either the left or the right wrist.
 
@@
 
-For face detection, MTCNN outperformed OpenPose when looking at the composite F-score (`r round(perfFacesOverall$f[1],2)` vs.
-`r round(perfFacesOverall$f[2],2)`); thus the face presence analyses to follow all result from MTCNN detections.
-We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN performs better on the high density sample evaluation (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
-ViolaJones performed quite poorly relative to the other detectors, finding much less faces in the random sample.
+For face detection, MTCNN outperformed OpenPose when taking into account only the composite F-score (`r round(perfFacesOverall$f[1],2)` MTCNN vs.
+`r round(perfFacesOverall$f[2],2)` OpenPose); thus the face presence analyses to follow all result from MTCNN detections.
+We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
+ViolaJones performed quite poorly relative to the other detectors especially on the random sample.
 
-For wrist detection, OpenPose performed moderately well (f = `r round(perfWristsOverall$f,2)`) with relatively high precision, but low recall on the randomly sampled set of frames (see Table 2).
+For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision, but low recall on the randomly sampled set of frames (see Table 2).
 We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of wrists overall in the datatset.

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
detector wording

@@
 informative features, suggesting that postural and locomotive
-development play a significant role in mediating children's emerging
+developments play a significant role in the emergence of children's
 linguistic and social capacities. Broadly, we suggest that the combined
@@
 Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
-frequencies of faces and hands (operationalized by the presence of
-wrists) in the child's visual environment, both overall and relative to
+frequencies of faces and hands (using wrist presence as a proxy for the
+latter) in the child's visual environment, both overall and relative to
 naming events by their caregivers. We hypothesized that there would be
@@
 
-\subsection{Face and Wrist Detection}\label{face-and-wrist-detection}
+\subsection{Face and Hand Detection}\label{face-and-hand-detection}
 
@@
 
-This second of the CNN-based algorithms - the pose detector - provided
-us with the location of an agent's wrists, which we used as a good proxy
-for the presence of a hand. We do so first because we did not capture
-the entire visual field experienced by the infants/children. Thus, even
-though the video may only show the presence of a wrist, the caregiver's
-hand may be within the child's field of view. Second, we do so because
-hands are often occluded by objects when caregivers are interacting with
-children, yet still visually accessible by the child and part of their
-joint interaction. For example, if a caregiver was holding a toy and
-presenting it to the infant, their wrists may be visible but their hand
-may not necessarily be (as it is occluded by the toy).
+This second of the CNN-based algorithms--the pose detector--provided us
+(along with other parts) with the location of an agent's wrists, which
+we used as a good proxy for the presence of a hand. We did so first
+because we did not capture the entire visual field experienced by the
+infants/children. Thus, even though the video may only show the presence
+of a wrist, the caregiver's hand may be within the child's field of
+view. Second, we did so because hands are often occluded by objects when
+caregivers are interacting with children, yet still visually accessible
+by the child and part of their joint interaction. For example, if a
+caregiver was holding a toy and presenting it to the infant, their
+wrists may be visible but their hand may not necessarily be (as it is
+occluded by the toy). Similarly, we used the nose (one of the 18 parts)
+to operationalize face presence, as any half of a face necessarily
+contains a nose.
 
@@
 face and hands (nose and wrists). We operationalized face detections as
-any frames containing a nose, as any half of a face necessarily contains
-a nose. We operationalized wrist detections as any frames containing
-either the left or the right wrist.
+any frames containing a nose, and hand detections as any frames
+containing either the left or the right wrist.
 

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
header edits

@@
 wrists may be visible but their hand may not necessarily be (as it is
-occluded by the toy). Similarly, we used the nose (one of the 18 parts)
-to operationalize face presence, as any half of a face necessarily
-contains a nose.
+occluded by the toy). Similarly, we used the nose (one of the 18 body
+parts detected) to operationalize face presence, as any half of a face
+necessarily contains a nose.
 
@@
 
-For face detection, MTCNN outperformed OpenPose when looking at the
-composite F-score (0.89 vs.~0.83); thus the face presence analyses to
-follow all result from MTCNN detections. We found that although MTCNN
-and OpenPose perform comparably with the random sample, MTCNN performs
-better on the high density sample evaluation (specifically looking at
-precision), suggesting that OpenPose generated more false positives than
-MTCNN. ViolaJones performed quite poorly relative to the other
-detectors, finding much less faces in the random sample.
+For face detection, MTCNN outperformed OpenPose when taking into account
+only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose); thus the face
+presence analyses to follow all result from MTCNN detections. We found
+that although MTCNN and OpenPose perform comparably with the random
+sample, MTCNN performed better on the high density sample (specifically
+looking at precision), suggesting that OpenPose generated more false
+positives than MTCNN. ViolaJones performed quite poorly relative to the
+other detectors especially on the random sample.
 
-For wrist detection, OpenPose performed moderately well (f = 0.72) with
+For wrist detection, OpenPose performed moderately well (F = 0.72) with
 relatively high precision, but low recall on the randomly sampled set of
@@
 
-\section{Changes in Access to Faces and
+\subsection{Changes in Access to Faces and
 Hands}\label{changes-in-access-to-faces-and-hands}
@@
 
-\subsection{Access to Faces and Wrists During Labeling
-Events}\label{access-to-faces-and-wrists-during-labeling-events}
+\subsection{Access to Faces and Hands During Labeling
+Events}\label{access-to-faces-and-hands-during-labeling-events}
 

### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
extr file

--- 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.Rmd
@@
 
-# Changes in Access to Faces and Hands
+## Changes in Access to Faces and Hands
 ```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age.
@@
 ```
-## Access to Faces and Wrists During Labeling Events
+## Access to Faces and Hands During Labeling Events
 Finally, we analyzed how face and wrists detections changed during object labeling events as a function of infant's posture and orientation.

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
figure changes for naming plot
more edits

@@
 
-However, as children are learning from others around them, their view of the world is also radically changing [@adolph2007].
+However, as children are learning from others around them, their view of the world is also changing radically [@adolph2007].
 Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently.
@@
 To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrist presence as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
-We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, and walking toddlers would have access to a richer visual landscape that would contain a larger portion of the social information in the environment.
+We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
+We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape thus enabling a larger portion of the social information in their environment to be accessible.
+
+that would contain a larger portion of the social information in the environment.
 
@@
   xlab("") +
-  ylim(0,.3) +
+  ylim(0,.25) +
   theme_few() + 
-  theme(legend.position="none")
-
+  theme(legend.position="none") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 p2=ggplot(aroundNamings_Orient_Faces, aes(x = orientation, y = mean, col = orientation)) +
@@
   xlab("") +
-  ylim(0,.3) +
+  ylim(0,.25) +
   theme_few() +
-  theme(legend.position="none")
-
+  theme(legend.position="none")  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 p3=ggplot(aroundNamings_Posture_Wrists, aes(x = posture, y = mean, col = posture)) +
@@
   xlab("") +
-  ylim(0,.3) +
+  ylim(0,.25) +
   theme_few() +
-  theme(legend.position="none")
-
+  theme(legend.position="none")  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 p4=ggplot(aroundNamings_Orient_Wrists, aes(x = orientation, y = mean, col = orientation)) +
@@
   xlab("") +
-  ylim(0,.3) +
+  ylim(0,.25) +
   theme_few() +
-  theme(legend.position="none")
-
+  theme(legend.position="none") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 ggarrange(p1,p2, p3, p4)

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
naming events redo

--- 2018-01-28 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
@@
 library(feather)
+library(assertthat)
 ```
@@
 A summary of the coefficients of a model with only main effects (and no interactions) model can be found in Table 3.
-When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p =`r round(faceGLMOut$coef[2,4],2)`).
+When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
 We found the same pattern of results for the OpenPose face detections, omitted here for brevity.
-We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also reduced any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)` , p =`r round(wristGLMOut$coef[2,4],2)`).
+We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also eliminated any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
 Thus, these results suggest that infants' access to social information is heavily influenced by their postural and locomotive development.
@@
   multi_boot_standard(col = "propWristsDetected") 
-
 ```
@@
   xlab("") +
-  ylim(0,.3) +
+  ylim(0,.25) +
   theme_few() + 
-  theme(legend.position="none")
+  theme(legend.position="none") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
   xlab("") +
-  ylim(0,.3) +
+  ylim(0,.25) +
   theme_few() +
-  theme(legend.position="none")
+  theme(legend.position="none")  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
   xlab("") +
-  ylim(0,.3) +
+  ylim(0,.25) +
   theme_few() +
-  theme(legend.position="none")
+  theme(legend.position="none") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 ```
+
+```{r echo=FALSE, include=FALSE}
+## Break down by before/after naming
+### First for wrists
+beforeNamings_Wrists <- d %>%
+  group_by(subid) %>%
+  mutate(detections = wristOP) %>%
+  do(summarize.naming(.,window = c(-2,0))) %>%
+  group_by(subid, age.grp)  %>%
+  summarise(avgDetections = mean(detections, na.rm=TRUE))  %>%
+  mutate(namingTime = "beforeNaming")
+
+namings_Wrists <- d %>%
+  group_by(subid) %>%
+  mutate(detections = wristOP) %>%
+  do(summarize.naming(.,window = c(0,2))) %>%
+  group_by(subid, age.grp)  %>%
+  summarise(avgDetections = mean(detections, na.rm=TRUE)) %>%
+  mutate(namingTime = "afterNaming") %>%
+  bind_rows(beforeNamings_Wrists)
+
+### Now for faces
+beforeNamings_Faces <- d %>%
+  group_by(subid) %>%
+  mutate(detections = faceMT) %>%
+  do(summarize.naming(.,window = c(-2,0))) %>%
+  group_by(subid, age.grp)  %>%
+  summarise(avgDetections = mean(detections, na.rm=TRUE))  %>%
+  mutate(namingTime = "beforeNaming")
+
+namings_Faces <- d %>%
+  group_by(subid) %>%
+  mutate(detections = faceMT) %>%
+  do(summarize.naming(.,window = c(0,2))) %>%
+  group_by(subid, age.grp)  %>%
+  summarise(avgDetections = mean(detections, na.rm=TRUE)) %>%
+  mutate(namingTime = "afterNaming") %>%
+  bind_rows(beforeNamings_Faces)
+
+# Plot if you want to look
+# p1=ggplot(namings_Faces, aes(x = factor(age.grp), y = avgDetections, col = age.grp)) +
+#   geom_boxplot() +
+#   ylab("Prop. faces") +
+#   xlab("") +
+#   theme_few() +
+#   theme(legend.position="none") +
+#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+#   facet_grid(~namingTime)
+# 
+# p2=ggplot(namings_Wrists, aes(x = factor(age.grp), y = avgDetections, col = age.grp)) +
+#   geom_boxplot() +
+#   ylab("Prop. wrists") +
+#   xlab("") +
+#   theme_few() +
+#   theme(legend.position="none") +
+#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+#   facet_grid(~namingTime)
+
+# ggarrange(p1,p2) # not shown in .rmd
+
+## output for text below \
+naming_Wrists_Summary <-  namings_Wrists %>%
+  spread(namingTime, avgDetections) %>%
+  mutate(diffWrists = afterNaming - beforeNaming) %>%
+  group_by(age.grp) %>%
+  summarize(mean = mean(diffWrists), sd=sd(diffWrists))
+
+naming_Faces_Summary <-  namings_Faces %>%
+  spread(namingTime, avgDetections) %>%
+  mutate(diffFaces = afterNaming - beforeNaming) %>%
+  group_by(age.grp) %>%
+  summarize(mean = mean(diffFaces), sd=sd(diffFaces)) # no room for SD maybe but still informative
+
+```
+
 ## Access to Faces and Hands During Labeling Events
-Finally, we analyzed how face and wrists detections changed during object labeling events as a function of infant's posture and orientation.
+Finally, we analyzed how face and wrist detections changed during object labeling events as a function of infant's posture and orientation.
 Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
-We again found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information.
+We found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information.
+However, this did not manifest in a direct change the presence of wrists/hands before vs.
+after naming events (average difference in proportion of wrists detected, 8 m.o.
+= `r round(naming_Wrists_Summary$mean[1],3)`, 12 m.o.
+= `r round(naming_Wrists_Summary$mean[2],3)`, 16 m.o.
+= `r round(naming_Wrists_Summary$mean[3],3)`) or in proportion of faces detected  (average difference in proportion of faces detected, 8 m.o.
+= `r round(naming_Faces_Summary$mean[1],3)`, 12 m.o.
+= `r round(naming_Faces_Summary$mean[2],3)`, 16 m.o.
+= `r round(naming_Faces_Summary$mean[3],3)`).
 
@@
 We used a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
-We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and wrists that were avaliable in the child's visual field.
-Thus, infants' postural and locomotive developments are mediating factors that explain some of the age-related changes in the proportion of faces and hands avaliable in infants' visual field.
+We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and wrists that were available in the child's visual field.
+Thus, infants' postural and locomotive developments are mediating factors that explain some of the age-related changes in the proportion of faces and wrists/hands that are visually available to infants.
 Broadly, this work suggests that motoric developments mediate how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
@@
 Videos from the infant perspective provided substantial challenges (e.g., partially occluded faces) for the classic models of face detection (e.g., ViolaJones) [@viola2004robust].
-Further, as the headcam technologies employed here were inexpensive (~$60 a camera) and the computer vision algorithms freely avaliable, this method is a promising avenue for quanitfying the visual and social information avaliable to infant learners.
+Further, as the headcam technologies employed here were inexpensive and the computer vision algorithms freely available, this method is a promising avenue for quantifying the visual and social information available to infant learners.
 
-Thus, we suggest that the combined use of these new tools can be leveraged to understand the changing infant perspecive on the visual world and the implications of these changes for both linguistic, cognitive, and social development.
+Broadly, we suggest that the combined use of these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex

--- 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
@@
 orientation, age no longer remained a significant predictor (b = -15.84,
-SE = 113.98, z = -0.14 , p =0.89). We found the same pattern of results
+SE = 113.98, z = -0.14 , p = 0.89). We found the same pattern of results
 for the OpenPose face detections, omitted here for brevity. We also
 found the same pattern with respect to infant's visual access to hands
-(see Table 4); including interaction terms also reduced any main effect
-of age on the proportion of hands detected (b = -15.7, SE = 113.98, z =
--0.14 , p =0.89). Thus, these results suggest that infants' access to
-social information is heavily influenced by their postural and
+(see Table 4); including interaction terms also eliminated any main
+effect of age on the proportion of hands detected (b = -15.7, SE =
+113.98, z = -0.14 , p = 0.89). Thus, these results suggest that infants'
+access to social information is heavily influenced by their postural and
 locomotive development.
@@
 
-Finally, we analyzed how face and wrists detections changed during
-object labeling events as a function of infant's posture and
-orientation. Specfically, we analyzed a four-second window around each
-labeling event (e.g., ``Look at the {[}zem{]}!''); these labeling events
-were hand-annotated and automatically synchronized with the
-frame-by-frame face detections. We again found that infants' posture and
-orientations impacted the degree to which they saw their caregiver's
-face and wrist during a labeling event; infants who were sitting or
-standing were more likely to have access to this social information.
+Finally, we analyzed how face and wrist detections changed during object
+labeling events as a function of infant's posture and orientation.
+Specfically, we analyzed a four-second window around each labeling event
+(e.g., ``Look at the {[}zem{]}!''); these labeling events were
+hand-annotated and automatically synchronized with the frame-by-frame
+face detections. We found that infants' posture and orientations
+impacted the degree to which they saw their caregiver's face and wrist
+during a labeling event; infants who were sitting or standing were more
+likely to have access to this social information. However, this did not
+manifest in a direct change the presence of wrists/hands before
+vs.~after naming events (average difference in proportion of wrists
+detected, 8 m.o. = -0.006,12 m.o. = -0.013 ,16 m.o. = -0.007 ) or in
+proportion of faces detected (average difference in proportion of faces
+detected, 8 m.o. = -0.002,12 m.o. = 0 ,16 m.o. = -0.004 ).
 
@@
 and that both of these factors dramatically impacted the proportion of
-faces and wrists that were avaliable in the child's visual field. Thus,
+faces and wrists that were available in the child's visual field. Thus,
 infants' postural and locomotive developments are mediating factors that
 explain some of the age-related changes in the proportion of faces and
-hands avaliable in infants' visual field. Broadly, this work suggests
-that motoric developments mediate how infants experience their visual
-world and the social information in it: infants that are sitting and
-standing have a different view of their world, the people in it, and the
-actions that are being performed.
+wrists/hands that are visually available to infants. Broadly, this work
+suggests that motoric developments mediate how infants experience their
+visual world and the social information in it: infants that are sitting
+and standing have a different view of their world, the people in it, and
+the actions that are being performed.
 
@@
 (e.g., ViolaJones) (Viola \& Jones, 2004). Further, as the headcam
-technologies employed here were inexpensive (\textasciitilde{}\$60 a
-camera) and the computer vision algorithms freely avaliable, this method
-is a promising avenue for quanitfying the visual and social information
-avaliable to infant learners.
+technologies employed here were inexpensive and the computer vision
+algorithms freely available, this method is a promising avenue for
+quantifying the visual and social information available to infant
+learners.
 
-Thus, we suggest that the combined use of these new tools can be
-leveraged to understand the changing infant perspecive on the visual
-world and the implications of these changes for both linguistic,
-cognitive, and social development.
+Broadly, we suggest that the combined use of these new tools can be
+leveraged to understand the changing infant perspective on the visual
+world and the implications of these changes for linguistic, cognitive,
+and social development.
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
random edits

--- 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
@@
 \begin{figure*}
+\centering
 \includegraphics[width=6in]{images/detector_samples_banner.pdf}
@@
 The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones.
-We used this as a benchmark for performance, and while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
+We used this as a benchmark for performance; while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
 We next capitalized on recent improvements in computer vision, testing the performance of two face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information.
@@
 
-This second of the CNN-based algorithms--the pose detector--provided us (along with other parts) with the location of an agent's wrists, which we used as a good proxy for the presence of a hand.
+This second of the CNN-based algorithms--the pose detector--provided us with the location of an agent's wrists, which we used as a proxy for the presence of a hand.
 We did so first because we did not capture the entire visual field experienced by the infants/children.
@@
 For example, if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy).
-Similarly, we used the nose (one of the 18 body parts detected) to operationalize face presence, as any half of a face necessarily contains a nose.
+Similarly, we used an agent's nose (one of the 18 body parts detected) to operationalize face presence, as any half of a face necessarily contains a nose.
 
@@
 Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 2.
-For wrist detections, the "gold set" was constructed in the same manner, with the exception that frames with a high density of wrists were taken by wrist detections made by OpenPose.
+For wrist detections, the "gold set" was constructed in the same manner, with the exception that frames with a high density of wrists were taken by pose detections made by OpenPose.
 
@@
 We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
-ViolaJones performed quite poorly relative to the other detectors especially on the random sample.
+ViolaJones performed quite poorly relative to the other detectors especially with respect to the random sample.
 
@@
 The proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with age.
-Both 8-month-olds and 12-month-olds spent equivalent amounts of time either lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time either sitting or standing (see Figure \ref{fig:posture}).
+Both 8 month-olds and 12 month-olds spent equivalent amounts of time either lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time either sitting or standing (see Figure \ref{fig:posture}).
 We also observed characteristic changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:posture}).

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
?

--- 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex
@@
 However, as children are learning from others around them, their view of
-the world is also radically changing (K. Adolph \& Berger, 2007).
+the world is also changing radically (K. Adolph \& Berger, 2007).
 Infants' motor abilities improve dramatically near the end of the first
@@
 Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
-frequencies of faces and hands (using wrist presence as a proxy for the
-latter) in the child's visual environment, both overall and relative to
-naming events by their caregivers. We hypothesized that there would be
+frequencies of faces and hands (using wrists as a proxy for the latter)
+in the child's visual environment, both overall and relative to naming
+events by their caregivers. We hypothesized that there would be
 differential access to social information based on children's postural
 developments: crawling infants would see fewer faces because they would
-primarily be looking at the ground, and walking toddlers would have
-access to a richer visual landscape that would contain a larger portion
-of the social information in the environment.
+primarily be looking at the ground, while walking toddlers would have
+access to a richer visual landscape thus enabling a larger portion of
+the social information in their environment to be accessible.
+
+that would contain a larger portion of the social information in the
+environment.
 
@@
 face detection algorithm: Viola-Jones. We used this as a benchmark for
-performance, and while it can achieve impressive accuracy in some
+performance; while it can achieve impressive accuracy in some
 situations, it is notoriously bad at dealing with occluded faces
@@
 This second of the CNN-based algorithms--the pose detector--provided us
-(along with other parts) with the location of an agent's wrists, which
-we used as a good proxy for the presence of a hand. We did so first
-because we did not capture the entire visual field experienced by the
-infants/children. Thus, even though the video may only show the presence
-of a wrist, the caregiver's hand may be within the child's field of
-view. Second, we did so because hands are often occluded by objects when
-caregivers are interacting with children, yet still visually accessible
-by the child and part of their joint interaction. For example, if a
-caregiver was holding a toy and presenting it to the infant, their
-wrists may be visible but their hand may not necessarily be (as it is
-occluded by the toy). Similarly, we used the nose (one of the 18 body
-parts detected) to operationalize face presence, as any half of a face
-necessarily contains a nose.
+with the location of an agent's wrists, which we used as a proxy for the
+presence of a hand. We did so first because we did not capture the
+entire visual field experienced by the infants/children. Thus, even
+though the video may only show the presence of a wrist, the caregiver's
+hand may be within the child's field of view. Second, we did so because
+hands are often occluded by objects when caregivers are interacting with
+children, yet still visually accessible by the child and part of their
+joint interaction. For example, if a caregiver was holding a toy and
+presenting it to the infant, their wrists may be visible but their hand
+may not necessarily be (as it is occluded by the toy). Similarly, we
+used an agent's nose (one of the 18 body parts detected) to
+operationalize face presence, as any half of a face necessarily contains
+a nose.
 
@@
 set'' was constructed in the same manner, with the exception that frames
-with a high density of wrists were taken by wrist detections made by
+with a high density of wrists were taken by pose detections made by
 OpenPose.
@@
 positives than MTCNN. ViolaJones performed quite poorly relative to the
-other detectors especially on the random sample.
+other detectors especially with respect to the random sample.
 
@@
 spent sitting decreased with age, and the proportion of time infants
-spent standing increased with age. Both 8-month-olds and 12-month-olds
+spent standing increased with age. Both 8 month-olds and 12 month-olds
 spent equivalent amounts of time either lying/crawling, which was

### 2018-01-29 analysis-main/main_analyses_bll.Rmd

@@
 ```{r}
-d <- read_csv("../data/consolidateddata/consolidated_data_4dets.csv") %>%
+d <- read_csv("../data/consolidated_data_4detectors.feather") %>%
   mutate(posture = factor(posture), 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
wrist detection corrections

--- 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
@@
 ### READ IN DETECTIONS
-getwd()
 d <- read_feather("../data/consolidated_data_4detectors.feather") %>%
@@
   ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
-  OpenPose-Wrists & High density & 0.64 & 1.00 & 0.78 \\ 
-  OpenPose-Wrists & Random & 0.76 & 0.26 & 0.39 \\ 
+  OpenPose-Wrists & High density & 0.66 & 1.00 & 0.79 \\ 
+  OpenPose-Wrists & Random & 0.88 & 0.29 & 0.43 \\ 
    \hline

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
wrist detection corrections

@@
 
-For wrist detection, OpenPose performed moderately well (F = 0.72) with
+For wrist detection, OpenPose performed moderately well (F = 0.74) with
 relatively high precision, but low recall on the randomly sampled set of
@@
 orientation, age no longer remained a significant predictor (b = -15.84,
-SE = 113.98, z = -0.14 , p =0.89). We found the same pattern of results
+SE = 113.98, z = -0.14 , p = 0.89). We found the same pattern of results
 for the OpenPose face detections, omitted here for brevity. We also
 found the same pattern with respect to infant's visual access to hands
-(see Table 4); including interaction terms also reduced any main effect
-of age on the proportion of hands detected (b = -15.7, SE = 113.98, z =
--0.14 , p =0.89). Thus, these results suggest that infants' access to
-social information is heavily influenced by their postural and
+(see Table 4); including interaction terms also eliminated any main
+effect of age on the proportion of hands detected (b = -15.7, SE =
+113.98, z = -0.14, p = 0.89). Thus, these results suggest that infants'
+access to social information is heavily influenced by their postural and
 locomotive development.
@@
 
-Finally, we analyzed how face and wrists detections changed during
-object labeling events as a function of infant's posture and
-orientation. Specfically, we analyzed a four-second window around each
-labeling event (e.g., ``Look at the {[}zem{]}!''); these labeling events
-were hand-annotated and automatically synchronized with the
-frame-by-frame face detections. We again found that infants' posture and
-orientations impacted the degree to which they saw their caregiver's
-face and wrist during a labeling event; infants who were sitting or
-standing were more likely to have access to this social information.
+Finally, we analyzed how face and wrist detections changed during object
+labeling events as a function of infant's posture and orientation.
+Specfically, we analyzed a four-second window around each labeling event
+(e.g., ``Look at the {[}zem{]}!''); these labeling events were
+hand-annotated and automatically synchronized with the frame-by-frame
+face detections. We found that infants' posture and orientations
+impacted the degree to which they saw their caregiver's face and wrist
+during a labeling event; infants who were sitting or standing were more
+likely to have access to this social information. However, this did not
+manifest in a direct change the presence of wrists/hands before
+vs.~after naming events (average difference in proportion of wrists
+detected, 8 m.o. = -0.006, 12 m.o. = -0.013, 16 m.o. = -0.007) or in
+proportion of faces detected (average difference in proportion of faces
+detected, 8 m.o. = -0.002, 12 m.o. = 0, 16 m.o. = -0.004).
 

### 2018-01-29 scripts/ground_truth_creation/validate.py

@@ -5,7 +5,7 @@ import xml.etree.ElementTree as ET
 
 # TODO take in arg for wrist or face and change the corresponding is_face?
 
-BRIA_ANNOTATIONS = "/Users/alessandro/Downloads/sample_face_images_openpose_bll"
+BRIA_ANNOTATIONS = "/Users/alessandro/Downloads/sample_hand_images_bll"
 
 if __name__ == "__main__":
 
@@ -33,7 +33,7 @@ if __name__ == "__main__":
 
     print(len(rows))
 
-    with open("ground_truth_face_openpose.csv", 'wb') as csvfile:
+    with open("ground_truth_wrists2.csv", 'wb') as csvfile:
         wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
         wr.writerow(['group', 'video', 'frame', 'is_face'])
         for key, val in rows.iteritems():

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
edits

@@
 We used a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
-We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and wrists that were available in the child's visual field.
+We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and hands that were available in the child's visual field.
 Thus, infants' postural and locomotive developments are mediating factors that explain some of the age-related changes in the proportion of faces and wrists/hands that are visually available to infants.

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
edits

@@
 and that both of these factors dramatically impacted the proportion of
-faces and wrists that were available in the child's visual field. Thus,
+faces and hands that were available in the child's visual field. Thus,
 infants' postural and locomotive developments are mediating factors that

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
smoothing edits

@@
 How do these postural and locomotive changes affect children's access to the social information relevant for word-learning?
-Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to detect the proportion of faces and hands in infants' environments.
-We find that infants' posture and orientation to their caregiver largely mediate infants' access to socially informative features, suggesting that postural and locomotive developments play a significant role in the emergence of children's linguistic and social capacities.
+Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to detect the proportion of faces and wrists in infants' environments.
+We find that infants' posture and orientation to their caregiver largely mediates their access to social information, suggesting that postural and locomotive developments play a significant role in the emergence of children's linguistic and social capacities.
 Broadly, we suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
@@
 Even newborns tend to prefer to look at faces with direct vs.
-averted gaze [@farroni2002eye] and follow overt gaze shifts by 6 months of age [@gredeback2008microstructure; @gredeback2010development].
-And when infants free-view videos, they tend to look mostly at faces at the expense of other visual information--though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
+averted gaze [@farroni2002eye] and young infants follow overt gaze shifts  [@gredeback2008microstructure; @gredeback2010development].
+Further, when infants free-view videos, they tend to look mostly at faces at the expense of other visual information--though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
 
-However, as children are learning from others around them, their view of the world is also changing radically [@adolph2007].
+However, as children are learning from others around them, their view of the world is also radically changing  [@adolph2007].
 Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently.
-These motor changes have drastic consequences for what children see; crawling and walking infants have radically different views of the world.
+These motor changes have significant consequences for what children see; crawling and walking infants simply have different views of the world.
 For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
@@
 Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
-On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access social information (e.g., faces, gaze) relative to children who are still crawling and sitting.
-In other words, the ability to walk and move around independently and subsequently access more detailed social information may accelerate infants’ ability to learn from the people in their environments.
+On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access social information (e.g., faces, gaze)  and in turn to accelerate their own learning.
 
 Recent technological developments allow for testing of this hypothesis by documenting the experiences of infants and children from their own perspective.
-By using head-mounted cameras, researchers have begun to document visual experiences of infants and children — which even for walking children are radically different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008].
-Children’s views tend to be restricted and to be dominated by objects and hands much more than that of adults [@yoshida2008], and both computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2012].
-Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing a close up view of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
+By using head-mounted cameras, researchers have begun to record the visual experiences of infants and children-- which even for walking children are extremely different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008].
+While children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], both computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2012].
+Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
 
-Here, we directly ask whether postural and locomotive developments change the availability of social information--the presence of faces and hands.
+Here, we directly examine whether postural and locomotive developments change the availability of social information--the presence of faces and hands.
 To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
 We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
-We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape thus enabling a larger portion of the social information in their environment to be accessible.
-
-that would contain a larger portion of the social information in the environment.
+We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus rendering accessible a larger portion of the social information in their environment.
 
@@
 Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's approximate vertical field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988].
-As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see the faces of adults around them.
-This choice allowed us to maximize our chances of capturing faces that the child would have seen during the play session.
+As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces.
 
 ## Procedure
-
 First, all parents signed consent documents in a waiting room where children were fitted with the headcam.
@@
 If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam [@yoshida2008].
-
-After the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session--the focus of the current study.
+After the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session.
 Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the novel toys.
-
 The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.
@@
 ## Face and Hand Detection
-We used three face detection systems to explore infants' changing access to social information and to avoid the cost of hand-annotating every frame.
+We used three face detection systems to measure infants' changing access to faces.
 The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones.
-We used this as a benchmark for performance; while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
-We next capitalized on recent improvements in computer vision, testing the performance of two face detectors that both made use of Convolutional Neural Networks (CNNs) to extract face information.
-The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts belonging to an agent.
+We used this as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
+We next capitalized tested the performance of two face detectors that both made use of recently developed Convolutional Neural Networks (CNNs) to extract face information.
+The first algorithm was specifically optimized for face detection, and the second algorithm was this to extract information about the position of 18 different body parts.
+For the second algorithm (called OpenPose) [@cao2017realtime], we used the agent's nose (one of the 18 body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
 
-This second of the CNN-based algorithms--the pose detector--provided us with the location of an agent's wrists, which we used as a proxy for the presence of a hand.
+The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for the presence of a hand.
 We did so first because we did not capture the entire visual field experienced by the infants/children.
@@
 For example, if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy).
-Similarly, we used an agent's nose (one of the 18 body parts detected) to operationalize face presence, as any half of a face necessarily contains a nose.
 
@@
 This detector provided information about the presence of a face as well as its size and position.
-
-The second algorithm was based on the work by @zhang2016 using multi-task cascaded convolutional neural networks (MTCNNs).
+The second algorithm (based on  work by @zhang2016) using multi-task cascaded convolutional neural networks (MTCNNs).
 The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
@@
 
-The third algorithm was a CNN-based pose detector that provided the locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm] avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The CNN-based pose detector (OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.)  avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
-The candidates are then matched to a single individual and finally assembled into a pose.
-For the purposes of our investigation we only made use of the body parts relevant to the face and hands (nose and wrists).
-We operationalized face detections as any frames containing a nose, and hand detections as any frames containing either the left or the right wrist.
+The candidates are then matched to a single individual and finally assembled into a pose; here, we only made use of the body parts relevant to the face and hands (nose and wrists).
+We operationalized face detections as any frames containing a nose, and hand detections as any frames containing either the left orright wrist.
 
 ### Detector evaluation
-To evaluate face detector performance, we needed to construct an unbiased "gold set" of labeled frames that made sure to account for the relatively rare appearance of faces in the dataset.
+To evaluate face detector performance, we constructed an unbiased "gold set" of labeled frames that accounted for the relatively rare appearance of faces in the dataset.
 We thus hand-labeled two samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
-Each sample was comprised of an equal number of frames taken from each participant (i.e., video).
+Each sample was comprised of an equal number of frames taken from each child's video.
 Faces were classifed as present in a frame if at least half of the face was showing.
@@
 For face detection, MTCNN outperformed OpenPose when taking into account only the composite F-score (`r round(perfFacesOverall$f[1],2)` MTCNN vs.
-`r round(perfFacesOverall$f[2],2)` OpenPose); thus the face presence analyses to follow all result from MTCNN detections.
-We found that although MTCNN and OpenPose perform comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
-ViolaJones performed quite poorly relative to the other detectors especially with respect to the random sample.
+`r round(perfFacesOverall$f[2],2)` OpenPose); we thus use MTCNN detections in the following analyses.
+Although MTCNN and OpenPose performed comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
+ViolaJones performed quite poorly relative to the other detectors, especially with respect to the random sample.
 
@@
 First, we report developmental shifts in infants' posture and their orientation relative to their caregiver.
-Then, we explore how these changes influence children's visual access to faces and hands across this developmental time range.
-Finally, we explore how these changes impact the accessibility of faces and hands during labeling events.
+Then, we explore how these changes influence children's visual access to faces and wrists/hands across this developmental time range.
+Finally, we explore how these changes impact the accessibility of faces and wrists/hands during labeling events.
 
@@
 ```
-
-We first examined the proportion of face and wrist detections across age; a full summary can be seen in Figure \ref{fig:detByAge}.
-When looking at face presence, we observed a slight U-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slighly fewer faces faces than 8 or 16-month-olds.
-
-However, we found that any age related effects were much smaller compared to the impact of postural and locomotive changes on children's visual access to faces.
-Children's posture was a major factor both in how many faces they saw during the play session.
-Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces.
-We also examined how the child's orientation relative to their caregiver impacted their visual access to faces: children who were far away from their caregiver were more likely to see faces than children who were close to their caregiver; this was true within all age groups and for both face detectors.
-
 ```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=4, fig.cap = "Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
@@
 ```
+
+We first examined the proportion of face and wrist detections across age; a full summary can be seen in Figure \ref{fig:detByAge}.
+When looking at face presence, we observed a slight U-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slighly fewer faces faces than 8 or 16-month-olds.
+
+However, we found that any age related effects were much smaller compared to the impact of postural and locomotive changes on children's visual access to faces.
+Children's posture was a major factor both in how many faces and wrists/hands they saw during the play session.
+Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces (Figure \ref{fig:detByPosOrient}, upper panel); this same pattern was also true for wrists/hand detectinos.
+We also examined how the child's orientation relative to their caregiver impacted their visual access to faces and hands.
+Children who were far away from their caregiver were more likely to see faces than children who were close to their caregiver; this was true within all age groups  (Figure \ref{fig:detByPosOrient}, lower panel).
 
@@
 When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
-We found the same pattern of results for the OpenPose face detections, omitted here for brevity.
 We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also eliminated any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
-Thus, these results suggest that infants' access to social information is heavily influenced by their postural and locomotive development.
+Thus, these results suggest that infants' access to social information is heavily influenced by their postural and locomotive developments.
 
@@
 ## Access to Faces and Hands During Labeling Events
-Finally, we analyzed how face and wrist detections changed during object labeling events as a function of infant's posture and orientation.
+Finally, we analyzed how face and wrist/hand detections changed during object labeling events as a function of infant's posture and orientation.
 Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
@@
 We used a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
-We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and hands that were available in the child's visual field.
+We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and wrists/hands that were available in the child's visual field.
 Thus, infants' postural and locomotive developments are mediating factors that explain some of the age-related changes in the proportion of faces and wrists/hands that are visually available to infants.

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
smoothing edits

@@
 infants' (8-16 months of age) egocentric visual perspective and use
-computer vision algorithms to detect the proportion of faces and hands
+computer vision algorithms to detect the proportion of faces and wrists
 in infants' environments. We find that infants' posture and orientation
-to their caregiver largely mediate infants' access to socially
-informative features, suggesting that postural and locomotive
-developments play a significant role in the emergence of children's
-linguistic and social capacities. Broadly, we suggest that the combined
-use of head-mounted cameras and the application of new computer vision
-techniques is a promising avenue for understanding the statistics of
-infants' visual and linguistic experience.
+to their caregiver largely mediates their access to social information,
+suggesting that postural and locomotive developments play a significant
+role in the emergence of children's linguistic and social capacities.
+Broadly, we suggest that the combined use of head-mounted cameras and
+the application of new computer vision techniques is a promising avenue
+for understanding the statistics of infants' visual and linguistic
+experience.
 
@@
 look at faces with direct vs.~averted gaze (Farroni, Csibra, Simion, \&
-Johnson, 2002) and follow overt gaze shifts by 6 months of age
-(Gredebäck, Fikke, \& Melinder, 2010; Gredebäck, Theuring, Hauf, \&
-Kenward, 2008). And when infants free-view videos, they tend to look
-mostly at faces at the expense of other visual information--though older
-infants start to look towards people's hands and the actions they are
-performing (Frank, Amso, \& Johnson, 2014; Frank, Vul, \& Saxe, 2012).
+Johnson, 2002) and young infants follow overt gaze shifts (Gredebäck,
+Fikke, \& Melinder, 2010; Gredebäck, Theuring, Hauf, \& Kenward, 2008).
+Further, when infants free-view videos, they tend to look mostly at
+faces at the expense of other visual information--though older infants
+start to look towards people's hands and the actions they are performing
+(Frank, Amso, \& Johnson, 2014; Frank, Vul, \& Saxe, 2012).
 
 However, as children are learning from others around them, their view of
-the world is also changing radically (K. Adolph \& Berger, 2007).
+the world is also radically changing (K. Adolph \& Berger, 2007).
 Infants' motor abilities improve dramatically near the end of the first
 year of life, allowing them to locomote independently. These motor
-changes have drastic consequences for what children see; crawling and
-walking infants have radically different views of the world. For
+changes have significant consequences for what children see; crawling
+and walking infants simply have different views of the world. For
 example, during spontaneous play in a laboratory playroom, toddlers are
@@
 and independently locomote may fundamentally change their ability to
-access social information (e.g., faces, gaze) relative to children who
-are still crawling and sitting. In other words, the ability to walk and
-move around independently and subsequently access more detailed social
-information may accelerate infants' ability to learn from the people in
-their environments.
+access social information (e.g., faces, gaze) and in turn to accelerate
+their own learning.
 
@@
 perspective. By using head-mounted cameras, researchers have begun to
-document visual experiences of infants and children --- which even for
-walking children are radically different from the adult perspective (and
+record the visual experiences of infants and children-- which even for
+walking children are extremely different from the adult perspective (and
 not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
 Yu, \& Smith, 2017; J. Franchak et al., 2011; Yoshida \& Smith, 2008).
-Children's views tend to be restricted and to be dominated by objects
-and hands much more than that of adults (Yoshida \& Smith, 2008), and
-both computational and empirical work suggest that this restricted
-viewpoint may be more effective for learning objects and their labels
-than the comparable adult perspective (Bambach, Crandall, Smith, \& Yu,
-2017; D. Yurovsky, Smith, \& Yu, in press). Further, recent work also
-suggests dramatic changes in the child's perspective over the first two
-years of life, as views transition from primarily containing a close up
-view of faces to capturing views of hands paired with the objects they
-are acting on (Fausey, Jayaraman, \& Smith, 2016).
+While children's views tend to be more restricted and dominated by
+objects and hands (Yoshida \& Smith, 2008), both computational and
+empirical work suggest that this restricted viewpoint may be more
+effective for learning objects and their labels than the comparable
+adult perspective (Bambach, Crandall, Smith, \& Yu, 2017; D. Yurovsky,
+Smith, \& Yu, in press). Further, recent work also suggests dramatic
+changes in the child's perspective over the first two years of life, as
+views transition from primarily containing close up views of faces to
+capturing views of hands paired with the objects they are acting on
+(Fausey, Jayaraman, \& Smith, 2016).
 
-Here, we directly ask whether postural and locomotive developments
+Here, we directly examine whether postural and locomotive developments
 change the availability of social information--the presence of faces and
@@
 primarily be looking at the ground, while walking toddlers would have
-access to a richer visual landscape thus enabling a larger portion of
-the social information in their environment to be accessible.
-
-that would contain a larger portion of the social information in the
-environment.
+access to a richer visual landscape, thus rendering accessible a larger
+portion of the social information in their environment.
 
@@
 view, we chose to orient the camera upwards to capture the entirety of
-the child's upper visual field where the child is likely to see the
-faces of adults around them. This choice allowed us to maximize our
-chances of capturing faces that the child would have seen during the
-play session.
+the child's upper visual field where the child is likely to see adult
+faces.
 
@@
 engaging toys to try to draw the child's focus away from the headcam
-(Yoshida \& Smith, 2008).
-
-After the child was comfortable wearing the headcam, the child and
-caregiver were shown to a playroom for the free-play session--the focus
-of the current study. Parents were shown a box containing three pairs of
+(Yoshida \& Smith, 2008). After the child was comfortable wearing the
+headcam, the child and caregiver were shown to a playroom for the
+free-play session. Parents were shown a box containing three pairs of
 novel and familiar objects (e.g., a ball and a microfiber duster, named
@@
 that their child had not previously seen the novel toys and were
-instructed to use the novel labels to refer to the novel toys.
-
-The experimenter then left the playroom for approximately 15 minutes,
-during which a tripod-mounted camera in the corner of the room recorded
-the session and the headcam captured video from the child's perspective.
+instructed to use the novel labels to refer to the novel toys. The
+experimenter then left the playroom for approximately 15 minutes, during
+which a tripod-mounted camera in the corner of the room recorded the
+session and the headcam captured video from the child's perspective.
 
@@
 
-We used three face detection systems to explore infants' changing access
-to social information and to avoid the cost of hand-annotating every
-frame. The first of these is the most commonly-used and widely available
-face detection algorithm: Viola-Jones. We used this as a benchmark for
-performance; while it can achieve impressive accuracy in some
-situations, it is notoriously bad at dealing with occluded faces
-(Scheirer, Anthony, Nakayama, \& Cox, 2014). We next capitalized on
-recent improvements in computer vision, testing the performance of two
-face detectors that both made use of Convolutional Neural Networks
-(CNNs) to extract face information. The first algorithm was specifically
-optimized for face detection, and the second algorithm was optimized to
-extract information about the position of 18 different body parts
-belonging to an agent.
+We used three face detection systems to measure infants' changing access
+to faces. The first of these is the most commonly-used and widely
+available face detection algorithm: Viola-Jones. We used this as a
+benchmark for performance, as while it can achieve impressive accuracy
+in some situations, it is notoriously bad at dealing with occluded faces
+(Scheirer, Anthony, Nakayama, \& Cox, 2014). We next capitalized tested
+the performance of two face detectors that both made use of recently
+developed Convolutional Neural Networks (CNNs) to extract face
+information. The first algorithm was specifically optimized for face
+detection, and the second algorithm was this to extract information
+about the position of 18 different body parts. For the second algorithm
+(called OpenPose) (Cao et al., 2017), we used the agent's nose (one of
+the 18 body parts detected) to operationalize the presence of faces, as
+any half of a face necessarily contains a nose.
 
-This second of the CNN-based algorithms--the pose detector--provided us
-with the location of an agent's wrists, which we used as a proxy for the
-presence of a hand. We did so first because we did not capture the
-entire visual field experienced by the infants/children. Thus, even
-though the video may only show the presence of a wrist, the caregiver's
-hand may be within the child's field of view. Second, we did so because
-hands are often occluded by objects when caregivers are interacting with
-children, yet still visually accessible by the child and part of their
-joint interaction. For example, if a caregiver was holding a toy and
-presenting it to the infant, their wrists may be visible but their hand
-may not necessarily be (as it is occluded by the toy). Similarly, we
-used an agent's nose (one of the 18 body parts detected) to
-operationalize face presence, as any half of a face necessarily contains
-a nose.
+The OpenPose detector also provided us with the location of an agent's
+wrists, which we used as a proxy for the presence of a hand. We did so
+first because we did not capture the entire visual field experienced by
+the infants/children. Thus, even though the video may only show the
+presence of a wrist, the caregiver's hand may be within the child's
+field of view. Second, we did so because hands are often occluded by
+objects when caregivers are interacting with children, yet still
+visually accessible by the child and part of their joint interaction.
+For example, if a caregiver was holding a toy and presenting it to the
+infant, their wrists may be visible but their hand may not necessarily
+be (as it is occluded by the toy).
 
@@
 individual frame. This detector provided information about the presence
-of a face as well as its size and position.
-
-The second algorithm was based on the work by K. Zhang et al. (2016)
-using multi-task cascaded convolutional neural networks (MTCNNs). The
-system was built using a novel cascaded CNN-based framework for joint
-detection and alignment, built to perform well in real-world
-environments where varying illuminations and occlusions are present. We
-used a Tensorflow implementation of this algorithm provided by
+of a face as well as its size and position. The second algorithm (based
+on work by K. Zhang et al. (2016)) using multi-task cascaded
+convolutional neural networks (MTCNNs). The system was built using a
+novel cascaded CNN-based framework for joint detection and alignment,
+built to perform well in real-world environments where varying
+illuminations and occlusions are present. We used a Tensorflow
+implementation of this algorithm provided by
 (\url{https://github.com/davidsandberg/facenet}). Like Viola-Jones, this
@@
 
-The third algorithm was a CNN-based pose detector that provided the
-locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose
-(Cao et al., 2017; Simon, Joo, Matthews, \& Sheikh, 2017; Wei,
-Ramakrishna, Kanade, \& Sheikh, 2016) avaliable at
+The CNN-based pose detector (OpenPose (Cao et al., 2017; Simon, Joo,
+Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016))
+provided the locations of 18 body parts (ears, nose, wrists, etc.)
+avaliable at
 \url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
@@
 series of body part candidates. The candidates are then matched to a
-single individual and finally assembled into a pose. For the purposes of
-our investigation we only made use of the body parts relevant to the
-face and hands (nose and wrists). We operationalized face detections as
-any frames containing a nose, and hand detections as any frames
-containing either the left or the right wrist.
+single individual and finally assembled into a pose; here, we only made
+use of the body parts relevant to the face and hands (nose and wrists).
+We operationalized face detections as any frames containing a nose, and
+hand detections as any frames containing either the left orright wrist.
 
@@
 
-To evaluate face detector performance, we needed to construct an
-unbiased ``gold set'' of labeled frames that made sure to account for
-the relatively rare appearance of faces in the dataset. We thus
-hand-labeled two samples: a sample containing a high density of faces
-(half reported by MTCNN, half by OpenPose) and a random sample from the
-remaining frames. Each sample was comprised of an equal number of frames
-taken from each participant (i.e., video). Faces were classifed as
-present in a frame if at least half of the face was showing. Precision
-(hits / hits + false alarms), recall (hits / hits + misses), and F-score
-(harmonic mean of precision and recall) were calculated for all
-detectors and are reported in Table 2. For wrist detections, the ``gold
-set'' was constructed in the same manner, with the exception that frames
-with a high density of wrists were taken by pose detections made by
-OpenPose.
+To evaluate face detector performance, we constructed an unbiased ``gold
+set'' of labeled frames that accounted for the relatively rare
+appearance of faces in the dataset. We thus hand-labeled two samples: a
+sample containing a high density of faces (half reported by MTCNN, half
+by OpenPose) and a random sample from the remaining frames. Each sample
+was comprised of an equal number of frames taken from each child's
+video. Faces were classifed as present in a frame if at least half of
+the face was showing. Precision (hits / hits + false alarms), recall
+(hits / hits + misses), and F-score (harmonic mean of precision and
+recall) were calculated for all detectors and are reported in Table 2.
+For wrist detections, the ``gold set'' was constructed in the same
+manner, with the exception that frames with a high density of wrists
+were taken by pose detections made by OpenPose.
 
 For face detection, MTCNN outperformed OpenPose when taking into account
-only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose); thus the face
-presence analyses to follow all result from MTCNN detections. We found
-that although MTCNN and OpenPose perform comparably with the random
-sample, MTCNN performed better on the high density sample (specifically
-looking at precision), suggesting that OpenPose generated more false
-positives than MTCNN. ViolaJones performed quite poorly relative to the
-other detectors especially with respect to the random sample.
+only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose); we thus use
+MTCNN detections in the following analyses. Although MTCNN and OpenPose
+performed comparably with the random sample, MTCNN performed better on
+the high density sample (specifically looking at precision), suggesting
+that OpenPose generated more false positives than MTCNN. ViolaJones
+performed quite poorly relative to the other detectors, especially with
+respect to the random sample.
 
@@
 orientation relative to their caregiver. Then, we explore how these
-changes influence children's visual access to faces and hands across
-this developmental time range. Finally, we explore how these changes
-impact the accessibility of faces and hands during labeling events.
+changes influence children's visual access to faces and wrists/hands
+across this developmental time range. Finally, we explore how these
+changes impact the accessibility of faces and wrists/hands during
+labeling events.
 
@@
 \end{figure}
+\end{CodeChunk}\begin{CodeChunk}
+\begin{figure*}[h]
+
+{\centering \includegraphics{figs/detByPosOrient-1}
+
+}
+
+\caption[Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant]{Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant.}\label{fig:detByPosOrient}
+\end{figure*}
 \end{CodeChunk}
@@
 visual access to faces. Children's posture was a major factor both in
-how many faces they saw during the play session. Infants who were
-sitting saw more faces than infants who were lying down or being
-carried, while infants who were standing saw the most faces. We also
-examined how the child's orientation relative to their caregiver
-impacted their visual access to faces: children who were far away from
-their caregiver were more likely to see faces than children who were
-close to their caregiver; this was true within all age groups and for
-both face detectors.
-
-\begin{CodeChunk}
-\begin{figure*}[h]
-
-{\centering \includegraphics{figs/detByPosOrient-1}
-
-}
-
-\caption[Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant]{Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant.}\label{fig:detByPosOrient}
-\end{figure*}
-\end{CodeChunk}
+how many faces and wrists/hands they saw during the play session.
+Infants who were sitting saw more faces than infants who were lying down
+or being carried, while infants who were standing saw the most faces
+(Figure \ref{fig:detByPosOrient}, upper panel); this same pattern was
+also true for wrists/hand detectinos. We also examined how the child's
+orientation relative to their caregiver impacted their visual access to
+faces and hands. Children who were far away from their caregiver were
+more likely to see faces than children who were close to their
+caregiver; this was true within all age groups (Figure
+\ref{fig:detByPosOrient}, lower panel).
 
@@
 orientation, age no longer remained a significant predictor (b = -15.84,
-SE = 113.98, z = -0.14 , p = 0.89). We found the same pattern of results
-for the OpenPose face detections, omitted here for brevity. We also
-found the same pattern with respect to infant's visual access to hands
-(see Table 4); including interaction terms also eliminated any main
-effect of age on the proportion of hands detected (b = -15.7, SE =
-113.98, z = -0.14, p = 0.89). Thus, these results suggest that infants'
-access to social information is heavily influenced by their postural and
-locomotive development.
+SE = 113.98, z = -0.14 , p = 0.89). We also found the same pattern with
+respect to infant's visual access to hands (see Table 4); including
+interaction terms also eliminated any main effect of age on the
+proportion of hands detected (b = -15.7, SE = 113.98, z = -0.14, p =
+0.89). Thus, these results suggest that infants' access to social
+information is heavily influenced by their postural and locomotive
+developments.
 
@@
 
-Finally, we analyzed how face and wrist detections changed during object
-labeling events as a function of infant's posture and orientation.
-Specfically, we analyzed a four-second window around each labeling event
-(e.g., ``Look at the {[}zem{]}!''); these labeling events were
-hand-annotated and automatically synchronized with the frame-by-frame
-face detections. We found that infants' posture and orientations
-impacted the degree to which they saw their caregiver's face and wrist
-during a labeling event; infants who were sitting or standing were more
-likely to have access to this social information. However, this did not
-manifest in a direct change the presence of wrists/hands before
-vs.~after naming events (average difference in proportion of wrists
-detected, 8 m.o. = -0.006, 12 m.o. = -0.013, 16 m.o. = -0.007) or in
-proportion of faces detected (average difference in proportion of faces
-detected, 8 m.o. = -0.002, 12 m.o. = 0, 16 m.o. = -0.004).
+Finally, we analyzed how face and wrist/hand detections changed during
+object labeling events as a function of infant's posture and
+orientation. Specfically, we analyzed a four-second window around each
+labeling event (e.g., ``Look at the {[}zem{]}!''); these labeling events
+were hand-annotated and automatically synchronized with the
+frame-by-frame face detections. We found that infants' posture and
+orientations impacted the degree to which they saw their caregiver's
+face and wrist during a labeling event; infants who were sitting or
+standing were more likely to have access to this social information.
+However, this did not manifest in a direct change the presence of
+wrists/hands before vs.~after naming events (average difference in
+proportion of wrists detected, 8 m.o. = -0.006, 12 m.o. = -0.013, 16
+m.o. = -0.007) or in proportion of faces detected (average difference in
+proportion of faces detected, 8 m.o. = -0.002, 12 m.o. = 0, 16 m.o. =
+-0.004).
 
@@
 and that both of these factors dramatically impacted the proportion of
-faces and hands that were available in the child's visual field. Thus,
-infants' postural and locomotive developments are mediating factors that
-explain some of the age-related changes in the proportion of faces and
-wrists/hands that are visually available to infants. Broadly, this work
-suggests that motoric developments mediate how infants experience their
-visual world and the social information in it: infants that are sitting
-and standing have a different view of their world, the people in it, and
-the actions that are being performed.
+faces and wrists/hands that were available in the child's visual field.
+Thus, infants' postural and locomotive developments are mediating
+factors that explain some of the age-related changes in the proportion
+of faces and wrists/hands that are visually available to infants.
+Broadly, this work suggests that motoric developments mediate how
+infants experience their visual world and the social information in it:
+infants that are sitting and standing have a different view of their
+world, the people in it, and the actions that are being performed.
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
redoing naming anlayses

@@
 
+###
 aroundNamings_Posture_Wrists <- d %>%
@@
 
-```{r echo=FALSE, include=FALSE}
-## Break down by before/after naming
-### First for wrists
-beforeNamings_Wrists <- d %>%
+```{r include=FALSE, echo=FALSE}
+# Compute proportion of faces/hands during naming events relative to baseline
+## Faces 
+faces_baseline<- d %>%
+  group_by(subid, age.grp) %>%
+  summarize(baselineDetections = mean(faceMT))
+#
+faces_naming <- d %>%
+  group_by(subid) %>%
+  mutate(detections = faceMT) %>%
+  do(summarize.naming(.)) %>%
+  group_by(subid, age.grp)  %>%
+  summarize(namingDetections = mean(detections,na.rm=TRUE)) %>% 
+  left_join(faces_baseline) %>%
+  mutate(baselineDiff = namingDetections - baselineDetections)  %>%
+  group_by(age.grp) %>%
+  summarize(avgBaselineDiff = mean(baselineDiff))
+## Wrists
+wrists_baseline<- d %>%
+  group_by(subid, age.grp) %>%
+  summarize(baselineDetections = mean(faceMT))
+#
+wrists_naming <- d %>%
   group_by(subid) %>%
   mutate(detections = wristOP) %>%
-  do(summarize.naming(.,window = c(-2,0))) %>%
+  do(summarize.naming(.)) %>%
   group_by(subid, age.grp)  %>%
-  summarise(avgDetections = mean(detections, na.rm=TRUE))  %>%
-  mutate(namingTime = "beforeNaming")
-
-namings_Wrists <- d %>%
-  group_by(subid) %>%
-  mutate(detections = wristOP) %>%
-  do(summarize.naming(.,window = c(0,2))) %>%
-  group_by(subid, age.grp)  %>%
-  summarise(avgDetections = mean(detections, na.rm=TRUE)) %>%
-  mutate(namingTime = "afterNaming") %>%
-  bind_rows(beforeNamings_Wrists)
-
-### Now for faces
-beforeNamings_Faces <- d %>%
-  group_by(subid) %>%
-  mutate(detections = faceMT) %>%
-  do(summarize.naming(.,window = c(-2,0))) %>%
-  group_by(subid, age.grp)  %>%
-  summarise(avgDetections = mean(detections, na.rm=TRUE))  %>%
-  mutate(namingTime = "beforeNaming")
-
-namings_Faces <- d %>%
-  group_by(subid) %>%
-  mutate(detections = faceMT) %>%
-  do(summarize.naming(.,window = c(0,2))) %>%
-  group_by(subid, age.grp)  %>%
-  summarise(avgDetections = mean(detections, na.rm=TRUE)) %>%
-  mutate(namingTime = "afterNaming") %>%
-  bind_rows(beforeNamings_Faces)
-
-# Plot if you want to look
-# p1=ggplot(namings_Faces, aes(x = factor(age.grp), y = avgDetections, col = age.grp)) +
-#   geom_boxplot() +
-#   ylab("Prop. faces") +
-#   xlab("") +
-#   theme_few() +
-#   theme(legend.position="none") +
-#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-#   facet_grid(~namingTime)
-# 
-# p2=ggplot(namings_Wrists, aes(x = factor(age.grp), y = avgDetections, col = age.grp)) +
-#   geom_boxplot() +
-#   ylab("Prop. wrists") +
-#   xlab("") +
-#   theme_few() +
-#   theme(legend.position="none") +
-#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-#   facet_grid(~namingTime)
-
-# ggarrange(p1,p2) # not shown in .rmd
-
-## output for text below \
-naming_Wrists_Summary <-  namings_Wrists %>%
-  spread(namingTime, avgDetections) %>%
-  mutate(diffWrists = afterNaming - beforeNaming) %>%
+  summarize(namingDetections = mean(detections, na.rm=TRUE)) %>%
+  left_join(wrists_baseline) %>%
+  mutate(baselineDiff = namingDetections - baselineDetections)  %>%
   group_by(age.grp) %>%
-  summarize(mean = mean(diffWrists), sd=sd(diffWrists))
-
-naming_Faces_Summary <-  namings_Faces %>%
-  spread(namingTime, avgDetections) %>%
-  mutate(diffFaces = afterNaming - beforeNaming) %>%
-  group_by(age.grp) %>%
-  summarize(mean = mean(diffFaces), sd=sd(diffFaces)) # no room for SD maybe but still informative
+  summarize(avgBaselineDiff = mean(baselineDiff))
 
@@
 ## Access to Faces and Hands During Labeling Events
-Finally, we analyzed how face and wrist/hand detections changed during object labeling events as a function of infant's posture and orientation.
+Finally, we explored how face and wrist/hand detections changed during object labeling events as a function of infant's posture and orientation.
 Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
-We found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information.
-However, this did not manifest in a direct change the presence of wrists/hands before vs.
-after naming events (average difference in proportion of wrists detected, 8 m.o.
-= `r round(naming_Wrists_Summary$mean[1],3)`, 12 m.o.
-= `r round(naming_Wrists_Summary$mean[2],3)`, 16 m.o.
-= `r round(naming_Wrists_Summary$mean[3],3)`) or in proportion of faces detected  (average difference in proportion of faces detected, 8 m.o.
-= `r round(naming_Faces_Summary$mean[1],3)`, 12 m.o.
-= `r round(naming_Faces_Summary$mean[2],3)`, 16 m.o.
-= `r round(naming_Faces_Summary$mean[3],3)`).
+As before, we found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information.
+A full summary of the results can be seen in Figure \ref{fig:detByNaming}.
+However, we did not find that infants saw more faces or more wrists/hands during naming events relative to baseline (average difference in proportion of wrists detected, 8 m.o.
+= `r round(wrists_naming$avgBaselineDiff[1],3)`, 12 m.o.
+= `r round(wrists_naming$avgBaselineDiff[2],3)`, 16 m.o.
+= `r round(wrists_naming$avgBaselineDiff[3],3)`;  average difference in proportion of faces detected, 8 m.o.
+= `r round(faces_naming$avgBaselineDiff[1],3)`, 12 m.o.
+= `r round(faces_naming$avgBaselineDiff[2],3)`, 16 m.o.
+= `r round(faces_naming$avgBaselineDiff[3],3)`).
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
fix typos

@@
 We used this as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
-We next capitalized tested the performance of two face detectors that both made use of recently developed Convolutional Neural Networks (CNNs) to extract face information.
-The first algorithm was specifically optimized for face detection, and the second algorithm was this to extract information about the position of 18 different body parts.
+We next tested the performance of two face detectors that both made use of recently developed Convolutional Neural Networks (CNNs) to extract face information.
+The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts.
 For the second algorithm (called OpenPose) [@cao2017realtime], we used the agent's nose (one of the 18 body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
@@
 This detector provided information about the presence of a face as well as its size and position.
-The second algorithm (based on  work by @zhang2016) using multi-task cascaded convolutional neural networks (MTCNNs).
+The second algorithm (based on  work by @zhang2016) uses multi-task cascaded convolutional neural networks (MTCNNs).
 The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
@@
 
-The CNN-based pose detector (OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.)  avaliable at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The CNN-based pose detector (OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.) available at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
 The candidates are then matched to a single individual and finally assembled into a pose; here, we only made use of the body parts relevant to the face and hands (nose and wrists).
-We operationalized face detections as any frames containing a nose, and hand detections as any frames containing either the left orright wrist.
+We operationalized face detections as any frames containing a nose, and hand detections as any frames containing either the left or right wrist.
 
@@
 To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
-A summary of the coefficients of a model with only main effects (and no interactions) model can be found in Table 3.
+A summary of the coefficients of a model with only main effects (and no interactions) can be found in Table 3.
 When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
fix typos

@@
 
-Finally, we analyzed how face and wrist/hand detections changed during
+Finally, we explored how face and wrist/hand detections changed during
 object labeling events as a function of infant's posture and
@@
 were hand-annotated and automatically synchronized with the
-frame-by-frame face detections. We found that infants' posture and
-orientations impacted the degree to which they saw their caregiver's
-face and wrist during a labeling event; infants who were sitting or
-standing were more likely to have access to this social information.
-However, this did not manifest in a direct change the presence of
-wrists/hands before vs.~after naming events (average difference in
-proportion of wrists detected, 8 m.o. = -0.006, 12 m.o. = -0.013, 16
-m.o. = -0.007) or in proportion of faces detected (average difference in
-proportion of faces detected, 8 m.o. = -0.002, 12 m.o. = 0, 16 m.o. =
--0.004).
+frame-by-frame face detections. As before, we found that infants'
+posture and orientations impacted the degree to which they saw their
+caregiver's face and wrist during a labeling event; infants who were
+sitting or standing were more likely to have access to this social
+information. A full summary of the results can be seen in Figure
+\ref{fig:detByNaming}. However, we did not find that infants saw more
+faces or more wrists/hands during naming events relative to baseline
+(average difference in proportion of wrists detected, 8 m.o. = 0, 12
+m.o. = -0.003, 16 m.o. = -0.014; average difference in proportion of
+faces detected, 8 m.o. = 0.003, 12 m.o. = 0.01, 16 m.o. = 0.021).
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
figure text

@@
 ```
-```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=4, fig.cap = "Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
+```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=4, fig.cap = "Proportion face and wrist/hand detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
 
@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture."}
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist/hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture."}
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
add conf intervals

@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist/hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture."}
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist/hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
+Error bars represent non-parametric bootstrapped 95% confidence intervals."}
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
caption + pdf

--- 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
@@
 ```
-```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=4, fig.cap = "Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
+```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=4, fig.cap = "Proportion face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
 
@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture."}
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture."}
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
caption + pdf

@@
 in some situations, it is notoriously bad at dealing with occluded faces
-(Scheirer, Anthony, Nakayama, \& Cox, 2014). We next capitalized tested
-the performance of two face detectors that both made use of recently
+(Scheirer, Anthony, Nakayama, \& Cox, 2014). We next tested the
+performance of two face detectors that both made use of recently
 developed Convolutional Neural Networks (CNNs) to extract face
 information. The first algorithm was specifically optimized for face
-detection, and the second algorithm was this to extract information
+detection, and the second algorithm was optimized to extract information
 about the position of 18 different body parts. For the second algorithm
@@
 of a face as well as its size and position. The second algorithm (based
-on work by K. Zhang et al. (2016)) using multi-task cascaded
+on work by K. Zhang et al. (2016)) uses multi-task cascaded
 convolutional neural networks (MTCNNs). The system was built using a
@@
 provided the locations of 18 body parts (ears, nose, wrists, etc.)
-avaliable at
+available at
 \url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
@@
 We operationalized face detections as any frames containing a nose, and
-hand detections as any frames containing either the left orright wrist.
+hand detections as any frames containing either the left or right wrist.
 
@@
 
-\caption[Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant]{Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant.}\label{fig:detByPosOrient}
+\caption[Proportion face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant]{Proportion face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant.}\label{fig:detByPosOrient}
 \end{figure*}
@@
 independent variables. A summary of the coefficients of a model with
-only main effects (and no interactions) model can be found in Table 3.
-When we did include interaction terms between age, posture, and
-orientation, age no longer remained a significant predictor (b = -15.84,
-SE = 113.98, z = -0.14 , p = 0.89). We also found the same pattern with
-respect to infant's visual access to hands (see Table 4); including
-interaction terms also eliminated any main effect of age on the
-proportion of hands detected (b = -15.7, SE = 113.98, z = -0.14, p =
-0.89). Thus, these results suggest that infants' access to social
-information is heavily influenced by their postural and locomotive
-developments.
+only main effects (and no interactions) can be found in Table 3. When we
+did include interaction terms between age, posture, and orientation, age
+no longer remained a significant predictor (b = -15.84, SE = 113.98, z =
+-0.14 , p = 0.89). We also found the same pattern with respect to
+infant's visual access to hands (see Table 4); including interaction
+terms also eliminated any main effect of age on the proportion of hands
+detected (b = -15.7, SE = 113.98, z = -0.14, p = 0.89). Thus, these
+results suggest that infants' access to social information is heavily
+influenced by their postural and locomotive developments.
 
@@
 
-\caption[Proportion face detections around a naming instance ('Look, a Zem']{Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.}\label{fig:detByNaming}
+\caption[Proportion face and wrist detections around a naming instance ('Look, a Zem']{Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.}\label{fig:detByNaming}
 \end{figure}

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd

@@
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture."}
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
+Error bars represent non-parametric bootstrapped 95% confidence intervals."}
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
fix dumb % sign
quotation

@@
 When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
-We also found the same pattern with respect to infant's visual access to hands (see Table 4); including interaction terms also eliminated any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
+We also found the same pattern with respect to infants' visual access to hands (see Table 4); including interaction terms also eliminated any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
 Thus, these results suggest that infants' access to social information is heavily influenced by their postural and locomotive developments.

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
changes

@@
     \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
-    \And {\large \bf Ally Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University
+    \And {\large \bf Allison M. Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University
     \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}
@@
 abstract: 
-  "The ability to process social cues--including eye gaze--is a critical component of children's early language and cognitive development.
+  "The ability to process social information--including eye gaze--is a critical component of children's early language and cognitive development.
 However, as children reach their first birthday, they begin to locomote themselves, walking and exploring their visual environment in an entirely new way.
@@
 ```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
-Error bars represent non-parametric bootstrapped 95% confidence intervals."}
+Error bars represent non-parametric bootstrapped 95 percent confidence intervals."}
 

### 2018-01-29 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
changes

@@
 -0.14 , p = 0.89). We also found the same pattern with respect to
-infant's visual access to hands (see Table 4); including interaction
+infants' visual access to hands (see Table 4); including interaction
 terms also eliminated any main effect of age on the proportion of hands
@@
 
-\caption[Proportion face and wrist detections around a naming instance ('Look, a Zem']{Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.}\label{fig:detByNaming}
+\caption[Proportion face and wrist detections around a naming instance ('Look, a Zem']{Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
+Error bars represent non-parametric bootstrapped 95 percent confidence intervals.}\label{fig:detByNaming}
 \end{figure}

### 2018-01-30 writeup-cogsci/helper.R
note to helper function

@@ -109,7 +109,7 @@ regularize.naming <- function (n) {
   
   # make the times seconds since onset - # if excel output is HMS vs MS - ugh! this was causing a big bug in analyses prior to jan 2018
   timeBefore=n$time
-  if (nchar(timeBefore[1])>8) {
+  if (nchar(timeBefore[1])>8) { ## hacky but works, sorry! will try to prettify soon.
     n$time=parse_date_time(timeBefore,"%H:%M:OS%")
   }
   else {
@@ -117,8 +117,7 @@ regularize.naming <- function (n) {
   }
   assert_that(sum(hour(n$time))==0)  # check there are no hours!all sessions <20 minutes.
   n$time <- minute(n$time)*60+second(n$time) # convert to seconds
-  test=c(n$time[1], timeBefore[1])
-  
+
   n <- subset(n,!is.na(name))
   n$naming.instance <- naming.instance(n$name)
   return(n)

### 2018-01-30 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd

@@
 These motor changes have significant consequences for what children see; crawling and walking infants simply have different views of the world.
-For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
+For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011head]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
 
@@
 ## Head-mounted camera
-```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for two different headcam configurations (we used the one below in our current study)."}
+```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for two different headcam configurations (we used the lower in our current study)."}
 img <- png::readPNG("images/viewangle_cut.png")

### 2018-01-30 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex

@@
 
-\author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Ally Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University \And {\large \bf Michael C.
+\author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Allison M.
+Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University \And {\large \bf Michael C.
 Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}
@@
 \begin{abstract}
-The ability to process social cues--including eye gaze--is a critical
-component of children's early language and cognitive development.
-However, as children reach their first birthday, they begin to locomote
-themselves, walking and exploring their visual environment in an
-entirely new way. How do these postural and locomotive changes affect
-children's access to the social information relevant for word-learning?
-Here, we explore this question by using head-mounted cameras to record
-infants' (8-16 months of age) egocentric visual perspective and use
-computer vision algorithms to detect the proportion of faces and wrists
-in infants' environments. We find that infants' posture and orientation
-to their caregiver largely mediates their access to social information,
-suggesting that postural and locomotive developments play a significant
-role in the emergence of children's linguistic and social capacities.
-Broadly, we suggest that the combined use of head-mounted cameras and
-the application of new computer vision techniques is a promising avenue
-for understanding the statistics of infants' visual and linguistic
-experience.
+The ability to process social information--including eye gaze--is a
+critical component of children's early language and cognitive
+development. However, as children reach their first birthday, they begin
+to locomote themselves, walking and exploring their visual environment
+in an entirely new way. How do these postural and locomotive changes
+affect children's access to the social information relevant for
+word-learning? Here, we explore this question by using head-mounted
+cameras to record infants' (8-16 months of age) egocentric visual
+perspective and use computer vision algorithms to detect the proportion
+of faces and wrists in infants' environments. We find that infants'
+posture and orientation to their caregiver largely mediates their access
+to social information, suggesting that postural and locomotive
+developments play a significant role in the emergence of children's
+linguistic and social capacities. Broadly, we suggest that the combined
+use of head-mounted cameras and the application of new computer vision
+techniques is a promising avenue for understanding the statistics of
+infants' visual and linguistic experience.
 
@@
 example, during spontaneous play in a laboratory playroom, toddlers are
-more likely to look at the floor while crawling than while walking (J.
-Franchak, Kretch, Soska, \& Adolph, 2011); in general, walking infants
+more likely to look at the floor while crawling than while walking
+(Franchak, Kretch, Soska, \& Adolph, 2011); in general, walking infants
 tend to have full visual access to their environment and the people in
-it, while crawling infants do not (K. S. Kretch, Franchak, \& Adolph,
-2014).
+it, while crawling infants do not (Kretch, Franchak, \& Adolph, 2014).
 
@@
 walking children are extremely different from the adult perspective (and
-not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
-Yu, \& Smith, 2017; J. Franchak et al., 2011; Yoshida \& Smith, 2008).
-While children's views tend to be more restricted and dominated by
-objects and hands (Yoshida \& Smith, 2008), both computational and
-empirical work suggest that this restricted viewpoint may be more
-effective for learning objects and their labels than the comparable
-adult perspective (Bambach, Crandall, Smith, \& Yu, 2017; D. Yurovsky,
-Smith, \& Yu, in press). Further, recent work also suggests dramatic
-changes in the child's perspective over the first two years of life, as
-views transition from primarily containing close up views of faces to
+not easily predicted by our own adult intuitions) ({\textbf{???}};
+Clerkin, Hart, Rehg, Yu, \& Smith, 2017; Yoshida \& Smith, 2008). While
+children's views tend to be more restricted and dominated by objects and
+hands (Yoshida \& Smith, 2008), both computational and empirical work
+suggest that this restricted viewpoint may be more effective for
+learning objects and their labels than the comparable adult perspective
+(Bambach, Crandall, Smith, \& Yu, 2017; D. Yurovsky, Smith, \& Yu, in
+press). Further, recent work also suggests dramatic changes in the
+child's perspective over the first two years of life, as views
+transition from primarily containing close up views of faces to
 capturing views of hands paired with the objects they are acting on
@@
 
-\hypertarget{ref-franchak2011}{}
-Franchak, J., Kretch, K., Soska, K., \& Adolph, K. (2011). Head-mounted
-eye tracking: A new method to describe infant looking. \emph{Child
-Development}.
+\hypertarget{ref-franchak2011head}{}
+Franchak, J. M., Kretch, K. S., Soska, K. C., \& Adolph, K. E. (2011).
+Head-mounted eye tracking: A new method to describe infant looking.
+\emph{Child Development}, \emph{82}(6), 1738--1750.
 

### 2018-01-30 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
importing all intro text from google doc & reformatting!

@@
 
-Children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like] and attend to the social information in their environment from their earliest days.
+From early infancy, children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like] and attend to the social information in their environment.
 Even newborns tend to prefer to look at faces with direct vs.
 averted gaze [@farroni2002eye] and young infants follow overt gaze shifts  [@gredeback2008microstructure; @gredeback2010development].
-Further, when infants free-view videos, they tend to look mostly at faces at the expense of other visual information--though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
+Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
 
-However, as children are learning from others around them, their view of the world is also radically changing  [@adolph2007].
+Then, however, their view of the world radically changes [@adolph2007].
 Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently.
 These motor changes have significant consequences for what children see; crawling and walking infants simply have different views of the world.
-For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011head]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
+For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
 
-One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities [@iverson2010].
+One possibility is that these motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities [@iverson2010].
 Indeed, these postural changes also impact how children interact with their mothers; walking (vs.
@@
 Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
-On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access social information (e.g., faces, gaze)  and in turn to accelerate their own learning.
+On their account, children’s ability to stand and independently locomote may change their ability to access social information (e.g., faces, gaze)  and in turn to accelerate their own learning.
 
 Recent technological developments allow for testing of this hypothesis by documenting the experiences of infants and children from their own perspective.
-By using head-mounted cameras, researchers have begun to record the visual experiences of infants and children-- which even for walking children are extremely different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008].
-While children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], both computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2012].
-Further, recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
+By using head-mounted cameras, researchers have begun to record the visual experiences of infants and children -- which even for walking children are extremely different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008].
+Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2012].
+Recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
 
@@
 Demographics and exclusion rates are given in Table \ref{tab:pop}.
-\begin{table}[H]
+\begin{table}[ht]
 \centering
@@
 ## Procedure
-First, all parents signed consent documents in a waiting room where children were fitted with the headcam.
-After the child was comfortable in the waiting room and with the experimenter, the experimenter placed the headcam on the child's head.
+First, all parents signed consent documents while children were fitted with the headcam.
 If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam [@yoshida2008].
@@
 ### Posture and Orientation Annotation
-We created a set of custom annotations that described the child's physical posture (e.g.
-standing) and the orientation of the caregiver relative to the child (e.g.
+We created custom annotations to describe the child's physical posture (e.g.
+standing) and the orientation of the child relative to the caregiver (e.g.
 far away).
 The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing.
-The caregiver's orientation was characterized as being close to the child, far from the child, and behind the child (independent of distance).
-For the first two annotations (close/far from the child), the caregiver could either be to the the front or to the side of the child.
-All annotations were made by a trained coder using the OpenSHAPA/Datavyu software  [@adolph2012], and times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.
+The caregiver's orientation was characterized as being close, far, and behind the child (independent of distance).
+For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child.
+All annotations were made by a trained coder using the OpenSHAPA/Datavyu software.
+[@adolph2012], and times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.
 
 ## Face and Hand Detection
-We used three face detection systems to measure infants' changing access to faces.
+We used three face detection systems to measure infants' access to faces.
 The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones.
@@
 The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts.
-For the second algorithm (called OpenPose) [@cao2017realtime], we used the agent's nose (one of the 18 body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
+For the second algorithm (called OpenPose) [@cao2017realtime], we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
 
-The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for the presence of a hand.
-We did so first because we did not capture the entire visual field experienced by the infants/children.
-Thus, even though the video may only show the presence of a wrist, the caregiver's hand may be within the child's field of view.
-Second, we did so because hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
-For example, if a caregiver was holding a toy and presenting it to the infant, their wrists may be visible but their hand may not necessarily be (as it is occluded by the toy).
-
+The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for hands for two reasons.
+First, as we did not capture childrens entire visual field.
+the presence of a wrist is likely often indicative of the presence of a hand within the field of view.
+Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
 ### Algorithms
 The first face detection system made use of a series of Haar feature-based cascade classifiers [@viola2004robust] applied to each individual frame.
-This detector provided information about the presence of a face as well as its size and position.
-The second algorithm (based on  work by @zhang2016) uses multi-task cascaded convolutional neural networks (MTCNNs).
-The system was built using a novel cascaded CNN-based framework for joint detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
-We used a Tensorflow implementation of this algorithm provided by (https://github.com/davidsandberg/facenet).
-Like Viola-Jones, this detector provided information about the presence of a face as well as its size and position.
+The second algorithm (based on work by @zhang2016) uses multi-task cascaded convolutional neural networks (MTCNNs)  for joint face detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
+We used a Tensorflow implementation of this algorithm avaliable at https://github.com/davidsandberg/facenet.
 
-The CNN-based pose detector (OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.) available at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The CNN-based pose detector (OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
@@
 ### Detector evaluation
-To evaluate face detector performance, we constructed an unbiased "gold set" of labeled frames that accounted for the relatively rare appearance of faces in the dataset.
-We thus hand-labeled two samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
+To evaluate face detector performance, we hand-labeled  an unbiased "gold set" of labeled frames.
+To account for the relatively rare appearance of faces in the dataset, we thus h abeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
 Each sample was comprised of an equal number of frames taken from each child's video.
-Faces were classifed as present in a frame if at least half of the face was showing.
+For wrist detections, the "gold set" was constructed in the same manner, except frames with a high density of wrists came only from detections made by OpenPose.
+Faces were classified as present if at least half of the face was showing; wrists were classified as present if any part of the wrist was showing.
 Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 2.
-For wrist detections, the "gold set" was constructed in the same manner, with the exception that frames with a high density of wrists were taken by pose detections made by OpenPose.
 
@@
 For face detection, MTCNN outperformed OpenPose when taking into account only the composite F-score (`r round(perfFacesOverall$f[1],2)` MTCNN vs.
-`r round(perfFacesOverall$f[2],2)` OpenPose); we thus use MTCNN detections in the following analyses.
+`r round(perfFacesOverall$f[2],2)` OpenPose).
 Although MTCNN and OpenPose performed comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
 ViolaJones performed quite poorly relative to the other detectors, especially with respect to the random sample.
+We thus use MTCNN detections in the following analyses.
 
-For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision, but low recall on the randomly sampled set of frames (see Table 2).
-We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of wrists overall in the datatset.
+For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision but low recall on the randomly sampled frames (see Table 2).
+We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of wrists in the dataset.
 
@@
 ## Changes in Access to Faces and Hands
-```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age.
-Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
-detectionsByAge <- dlong %>%
-  filter(detectorType!="OpenPose-Faces") %>% # for brevity
-  filter(detectorType!="ViolaJones-Faces") %>% # low performance
-  filter(!is.na(age.at.test)) %>%
-  group_by(subid, age.at.test, detectorType) %>%
-  summarise(detections = mean(detection)) %>%
-  left_join(inc_subs) %>%
-  mutate(len = as.numeric(len.min))
 
-ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
-    geom_jitter(width = .25) +
-    geom_smooth(size=1) +
-    scale_x_continuous(breaks=c(8,12,16)) +
-    scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
-    labs(y = "Prop. detected", x = element_blank()) +
-    theme_few() + 
-    theme(legend.position="none") +
-    facet_grid(~detectorType) 
-```
-```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=4, fig.cap = "Proportion face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
+We first examined the proportion of face and wrist detections across age; a summary can be seen in Figure \ref{fig:detByAge}.
+We observed a slight U-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slightly fewer faces faces than 8 or 16-month-olds.
 
-dlongPlot <-dlong %>%
-    filter(detectorType!="OpenPose-Faces") %>%
-    filter(detectorType!="ViolaJones-Faces")
-
-detectionsByPosture <- dlongPlot %>%
-  filter(!is.na(posture)) %>%
-  group_by(age.grp, posture, subid,detectorType) %>%
-  summarise(face = mean(detection)) %>%
-  group_by(age.grp)
-
-p1=ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
-  geom_boxplot(outlier.size = 0.5) + 
-  labs(y = "Prop. detected", x = element_blank()) +
-  facet_grid(~detectorType, scales="free_y") +
-  theme_few() +
-  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
-
-detectionsByOrientation <- dlongPlot %>%
-  filter(!is.na(orientation)) %>%
-  group_by(age.grp, orientation, subid,detectorType) %>%
-  summarise(detections = mean(detection)) %>%
-  group_by(age.grp)
-
-p2=ggplot(detectionsByOrientation, aes(x = factor(age.grp), y = detections, col = orientation)) + 
-  geom_boxplot(outlier.size = 0.5) + 
-  labs(y = "Prop. detected", x = element_blank()) +
-  facet_grid(~detectorType, scales="free_y") +
-  theme_few() +
-  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
-
-ggarrange(p1, p2)
-
-```
-
-We first examined the proportion of face and wrist detections across age; a full summary can be seen in Figure \ref{fig:detByAge}.
-When looking at face presence, we observed a slight U-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slighly fewer faces faces than 8 or 16-month-olds.
-
-However, we found that any age related effects were much smaller compared to the impact of postural and locomotive changes on children's visual access to faces.
+However, we found that any age related effects were much smaller compared to the impact of postural and locomotive changes on children's visual access to faces and wrists/hands.
 Children's posture was a major factor both in how many faces and wrists/hands they saw during the play session.
-Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces (Figure \ref{fig:detByPosOrient}, upper panel); this same pattern was also true for wrists/hand detectinos.
-We also examined how the child's orientation relative to their caregiver impacted their visual access to faces and hands.
-Children who were far away from their caregiver were more likely to see faces than children who were close to their caregiver; this was true within all age groups  (Figure \ref{fig:detByPosOrient}, lower panel).
+Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces (Figure \ref{fig:detByPosOrient}, upper panel); this same pattern was also true for wrists/hand detection.
+Children’s orientation also impacted their visual access to faces and hands.
+Children who were far away from their caregiver were more likely to see faces/hands than children who were close to their caregiver (Figure \ref{fig:detByPosOrient}, lower panel).
 
@@
 ```
+```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age.
+Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
+detectionsByAge <- dlong %>%
+  filter(detectorType!="OpenPose-Faces") %>% # for brevity
+  filter(detectorType!="ViolaJones-Faces") %>% # low performance
+  filter(!is.na(age.at.test)) %>%
+  group_by(subid, age.at.test, detectorType) %>%
+  summarise(detections = mean(detection)) %>%
+  left_join(inc_subs) %>%
+  mutate(len = as.numeric(len.min))
 
-```{r results="asis"}
-#print(faceGLMOut_noInt, type="latex", comment = F, table.placement = "H")
+ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
+    geom_jitter(width = .25) +
+    geom_smooth(size=1) +
+    scale_x_continuous(breaks=c(8,12,16)) +
+    scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
+    labs(y = "Prop. detected", x = element_blank()) +
+    theme_few() + 
+    theme(legend.position="none") +
+    facet_grid(~detectorType) 
 ```
 
+```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
+
+dlongPlot <-dlong %>%
+    filter(detectorType!="OpenPose-Faces") %>%
+    filter(detectorType!="ViolaJones-Faces")
+
+detectionsByPosture <- dlongPlot %>%
+  filter(!is.na(posture)) %>%
+  group_by(age.grp, posture, subid,detectorType) %>%
+  summarise(face = mean(detection)) %>%
+  group_by(age.grp)
+
+p1=ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
+  geom_boxplot(outlier.size = 0.5) + 
+  labs(y = "Prop. detected", x = element_blank()) +
+  facet_grid(~detectorType, scales="free_y") +
+  theme_few() +
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
+
+detectionsByOrientation <- dlongPlot %>%
+  filter(!is.na(orientation)) %>%
+  group_by(age.grp, orientation, subid,detectorType) %>%
+  summarise(detections = mean(detection)) %>%
+  group_by(age.grp)
+
+p2=ggplot(detectionsByOrientation, aes(x = factor(age.grp), y = detections, col = orientation)) + 
+  geom_boxplot(outlier.size = 0.5) + 
+  labs(y = "Prop. detected", x = element_blank()) +
+  facet_grid(~detectorType, scales="free_y") +
+  theme_few() +
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
+
+ggarrange(p1, p2)
+
+```
 \begin{table}[H]
@@
   \hline
+  \textbf{Faces} \\
   Intercept & -5.2469 & 0.0584 & -89.88 & 0.0000 \\ 
@@
   \hline
-\end{tabular}
-\caption{Model coefficients from a generalized linear model predicting the proportion of faces seen by infants.} 
-\end{table}
-
-```{r results="asis"}
-#print(wristGLMOut_noInt, type="latex", comment = F, table.placement = "H")
-```
-
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
   & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
   \hline
+  \textbf{Wrists} \\ 
   Intercept & -5.0818 & 0.0776 & -65.46 & 0.0000 \\ 
@@
 \end{tabular}
-\caption{Model coefficients from a generalized linear model predicting the proportion of wrists seen by infants.} 
+\caption{Model coefficients from generalized linear models predicting the proportion of faces (upper panel) and wrists (lower panel) seen by children.} 
 \end{table}
+
+To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
+A summary of the coefficients of a model with only main effects (and no interactions) can be found in Table 3.
+When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
+We also found the same pattern with respect to infants' visual access to hands (see Table 4); including interaction terms also eliminated any main effect of age on the proportion of wrists detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
+These results suggest that infants' visual access to social information is heavily influenced by their postural and locomotive developments.
+
+```{r results="asis"}
+#print(faceGLMOut_noInt, type="latex", comment = F, table.placement = "H")
+```
+
+```{r results="asis"}
+#print(wristGLMOut_noInt, type="latex", comment = F, table.placement = "H")
+```
 
@@
 ```
-
-To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
-A summary of the coefficients of a model with only main effects (and no interactions) can be found in Table 3.
-When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
-We also found the same pattern with respect to infants' visual access to hands (see Table 4); including interaction terms also eliminated any main effect of age on the proportion of hands detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
-Thus, these results suggest that infants' access to social information is heavily influenced by their postural and locomotive developments.
 
@@
 
+```{r include=FALSE, echo=FALSE}
+# Compute proportion of faces/hands during naming events relative to baseline
+## Faces 
+faces_baseline<- d %>%
+  group_by(subid, age.grp) %>%
+  summarize(baselineDetections = mean(faceMT))
+#
+faces_naming <- d %>%
+  group_by(subid) %>%
+  mutate(detections = faceMT) %>%
+  do(summarize.naming(.)) %>%
+  group_by(subid, age.grp)  %>%
+  summarize(namingDetections = mean(detections,na.rm=TRUE)) %>% 
+  left_join(faces_baseline) %>%
+  mutate(baselineDiff = namingDetections - baselineDetections)  %>%
+  group_by(age.grp) %>%
+  summarize(avgBaselineDiff = mean(baselineDiff))
+## Wrists
+wrists_baseline<- d %>%
+  group_by(subid, age.grp) %>%
+  summarize(baselineDetections = mean(faceMT))
+#
+wrists_naming <- d %>%
+  group_by(subid) %>%
+  mutate(detections = wristOP) %>%
+  do(summarize.naming(.)) %>%
+  group_by(subid, age.grp)  %>%
+  summarize(namingDetections = mean(detections, na.rm=TRUE)) %>%
+  left_join(wrists_baseline) %>%
+  mutate(baselineDiff = namingDetections - baselineDetections)  %>%
+  group_by(age.grp) %>%
+  summarize(avgBaselineDiff = mean(baselineDiff))
+
+```
+
+## Access to Faces and Hands During Labeling Events
 ```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
@@
 
-```{r include=FALSE, echo=FALSE}
-# Compute proportion of faces/hands during naming events relative to baseline
-## Faces 
-faces_baseline<- d %>%
-  group_by(subid, age.grp) %>%
-  summarize(baselineDetections = mean(faceMT))
-#
-faces_naming <- d %>%
-  group_by(subid) %>%
-  mutate(detections = faceMT) %>%
-  do(summarize.naming(.)) %>%
-  group_by(subid, age.grp)  %>%
-  summarize(namingDetections = mean(detections,na.rm=TRUE)) %>% 
-  left_join(faces_baseline) %>%
-  mutate(baselineDiff = namingDetections - baselineDetections)  %>%
-  group_by(age.grp) %>%
-  summarize(avgBaselineDiff = mean(baselineDiff))
-## Wrists
-wrists_baseline<- d %>%
-  group_by(subid, age.grp) %>%
-  summarize(baselineDetections = mean(faceMT))
-#
-wrists_naming <- d %>%
-  group_by(subid) %>%
-  mutate(detections = wristOP) %>%
-  do(summarize.naming(.)) %>%
-  group_by(subid, age.grp)  %>%
-  summarize(namingDetections = mean(detections, na.rm=TRUE)) %>%
-  left_join(wrists_baseline) %>%
-  mutate(baselineDiff = namingDetections - baselineDetections)  %>%
-  group_by(age.grp) %>%
-  summarize(avgBaselineDiff = mean(baselineDiff))
-
-```
-
-## Access to Faces and Hands During Labeling Events
 Finally, we explored how face and wrist/hand detections changed during object labeling events as a function of infant's posture and orientation.
-Specfically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
-As before, we found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrist during a labeling event; infants who were sitting or standing were more likely to have access to this social information.
+Specifically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
+As before, we found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrists during a labeling event; infants who were sitting or standing were more likely to have visual access to this social information.
 A full summary of the results can be seen in Figure \ref{fig:detByNaming}.
-However, we did not find that infants saw more faces or more wrists/hands during naming events relative to baseline (average difference in proportion of wrists detected, 8 m.o.
+However, we did not find that infants saw more faces or wrists/hands during naming events relative to baseline (average difference in proportion of wrists detected, 8 m.o.
 = `r round(wrists_naming$avgBaselineDiff[1],3)`, 12 m.o.
@@
 # General Discussion
-We used a head-mounted camera to explore how children's postural and locomotive development directly impacts their access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
-We found that children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically impacted the proportion of faces and wrists/hands that were available in the child's visual field.
-Thus, infants' postural and locomotive developments are mediating factors that explain some of the age-related changes in the proportion of faces and wrists/hands that are visually available to infants.
+We used a head-mounted camera to explore how children's postural and locomotive development impacts their visual access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
+Children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically influenced the proportion of faces and wrists/hands that were available in the child's visual field.
 Broadly, this work suggests that motoric developments mediate how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
@@
 This work also deploys novel advancements in computer vision to the study of developmental psychology.
-The field of object detection and recognition has advanced dramatically in the past five years since the re-birth of deep learning algorithms [@krizhevsky2012imagenet], creating a new generation of algorithmic tools.
+The field of object detection and recognition has advanced dramatically in the past five years, creating a new generation of algorithmic tools.
 These tools are substantially better equipped to deal with noisier, more complicated datasets and can extract richer and more detailed information.
-Videos from the infant perspective provided substantial challenges (e.g., partially occluded faces) for the classic models of face detection (e.g., ViolaJones) [@viola2004robust].
-Further, as the headcam technologies employed here were inexpensive and the computer vision algorithms freely available, this method is a promising avenue for quantifying the visual and social information available to infant learners.
+As the headcams we used were inexpensive and the computer vision algorithms freely available, this method is a promising avenue for quantifying the visual and social information available to infant learners.
 
-Broadly, we suggest that the combined use of these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
+We suggest that the combined use of these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
 

### 2018-01-30 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
incorporating of mike's remaining comments & a few trimming tries

@@
 
-From early infancy, children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like] and attend to the social information in their environment.
+From early infancy, children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like].
 Even newborns tend to prefer to look at faces with direct vs.
@@
 Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2012].
-Recent work also suggests dramatic changes in the child’s perspective over the first two years of life, as views transition from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
+This perspective changes over the first two years of life, as views transition from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
 
@@
 `r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
-`r sum(all_subs$reason=="infant crying")` for fussiness, and one excluded for sample symmetry.
+`r sum(all_subs$reason=="infant crying")` for fussiness, and one for sample symmetry.
 All inclusion decisions were made independent of the results of subsequent analyses.
-Some of these data were also used in @frank2013.
+Some of these data were also analyzed in @frank2013.
 
 ## Head-mounted camera
-```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Field of view for two different headcam configurations (we used the lower in our current study)."}
+```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}
 img <- png::readPNG("images/viewangle_cut.png")
@@
 Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
-A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, left).
+A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, bottom).
 
 Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's approximate vertical field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988].
-As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces.
+As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field)
 
 ## Procedure
-First, all parents signed consent documents while children were fitted with the headcam.
+All parents signed consent documents while children were fitted with the headcam.
 If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam [@yoshida2008].
-After the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session.
+When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session.
 Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the novel toys.
@@
 The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing.
-The caregiver's orientation was characterized as being close, far, and behind the child (independent of distance).
+The caregiver's orientation was characterized as being close, far, or behind the child (independent of distance).
 For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child.
-All annotations were made by a trained coder using the OpenSHAPA/Datavyu software.
-[@adolph2012], and times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.
+All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012].
+Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.
 
@@
 The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones.
-We used this as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
+We used this algorithm as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
 We next tested the performance of two face detectors that both made use of recently developed Convolutional Neural Networks (CNNs) to extract face information.
 The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts.
-For the second algorithm (called OpenPose) [@cao2017realtime], we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
+For the second algorithm [OpenPose, @cao2017realtime], we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
 
@@
 Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
+For example, if a caregiver was holding a toy andpresenting it to the infant, their wrists may be visible but theirhand may not necessarily be (as it is occluded by the toy).
+
 ### Algorithms
@@
 
-The CNN-based pose detector (OpenPose [@cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The CNN-based pose detector  [OpenPose, @cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
@@
 ### Detector evaluation
-To evaluate face detector performance, we hand-labeled  an unbiased "gold set" of labeled frames.
-To account for the relatively rare appearance of faces in the dataset, we thus h abeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
+To evaluate face detector performance, we hand-labeled  an "gold set" of labeled frames.
+To account for the relatively rare appearance of faces in the dataset, we  hand abeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
 Each sample was comprised of an equal number of frames taken from each child's video.
@@
 We thus use MTCNN detections in the following analyses.
-
 For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision but low recall on the randomly sampled frames (see Table 2).
-We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of wrists in the dataset.
+We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of hands in the dataset.
 
@@
 \end{tabular}
-\caption{Detector performance on both high density samples (where proportion of faces detected was high) and random samples (where frames were randomly selected).
+\caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected).
 P, R, and F denote precision, recall, and F-score, respectively.
@@
 ```
-We noted characteristic changes in infants' posture and orientation across this developmental time range.
 The proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with age.
-Both 8 month-olds and 12 month-olds spent equivalent amounts of time either lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time either sitting or standing (see Figure \ref{fig:posture}).
-We also observed characteristic changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:posture}).
+Both 8 month-olds and 12 month-olds spent equivalent amounts of time lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time sitting or standing (see Figure \ref{fig:posture}).
+We also observed changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:posture}).
 
@@
 
-We first examined the proportion of face and wrist detections across age; a summary can be seen in Figure \ref{fig:detByAge}.
-We observed a slight U-shaped function both when analyzing the output of the MTCNN and OpenPose detectors, such that 12-month-olds appeared to experience slightly fewer faces faces than 8 or 16-month-olds.
+We examined the proportion of face and wrist detections across age (see Figure \ref{fig:detByAge}).
+We observed a slight U-shaped function in face detections, such that 12-month-olds appeared to have visual acces to slightly fewer faces faces than 8 or 16-month-olds; conversely, wrist detections perhaps appeared to generally increase with age.
 
-However, we found that any age related effects were much smaller compared to the impact of postural and locomotive changes on children's visual access to faces and wrists/hands.
+Age related effects were much smaller than postural and locomotive changes on children's visual access to faces and wrists/hands.
 Children's posture was a major factor both in how many faces and wrists/hands they saw during the play session.
-Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces (Figure \ref{fig:detByPosOrient}, upper panel); this same pattern was also true for wrists/hand detection.
-Children’s orientation also impacted their visual access to faces and hands.
-Children who were far away from their caregiver were more likely to see faces/hands than children who were close to their caregiver (Figure \ref{fig:detByPosOrient}, lower panel).
+Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces (Figure \ref{fig:detByPosOrient}, upper panel); this same pattern was also true for wrists/hand detections.
+Children’s orientation also impacted their visual access to faces and hands: children who were far away from their caregiver were more likely to see faces/hands than children who were close to their caregiver (Figure \ref{fig:detByPosOrient}, lower panel).
 
@@
 When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
-We also found the same pattern with respect to infants' visual access to hands (see Table 4); including interaction terms also eliminated any main effect of age on the proportion of wrists detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
-These results suggest that infants' visual access to social information is heavily influenced by their postural and locomotive developments.
+We also found the same pattern with respect to infants' visual access to hands (see Table 3); including interaction terms also eliminated any main effect of age on the proportion of wrists detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
+Thus, these results suggest that infants' visual access to social information is mediated by their posture and orientation, which is in turn a function of their general locomotor development (as indexed by chronological age).
 
@@
 
-Finally, we explored how face and wrist/hand detections changed during object labeling events as a function of infant's posture and orientation.
-Specifically, we analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and automatically synchronized with the frame-by-frame face detections.
-As before, we found that infants' posture and orientations impacted the degree to which they saw their caregiver's face and wrists during a labeling event; infants who were sitting or standing were more likely to have visual access to this social information.
-A full summary of the results can be seen in Figure \ref{fig:detByNaming}.
-However, we did not find that infants saw more faces or wrists/hands during naming events relative to baseline (average difference in proportion of wrists detected, 8 m.o.
+Finally, we explored how face and wrist detections changed during object labeling events as a function of infant's posture and orientation.
+We analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and synchronized with the frame-by-frame face/wrist detections.
+We found that infants' posture and orientations mediated their visual access to faces and wrists/hands during labeling events; infants who were sitting or standing were more likely to have visual access to this social information (see Figure \ref{fig:detByNaming}).
+However, we did not find that infants saw particularily more faces or wrists/hands during naming events relative to baseline (avg.
+difference in proportion of wrists, 8 m.o.
 = `r round(wrists_naming$avgBaselineDiff[1],3)`, 12 m.o.
 = `r round(wrists_naming$avgBaselineDiff[2],3)`, 16 m.o.
-= `r round(wrists_naming$avgBaselineDiff[3],3)`;  average difference in proportion of faces detected, 8 m.o.
+= `r round(wrists_naming$avgBaselineDiff[3],3)`;  avg.
+difference in proportion of faces, 8 m.o.
 = `r round(faces_naming$avgBaselineDiff[1],3)`, 12 m.o.
@@
 Children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically influenced the proportion of faces and wrists/hands that were available in the child's visual field.
-Broadly, this work suggests that motoric developments mediate how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
+Broadly, this work suggests that motor development mediates how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
+More work is needed to understand how these results relate to the pattern of activities and social interctions observed in children's home experiences [@fausey2016].
 
@@
 The field of object detection and recognition has advanced dramatically in the past five years, creating a new generation of algorithmic tools.
-These tools are substantially better equipped to deal with noisier, more complicated datasets and can extract richer and more detailed information.
-As the headcams we used were inexpensive and the computer vision algorithms freely available, this method is a promising avenue for quantifying the visual and social information available to infant learners.
-
-We suggest that the combined use of these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
+These tools are substantially better equipped to deal with noisier, more complicated datasets and can extract richer and more detailed information than classic models of face detection (i.e., ViolaJones).
+As the headcams we used were inexpensive and the computer vision algorithms freely available,  suggest that the combined use of these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
 
@@
 
-Thanks to Kaia Simmons, Kathy Woo, Aditi Maliwal, and other members of the Language and Cognition Lab for help in recruitment, data collection, and annotation.
-This research was supported by a John Merck Scholars grant to MCF.
+Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation.
 An earlier version of this work was presented to the Cognitive Science Society in @frank2013.
-Please address correspondence to Michael C.
-Frank, Department of Psychology, Stanford University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel: (650) 724-4003, email: \texttt{mcfrank@stanford.edu}.
-
 # References

### 2018-01-30 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
rearranging text and cuts to make 6 pages - almost there!

@@
 author-information: > 
-    \author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University
-    \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
+    \author{{\large \bf Alessandro Sanchez*} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University
+    \And {\large \bf Bria Long*}  \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
     \And {\large \bf Allison M. Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University
@@
 # Introduction
+[^1]: *These authors contributed equally to this work.
 
-From early infancy, children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like].
+[^1] From early infancy, children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like].
 Even newborns tend to prefer to look at faces with direct vs.
-averted gaze [@farroni2002eye] and young infants follow overt gaze shifts  [@gredeback2008microstructure; @gredeback2010development].
+averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010development].
 Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
@@
 # Methods
-## Participants
+### Participants
 ```{r echo=FALSE, include=FALSE}
@@
 
-Our final sample consisted of 36 infants and children, with 12 participants in three age groups: 8 months (6 females), 12 months (7 females), and 16 months (6 females).
+Our final sample consisted of 36 infants and children, with 12 participants in three age groups: 8 months (6 F), 12 months (7 F), and 16 months (6 F).
 Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80\% English at home.
@@
 
-## Head-mounted camera
-```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=3, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}
+### Head-mounted camera
+```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}
 img <- png::readPNG("images/viewangle_cut.png")
@@
 We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband.
-Videos captured by the headcam were 720x480 pixels with 25 frames per second.
-Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
+Videos captured by the headcam were 720x480 pixels with 25 frames per second [^2].
 A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, bottom).
+
+[^2]: Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
 
@@
 
-## Procedure
+### Procedure
 All parents signed consent documents while children were fitted with the headcam.
@@
 
-## Data Processing and Annotation
-
-\begin{figure*}
-\centering
-\includegraphics[width=6in]{images/detector_samples_banner.pdf}
-\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
-The last column features a false positive from OpenPose and a false negative from MTCNN.}
-\end{figure*}
-
+### Data Processing and Annotation
 Headcam videos were trimmed such that they excluded the instruction phase when the experimenter was in the room and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software.
@@
 Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
-For example, if a caregiver was holding a toy andpresenting it to the infant, their wrists may be visible but theirhand may not necessarily be (as it is occluded by the toy).
+\begin{figure*}
+\centering
+\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
+\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
+The last column features a false positive from OpenPose and a false negative from MTCNN.}
+\end{figure*}
 
@@
 # Results
-First, we report developmental shifts in infants' posture and their orientation relative to their caregiver.
+<!-- First, we report developmental shifts in infants' posture and their orientation relative to their caregiver.
 Then, we explore how these changes influence children's visual access to faces and wrists/hands across this developmental time range.
 Finally, we explore how these changes impact the accessibility of faces and wrists/hands during labeling events.
-
-## Changes in Posture and Orientation
+-->
 
@@
 
-```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3.5, fig.height=4, fig.cap = "Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver." }
+```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3.5, fig.height=3.5, fig.cap = "Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver." }
 p1=ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
@@
 ```
+## Changes in Posture and Orientation
 The proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with age.
@@
 ## Changes in Access to Faces and Hands
-
 We examined the proportion of face and wrist detections across age (see Figure \ref{fig:detByAge}).
@@
 ```
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-  \textbf{Faces} \\
-  Intercept & -5.2469 & 0.0584 & -89.88 & 0.0000 \\ 
-  Age & 0.0847 & 0.0041 & 20.89 & 0.0000 \\ 
-  Prone & 0.2015 & 0.0564 & 3.57 & 0.0004 \\ 
-  Sit & 1.4053 & 0.0541 & 25.98 & 0.0000 \\ 
-  Stand & 1.4272 & 0.0542 & 26.33 & 0.0000 \\ 
-  Close & 1.8239 & 0.0230 & 79.17 & 0.0000 \\ 
-  Far & 2.5479 & 0.0239 & 106.42 & 0.0000 \\ 
-  \hline
-  & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-  \textbf{Wrists} \\ 
-  Intercept & -5.0818 & 0.0776 & -65.46 & 0.0000 \\ 
-  Age & 0.0564 & 0.0050 & 11.35 & 0.0000 \\ 
-  Prone & 0.9499 & 0.0774 & 12.27 & 0.0000 \\ 
-  Sit & 1.7282 & 0.0758 & 22.79 & 0.0000 \\ 
-  Stand & 1.6618 & 0.0760 & 21.88 & 0.0000 \\ 
-  Close & 0.7472 & 0.0188 & 39.81 & 0.0000 \\ 
-  Far & 1.6357 & 0.0200 & 81.61 & 0.0000 \\ 
-   \hline
-\end{tabular}
-\caption{Model coefficients from generalized linear models predicting the proportion of faces (upper panel) and wrists (lower panel) seen by children.} 
-\end{table}
-
-To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
-A summary of the coefficients of a model with only main effects (and no interactions) can be found in Table 3.
-When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
-We also found the same pattern with respect to infants' visual access to hands (see Table 3); including interaction terms also eliminated any main effect of age on the proportion of wrists detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
-Thus, these results suggest that infants' visual access to social information is mediated by their posture and orientation, which is in turn a function of their general locomotor development (as indexed by chronological age).
-
-```{r results="asis"}
-#print(faceGLMOut_noInt, type="latex", comment = F, table.placement = "H")
-```
-
-```{r results="asis"}
-#print(wristGLMOut_noInt, type="latex", comment = F, table.placement = "H")
-```
-
-```{r echo=FALSE, include=FALSE}
-## Mixed effect model tries - many convergence issues
-
- # Interactions fail to converge, as does the one below 
- # fullModelminsInt_Faces_MTCNN <- glmer(faceMT ~ scale(age.at.test)+(posture)+(orientation) + (1|subid), 
- #        data = d, 
- #        family = "binomial")
-
-## Reading https://rdrr.io/cran/lme4/man/convergence.html 
-## Check for singularity - not an issue 
-# fm1=fullModelminsInt_Faces_MTCNN 
-# diag.vals <- getME(fm1,"theta")[getME(fm1,"lower") == 0] 
-# any(diag.vals < 1e-6) # FALSE
-
-## But has issues with every optimizer... -->
-#source(system.file("utils", "allFit.R", package="lme4")) 
-#fm1.all <- allFit(fm1)
-#ss <- summary(fm1.all)
-
-#No convergence issues with open pose detectors for faces, but for wrists
-#fullModelminsInt_Faces_OP <- glmer(faceOP ~ scale(age.at.test) + posture + orientation + (1|subid), 
-#      data = d, 
-#        family = "binomial") 
-```
 
@@
 
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+  \textbf{Faces} \\
+  Intercept & -5.2469 & 0.0584 & -89.88 & 0.0000 \\ 
+  Age & 0.0847 & 0.0041 & 20.89 & 0.0000 \\ 
+  Prone & 0.2015 & 0.0564 & 3.57 & 0.0004 \\ 
+  Sit & 1.4053 & 0.0541 & 25.98 & 0.0000 \\ 
+  Stand & 1.4272 & 0.0542 & 26.33 & 0.0000 \\ 
+  Close & 1.8239 & 0.0230 & 79.17 & 0.0000 \\ 
+  Far & 2.5479 & 0.0239 & 106.42 & 0.0000 \\ 
+  \hline
+  & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+  \textbf{Wrists} \\ 
+  Intercept & -5.0818 & 0.0776 & -65.46 & 0.0000 \\ 
+  Age & 0.0564 & 0.0050 & 11.35 & 0.0000 \\ 
+  Prone & 0.9499 & 0.0774 & 12.27 & 0.0000 \\ 
+  Sit & 1.7282 & 0.0758 & 22.79 & 0.0000 \\ 
+  Stand & 1.6618 & 0.0760 & 21.88 & 0.0000 \\ 
+  Close & 0.7472 & 0.0188 & 39.81 & 0.0000 \\ 
+  Far & 1.6357 & 0.0200 & 81.61 & 0.0000 \\ 
+   \hline
+\end{tabular}
+\caption{Model coefficients from generalized linear models predicting the proportion of faces (upper panel) and wrists (lower panel) seen by children.} 
+\end{table}
+
+To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
+A summary of the coefficients of a model with only main effects (and no interactions) can be found in Table 3.
+When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
+We also found the same pattern with respect to infants' visual access to hands (see Table 3); including interaction terms also eliminated any main effect of age on the proportion of wrists detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
+Thus, these results suggest that infants' visual access to social information is mediated by their posture and orientation, which is in turn a function of their general locomotor development (as indexed by chronological age).
+
 ## Access to Faces and Hands During Labeling Events
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
-Error bars represent non-parametric bootstrapped 95 percent confidence intervals."}
-
-p1=ggplot(aroundNamings_Posture_Faces, aes(x = posture, y = mean, col = posture)) +
-  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
-  ylab("Prop. faces") +
-  xlab("") +
-  ylim(0,.25) +
-  theme_few() + 
-  theme(legend.position="none") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-p2=ggplot(aroundNamings_Orient_Faces, aes(x = orientation, y = mean, col = orientation)) +
-  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
-  ylab("Prop. faces") +
-  xlab("") +
-  ylim(0,.25) +
-  theme_few() +
-  theme(legend.position="none")  +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-p3=ggplot(aroundNamings_Posture_Wrists, aes(x = posture, y = mean, col = posture)) +
-  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
-  ylab("Prop. wrists") +
-  xlab("") +
-  ylim(0,.25) +
-  theme_few() +
-  theme(legend.position="none")  +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-p4=ggplot(aroundNamings_Orient_Wrists, aes(x = orientation, y = mean, col = orientation)) +
-  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
-  ylab("Prop. wrists") +
-  xlab("") +
-  ylim(0,.25) +
-  theme_few() +
-  theme(legend.position="none") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-ggarrange(p1,p2, p3, p4)
-```
 
@@
 
+```{r results="asis"}
+#print(faceGLMOut_noInt, type="latex", comment = F, table.placement = "H")
+```
+
+```{r results="asis"}
+#print(wristGLMOut_noInt, type="latex", comment = F, table.placement = "H")
+```
+
+```{r echo=FALSE, include=FALSE}
+## Mixed effect model tries - many convergence issues
+
+ # Interactions fail to converge, as does the one below 
+ # fullModelminsInt_Faces_MTCNN <- glmer(faceMT ~ scale(age.at.test)+(posture)+(orientation) + (1|subid), 
+ #        data = d, 
+ #        family = "binomial")
+
+## Reading https://rdrr.io/cran/lme4/man/convergence.html 
+## Check for singularity - not an issue 
+# fm1=fullModelminsInt_Faces_MTCNN 
+# diag.vals <- getME(fm1,"theta")[getME(fm1,"lower") == 0] 
+# any(diag.vals < 1e-6) # FALSE
+
+## But has issues with every optimizer... -->
+#source(system.file("utils", "allFit.R", package="lme4")) 
+#fm1.all <- allFit(fm1)
+#ss <- summary(fm1.all)
+
+#No convergence issues with open pose detectors for faces, but for wrists
+#fullModelminsInt_Faces_OP <- glmer(faceOP ~ scale(age.at.test) + posture + orientation + (1|subid), 
+#      data = d, 
+#        family = "binomial") 
+```
+
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
+Error bars represent non-parametric bootstrapped 95 percent confidence intervals."}
+
+p1=ggplot(aroundNamings_Posture_Faces, aes(x = posture, y = mean, col = posture)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. faces") +
+  xlab("") +
+  theme_few() + 
+  theme(legend.position="none") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
+
+p2=ggplot(aroundNamings_Orient_Faces, aes(x = orientation, y = mean, col = orientation)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. faces") +
+  xlab("") +
+  theme_few() +
+  theme(legend.position="none")  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
+
+p3=ggplot(aroundNamings_Posture_Wrists, aes(x = posture, y = mean, col = posture)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. wrists") +
+  xlab("") +
+  theme_few() +
+  theme(legend.position="none")  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
+
+p4=ggplot(aroundNamings_Orient_Wrists, aes(x = orientation, y = mean, col = orientation)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. wrists") +
+  xlab("") +
+  theme_few() +
+  theme(legend.position="none") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
+
+ggarrange(p1,p2, p3, p4)
+```
+
 # General Discussion
 We used a head-mounted camera to explore how children's postural and locomotive development impacts their visual access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
-Children's posture and orientation towards their caregiver changed systematically across age, and that both of these factors dramatically influenced the proportion of faces and wrists/hands that were available in the child's visual field.
-Broadly, this work suggests that motor development mediates how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
+Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors dramatically influenced the proportion of faces and wrists/hands that were available in the child's visual field.
+This work suggests that motor development mediates how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
 More work is needed to understand how these results relate to the pattern of activities and social interctions observed in children's home experiences [@fausey2016].
 
-This work also deploys novel advancements in computer vision to the study of developmental psychology.
-The field of object detection and recognition has advanced dramatically in the past five years, creating a new generation of algorithmic tools.
-These tools are substantially better equipped to deal with noisier, more complicated datasets and can extract richer and more detailed information than classic models of face detection (i.e., ViolaJones).
-As the headcams we used were inexpensive and the computer vision algorithms freely available,  suggest that the combined use of these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
+This work also integrates novel advancements in computer vision with developmental psychology.
+The field of object detection/recognition has advanced dramatically in the past five years, creating a new generation of algorithmic tools.
+These tools are substantially better equipped to deal with noisier, more complicated datasets and to extract richer and more detailed information than classic models of face detection (i.e., ViolaJones).
+As the headcams we used were inexpensive and the computer vision algorithms freely available,  we suggest that the combined use of these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
 
@@
 An earlier version of this work was presented to the Cognitive Science Society in @frank2013.
+
 # References

### 2018-01-30 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
< six pages!

@@
   print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
+
+# <!-- \caption{\label{tab:pop} Demographics by age group.} -->
 ```
@@
 Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80\% English at home.
-Demographics and exclusion rates are given in Table \ref{tab:pop}.
+Demographics and exclusion rates are given in the table below.
 \begin{table}[ht]
@@
 \end{tabular}
-\caption{\label{tab:pop} Demographics by age group.}
 \end{table}
@@
 All parents signed consent documents while children were fitted with the headcam.
-If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam [@yoshida2008].
+If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam.
 When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session.
@@
 ```
-
+To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
 \begin{table}[H]
@@
 
-To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
-A summary of the coefficients of a model with only main effects (and no interactions) can be found in Table 3.
+A summary of the coefficients of a model with only main effects (and no interactions) can be found in Table 2.
 When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
 We also found the same pattern with respect to infants' visual access to hands (see Table 3); including interaction terms also eliminated any main effect of age on the proportion of wrists detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
-Thus, these results suggest that infants' visual access to social information is mediated by their posture and orientation, which is in turn a function of their general locomotor development (as indexed by chronological age).
+Thus, these results suggest that infants' visual access to social information is mediated by their posture and orientation, which is in turn a function of their general locomotor development.
 
@@
 
+# General Discussion
+We used a head-mounted camera to explore how children's postural and locomotive development impacts their visual access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
+Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors dramatically influenced the proportion of faces and wrists/hands that were available in the child's visual field.
+This work suggests that motor development mediates how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
+More work is needed to understand how these results relate to the pattern of activities and social interctions observed in children's home experiences [@fausey2016].
+
 ```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
@@
 
-# General Discussion
-We used a head-mounted camera to explore how children's postural and locomotive development impacts their visual access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
-Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors dramatically influenced the proportion of faces and wrists/hands that were available in the child's visual field.
-This work suggests that motor development mediates how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
-More work is needed to understand how these results relate to the pattern of activities and social interctions observed in children's home experiences [@fausey2016].
-
 This work also integrates novel advancements in computer vision with developmental psychology.

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
discussion edits & new modeling section

@@
 
-abstract: 
-  "The ability to process social information--including eye gaze--is a critical component of children's early language and cognitive development.
+abstract: "The ability to process social information--including eye gaze--is a critical component of children's early language and cognitive development.
 However, as children reach their first birthday, they begin to locomote themselves, walking and exploring their visual environment in an entirely new way.
@@
 
-keywords:
-    "social cognition; face-perception; infancy; locomotion; head-cameras; deep learning"
-
+keywords: "social cognition; face-perception; infancy; locomotion; head-cameras; deep learning"
 output: cogsci2016::cogsci_paper
@@
 ```{r echo=FALSE, include=FALSE}
-#  Try a normal logistic regressino, grouping by subject / posture / orientation -- model convergance errors for glmer.
 
@@
   group_by(subid, posture, orientation) %>%
-  summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1])
+  summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1]) %>%
+  mutate(propFaces = countFaces / (countFaces + countNotFaces))
 
@@
   group_by(subid, posture, orientation) %>%
-  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE),   age.at.test=age.at.test[1])
+  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE),     age.at.test=age.at.test[1])  %>%
+  mutate(propWrists = countWrists / (countWrists + countNotWrists))
 
-# scale cont predictors
-dGLM_faces$age.at.test=scale(dGLM_faces$age.at.test)
-dGLM_wrists$age.at.test=scale(dGLM_wrists$age.at.test)
+## run models
+faceGLM <- glmer(cbind(countFaces,countNotFaces) ~ orientation + posture + scale(age.at.test) + 
+               (orientation | subid), 
+             dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
 
-# successes / failures data structure for glm
-y_faces=cbind(dGLM_faces$countFaces, dGLM_faces$countNotFaces)  
-y_wrists=cbind(dGLM_wrists$countWrists, dGLM_wrists$countNotWrists)
+wristGLM <- glmer(cbind(countWrists,countNotWrists) ~ orientation + posture + scale(age.at.test) + 
+               (orientation | subid), 
+             dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
 
-# run models
-glm_faces<-glm(y_faces ~ dGLM_faces$age.at.test * dGLM_faces$posture * dGLM_faces$orientation,family=binomial(logit))
+## output
+cleanedNames_NoInt = c("Intercept","Close","Far","Prone","Sit","Stand","Age")
+faceGLMOut=(summary(faceGLM))
+rownames(faceGLMOut$coefficients)<-cleanedNames_NoInt
 
-glm_faces_noInt<-glm(y_faces ~ dGLM_faces$age.at.test + dGLM_faces$posture + dGLM_faces$orientation,family=binomial(logit))
+wristGLMOut=(summary(wristGLM))
+rownames(wristGLMOut$coefficients)<-cleanedNames_NoInt
 
-glm_wrists<-glm(y_wrists ~ dGLM_wrists$age.at.test * dGLM_wrists$posture * dGLM_wrists$orientation,family=binomial(logit))
-
-glm_wrists_noInt<-glm(y_wrists ~ dGLM_wrists$age.at.test + dGLM_wrists$posture + dGLM_wrists$orientation,family=binomial(logit))
-
-## OUTPUT GLM RESULTS
-cleanedNames_NoInt = c("Intercept","Age","Prone","Sit","Stand","Close","Far")
-cleanedNames = c("Intercept","Age","Prone","Sit","Stand","Close","Far", "Age:Prone","Age:Sit","Age:Stand","Age:Close",
-          "Age:Far","Prone:Close","Sit:Close","Stand:Close","Prone:Far","Sit:Far","Stand:Far","Age:Prone:Close",
-          "Age:Sit:Close","Age:Prone:Far","Age:Sit:Far")
-
-faceGLMOut=(summary(glm_faces))
-# rownames(faceGLMOut)<-cleanedNames
-
-wristGLMOut=(summary(glm_wrists))
-# rownames(wristGLMOut)<-cleanedNames
-
-wristGLMOut_noInt=xtable(summary(glm_wrists_noInt), caption="Model coefficients from a generalized linear model predicting the proportion of wrists seen by infants.")
-rownames(wristGLMOut_noInt)<-cleanedNames_NoInt
-
-faceGLMOut_noInt=xtable(summary(glm_faces_noInt), caption="Model coefficients from a generalized linear model predicting the proportion of wrists seen by infants.")
-rownames(faceGLMOut_noInt)<-cleanedNames_NoInt
+xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed model predicting the proportion of faces vs. wrists seen by infants.") 
+xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3))
 
 ```
+
 ```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age.
@@
 ```
-To formalize these observations, we fit a generalized linear model to the proportion of faces infants saw in each posture and orientation (as detected by MTCNN), with participant's age, orientation, and posture as independent variables.
-\begin{table}[H]
+To formalize these observations, we fit two generalized linear mixed-effect models for the proportion of faces and the proportion of hands infants saw in each posture and orientation, with participant's age, orientation, and posture as fixed effects and with random slopes for infant's orientation.
+Random slopes for posture or interactions between our fixed effects caused the models to fail to converge.
+
+\begin{table}[ht]
 \centering
@@
   \textbf{Faces} \\
-  Intercept & -5.2469 & 0.0584 & -89.88 & 0.0000 \\ 
-  Age & 0.0847 & 0.0041 & 20.89 & 0.0000 \\ 
-  Prone & 0.2015 & 0.0564 & 3.57 & 0.0004 \\ 
-  Sit & 1.4053 & 0.0541 & 25.98 & 0.0000 \\ 
-  Stand & 1.4272 & 0.0542 & 26.33 & 0.0000 \\ 
-  Close & 1.8239 & 0.0230 & 79.17 & 0.0000 \\ 
-  Far & 2.5479 & 0.0239 & 106.42 & 0.0000 \\ 
+Intercept & -5.78 & 0.33 & -17.55 & 0.000 \\ 
+  Close & 2.09 & 0.31 & 6.70 & 0.000 \\ 
+  Far & 2.58 & 0.34 & 7.54 & 0.000 \\ 
+  Prone & -0.18 & 0.08 & -2.39 & 0.017 \\ 
+  Sit & 1.00 & 0.07 & 13.54 & 0.000 \\ 
+  Stand & 1.50 & 0.08 & 19.98 & 0.000 \\ 
+  Age & 0.17 & 0.17 & 0.97 & 0.334 \\ 
   \hline
-  & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-  \textbf{Wrists} \\ 
-  Intercept & -5.0818 & 0.0776 & -65.46 & 0.0000 \\ 
-  Age & 0.0564 & 0.0050 & 11.35 & 0.0000 \\ 
-  Prone & 0.9499 & 0.0774 & 12.27 & 0.0000 \\ 
-  Sit & 1.7282 & 0.0758 & 22.79 & 0.0000 \\ 
-  Stand & 1.6618 & 0.0760 & 21.88 & 0.0000 \\ 
-  Close & 0.7472 & 0.0188 & 39.81 & 0.0000 \\ 
-  Far & 1.6357 & 0.0200 & 81.61 & 0.0000 \\ 
+  \textbf{Hands} \\
+  Intercept & -4.84 & 0.19 & -25.40 & 0.000 \\ 
+  Close & 0.88 & 0.17 & 5.25 & 0.000 \\ 
+  Far & 1.77 & 0.27 & 6.45 & 0.000 \\ 
+  Prone & 0.58 & 0.10 & 5.73 & 0.000 \\ 
+  Sit & 1.18 & 0.10 & 12.04 & 0.000 \\ 
+  Stand & 1.01 & 0.10 & 10.22 & 0.000 \\ 
+  Age & 0.14 & 0.10 & 1.40 & 0.161 \\ 
    \hline
@@
 
-A summary of the coefficients of a model with only main effects (and no interactions) can be found in Table 2.
-When we did include interaction terms between age, posture, and orientation, age no longer remained a significant predictor (b = `r round(faceGLMOut$coef[2,1],2)`, SE = `r round(faceGLMOut$coef[2,2],2)`, z = `r round(faceGLMOut$coef[2,3],2)` , p = `r round(faceGLMOut$coef[2,4],2)`).
-We also found the same pattern with respect to infants' visual access to hands (see Table 3); including interaction terms also eliminated any main effect of age on the proportion of wrists detected (b = `r round(wristGLMOut$coef[2,1],2)`, SE = `r round(wristGLMOut$coef[2,2],2)`, z = `r round(wristGLMOut$coef[2,3],2)`, p = `r round(wristGLMOut$coef[2,4],2)`).
+A summary of the coefficients of a models can be found in Table 2.
+While infant's posture and orientation significantly impacted the proportion of faces and hands that infants saw, age was not a significant predictor in either model.
 Thus, these results suggest that infants' visual access to social information is mediated by their posture and orientation, which is in turn a function of their general locomotor development.
@@
 # General Discussion
-We used a head-mounted camera to explore how children's postural and locomotive development impacts their visual access to social information, here operationalized as the presence of the faces and wrists of their caregiver.
-Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors dramatically influenced the proportion of faces and wrists/hands that were available in the child's visual field.
-This work suggests that motor development mediates how infants experience their visual world and the social information in it: infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
-More work is needed to understand how these results relate to the pattern of activities and social interctions observed in children's home experiences [@fausey2016].
+
+We used a head-mounted camera to explore how children's postural and locomotor development affects their visual access to social information, here operationalized as the presence of the faces and hands of their caregiver.
+Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors influenced the proportion of faces and hands that were available in the child's visual field.
+This work suggests that motor development mediates how infants experience their visual world and the social information in it: Infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
 
@@
 
-This work also integrates novel advancements in computer vision with developmental psychology.
-The field of object detection/recognition has advanced dramatically in the past five years, creating a new generation of algorithmic tools.
-These tools are substantially better equipped to deal with noisier, more complicated datasets and to extract richer and more detailed information than classic models of face detection (i.e., ViolaJones).
-As the headcams we used were inexpensive and the computer vision algorithms freely available,  we suggest that the combined use of these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
+We created a situation in the lab in which the context of interaction was tightly controlled, but as children grow and change, the activities in which they engage with their caregivers also vary, leading to differences in the distribution of contexts they experience.
+Thus, more work is needed to understand how our results relate to children's home experiences [@fausey2016].
+The ability to walk is part of a cascade of changes in children's  experience, and our study captures only a cross-sectional slice of this broader, multifaceted trajectory.
+
+Understanding these changes has been a persistent challenge for developmental psychology, but the field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools.
+These tools deal better with noisier, more complicated datasets and extract richer and more detailed information than previous systems.
+We hope that these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
 

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
rewording, edits, typos

@@
 ---
-title: "Postural developments mediate children's visual access to social information"
+title: "Postural developments modulate children's visual access to social information"
 bibliography: xsface.bib
@@
 author-information: > 
-    \author{{\large \bf Alessandro Sanchez*} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University
-    \And {\large \bf Bria Long*}  \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
+    \author{{\large \bf Alessandro Sanchez\thanks{These authors contributed equally to this work.}} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University
+    \And {\large \bf Bria Long\footnotemark[1]}  \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University
     \And {\large \bf Allison M. Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University
@@
 
-abstract: "The ability to process social information--including eye gaze--is a critical component of children's early language and cognitive development.
-However, as children reach their first birthday, they begin to locomote themselves, walking and exploring their visual environment in an entirely new way.
+abstract: "The ability to process social information is a critical component of children's early language and cognitive development.
+However, as children reach their first birthday, they begin to locomote themselves, dramatically affecting their visual access to this information.
 How do these postural and locomotive changes affect children's access to the social information relevant for word-learning?
-Here, we explore this question by using head-mounted cameras to record infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to detect the proportion of faces and wrists in infants' environments.
-We find that infants' posture and orientation to their caregiver largely mediates their access to social information, suggesting that postural and locomotive developments play a significant role in the emergence of children's linguistic and social capacities.
-Broadly, we suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
+Here, we explore this question by using head-mounted cameras to record 36 infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to estimate the proportion of faces and hands in infants' environments.
+We find that infants' posture and orientation to their caregiver modulates their access to social information, confirming previous work that suggests motoric developments play a significant role in shaping the input to children's developing linguistic and social capacities.
+We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
 
@@
 # Introduction
-[^1]: *These authors contributed equally to this work.
+[^1]: These authors contributed equally to this work.
 
-[^1] From early infancy, children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like].
+From early infancy, children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like].
 Even newborns tend to prefer to look at faces with direct vs.
@@
 These motor changes have significant consequences for what children see; crawling and walking infants simply have different views of the world.
-For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
+For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011].
+In general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
 
 One possibility is that these motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities [@iverson2010].
-Indeed, these postural changes also impact how children interact with their mothers; walking (vs.
+These postural changes also impact how children interact with their mothers: walking (vs.
 crawling) infants make different kinds of object-related bids for attention from their mothers and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].
@@
 This perspective changes over the first two years of life, as views transition from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
+These changes may be driven by – or independent from – the concurrent motoric changes described above.
 
-Here, we directly examine whether postural and locomotive developments change the availability of social information--the presence of faces and hands.
+Here, we directly examine whether postural and locomotive developments change the availability of social information -- the presence of faces and hands.
 To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
 We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
-We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus rendering accessible a larger portion of the social information in their environment.
+We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus providing children with greater access to the social information in their environment.
 
@@
 
-Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's approximate vertical field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988].
-As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field)
+Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988].
+As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field).
 
@@
 When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session.
-Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the novel toys.
+Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys.
 The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.
@@
 These sessions yielded videos of
-`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a milion frames), with an average video length of
+`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a million frames), with an average video length of
 `r round(sum(inc_subs$len.min)/60, digits = 2)` minutes
@@
 The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts.
-For the second algorithm [OpenPose, @cao2017realtime], we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
+For the second algorithm [OpenPose; @cao2017realtime], we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
 
 The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for hands for two reasons.
-First, as we did not capture childrens entire visual field.
-the presence of a wrist is likely often indicative of the presence of a hand within the field of view.
+First, as we did not capture children's entire visual field, the presence of a wrist is likely often indicative of the presence of a hand within the field of view.
 Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
@@
 
-The CNN-based pose detector  [OpenPose, @cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The CNN-based pose detector  [OpenPose; @cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
 The candidates are then matched to a single individual and finally assembled into a pose; here, we only made use of the body parts relevant to the face and hands (nose and wrists).
-We operationalized face detections as any frames containing a nose, and hand detections as any frames containing either the left or right wrist.
 
 ### Detector evaluation
-To evaluate face detector performance, we hand-labeled  an "gold set" of labeled frames.
-To account for the relatively rare appearance of faces in the dataset, we  hand abeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
+To evaluate face detector performance, we hand-labeled a "gold set" of labeled frames.
+To account for the relatively rare appearance of faces in the dataset, we hand-labeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
 Each sample was comprised of an equal number of frames taken from each child's video.
@@
 
-```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3.5, fig.height=3.5, fig.cap = "Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver." }
+```{r posture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3.5, fig.height=3.5, fig.cap = "Proportion of time that infants in each age group spent in each posture/orientation relative to their caregiver." }
 p1=ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
@@
 The proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with age.
-Both 8 month-olds and 12 month-olds spent equivalent amounts of time lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time sitting or standing (see Figure \ref{fig:posture}).
+Both 8-month-olds and 12-month-olds spent equivalent amounts of time lying/crawling, which was markedly decreased in the 16-month-olds, who spent most of their time sitting or standing (see Figure \ref{fig:posture}).
 We also observed changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions (see Figure \ref{fig:posture}).
@@
 ## Changes in Access to Faces and Hands
-We examined the proportion of face and wrist detections across age (see Figure \ref{fig:detByAge}).
-We observed a slight U-shaped function in face detections, such that 12-month-olds appeared to have visual acces to slightly fewer faces faces than 8 or 16-month-olds; conversely, wrist detections perhaps appeared to generally increase with age.
+We examined the proportion of face and hand detections across age (see Figure \ref{fig:detByAge}).
+We observed a slight U-shaped function in face detections, such that 12-month-olds appeared to have visual access to slightly fewer faces faces than 8 or 16-month-olds; conversely, hand detections perhaps appeared to generally increase with age.
 
-Age related effects were much smaller than postural and locomotive changes on children's visual access to faces and wrists/hands.
-Children's posture was a major factor both in how many faces and wrists/hands they saw during the play session.
-Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces (Figure \ref{fig:detByPosOrient}, upper panel); this same pattern was also true for wrists/hand detections.
+Age related effects were much smaller than postural and locomotive changes on children's visual access to faces and hands.
+Children's posture was a major factor both in how many faces and hands they saw during the play session.
+Infants who were sitting saw more faces than infants who were lying down or being carried, while infants who were standing saw the most faces (Figure \ref{fig:detByPosOrient}, upper panel); this same pattern was also true for hand detections.
 Children’s orientation also impacted their visual access to faces and hands: children who were far away from their caregiver were more likely to see faces/hands than children who were close to their caregiver (Figure \ref{fig:detByPosOrient}, lower panel).
@@
 
-```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
+```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
 
@@
 
-A summary of the coefficients of a models can be found in Table 2.
+A summary of the coefficients of the models can be found in Table 2.
 While infant's posture and orientation significantly impacted the proportion of faces and hands that infants saw, age was not a significant predictor in either model.
-Thus, these results suggest that infants' visual access to social information is mediated by their posture and orientation, which is in turn a function of their general locomotor development.
+Thus, these results suggest that infants' visual access to social information is modulated by their posture and orientation, which is in turn a function of their general locomotor development.
 
@@
 
-Finally, we explored how face and wrist detections changed during object labeling events as a function of infant's posture and orientation.
-We analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and synchronized with the frame-by-frame face/wrist detections.
-We found that infants' posture and orientations mediated their visual access to faces and wrists/hands during labeling events; infants who were sitting or standing were more likely to have visual access to this social information (see Figure \ref{fig:detByNaming}).
-However, we did not find that infants saw particularily more faces or wrists/hands during naming events relative to baseline (avg.
-difference in proportion of wrists, 8 m.o.
+Finally, we explored how face and hand detections changed during object labeling events as a function of infants' posture and orientation.
+We analyzed a four-second window around each labeling event (e.g., "Look at the [zem]!"); these labeling events were hand-annotated and synchronized with the frame-by-frame face/hand detections.
+We found that infants' posture and orientations modulated their visual access to faces and hands during labeling events; infants who were sitting or standing were more likely to have visual access to this social information (see Figure \ref{fig:detByNaming}).
+However, we did not find that infants saw particularly more faces or hands during naming events relative to baseline (avg.
+difference in proportion of hands, 8 m.o.
 = `r round(wrists_naming$avgBaselineDiff[1],3)`, 12 m.o.
@@
 Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors influenced the proportion of faces and hands that were available in the child's visual field.
-This work suggests that motor development mediates how infants experience their visual world and the social information in it: Infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
+This work suggests that motor development modulates how infants experience their visual world and the social information in it: Infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
 
-```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
+```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion of face and hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
 Error bars represent non-parametric bootstrapped 95 percent confidence intervals."}

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
rewording, edits, typos

@@
 
-\title{Postural developments mediate children's visual access to social
+\title{Postural developments modulate children's visual access to social
 information}
 
-\author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Allison M.
+\author{{\large \bf Alessandro Sanchez\thanks{These authors contributed equally to this work.}} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long\footnotemark[1]}  \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Allison M.
 Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University \And {\large \bf Michael C.
@@
 \begin{abstract}
-The ability to process social information--including eye gaze--is a
-critical component of children's early language and cognitive
-development. However, as children reach their first birthday, they begin
-to locomote themselves, walking and exploring their visual environment
-in an entirely new way. How do these postural and locomotive changes
-affect children's access to the social information relevant for
-word-learning? Here, we explore this question by using head-mounted
-cameras to record infants' (8-16 months of age) egocentric visual
-perspective and use computer vision algorithms to detect the proportion
-of faces and wrists in infants' environments. We find that infants'
-posture and orientation to their caregiver largely mediates their access
-to social information, suggesting that postural and locomotive
-developments play a significant role in the emergence of children's
-linguistic and social capacities. Broadly, we suggest that the combined
-use of head-mounted cameras and the application of new computer vision
-techniques is a promising avenue for understanding the statistics of
-infants' visual and linguistic experience.
+The ability to process social information is a critical component of
+children's early language and cognitive development. However, as
+children reach their first birthday, they begin to locomote themselves,
+dramatically affecting their visual access to this information. How do
+these postural and locomotive changes affect children's access to the
+social information relevant for word-learning? Here, we explore this
+question by using head-mounted cameras to record 36 infants' (8-16
+months of age) egocentric visual perspective and use computer vision
+algorithms to estimate the proportion of faces and hands in infants'
+environments. We find that infants' posture and orientation to their
+caregiver modulates their access to social information, confirming
+previous work that suggests motoric developments play a significant role
+in shaping the input to children's developing linguistic and social
+capacities. We suggest that the combined use of head-mounted cameras and
+the application of new computer vision techniques is a promising avenue
+for understanding the statistics of infants' visual and linguistic
+experience.
 
@@
 
-Children are deeply engaged in learning from others (Csibra \& Gergely,
-2009; Meltzoff, 2007) and attend to the social information in their
-environment from their earliest days. Even newborns tend to prefer to
-look at faces with direct vs.~averted gaze (Farroni, Csibra, Simion, \&
-Johnson, 2002) and young infants follow overt gaze shifts (Gredebäck,
-Fikke, \& Melinder, 2010; Gredebäck, Theuring, Hauf, \& Kenward, 2008).
-Further, when infants free-view videos, they tend to look mostly at
-faces at the expense of other visual information--though older infants
-start to look towards people's hands and the actions they are performing
-(Frank, Amso, \& Johnson, 2014; Frank, Vul, \& Saxe, 2012).
+From early infancy, children are deeply engaged in learning from others
+(Csibra \& Gergely, 2009; Meltzoff, 2007). Even newborns tend to prefer
+to look at faces with direct vs.~averted gaze (Farroni, Csibra, Simion,
+\& Johnson, 2002) and young infants follow overt gaze shifts (Gredebäck,
+Fikke, \& Melinder, 2010). Further, when infants attend to video
+stimuli, they tend to look mostly at faces at the expense of other
+visual information -- though older infants start to look towards
+people's hands and the actions they are performing (Frank, Amso, \&
+Johnson, 2014; Frank, Vul, \& Saxe, 2012).
 
-However, as children are learning from others around them, their view of
-the world is also radically changing (K. Adolph \& Berger, 2007).
-Infants' motor abilities improve dramatically near the end of the first
-year of life, allowing them to locomote independently. These motor
-changes have significant consequences for what children see; crawling
-and walking infants simply have different views of the world. For
-example, during spontaneous play in a laboratory playroom, toddlers are
-more likely to look at the floor while crawling than while walking
-(Franchak, Kretch, Soska, \& Adolph, 2011); in general, walking infants
+Then, however, their view of the world radically changes (K. Adolph \&
+Berger, 2007). Infants' motor abilities improve dramatically near the
+end of the first year of life, allowing them to locomote independently.
+These motor changes have significant consequences for what children see;
+crawling and walking infants simply have different views of the world.
+For example, during spontaneous play in a laboratory playroom, toddlers
+are more likely to look at the floor while crawling than while walking
+(Franchak, Kretch, Soska, \& Adolph, 2011). In general, walking infants
 tend to have full visual access to their environment and the people in
@@
 
-One possibility is that these motor improvements have strong
-developmental cascades, impacting children's emerging social, cognitive,
-and linguistic abilities (Iverson, 2010). Indeed, these postural changes
-also impact how children interact with their mothers; walking
-(vs.~crawling) infants make different kinds of object-related bids for
-attention from their mothers and tend to hear more action directed
-statements (e.g., ``open it'') (Karasik, Tamis-LeMonda, \& Adolph,
-2014). Further, in an observational study, Walle \& Campos (2014) found
-that children who were able to walk had both higher receptive and
-productive vocabularies. On their account, children's ability to stand
-and independently locomote may fundamentally change their ability to
-access social information (e.g., faces, gaze) and in turn to accelerate
-their own learning.
+One possibility is that these motor improvements lead to developmental
+cascades that impact children's emerging social, cognitive, and
+linguistic abilities (Iverson, 2010). These postural changes also impact
+how children interact with their mothers: walking (vs.~crawling) infants
+make different kinds of object-related bids for attention from their
+mothers and tend to hear more action directed statements (e.g., ``open
+it'') (Karasik, Tamis-LeMonda, \& Adolph, 2014). Further, in an
+observational study, Walle \& Campos (2014) found that children who were
+able to walk had both higher receptive and productive vocabularies. On
+their account, children's ability to stand and independently locomote
+may change their ability to access social information (e.g., faces,
+gaze) and in turn to accelerate their own learning.
 
@@
 perspective. By using head-mounted cameras, researchers have begun to
-record the visual experiences of infants and children-- which even for
+record the visual experiences of infants and children -- which even for
 walking children are extremely different from the adult perspective (and
-not easily predicted by our own adult intuitions) ({\textbf{???}};
-Clerkin, Hart, Rehg, Yu, \& Smith, 2017; Yoshida \& Smith, 2008). While
-children's views tend to be more restricted and dominated by objects and
-hands (Yoshida \& Smith, 2008), both computational and empirical work
+not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
+Yu, \& Smith, 2017; Franchak et al., 2011; Yoshida \& Smith, 2008).
+Children's views tend to be more restricted and dominated by objects and
+hands (Yoshida \& Smith, 2008), and computational and empirical work
 suggest that this restricted viewpoint may be more effective for
@@
 (Bambach, Crandall, Smith, \& Yu, 2017; D. Yurovsky, Smith, \& Yu, in
-press). Further, recent work also suggests dramatic changes in the
-child's perspective over the first two years of life, as views
-transition from primarily containing close up views of faces to
+press). This perspective changes over the first two years of life, as
+views transition from primarily containing close up views of faces to
 capturing views of hands paired with the objects they are acting on
-(Fausey, Jayaraman, \& Smith, 2016).
+(Fausey, Jayaraman, \& Smith, 2016). These changes may be driven by --
+or independent from -- the concurrent motoric changes described above.
 
 Here, we directly examine whether postural and locomotive developments
-change the availability of social information--the presence of faces and
-hands. To do so, we recorded the visual experience of a group of infants
-in three age ranges (8,12, and 16 months) using head-mounted cameras
-during a brief laboratory free-play session; children's posture and
-orientation relative to their caregiver were also recorded from a
+change the availability of social information -- the presence of faces
+and hands. To do so, we recorded the visual experience of a group of
+infants in three age ranges (8,12, and 16 months) using head-mounted
+cameras during a brief laboratory free-play session; children's posture
+and orientation relative to their caregiver were also recorded from a
 third-person perspective and hand-annotated. We then capitalize on
@@
 primarily be looking at the ground, while walking toddlers would have
-access to a richer visual landscape, thus rendering accessible a larger
-portion of the social information in their environment.
+access to a richer visual landscape, thus providing children with
+greater access to the social information in their environment.
 
@@
 
-\subsection{Participants}\label{participants}
+\subsubsection{Participants}\label{participants}
 
 Our final sample consisted of 36 infants and children, with 12
-participants in three age groups: 8 months (6 females), 12 months (7
-females), and 16 months (6 females). Participants were recruited from
-the surrounding community via state birth records, had no documented
-disabilities, and were reported to hear at least 80\% English at home.
-Demographics and exclusion rates are given in Table \ref{tab:pop}.
+participants in three age groups: 8 months (6 F), 12 months (7 F), and
+16 months (6 F). Participants were recruited from the surrounding
+community via state birth records, had no documented disabilities, and
+were reported to hear at least 80\% English at home. Demographics and
+exclusion rates are given in the table below.
 
-\begin{table}[H]
+\begin{table}[ht]
 \centering
@@
 missing Communicative Development Inventory (CDI) data, 2 for missing
-scene camera footage, 1 for fussiness, and one excluded for sample
-symmetry. All inclusion decisions were made independent of the results
-of subsequent analyses. Some of these data were also used in Frank,
-Simmons, Yurovsky, \& Pusiol (2013).
+scene camera footage, 1 for fussiness, and one for sample symmetry. All
+inclusion decisions were made independent of the results of subsequent
+analyses. Some of these data were also analyzed in Frank, Simmons,
+Yurovsky, \& Pusiol (2013).
 
-\subsection{Head-mounted camera}\label{head-mounted-camera}
+\subsubsection{Head-mounted camera}\label{head-mounted-camera}
 
@@
 
-\caption[Field of view for two different headcam configurations (we used the one below in our current study)]{Field of view for two different headcam configurations (we used the one below in our current study).}\label{fig:headcam}
+\caption[Vertical field of view for two different headcam configurations (we used the lower in our current study)]{Vertical field of view for two different headcam configurations (we used the lower in our current study).}\label{fig:headcam}
 \end{figure}
@@
 from a MD80 model camera attached to a soft elastic headband. Videos
-captured by the headcam were 720x480 pixels with 25 frames per second.
-Detailed instructions for creating this headcam can be found at
-\url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
+captured by the headcam were 720x480 pixels with 25 frames per
+second.\footnote{Detailed instructions for creating this headcam can be
+  found at
+  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
 A fisheye lens was attached to the camera to increase the view angle
@@
 \(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
-\ref{fig:headcam}, left).
+\ref{fig:headcam}, bottom).
 
 Even with the fish-eye lens, the vertical field of view of the camera is
-still considerably reduced compared to the child's approximate vertical
-field of view, which spans around 100--120\(^{\circ}\) in the vertical
-dimension by 6-7 months of age (Cummings, Van Hof-Van Duin, Mayer,
-Hansen, \& Fulton, 1988; Mayer, Fulton, \& Cummings, 1988). As we were
-primarily interested in the presence of faces in the child's field of
-view, we chose to orient the camera upwards to capture the entirety of
-the child's upper visual field where the child is likely to see adult
-faces.
+still considerably reduced compared to the child's field of view, which
+spans around 100--120\(^{\circ}\) in the vertical dimension by 6-7
+months of age (Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton,
+1988; Mayer, Fulton, \& Cummings, 1988). As we were primarily interested
+in the presence of faces in the child's field of view, we chose to
+orient the camera upwards to capture the entirety of the child's upper
+visual field where the child is likely to see adult faces, understanding
+that this decision limited our ability to detect hands (especially those
+of the child, which are typically found at the bottom of the visual
+field).
 
-\subsection{Procedure}\label{procedure}
+\subsubsection{Procedure}\label{procedure}
 
-First, all parents signed consent documents in a waiting room where
-children were fitted with the headcam. After the child was comfortable
-in the waiting room and with the experimenter, the experimenter placed
-the headcam on the child's head. If the child was uninterested in
-wearing the headcam or tried to take it off, the experimenter presented
-engaging toys to try to draw the child's focus away from the headcam
-(Yoshida \& Smith, 2008). After the child was comfortable wearing the
-headcam, the child and caregiver were shown to a playroom for the
-free-play session. Parents were shown a box containing three pairs of
-novel and familiar objects (e.g., a ball and a microfiber duster, named
-a ``zem''), and were instructed to play with the object pairs with their
-child one at a time, ``as they typically would.'' All parents confirmed
-that their child had not previously seen the novel toys and were
-instructed to use the novel labels to refer to the novel toys. The
-experimenter then left the playroom for approximately 15 minutes, during
-which a tripod-mounted camera in the corner of the room recorded the
-session and the headcam captured video from the child's perspective.
+All parents signed consent documents while children were fitted with the
+headcam. If the child was uninterested in wearing the headcam or tried
+to take it off, the experimenter presented engaging toys to try to draw
+the child's focus away from the headcam. When the child was comfortable
+wearing the headcam, the child and caregiver were shown to a playroom
+for the free-play session. Parents were shown a box containing three
+pairs of novel and familiar objects (e.g., a ball and a microfiber
+duster, named a ``zem''), and were instructed to play with the object
+pairs with their child one at a time, ``as they typically would.'' All
+parents confirmed that their child had not previously seen the novel
+toys and were instructed to use the novel labels to refer to the toys.
+The experimenter then left the playroom for approximately 15 minutes,
+during which a tripod-mounted camera in the corner of the room recorded
+the session and the headcam captured video from the child's perspective.
 
-\subsection{Data Processing and
+\subsubsection{Data Processing and
 Annotation}\label{data-processing-and-annotation}
-
-\begin{figure*}
-\centering
-\includegraphics[width=6in]{images/detector_samples_banner.pdf}
-\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
-The last column features a false positive from OpenPose and a false negative from MTCNN.}
-\end{figure*}
 
@@
 synchronized with the tripod-mounted videos using FinalCut Pro Software.
-These sessions yielded videos of 516 minutes (almost a milion frames),
+These sessions yielded videos of 516 minutes (almost a million frames),
 with an average video length of 8.6 minutes (min = 4.53, max = 19.35).
@@
 
-We created a set of custom annotations that described the child's
-physical posture (e.g.~standing) and the orientation of the caregiver
-relative to the child (e.g.~far away). The child's posture was
-categorized as being held/carried, prone (crawling or lying), sitting,
-or standing. The caregiver's orientation was characterized as being
-close to the child, far from the child, and behind the child
-(independent of distance). For the first two annotations (close/far from
-the child), the caregiver could either be to the the front or to the
-side of the child. All annotations were made by a trained coder using
+We created custom annotations to describe the child's physical posture
+(e.g.~standing) and the orientation of the child relative to the
+caregiver (e.g.~far away). The child's posture was categorized as being
+held/carried, prone (crawling or lying), sitting, or standing. The
+caregiver's orientation was characterized as being close, far, or behind
+the child (independent of distance). For the first two annotations
+(close/far from the child), the caregiver could either be to the front
+or side of the child. All annotations were made by a trained coder using
 the OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman, Sanderson,
-\& Millman, 2012), and times when the child was out of view of the
-tripod camera were marked as uncodable and were excluded from these
+\& Millman, 2012). Times when the child was out of view of the tripod
+camera were marked as uncodable and were excluded from these
 annotations.
@@
 
-We used three face detection systems to measure infants' changing access
-to faces. The first of these is the most commonly-used and widely
-available face detection algorithm: Viola-Jones. We used this as a
+We used three face detection systems to measure infants' access to
+faces. The first of these is the most commonly-used and widely available
+face detection algorithm: Viola-Jones. We used this algorithm as a
 benchmark for performance, as while it can achieve impressive accuracy
@@
 about the position of 18 different body parts. For the second algorithm
-(called OpenPose) (Cao et al., 2017), we used the agent's nose (one of
-the 18 body parts detected) to operationalize the presence of faces, as
-any half of a face necessarily contains a nose.
+(OpenPose; Cao et al., 2017), we used the agent's nose (one of the body
+parts detected) to operationalize the presence of faces, as any half of
+a face necessarily contains a nose.
 
 The OpenPose detector also provided us with the location of an agent's
-wrists, which we used as a proxy for the presence of a hand. We did so
-first because we did not capture the entire visual field experienced by
-the infants/children. Thus, even though the video may only show the
-presence of a wrist, the caregiver's hand may be within the child's
-field of view. Second, we did so because hands are often occluded by
-objects when caregivers are interacting with children, yet still
-visually accessible by the child and part of their joint interaction.
-For example, if a caregiver was holding a toy and presenting it to the
-infant, their wrists may be visible but their hand may not necessarily
-be (as it is occluded by the toy).
+wrists, which we used as a proxy for hands for two reasons. First, as we
+did not capture children's entire visual field. the presence of a wrist
+is likely often indicative of the presence of a hand within the field of
+view. Second, hands are often occluded by objects when caregivers are
+interacting with children, yet still visually accessible by the child
+and part of their joint interaction.
+
+\begin{figure*}
+\centering
+\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
+\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
+The last column features a false positive from OpenPose and a false negative from MTCNN.}
+\end{figure*}
 
@@
 feature-based cascade classifiers (Viola \& Jones, 2004) applied to each
-individual frame. This detector provided information about the presence
-of a face as well as its size and position. The second algorithm (based
-on work by K. Zhang et al. (2016)) uses multi-task cascaded
-convolutional neural networks (MTCNNs). The system was built using a
-novel cascaded CNN-based framework for joint detection and alignment,
-built to perform well in real-world environments where varying
-illuminations and occlusions are present. We used a Tensorflow
-implementation of this algorithm provided by
-(\url{https://github.com/davidsandberg/facenet}). Like Viola-Jones, this
-detector provided information about the presence of a face as well as
-its size and position.
+individual frame. The second algorithm (based on work by K. Zhang et al.
+(2016)) uses multi-task cascaded convolutional neural networks (MTCNNs)
+for joint face detection and alignment, built to perform well in
+real-world environments where varying illuminations and occlusions are
+present. We used a Tensorflow implementation of this algorithm avaliable
+at \url{https://github.com/davidsandberg/facenet}.
 
-The CNN-based pose detector (OpenPose (Cao et al., 2017; Simon, Joo,
+The CNN-based pose detector (OpenPose, Cao et al., 2017; Simon, Joo,
 Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016))
-provided the locations of 18 body parts (ears, nose, wrists, etc.)
-available at
+provided the locations of 18 body parts (ears, nose, wrists, etc.) and
+is available at
 \url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
@@
 use of the body parts relevant to the face and hands (nose and wrists).
-We operationalized face detections as any frames containing a nose, and
-hand detections as any frames containing either the left or right wrist.
 
@@
 
-To evaluate face detector performance, we constructed an unbiased ``gold
-set'' of labeled frames that accounted for the relatively rare
-appearance of faces in the dataset. We thus hand-labeled two samples: a
-sample containing a high density of faces (half reported by MTCNN, half
-by OpenPose) and a random sample from the remaining frames. Each sample
-was comprised of an equal number of frames taken from each child's
-video. Faces were classifed as present in a frame if at least half of
-the face was showing. Precision (hits / hits + false alarms), recall
-(hits / hits + misses), and F-score (harmonic mean of precision and
-recall) were calculated for all detectors and are reported in Table 2.
-For wrist detections, the ``gold set'' was constructed in the same
-manner, with the exception that frames with a high density of wrists
-were taken by pose detections made by OpenPose.
+To evaluate face detector performance, we hand-labeled a ``gold set'' of
+labeled frames. To account for the relatively rare appearance of faces
+in the dataset, we hand abeled two types of samples: a sample containing
+a high density of faces (half reported by MTCNN, half by OpenPose) and a
+random sample from the remaining frames. Each sample was comprised of an
+equal number of frames taken from each child's video. For wrist
+detections, the ``gold set'' was constructed in the same manner, except
+frames with a high density of wrists came only from detections made by
+OpenPose. Faces were classified as present if at least half of the face
+was showing; wrists were classified as present if any part of the wrist
+was showing. Precision (hits / hits + false alarms), recall (hits / hits
++ misses), and F-score (harmonic mean of precision and recall) were
+calculated for all detectors and are reported in Table 2.
 
 For face detection, MTCNN outperformed OpenPose when taking into account
-only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose); we thus use
-MTCNN detections in the following analyses. Although MTCNN and OpenPose
-performed comparably with the random sample, MTCNN performed better on
-the high density sample (specifically looking at precision), suggesting
-that OpenPose generated more false positives than MTCNN. ViolaJones
-performed quite poorly relative to the other detectors, especially with
-respect to the random sample.
-
-For wrist detection, OpenPose performed moderately well (F = 0.74) with
-relatively high precision, but low recall on the randomly sampled set of
-frames (see Table 2). We thus analyze wrist detections, with the caveat
-that we are likely underestimating the proportion of wrists overall in
-the datatset.
+only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose). Although
+MTCNN and OpenPose performed comparably with the random sample, MTCNN
+performed better on the high density sample (specifically looking at
+precision), suggesting that OpenPose generated more false positives than
+MTCNN. ViolaJones performed quite poorly relative to the other
+detectors, especially with respect to the random sample. We thus use
+MTCNN detections in the following analyses. For wrist detection,
+OpenPose performed moderately well (F = 0.74) with relatively high
+precision but low recall on the randomly sampled frames (see Table 2).
+We thus analyze wrist detections, with the caveat that we are likely
+underestimating the proportion of hands in the dataset.
 
@@
 \section{Results}\label{results}
-
-First, we report developmental shifts in infants' posture and their
-orientation relative to their caregiver. Then, we explore how these
-changes influence children's visual access to faces and wrists/hands
-across this developmental time range. Finally, we explore how these
-changes impact the accessibility of faces and wrists/hands during
-labeling events.
-
-\subsection{Changes in Posture and
-Orientation}\label{changes-in-posture-and-orientation}
 
@@
 
-\caption[Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver]{Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver.}\label{fig:posture}
+\caption[Proportion of time that infants in each age group spent in each posture/orientation relative to their caregiver]{Proportion of time that infants in each age group spent in each posture/orientation relative to their caregiver.}\label{fig:posture}
 \end{figure}
@@
 
-We noted characteristic changes in infants' posture and orientation
-across this developmental time range. The proportion of time infants
-spent sitting decreased with age, and the proportion of time infants
-spent standing increased with age. Both 8 month-olds and 12 month-olds
-spent equivalent amounts of time either lying/crawling, which was
-markedly decreased in the 16-month-olds, who spent most of their time
-either sitting or standing (see Figure \ref{fig:posture}). We also
-observed characteristic changes in children's orientation relative to
-their caregivers: the 8-month-olds spent more time with their caregiver
-behind them supporting their sitting positions (see Figure
-\ref{fig:posture}).
+\subsection{Changes in Posture and
+Orientation}\label{changes-in-posture-and-orientation}
+
+The proportion of time infants spent sitting decreased with age, and the
+proportion of time infants spent standing increased with age. Both 8
+month-olds and 12 month-olds spent equivalent amounts of time
+lying/crawling, which was markedly decreased in the 16-month-olds, who
+spent most of their time sitting or standing (see Figure
+\ref{fig:posture}). We also observed changes in children's orientation
+relative to their caregivers: the 8-month-olds spent more time with
+their caregiver behind them supporting their sitting positions (see
+Figure \ref{fig:posture}).
 
@@
 Hands}\label{changes-in-access-to-faces-and-hands}
+
+We examined the proportion of face and hand detections across age (see
+Figure \ref{fig:detByAge}). We observed a slight U-shaped function in
+face detections, such that 12-month-olds appeared to have visual acces
+to slightly fewer faces faces than 8 or 16-month-olds; conversely, hand
+detections perhaps appeared to generally increase with age.
+
+Age related effects were much smaller than postural and locomotive
+changes on children's visual access to faces and hands. Children's
+posture was a major factor both in how many faces and hands they saw
+during the play session. Infants who were sitting saw more faces than
+infants who were lying down or being carried, while infants who were
+standing saw the most faces (Figure \ref{fig:detByPosOrient}, upper
+panel); this same pattern was also true for hand detections. Children's
+orientation also impacted their visual access to faces and hands:
+children who were far away from their caregiver were more likely to see
+faces/hands than children who were close to their caregiver (Figure
+\ref{fig:detByPosOrient}, lower panel).
 
@@
 \end{figure}
-\end{CodeChunk}\begin{CodeChunk}
+\end{CodeChunk}
+
+\begin{CodeChunk}
 \begin{figure*}[h]
@@
 
-\caption[Proportion face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant]{Proportion face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant.}\label{fig:detByPosOrient}
+\caption[Proportion of face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant]{Proportion of face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant.}\label{fig:detByPosOrient}
 \end{figure*}
@@
 
-We first examined the proportion of face and wrist detections across
-age; a full summary can be seen in Figure \ref{fig:detByAge}. When
-looking at face presence, we observed a slight U-shaped function both
-when analyzing the output of the MTCNN and OpenPose detectors, such that
-12-month-olds appeared to experience slighly fewer faces faces than 8 or
-16-month-olds.
+To formalize these observations, we fit two generalized linear
+mixed-effect models for the proportion of faces and the proportion of
+hands infants saw in each posture and orientation, with participant's
+age, orientation, and posture as fixed effects and with random slopes
+for infant's orientation. Random slopes for posture or interactions
+between our fixed effects caused the models to fail to converge.
 
-However, we found that any age related effects were much smaller
-compared to the impact of postural and locomotive changes on children's
-visual access to faces. Children's posture was a major factor both in
-how many faces and wrists/hands they saw during the play session.
-Infants who were sitting saw more faces than infants who were lying down
-or being carried, while infants who were standing saw the most faces
-(Figure \ref{fig:detByPosOrient}, upper panel); this same pattern was
-also true for wrists/hand detectinos. We also examined how the child's
-orientation relative to their caregiver impacted their visual access to
-faces and hands. Children who were far away from their caregiver were
-more likely to see faces than children who were close to their
-caregiver; this was true within all age groups (Figure
-\ref{fig:detByPosOrient}, lower panel).
-
-\begin{table}[H]
+\begin{table}[ht]
 \centering
@@
   \hline
-Intercept & -5.2469 & 0.0584 & -89.88 & 0.0000 \\ 
-  Age & 0.0847 & 0.0041 & 20.89 & 0.0000 \\ 
-  Prone & 0.2015 & 0.0564 & 3.57 & 0.0004 \\ 
-  Sit & 1.4053 & 0.0541 & 25.98 & 0.0000 \\ 
-  Stand & 1.4272 & 0.0542 & 26.33 & 0.0000 \\ 
-  Close & 1.8239 & 0.0230 & 79.17 & 0.0000 \\ 
-  Far & 2.5479 & 0.0239 & 106.42 & 0.0000 \\ 
+  \textbf{Faces} \\
+Intercept & -5.78 & 0.33 & -17.55 & 0.000 \\ 
+  Close & 2.09 & 0.31 & 6.70 & 0.000 \\ 
+  Far & 2.58 & 0.34 & 7.54 & 0.000 \\ 
+  Prone & -0.18 & 0.08 & -2.39 & 0.017 \\ 
+  Sit & 1.00 & 0.07 & 13.54 & 0.000 \\ 
+  Stand & 1.50 & 0.08 & 19.98 & 0.000 \\ 
+  Age & 0.17 & 0.17 & 0.97 & 0.334 \\ 
+  \hline
+  \textbf{Hands} \\
+  Intercept & -4.84 & 0.19 & -25.40 & 0.000 \\ 
+  Close & 0.88 & 0.17 & 5.25 & 0.000 \\ 
+  Far & 1.77 & 0.27 & 6.45 & 0.000 \\ 
+  Prone & 0.58 & 0.10 & 5.73 & 0.000 \\ 
+  Sit & 1.18 & 0.10 & 12.04 & 0.000 \\ 
+  Stand & 1.01 & 0.10 & 10.22 & 0.000 \\ 
+  Age & 0.14 & 0.10 & 1.40 & 0.161 \\ 
    \hline
 \end{tabular}
-\caption{Model coefficients from a generalized linear model predicting the proportion of faces seen by infants.} 
+\caption{Model coefficients from generalized linear models predicting the proportion of faces (upper panel) and wrists (lower panel) seen by children.} 
 \end{table}
 
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-Intercept & -5.0818 & 0.0776 & -65.46 & 0.0000 \\ 
-  Age & 0.0564 & 0.0050 & 11.35 & 0.0000 \\ 
-  Prone & 0.9499 & 0.0774 & 12.27 & 0.0000 \\ 
-  Sit & 1.7282 & 0.0758 & 22.79 & 0.0000 \\ 
-  Stand & 1.6618 & 0.0760 & 21.88 & 0.0000 \\ 
-  Close & 0.7472 & 0.0188 & 39.81 & 0.0000 \\ 
-  Far & 1.6357 & 0.0200 & 81.61 & 0.0000 \\ 
-   \hline
-\end{tabular}
-\caption{Model coefficients from a generalized linear model predicting the proportion of wrists seen by infants.} 
-\end{table}
+A summary of the coefficients of a models can be found in Table 2. While
+infant's posture and orientation significantly impacted the proportion
+of faces and hands that infants saw, age was not a significant predictor
+in either model. Thus, these results suggest that infants' visual access
+to social information is modulated by their posture and orientation,
+which is in turn a function of their general locomotor development.
 
-To formalize these observations, we fit a generalized linear model to
-the proportion of faces infants saw in each posture and orientation (as
-detected by MTCNN), with participant's age, orientation, and posture as
-independent variables. A summary of the coefficients of a model with
-only main effects (and no interactions) can be found in Table 3. When we
-did include interaction terms between age, posture, and orientation, age
-no longer remained a significant predictor (b = -15.84, SE = 113.98, z =
--0.14 , p = 0.89). We also found the same pattern with respect to
-infants' visual access to hands (see Table 4); including interaction
-terms also eliminated any main effect of age on the proportion of hands
-detected (b = -15.7, SE = 113.98, z = -0.14, p = 0.89). Thus, these
-results suggest that infants' access to social information is heavily
-influenced by their postural and locomotive developments.
+\subsection{Access to Faces and Hands During Labeling
+Events}\label{access-to-faces-and-hands-during-labeling-events}
+
+Finally, we explored how face and hand detections changed during object
+labeling events as a function of infant's posture and orientation. We
+analyzed a four-second window around each labeling event (e.g., ``Look
+at the {[}zem{]}!''); these labeling events were hand-annotated and
+synchronized with the frame-by-frame face/hand detections. We found that
+infants' posture and orientations modulated their visual access to faces
+and hands during labeling events; infants who were sitting or standing
+were more likely to have visual access to this social information (see
+Figure \ref{fig:detByNaming}). However, we did not find that infants saw
+particularily more faces or hands during naming events relative to
+baseline (avg. difference in proportion of hands, 8 m.o. = 0, 12 m.o. =
+-0.003, 16 m.o. = -0.014; avg. difference in proportion of faces, 8 m.o.
+= 0.003, 12 m.o. = 0.01, 16 m.o. = 0.021).
+
+\section{General Discussion}\label{general-discussion}
+
+We used a head-mounted camera to explore how children's postural and
+locomotor development affects their visual access to social information,
+here operationalized as the presence of the faces and hands of their
+caregiver. Children's posture and orientation towards their caregiver
+changed systematically across age, and both of these factors influenced
+the proportion of faces and hands that were available in the child's
+visual field. This work suggests that motor development modulates how
+infants experience their visual world and the social information in it:
+Infants that are sitting and standing have a different view of their
+world, the people in it, and the actions that are being performed.
 
@@
 
-\caption[Proportion face and wrist detections around a naming instance ('Look, a Zem']{Proportion face and wrist detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
+\caption[Proportion face and hand detections around a naming instance ('Look, a Zem']{Proportion face and hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
 Error bars represent non-parametric bootstrapped 95 percent confidence intervals.}\label{fig:detByNaming}
@@
 
-\subsection{Access to Faces and Hands During Labeling
-Events}\label{access-to-faces-and-hands-during-labeling-events}
+We created a situation in the lab in which the context of interaction
+was tightly controlled, but as children grow and change, the activities
+in which they engage with their caregivers also vary, leading to
+differences in the distribution of contexts they experience. Thus, more
+work is needed to understand how our results relate to children's home
+experiences (Fausey et al., 2016). The ability to walk is part of a
+cascade of changes in children's experience, and our study captures only
+a cross-sectional slice of this broader, multifaceted trajectory.
 
-Finally, we explored how face and wrist/hand detections changed during
-object labeling events as a function of infant's posture and
-orientation. Specfically, we analyzed a four-second window around each
-labeling event (e.g., ``Look at the {[}zem{]}!''); these labeling events
-were hand-annotated and automatically synchronized with the
-frame-by-frame face detections. As before, we found that infants'
-posture and orientations impacted the degree to which they saw their
-caregiver's face and wrist during a labeling event; infants who were
-sitting or standing were more likely to have access to this social
-information. A full summary of the results can be seen in Figure
-\ref{fig:detByNaming}. However, we did not find that infants saw more
-faces or more wrists/hands during naming events relative to baseline
-(average difference in proportion of wrists detected, 8 m.o. = 0, 12
-m.o. = -0.003, 16 m.o. = -0.014; average difference in proportion of
-faces detected, 8 m.o. = 0.003, 12 m.o. = 0.01, 16 m.o. = 0.021).
-
-\section{General Discussion}\label{general-discussion}
-
-We used a head-mounted camera to explore how children's postural and
-locomotive development directly impacts their access to social
-information, here operationalized as the presence of the faces and
-wrists of their caregiver. We found that children's posture and
-orientation towards their caregiver changed systematically across age,
-and that both of these factors dramatically impacted the proportion of
-faces and wrists/hands that were available in the child's visual field.
-Thus, infants' postural and locomotive developments are mediating
-factors that explain some of the age-related changes in the proportion
-of faces and wrists/hands that are visually available to infants.
-Broadly, this work suggests that motoric developments mediate how
-infants experience their visual world and the social information in it:
-infants that are sitting and standing have a different view of their
-world, the people in it, and the actions that are being performed.
-
-This work also deploys novel advancements in computer vision to the
-study of developmental psychology. The field of object detection and
-recognition has advanced dramatically in the past five years since the
-re-birth of deep learning algorithms (Krizhevsky, Sutskever, \& Hinton,
-2012), creating a new generation of algorithmic tools. These tools are
-substantially better equipped to deal with noisier, more complicated
-datasets and can extract richer and more detailed information. Videos
-from the infant perspective provided substantial challenges (e.g.,
-partially occluded faces) for the classic models of face detection
-(e.g., ViolaJones) (Viola \& Jones, 2004). Further, as the headcam
-technologies employed here were inexpensive and the computer vision
-algorithms freely available, this method is a promising avenue for
-quantifying the visual and social information available to infant
-learners.
-
-Broadly, we suggest that the combined use of these new tools can be
-leveraged to understand the changing infant perspective on the visual
-world and the implications of these changes for linguistic, cognitive,
-and social development.
+Understanding these changes has been a persistent challenge for
+developmental psychology, but the field of computer vision has advanced
+dramatically in recent years, creating a new generation of algorithmic
+tools. These tools deal better with noisier, more complicated datasets
+and extract richer and more detailed information than previous systems.
+We hope that these new tools can be leveraged to understand the changing
+infant perspective on the visual world and the implications of these
+changes for linguistic, cognitive, and social development.
 
@@
 
-Thanks to Kaia Simmons, Kathy Woo, Aditi Maliwal, and other members of
-the Language and Cognition Lab for help in recruitment, data collection,
-and annotation. This research was supported by a John Merck Scholars
-grant to MCF. An earlier version of this work was presented to the
-Cognitive Science Society in Frank et al. (2013). Please address
-correspondence to Michael C. Frank, Department of Psychology, Stanford
-University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel:
-(650) 724-4003, email: \texttt{mcfrank@stanford.edu}.
+Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in
+recruitment, data collection, and annotation. An earlier version of this
+work was presented to the Cognitive Science Society in Frank et al.
+(2013).
 
@@
 
-\hypertarget{ref-franchak2011head}{}
+\hypertarget{ref-franchak2011}{}
 Franchak, J. M., Kretch, K. S., Soska, K. C., \& Adolph, K. E. (2011).
@@
 
-\hypertarget{ref-gredeback2008microstructure}{}
-Gredebäck, G., Theuring, C., Hauf, P., \& Kenward, B. (2008). The
-microstructure of infants' gaze as they view adult shifts in overt
-attention. \emph{Infancy}, \emph{13}(5), 533--543.
-
 \hypertarget{ref-iverson2010}{}
@@
 \emph{85}(4), 1503--1518.
-
-\hypertarget{ref-krizhevsky2012imagenet}{}
-Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012). Imagenet
-classification with deep convolutional neural networks. In
-\emph{Advances in neural information processing systems} (pp.
-1097--1105).
 

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
edits and bib

@@
 # Introduction
-[^1]: These authors contributed equally to this work.
 
@@
 
-Then, however, their view of the world radically changes [@adolph2007].
+Then, however, their view of the world radically changes [@adolph2006motor].
 Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently.
@@
 
-One possibility is that these motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities [@iverson2010].
-These postural changes also impact how children interact with their mothers: walking (vs.
+These motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities in various ways [@iverson2010].
+For example, postural changes impact how children interact with their mothers: walking (vs.
 crawling) infants make different kinds of object-related bids for attention from their mothers and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].
@@
 For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child.
-All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012].
+All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012toward].
 Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
edits and bib

@@
 
-Then, however, their view of the world radically changes (K. Adolph \&
-Berger, 2007). Infants' motor abilities improve dramatically near the
+Then, however, their view of the world radically changes (Adolph \&
+Berger, 2006). Infants' motor abilities improve dramatically near the
 end of the first year of life, allowing them to locomote independently.
@@
 
-One possibility is that these motor improvements lead to developmental
-cascades that impact children's emerging social, cognitive, and
-linguistic abilities (Iverson, 2010). These postural changes also impact
-how children interact with their mothers: walking (vs.~crawling) infants
+These motor improvements lead to developmental cascades that impact
+children's emerging social, cognitive, and linguistic abilities in
+various ways (Iverson, 2010). For example, postural changes impact how
+children interact with their mothers: walking (vs.~crawling) infants
 make different kinds of object-related bids for attention from their
@@
 or side of the child. All annotations were made by a trained coder using
-the OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman, Sanderson,
-\& Millman, 2012). Times when the child was out of view of the tripod
+the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
+Millman, 2012). Times when the child was out of view of the tripod
 camera were marked as uncodable and were excluded from these
@@
 wrists, which we used as a proxy for hands for two reasons. First, as we
-did not capture children's entire visual field. the presence of a wrist
+did not capture children's entire visual field, the presence of a wrist
 is likely often indicative of the presence of a hand within the field of
@@
 
-The CNN-based pose detector (OpenPose, Cao et al., 2017; Simon, Joo,
+The CNN-based pose detector (OpenPose; Cao et al., 2017; Simon, Joo,
 Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016))
@@
 labeled frames. To account for the relatively rare appearance of faces
-in the dataset, we hand abeled two types of samples: a sample containing
-a high density of faces (half reported by MTCNN, half by OpenPose) and a
-random sample from the remaining frames. Each sample was comprised of an
-equal number of frames taken from each child's video. For wrist
-detections, the ``gold set'' was constructed in the same manner, except
-frames with a high density of wrists came only from detections made by
-OpenPose. Faces were classified as present if at least half of the face
-was showing; wrists were classified as present if any part of the wrist
-was showing. Precision (hits / hits + false alarms), recall (hits / hits
-+ misses), and F-score (harmonic mean of precision and recall) were
-calculated for all detectors and are reported in Table 2.
+in the dataset, we hand-labeled two types of samples: a sample
+containing a high density of faces (half reported by MTCNN, half by
+OpenPose) and a random sample from the remaining frames. Each sample was
+comprised of an equal number of frames taken from each child's video.
+For wrist detections, the ``gold set'' was constructed in the same
+manner, except frames with a high density of wrists came only from
+detections made by OpenPose. Faces were classified as present if at
+least half of the face was showing; wrists were classified as present if
+any part of the wrist was showing. Precision (hits / hits + false
+alarms), recall (hits / hits + misses), and F-score (harmonic mean of
+precision and recall) were calculated for all detectors and are reported
+in Table 2.
 
@@
 The proportion of time infants spent sitting decreased with age, and the
-proportion of time infants spent standing increased with age. Both 8
-month-olds and 12 month-olds spent equivalent amounts of time
+proportion of time infants spent standing increased with age. Both
+8-month-olds and 12-month-olds spent equivalent amounts of time
 lying/crawling, which was markedly decreased in the 16-month-olds, who
@@
 Figure \ref{fig:detByAge}). We observed a slight U-shaped function in
-face detections, such that 12-month-olds appeared to have visual acces
+face detections, such that 12-month-olds appeared to have visual access
 to slightly fewer faces faces than 8 or 16-month-olds; conversely, hand
@@
 
-A summary of the coefficients of a models can be found in Table 2. While
-infant's posture and orientation significantly impacted the proportion
-of faces and hands that infants saw, age was not a significant predictor
-in either model. Thus, these results suggest that infants' visual access
-to social information is modulated by their posture and orientation,
-which is in turn a function of their general locomotor development.
+A summary of the coefficients of the models can be found in Table 2.
+While infant's posture and orientation significantly impacted the
+proportion of faces and hands that infants saw, age was not a
+significant predictor in either model. Thus, these results suggest that
+infants' visual access to social information is modulated by their
+posture and orientation, which is in turn a function of their general
+locomotor development.
 
@@
 Finally, we explored how face and hand detections changed during object
-labeling events as a function of infant's posture and orientation. We
+labeling events as a function of infants' posture and orientation. We
 analyzed a four-second window around each labeling event (e.g., ``Look
@@
 Figure \ref{fig:detByNaming}). However, we did not find that infants saw
-particularily more faces or hands during naming events relative to
+particularly more faces or hands during naming events relative to
 baseline (avg. difference in proportion of hands, 8 m.o. = 0, 12 m.o. =
@@
 
-\caption[Proportion face and hand detections around a naming instance ('Look, a Zem']{Proportion face and hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
+\caption[Proportion of face and hand detections around a naming instance ('Look, a Zem']{Proportion of face and hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
 Error bars represent non-parametric bootstrapped 95 percent confidence intervals.}\label{fig:detByNaming}
@@
 \hypertarget{refs}{}
-\hypertarget{ref-adolph2007}{}
-Adolph, K., \& Berger, S. (2007). Motor development. In \emph{Handbook
-of child psychology}. Wiley Online Library.
+\hypertarget{ref-adolph2006motor}{}
+Adolph, K. E., \& Berger, S. E. (2006). Motor development.
+\emph{Handbook of Child Psychology}.
 
-\hypertarget{ref-adolph2012}{}
-Adolph, K., Gilmore, R., Freeman, C., Sanderson, P., \& Millman, D.
-(2012). Toward open behavioral science. \emph{Psychological Inquiry},
+\hypertarget{ref-adolph2012toward}{}
+Adolph, K. E., Gilmore, R. O., Freeman, C., Sanderson, P., \& Millman,
+D. (2012). Toward open behavioral science. \emph{Psychological Inquiry},
 \emph{23}(3), 244--247.

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
one more change

@@
 Here, we explore this question by using head-mounted cameras to record 36 infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to estimate the proportion of faces and hands in infants' environments.
-We find that infants' posture and orientation to their caregiver modulates their access to social information, confirming previous work that suggests motoric developments play a significant role in shaping the input to children's developing linguistic and social capacities.
+We find that infants' posture and orientation to their caregiver modulates their access to social information, confirming previous work that suggests motoric developments play a significant role in the emergence of children's linguistic and social capacities.
 We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience."

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
one more change

@@
 previous work that suggests motoric developments play a significant role
-in shaping the input to children's developing linguistic and social
-capacities. We suggest that the combined use of head-mounted cameras and
-the application of new computer vision techniques is a promising avenue
-for understanding the statistics of infants' visual and linguistic
+in the emergence of children's linguistic and social capacities. We
+suggest that the combined use of head-mounted cameras and the
+application of new computer vision techniques is a promising avenue for
+understanding the statistics of infants' visual and linguistic
 experience.

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
fixed bib and special characters

@@
 output: cogsci2016::cogsci_paper
+
 ---
-
+\newcommand{\wrapmf}[1]{#1}
+\vspace{-1em}
 ```{r global_options, include=FALSE}
@@
 library(assertthat)
+
 ```
 # Introduction
-
 From early infancy, children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like].
 Even newborns tend to prefer to look at faces with direct vs.
-averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010development].
+averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010].
 Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
@@
 By using head-mounted cameras, researchers have begun to record the visual experiences of infants and children -- which even for walking children are extremely different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008].
-Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2012].
+Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2013statistical].
 This perspective changes over the first two years of life, as views transition from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
@@
 }
+```
 
-## EVALUATE FACE DETECTORS
+```{r  include=FALSE, echo=FALSE}
+## Evalute face and wrist detectors
 performanceFaces <- ground_truth_faces  %>%
@@
 Scores in bold are the highest F-scores for each sample type.}
+\vspace{-1em}
 \end{table}
@@
 \caption{Model coefficients from generalized linear models predicting the proportion of faces (upper panel) and wrists (lower panel) seen by children.} 
+\vspace{-1em}
 \end{table}
@@
 Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation.
+This work was funded by a Jacobs Foundation Fellowship and a John Merck Scholars award to MCF.
 An earlier version of this work was presented to the Cognitive Science Society in @frank2013.
+
+\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/amsan7/xs-face}}} \vspace{1em}
 

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
fixed video length & bibliography things

@@
 \newcommand{\wrapmf}[1]{#1}
-\vspace{-1em}
+
 ```{r global_options, include=FALSE}
@@
 Even newborns tend to prefer to look at faces with direct vs.
-averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010].
+averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010development].
 Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
@@
   left_join(inc_subs) %>%
+  filter(inc==TRUE) %>%
   group_by(age_group) %>%
@@
 ```{r headcam, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2.5, set.cap.width=T, num.cols.cap=1, fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}
+
 img <- png::readPNG("images/viewangle_cut.png")
@@
 Headcam videos were trimmed such that they excluded the instruction phase when the experimenter was in the room and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software.
-These sessions yielded videos of
-`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a million frames), with an average video length of
-`r round(sum(inc_subs$len.min)/60, digits = 2)` minutes
+These sessions yielded
+`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a million frames) of video, with an average video length of
+`r round(mean(inc_subs$len.min), digits = 2)` minutes
 (min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
minor edits and back to 6 pages

@@
 However, as children reach their first birthday, they begin to locomote themselves, dramatically affecting their visual access to this information.
-How do these postural and locomotive changes affect children's access to the social information relevant for word-learning?
+How do these postural and locomotor changes affect children's access to the social information relevant for word-learning?
 Here, we explore this question by using head-mounted cameras to record 36 infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to estimate the proportion of faces and hands in infants' environments.
@@
 
-These motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities in various ways [@iverson2010].
+These motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities in various ways.
 For example, postural changes impact how children interact with their mothers: walking (vs.
@@
 Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
-On their account, children’s ability to stand and independently locomote may change their ability to access social information (e.g., faces, gaze)  and in turn to accelerate their own learning.
+Thus, children’s ability independently locomote may change their ability to access social information (e.g., faces, gaze) and in turn accelerate their learning.
 
@@
 
-Here, we directly examine whether postural and locomotive developments change the availability of social information -- the presence of faces and hands.
+Here, we directly examine whether postural and locomotor developments change the availability of social information -- the presence of faces and hands.
 To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
@@
 
-The CNN-based pose detector  [OpenPose; @cao2017realtime; @simon2017hand; @wei2016cpm]) provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
+The CNN-based pose detector  [OpenPose; @cao2017realtime; @simon2017hand; @wei2016cpm] provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose.
 The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates.
@@
 
-Age related effects were much smaller than postural and locomotive changes on children's visual access to faces and hands.
+Age related effects were much smaller than postural and locomotor changes on children's visual access to faces and hands.
 Children's posture was a major factor both in how many faces and hands they saw during the play session.
@@
 
-```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=3, fig.cap = "Proportion of face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant."}
+```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=6, fig.height=3, fig.cap = "Proportion of face and wrist detections in each age group as a function of children's posture (top panel) and orientation (bottom panel)."}
 
@@
   labs(y = "Prop. detected", x = element_blank()) +
-  facet_grid(~detectorType, scales="free_y") +
+  facet_grid(~detectorType) +
   theme_few() +
-  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
+  scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6))
 
@@
   labs(y = "Prop. detected", x = element_blank()) +
-  facet_grid(~detectorType, scales="free_y") +
+  facet_grid(~detectorType) +
   theme_few() +
-  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))  +
+  scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6))
 
@@
 
-```{r results="asis"}
-#print(faceGLMOut_noInt, type="latex", comment = F, table.placement = "H")
-```
-
-```{r results="asis"}
-#print(wristGLMOut_noInt, type="latex", comment = F, table.placement = "H")
-```
-
-```{r echo=FALSE, include=FALSE}
-## Mixed effect model tries - many convergence issues
-
-# Interactions fail to converge, as does the one below 
-# fullModelminsInt_Faces_MTCNN <- glmer(faceMT ~ scale(age.at.test)+(posture)+(orientation) + (1|subid), 
-#        data = d, 
-#        family = "binomial")
-
-## Reading https://rdrr.io/cran/lme4/man/convergence.html 
-## Check for singularity - not an issue 
-# fm1=fullModelminsInt_Faces_MTCNN 
-# diag.vals <- getME(fm1,"theta")[getME(fm1,"lower") == 0] 
-# any(diag.vals < 1e-6) # FALSE
-
-## But has issues with every optimizer... -->
-#source(system.file("utils", "allFit.R", package="lme4")) 
-#fm1.all <- allFit(fm1)
-#ss <- summary(fm1.all)
-
-#No convergence issues with open pose detectors for faces, but for wrists
-#fullModelminsInt_Faces_OP <- glmer(faceOP ~ scale(age.at.test) + posture + orientation + (1|subid), 
-#      data = d, 
-#        family = "binomial") 
-```
-
 # General Discussion
-
 We used a head-mounted camera to explore how children's postural and locomotor development affects their visual access to social information, here operationalized as the presence of the faces and hands of their caregiver.
@@
 ```{r detByNaming, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3, fig.height=3, fig.cap = "Proportion of face and hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
-Error bars represent non-parametric bootstrapped 95 percent confidence intervals."}
+Error bars represent non-parametric bootstrapped 95 \\% confidence intervals."}
 
@@
 Understanding these changes has been a persistent challenge for developmental psychology, but the field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools.
-These tools deal better with noisier, more complicated datasets and extract richer and more detailed information than previous systems.
+These tools deal better with noisier, more complicated datasets and extract richer information than previous systems.
 We hope that these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
grant acknowledgement add
oops reworded to get back that last line
more typo fixes

@@
 Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
-Thus, children’s ability independently locomote may change their ability to access social information (e.g., faces, gaze) and in turn accelerate their learning.
+Thus, children’s ability to independently locomote may change their ability to access social information (e.g., faces, gaze) and in turn accelerate their learning.
 
@@
 Here, we directly examine whether postural and locomotor developments change the availability of social information -- the presence of faces and hands.
-To do so, we recorded the visual experience of a group of infants in three age ranges (8,12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
+To do so, we recorded the visual experience of a group of infants in three age ranges (8, 12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
 We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
@@
 Videos captured by the headcam were 720x480 pixels with 25 frames per second [^2].
-A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, bottom).
+A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, above).
 
@@
 Faces were classified as present if at least half of the face was showing; wrists were classified as present if any part of the wrist was showing.
-Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 2.
+Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 1.
 
@@
 We thus use MTCNN detections in the following analyses.
-For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision but low recall on the randomly sampled frames (see Table 2).
+For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision but low recall on the randomly sampled frames (see Table 1).
 We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of hands in the dataset.
@@
 We examined the proportion of face and hand detections across age (see Figure \ref{fig:detByAge}).
-We observed a slight U-shaped function in face detections, such that 12-month-olds appeared to have visual access to slightly fewer faces faces than 8 or 16-month-olds; conversely, hand detections perhaps appeared to generally increase with age.
+We observed a slight U-shaped function in face detections, such that 12-month-olds appeared to have visual access to slightly fewer faces than 8 or 16-month-olds; conversely, hand detections perhaps appeared to generally increase with age.
 
@@
 ```
-To formalize these observations, we fit two generalized linear mixed-effect models for the proportion of faces and the proportion of hands infants saw in each posture and orientation, with participant's age, orientation, and posture as fixed effects and with random slopes for infant's orientation.
+To formalize these observations, we fit two generalized linear mixed-effect models for the proportion of faces and the proportion of hands infants saw in each posture and orientation, with participant's age, orientation, and posture as fixed effects and with random slopes for infants' orientation.
 Random slopes for posture or interactions between our fixed effects caused the models to fail to converge.
@@
 A summary of the coefficients of the models can be found in Table 2.
-While infant's posture and orientation significantly impacted the proportion of faces and hands that infants saw, age was not a significant predictor in either model.
+While infants' posture and orientation significantly impacted the proportion of faces and hands that infants saw, age was not a significant predictor in either model.
 Thus, these results suggest that infants' visual access to social information is modulated by their posture and orientation, which is in turn a function of their general locomotor development.
@@
 Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation.
-This work was funded by a Jacobs Foundation Fellowship and a John Merck Scholars award to MCF.
+This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and a NSF #1714726 to BLL.
 An earlier version of this work was presented to the Cognitive Science Society in @frank2013.

### 2018-01-31 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
more typo fixes

@@
 dramatically affecting their visual access to this information. How do
-these postural and locomotive changes affect children's access to the
+these postural and locomotor changes affect children's access to the
 social information relevant for word-learning? Here, we explore this
@@
 
+\newcommand{\wrapmf}[1]{#1}
+
 \section{Introduction}\label{introduction}
@@
 to look at faces with direct vs.~averted gaze (Farroni, Csibra, Simion,
-\& Johnson, 2002) and young infants follow overt gaze shifts (Gredebäck,
-Fikke, \& Melinder, 2010). Further, when infants attend to video
-stimuli, they tend to look mostly at faces at the expense of other
-visual information -- though older infants start to look towards
-people's hands and the actions they are performing (Frank, Amso, \&
-Johnson, 2014; Frank, Vul, \& Saxe, 2012).
+\& Johnson, 2002) and young infants follow overt gaze shifts
+(Gredeb\wrapmf{\"{a}}ck, Fikke, \& Melinder, 2010). Further, when
+infants attend to video stimuli, they tend to look mostly at faces at
+the expense of other visual information -- though older infants start to
+look towards people's hands and the actions they are performing (Frank,
+Amso, \& Johnson, 2014; Frank, Vul, \& Saxe, 2012).
 
@@
 children's emerging social, cognitive, and linguistic abilities in
-various ways (Iverson, 2010). For example, postural changes impact how
-children interact with their mothers: walking (vs.~crawling) infants
-make different kinds of object-related bids for attention from their
-mothers and tend to hear more action directed statements (e.g., ``open
-it'') (Karasik, Tamis-LeMonda, \& Adolph, 2014). Further, in an
-observational study, Walle \& Campos (2014) found that children who were
-able to walk had both higher receptive and productive vocabularies. On
-their account, children's ability to stand and independently locomote
-may change their ability to access social information (e.g., faces,
-gaze) and in turn to accelerate their own learning.
+various ways. For example, postural changes impact how children interact
+with their mothers: walking (vs.~crawling) infants make different kinds
+of object-related bids for attention from their mothers and tend to hear
+more action directed statements (e.g., ``open it'') (Karasik,
+Tamis-LeMonda, \& Adolph, 2014). Further, in an observational study,
+Walle \& Campos (2014) found that children who were able to walk had
+both higher receptive and productive vocabularies. Thus, children's
+ability independently locomote may change their ability to access social
+information (e.g., faces, gaze) and in turn accelerate their learning.
 
@@
 learning objects and their labels than the comparable adult perspective
-(Bambach, Crandall, Smith, \& Yu, 2017; D. Yurovsky, Smith, \& Yu, in
-press). This perspective changes over the first two years of life, as
-views transition from primarily containing close up views of faces to
+(Bambach, Crandall, Smith, \& Yu, 2017; Yurovsky, Smith, \& Yu, 2013).
+This perspective changes over the first two years of life, as views
+transition from primarily containing close up views of faces to
 capturing views of hands paired with the objects they are acting on
@@
 
-Here, we directly examine whether postural and locomotive developments
+Here, we directly examine whether postural and locomotor developments
 change the availability of social information -- the presence of faces
@@
 synchronized with the tripod-mounted videos using FinalCut Pro Software.
-These sessions yielded videos of 516 minutes (almost a million frames),
-with an average video length of 8.6 minutes (min = 4.53, max = 19.35).
+These sessions yielded 516 minutes (almost a million frames) of video,
+with an average video length of 14.33 minutes (min = 4.53, max = 19.35).
 
@@
 The CNN-based pose detector (OpenPose; Cao et al., 2017; Simon, Joo,
-Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016))
+Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016)
 provided the locations of 18 body parts (ears, nose, wrists, etc.) and
@@
 
-Age related effects were much smaller than postural and locomotive
+Age related effects were much smaller than postural and locomotor
 changes on children's visual access to faces and hands. Children's
@@
 
-\caption[Proportion of face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant]{Proportion of face and wrist detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant.}\label{fig:detByPosOrient}
+\caption[Proportion of face and wrist detections in each age group as a function of children's posture (top panel) and orientation (bottom panel)]{Proportion of face and wrist detections in each age group as a function of children's posture (top panel) and orientation (bottom panel).}\label{fig:detByPosOrient}
 \end{figure*}
@@
 \caption[Proportion of face and hand detections around a naming instance ('Look, a Zem']{Proportion of face and hand detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.
-Error bars represent non-parametric bootstrapped 95 percent confidence intervals.}\label{fig:detByNaming}
+Error bars represent non-parametric bootstrapped 95 \% confidence intervals.}\label{fig:detByNaming}
 \end{figure}
@@
 tools. These tools deal better with noisier, more complicated datasets
-and extract richer and more detailed information than previous systems.
-We hope that these new tools can be leveraged to understand the changing
-infant perspective on the visual world and the implications of these
-changes for linguistic, cognitive, and social development.
+and extract richer information than previous systems. We hope that these
+new tools can be leveraged to understand the changing infant perspective
+on the visual world and the implications of these changes for
+linguistic, cognitive, and social development.
 
@@
 Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in
-recruitment, data collection, and annotation. An earlier version of this
-work was presented to the Cognitive Science Society in Frank et al.
-(2013).
+recruitment, data collection, and annotation. This work was funded by a
+Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF,
+and a NSF \#1714726 to BLL. An earlier version of this work was
+presented to the Cognitive Science Society in Frank et al. (2013).
+
+\vspace{1em}
+
+\fbox{\parbox[b][][c]{7.3cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/amsan7/xs-face}}}
+\vspace{1em}
 
@@
 \hypertarget{ref-gredeback2010development}{}
-Gredebäck, G., Fikke, L., \& Melinder, A. (2010). The development of
-joint visual attention: A longitudinal study of gaze following during
-interactions with mothers and strangers. \emph{Developmental Science},
-\emph{13}(6), 839--848.
-
-\hypertarget{ref-iverson2010}{}
-Iverson, J. M. (2010). Developing language in a developing body: The
-relationship between motor development and language development.
-\emph{Journal of Child Language}, \emph{37}(2), 229--261.
+Gredeb\wrapmf{\"{a}}ck, G., Fikke, L., \& Melinder, A. (2010). The
+development of joint visual attention: A longitudinal study of gaze
+following during interactions with mothers and strangers.
+\emph{Developmental Science}, \emph{13}(6), 839--848.
 
@@
 
-\hypertarget{ref-yurovsky2012}{}
-Yurovsky, D., Smith, L., \& Yu, C. (in press). Statistical word learning
-at scale: The baby's view is better. \emph{Developmental Science}.
+\hypertarget{ref-yurovsky2013statistical}{}
+Yurovsky, D., Smith, L. B., \& Yu, C. (2013). Statistical word learning
+at scale: The baby's view is better. \emph{Developmental Science},
+\emph{16}(6), 959--966.
 

### 2018-02-01 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
mikes last edis

@@
 author-information: > 
+    \renewcommand*{\thefootnote}{\fnsymbol{footnote}} 
     \author{{\large \bf Alessandro Sanchez\thanks{These authors contributed equally to this work.}} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University
@@
 \newcommand{\wrapmf}[1]{#1}
+\renewcommand*{\thefootnote}{\arabic{footnote}}
+\setcounter{footnote}{0}
 
@@
 By using head-mounted cameras, researchers have begun to record the visual experiences of infants and children -- which even for walking children are extremely different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008].
-Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggest that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2013statistical].
-This perspective changes over the first two years of life, as views transition from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
-These changes may be driven by – or independent from – the concurrent motoric changes described above.
+Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggests that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2013statistical].
+This perspective changes over the first two years of life, as visual input transitions from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
+It is uncertain how much these changes are driven by the concurrent motoric changes described above versus by other developmental factors.
 
@@
 Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation.
-This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and a NSF #1714726 to BLL.
+This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and NSF #1714726 to BLL.
 An earlier version of this work was presented to the Cognitive Science Society in @frank2013.

### 2018-02-01 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
mikes last edis

@@
 
+\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
+
 \author{{\large \bf Alessandro Sanchez\thanks{These authors contributed equally to this work.}} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long\footnotemark[1]}  \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Allison M.
@@
 \newcommand{\wrapmf}[1]{#1}
+\renewcommand*{\thefootnote}{\arabic{footnote}}
+
+\setcounter{footnote}{0}
 
@@
 both higher receptive and productive vocabularies. Thus, children's
-ability independently locomote may change their ability to access social
-information (e.g., faces, gaze) and in turn accelerate their learning.
+ability to independently locomote may change their ability to access
+social information (e.g., faces, gaze) and in turn accelerate their
+learning.
 
@@
 hands (Yoshida \& Smith, 2008), and computational and empirical work
-suggest that this restricted viewpoint may be more effective for
+suggests that this restricted viewpoint may be more effective for
 learning objects and their labels than the comparable adult perspective
 (Bambach, Crandall, Smith, \& Yu, 2017; Yurovsky, Smith, \& Yu, 2013).
-This perspective changes over the first two years of life, as views
-transition from primarily containing close up views of faces to
+This perspective changes over the first two years of life, as visual
+input transitions from primarily containing close up views of faces to
 capturing views of hands paired with the objects they are acting on
-(Fausey, Jayaraman, \& Smith, 2016). These changes may be driven by --
-or independent from -- the concurrent motoric changes described above.
+(Fausey, Jayaraman, \& Smith, 2016). It is uncertain how much these
+changes are driven by the concurrent motoric changes described above
+versus by other developmental factors.
 
@@
 and hands. To do so, we recorded the visual experience of a group of
-infants in three age ranges (8,12, and 16 months) using head-mounted
+infants in three age ranges (8, 12, and 16 months) using head-mounted
 cameras during a brief laboratory free-play session; children's posture
@@
 \(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
-\ref{fig:headcam}, bottom).
+\ref{fig:headcam}, above).
 
@@
 precision and recall) were calculated for all detectors and are reported
-in Table 2.
+in Table 1.
 
@@
 OpenPose performed moderately well (F = 0.74) with relatively high
-precision but low recall on the randomly sampled frames (see Table 2).
+precision but low recall on the randomly sampled frames (see Table 1).
 We thus analyze wrist detections, with the caveat that we are likely
@@
 face detections, such that 12-month-olds appeared to have visual access
-to slightly fewer faces faces than 8 or 16-month-olds; conversely, hand
+to slightly fewer faces than 8 or 16-month-olds; conversely, hand
 detections perhaps appeared to generally increase with age.
@@
 age, orientation, and posture as fixed effects and with random slopes
-for infant's orientation. Random slopes for posture or interactions
+for infants' orientation. Random slopes for posture or interactions
 between our fixed effects caused the models to fail to converge.
@@
 A summary of the coefficients of the models can be found in Table 2.
-While infant's posture and orientation significantly impacted the
+While infants' posture and orientation significantly impacted the
 proportion of faces and hands that infants saw, age was not a
@@
 Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF,
-and a NSF \#1714726 to BLL. An earlier version of this work was
-presented to the Cognitive Science Society in Frank et al. (2013).
+and NSF \#1714726 to BLL. An earlier version of this work was presented
+to the Cognitive Science Society in Frank et al. (2013).
 

### 2018-02-01 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
last change

@@
 \end{tabular}
-\caption{Model coefficients from generalized linear models predicting the proportion of faces (upper panel) and wrists (lower panel) seen by children.} 
+\caption{Model coefficients from generalized linear models predicting the proportion of faces (upper panel) and hands (lower panel) seen by children.} 
 \vspace{-1em}

### 2018-03-13 analysis-main/main_analyses_bll.Rmd
cleanup of detector/data merge scripts

@@
----
+a---
 title: "XSFACE Analyses"

### 2018-03-13 analysis-main/helper.R

@@ -1,23 +1,30 @@
 library(stringr)
+library(assertthat)
 
 ######## MERGE IN EXACT TIMES FOR FRAMES #########
 add.times <- function(x) {
   print(x$subid[1])
-  fname <- paste0("../data/frame_times/",
+
+  fname <- paste0("../data/frame_times_redone/",
                  str_sub(as.character(x$subid[1]),start=4,end=8),
                  ".csv")
+  
+  
   if (file.exists(fname)) {
     times <- read_csv(fname)
+    times$time <- times$best_time ## best estimate of timestamp
+    assert_that(sum(is.na(times$time))==0)
     
-#     x <- merge(x,times, by.x = "frame", by.y = "frame", all.x = TRUE, all.y = FALSE)
-    x <- left_join(x, times)
-    x$dt <- c(diff(x$time),.032) ## assumes data is sorted by frames! also not sure why .032?
+    print(paste0("**** ", x$subid[1], 'det frame count', max(x$frame), 'time frame count', max(times$frame), "****"))
+    ## 
+    x<-left_join(x, times)
+    x$dt <- c(diff(x$time),.032)
+    
   } else {
     print(paste0("**** ", x$subid[1], " is missing frames ****"))
     x$time <- NA
     x$dt <- NA
   }
-  
   return(x)
 }
 
@@ -97,7 +104,6 @@ regularize.naming <- function (n) {
   n$name[n$name == "truck"] <- "car"
   n$name[n$name == "kittycat"] <- "cat"
   n$name[n$name == "bobcat"] <- "cat"
-
   n$name <- factor(n$name, levels=words)
   
   n$familiarity <- "familiar"
@@ -108,12 +114,23 @@ regularize.naming <- function (n) {
     n$name == "manu"] <- "novel"
   n$familiarity <- factor(n$familiarity)
   
-  # make the times seconds since onset
-  n$time <- as.numeric(difftime(strptime(n$time,"%M:%OS"),
-                                strptime("00:00:00.0","%H:%M:%OS"),units="secs"))
-  
+  # make the times seconds since onset - # if excel output is HMS vs MS - ugh! this was causing a big bug in analyses prior to jan 2018
+  # -- not evident in excel files can see in sublime -- different subs timestamps have different formats
+  op <- options(digits.secs = 3) ## needs this to grab milliseconds precision if it not called beforehand
+  timeBefore=n$time
[…]

### 2018-03-13 analysis-main/merge_all_data.R

@@ -1,3 +1,5 @@
+
+
 ## main data merge script - prerequisite for all other analyses
 rm(list=ls())
 library(dplyr)
@@ -7,14 +9,14 @@ library(tidyr)
 library(assertthat)
 source("helper.R")
 
-## load detectors for mtcnn2
+## load detectors for mtcnn
 dets <- read_csv("../data/final_output/mtcnn3.csv") %>%
   mutate(subid = video) %>%
   mutate(frame = as.numeric(frame)) %>%
   mutate(faceMT = as.logical(is_face)) %>%
   distinct(video,frame,.keep_all=TRUE)   
 
-# open pose detectorsa
+# open pose detections - faces and wrists
 detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") 
 detsOpenPose <- detsOpenPose %>%
   mutate(video = name) %>%
@@ -23,28 +25,31 @@ detsOpenPose <- detsOpenPose %>%
   mutate(faceOP = Nose_conf!=0)  %>%
   mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
 
-# open pose detectors
-detsOpenPose_Hands <- read_csv("../data/final_output/openpose_results_with_detail_hands.csv") 
-detsOpenPose_HandConf <- detsOpenPose_Hands %>%
-  mutate(video = name) %>%
-  distinct(video,frame,.keep_all=TRUE)  %>%
-  mutate(frame = as.numeric(frame) + 1) %>%
-  select(video,frame,ends_with("conf")) %>%
-  mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
-  mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
-  mutate(handOP = confRight>0 | confLeft>0 )
+# # open pose detectors - hands -- not using because poor accuracy
+# detsOpenPose_Hands <- read_csv("../data/final_output/openpose_results_with_detail_hands.csv") 
+# detsOpenPose_HandConf <- detsOpenPose_Hands %>%
+#   mutate(video = name) %>%
+#   distinct(video,frame,.keep_all=TRUE)  %>%
+#   mutate(frame = as.numeric(frame) + 1) %>%
+#   select(video,frame,ends_with("conf")) %>%
+#   mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
+#   mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
[…]

### 2018-03-13 analysis-validation/fps_sanity_checks.Rmd
expand sanity check frame scripts

---
title: "FPS Sanity Checks"
output: html_notebook
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
```

```{r, message=FALSE, warning=FALSE}
new <- read_csv("../data/video_stats/new_movie_lengths.csv")
old <- read_csv("../data/video_stats/old_movie_lengths.csv")
```

```{r}
new %>%
  rename(new_fps = fps, new_length = length) %>%
  inner_join(old %>%
              mutate(old_fps = last.frame / length) %>%
              rename(old_length = length), 
            by = "subid") %>%
  mutate(length_close_enough = abs(new_length - old_length) <= 1) %>%
  mutate(fps_close_enough =  abs(new_fps - old_fps) <= 1) %>%
  kable
```
Notes:

XS_1229: Was trimmed from when headcam fell out and therefore is a different lengthj
XS_1628: Was exported with wrong frame count in Oct 2017; error was not caught until March 2018 (error was in CogSCi 2018
initial submission, corrected for revisions)

### 2018-03-13 analysis-validation/frame_validation/check_frame_counts_post_detection.Rmd
expand sanity check frame scripts

@@
 ```{r}
+
 library(tidyverse)
 library(knitr)
+
 ```
@@
 ```{r}
-ffprobe_frame_counts <- read_csv("data/video_stats/ffprobe_frame_count.csv")
-extracted_frame_counts <- read_csv("data/video_stats/extracted_frame_count.csv")
-mtcnn_data <- read_csv("data/final_output/mtcnn_face_detections.csv")
+ffprobe_frame_counts <- read_csv("../../data/video_stats/ffprobe_frame_count.csv")
+extracted_frame_counts <- read_csv("../../data/video_stats/extracted_frame_count.csv")
+mtcnn_data <- read_csv("../../data/final_output/mtcnn3.csv")
 # openpose_data 
@@
 ```{r}
-ffprobe_frame_counts %>%
-  all.equal(extracted_frame_counts)
+s
 ```

### 2018-03-13 analysis-validation/length_sanity_check.Rmd
expand sanity check frame scripts

@@
 title: "xsface - file length sanity checking"
-author: "Mike Frank"
+author: "Mike Frank, continued by Bria Long & Alessandro Sanchez"
 date: "July 11, 2016"
@@
 2. demographic sheet (containing session lengths)
-3. detector outputs
+3. detector outputs (MTCNN/ OpenPose)
+4. timing outputs from ffprobe
 
@@
 library(lubridate)
+library(stringr)
+
 ```
 
-read in lengths on movies
+read in lengths on movies, demographics, detections, and frame time outputs
 
@@
 movies <- read_csv("../data/video_stats/new_movie_lengths.csv") 
-dets <- read_csv("../data/final_output/mtcnn3.csv")
+MTCNN <- read_csv("../data/final_output/mtcnn3.csv")
+openPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv")
+frameTimes <- read_csv("../data/video_stats/frametimechecks_showframes.csv")
 ```
@@
 # standardize
-dets$frame=as.numeric(dets$frame)
-dets$subid=dets$video
+MTCNN$frame=as.numeric(MTCNN$frame)
+MTCNN$subid=MTCNN$video
+openPose$frame=as.numeric(openPose$frame)
+openPose$subid=openPose$name
 
-det.lens <- dets %>%
+## get out max frame
+MTCNN.lens <- MTCNN %>%
   group_by(subid) %>%
-  summarise(frames_dets = max(frame))
+  summarise(frames_dets_mtcnn = max(frame))
+
+openPose.lens <- openPose %>%
+  group_by(subid) %>%
+  summarise(frames_dets_op = max(frame))
+
+## get frames from frame time output (ffprobe)
+frameTimes <- frameTimes %>%
+  mutate(last_frame_times = last.frame) %>%
+  mutate(subid = str_replace_all(subid,"-","_")) %>%
+  mutate(subid = str_sub(as.character(subid),start=1,end=7))
+
 ```
@@
 all_data <- left_join(demo, movies) %>%
-  left_join(det.lens) %>%
-  select(subid, recorded_len, last_frame_movie, length_movie, frames_dets) %>%
-  mutate(det_frame_err = abs(last_frame_movie - frames_dets) > 100,
-         demo_time_err = abs(recorded_len - length_movie) > 1) %>%
+  left_join(MTCNN.lens) %>%
+  left_join(openPose.lens) %>%
+  left_join(frameTimes) %>%
+  select(subid, recorded_len, length_movie, last_frame_movie, last_frame_times, frames_dets_mtcnn, frames_dets_op) %>%
+  mutate(det_frame_err = abs(last_frame_movie - frames_dets_mtcnn) > 100,
+         demo_time_err = abs(recorded_len - length_movie) > 1,
+         frame_time_diff = last_frame_times - frames_dets_mtcnn) %>%
   data.frame
+
+sum(all_data$frame_time_diff)
 ```
@@
 
-Explanations of discrepancies:
+XS_1229 has been trimmed (headcam came off) and therefore there are different movie lenghts.
+
+=====
+
+Mike's notes (Explanations of old discrepancies :)
 

### 2018-03-13 data/frame_times_redone/scripts/show_packets.py
reorganize frame time extractor scripts/folders

@@ -1,0 +1,39 @@
+#!/usr/bin/env python
+
+# run this locally to take FFMPEG packet files and get the actual timestamps for frames
+
+import os, sys, re
+#from PyRTF import *
+
+#target = sys.argv[1]
+#targetfiles = os.listdir(target + "/raw")
+# print target
+# print targetfiles
+
+#v = "1614"
+
+checkfile = open("checkfile.csv", "a")
+checkfile.write("subid,last.frame,length\n")
+
+#for f in filter(lambda a: a.find(".txt") != -1, targetfiles):
+# print f
+for filename in os.listdir("raw"):
+
+   fin = open("raw/" + filename, "r")
+   short_f = filename.split("_")[1] + ".csv"
+   fout = open(short_f, "w")
+   fout.write("frame,pkt_time,best_time\n")
+
+   contents = fin.read().split("[FRAME]")
+   m = None
+   i = 1
+   for c in contents:
+       if c.find("media_type=video") != -1:
+           m = re.search(r"pkt_pts_time=\d+.\d+", c) or None
+	   n = re.search(r"best_effort_timestamp_time=\d+.\d+", c) or None
+           fout.write(str(i) + "," + m.group(0).replace("pkt_pts_time=", "") + "," + n.group(0).replace("best_effort_timestamp_time=","") + "\n")
+           i = i + 1
+
+   checkfile.write("XS-" + short_f + "," + str(i - 1) + "," + m.group(0).replace("pkt_pts_time=", "") + "\n")
+
+

### 2018-03-13 data/frame_times_redone/scripts/wrap_show_packets.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+for vid in $PI_HOME/xs-face/headcam_videos_original/*; do
+	sbatch -p normal,hns -t 35:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+        --wrap="ffprobe -show_frames -select_streams v:0 $vid > $(basename $vid).frames.txt"
+	sleep 1
+done

### 2018-03-13 scripts/extractframes/check_data_completeness.py

@@ -1,0 +1,46 @@
+#!/usr/bin/python
+
+## checks data completeness for all codings
+
+import os, sys, re, csv, glob
+from PyRTF import *
+
+subs = []
+
+with open("../data/demographics/demographics.csv", 'rU') as f:
+    reader = csv.reader(f)
+    next(reader, None)  # skip the headers
+    for row in reader:
+        subs.append(row[0])
+
+# print subs
+print "sub\tdets\tposture\tnaming\ttranscript"
+
+for sub in subs:
+    msg = sub + "\t"
+
+    # check detectors
+    if len(glob.glob("../data/detectors/*" + sub[3:10] + "*")) != 0:
+        msg += "x"
+        # print "x"
+
+    msg += "\t"
+               
+    # check posture
+    if len(glob.glob("../data/posture/raw/*" + sub + "*")) != 0:
+        msg += "x"
+
+    msg += "\t"
+    
+    # check namings
+    if len(glob.glob("../data/naming/raw/*" + sub + "*")) != 0:
+        msg += "x"
+
+    msg += "\t"
+    
+    # check transcript
+    if len(glob.glob("../data/naming/raw/*" + sub + "*")) != 0:
+        msg += "x"
+
+    print msg
+

### 2018-03-13 scripts/extractframes/consolidate_detectors.py

@@ -1,0 +1,29 @@
+#!/usr/bin/env python
+# run this to consolidate face detectors
+
+import glob
+
+target = "../data/detectors/*.txt"
+destination = "../data/all_detectors.csv"
+
+outfile = open(destination,"w")
+outfile.write("subid,frame,face,x,y,w,h\n")
+
+for fname in glob.glob(target):
+	print fname
+	infile = open(fname, "r")
+	subid = "XS_" + fname[-8:-4]
+
+	for line in infile.readlines():
+		[f, x, y, w, h, conf] = line.replace("NaN","NA").split(" ")
+
+
+		if x == "NA":
+			face = "FALSE"
+		else:
+			face = "TRUE"
+
+		outfile.write("{},{},{},{},{},{},{}\n".format(subid, f, face, x, y, w, h))
+
+
+	infile.close()

### 2018-03-13 scripts/extractframes/extract_frametimes.py

@@ -1,0 +1,36 @@
+#!/usr/bin/env python
+
+# run this locally to take FFMPEG packet files and get the actual timestamps for frames
+
+import os, sys, re
+from PyRTF import *
+
+target = sys.argv[1]
+targetfiles = os.listdir(target + "/raw")
+#print target
+#print targetfiles
+
+checkfile = open("../data/movie_lengths.csv","a")
+checkfile.write("subid,last.frame,length\n")
+
+for f in filter (lambda a: a.find(".txt") != -1, targetfiles):
+    # print f
+    fin = open(target + "/raw/" + f,"r")
+    short_f = f[3:7] + ".csv"
+    fout = open(target + "/" + short_f,"w")
+    fout.write("frame,time\n")
+    
+    contents = fin.read().split("[PACKET]")
+
+    i = 1
+    for c in contents:
+        if c.find("codec_type=video") != -1:
+            m = re.search(r"pts_time=\d+.\d+",c)
+            if m != None:
+                fout.write(str(i) + "," + m.group(0).replace("pts_time=","") + "\n")
+                i = i + 1
+
+    checkfile.write("XS-" + short_f + "," + str(i-1) + "," + m.group(0).replace("pts_time=","") + "\n")
+        
+    
+    

### 2018-03-13 scripts/extractframes/get_frametimes.py

@@ -1,0 +1,24 @@
+#!/usr/bin/env python                                                          
+
+# run this on the server for all movies to get parsed FFMPEG output files
+
+import sys, os, subprocess
+
+target = sys.argv[1]
+destination = sys.argv[2]
+
+print("target: " + target + "\n")
+print("destination: " + destination + "\n")
+
+targetfiles = os.listdir(target)
+
+for file in targetfiles:
+    if file.find(".mp4") > 0: # if this is a movie                             \
+                                                                                
+	dirname = file.replace(".mp4","")
+	print "***********" + dirname + "**********"
+
+	f = open(destination + "/" + file + ".frames.txt","w")
+	subprocess.call(["ffprobe",target + "/" + file,"-show_packets"],stdout=\
+\
+f)

### 2018-03-13 writeup-cogsci/cogsci2018_march6.Rmd
reorganize writeup

--- 2018-02-01 writeup-cogsci/cogsci2018_faceOnly_alessandro.Rmd
@@
 8 mo. &   12 & 0.46 & 8.71 & 14.41 \\ 
-12 mo. &  12 & 0.40 & 12.62 & 13.48 \\ 
-16 mo. &  12 & 0.31 & 16.29 & 15.00\\ 
+12 mo. &  12 & 0.40 & 12.62 & 12.71 \\ 
+16 mo. &  12 & 0.31 & 16.29 & 15.10\\ 
 \hline
@@
 ### READ IN DETECTIONS
-d <- read_feather("../data/consolidated_data_4detectors.feather") %>%
+d <- read_csv("../data/consolidateddata/consolidated_data_4detectors_march7.csv") %>%
   mutate(posture = factor(posture), orientation = factor(orientation))
+
+# initial submission data - in case you want to reproduce.
+# d <- read_feather("../data/consolidated_data_4detectors.feather") %>%
+#   mutate(posture = factor(posture), orientation = factor(orientation))
 
@@
 levels(dlong$detectorType) 
-#[1] "faceMT"  "faceOP"  "handOP"  "wristOP"
+# [1] "faceMT"  "faceOP"  "faceVJ"  "wristOP"
 levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces","OpenPose-Wrists")
@@
 ## output
-cleanedNames_NoInt = c("Intercept","Close","Far","Prone","Sit","Stand","Age")
+cleanedNames_NoInt = c("Intercept","Close","Far","Sit","Stand","Age") # march 6th - prone is dropped from GLMs
+# cleanedNames_NoInt = c("Intercept","Close","Far","Prone","Sit","Stand","Age") # from initial submission
+
 faceGLMOut=(summary(faceGLM))
@@
   \textbf{Faces} \\
-Intercept & -5.78 & 0.33 & -17.55 & 0.000 \\ 
-  Close & 2.09 & 0.31 & 6.70 & 0.000 \\ 
-  Far & 2.58 & 0.34 & 7.54 & 0.000 \\ 
-  Prone & -0.18 & 0.08 & -2.39 & 0.017 \\ 
-  Sit & 1.00 & 0.07 & 13.54 & 0.000 \\ 
-  Stand & 1.50 & 0.08 & 19.98 & 0.000 \\ 
-  Age & 0.17 & 0.17 & 0.97 & 0.334 \\ 
+  Intercept & -5.96 & 0.29 & -20.77 & 0.000 \\ 
+  Close & 2.15 & 0.28 & 7.54 & 0.000 \\ 
+  Far & 2.63 & 0.31 & 8.42 & 0.000 \\ 
+  Sit & 1.14 & 0.02 & 60.67 & 0.000 \\ 
+  Stand & 1.53 & 0.02 & 71.34 & 0.000 \\ 
+  Age & 0.19 & 0.17 & 1.11 & 0.268 \\ 
   \hline
   \textbf{Hands} \\
-  Intercept & -4.84 & 0.19 & -25.40 & 0.000 \\ 
-  Close & 0.88 & 0.17 & 5.25 & 0.000 \\ 
-  Far & 1.77 & 0.27 & 6.45 & 0.000 \\ 
-  Prone & 0.58 & 0.10 & 5.73 & 0.000 \\ 
-  Sit & 1.18 & 0.10 & 12.04 & 0.000 \\ 
-  Stand & 1.01 & 0.10 & 10.22 & 0.000 \\ 
-  Age & 0.14 & 0.10 & 1.40 & 0.161 \\ 
+  Intercept & -4.29 & 0.16 & -26.10 & 0.000 \\ 
+  Close & 0.87 & 0.17 & 5.28 & 0.000 \\ 
+  Far & 1.78 & 0.27 & 6.60 & 0.000 \\ 
+  Sit & 0.63 & 0.02 & 30.97 & 0.000 \\ 
+  Stand & 0.44 & 0.02 & 19.32 & 0.000 \\ 
+  Age & 0.15 & 0.10 & 1.50 & 0.134 \\ 
    \hline
 \end{tabular}
-\caption{Model coefficients from generalized linear models predicting the proportion of faces (upper panel) and wrists (lower panel) seen by children.} 
+\caption{Model coefficients from generalized linear models predicting the proportion of faces (upper panel) and hands (lower panel) seen by children.} 
 \vspace{-1em}

### 2018-03-13 writeup-cogsci/cogsci2018_march6.tex
reorganize writeup

--- 2018-02-01 writeup-cogsci/cogsci2018_faceOnly_alessandro.tex
@@
   \textbf{Faces} \\
-Intercept & -5.78 & 0.33 & -17.55 & 0.000 \\ 
-  Close & 2.09 & 0.31 & 6.70 & 0.000 \\ 
-  Far & 2.58 & 0.34 & 7.54 & 0.000 \\ 
-  Prone & -0.18 & 0.08 & -2.39 & 0.017 \\ 
-  Sit & 1.00 & 0.07 & 13.54 & 0.000 \\ 
-  Stand & 1.50 & 0.08 & 19.98 & 0.000 \\ 
-  Age & 0.17 & 0.17 & 0.97 & 0.334 \\ 
+  Intercept & -5.96 & 0.29 & -20.77 & 0.000 \\ 
+  Close & 2.15 & 0.28 & 7.54 & 0.000 \\ 
+  Far & 2.63 & 0.31 & 8.42 & 0.000 \\ 
+  Sit & 1.14 & 0.02 & 60.67 & 0.000 \\ 
+  Stand & 1.53 & 0.02 & 71.34 & 0.000 \\ 
+  Age & 0.19 & 0.17 & 1.11 & 0.268 \\ 
   \hline
   \textbf{Hands} \\
-  Intercept & -4.84 & 0.19 & -25.40 & 0.000 \\ 
-  Close & 0.88 & 0.17 & 5.25 & 0.000 \\ 
-  Far & 1.77 & 0.27 & 6.45 & 0.000 \\ 
-  Prone & 0.58 & 0.10 & 5.73 & 0.000 \\ 
-  Sit & 1.18 & 0.10 & 12.04 & 0.000 \\ 
-  Stand & 1.01 & 0.10 & 10.22 & 0.000 \\ 
-  Age & 0.14 & 0.10 & 1.40 & 0.161 \\ 
+  Intercept & -4.29 & 0.16 & -26.10 & 0.000 \\ 
+  Close & 0.87 & 0.17 & 5.28 & 0.000 \\ 
+  Far & 1.78 & 0.27 & 6.60 & 0.000 \\ 
+  Sit & 0.63 & 0.02 & 30.97 & 0.000 \\ 
+  Stand & 0.44 & 0.02 & 19.32 & 0.000 \\ 
+  Age & 0.15 & 0.10 & 1.50 & 0.134 \\ 
    \hline
@@
 baseline (avg. difference in proportion of hands, 8 m.o. = 0, 12 m.o. =
--0.003, 16 m.o. = -0.014; avg. difference in proportion of faces, 8 m.o.
-= 0.003, 12 m.o. = 0.01, 16 m.o. = 0.021).
+-0.004, 16 m.o. = -0.014; avg. difference in proportion of faces, 8 m.o.
+= 0.003, 12 m.o. = 0.01, 16 m.o. = 0.018).
 

### 2018-03-13 writeup-cogsci/helper.R

@@ -1,23 +1,28 @@
 library(stringr)
 
 ######## MERGE IN EXACT TIMES FOR FRAMES #########
+
 add.times <- function(x) {
   print(x$subid[1])
-  fname <- paste0("../data/frame_times/",
-                 str_sub(as.character(x$subid[1]),start=4,end=8),
-                 ".csv")
+  ## from initial submission
+  # fname <- paste0("../data/frame_times/",
+  #                str_sub(as.character(x$subid[1]),start=4,end=8),
+  #                ".csv")
+  
+  ## redone march 6th
+  fname <- paste0("../data/frame_times_redone/",
+                  str_sub(as.character(x$subid[1])),
+                  "_objs.mov.frames.txt.csv")
+  
   if (file.exists(fname)) {
     times <- read_csv(fname)
-    
-#     x <- merge(x,times, by.x = "frame", by.y = "frame", all.x = TRUE, all.y = FALSE)
     x <- left_join(x, times)
-    x$dt <- c(diff(x$time),.032) ## assumes data is sorted by frames! also not sure why .032?
+    x$dt <- c(diff(x$time),.032) 
   } else {
     print(paste0("**** ", x$subid[1], " is missing frames ****"))
     x$time <- NA
     x$dt <- NA
   }
-  
   return(x)
 }
 
@@ -109,13 +114,14 @@ regularize.naming <- function (n) {
   
   # make the times seconds since onset - # if excel output is HMS vs MS - ugh! this was causing a big bug in analyses prior to jan 2018
   timeBefore=n$time
-  if (nchar(timeBefore[1])>8) { ## hacky but works, sorry! will try to prettify soon.
+  if (nchar(timeBefore[1])>8) { ## hacky but works.
     n$time=parse_date_time(timeBefore,"%H:%M:OS%")
   }
   else {
     n$time=parse_date_time(timeBefore,"%M:OS%")
   }
   assert_that(sum(hour(n$time))==0)  # check there are no hours!all sessions <20 minutes.
+  browser()
   n$time <- minute(n$time)*60+second(n$time) # convert to seconds
 
   n <- subset(n,!is.na(name))
@@ -147,7 +153,7 @@ summarize.naming <- function (x, window = c(-2,2)) {
                       sep=""),
                 stringsAsFactors=FALSE)
   
-  # print(x$subid[1])
+  print(x$subid[1])
   
   # rectify the coding
   namings <- regularize.naming(namings)

### 2018-05-03 writeup-cogsci/cogsci2018_march6.Rmd
cogsci revisions minor

@@
   summarize(avgBaselineDiff = mean(baselineDiff))
+
 ## Wrists

### 2018-05-03 writeup-cogsci/cogsci2018_march6.Rmd
add franchak ref

@@
 This perspective changes over the first two years of life, as visual input transitions from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
-It is uncertain how much these changes are driven by the concurrent motoric changes described above versus by other developmental factors.
+While it is uncertain how much these changes are driven by the concurrent motoric changes described above versus by other developmental factors, recent work has found that whether an infant is crawling, sitting, or standing predicts how often they gaze at faces and bodies [@franchak2017see].
 
-Here, we directly examine whether postural and locomotor developments change the availability of social information -- the presence of faces and hands.
+Here, we directly examine whether postural and locomotor developments change the avaliability of social information -- the presence of faces and hands.
 To do so, we recorded the visual experience of a group of infants in three age ranges (8, 12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
@@
 Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors influenced the proportion of faces and hands that were available in the child's visual field.
+Though we oriented our cameras to maximally capture face information, we nonetheless saw an influence or orientation and posture on visual access to hands.
 This work suggests that motor development modulates how infants experience their visual world and the social information in it: Infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.

### 2018-05-10 writeup-cogsci/cogsci2018_march6.Rmd
wording changes

--- 2018-05-03 writeup-cogsci/cogsci2018_march6.Rmd
@@
 crawling) infants make different kinds of object-related bids for attention from their mothers and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].
+Infants also increase interactive behaviors in general and explore their surroundings more [@thurman2017spatial].
 Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
@@
 This perspective changes over the first two years of life, as visual input transitions from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
-It is uncertain how much these changes are driven by the concurrent motoric changes described above versus by other developmental factors.
+
+While it is uncertain how much these changes are driven by the concurrent motoric changes described above versus by other developmental factors, recent work has found that whether an infant is crawling, sitting, or standing predicts how often they gaze at faces and bodies [@franchak2017see].
 
@@
 To do so, we recorded the visual experience of a group of infants in three age ranges (8, 12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
+We then capitalized on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
+Only one piece of work to date has directly examined the impact of postural development to social information access in infants: a study by @franchak2017see found support for the motoric hypothesis presented above using a similar free-play experimental paradigm with head-mounted cameras on infants.
+This study required manual coding of the first 5 minutes of 15 minute videos from a cohort of 12-month olds; here we replicate and extend their findings using novel computer vision techniques to allow for video analysis at scale.
+We also analyze infants from a larger developmental age range (8, 12, and 16 month-olds).
+
 We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus providing children with greater access to the social information in their environment.
@@
 Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors influenced the proportion of faces and hands that were available in the child's visual field.
+Though we oriented our cameras to maximally capture face information, we nonetheless saw an influence of orientation and posture on visual access to hands.
 This work suggests that motor development modulates how infants experience their visual world and the social information in it: Infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.

### 2018-05-11 analysis-main/helper.R
reorg main analysis file

@@ -1,194 +1,0 @@
-library(stringr)
-library(assertthat)
-
-######## MERGE IN EXACT TIMES FOR FRAMES #########
-add.times <- function(x) {
-  print(x$subid[1])
-
-  fname <- paste0("../data/frame_times_redone/",
-                 str_sub(as.character(x$subid[1]),start=4,end=8),
-                 ".csv")
-  
-  
-  if (file.exists(fname)) {
-    times <- read_csv(fname)
-    times$time <- times$best_time ## best estimate of timestamp
-    assert_that(sum(is.na(times$time))==0)
-    
-    print(paste0("**** ", x$subid[1], 'det frame count', max(x$frame), 'time frame count', max(times$frame), "****"))
-    ## 
-    x<-left_join(x, times)
-    x$dt <- c(diff(x$time),.032)
-    
-  } else {
-    print(paste0("**** ", x$subid[1], " is missing frames ****"))
-    x$time <- NA
-    x$dt <- NA
-  }
-  return(x)
-}
-
-######## ADD POSTURE CODING #########
-add.posture <- function(x) {
-  fname <- paste("../data/posture/",
-                 x$subid[1],
-                 ".csv",
-                 sep="")
-  
-  if (file.exists(fname)) {
-    print(fname)
-    postures <- read_csv(fname)
-    
-    postures <- regularize.postures(subset(postures,code=="posture"))
-    
-    x$posture <- factor(NA,levels=levels(postures$posture))
-    x$orientation <- factor(NA,levels=levels(postures$orientation))
-    
-    # for each row of p, populate posture and orientation to x
-    for (i in 1:nrow(postures)) {
-      range <- x$time > postures$start[i] & x$time <= postures$end[i]
-      x$posture[range] <- postures$posture[i]
-      x$orientation[range] <- postures$orientation[i]
-    }
-  } else {
-    print(paste0("**** ", x$subid[1], " is missing posture ****"))
-    x$posture <- NA
-    x$orientation <- NA
-  }
-  
-  return(x)
-}
-
-######## GET POSTURE CODING SORTED OUT #########
-regularize.postures <- function (p) {
-  original.postures <- c("p","si","st","l","c","w","NIF")
-  
-  # clean up postures
-  p$type <- factor(p$type,levels=original.postures)
-  p$type <- plyr::revalue(p$type,c("p"="prone",
-                   "si"="sit",
[…]

### 2018-05-11 analysis-main/merge_all_data.R

@@ -1,109 +1,0 @@
-
-
-## main data merge script - prerequisite for all other analyses
-rm(list=ls())
-library(dplyr)
-library(readr)
-library(magrittr)
-library(tidyr)
-library(assertthat)
-source("helper.R")
-
-## load detectors for mtcnn
-dets <- read_csv("../data/final_output/mtcnn3.csv") %>%
-  mutate(subid = video) %>%
-  mutate(frame = as.numeric(frame)) %>%
-  mutate(faceMT = as.logical(is_face)) %>%
-  distinct(video,frame,.keep_all=TRUE)   
-
-# open pose detections - faces and wrists
-detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") 
-detsOpenPose <- detsOpenPose %>%
-  mutate(video = name) %>%
-  distinct(video,frame,.keep_all=TRUE)  %>%
-  mutate(frame = as.numeric(frame) + 1) %>%
-  mutate(faceOP = Nose_conf!=0)  %>%
-  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
-
-# # open pose detectors - hands -- not using because poor accuracy
-# detsOpenPose_Hands <- read_csv("../data/final_output/openpose_results_with_detail_hands.csv") 
-# detsOpenPose_HandConf <- detsOpenPose_Hands %>%
-#   mutate(video = name) %>%
-#   distinct(video,frame,.keep_all=TRUE)  %>%
-#   mutate(frame = as.numeric(frame) + 1) %>%
-#   select(video,frame,ends_with("conf")) %>%
-#   mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
-#   mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
-#   mutate(handOP = confRight>0 | confLeft>0 )
-
-# viola jones face detectors
-detsViola <- read_csv("../data/final_output/viola.csv") 
-detsViola <- detsViola %>%
-  distinct(video,frame,.keep_all=TRUE)  %>%
-  mutate(faceVJ = as.logical(is_face))  %>%
-  mutate(frame = as.numeric(frame)) 
-
-## check sizes
-assert_that(dim(dets)[1]==dim(detsViola)[1])
-# assert_that(dim(dets)[1]==dim(detsOpenPose)[1]) ## slightly different.
-
-# merge all three detectors
-alldets=left_join(dets,detsViola[,c("video","frame","faceVJ")]) 
[…]

### 2018-05-11 preprocessing/show_packets_old/scripts/show_packets.py
delete old files

@@ -1,39 +1,0 @@
-#!/usr/bin/env python
-
-# run this locally to take FFMPEG packet files and get the actual timestamps for frames
-
-import os, sys, re
-#from PyRTF import *
-
-#target = sys.argv[1]
-#targetfiles = os.listdir(target + "/raw")
-# print target
-# print targetfiles
-
-#v = "1614"
-
-checkfile = open("checkfile.csv", "a")
-checkfile.write("subid,last.frame,length\n")
-
-#for f in filter(lambda a: a.find(".txt") != -1, targetfiles):
-# print f
-for filename in os.listdir("get_packet_info"):
-
-   fin = open("get_packet_info/" + filename, "r")
-   short_f = filename + ".csv"
-   fout = open("junk/" + short_f, "w")
-   fout.write("frame,time\n")
-
-   contents = fin.read().split("[PACKET]")
-
-   i = 1
-   for c in contents:
-       if c.find("codec_type=video") != -1:
-           m = re.search(r"pts_time=\d+.\d+", c)
-           if m != None:
-               fout.write(str(i) + "," + m.group(0).replace("pts_time=", "") + "\n")
-               i = i + 1
-
-   checkfile.write("XS-" + short_f + "," + str(i - 1) + "," + m.group(0).replace("pts_time=", "") + "\n")
-
-

### 2018-05-11 preprocessing/show_packets_old/scripts/wrap_show_packets.sh

@@ -1,7 +1,0 @@
-#!/bin/bash
-
-for vid in $PI_HOME/xs-face/headcam_videos_original/*; do
-	sbatch -p normal,hns -t 35:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
-        --wrap="ffprobe -show_packets $vid > $(basename $vid).frames.txt"
-	sleep 1
-done

### 2018-05-11 writeup-cogsci/helper.R

@@ -121,7 +121,6 @@ regularize.naming <- function (n) {
     n$time=parse_date_time(timeBefore,"%M:OS%")
   }
   assert_that(sum(hour(n$time))==0)  # check there are no hours!all sessions <20 minutes.
-  browser()
   n$time <- minute(n$time)*60+second(n$time) # convert to seconds
 
   n <- subset(n,!is.na(name))

### 2018-05-11 writeup-cogsci/merge_all_data.R

@@ -1,0 +1,109 @@
+
+
+## main data merge script - prerequisite for all other analyses
+rm(list=ls())
+library(dplyr)
+library(readr)
+library(magrittr)
+library(tidyr)
+library(assertthat)
+source("helper.R")
+
+## load detectors for mtcnn
+dets <- read_csv("../data/final_output/mtcnn3.csv") %>%
+  mutate(subid = video) %>%
+  mutate(frame = as.numeric(frame)) %>%
+  mutate(faceMT = as.logical(is_face)) %>%
+  distinct(video,frame,.keep_all=TRUE)   
+
+# open pose detections - faces and wrists
+detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") 
+detsOpenPose <- detsOpenPose %>%
+  mutate(video = name) %>%
+  distinct(video,frame,.keep_all=TRUE)  %>%
+  mutate(frame = as.numeric(frame) + 1) %>%
+  mutate(faceOP = Nose_conf!=0)  %>%
+  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
+
+# # open pose detectors - hands -- not using because poor accuracy
+# detsOpenPose_Hands <- read_csv("../data/final_output/openpose_results_with_detail_hands.csv") 
+# detsOpenPose_HandConf <- detsOpenPose_Hands %>%
+#   mutate(video = name) %>%
+#   distinct(video,frame,.keep_all=TRUE)  %>%
+#   mutate(frame = as.numeric(frame) + 1) %>%
+#   select(video,frame,ends_with("conf")) %>%
+#   mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
+#   mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
+#   mutate(handOP = confRight>0 | confLeft>0 )
+
+# viola jones face detectors
+detsViola <- read_csv("../data/final_output/viola.csv") 
+detsViola <- detsViola %>%
+  distinct(video,frame,.keep_all=TRUE)  %>%
+  mutate(faceVJ = as.logical(is_face))  %>%
+  mutate(frame = as.numeric(frame)) 
+
+## check sizes
+assert_that(dim(dets)[1]==dim(detsViola)[1])
+# assert_that(dim(dets)[1]==dim(detsOpenPose)[1]) ## slightly different -- will get to this below.
+
+# merge all three detectors
[…]

### 2018-05-11 writeup-cogsci/cogsci2018_cameraReady.Rmd
cogsci revisions

--- 2018-05-03 writeup-cogsci/cogsci2018_march6.Rmd
@@
 Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently.
-These motor changes have significant consequences for what children see; crawling and walking infants simply have different views of the world.
-For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011].
-In general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
+These motor changes have significant consequences for what children see.
+For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
 
@@
 For example, postural changes impact how children interact with their mothers: walking (vs.
-crawling) infants make different kinds of object-related bids for attention from their mothers and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].
-Further, in an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
+crawling) infants make different kinds of object-related bids for attention from their mothers, tend to hear more action directed statements (e.g., “open it”), and tend to explore their surroundings more [@karasik2014; @thurman2017spatial].
+In an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
 Thus, children’s ability to independently locomote may change their ability to access social information (e.g., faces, gaze) and in turn accelerate their learning.
@@
 This perspective changes over the first two years of life, as visual input transitions from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
-It is uncertain how much these changes are driven by the concurrent motoric changes described above versus by other developmental factors.
 
 Here, we directly examine whether postural and locomotor developments change the availability of social information -- the presence of faces and hands.
-To do so, we recorded the visual experience of a group of infants in three age ranges (8, 12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
-We then capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
-We hypothesized that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus providing children with greater access to the social information in their environment.
+We recorded the visual experience of a group of infants in three age ranges (8, 12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
+Recent work directly examined how postural developments change access to social information in 12 month-olds in a similar paradigm, finding support for the motoric hypothesis [@franchak2017see].
+Here we replicate and extend their main finding using novel computer vision techniques to allow for video analysis at scale.
+In particular, we capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
+We predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus providing children with greater access to the social information in their environment.
 
@@
 
-# initial submission data - in case you want to reproduce.
-# d <- read_feather("../data/consolidated_data_4detectors.feather") %>%
-#   mutate(posture = factor(posture), orientation = factor(orientation))
-
 ## make long form version for plotting both face detections on same graph
@@
   else {
-    if (a == b) return("TN") # was not face/wrist, detected face/wrist
-    else return("FP") # was not face/wrist, said face/wrist
+    if (a == b) return("TN") # was not face/wrist, did not detect face/wrist
+    else return("FP") # was not face/wrist, detected face/wrist
   }
@@
   rowwise() %>%
-  mutate(result = evaluate(is_face, detector_is_face))
+  mutate(result = evaluate(is_face, detector_is_face)) # always have the gronud truth first in evaluate argument
 
@@
   filter(detectorType=="OpenPose-Wrists") %>%
-  mutate(detector_is_wrist = detection, is_wrist = as.logical(is_face)) %>%
+  mutate(detector_is_wrist = detection, is_wrist = as.logical(is_wrist)) %>% 
   rowwise() %>%
@@
 #wrangle data
+## Using MTCNN detections for GLM since it was the best
 dGLM_faces <- d %>%
@@
   summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1]) %>%
-  mutate(propFaces = countFaces / (countFaces + countNotFaces))
+  ungroup()
 
@@
   summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
-  mutate(propWrists = countWrists / (countWrists + countNotWrists))
+  ungroup()
 
 ## run models 
-faceGLM <- glmer(cbind(countFaces,countNotFaces) ~ orientation + posture + scale(age.at.test) + 
+faceGLM <- glmer(cbind(countFaces,countNotFaces) ~ posture + orientation + scale(age.at.test) + 
                (orientation | subid), 
@@
 ## output
-cleanedNames_NoInt = c("Intercept","Close","Far","Sit","Stand","Age") # march 6th - prone is dropped from GLMs
-# cleanedNames_NoInt = c("Intercept","Close","Far","Prone","Sit","Stand","Age") # from initial submission
+cleanedNames_NoInt = c("Intercept","Close","Far","Sit","Stand","Age")
 
@@
   left_join(inc_subs) %>%
-  mutate(len = as.numeric(len.min))
+  mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video
 
@@
   group_by(subid) %>%
-  mutate(detections = faceMT) %>%
+  mutate(detections = faceMT) %>% ## for covergence with helper.R file, which looks for this variable name
   do(summarize.naming(.)) %>%
@@
   group_by(subid, age.grp) %>%
-  summarize(baselineDetections = mean(faceMT))
+  summarize(baselineDetections = mean(wristOP))
 #

### 2018-05-11 writeup-cogsci/cogsci2018_cameraReady.tex
cogsci revisions

--- 2018-03-13 writeup-cogsci/cogsci2018_march6.tex
@@
 end of the first year of life, allowing them to locomote independently.
-These motor changes have significant consequences for what children see;
-crawling and walking infants simply have different views of the world.
+These motor changes have significant consequences for what children see.
 For example, during spontaneous play in a laboratory playroom, toddlers
 are more likely to look at the floor while crawling than while walking
-(Franchak, Kretch, Soska, \& Adolph, 2011). In general, walking infants
+(Franchak, Kretch, Soska, \& Adolph, 2011); in general, walking infants
 tend to have full visual access to their environment and the people in
@@
 with their mothers: walking (vs.~crawling) infants make different kinds
-of object-related bids for attention from their mothers and tend to hear
-more action directed statements (e.g., ``open it'') (Karasik,
-Tamis-LeMonda, \& Adolph, 2014). Further, in an observational study,
-Walle \& Campos (2014) found that children who were able to walk had
-both higher receptive and productive vocabularies. Thus, children's
-ability to independently locomote may change their ability to access
-social information (e.g., faces, gaze) and in turn accelerate their
-learning.
+of object-related bids for attention from their mothers, tend to hear
+more action directed statements (e.g., ``open it''), and tend to explore
+their surroundings more (Karasik, Tamis-LeMonda, \& Adolph, 2014;
+Thurman \& Corbetta, 2017). In an observational study, Walle \& Campos
+(2014) found that children who were able to walk had both higher
+receptive and productive vocabularies. Thus, children's ability to
+independently locomote may change their ability to access social
+information (e.g., faces, gaze) and in turn accelerate their learning.
 
@@
 capturing views of hands paired with the objects they are acting on
-(Fausey, Jayaraman, \& Smith, 2016). It is uncertain how much these
-changes are driven by the concurrent motoric changes described above
-versus by other developmental factors.
+(Fausey, Jayaraman, \& Smith, 2016).
 
@@
 change the availability of social information -- the presence of faces
-and hands. To do so, we recorded the visual experience of a group of
-infants in three age ranges (8, 12, and 16 months) using head-mounted
-cameras during a brief laboratory free-play session; children's posture
-and orientation relative to their caregiver were also recorded from a
-third-person perspective and hand-annotated. We then capitalize on
+and hands. We recorded the visual experience of a group of infants in
+three age ranges (8, 12, and 16 months) using head-mounted cameras
+during a brief laboratory free-play session; children's posture and
+orientation relative to their caregiver were also recorded from a
+third-person perspective and hand-annotated. Recent work directly
+examined how postural developments change access to social information
+in 12 month-olds in a similar paradigm, finding support for the motoric
+hypothesis (Franchak, Kretch, \& Adolph, 2017). Here we replicate and
+extend their main finding using novel computer vision techniques to
+allow for video analysis at scale. In particular, we capitalize on
 recent improvements in face and pose detection algorithms (Cao, Simon,
@@
 in the child's visual environment, both overall and relative to naming
-events by their caregivers. We hypothesized that there would be
+events by their caregivers. We predicted that there would be
 differential access to social information based on children's postural
@@
 synchronized with the tripod-mounted videos using FinalCut Pro Software.
-These sessions yielded 516 minutes (almost a million frames) of video,
-with an average video length of 14.33 minutes (min = 4.53, max = 19.35).
+These sessions yielded 507 minutes (almost a million frames) of video,
+with an average video length of 14.07 minutes (min = 4.53, max = 19.35).
 
@@
 particularly more faces or hands during naming events relative to
-baseline (avg. difference in proportion of hands, 8 m.o. = 0, 12 m.o. =
--0.004, 16 m.o. = -0.014; avg. difference in proportion of faces, 8 m.o.
-= 0.003, 12 m.o. = 0.01, 16 m.o. = 0.018).
+baseline (avg. difference in proportion of hands, 8 m.o. = 0.012, 12
+m.o. = 0.012, 16 m.o. = 0.013; avg. difference in proportion of faces, 8
+m.o. = 0.003, 12 m.o. = 0.01, 16 m.o. = 0.018).
 
@@
 the proportion of faces and hands that were available in the child's
-visual field. This work suggests that motor development modulates how
-infants experience their visual world and the social information in it:
-Infants that are sitting and standing have a different view of their
-world, the people in it, and the actions that are being performed.
+visual field. Though we oriented our cameras to maximally capture face
+information, we nonetheless saw an influence of orientation and posture
+on visual access to hands. This work suggests that motor development
+modulates how infants experience their visual world and the social
+information in it: Infants that are sitting and standing have a
+different view of their world, the people in it, and the actions that
+are being performed.
 
@@
 
+\hypertarget{ref-franchak2017see}{}
+Franchak, J. M., Kretch, K. S., \& Adolph, K. E. (2017). See and be
+seen: Infant--caregiver social looking during locomotor free play.
+\emph{Developmental Science}.
+
 \hypertarget{ref-franchak2011}{}
@@
 
+\hypertarget{ref-thurman2017spatial}{}
+Thurman, S. L., \& Corbetta, D. (2017). Spatial exploration and changes
+in infant--mother dyads around transitions in infant locomotion.
+\emph{Developmental Psychology}, \emph{53}(7), 1207.
+
 \hypertarget{ref-viola2004robust}{}

### 2018-05-11 writeup-cogsci/cogsci2018_cameraReady.Rmd
still a few lines over but much closer

@@
 # Introduction
-From early infancy, children are deeply engaged in learning from others [@csibra2009natural; @meltzoff2007like].
+From early infancy, children are deeply engaged in learning from others.
 Even newborns tend to prefer to look at faces with direct vs.
@@
 
-```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=6, fig.height=3, fig.cap = "Proportion of face and wrist detections in each age group as a function of children's posture (top panel) and orientation (bottom panel)."}
+```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=6, fig.height=3, fig.cap = "Proportion of face / wrist detections by children's age, posture (top panel), and orientation (bottom panel)."}
 
@@
 We used a head-mounted camera to explore how children's postural and locomotor development affects their visual access to social information, here operationalized as the presence of the faces and hands of their caregiver.
-Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors influenced the proportion of faces and hands that were available in the child's visual field.
-Though we oriented our cameras to maximally capture face information, we nonetheless saw an influence of orientation and posture on visual access to hands.
-This work suggests that motor development modulates how infants experience their visual world and the social information in it: Infants that are sitting and standing have a different view of their world, the people in it, and the actions that are being performed.
+Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors influenced the proportion of faces and hands in the child's visual field.
+Though our cameras were oriented to capture face information, we nonetheless saw an influence of orientation and posture on visual access to hands.
+Motor development appears to modulate how infants experience their visual world and the social information in it.
 

### 2018-05-11 writeup-cogsci/cogsci2018_cameraReady.tex
still a few lines over but much closer

@@
 
-From early infancy, children are deeply engaged in learning from others
-(Csibra \& Gergely, 2009; Meltzoff, 2007). Even newborns tend to prefer
-to look at faces with direct vs.~averted gaze (Farroni, Csibra, Simion,
-\& Johnson, 2002) and young infants follow overt gaze shifts
-(Gredeb\wrapmf{\"{a}}ck, Fikke, \& Melinder, 2010). Further, when
-infants attend to video stimuli, they tend to look mostly at faces at
-the expense of other visual information -- though older infants start to
-look towards people's hands and the actions they are performing (Frank,
-Amso, \& Johnson, 2014; Frank, Vul, \& Saxe, 2012).
+From early infancy, children are deeply engaged in learning from others.
+Even newborns tend to prefer to look at faces with direct vs.~averted
+gaze (Farroni, Csibra, Simion, \& Johnson, 2002) and young infants
+follow overt gaze shifts (Gredeb\wrapmf{\"{a}}ck, Fikke, \& Melinder,
+2010). Further, when infants attend to video stimuli, they tend to look
+mostly at faces at the expense of other visual information -- though
+older infants start to look towards people's hands and the actions they
+are performing (Frank, Amso, \& Johnson, 2014; Frank, Vul, \& Saxe,
+2012).
 
@@
 
-\caption[Proportion of face and wrist detections in each age group as a function of children's posture (top panel) and orientation (bottom panel)]{Proportion of face and wrist detections in each age group as a function of children's posture (top panel) and orientation (bottom panel).}\label{fig:detByPosOrient}
+\caption[Proportion of face / wrist detections by children's age, posture (top panel), and orientation (bottom panel)]{Proportion of face / wrist detections by children's age, posture (top panel), and orientation (bottom panel).}\label{fig:detByPosOrient}
 \end{figure*}
@@
 changed systematically across age, and both of these factors influenced
-the proportion of faces and hands that were available in the child's
-visual field. Though we oriented our cameras to maximally capture face
-information, we nonetheless saw an influence of orientation and posture
-on visual access to hands. This work suggests that motor development
-modulates how infants experience their visual world and the social
-information in it: Infants that are sitting and standing have a
-different view of their world, the people in it, and the actions that
-are being performed.
+the proportion of faces and hands in the child's visual field. Though
+our cameras were oriented to capture face information, we nonetheless
+saw an influence of orientation and posture on visual access to hands.
+Motor development appears to modulate how infants experience their
+visual world and the social information in it.
 
@@
 
-\hypertarget{ref-csibra2009natural}{}
-Csibra, G., \& Gergely, G. (2009). Natural pedagogy. \emph{Trends in
-Cognitive Sciences}, \emph{13}(4), 148--153.
-
 \hypertarget{ref-cummings1988}{}
@@
 
-\hypertarget{ref-meltzoff2007like}{}
-Meltzoff, A. N. (2007). ``Like me'': A foundation for social cognition.
-\emph{Developmental Science}, \emph{10}(1), 126--134.
-
 \hypertarget{ref-scheirer2014perceptual}{}

### 2018-05-11 writeup-cogsci/cogsci2018_cameraReady.Rmd
cut a few last words, references, and rearranged

@@
 averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010development].
-Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring; @frank2014visual].
+Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring].
 
@@
 
-We created a situation in the lab in which the context of interaction was tightly controlled, but as children grow and change, the activities in which they engage with their caregivers also vary, leading to differences in the distribution of contexts they experience.
+We created a situation in which the context of interaction was tightly controlled, but as children grow and change, the activities in which they engage with their caregivers also vary, leading to differences in the distribution of contexts they experience.
 Thus, more work is needed to understand how our results relate to children's home experiences [@fausey2016].

### 2018-05-11 writeup-cogsci/cogsci2018_cameraReady.tex
cut a few last words, references, and rearranged

@@
 older infants start to look towards people's hands and the actions they
-are performing (Frank, Amso, \& Johnson, 2014; Frank, Vul, \& Saxe,
-2012).
+are performing (Frank, Vul, \& Saxe, 2012).
 
@@
 
-We created a situation in the lab in which the context of interaction
-was tightly controlled, but as children grow and change, the activities
-in which they engage with their caregivers also vary, leading to
-differences in the distribution of contexts they experience. Thus, more
-work is needed to understand how our results relate to children's home
-experiences (Fausey et al., 2016). The ability to walk is part of a
-cascade of changes in children's experience, and our study captures only
-a cross-sectional slice of this broader, multifaceted trajectory.
+We created a situation in which the context of interaction was tightly
+controlled, but as children grow and change, the activities in which
+they engage with their caregivers also vary, leading to differences in
+the distribution of contexts they experience. Thus, more work is needed
+to understand how our results relate to children's home experiences
+(Fausey et al., 2016). The ability to walk is part of a cascade of
+changes in children's experience, and our study captures only a
+cross-sectional slice of this broader, multifaceted trajectory.
 
@@
 
-\hypertarget{ref-frank2014visual}{}
-Frank, M. C., Amso, D., \& Johnson, S. P. (2014). Visual search and
-attention to faces during early infancy. \emph{Journal of Experimental
-Child Psychology}, \emph{118}, 13--26.
-
 \hypertarget{ref-frank2013}{}

### 2018-07-17 writeup-cogsci/cogsci2018_remakePlots.Rmd
script to remake a few plots

--- 2018-05-11 writeup-cogsci/cogsci2018_cameraReady.Rmd
@@
 knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, fig.pos = "tb", fig.path='figs/',
-                      echo=F, warning=F, cache=F, message=F, sanitize = T)
+                      echo=F, warning=F, cache=F, message=F, sanitize = T, dev = "cairo_pdf")
 ```
@@
 ggarrange(p1,p2)
+fig1=ggarrange(p1,p2)
+ggsave("Fig3-TimebyPos.png", width = 5, height = 5,unit =  "in", plot = fig1, path="./figs/", device = "png",dpi = 300)
 ```
@@
 
-ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
+fig4=ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
   geom_jitter(width = .25) +
@@
   facet_grid(~detectorType)
+
+ggsave("Fig4-DetbyAge.png", width = 5, height = 4,unit =  "in", plot = fig4, path="./figs/", device = "png",dpi = 300)
 ```
@@
 ggarrange(p1, p2)
+fig5=ggarrange(p1, p2)
+ggsave("Fig5-DetbyOrientation.png", width = 6, height = 3,unit =  "in", plot = fig5, path="./figs/", device = "png",dpi = 300)
 
@@
 
-ggarrange(p1,p2, p3, p4)
+fig7=ggarrange(p1,p2, p3, p4)
+
+ggsave("Fig7-Naming.png", width = 3, height = 3,unit =  "in", plot = fig7, path="./figs/", device = "png",dpi = 300)
 ```

### 2018-10-12 writeup-cogsci/helper.R
fix typo in reading raw filenames

@@ -5,14 +5,13 @@ library(stringr)
 add.times <- function(x) {
   print(x$subid[1])
   ## from initial submission
-  # fname <- paste0("../data/frame_times/",
-  #                str_sub(as.character(x$subid[1]),start=4,end=8),
-  #                ".csv")
-  
-  ## redone march 6th
   fname <- paste0("../data/frame_times_redone/",
-                  str_sub(as.character(x$subid[1])),
-                  "_objs.mov.frames.txt.csv")
+                 str_sub(as.character(x$subid[1]),start=4,end=8),
+                 ".csv")
+  
+  # ## redone march 6th / typo fixed oct 12 2018
+  # fname <- paste0("../data/frame_times_redone/",
+  #                 str_sub(as.character(x$subid[1])),".csv")
   
   if (file.exists(fname)) {
     times <- read_csv(fname)

### 2018-10-12 writeup-cogsci/cogsci2018_remakePlots.Rmd
r file that remakes plots for poster

@@
 
-ggarrange(p1,p2)
-fig1=ggarrange(p1,p2)
-ggsave("Fig3-TimebyPos.png", width = 5, height = 5,unit =  "in", plot = fig1, path="./figs/", device = "png",dpi = 300)
+ggarrange(p1,p2, nrow=1)
+fig1=ggarrange(p1,p2, ncol=1)
+ggsave("Fig3-TimebyPos.png", width = 4, height = 4,unit =  "in", plot = fig1, path="./figs/", device = "png",dpi = 300)
 ```
@@
   scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
+  scale_colour_grey(start=.2, end=.6) +
   labs(y = "Prop. detected", x = element_blank()) +
@@
   theme(legend.position="none") +
-  facet_grid(~detectorType)
+  facet_wrap(~detectorType, ncol=1)
 
-ggsave("Fig4-DetbyAge.png", width = 5, height = 4,unit =  "in", plot = fig4, path="./figs/", device = "png",dpi = 300)
+ggsave("Fig4-DetbyAge.png", width = 3, height = 4,unit =  "in", plot = fig4, path="./figs/", device = "png",dpi = 300)
 ```

### 2018-11-06 writeup-paper/helper.R
create new writeup folder for paper

@@ -1,0 +1,184 @@
+library(stringr)
+
+######## MERGE IN EXACT TIMES FOR FRAMES #########
+
+add.times <- function(x) {
+  print(x$subid[1])
+  ## from initial submission
+  fname <- paste0("../data/frame_times_redone/",
+                 str_sub(as.character(x$subid[1]),start=4,end=8),
+                 ".csv")
+  
+  if (file.exists(fname)) {
+    times <- read_csv(fname)
+    times$time <- times$best_time ## comes from show_frames python script
+    x <- left_join(x, times)
+    x$dt <- c(diff(x$time),.032) 
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing frames ****"))
+    x$time <- NA
+    x$dt <- NA
+  }
+  return(x)
+}
+
+######## ADD POSTURE CODING #########
+add.posture <- function(x) {
+  fname <- paste("../data/posture/",
+                 x$subid[1],
+                 ".csv",
+                 sep="")
+  
+  if (file.exists(fname)) {
+    print(fname)
+    postures <- read_csv(fname)
+    postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order
+    postures <- regularize.postures(subset(postures,code=="posture"))
+    x$posture <- factor(NA,levels=levels(postures$posture))
+    x$orientation <- factor(NA,levels=levels(postures$orientation))
+    
+    # for each row of p, populate posture and orientation to x
+    for (i in 1:nrow(postures)) {
+      range <- x$time > postures$start[i] & x$time <= postures$end[i]
+      x$posture[range] <- postures$posture[i]
+      x$orientation[range] <- postures$orientation[i]
+    }
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing posture ****"))
+    x$posture <- NA
+    x$orientation <- NA
+  }
+  
+  return(x)
+}
+
+######## GET POSTURE CODING SORTED OUT #########
+regularize.postures <- function (p) {
+  original.postures <- c("p","si","st","l","c","w","NIF")
+  
+  # clean up postures
+  p$type <- factor(p$type,levels=original.postures)
+  p$type <- plyr::revalue(p$type,c("p"="prone",
+                   "si"="sit",
+                   "st"="stand",
+                   "l"="prone",
+                   "c"="carry",
[…]

### 2018-11-06 writeup-paper/images/frames_with_overlayed_coords/mtcnn_correct/draw_rectangles.py

@@ -1,0 +1,32 @@
+import os
+os.system("module load opencv")
+import cv2
+import ntpath
+
+# x, y, w, h
+
+# m = {
+# 	"09551":[120, 82, 279, 265],
+# 	"07106":[217, 300, 338, 444],
+# 	"10506":[441, 296, 692, 480]
+# }
+
+base = "/share/PI/mcfrank/frames_rotated/"
+m2 = {
+	base + "XS_0803/image-09551.jpg" : [120, 82, 279, 265],
+	base + "XS_1222/image-07106.jpg" : [217, 300, 338, 444],
+	base + "XS_1619/image-10506.jpg" : [441, 296, 692, 480]
+}
+
+if __name__ == "__main__":
+	base = "/share/PI/mcfrank/"
+	d = [base + "XS_0803/image-09551.jpg"]
+
+	for k, v in m2.iteritems():
+		image = cv2.imread(k)
+
+		x, y, x2, y2 = m2[k]
+
+		cv2.rectangle(image, (x, y), (x2, y2), (0, 255, 0), 3)
+
+		cv2.imwrite(ntpath.basename(k).split(".")[0] + "_with_box.jpg", image)

### 2018-11-06 writeup-paper/merge_all_data.R

@@ -1,0 +1,109 @@
+
+
+## main data merge script - prerequisite for all other analyses
+rm(list=ls())
+library(dplyr)
+library(readr)
+library(magrittr)
+library(tidyr)
+library(assertthat)
+source("helper.R")
+
+## load detectors for mtcnn
+dets <- read_csv("../data/final_output/mtcnn3.csv") %>%
+  mutate(subid = video) %>%
+  mutate(frame = as.numeric(frame)) %>%
+  mutate(faceMT = as.logical(is_face)) %>%
+  distinct(video,frame,.keep_all=TRUE)   
+
+# open pose detections - faces and wrists
+detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") 
+detsOpenPose <- detsOpenPose %>%
+  mutate(video = name) %>%
+  distinct(video,frame,.keep_all=TRUE)  %>%
+  mutate(frame = as.numeric(frame) + 1) %>%
+  mutate(faceOP = Nose_conf!=0)  %>%
+  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
+
+# # open pose detectors - hands -- not using because poor accuracy
+# detsOpenPose_Hands <- read_csv("../data/final_output/openpose_results_with_detail_hands.csv") 
+# detsOpenPose_HandConf <- detsOpenPose_Hands %>%
+#   mutate(video = name) %>%
+#   distinct(video,frame,.keep_all=TRUE)  %>%
+#   mutate(frame = as.numeric(frame) + 1) %>%
+#   select(video,frame,ends_with("conf")) %>%
+#   mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
+#   mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
+#   mutate(handOP = confRight>0 | confLeft>0 )
+
+# viola jones face detectors
+detsViola <- read_csv("../data/final_output/viola.csv") 
+detsViola <- detsViola %>%
+  distinct(video,frame,.keep_all=TRUE)  %>%
+  mutate(faceVJ = as.logical(is_face))  %>%
+  mutate(frame = as.numeric(frame)) 
+
+## check sizes
+assert_that(dim(dets)[1]==dim(detsViola)[1])
+# assert_that(dim(dets)[1]==dim(detsOpenPose)[1]) ## slightly different -- will get to this below.
+
+# merge all three detectors
[…]

### 2018-11-06 scripts/ground_truth_creation/download.py
move scripts folder & contents to preprocessing folder; reorg

@@ -1,10 +1,0 @@
-import os
-import csv
-
-with open("../../data/ground_truth/gold_set_candidates_from_openpose.csv", 'rb') as img_paths_f:
-        img_paths = list(csv.reader(img_paths_f, delimiter=','))
-
-for path in img_paths[1:]:
-    os.system("mkdir -p ~/xsface/sample_face_images_openpose/%s" % (path[1]))
-    os.system("scp sanchez7@sherlock.stanford.edu:/share/PI/mcfrank/frames_rotated/%s ~/xsface/sample_face_images_openpose/%s" % (path[8], path[1]))
-

### 2018-11-06 scripts/ground_truth_creation/validate.py

@@ -1,44 +1,0 @@
-import os
-import csv
-import ntpath
-import xml.etree.ElementTree as ET
-
-# TODO take in arg for wrist or face and change the corresponding is_face?
-
-BRIA_ANNOTATIONS = "/Users/alessandro/Downloads/sample_hand_images_bll"
-
-if __name__ == "__main__":
-
-    rows = {}
-
-    for root, subdirectory, filenames in os.walk(BRIA_ANNOTATIONS):
-
-        if root[len(root) - 1].isdigit(): # inside video directory
-            for filename in filenames:
-                if filename == ".DS_Store": continue
-                video = ntpath.basename(root)
-                group = str(int(video.split("_")[1][:2]))
-                frame = filename.split(".")[0].split("-")[1]
-
-                rows[video + filename.split(".")[0]] = [group, video, frame, "False"]
-
-        elif ntpath.basename(root) == "annotations": # inside annotations
-            video = ntpath.basename(os.path.dirname(root))
-            for filename in filenames:
-                xmlroot = ET.parse(os.path.join(root, filename)).getroot()
-                if xmlroot.findall("object"):
-                    arr = rows[video + filename.split(".")[0]]
-                    arr[3] = "True"
-                    rows[video + filename.split(".")[0]] = arr
-
-    print(len(rows))
-
-    with open("ground_truth_wrists2.csv", 'wb') as csvfile:
-        wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
-        wr.writerow(['group', 'video', 'frame', 'is_face'])
-        for key, val in rows.iteritems():
-            wr.writerow(val)
-
-
-
-

### 2018-11-06 scripts/mtcnn_face_detection/wrap_mtcnn_detect_faces.sh

@@ -1,16 +1,0 @@
-#!/bin/sh
-
-module load tensorflow
-module load opencv/3.0.0
-
-for vid in $PI_HOME/frames_rotated/*; do
-	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
-	    echo "Large video (> 10GB): $vid"
-		sbatch -p hns,normal -c 16 -t 20:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
-		    --wrap="python detect_faces.py $vid"
-	else
-		sbatch -p hns,normal -c 8 -t 10:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
-		    --wrap="python detect_faces.py $vid"
-	fi
-	sleep 1
-done

### 2018-11-06 scripts/openpose/check_output.py

@@ -1,1 +1,0 @@
-# check # of frames along each step in the pipeline

### 2018-11-06 scripts/openpose/wrap_json_to_csv.sh

@@ -1,17 +1,0 @@
-#!/bin/sh
-
-for vid in $PI_HOME/openpose_output6/*; do
-	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
-	    echo "Large video (> 10GB): $vid"
-	    time=3:00:00
-	    memory=12G
-	else
-	    echo "Small vid: $vid"
-	    time=2:00:00
-	    memory=4G
-	fi
-
-	sbatch -p normal,hns -t $time --mem $memory --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
-	    --wrap="python json_to_csv.py $vid"
-    sleep 1
-done

### 2018-11-06 scripts/openpose/wrap_openpose.sh

@@ -1,37 +1,0 @@
-#!/bin/sh
-
-module load system
-module load singularity/2.4
-module load cuda/8.0.61
-module load caffe2/0.8.1
-module load cudnn/5.1
-
-mkdir -p $PI_HOME/openpose_json_output
-
-for vid in $SCRATCH/globus/*/*; do
-	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
-	    echo "Large video (> 10GB): $vid"
-	    time=10:00:00
-	    memory=64G
-	else
-	    echo "Small vid: $vid"
-	    time=5:00:00
-	    memory=8G
-	fi
-
-    output_dir=$PI_HOME/openpose_output6/$(basename $vid)
-	mkdir -p $output_dir
-
-	sbatch -p gpu --gres gpu:1 -t $time --mem $memory --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
-	    --wrap="singularity exec --nv $SINGULARITY_CACHEDIR/openpose-latest.img bash -c \
-	        'cd /openpose-master && ./build/examples/openpose/openpose.bin \
-	        --no_display true \
-            --render_pose 0 \
-            --video $vid \
-            --frame_rotate 180 \
-	    --face \
-	    --hand \
-            --write_keypoint_json $output_dir'"
-
-	sleep 1
-done

### 2018-11-06 scripts/viola_jones/viola_jones_detector.py

@@ -1,69 +1,0 @@
-import os
-import re
-import sys
-import csv
-import ntpath
-import traceback
-import multiprocessing
-sys.path.append('/home/sanchez7/local')
-import cv2
-
-OUTPUT_FILE = 'tmp/face_detection_%s.csv'
-
-
-def detect_faces(gray_img):
-    cascPath = "haarcascade_frontalface_default.xml"
-    faceCascade = cv2.CascadeClassifier(cascPath)
-    return faceCascade.detectMultiScale(
-        gray_img,
-        scaleFactor=1.1,
-        minNeighbors=5,
-        minSize=(30, 30)
-    )
-
-
-def process_img(imgpath):
-    # print('batch {0}: {1}%'.format(arg, i/1075.0 * 100))
-    print imgpath
-    gray_img = cv2.cvtColor(cv2.imread(imgpath), cv2.COLOR_BGR2GRAY)
-
-    faces = detect_faces(gray_img)
-
-    name = ntpath.basename(os.path.dirname(imgpath))
-    group = str(int(name.split("_")[1][:2]))
-    frame = re.search('image-(.*).jpg', ntpath.basename(imgpath)).group(1)
-
-    rows = []
-
-    if len(faces) == 0:
-        rows.append([group, name, frame, False, None, None, None, None])
-    else:
-        for (x, y, w, h) in faces:
-            rows.append([group, name, frame, True, x, y, w, h])
-
-    return rows
-
-
-if __name__ == "__main__":
-    vid_folder = sys.argv[1]
-    vid = ntpath.basename(vid_folder)
-
-    pool = multiprocessing.Pool()
-    results = []
-
-    for image in os.listdir(vid_folder):
-        results.append(pool.apply_async(process_img, args=(os.path.join(vid_folder, image),)))
-
-    pool.close()
-
-    with open(OUTPUT_FILE % vid, 'wb') as csvfile:
-        wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
-        # wr.writerow(['group', 'video', 'frame', 'is_face', 'x', 'y', 'w', 'h'])
-
-        for result in results:
-            try:
-                rows = result.get()
-                for row in rows:
-                    wr.writerow(row)
-            except:
-                traceback.print_exc()

### 2018-11-06 scripts/viola_jones/wrap_viola_jones_detector.sh

@@ -1,12 +1,0 @@
-#!/bin/sh
-for vid in $PI_HOME/frames_rotated/*; do
-	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
-	    echo "Large video (> 10GB): $vid"
-		sbatch -p hns,normal -c 16 -t 10:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
-		    --wrap="python viola_jones_detector.py $vid"
-	else
-		sbatch -p hns,normal -c 8 -t 5:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
-		    --wrap="python viola_jones_detector.py $vid"
-	fi
-	sleep 1
-done

### 2018-11-06 analysis-validation/z-old/check_frame_counts_post_detection.Rmd
Get rid of duplicate sanity check files / old unused scripts

---
title: "Check frame counts after MTCNN detection"
output: html_notebook
---

```{r}
library(tidyverse)
library(knitr)
```

Load data
```{r}
showframes_frame_counts <- read_csv("../data/video_stats/frametimechecks_showframes.csv")

mtcnn_data <- read_csv("../data/final_detector_output/mtcnn3.csv")
```

Check mtcnn results frame counts against extracted frames
```{r}

showframes_frame_counts <- showframes_frame_counts %>%
  group_by(subid) %>%
  mutate(video = strsplit(subid,'.csv',2)) %>%
  mutate(video = str_replace(video,'-','_'))

check_frame_times <- mtcnn_data %>%
  distinct(video, frame) %>%
  group_by(video) %>%
  summarise(num_frames = n()) %>%
  left_join(showframes_frame_counts) %>%
  mutate(checkOK = num_frames == last.frame)

kable(check_frame_times)
```

### 2018-11-06 preprocessing/detector_scripts/ground_truth_creation/get_gold_set_images.Rmd
reorganize preprocessing folder, part 2

--- 2018-01-11 scripts/ground_truth_creation/get_gold_set_images.Rmd
@@
 ```{r}
-data <- read_csv("data/final_output/mtcnn3.csv") %>%
-  distinct(video, frame, .keep_all = TRUE)
+data <- read_csv("~/Github/xs-face/data/final_output/openpose_results_truncated_2.csv") %>%
+  distinct(name, frame, .keep_all = TRUE)
+```
+
+Filter images from previous ground truth (openpose only)
+```{r}
+old_ground_truth <- read_csv("~/Github/xs-face/data/ground_truth/gold_set_candidates2.csv")
+data %<>%
+  rename(video = name) %>%
+  mutate(frame = sprintf("%05d", as.numeric(frame) + 1)) %>%
+  anti_join(old_ground_truth) %>%
+  mutate(is_face = ifelse(Nose_conf != 0, "True", "False")) %>%
+  select(group, video, frame, Nose_x, Nose_y, Nose_conf, is_face)
 ```
@@
 
-write_csv(gold_set_candidates, "data/ground_truth/gold_set_candidates.csv")
+write_csv(gold_set_candidates, "~/Github/xs-face/data/ground_truth/gold_set_candidates_from_openpose.csv")
 ```

### 2018-11-06 preprocessing/detector_scripts/ground_truth_creation/download.py

@@ -1,0 +1,10 @@
+import os
+import csv
+
+with open("../../data/ground_truth/gold_set_candidates_from_openpose.csv", 'rb') as img_paths_f:
+        img_paths = list(csv.reader(img_paths_f, delimiter=','))
+
+for path in img_paths[1:]:
+    os.system("mkdir -p ~/xsface/sample_face_images_openpose/%s" % (path[1]))
+    os.system("scp sanchez7@sherlock.stanford.edu:/share/PI/mcfrank/frames_rotated/%s ~/xsface/sample_face_images_openpose/%s" % (path[8], path[1]))
+

### 2018-11-06 preprocessing/detector_scripts/ground_truth_creation/validate.py

@@ -1,0 +1,44 @@
+import os
+import csv
+import ntpath
+import xml.etree.ElementTree as ET
+
+# TODO take in arg for wrist or face and change the corresponding is_face?
+
+BRIA_ANNOTATIONS = "/Users/alessandro/Downloads/sample_hand_images_bll"
+
+if __name__ == "__main__":
+
+    rows = {}
+
+    for root, subdirectory, filenames in os.walk(BRIA_ANNOTATIONS):
+
+        if root[len(root) - 1].isdigit(): # inside video directory
+            for filename in filenames:
+                if filename == ".DS_Store": continue
+                video = ntpath.basename(root)
+                group = str(int(video.split("_")[1][:2]))
+                frame = filename.split(".")[0].split("-")[1]
+
+                rows[video + filename.split(".")[0]] = [group, video, frame, "False"]
+
+        elif ntpath.basename(root) == "annotations": # inside annotations
+            video = ntpath.basename(os.path.dirname(root))
+            for filename in filenames:
+                xmlroot = ET.parse(os.path.join(root, filename)).getroot()
+                if xmlroot.findall("object"):
+                    arr = rows[video + filename.split(".")[0]]
+                    arr[3] = "True"
+                    rows[video + filename.split(".")[0]] = arr
+
+    print(len(rows))
+
+    with open("ground_truth_wrists2.csv", 'wb') as csvfile:
+        wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
+        wr.writerow(['group', 'video', 'frame', 'is_face'])
+        for key, val in rows.iteritems():
+            wr.writerow(val)
+
+
+
+

### 2018-11-06 preprocessing/detector_scripts/mtcnn_face_detection/wrap_mtcnn_detect_faces.sh

@@ -1,0 +1,16 @@
+#!/bin/sh
+
+module load tensorflow
+module load opencv/3.0.0
+
+for vid in $PI_HOME/frames_rotated/*; do
+	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
+	    echo "Large video (> 10GB): $vid"
+		sbatch -p hns,normal -c 16 -t 20:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+		    --wrap="python detect_faces.py $vid"
+	else
+		sbatch -p hns,normal -c 8 -t 10:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+		    --wrap="python detect_faces.py $vid"
+	fi
+	sleep 1
+done

### 2018-11-06 preprocessing/detector_scripts/openpose/wrap_json_to_csv.sh

@@ -1,0 +1,17 @@
+#!/bin/sh
+
+for vid in $PI_HOME/openpose_output6/*; do
+	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
+	    echo "Large video (> 10GB): $vid"
+	    time=3:00:00
+	    memory=12G
+	else
+	    echo "Small vid: $vid"
+	    time=2:00:00
+	    memory=4G
+	fi
+
+	sbatch -p normal,hns -t $time --mem $memory --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+	    --wrap="python json_to_csv.py $vid"
+    sleep 1
+done

### 2018-11-06 preprocessing/detector_scripts/openpose/wrap_openpose.sh

@@ -1,0 +1,37 @@
+#!/bin/sh
+
+module load system
+module load singularity/2.4
+module load cuda/8.0.61
+module load caffe2/0.8.1
+module load cudnn/5.1
+
+mkdir -p $PI_HOME/openpose_json_output
+
+for vid in $SCRATCH/globus/*/*; do
+	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
+	    echo "Large video (> 10GB): $vid"
+	    time=10:00:00
+	    memory=64G
+	else
+	    echo "Small vid: $vid"
+	    time=5:00:00
+	    memory=8G
+	fi
+
+    output_dir=$PI_HOME/openpose_output6/$(basename $vid)
+	mkdir -p $output_dir
+
+	sbatch -p gpu --gres gpu:1 -t $time --mem $memory --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+	    --wrap="singularity exec --nv $SINGULARITY_CACHEDIR/openpose-latest.img bash -c \
+	        'cd /openpose-master && ./build/examples/openpose/openpose.bin \
+	        --no_display true \
+            --render_pose 0 \
+            --video $vid \
+            --frame_rotate 180 \
+	    --face \
+	    --hand \
+            --write_keypoint_json $output_dir'"
+
+	sleep 1
+done

### 2018-11-06 preprocessing/detector_scripts/viola_jones/viola_jones_detector.py

@@ -1,0 +1,69 @@
+import os
+import re
+import sys
+import csv
+import ntpath
+import traceback
+import multiprocessing
+sys.path.append('/home/sanchez7/local')
+import cv2
+
+OUTPUT_FILE = 'tmp/face_detection_%s.csv'
+
+
+def detect_faces(gray_img):
+    cascPath = "haarcascade_frontalface_default.xml"
+    faceCascade = cv2.CascadeClassifier(cascPath)
+    return faceCascade.detectMultiScale(
+        gray_img,
+        scaleFactor=1.1,
+        minNeighbors=5,
+        minSize=(30, 30)
+    )
+
+
+def process_img(imgpath):
+    # print('batch {0}: {1}%'.format(arg, i/1075.0 * 100))
+    print imgpath
+    gray_img = cv2.cvtColor(cv2.imread(imgpath), cv2.COLOR_BGR2GRAY)
+
+    faces = detect_faces(gray_img)
+
+    name = ntpath.basename(os.path.dirname(imgpath))
+    group = str(int(name.split("_")[1][:2]))
+    frame = re.search('image-(.*).jpg', ntpath.basename(imgpath)).group(1)
+
+    rows = []
+
+    if len(faces) == 0:
+        rows.append([group, name, frame, False, None, None, None, None])
+    else:
+        for (x, y, w, h) in faces:
+            rows.append([group, name, frame, True, x, y, w, h])
+
+    return rows
+
+
+if __name__ == "__main__":
+    vid_folder = sys.argv[1]
+    vid = ntpath.basename(vid_folder)
+
+    pool = multiprocessing.Pool()
+    results = []
+
+    for image in os.listdir(vid_folder):
+        results.append(pool.apply_async(process_img, args=(os.path.join(vid_folder, image),)))
+
+    pool.close()
+
+    with open(OUTPUT_FILE % vid, 'wb') as csvfile:
+        wr = csv.writer(csvfile, quoting=csv.QUOTE_ALL)
+        # wr.writerow(['group', 'video', 'frame', 'is_face', 'x', 'y', 'w', 'h'])
+
+        for result in results:
+            try:
+                rows = result.get()
+                for row in rows:
+                    wr.writerow(row)
+            except:
+                traceback.print_exc()

### 2018-11-06 preprocessing/detector_scripts/viola_jones/wrap_viola_jones_detector.sh

@@ -1,0 +1,12 @@
+#!/bin/sh
+for vid in $PI_HOME/frames_rotated/*; do
+	if [[ $vid == *1205* || $vid == *1211* || $vid == *1225* || $vid == *1628*  ]]; then
+	    echo "Large video (> 10GB): $vid"
+		sbatch -p hns,normal -c 16 -t 10:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+		    --wrap="python viola_jones_detector.py $vid"
+	else
+		sbatch -p hns,normal -c 8 -t 5:00:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
+		    --wrap="python viola_jones_detector.py $vid"
+	fi
+	sleep 1
+done

### 2018-11-06 preprocessing/show_frames/scripts/show_frames.py

@@ -1,39 +1,0 @@
-#!/usr/bin/env python
-
-# run this locally to take FFMPEG packet files and get the actual timestamps for frames
-
-import os, sys, re
-#from PyRTF import *
-
-#target = sys.argv[1]
-#targetfiles = os.listdir(target + "/raw")
-# print target
-# print targetfiles
-
-#v = "1614"
-
-checkfile = open("checkfile.csv", "a")
-checkfile.write("subid,last.frame,length\n")
-
-#for f in filter(lambda a: a.find(".txt") != -1, targetfiles):
-# print f
-for filename in os.listdir("raw"):
-
-   fin = open("raw/" + filename, "r")
-   short_f = filename.split("_")[1] + ".csv"
-   fout = open(short_f, "w")
-   fout.write("frame,pkt_time,best_time\n")
-
-   contents = fin.read().split("[FRAME]")
-   m = None
-   i = 1
-   for c in contents:
-       if c.find("media_type=video") != -1:
-           m = re.search(r"pkt_pts_time=\d+.\d+", c) or None
-	   n = re.search(r"best_effort_timestamp_time=\d+.\d+", c) or None
-           fout.write(str(i) + "," + m.group(0).replace("pkt_pts_time=", "") + "," + n.group(0).replace("best_effort_timestamp_time=","") + "\n")
-           i = i + 1
-
-   checkfile.write("XS-" + short_f + "," + str(i - 1) + "," + m.group(0).replace("pkt_pts_time=", "") + "\n")
-
-

### 2018-11-06 preprocessing/show_frames/scripts/wrap_show_frames.sh

@@ -1,7 +1,0 @@
-#!/bin/bash
-
-for vid in $PI_HOME/xs-face/headcam_videos_original/*; do
-	sbatch -p normal,hns -t 35:00 --mail-type=FAIL --mail-user=sanchez7@stanford.edu \
-        --wrap="ffprobe -show_frames -select_streams v:0 $vid > $(basename $vid).frames.txt"
-	sleep 1
-done

### 2018-11-06 writeup-paper/draft_1_from_CogSci.Rmd
rename draft

--- 2018-11-06 writeup-paper/cogsci2018_cameraReady.Rmd
@@
 library(assertthat)
-
 ```

### 2018-11-06 writeup-paper/helper.R
comment out excessive printing

@@ -1,7 +1,6 @@
 library(stringr)
 
 ######## MERGE IN EXACT TIMES FOR FRAMES #########
-
 add.times <- function(x) {
   print(x$subid[1])
   ## from initial submission
@@ -82,7 +81,7 @@ regularize.postures <- function (p) {
   
   # re-zero the times
   begin <- p$start[1]
-  print(paste0("**** zero-time", begin, "*** min time:", min(p$start)))
+  # print(paste0("**** zero-time", begin, "*** min time:", min(p$start)))
   ## make sure that this value is the smallest that we're reading
   assert_that(begin==min(p$start))
 

### 2018-11-06 writeup-paper/step1_merge_data.Rmd
Replace merge_all_data.r with annotated r-markdown file

---
title: 'XS-Face: Merge Raw Data'
author: "Bria Long, continued from MCF document merge_all_data.r"
date: "11/6/2018"
output: html_document
---
# Step 0: Libraries and helper functions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) # remove all objects from current workspace
library(dplyr)
library(readr)
library(magrittr)
library(tidyr)
library(assertthat)
source("helper.R") # Contains critical helper functions for merging annotations
```

## Step 1: Load detectors
### Load detectors from MTCNN
```{r}
dets_MTCNN <- read_csv("../data/final_detector_output/mtcnn3.csv") %>%
  mutate(subid = video) %>%
  mutate(frame = as.numeric(frame)) %>%
  mutate(faceMT = as.logical(is_face)) %>%
  distinct(video,frame,.keep_all=TRUE)   
```

### Load Viola-Jones detectors
```{r}
detsViola <- read_csv("../data/final_detector_output/viola.csv") 
detsViola <- detsViola %>%
  distinct(video,frame,.keep_all=TRUE)  %>%
  mutate(faceVJ = as.logical(is_face))  %>%
  mutate(frame = as.numeric(frame)) 
```

### Load OpenPose detections (face and wrists)
Wrists: if right or left wrist was ever detected
Faces: If nose was ever detected

```{r}
detsOpenPose <- read_csv("../data/final_detector_output/openpose_results_truncated_2.csv") 
detsOpenPose <- detsOpenPose %>%
  mutate(video = name) %>%
  distinct(video,frame,.keep_all=TRUE)  %>%
  mutate(frame = as.numeric(frame) + 1) %>% ## Frame output is off by one since we extracted from videos vs. frames.
  mutate(faceOP = Nose_conf!=0)  %>%  
  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
```

## Step 2:  Merge detectors and sanity check
```{r}
assert_that(dim(dets_MTCNN)[1]==dim(detsViola)[1]) 
alldets=left_join(dets_MTCNN,detsViola[,c("video","frame","faceVJ")])  ## Join MTCNN + Viola Jones
alldets=left_join(alldets,detsOpenPose[,c("video","frame","faceOP","wristOP")])  ##  ++ OpenPose
```

### Sanity checks for merging detectors
We know that openPose works on videos and so might notget all frames. Let's check out the frames it doesn't have

```{r}
# First -- MTCNN v Viola Joes
# These were done on the same frames, so should work

## 
missingInd=is.na(alldets$faceOP)
assert_that(sum(is.na(alldets$faceOP))==45)

maxFrameNum <- alldets %>%
  group_by(video) %>%
  summarize(maxFrame = max(frame))

## Let's look at which frames these are
missingOP<- alldets %>% 
  filter(missingInd) %>%
  group_by(video) %>%
  left_join(maxFrameNum) %>%
  summarize(numPerSub = n(), isFaceMTCNN = sum(faceMT), isFaceJF = sum(faceVJ), missingFrameStart = min(frame), maxFrameNum = maxFrame[1])

```

Open pose seems to be skipping between 2-13 frames for 7 subjects for a total of 45 frames. There's no way to gaurentee that these actually come from the end of the videos, 45 frames is less than 2 seconds of video overall so we're OK with that margin of error

#### None of the missing frames are faces. Replace OpenPose "NAN" detections with "FALSE"
```{r}
alldets$faceOP[missingInd]=FALSE 
alldets$wristOP[missingInd]=FALSE

# detectors are reasonably colinear
MTvsOP = mean(alldets$faceOP==alldets$faceMT)
VJvsOP = mean(alldets$faceOP==alldets$faceVJ)
```

#### Rearrange alldets we make sure it is in the order of the frames
```{r}
alldets<-arrange(alldets,video,frame) 
```

## Step 3. Merge demographics and manual annotations
```{r}
# load demographics and pose / orientation / timing
demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
  select(-ra, -assist, -len) ## -dot -dib fields did not exist, deleted bll
```

### Merge in frame times, postures, and orientations
```{r}
# calls helper functions in helper.r to get times and posture coding integrated
d <- alldets %>%
  select(-c(is_face,video)) %>% # redudant with faceMT & subid, dropping
  left_join(demo.data) %>% # join demographics info 
  group_by(subid) %>% # get subid
  do(add.times(.)) %>%  # calls helper functions to add times
  group_by(subid) %>%
  do(add.posture(.)) ## adds both postures and orientatinos
```

#### Check that we synced everything right -- no duplicate timestamps
```{r}
test <- d %>%
  mutate(isDuplicate = (dt==0)) %>%
  group_by(subid) %>%
  summarize(duplicateCount = sum(isDuplicate))

assert_that(sum(is.na(d$dt))==0)
assert_that(sum(test$duplicateCount)==0)
assert_that(sum(is.na(d$time))==0)
```

### Complete the data frame 
So that zeros get counted: expands so that each subid includes a zero length row for each posture and orientation.
This was annoying.
```{r}
ages <- d %>%
  group_by(subid) %>%
  summarise(age.grp = mean(age.grp))

complete_combos <- tidyr::expand(d, nesting(posture, orientation), subid) %>% 
  mutate(dt = 0, faceMT = FALSE, faceVJ = FALSE, faceOP = FALSE, handOP = FALSE, wristOP = FALSE) %>%
  left_join(ages)

d <- bind_rows(d, complete_combos) ## bind these 251 data points
```

# Step 4: Save out the data
```{r}
# write_csv(d, "../data/consolidated_data/consolidated_data_4detectors_test.csv") #
```

### 2018-11-06 writeup-paper/merge_all_data.R

@@ -1,109 +1,0 @@
-
-
-## main data merge script - prerequisite for all other analyses
-rm(list=ls())
-library(dplyr)
-library(readr)
-library(magrittr)
-library(tidyr)
-library(assertthat)
-source("helper.R")
-
-## load detectors for mtcnn
-dets <- read_csv("../data/final_output/mtcnn3.csv") %>%
-  mutate(subid = video) %>%
-  mutate(frame = as.numeric(frame)) %>%
-  mutate(faceMT = as.logical(is_face)) %>%
-  distinct(video,frame,.keep_all=TRUE)   
-
-# open pose detections - faces and wrists
-detsOpenPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv") 
-detsOpenPose <- detsOpenPose %>%
-  mutate(video = name) %>%
-  distinct(video,frame,.keep_all=TRUE)  %>%
-  mutate(frame = as.numeric(frame) + 1) %>%
-  mutate(faceOP = Nose_conf!=0)  %>%
-  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
-
-# # open pose detectors - hands -- not using because poor accuracy
-# detsOpenPose_Hands <- read_csv("../data/final_output/openpose_results_with_detail_hands.csv") 
-# detsOpenPose_HandConf <- detsOpenPose_Hands %>%
-#   mutate(video = name) %>%
-#   distinct(video,frame,.keep_all=TRUE)  %>%
-#   mutate(frame = as.numeric(frame) + 1) %>%
-#   select(video,frame,ends_with("conf")) %>%
-#   mutate(confLeft=rowSums(.[grep("hand_left", names(.))])/21)  %>%
-#   mutate(confRight=rowSums(.[grep("hand_right", names(.))])/21)  %>%
-#   mutate(handOP = confRight>0 | confLeft>0 )
-
-# viola jones face detectors
-detsViola <- read_csv("../data/final_output/viola.csv") 
-detsViola <- detsViola %>%
-  distinct(video,frame,.keep_all=TRUE)  %>%
-  mutate(faceVJ = as.logical(is_face))  %>%
-  mutate(frame = as.numeric(frame)) 
-
-## check sizes
-assert_that(dim(dets)[1]==dim(detsViola)[1])
-# assert_that(dim(dets)[1]==dim(detsOpenPose)[1]) ## slightly different -- will get to this below.
-
-# merge all three detectors
[…]

### 2020-04-10 analysis-franchak/franchak_reanalysos.Rmd
add franchak metadata and analysis code

---
title: "Franchak_ReAnalysis"
author: "Bria Long"
date: "3/5/2020"
output: html_document
---

# Preprocessing
## Libraries
```{r}
rm(list=ls()) # remove all objects from current workspace
library(ggthemes)
library(langcog)
library(tidyverse)
library(cowplot)

```

## Read in merging filenames data
```{r}
filenames <- read.csv('../data/franchak_dataset/meta/subID_matching.csv') %>%
  mutate(vid_name = str_split_fixed(VideoName,'.mp4',2)[,1]) %>%
  mutate(sub_id = SubID) %>%
  select(vid_name, sub_id)
```

## Read in position data and preprocess
```{r}
files = dir('../data/franchak_dataset/posture/')
all_pos_data=tibble()

for (s in files) {
  sub_id_short_temp = str_split_fixed(s, '-',2)[,2]
  sub_id_short = str_split_fixed(sub_id_short_temp, '_pos.csv',2)[,1]

this_sub_position_data <- read.csv(paste0('../data/franchak_dataset/posture/', s)) %>%
  as.tibble %>%
  mutate(frame = row_number(), sub_id = sub_id_short) %>%
  mutate(frame = frame - 1)  %>% ## openpose is 0 indexed
  rename('infant_posture' = 'NaN.', 'caregiver_posture' = 'NaN..1') %>%
  filter(!is.na(infant_posture))

all_pos_data <- bind_rows(all_pos_data,this_sub_position_data)
}

all_pos_data <- all_pos_data %>%
  mutate(sub_id = as.factor(sub_id))
```

## Read in detections and merge with subid from above
```{r}
detections <- read.csv('../data/franchak_dataset/openpose_detections/franchak_frames_openpose_master_frames.csv') 
dets <- detections %>%
  left_join(filenames) %>%
  mutate(sub_id = as.factor(sub_id))
```

# Merge detections with position data (should merge on subid, frame)
```{r}
pos_detections  <- all_pos_data %>%
  left_join(dets) 
```

### Recode variables so are redable
```{R}
level_key = c('0' = "Prone", '1' = "Sitting", '2' = "Upright")
pos_detections$caregiver_posture_label = recode(pos_detections$caregiver_posture, !!!level_key) 
pos_detections$infant_posture_label = recode(pos_detections$infant_posture, !!!level_key)

```

### Save out preprocessed data
```{r}
save(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections.RData', pos_detections)
```

# Basic analyses
## Compute faces by infant posture
```{r}
faces_by_pos  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_faces = mean(face_openpose)) %>%
  group_by(infant_posture_label) %>%
  multi_boot_standard(col = 'prop_faces')

faces_by_pos_by_sub  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_faces = mean(face_openpose))
```

## Plot detections by posture
```{r}
(faces_by_pos_plot <- ggplot(faces_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
   geom_point(data=faces_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label), alpha=.2) +
  theme_few(base_size=18) +
  ylab('Prop faces detected') +
  xlab('Infant posture') + 
  ylim(c(0,.3)) +
  theme(legend.position = 'none') +
  scale_color_manual(name = "Posture", values=c("#339966","#eb7928","#0099cc")))

```

## Faces by caregiver posture 
```{r}
faces_by_CG_pos  <- pos_detections %>%
  group_by(sub_id, caregiver_posture_label) %>%
  summarize(prop_faces = mean(face_openpose)) %>%
  group_by(caregiver_posture_label) %>%
  multi_boot_standard(col = 'prop_faces')

(ggplot(faces_by_CG_pos, aes(x=caregiver_posture_label, y=mean, col=caregiver_posture_label)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  theme_few())
```

## Look at hands by infant posture as well
```{r}
hands_by_pos  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_hands = mean(hand_openpose)) %>%
  group_by(infant_posture_label) %>%
  multi_boot_standard(col = 'prop_hands')

hands_by_pos_by_sub  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_hands = mean(hand_openpose))

```

### Plot hands by infant posture
```{r}
(hands_by_pos_plot <- ggplot(hands_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  geom_point(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label), alpha=.2) +
  theme_few(base_size=18) +
  ylab('Prop hands detected') +
   xlab('Infant posture') + 
  ylim(c(0,.4)) +
  theme(legend.position = 'none') +
  scale_color_manual(name = "Posture", values=c("#339966","#eb7928","#0099cc")))
)

```

```{r}
cowplot::plot_grid(faces_by_pos_plot,hands_by_pos_plot)
```

### 2020-04-10 writeup-paper/draft_apa_march2020.Rmd
add writeup folder for apa draft

---
title             : "Postural developments modulate children's visual access to social information"
shorttitle        : "Posture modulates social access"

author: 
  - name          : "Bria L. Long"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Serra Mall, Stanford CA 94305"
    email         : "bria@stanford.edu"
  - name          : "Alessandro Sanchez"
    affiliation   : "1"
  - name          : "Allison M. Kraus"
    affiliation   : "1"
  - name          : "Ketan Agrawal"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Stanford University"

# authornote: |
#   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
# 
#   Enter author note here.

abstract: |
  The ability to process social information is a critical component of children's early language and cognitive development.
  However, as children reach their first birthday, they begin to locomote themselves, dramatically affecting their visual access to this information. 
  How do these postural and locomotor changes affect children's access to the social information relevant for word-learning? 
  Here, we explore this question by using head-mounted cameras to record 36 infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to estimate the proportion of faces and hands in infants' environments. 
  We find that infants' posture and orientation to their caregiver modulates their access to social information, confirming previous work that suggests motoric developments play a significant role in the emergence of children's linguistic and social capacities. 
  In addition, we applied our automated analysis to the headcam data from a recent paper showing similar effects postre on visual attention (Franchak et al., 2017), finding convergence across methodologies, laboratories, and datasets.
  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience as they change over development.

keywords          : "Postural developments modulate children's visual access to social information"
wordcount         : "X"

bibliography      : ["xsface.bib"]

floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : yes

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(lme4)
library(gridExtra)
library(ggridges)
library(assertthat)
library(tidyverse)
library(viridis)
library(papaja)
```

From their earliest months, infants are deeply engaged in learning from others.  Even newborns tend to prefer to look at faces with direct vs. averted gaze [@farroni2002eye] and young infants follow overt gaze shifts. And as infants learn to parse and segment their native language(s), the social cues provided by speakers (e.g., eye-gaze) may provide strong scaffolding for early word learning. Indeed, empirical work suggests that children’s ability to process social cues is a key factor in early language development: in a longitudinal study, children’s level of joint engagement with their mother at 9-12 months was found to predict both their receptive and productive vocabularies [@carpenter1998]. More specifically, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months and throughout the second year of life [@brooks2005; @brooks2008].

Around their first birthday, however, children’s ability to interact with their world radically changes [@adolph2006motor] as they are no longer constrained to the same spot that their caregivers last placed them in. Before the first birthday, children often begin crawling; soon after, they begin to walk independently. While not all children crawl [@adolph1998]—or crawl in the same way—children tend to find creative ways of moving (e.g., scooting) or getting others to help them move (e.g cruising) to begin to explore their world on their own [@patrick2012]. And as independent agents in the world, children are more active in the construction of their own learning environments. Yet while toddlers do determine their own input to a much greater degree, they spend much of their time in a world primarily populated by knees

One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities [@iverson2010]. Indeed, these postural changes change how children interact with their mothers; walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”) [@karasik2014]. In an observational study, @walle2014 also found that children who were able to walk had both higher receptive and productive vocabularies. On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access the social information (e.g., faces, gaze) relative to children who are still crawling and sitting. In other words, the ability of walking infants to access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.

Over the past decade, researchers have started to use egocentric, head-mounted cameras to document the social information that infants and children have access to across early development, and to understand the degree to which there are substantial shifts in their viewpoints that may have downstream developmental consequences [@Yoshida2008]. Indeed, the infant visual perspective has proved to be both remarkably different than the adult perspective and not easily predicted by our own intuitions [@franchak2011, @yoshida2008; @clerkin2017]. Over the first two years of life, the infant viewpoint transitions from primarily containing close up view of faces to capturing relatively restricted views of hands paired with the objects they are acting on [@fausey2016]; both computational and empirical work suggest that this restricted viewpoint is more effective for learning about objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017].

Here, we directly examine the role that postural developments that infants’ experience as they reach their first birthday change the social information in the infant view. We hypothesize that children’s postural developments may be partially responsible for some of these changes in the infant perspective. During spontaneous play, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011), when they have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014). More recently, when 12-month-olds [@Franchak2018] participated in a free-play session with their caregivers, their own posture and as well as their caregiver’s posture influenced the proportion of time they spent looking at the faces and bodies of their caregivers [@Franchak2017].

Here, we expand on these previous findings in three key ways. First, using head-mounted cameras during a brief laboratory free-play session, we record the visual experience of three groups of children at 8, 12, and 16 months-of-age, covering a broad range of ages and locomotor abilities around the first birthday; Children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated. This cross-sectional design thus allows us to directly examine the relative contributions of age vs. postural developments on children’s visual access to social information. Second, we use a modern computer vision algorithm for the automated detection of faces and hands in the egocentric viewpoints of infants.  In particular, we capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands in the child’s visual environment, both overall and relative to naming events by their caregivers. Use of this emerging technology allows us to annotate the entirety of the dataset, allowing a more complete picture of the changing infant perspective. In addition, we extend this technique to the same dataset used in @Franchak2017, replicating their main pattern of results using this novel methodology and finding convergence between our datasets with different play sessions, populations, and head-mounted cameras. Finally, as parents were presented with pairs of novel and familiar objects with their children during the play sessions, we directly examine the availability of social signals during naming events.

Thus, our current dataset allows us to analyze changes in the visual access to faces and hands according to children's age, posture, and linguistic input. We predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces/hands because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape with greater access to the social information in their environment. [language predictions?]

# Methods
### Participants
```{r echo=FALSE, include=FALSE}
all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
  rename(subid = subID)
inc_subs <- read.csv("../data/demographics/demographics.csv")
all_subs$inc <- all_subs$reason=="included"

inc_subs$len.min <- sapply(strsplit(as.character(inc_subs$len),":"),
                           function(x) {
                             x <- as.numeric(x)
                             x[1]+x[2]/60
                           }
)

all_subs %>% 
  group_by(reason) %>%
  summarise(n = n()) %>% 
  kable

all_subs %>%
  left_join(inc_subs) %>%
  filter(inc==TRUE) %>%
  group_by(age_group) %>%
  summarise(n = sum(inc),
            percent_included = mean(inc),
            age = mean(age.at.test, na.rm=TRUE),
            video_length = mean(len.min, na.rm=TRUE),
            female = sum(gender=="female", na.rm=TRUE)) 
  # xtable %>%
  # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)

# <!-- \caption{\label{tab:pop} Demographics by age group.} -->
```

Our final sample consisted of 36 infants and children, with 12 participants in three age groups: 8 months (6 F), 12 months (7 F), and 16 months (6 F). Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80 percent English at home. Demographics and exclusion rates are  given in the table below.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
\hline
Group & N & \% incl. & Avg age & Avg video length (min) \\
\hline
8 mo. &   12 & 0.46 & 8.71 & 14.41 \\
12 mo. &  12 & 0.40 & 12.62 & 12.71 \\
16 mo. &  12 & 0.31 & 16.29 & 15.10\\
\hline
\end{tabular}
\end{table}

To obtain this final sample, we tested `r length(all_subs$reason)` children, excluding `r sum(all_subs$reason!="included")` children for the following reasons:
`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
`r sum(all_subs$reason=="no CDI")` for missing Communicative Development Inventory (CDI) data,
`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
`r sum(all_subs$reason=="infant crying")` for fussiness, and one for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses. Some of these data were also analyzed in @frank2013.

### Procedure
All parents signed consent documents while children were fitted with the headcam. If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam. When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session. Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys. The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.

### Data Processing and Annotation
Headcam videos were trimmed such that they excluded the instruction phase when the experimenter was in the room and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software. These sessions yielded 
`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a million frames) of video, with an average video length of
`r round(mean(inc_subs$len.min), digits = 2)` minutes
(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).

### Posture and Orientation Annotation
We created custom annotations to describe the child's physical posture (e.g. standing) and the orientation of the child relative to the caregiver (e.g. far away). The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing. The caregiver's orientation was characterized as being close, far, or behind the child (independent of distance). For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child. All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012toward]. Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.

## Face and Hand Detection
We used three face detection systems to measure infants' access to faces. The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones. We used this algorithm as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual]. We next tested the performance of two face detectors that both made use of recently developed Convolutional Neural Networks (CNNs) to extract face information. The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts. For the second algorithm [OpenPose; @cao2017realtime], we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.

The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for hands for two reasons. First, as we did not capture children's entire visual field, the presence of a wrist is likely often indicative of the presence of a hand within the field of view. Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.

\begin{figure*}
\centering
\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

### Algorithms
The first face detection system made use of a series of Haar feature-based cascade classifiers [@viola2004robust] applied to each individual frame. The second algorithm (based on work by @zhang2016) uses multi-task cascaded convolutional neural networks (MTCNNs) for joint face detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present. We used a Tensorflow implementation of this algorithm avaliable at https://github.com/davidsandberg/facenet.

The CNN-based pose detector  [OpenPose; @cao2017realtime; @simon2017hand; @wei2016cpm] provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose. The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose; here, we only made use of the body parts relevant to the face and hands (nose and wrists).

### Detector evaluation
To evaluate face detector performance, we hand-labeled a "gold set" of labeled frames. To account for the relatively rare appearance of faces in the dataset, we hand-labeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames. Each sample was comprised of an equal number of frames taken from each child's video. For wrist detections, the "gold set" was constructed in the same manner, except frames with a high density of wrists came only from detections made by OpenPose.  Faces were classified as present if at least half of the face was showing; wrists were classified as present if any part of the wrist was showing. Two authors labelled the frames independently and resolved disagreements on a case-by-case basis. Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 1.

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN DETECTIONS
d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_march2020.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

## make long form version for plotting both face detections on same graph
dlong <- d %>%
  mutate(id = paste(subid,frame)) %>%
  gather(key = detectorType, value = detection, faceMT, faceOP, faceVJ, wristOP) %>%
  mutate(detectorType=as.factor(detectorType))

# for better plotting names
levels(dlong$detectorType) 
# [1] "faceMT"  "faceOP"  "faceVJ"  "wristOP"
levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces","OpenPose-Wrists")

```

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN GROUND TRUTH and load function
ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv") %>%
  mutate(sample_type=as.factor(sample_type))

ground_truth_wrists <- read_csv("../data/ground_truth/ground_truth_wrists.csv") %>%
  mutate(sample_type=as.factor(sample_type))

# Renaming for table output
levels(ground_truth_faces$sample_type)=c("High density", "Random")
levels(ground_truth_wrists$sample_type)=c("High density", "Random")

# Function to evaluate detectors
evaluate <- function(a, b) {
  if (a == TRUE) {
    if (a == b) return ("TP") # was face/wrist, detected face/wrist
    else return("FN") # was face/wrist, missed face/wrist
  }
  else {
    if (a == b) return("TN") # was not face/wrist, did not detect face/wrist
    else return("FP") # was not face/wrist, detected face/wrist
  }
}
```

```{r  include=FALSE, echo=FALSE}
## Evalute face and wrist detectors
performanceFaces <- ground_truth_faces  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong)   %>%
  filter(detectorType!="OpenPose-Wrists") %>%
  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face)) # always have the gronud truth first in evaluate argument

perfFacesbySampleType <- performanceFaces %>%
  group_by(detectorType, sample_type) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN"))  %>%
  group_by(detectorType, sample_type) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
  # xtable %>%
  # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)

perfFacesOverall <- performanceFaces %>%
  group_by(detectorType) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  group_by(detectorType) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))

## EVALUATE WRIST DETECTOR
performanceWrists <- ground_truth_wrists  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong) %>%
  filter(detectorType=="OpenPose-Wrists") %>%
  mutate(detector_is_wrist = detection, is_wrist = as.logical(is_wrist)) %>% 
  rowwise() %>%
  mutate(result = evaluate(is_wrist, detector_is_wrist))

perfWristsbySampleType<- performanceWrists %>%
  group_by(detectorType, sample_type) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  group_by(detectorType, sample_type) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
  xtable %>%
  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)

perfWristsOverall<- performanceWrists %>%
  group_by(detectorType) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))

## get rid of ViolaJones detections from data structure for future analyses
dlong <- dlong %>%
  filter(detectorType!="faceVJ")

```

For face detection, MTCNN outperformed OpenPose when taking into account only the composite F-score (`r round(perfFacesOverall$f[1],2)` MTCNN vs. `r round(perfFacesOverall$f[2],2)` OpenPose). Although MTCNN and OpenPose performed comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN. ViolaJones performed quite poorly relative to the other detectors, especially with respect to the random sample. We thus use MTCNN detections in the following analyses. For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision but low recall on the randomly sampled frames (see Table 1). We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of hands in the dataset.

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
\hline
Algorithm & Sample\ Type & P & R & F \\ 
\hline
MTCNN-Faces & High density & 0.89 & 0.92 & \textbf{0.90} \\ 
MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
OpenPose-Faces & Random & 0.72 & 0.80 & \textbf{0.76} \\ 
ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
OpenPose-Wrists & High density & 0.66 & 1.00 & 0.79 \\ 
OpenPose-Wrists & Random & 0.88 & 0.29 & 0.44 \\ 
\hline
\end{tabular}
\caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected). P, R, and F denote precision, recall, and F-score, respectively. Scores in bold are the highest F-scores for each sample type.} 
\vspace{-1em}
\end{table}

# Results
First, we report developmental shifts in infants' posture and their orientation relative to their caregiver, consistent with previous literature [@adolph2006; @franchak2017]. Then, we examine how these changes influence children's visual access to faces and wrists/hands across this developmental time range, examining the relative contributions of age vs. postural developments. We also explore how these changes impact the accessibility of faces and wrists/hands during labeling events, both when parents were labeling familiar objects (e.g., 'cat') as well as novel objects (e.g., 'zem'). Finally, we apply the same automated detection method to a egocentric video dataset collected by a different lab [@franch2017] during which children's in-the-moment posture was annotated, replicating their existing results and finding converging support across methods, labs, and paradigms for the impact of postural developments on children's visual access to social information.

```{r}
## Calculate distributions of infants spent in each category separately for posture / orientation
ages_inc_subs <- inc_subs %>%
  select(age.at.test, age.grp, subid)

orientation <- d %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, subid, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(subid, age.grp, orientation) %>%
  left_join(ages_inc_subs)

posture <- d %>%
  filter(!is.na(posture)) %>%
  group_by(age.grp, subid, posture) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(subid, age.grp, posture) %>%
  left_join(ages_inc_subs)
```

```{r}
## Make plots of time spent in each orientation / posture
orientation_labels <- c(
  behind ='Caregiver behind',
  close = 'Caregiver close',
  far = 'Caregiver far')

posture_labels = c(
  carry = 'Infant carried',
  prone = 'Infant prone',
  sit = 'Infant sitting',
  stand ='Infant standing'
)

pos_plot <- ggplot(posture, aes(x = age.at.test, y = prop.time, col = posture, size=time)) + 
  geom_point(alpha=.6) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  theme_few(base_size=12) +
  ylim(0,1) +
  facet_grid(~posture, labeller = labeller(posture = posture_labels)) +
  geom_smooth(span=5, alpha=.2) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_color_manual(name = "Posture", values=c("#8878cc","#0099cc","#339966","#eb7928")) +
  theme(legend.position = 'none')

orient_plot <- ggplot(orientation, aes(x = age.at.test, y = prop.time, col = orientation, size=time)) + 
  geom_point(alpha=.8) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  theme_few(base_size=12) +
  ylim(0,1) +
  facet_grid(~orientation,  labeller = labeller(orientation= orientation_labels)) +
  geom_smooth(span=10, alpha=.2) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_color_manual(name = "CG", values=c("#171615","#5c5753","#99928d")) +
  theme(legend.position = 'none')
```

```{r posture, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
cowplot::plot_grid(pos_plot, orient_plot, nrow=2)
```

```{r}
### Calculate individual amounts of time spent in each orientation by subject
posture_by_orientation <- d %>%
  filter(!is.na(posture)) %>%
  # filter(!is.na(orientation)) %>%
  group_by(subid, posture, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  group_by(subid) %>%
  mutate(prop.time = time/sum(time)) %>%
  mutate(pos_orient = paste0('Infant ',posture, ', CG ', orientation))

posture_by_orientation <- posture_by_orientation %>%
  left_join(ages_inc_subs) %>%
  ungroup() %>%
  mutate(subid = fct_reorder(subid, age.at.test))
```

```{r indivPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}

## should be a two column figure
ggplot(posture_by_orientation, aes(x = subid, y = prop.time)) +
  geom_col(aes(fill = pos_orient), position=position_stack()) +
  ylab("Proportion Time") +
  xlab("Individual subjects (ordered by age)") +
  ylim(0,1) +
  theme_few(base_size=18) +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  scale_fill_manual(name = "Posture and CG Orientation", values=c("#8878cc","#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc", "#98a69f", "#e67300","#ffa64d", "#ffe6cc","#d1b673"))
```

## Changes in Posture and Orientation
How do childen's in-the-moment posture and orientation change across age? In our dataset, the proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with infants' age. As children got older, their locomotive abilities allowed them to become more independent. Both 8-month-olds and 12-month-olds spent relatively equivalent amounts of time lying/crawling (i.e., 'prone') which was markedly decreased in the 16-month-olds, who spent most of their time sitting or standing (see Figure \ref{fig:posture}). We also observed changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions than did children at other ages (see Figure \ref{fig:posture}). However, we also saw considerable variability across children: some infants spent almost their entire time sitting at a close distance from their caregiver, whereas others showed more considerable variability (see Figure \ref{fig:indivPosOrient}).

## Changes in Access to Faces and Hands
We first examined the proportion of face and hand detections as a function of children's age without considering their posture (see Figure \ref{fig:detByAge}). While faces tended to be in the field-of-view overall more often than hands, children's head-mounted cameras were angled slightly upward to capture the presence of faces, and hand detections suffered from somewhat lower recall than face detections. Going forward, we thus analyze differences in the relative proportion of faces or hands in view as a function of age, posture, and orientation, rather than comparing them directly. Overall, we observed that 12-month-olds appeared to have visual access to slightly fewer faces than 8 or 16-month-olds, creating a slight U-shaped function in face detections; conversely, hand detections were showed a slight increase across this age range, as reported in prior literature [@Fausey2016].

However, these age-related trends were much smaller than the effect of infant's postural developments on children's visual access to faces and hands. Children's in-the-moment posture was a major factor both in how many faces and hands were in view during the play session, as was their orientation relative to their caregiver. Infants who were sitting or standing had more faces in view than infants who were lying down/crawling (i.e. prone), which was most frequent among 12-month-olds relative to the other age groups  (Figure \ref{fig:detByPosOrient}). When caregivers were behind their children, supporting their children's sitting or standing positions, children saw fewer faces and wrists. In particular, children who were sitting or in front of their caregiver had a high proportion of faces and hands in their field of view. (Figure \ref{fig:detByPosOrient}, lower panel).

These trends were quantified using two generalized linear mixed-effect models estimating the proportion of faces and hands that were in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture. A summary of the coefficients of the models can be found in Table 2, confirming that infants who were sitting/standing and in front of their caregivers saw the most faces and hands. When modeling the proportion of faces seen, age was not a significant predictor; however, age remained a significant predictor when modeling the proportion of hands seen by infants. Overall, these results suggest that infants' visual access to social information is largely modulated by their posture and orientation to their caregiver, which is in turn a function of their general locomotor development. Nonetheless, these results also suggest that infants may still increase their attention to hands throughout this age range as they continue to learn about objects, their functions, and their names.

## Access to Faces and Hands During Labeling Events
Our play session was designed to provide parents with opporunities to label objects--both familiar and novel--such that we could examine whether children seek out different kinds of social information around  naming events. We thus explored how face and hand detections changed during object labeling events, analyzing a four-second window around each labeling event (e.g., "Look at the [zem]!"). Every utterance of one of the labelled objects (e.g., 'ball') was counted as a 'labeling event'; timestamps of the beginning of each word were hand-annotated from transcripts ad synchronized with the frame-by-frame detections. Overall, we did not find major differences in face/hand detections during naming events relative to baseline, either for novel objects or for familiar objects (see Figure \ref{fig:detByNaming}). This was true both when we analyzed naming rates with or without taking into account children's posture and orientation relative to their caregiver. These results suggest that either children did not actively seek out social information during this particular play session around naming events, perhaps because were a limited number of possible referents at any given time.

```{r eval=FALSE, include=FALSE}
#wrangle data
## Using MTCNN detections for GLM since it was the best
dGLM_faces <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

dGLM_wrists <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

## run models 
faceGLM <- glmer(cbind(countFaces,countNotFaces) ~ posture * orientation + scale(age.at.test) + 
               (posture + orientation | subid), 
             dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

wristGLM <- glmer(cbind(countWrists,countNotWrists) ~ orientation * posture + scale(age.at.test) + 
               (posture + orientation | subid), 
             dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

## output ---- no interaction
# cleanedNames_NoInt = c("Intercept","Close","Far","Sit","Stand","Age") 
# 
# faceGLMOut=(summary(faceGLM))
# rownames(faceGLMOut$coefficients)<-cleanedNames_NoInt
# 
# wristGLMOut=(summary(wristGLM))
# rownames(wristGLMOut$coefficients)<-cleanedNames_NoInt
# 
# xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed model predicting the proportion of faces vs. wrists seen by infants.") 
# xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3))

```

```{r}
detectionsByAge <- dlong %>%
  filter(detectorType!="MTCNN-Faces") %>% # for brevity
  filter(detectorType!="ViolaJones-Faces") %>% # low performance
  filter(!is.na(age.at.test)) %>% # filters NAs for combos of orientation/posture that were empty (see merge_data.Rmd)
  group_by(subid, age.at.test, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  left_join(inc_subs) %>%
  mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video
```

```{r detByAge, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=3, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data."}

ggplot(detectionsByAge, aes(x = age.at.test, y = prop_detected, color = detectorType, size=num_detections)) +
  theme_few(base_size=12) + 
  geom_point(alpha=.5) +
  geom_smooth(size=1, alpha=.3, span=10) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
  scale_size_continuous(name = "Number of frames") +
  labs(y = "Proportion detected", x = element_blank()) +
  scale_colour_discrete(name = element_blank()) +
  theme(legend.position="none") +
  facet_grid(~detectorType)

```

```{r}
### Make dataframes for main plot (detections by orientation & posture)
# filter out detectors we aren't plotting
# filter out detections w/o posture/orientation codings (wasn't always in view) -- this elimiantes "carry"
dlongPlot <-dlong %>%
  filter(detectorType!="MTCNN-Faces") %>%
  filter(detectorType!="ViolaJones-Faces") %>%
  filter(!is.na(posture)) %>%
  filter(!is.na(orientation))

# pop out of actual ages of included subs
ages_inc_subs <- inc_subs %>%
  select(subid, age.at.test, age.grp)

# raw data
detectionsByPostureAndOrient <- dlongPlot %>%
  group_by(age.grp, posture, orientation, subid, detectorType) %>%
  summarise(prop_detected = mean(detection, na.rm=TRUE), num_detections = length(detection)) %>%
  left_join(ages_inc_subs) %>%
  mutate(age.at.test = as.numeric(age.at.test))

# with CIs
detectionsByPostureAndOrientByGroup <- dlongPlot %>%
  group_by(age.grp, orientation, posture, subid, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  group_by(age.grp, orientation, posture, detectorType) %>%
  multi_boot_standard(col = 'prop_detected')

```

```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\\% bootstrapped confidence intervals"}

# better labels 
orientation_labels <- c(
  behind ='Caregiver behind',
  close = 'Caregiver close',
  far = 'Caregiver far')

posture_labels = c(
  prone = 'Infant prone',
  sit = 'Infant sitting',
  stand ='Infant standing'
)

 ## plot data all together now, with CIs and individual points from each infant
ggplot(detectionsByPostureAndOrientByGroup, aes(x = age.grp, y = mean, col = detectorType)) + 
  # plot raw data
  geom_point(data=detectionsByPostureAndOrient, alpha=.2, aes(x = age.at.test, y=prop_detected, size=num_detections)) +
  # plot trend lines & SEs of trends based on raw data
  geom_smooth(data=detectionsByPostureAndOrient, aes(x = age.at.test, y=prop_detected, weight = num_detections), alpha=0.1, size=0, span=5) +
  stat_smooth(data=detectionsByPostureAndOrient, aes(x = age.at.test, y=prop_detected, weight = num_detections), geom="line", alpha=0.5, size=1, span=5) +
  # plot means and CIs based on age groups
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=1), alpha=.9) +
  # fix up the plot to look nice
  labs(y = "Proportion detections", x = "Age (months)") +
  theme_few(base_size=16) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 12)) +
  scale_colour_discrete(name = 'Detection type') +
  scale_size_continuous(name = 'Number of frames') +
  scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  facet_grid(orientation ~ posture, labeller = labeller(posture = posture_labels, orientation= orientation_labels))

```

```{r}
source("helper.R")
## summarize detections around naming events
aroundNamings_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.))

aroundNamings_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceOP) %>%
  do(summarize.naming(.)) 
```

```{r}
### Analyze detections around naming events (without posture/orientation)

### first by faces
aroundNamings_Detections_BySub_Faces <- aroundNamings_Faces %>%
  group_by(subid, familiarity, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE))

aroundNamings_Detections_Grouped_Faces <- aroundNamings_Detections_BySub_Faces %>%
  group_by(familiarity, age.grp) %>%
  multi_boot_standard(col = 'prop_detected')

### now by wrists 
aroundNamings_Detections_BySub_Wrists <- aroundNamings_Wrists %>%
  group_by(subid, familiarity, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE))

aroundNamings_Detections_Grouped_Wrists <- aroundNamings_Detections_BySub_Wrists %>%
  group_by(familiarity, age.grp) %>%
  multi_boot_standard(col = 'prop_detected')

### get individual points
detectionsByGroup <- dlongPlot %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  group_by(age.grp, detectorType) %>%
  multi_boot_standard(col = 'prop_detected')
```

```{r}
## Plot faces/wrists around namings by age

faces_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
  # plot baseline
  geom_pointrange(data = (detectionsByGroup %>% filter(detectorType == 'OpenPose-Faces')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=3), alpha=.9) +
  # plot averages by familiarity 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
  # plot familairity detections for each sub
  geom_point(data=aroundNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
  theme_few() +
  labs(y = "Prop. face detections", x = "Age (months)") +
  theme_few(base_size=16) +
  ylim(0,.6) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")

## Faces
wrists_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Wrists, aes(x=age.grp, y=mean, color=familiarity)) +
  # plot baseline
  geom_pointrange(data = (detectionsByGroup %>% filter(detectorType == 'OpenPose-Wrists')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=3), alpha=.9) +
  # plot averages by familiarity 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
  # plot familairity detections for each sub
  geom_point(data=aroundNamings_Detections_BySub_Wrists, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
  theme_few() +
  labs(y = "Prop. wrist detections", x = "Age (months)") +
  theme_few(base_size=16) +
  ylim(0,.6) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")

```

```{r detByNaming, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\\% bootstrapped confidence intervals."}
cowplot::plot_grid(faces_aroundnamings_plot, wrists_aroundnamings_plot)
```

# Extension to Franchak et al., 2017
In the present work, we found that infant's in-the-moment posture changed with their age, as did infants' orientation relative their caregiver. In a related study with 12-month-olds, @franchak2017 also that children's in-the-moment posture changed the amount of time infants spent looking at faces. Here, we sought to replicate the findings from @franchak2017 using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras (shared on Databrary, @Databrary).

This dataset differed from the present in two key ways that could present a challenge for our automated methodology. First, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to explore, unlike the present room  dataset which was relatively small (approximately 10' x 10'). Thus, our automated methodology could fail to generalize to scenes from these more complex environments, where detecting faces and hands could arguably be a much harder task. Second, @franchak2017 used a head-mounted eye-tracker, finding that children's posture affected where children were looking within their visual field. Nonetheless, if children are often orienting their heads towards where they are allocating their attention  [@Yoshida20008] then we should expect to find the same pattern of results only when analyzing the information in view.

```{r}
# loads pos_detections used below for franchak dataset
load(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections.RData')
```

```{r}
faces_by_pos  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_faces = mean(face_openpose)) %>%
  group_by(infant_posture_label) %>%
  multi_boot_standard(col = 'prop_faces')

faces_by_pos_by_sub  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_faces = mean(face_openpose), faces = sum(face_openpose), no_faces = sum(face_openpose==0))

hands_by_pos  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_hands = mean(hand_openpose)) %>%
  group_by(infant_posture_label) %>%
  multi_boot_standard(col = 'prop_hands')

hands_by_pos_by_sub  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_hands = mean(hand_openpose), hands = sum(hand_openpose), no_hands = sum(hand_openpose==0))
```

```{r}
# glmers for franchak dataset
posture_model_faces_franchak <- glmer(cbind(faces,no_faces) ~ infant_posture_label + (1|sub_id), data=faces_by_pos_by_sub, family="binomial")

posture_model_hands_franchak <- glmer(cbind(hands,no_hands) ~ infant_posture_label + (1|sub_id), data=hands_by_pos_by_sub, family="binomial")
```

```{r}
faces_by_pos_plot <- ggplot(faces_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
   geom_point(data=faces_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label), alpha=.1) +
  geom_line(data=faces_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label, group=sub_id), alpha=.1) +
  theme_few(base_size=18) +
  ylab('Prop faces detected') +
  xlab('Infant posture') + 
  ylim(c(0,.3)) +
  theme(legend.position = 'none') +
  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928"))

hands_by_pos_plot <- ggplot(hands_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  geom_point(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_hands, col=infant_posture_label), alpha=.1) +
  geom_line(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_hands, col=infant_posture_label, group=sub_id), alpha=.1) +
  theme_few(base_size=18) +
  ylab('Prop hands detected') +
   xlab('Infant posture') + 
  ylim(c(0,.4)) +
  theme(legend.position = 'none') +
  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928")) 
```

```{r franchak, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals."}
cowplot::plot_grid(faces_by_pos_plot,hands_by_pos_plot, nrow=1)
```

Overall, we found convergence between our two methodologies: we replicated @Franchak2017's main results, finding that the proportion of faces in view was greater when infants were sitting or standing vs. prone. We also found the same patterns of results with hands (i.e., wrist detectinos), which were not originally annotated in @Franch2017 (see Figure \ref{fig:franchak}); results were validated with generalized mixed effect models as in our previous analysis. Interestingly, there was a relatively higher proportion of hands in view in @Franchak2017 relative to the present dataset; this is likely mostly due to the fact that there were often multiple people in view: experimenters stayed in the room to film children and assist with technical difficulties, while in our present dataset the experiments left the room for the majority of the play session. Nonetheless, we still found that posture modulated the hands in view as well, suggesting that this is a major factor that structures infants' access to visual information.

# General Discussion
We used a head-mounted camera to explore how children's postural and locomotor development affects their visual access to social information, here operationalized as the presence of the faces and hands of their caregiver. Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors influenced the proportion of faces and hands in the child's visual field. Though our cameras were oriented to capture face information, we nonetheless saw an influence of orientation and posture on visual access to hands. Motor development appears to modulate how infants experience their visual world and the social information in it.

We created a situation in which the context of interaction was tightly controlled, but as children grow and change, the activities in which they engage with their caregivers also vary, leading to differences in the distribution of contexts they experience. Thus, more work is needed to understand how our results relate to children's home experiences [@fausey2016]. The ability to walk is part of a cascade of changes in children's  experience, and our study captures only a cross-sectional slice of this broader, multifaceted trajectory.

Understanding these changes has been a persistent challenge for developmental psychology, but the field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools. These tools deal better with noisier, more complicated datasets and extract richer information than previous systems. We hope that these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.

# Acknowledgements

Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation. This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and NSF #1714726 to BLL. An earlier version of this work was presented to the Cognitive Science Society in @frank2013 and @sanchez2018.

\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/amsan7/xs-face}}} \vspace{1em}

\newpage

# References
```{r create_r-references}
# r_refs(file = "xsface.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2020-04-10 writeup-paper/draft_apa_march2020.tex
add writeup folder for apa draft

\documentclass[english,man,floatsintext,draftall]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Postural developments modulate children's visual access to social information},
            pdfkeywords={Postural developments modulate children's visual access to social
information},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Posture modulates social access}
\author{Bria L. Long\textsuperscript{1}, Alessandro Sanchez\textsuperscript{1}, Allison M. Kraus\textsuperscript{1}, Ketan Agrawal\textsuperscript{1}, \& Michael C. Frank\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\authornote{

Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
\keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}

\title{Postural developments modulate children's visual access to social
information}

\date{}

\abstract{
The ability to process social information is a critical component of
children's early language and cognitive development. However, as
children reach their first birthday, they begin to locomote themselves,
dramatically affecting their visual access to this information. How do
these postural and locomotor changes affect children's access to the
social information relevant for word-learning? Here, we explore this
question by using head-mounted cameras to record 36 infants' (8-16
months of age) egocentric visual perspective and use computer vision
algorithms to estimate the proportion of faces and hands in infants'
environments. We find that infants' posture and orientation to their
caregiver modulates their access to social information, confirming
previous work that suggests motoric developments play a significant role
in the emergence of children's linguistic and social capacities. In
addition, we applied our automated analysis to the headcam data from a
recent paper showing similar effects postre on visual attention
(Franchak et al., 2017), finding convergence across methodologies,
laboratories, and datasets. We suggest that the combined use of
head-mounted cameras and the application of new computer vision
techniques is a promising avenue for understanding the statistics of
infants' visual and linguistic experience as they change over
development.
}

\begin{document}
\maketitle

From their earliest months, infants are deeply engaged in learning from
others. Even newborns tend to prefer to look at faces with direct
vs.~averted gaze (Farroni, Csibra, Simion, \& Johnson, 2002) and young
infants follow overt gaze shifts. And as infants learn to parse and
segment their native language(s), the social cues provided by speakers
(e.g., eye-gaze) may provide strong scaffolding for early word learning.
Indeed, empirical work suggests that children's ability to process
social cues is a key factor in early language development: in a
longitudinal study, children's level of joint engagement with their
mother at 9-12 months was found to predict both their receptive and
productive vocabularies (Carpenter, Nagell, \& Tomasello, 1998). More
specifically, 10 month-olds who follow an adult's gaze in an
experimental context have larger vocabularies at 18 months and
throughout the second year of life (R. Brooks \& Meltzoff, 2005; Rechele
Brooks \& Meltzoff, 2008).

Around their first birthday, however, children's ability to interact
with their world radically changes (Adolph \& Berger, 2006) as they are
no longer constrained to the same spot that their caregivers last placed
them in. Before the first birthday, children often begin crawling; soon
after, they begin to walk independently. While not all children crawl
({\textbf{???}})---or crawl in the same way---children tend to find
creative ways of moving (e.g., scooting) or getting others to help them
move (e.g cruising) to begin to explore their world on their own
({\textbf{???}}). And as independent agents in the world, children are
more active in the construction of their own learning environments. Yet
while toddlers do determine their own input to a much greater degree,
they spend much of their time in a world primarily populated by knees

One possibility is that these motor improvements have strong
developmental cascades, impacting children's emerging social, cognitive,
and linguistic abilities (Iverson, 2010). Indeed, these postural changes
change how children interact with their mothers; walking infants make
different kinds of object-related bids for attention from their mothers
than crawling infants, and tend to hear more action directed statements
(e.g., \enquote{open it}) (Karasik, Tamis-LeMonda, \& Adolph, 2014). In
an observational study, Walle and Campos (2014) also found that children
who were able to walk had both higher receptive and productive
vocabularies. On their account, children's ability to stand and
independently locomote may fundamentally change their ability to access
the social information (e.g., faces, gaze) relative to children who are
still crawling and sitting. In other words, the ability of walking
infants to access more detailed social information may allow infants to
learn words quicker and more efficiently, facilitating language growth.

Over the past decade, researchers have started to use egocentric,
head-mounted cameras to document the social information that infants and
children have access to across early development, and to understand the
degree to which there are substantial shifts in their viewpoints that
may have downstream developmental consequences ({\textbf{???}}). Indeed,
the infant visual perspective has proved to be both remarkably different
than the adult perspective and not easily predicted by our own
intuitions (Clerkin, Hart, Rehg, Yu, \& Smith, 2017; Franchak, Kretch,
Soska, \& Adolph, 2011, Yoshida and Smith (2008)). Over the first two
years of life, the infant viewpoint transitions from primarily
containing close up view of faces to capturing relatively restricted
views of hands paired with the objects they are acting on (Fausey,
Jayaraman, \& Smith, 2016); both computational and empirical work
suggest that this restricted viewpoint is more effective for learning
about objects and their labels than the comparable adult perspective (D.
Yurovsky, Smith, \& Yu, in press, Bambach, Crandall, Smith, and Yu
(2017)).

Here, we directly examine the role that postural developments that
infants' experience as they reach their first birthday change the social
information in the infant view. We hypothesize that children's postural
developments may be partially responsible for some of these changes in
the infant perspective. During spontaneous play, toddlers are more
likely to look at the floor while crawling than while walking (Franchak
et al., 2011), when they have full visual access to their environment
and the people in it (Kretch, Franchak, and Adolph, 2014). More
recently, when 12-month-olds ({\textbf{???}}) participated in a
free-play session with their caregivers, their own posture and as well
as their caregiver's posture influenced the proportion of time they
spent looking at the faces and bodies of their caregivers
({\textbf{???}}).

Here, we expand on these previous findings in three key ways. First,
using head-mounted cameras during a brief laboratory free-play session,
we record the visual experience of three groups of children at 8, 12,
and 16 months-of-age, covering a broad range of ages and locomotor
abilities around the first birthday; Children's posture and orientation
relative to their caregiver were also recorded from a third-person
perspective and hand-annotated. This cross-sectional design thus allows
us to directly examine the relative contributions of age vs.~postural
developments on children's visual access to social information. Second,
we use a modern computer vision algorithm for the automated detection of
faces and hands in the egocentric viewpoints of infants. In particular,
we capitalize on recent improvements in face and pose detection
algorithms (Cao, Simon, Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \&
Qiao, 2016) to analyze the frequencies of faces and hands in the child's
visual environment, both overall and relative to naming events by their
caregivers. Use of this emerging technology allows us to annotate the
entirety of the dataset, allowing a more complete picture of the
changing infant perspective. In addition, we extend this technique to
the same dataset used in ({\textbf{???}}), replicating their main
pattern of results using this novel methodology and finding convergence
between our datasets with different play sessions, populations, and
head-mounted cameras. Finally, as parents were presented with pairs of
novel and familiar objects with their children during the play sessions,
we directly examine the availability of social signals during naming
events.

Thus, our current dataset allows us to analyze changes in the visual
access to faces and hands according to children's age, posture, and
linguistic input. We predicted that there would be differential access
to social information based on children's postural developments:
crawling infants would see fewer faces/hands because they would
primarily be looking at the ground, while walking toddlers would have
access to a richer visual landscape with greater access to the social
information in their environment. {[}language predictions?{]}

\section{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

Our final sample consisted of 36 infants and children, with 12
participants in three age groups: 8 months (6 F), 12 months (7 F), and
16 months (6 F). Participants were recruited from the surrounding
community via state birth records, had no documented disabilities, and
were reported to hear at least 80 percent English at home. Demographics
and exclusion rates are given in the table below.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
\hline
Group & N & \% incl. & Avg age & Avg video length (min) \\
\hline
8 mo. &   12 & 0.46 & 8.71 & 14.41 \\
12 mo. &  12 & 0.40 & 12.62 & 12.71 \\
16 mo. &  12 & 0.31 & 16.29 & 15.10\\
\hline
\end{tabular}
\end{table}

To obtain this final sample, we tested 95 children, excluding 59
children for the following reasons: 20 for technical issues related to
the headcam, 15 for failing to wear the headcam, 10 for fewer than 4
minutes of headcam footage, 5 for having multiple adults present, 5 for
missing Communicative Development Inventory (CDI) data, 2 for missing
scene camera footage, 1 for fussiness, and one for sample symmetry. All
inclusion decisions were made independent of the results of subsequent
analyses. Some of these data were also analyzed in Frank, Simmons,
Yurovsky, and Pusiol (2013).

\subsubsection{Procedure}\label{procedure}

All parents signed consent documents while children were fitted with the
headcam. If the child was uninterested in wearing the headcam or tried
to take it off, the experimenter presented engaging toys to try to draw
the child's focus away from the headcam. When the child was comfortable
wearing the headcam, the child and caregiver were shown to a playroom
for the free-play session. Parents were shown a box containing three
pairs of novel and familiar objects (e.g., a ball and a microfiber
duster, named a \enquote{zem}), and were instructed to play with the
object pairs with their child one at a time, \enquote{as they typically
would.} All parents confirmed that their child had not previously seen
the novel toys and were instructed to use the novel labels to refer to
the toys. The experimenter then left the playroom for approximately 15
minutes, during which a tripod-mounted camera in the corner of the room
recorded the session and the headcam captured video from the child's
perspective.

\subsubsection{Data Processing and
Annotation}\label{data-processing-and-annotation}

Headcam videos were trimmed such that they excluded the instruction
phase when the experimenter was in the room and were automatically
synchronized with the tripod-mounted videos using FinalCut Pro Software.
These sessions yielded 507 minutes (almost a million frames) of video,
with an average video length of 14.07 minutes (min = 4.53, max = 19.35).

\subsubsection{Posture and Orientation
Annotation}\label{posture-and-orientation-annotation}

We created custom annotations to describe the child's physical posture
(e.g.~standing) and the orientation of the child relative to the
caregiver (e.g.~far away). The child's posture was categorized as being
held/carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations.

\subsection{Face and Hand Detection}\label{face-and-hand-detection}

We used three face detection systems to measure infants' access to
faces. The first of these is the most commonly-used and widely available
face detection algorithm: Viola-Jones. We used this algorithm as a
benchmark for performance, as while it can achieve impressive accuracy
in some situations, it is notoriously bad at dealing with occluded faces
(Scheirer, Anthony, Nakayama, \& Cox, 2014). We next tested the
performance of two face detectors that both made use of recently
developed Convolutional Neural Networks (CNNs) to extract face
information. The first algorithm was specifically optimized for face
detection, and the second algorithm was optimized to extract information
about the position of 18 different body parts. For the second algorithm
(OpenPose; Cao et al., 2017), we used the agent's nose (one of the body
parts detected) to operationalize the presence of faces, as any half of
a face necessarily contains a nose.

The OpenPose detector also provided us with the location of an agent's
wrists, which we used as a proxy for hands for two reasons. First, as we
did not capture children's entire visual field, the presence of a wrist
is likely often indicative of the presence of a hand within the field of
view. Second, hands are often occluded by objects when caregivers are
interacting with children, yet still visually accessible by the child
and part of their joint interaction.

\begin{figure*}
\centering
\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

\subsubsection{Algorithms}\label{algorithms}

The first face detection system made use of a series of Haar
feature-based cascade classifiers (Viola \& Jones, 2004) applied to each
individual frame. The second algorithm (based on work by K. Zhang et al.
(2016)) uses multi-task cascaded convolutional neural networks (MTCNNs)
for joint face detection and alignment, built to perform well in
real-world environments where varying illuminations and occlusions are
present. We used a Tensorflow implementation of this algorithm avaliable
at \url{https://github.com/davidsandberg/facenet}.

The CNN-based pose detector (OpenPose; Cao et al., 2017; Simon, Joo,
Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016)
provided the locations of 18 body parts (ears, nose, wrists, etc.) and
is available at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose; here, we only made
use of the body parts relevant to the face and hands (nose and wrists).

\subsubsection{Detector evaluation}\label{detector-evaluation}

To evaluate face detector performance, we hand-labeled a \enquote{gold
set} of labeled frames. To account for the relatively rare appearance of
faces in the dataset, we hand-labeled two types of samples: a sample
containing a high density of faces (half reported by MTCNN, half by
OpenPose) and a random sample from the remaining frames. Each sample was
comprised of an equal number of frames taken from each child's video.
For wrist detections, the \enquote{gold set} was constructed in the same
manner, except frames with a high density of wrists came only from
detections made by OpenPose. Faces were classified as present if at
least half of the face was showing; wrists were classified as present if
any part of the wrist was showing. Two authors labelled the frames
independently and resolved disagreements on a case-by-case basis.
Precision (hits / hits + false alarms), recall (hits / hits + misses),
and F-score (harmonic mean of precision and recall) were calculated for
all detectors and are reported in Table 1.

For face detection, MTCNN outperformed OpenPose when taking into account
only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose). Although
MTCNN and OpenPose performed comparably with the random sample, MTCNN
performed better on the high density sample (specifically looking at
precision), suggesting that OpenPose generated more false positives than
MTCNN. ViolaJones performed quite poorly relative to the other
detectors, especially with respect to the random sample. We thus use
MTCNN detections in the following analyses. For wrist detection,
OpenPose performed moderately well (F = 0.74) with relatively high
precision but low recall on the randomly sampled frames (see Table 1).
We thus analyze wrist detections, with the caveat that we are likely
underestimating the proportion of hands in the dataset.

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
\hline
Algorithm & Sample\ Type & P & R & F \\ 
\hline
MTCNN-Faces & High density & 0.89 & 0.92 & \textbf{0.90} \\ 
MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
OpenPose-Faces & Random & 0.72 & 0.80 & \textbf{0.76} \\ 
ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
OpenPose-Wrists & High density & 0.66 & 1.00 & 0.79 \\ 
OpenPose-Wrists & Random & 0.88 & 0.29 & 0.44 \\ 
\hline
\end{tabular}
\caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected). P, R, and F denote precision, recall, and F-score, respectively. Scores in bold are the highest F-scores for each sample type.} 
\vspace{-1em}
\end{table}

\section{Results}\label{results}

First, we report developmental shifts in infants' posture and their
orientation relative to their caregiver, consistent with previous
literature ({\textbf{???}}, {\textbf{???}}). Then, we examine how these
changes influence children's visual access to faces and wrists/hands
across this developmental time range, examining the relative
contributions of age vs.~postural developments. We also explore how
these changes impact the accessibility of faces and wrists/hands during
labeling events, both when parents were labeling familiar objects (e.g.,
\enquote{cat}) as well as novel objects (e.g., \enquote{zem}). Finally,
we apply the same automated detection method to a egocentric video
dataset collected by a different lab ({\textbf{???}}) during which
children's in-the-moment posture was annotated, replicating their
existing results and finding converging support across methods, labs,
and paradigms for the impact of postural developments on children's
visual access to social information.

\begin{figure*}[h]

{\centering \includegraphics{figs/posture-1}

}

\caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:posture}
\end{figure*}

\begin{figure*}[h]

{\centering \includegraphics{figs/indivPosOrient-1}

}

\caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:indivPosOrient}
\end{figure*}

\subsection{Changes in Posture and
Orientation}\label{changes-in-posture-and-orientation}

How do childen's in-the-moment posture and orientation change across
age? In our dataset, the proportion of time infants spent sitting
decreased with age, and the proportion of time infants spent standing
increased with infants' age. As children got older, their locomotive
abilities allowed them to become more independent. Both 8-month-olds and
12-month-olds spent relatively equivalent amounts of time lying/crawling
(i.e., \enquote{prone}) which was markedly decreased in the
16-month-olds, who spent most of their time sitting or standing (see
Figure \ref{fig:posture}). We also observed changes in children's
orientation relative to their caregivers: the 8-month-olds spent more
time with their caregiver behind them supporting their sitting positions
than did children at other ages (see Figure \ref{fig:posture}). However,
we also saw considerable variability across children: some infants spent
almost their entire time sitting at a close distance from their
caregiver, whereas others showed more considerable variability (see
Figure \ref{fig:indivPosOrient}).

\subsection{Changes in Access to Faces and
Hands}\label{changes-in-access-to-faces-and-hands}

We first examined the proportion of face and hand detections as a
function of children's age without considering their posture (see Figure
\ref{fig:detByAge}). While faces tended to be in the field-of-view
overall more often than hands, children's head-mounted cameras were
angled slightly upward to capture the presence of faces, and hand
detections suffered from somewhat lower recall than face detections.
Going forward, we thus analyze differences in the relative proportion of
faces or hands in view as a function of age, posture, and orientation,
rather than comparing them directly. Overall, we observed that
12-month-olds appeared to have visual access to slightly fewer faces
than 8 or 16-month-olds, creating a slight U-shaped function in face
detections; conversely, hand detections were showed a slight increase
across this age range, as reported in prior literature ({\textbf{???}}).

However, these age-related trends were much smaller than the effect of
infant's postural developments on children's visual access to faces and
hands. Children's in-the-moment posture was a major factor both in how
many faces and hands were in view during the play session, as was their
orientation relative to their caregiver. Infants who were sitting or
standing had more faces in view than infants who were lying
down/crawling (i.e.~prone), which was most frequent among 12-month-olds
relative to the other age groups (Figure \ref{fig:detByPosOrient}). When
caregivers were behind their children, supporting their children's
sitting or standing positions, children saw fewer faces and wrists. In
particular, children who were sitting or in front of their caregiver had
a high proportion of faces and hands in their field of view. (Figure
\ref{fig:detByPosOrient}, lower panel).

These trends were quantified using two generalized linear mixed-effect
models estimating the proportion of faces and hands that were in view,
with orientation, posture, their interaction, and scaled participant's
age as fixed effects, and with random slopes for infants' orientation
and posture. A summary of the coefficients of the models can be found in
Table 2, confirming that infants who were sitting/standing and in front
of their caregivers saw the most faces and hands. When modeling the
proportion of faces seen, age was not a significant predictor; however,
age remained a significant predictor when modeling the proportion of
hands seen by infants. Overall, these results suggest that infants'
visual access to social information is largely modulated by their
posture and orientation to their caregiver, which is in turn a function
of their general locomotor development. Nonetheless, these results also
suggest that infants may still increase their attention to hands
throughout this age range as they continue to learn about objects, their
functions, and their names.

\subsection{Access to Faces and Hands During Labeling
Events}\label{access-to-faces-and-hands-during-labeling-events}

Our play session was designed to provide parents with opporunities to
label objects--both familiar and novel--such that we could examine
whether children seek out different kinds of social information around
naming events. We thus explored how face and hand detections changed
during object labeling events, analyzing a four-second window around
each labeling event (e.g., \enquote{Look at the {[}zem{]}!}). Every
utterance of one of the labelled objects (e.g., \enquote{ball}) was
counted as a \enquote{labeling event}; timestamps of the beginning of
each word were hand-annotated from transcripts ad synchronized with the
frame-by-frame detections. Overall, we did not find major differences in
face/hand detections during naming events relative to baseline, either
for novel objects or for familiar objects (see Figure
\ref{fig:detByNaming}). This was true both when we analyzed naming rates
with or without taking into account children's posture and orientation
relative to their caregiver. These results suggest that either children
did not actively seek out social information during this particular play
session around naming events, perhaps because were a limited number of
possible referents at any given time.

\begin{figure*}[h]

{\centering \includegraphics{figs/detByAge-1}

}

\caption{Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data.}\label{fig:detByAge}
\end{figure*}

\begin{figure*}[h]

{\centering \includegraphics{figs/detByPosOrient-1}

}

\caption{Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\% bootstrapped confidence intervals}\label{fig:detByPosOrient}
\end{figure*}

\begin{figure*}[h]

{\centering \includegraphics{figs/detByNaming-1}

}

\caption{Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\% bootstrapped confidence intervals.}\label{fig:detByNaming}
\end{figure*}

\section{Extension to Franchak et al.,
2017}\label{extension-to-franchak-et-al.-2017}

In the present work, we found that infant's in-the-moment posture
changed with their age, as did infants' orientation relative their
caregiver. In a related study with 12-month-olds, ({\textbf{???}}) also
that children's in-the-moment posture changed the amount of time infants
spent looking at faces. Here, we sought to replicate the findings from
({\textbf{???}}) using our automated methodology (OpenPose detections),
using the footage from their head-mounted cameras (shared on Databrary,
({\textbf{???}})).

This dataset differed from the present in two key ways that could
present a challenge for our automated methodology. First, the
environment that infants were immersed in with their caregivers (and
experimenters) was much larger and more varied than the play room used
in the present dataset, containing multiple structures and toys in
different parts of the room for infants to explore, unlike the present
room dataset which was relatively small (approximately 10' x 10'). Thus,
our automated methodology could fail to generalize to scenes from these
more complex environments, where detecting faces and hands could
arguably be a much harder task. Second, ({\textbf{???}}) used a
head-mounted eye-tracker, finding that children's posture affected where
children were looking within their visual field. Nonetheless, if
children are often orienting their heads towards where they are
allocating their attention ({\textbf{???}}) then we should expect to
find the same pattern of results only when analyzing the information in
view.

\begin{figure*}[h]

{\centering \includegraphics{figs/franchak-1}

}

\caption{Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals.}\label{fig:franchak}
\end{figure*}

Overall, we found convergence between our two methodologies: we
replicated ({\textbf{???}})\enquote{s main results, finding that the
proportion of faces in view was greater when infants were sitting or
standing vs.~prone. We also found the same patterns of results with
hands (i.e., wrist detectinos), which were not originally annotated in
({\textbf{???}}) (see Figure \ref{fig:franchak}); results were validated
with generalized mixed effect models as in our previous analysis.
Interestingly, there was a relatively higher proportion of hands in view
in ({\textbf{???}}) relative to the present dataset; this is likely
mostly due to the fact that there were often multiple people in view:
experimenters stayed in the room to film children and assist with
technical difficulties, while in our present dataset the experiments
left the room for the majority of the play session. Nonetheless, we
still found that posture modulated the hands in view as well, suggesting
that this is a major factor that structures infants} access to visual
information.

\section{Discussion}\label{discussion}

\section{Discussion}\label{discussion-1}

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-adolph2006motor}{}
Adolph, K. E., \& Berger, S. E. (2006). Motor development.
\emph{Handbook of Child Psychology}.

\hypertarget{ref-adolph2012toward}{}
Adolph, K. E., Gilmore, R. O., Freeman, C., Sanderson, P., \& Millman,
D. (2012). Toward open behavioral science. \emph{Psychological Inquiry},
\emph{23}(3), 244--247.

\hypertarget{ref-bambach2017}{}
Bambach, S., Crandall, D. J., Smith, L. B., \& Yu, C. (2017). An
egocentric perspective on active vision and visual object learning in
toddlers. In \emph{Proceedings of the seventh joint ieee conference on
development and learning and on epigenetic robotics}.

\hypertarget{ref-brooks2005}{}
Brooks, R., \& Meltzoff, A. (2005). The development of gaze following
and its relation to language. \emph{Developmental Science}, \emph{8}(6),
535--543.

\hypertarget{ref-brooks2008}{}
Brooks, R., \& Meltzoff, A. N. (2008). Infant gaze following and
pointing predict accelerated vocabulary growth through two years of age:
A longitudinal, growth curve modeling study. \emph{Journal of Child
Language}, \emph{35}(1), 207--220.

\hypertarget{ref-cao2017realtime}{}
Cao, Z., Simon, T., Wei, S.-E., \& Sheikh, Y. (2017). Realtime
multi-person 2D pose estimation using part affinity fields. In
\emph{CVPR}.

\hypertarget{ref-carpenter1998}{}
Carpenter, M., Nagell, K., \& Tomasello, M. (1998). Social cognition,
joint attention, and communicative competence from 9 to 15 months of
age. \emph{Monographs of the Society for Research in Child Development},
\emph{63}(4).

\hypertarget{ref-clerkin2017}{}
Clerkin, E. M., Hart, E., Rehg, J. M., Yu, C., \& Smith, L. B. (2017).
Real-world visual statistics and infants' first-learned object names.
\emph{Phil. Trans. R. Soc. B}, \emph{372}(1711), 20160055.

\hypertarget{ref-farroni2002eye}{}
Farroni, T., Csibra, G., Simion, F., \& Johnson, M. H. (2002). Eye
contact detection in humans from birth. \emph{Proceedings of the
National Academy of Sciences}, \emph{99}(14), 9602--9605.

\hypertarget{ref-fausey2016}{}
Fausey, C. M., Jayaraman, S., \& Smith, L. B. (2016). From faces to
hands: Changing visual input in the first two years. \emph{Cognition},
\emph{152}, 101--107.

\hypertarget{ref-franchak2011}{}
Franchak, J. M., Kretch, K. S., Soska, K. C., \& Adolph, K. E. (2011).
Head-mounted eye tracking: A new method to describe infant looking.
\emph{Child Development}, \emph{82}(6), 1738--1750.

\hypertarget{ref-frank2013}{}
Frank, M. C., Simmons, K., Yurovsky, D., \& Pusiol, G. (2013).
Developmental and postural changes in children's visual access to faces.
In \emph{Proceedings of the 35th annual meeting of the cognitive science
society} (pp. 454--459).

\hypertarget{ref-iverson2010}{}
Iverson, J. M. (2010). Developing language in a developing body: The
relationship between motor development and language development.
\emph{Journal of Child Language}, \emph{37}(2), 229--261.

\hypertarget{ref-karasik2014}{}
Karasik, L. B., Tamis-LeMonda, C. S., \& Adolph, K. E. (2014). Crawling
and walking infants elicit different verbal responses from mothers.
\emph{Developmental Science}, \emph{17}(3), 388--395.

\hypertarget{ref-scheirer2014perceptual}{}
Scheirer, W. J., Anthony, S. E., Nakayama, K., \& Cox, D. D. (2014).
Perceptual annotation: Measuring human vision to improve computer
vision. \emph{IEEE Transactions on Pattern Analysis and Machine
Intelligence}, \emph{36}(8), 1679--1686.

\hypertarget{ref-simon2017hand}{}
Simon, T., Joo, H., Matthews, I., \& Sheikh, Y. (2017). Hand keypoint
detection in single images using multiview bootstrapping. In
\emph{CVPR}.

\hypertarget{ref-viola2004robust}{}
Viola, P., \& Jones, M. J. (2004). Robust real-time face detection.
\emph{International Journal of Computer Vision}, \emph{57}(2), 137--154.

\hypertarget{ref-walle2014}{}
Walle, E. A., \& Campos, J. J. (2014). Infant language development is
related to the acquisition of walking. \emph{Developmental Psychology},
\emph{50}(2), 336.

\hypertarget{ref-wei2016cpm}{}
Wei, S.-E., Ramakrishna, V., Kanade, T., \& Sheikh, Y. (2016).
Convolutional pose machines. In \emph{CVPR}.

\hypertarget{ref-yoshida2008}{}
Yoshida, H., \& Smith, L. (2008). What's in view for toddlers? Using a
head camera to study visual experience. \emph{Infancy}, \emph{13},
229--248.

\hypertarget{ref-yurovsky2012}{}
Yurovsky, D., Smith, L., \& Yu, C. (in press). Statistical word learning
at scale: The baby's view is better. \emph{Developmental Science}.

\hypertarget{ref-zhang2016}{}
Zhang, K., Zhang, Z., Li, Z., \& Qiao, Y. (2016). Joint face detection
and alignment using multitask cascaded convolutional networks.
\emph{IEEE Signal Processing Letters}, \emph{23}(10), 1499--1503.

\endgroup

\end{document}

### 2020-04-10 writeup-paper/old-drafts/draft_1_from_CogSci.Rmd
add writeup folder for apa draft

--- 2018-11-06 writeup-paper/cogsci2018_cameraReady.Rmd
@@
 library(assertthat)
-
+source("helper.R") ## script for merging in naming annotations.
 ```
 # Introduction
-From early infancy, children are deeply engaged in learning from others.
-Even newborns tend to prefer to look at faces with direct vs.
-averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010development].
-Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring].
-
-Then, however, their view of the world radically changes [@adolph2006motor].
-Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently.
-These motor changes have significant consequences for what children see.
-For example, during spontaneous play in a laboratory playroom, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011]; in general, walking infants tend to have full visual access to their environment and the people in it, while crawling infants do not [@kretch2014].
-
-These motor improvements lead to developmental cascades that impact children’s emerging social, cognitive, and linguistic abilities in various ways.
-For example, postural changes impact how children interact with their mothers: walking (vs.
-crawling) infants make different kinds of object-related bids for attention from their mothers, tend to hear more action directed statements (e.g., “open it”), and tend to explore their surroundings more [@karasik2014; @thurman2017spatial].
-In an observational study, @walle2014 found that children who were able to walk had both higher receptive and productive vocabularies.
-Thus, children’s ability to independently locomote may change their ability to access social information (e.g., faces, gaze) and in turn accelerate their learning.
-
-Recent technological developments allow for testing of this hypothesis by documenting the experiences of infants and children from their own perspective.
-By using head-mounted cameras, researchers have begun to record the visual experiences of infants and children -- which even for walking children are extremely different from the adult perspective (and not easily predicted by our own adult intuitions) [@clerkin2017;@franchak2011;@yoshida2008].
-Children’s views tend to be more restricted and dominated by objects and hands  [@yoshida2008], and computational and empirical work suggests that this restricted viewpoint may be more effective for learning objects and their labels than the comparable adult perspective [@bambach2017;@yurovsky2013statistical].
-This perspective changes over the first two years of life, as visual input transitions from primarily containing close up views of faces to capturing views of hands paired with the objects they are acting on [@fausey2016].
-
-Here, we directly examine whether postural and locomotor developments change the availability of social information -- the presence of faces and hands.
-We recorded the visual experience of a group of infants in three age ranges (8, 12, and 16 months) using head-mounted cameras during a brief laboratory free-play session; children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
-Recent work directly examined how postural developments change access to social information in 12 month-olds in a similar paradigm, finding support for the motoric hypothesis [@franchak2017see].
-Here we replicate and extend their main finding using novel computer vision techniques to allow for video analysis at scale.
-In particular, we capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands (using wrists as a proxy for the latter) in the child’s visual environment, both overall and relative to naming events by their caregivers.
-We predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape, thus providing children with greater access to the social information in their environment.
-
 # Methods
@@
 When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session.
-Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys.
+Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." These three pairs were taken from a superset of five paired groups of novel and familiar objects (familiar objects: cat, toothbrush, car, rubber duck, toy car; all stimuli avaliable at XXX.
+All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys.
 The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.
@@
 far away).
-The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing.
-The caregiver's orientation was characterized as being close, far, or behind the child (independent of distance).
-For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child.
+The child's posture was categorized as being held/carried by the parent, prone (crawling or lying on back or front), sitting, or standing.
+The caregiver's orientation was characterized as being far away from the child (out of reaching distance), behind the child (and within reaching distance) or close to the child (within reaching distnace either to the front or the side of the child).
+Activities were coded in 2-3 second windows, meaning that if a child briefly sat for 1 second and then went back to standing, this brief transition was ignored.
 All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012toward].
-Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.
+Times when the child was out of view of the tripod camera or when children's action were highly sporadic  were marked as uncodable and were excluded from both of these these annotations.
+
+### Naming annotations
+Naming events were coded by a trained coder who listened  to the audio from the head-mounted camera video (using the Transcriptions software for MacOS).
+These naming events were coded relative to word onset for each of eight possible objects (ball, cat, brush,  car,  duck, zem, gimo,  tima, manu, and gazzer) and only proper names were coded (e.g., "brush" and zem" were both coded, "the yellow one" was not).
+If a parent used a variant of the object name (e.g., "kitty" instead of "cat"), these naming events were recorded as the standard name for the object.
+Timestamps from these manual codings were automatically synced to the timestamps for each frame in each video.
+Full transcriptions are hosted in the publicly avaliable repository for this manuscript.
 
@@
 The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones.
-We used this algorithm as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual].
-We next tested the performance of two face detectors that both made use of recently developed Convolutional Neural Networks (CNNs) to extract face information.
+We used this algorithm as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual] which are likely to be common in video data from these head-mounted cameras.
+We next tested the performance of two face detectors that both make use of recently developed Convolutional Neural Networks (CNNs) to extract face information.
 The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts.
-For the second algorithm [OpenPose; @cao2017realtime], we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
 
+We used the outputs of the second algorithm [OpenPose; @cao2017realtime] to create proxies for both face and hand detection.
+As OpenPose seeks to estimate a body frame and then derive estimates of the face/hand positions from this body frame, we reasoned that it might do a better job of detection face/hand information in situations when part of the hand/face are not visible given that  the camera does not entirely capture the child's field of view.
+Thus, for face detection we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.
 The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for hands for two reasons.
@@
 Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.
+
 \begin{figure*}
@@
 ### Detector evaluation
-To evaluate face detector performance, we hand-labeled a "gold set" of labeled frames.
-To account for the relatively rare appearance of faces in the dataset, we hand-labeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames.
-Each sample was comprised of an equal number of frames taken from each child's video.
+To evaluate face detector performance, we hand-labeled a "gold set" of labeled frames (using RectLabel software for MacOS).
+To account for the relatively rare appearance of faces in the dataset, we hand-labeled two types of samples: a sample containing a high density of faces as reported by our two new detectors (half detections by MTCNN, half detections by OpenPose) and a random sample from the remaining frames.
+Each sample was comprised of an equal number of frames taken from each child's video, comprising roughly 1000 frames overall (14 frames for each sample type for each of 36 children).
 For wrist detections, the "gold set" was constructed in the same manner, except frames with a high density of wrists came only from detections made by OpenPose.
-Faces were classified as present if at least half of the face was showing; wrists were classified as present if any part of the wrist was showing.
+For hand-annotation, faces were classified as present if at least half of the face was showing; while wrists were classified as present if any part of the wrist was showing.
 Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 1.
@@
 ### READ IN DETECTIONS
-d <- read_csv("../data/consolidateddata/consolidated_data_4detectors_march7.csv") %>%
-  mutate(posture = factor(posture), orientation = factor(orientation))
+d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_march7.csv") %>%
+  mutate(posture = factor(posture), orientation = factor(orientation)) %>%
+  select(-handOP)
 
@@
 # Results
-<!-- First, we report developmental shifts in infants' posture and their orientation relative to their caregiver.
+First, we report developmental shifts in infants' posture and their orientation relative to their caregiver.
 Then, we explore how these changes influence children's visual access to faces and wrists/hands across this developmental time range.
 Finally, we explore how these changes impact the accessibility of faces and wrists/hands during labeling events.
--->
 
@@
 
-```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right) as a function of child's age.
-Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
-detectionsByAge <- dlong %>%
+### extra analyses not included right now
+```{r facesAndHands, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "."}
+facesAndHands <- d %>%
+  mutate(faceAndHand = faceMT & wristOP) %>%
+  filter(!is.na(age.at.test)) %>%
+  filter(!is.na(posture)) %>%
+  group_by(subid, age.at.test, posture) %>%
+  summarise(faceDetect = sum(faceMT), faceAndHand = sum(faceAndHand), propBoth =faceAndHand /  faceDetect) %>%
+  left_join(inc_subs) 
+  # mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video
+
+ggplot(facesAndHands, aes(x = factor(age.grp), y = propBoth, col = posture)) + 
+  geom_boxplot(outlier.size = 0.5) + 
+  labs(y = "Log face size", x = element_blank()) +
+  theme_few() +
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
+  scale_y_continuous(breaks=seq(0,.60,.20))
+
+# 
+# ggplot(facesAndHands, aes(x = age.at.test, y = propBoth, size=len.min)) +
+#   geom_jitter(width = .25) +
+#   geom_smooth(size=1) +
+#   scale_x_continuous(breaks=c(8,12,16)) +
+#   scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
+#   labs(y = "Prop. both faces & hands", x = element_blank()) +
+#   theme_few() + 
+#   theme(legend.position="none") 
+```
+
+```{r faceSizeByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap ="}
+faceSizeByAge <- dlong %>%
   filter(detectorType!="OpenPose-Faces") %>% # for brevity
   filter(detectorType!="ViolaJones-Faces") %>% # low performance
+  filter(detection == TRUE) %>%
   filter(!is.na(age.at.test)) %>%
+  mutate(logFaceSize = log(w*h)) %>%
   group_by(subid, age.at.test, detectorType) %>%
-  summarise(detections = mean(detection)) %>%
+  summarise(logFaceSize = mean(logFaceSize)) %>%
   left_join(inc_subs) %>%
@@
 
-ggplot(detectionsByAge, aes(x = age.at.test, y = detections, color = detectorType, size=len.min)) +
+ggplot(faceSizeByAge, aes(x = age.at.test, y = logFaceSize, color = detectorType, size=len.min)) +
   geom_jitter(width = .25) +
@@
   scale_x_continuous(breaks=c(8,12,16)) +
-  scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
-  labs(y = "Prop. detected", x = element_blank()) +
+  # scale_size_continuous(range=c(.01,2),name = "Video Length (mins)") +
+  labs(y = "Face size", x = element_blank()) +
   theme_few() + 
@@
   facet_grid(~detectorType) 
+```
+
+```{r faceSizeByPosture, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "."}
+faceSizeByPosture <- dlong %>%
+  filter(detectorType=="MTCNN-Faces") %>% # has bounding box
+  filter(detection == TRUE) %>%
+  filter(!is.na(age.at.test)) %>%
+  filter(!is.na(posture)) %>%
+  mutate(logFaceSize = log(w*h)) %>%
+  group_by(subid, age.at.test, posture) %>%
+  summarise(logFaceSize = mean(logFaceSize), meanVert = mean(y)) %>%
+  left_join(inc_subs)
+
+ggplot(faceSizeByPosture, aes(x = factor(age.grp), y = logFaceSize, col = posture)) + 
+  geom_boxplot(outlier.size = 0.5) + 
+  labs(y = "Log face size", x = element_blank()) +
+  theme_few() +
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
+  scale_y_continuous(breaks=seq(0,.60,.20))
+
+```
+
+```{r faceSizeByOrientation}
+
+faceSizeByOrientation <- dlong %>%
+  filter(detectorType=="MTCNN-Faces") %>% # has bounding box
+  filter(detection == TRUE) %>%
+  filter(!is.na(age.at.test)) %>%
+  filter(!is.na(orientation)) %>%
+  mutate(logFaceSize = log(w*h)) %>%
+  group_by(subid, age.at.test, orientation) %>%
+  summarise(logFaceSize = mean(logFaceSize), meanVert = mean(y)) %>%
+  left_join(inc_subs)
+
+ggplot(faceSizeByOrientation, aes(x = factor(age.grp), y = logFaceSize, col = orientation)) + 
+  geom_boxplot(outlier.size = 0.5) + 
+  labs(y = "Log face size", x = element_blank()) +
+  theme_few() +
+  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
+  scale_y_continuous(breaks=seq(0,.60,.20))
 ```
@@
 
-```{r echo=FALSE, include=FALSE }
-source("helper.R")
+```{r}
+#### Face detections during naming events vs. baseline by posture
 
@@
   group_by(posture) %>%
-  multi_boot_standard(col = "propFacesDetected")
+  multi_boot_standard(col = "propFacesDetected")  %>%
+  mutate(event_type = "naming")
 
+Posture_Faces_Baseline <- d %>%
+  group_by(subid) %>%
+  mutate(detections = faceMT) %>%
+  group_by(subid, posture) %>%
+  summarize(propFacesDetected = mean(detections)) %>%
+  filter(!is.na(propFacesDetected)) %>%
+  filter(!is.na(posture)) %>%
+  group_by(posture) %>%
+  multi_boot_standard(col = "propFacesDetected") %>%
+  mutate(event_type = "baseline")
+
+namingVsBaseline_PostueFaces <- Posture_Faces_Baseline %>%
+  full_join(aroundNamings_Posture_Faces)
+
+ggplot(namingVsBaseline_PostueFaces, aes(x = posture, y = mean, col = event_type)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. faces") +
+  xlab("") +
+  theme_few() +
+  theme(legend.position="left")  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
+```
+
+```{r}
+#### Face detections during naming events vs. baseline by orientation
 aroundNamings_Orient_Faces <- d %>%
@@
   group_by(orientation) %>%
-  multi_boot_standard(col = "propFacesDetected")
+  multi_boot_standard(col = "propFacesDetected") %>%
+  mutate(event_type = "naming")
+
+Orient_Faces_Baseline <- d %>%
+  group_by(subid) %>%
+  mutate(detections = faceMT) %>%
+  group_by(subid, orientation) %>%
+  summarize(propFacesDetected = mean(detections)) %>%
+  filter(!is.na(propFacesDetected)) %>%
+  filter(!is.na(orientation)) %>%
+  group_by(orientation) %>%
+  multi_boot_standard(col = "propFacesDetected") %>%
+  mutate(event_type = "baseline")
 
 ###
+namingVsBaseline_Orientation_Faces <- Orient_Faces_Baseline %>%
+  full_join(aroundNamings_Orient_Faces)
+
+ggplot(namingVsBaseline_Orientation_Faces, aes(x = orientation, y = mean, col = event_type)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. faces") +
+  xlab("") +
+  theme_few() +
+  theme(legend.position="left")  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
+
+```
+
+```{r}
+#### Wrist detections during naming events vs. baseline by posture
 aroundNamings_Posture_Wrists <- d %>%
@@
   group_by(posture) %>%
-  multi_boot_standard(col = "propWristsDetected")
+  multi_boot_standard(col = "propWristsDetected")  %>%
+  mutate(event_type = "naming")
 
+Posture_Wrists_Baseline <- d %>%
+  group_by(subid) %>%
+  mutate(detections = wristOP) %>%
+  group_by(subid, posture) %>%
+  summarize(propWristsDetected = mean(detections)) %>%
+  filter(!is.na(propWristsDetected)) %>%
+  filter(!is.na(posture)) %>%
+  group_by(posture) %>%
+  multi_boot_standard(col = "propWristsDetected") %>%
+  mutate(event_type = "baseline")
+
+namingVsBaseline_Posture_Wrists <- Posture_Wrists_Baseline %>%
+  full_join(aroundNamings_Posture_Wrists)
+
+ggplot(namingVsBaseline_Posture_Wrists, aes(x = posture, y = mean, col = event_type)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. wrists") +
+  xlab("") +
+  theme_few() +
+  # theme(legend.position="left")  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
+```
+
+```{r namingVBasline_Orientation_Wrists}
+#### Wrist detections during naming events vs. baseline by orientation
 aroundNamings_Orient_Wrists <- d %>%
@@
   group_by(orientation) %>%
-  multi_boot_standard(col = "propWristsDetected") 
+  multi_boot_standard(col = "propWristsDetected") %>%
+  mutate(event_type = "naming")
+
+Orient_Wrists_Baseline <- d %>%
+  group_by(subid) %>%
+  mutate(detections = wristOP) %>%
+  group_by(subid, orientation) %>%
+  summarize(propWristsDetected = mean(detections)) %>%
+  filter(!is.na(propWristsDetected)) %>%
+  filter(!is.na(orientation)) %>%
+  group_by(orientation) %>%
+  multi_boot_standard(col = "propWristsDetected") %>%
+  mutate(event_type = "baseline")
+
+namingVsBaseline_Orient_Wrists <- Orient_Wrists_Baseline %>%
+  full_join(aroundNamings_Orient_Wrists)
+
+p2=ggplot(namingVsBaseline_Orient_Wrists, aes(x = orientation, y = mean, col = event_type)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  ylab("Prop. wrists") +
+  xlab("") +
+  theme_few() +
+  theme(legend.position="left")  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_y_continuous(breaks=seq(0,.20,.10), limits=c(0, .25))
+
 ```

### 2020-04-10 writeup-paper/old-drafts/draft_march2020.Rmd
add writeup folder for apa draft

--- 2018-11-06 writeup-paper/cogsci2018_cameraReady.Rmd
@@
 We find that infants' posture and orientation to their caregiver modulates their access to social information, confirming previous work that suggests motoric developments play a significant role in the emergence of children's linguistic and social capacities.
-We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience."
+We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience as they change over development."
 
@@
 library(assertthat)
+```
 
-```
 # Introduction
-From early infancy, children are deeply engaged in learning from others.
-Even newborns tend to prefer to look at faces with direct vs.
-averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010development].
-Further, when infants attend to video stimuli, they tend to look mostly at faces at the expense of other visual information -- though older infants start to look towards people's hands and the actions they are performing [@frank2012measuring].
+Children are remarkably skilled language learners, connecting arbitrary labels (“cup”) with specific visual concepts at a rapid pace through the first two years of life.
+Yet children do not learn words in a vacuum but rather in a rich social environment, where social cues provided by speakers (e.g., eye-gaze) provide strong scaffolding for this learning process.
+Indeed, some empirical work suggests children’s ability to effectively process these social cues may be a key factor in their early language development.
+In a longitudinal study, children’s level of joint engagement with their mother was found to predict both their receptive and productive vocabularies [@carpenter1998].
+Furthermore, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months [@brooks2005] and throughout the second year of life [@brooks2008].
 
-Then, however, their view of the world radically changes [@adolph2006motor].
-Infants' motor abilities improve dramatically near the end of the first year of life, allowing them to locomote independently.
+Around their first birthday, however, their view of the world radically changes [@adolph2006motor] as their motor abilities improve dramatically, allowing them to locomote independently for the first time.
 These motor changes have significant consequences for what children see.
@@
 ### READ IN DETECTIONS
-d <- read_csv("../data/consolidateddata/consolidated_data_4detectors_march7.csv") %>%
+d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_march7.csv") %>%
   mutate(posture = factor(posture), orientation = factor(orientation))

### 2020-04-10 writeup-paper/old-drafts/step1_merge_data 2.Rmd
add writeup folder for apa draft

--- 2018-11-06 writeup-paper/step1_merge_data.Rmd
@@
 ```
+### Recheck posture syncing times
+```{r}
+sync_times_check <- read_csv("../data/manual_sync_times/video_sync_times.csv") 
+s <- sync_times_check %>%
+  rename(subid = sid) %>%
+  group_by(subid) %>%
+  mutate(sync_time_string =toString(sync_time_stamp)) %>%
+  mutate(frames = str_split_fixed(sync_time_string,":",3)[,3]) %>%
+  mutate(sec = str_split_fixed(sync_time_string,":",3)[,2]) %>%
+  mutate(min = str_split_fixed(sync_time_string,":",3)[,1]) %>%
+  mutate(sync_time_ms = as.numeric(min)*6000 + as.numeric(sec)*1000 + as.numeric(frames)*33.33)
+
+posture.sync <- function(x) {
+  fname <- paste("../data/posture/",
+                 x$subid[1],
+                 ".csv",
+                 sep="")
+   if (file.exists(fname)) {
+
+    postures <- read_csv(fname)
+    postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order -- the case for some kids
+    x$posture_start_time = postures$start[1]
+    print(paste0("**** ", x$subid[1], " has time",x$posture_start_time))
+   } else {
+     print(paste0("**** ", x$subid[1], " is missing posture ****"))
+   }
+  return(x)
+}
+
+s <- s %>%
+  group_by(subid) %>%
+  do(posture.sync(.))
+
+## problematic kids
+issues = (s$posture_start_time<s$sync_time_ms)
+diffs =  s$posture_start_time-s$sync_time_ms
+off_by = diffs[issues]
+seconds_off = round(off_by / 1000*-1,2)
+```
 

### 2020-04-10 writeup-paper/old-drafts/step1_merge_data.Rmd
add writeup folder for apa draft

--- 2018-11-06 writeup-paper/step1_merge_data.Rmd
@@
   mutate(faceOP = Nose_conf!=0)  %>%  
-  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
+  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )%>%
+  mutate(armOP = LElbow_conf!=0 | RElbow_conf!=0)
 ```
@@
   group_by(subid) %>%
-  summarise(age.grp = mean(age.grp))
+  summarise(age.grp = age.grp[1], age.at.test = age.at.test[1])  ## added age.at.test
 

### 2020-04-10 writeup-paper/helper.R

@@ -12,7 +12,7 @@ add.times <- function(x) {
     times <- read_csv(fname)
     times$time <- times$best_time ## comes from show_frames python script
     x <- left_join(x, times)
-    x$dt <- c(diff(x$time),.032) 
+    x$dt <- c(diff(x$time),.032) ## adds in 33ms for the last frame but otherwise goes on actual frame times, making an fps rate differences not a problem
   } else {
     print(paste0("**** ", x$subid[1], " is missing frames ****"))
     x$time <- NA
@@ -32,7 +32,9 @@ add.posture <- function(x) {
     print(fname)
     postures <- read_csv(fname)
     postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order
-    postures <- regularize.postures(subset(postures,code=="posture"))
+   
+    ## this reorders the postures so that they are "zeroed" relative to the beginning of the headcam videos
+     postures <- regularize.postures(subset(postures,code=="posture"),x$subid[1]) ## needs subid to get real sync time later 
     x$posture <- factor(NA,levels=levels(postures$posture))
     x$orientation <- factor(NA,levels=levels(postures$orientation))
     
@@ -51,8 +53,23 @@ add.posture <- function(x) {
   return(x)
 }
 
+get_sub_sync_time <- function(this_sub_id){
+  sync_times_check_file <- read_csv("../data/manual_sync_times/video_sync_times.csv") 
+  s <- sync_times_check_file %>%
+    filter(sid == this_sub_id) %>%
+    mutate(sync_time_string =toString(sync_time_stamp)) %>%
+    mutate(frames = str_split_fixed(sync_time_string,":",3)[,3]) %>%
+    mutate(sec = str_split_fixed(sync_time_string,":",3)[,2]) %>%
+    mutate(min = str_split_fixed(sync_time_string,":",3)[,1]) %>%
+    mutate(sync_time_ms = as.numeric(min)*6000 + as.numeric(sec)*1000 + as.numeric(frames)*33.33)
+
+  sync_time = s$sync_time_ms
+  return(sync_time)
+}
+
+
 ######## GET POSTURE CODING SORTED OUT #########
-regularize.postures <- function (p) {
+regularize.postures <- function (p,sub_id) {
   original.postures <- c("p","si","st","l","c","w","NIF")
   
   # clean up postures
@@ -79,14 +96,27 @@ regularize.postures <- function (p) {
                         "3"="behind",
                         "5"="other"))
   
[…]

### 2020-04-29 writeup-paper/draft_apa_march2020.Rmd
clean up writing repository

@@
 abstract: |
-  The ability to process social information is a critical component of children's early language and cognitive development.
-  However, as children reach their first birthday, they begin to locomote themselves, dramatically affecting their visual access to this information. 
-  How do these postural and locomotor changes affect children's access to the social information relevant for word-learning? 
-  Here, we explore this question by using head-mounted cameras to record 36 infants' (8-16 months of age) egocentric visual perspective and use computer vision algorithms to estimate the proportion of faces and hands in infants' environments.
-  We find that infants' posture and orientation to their caregiver modulates their access to social information, confirming previous work that suggests motoric developments play a significant role in the emergence of children's linguistic and social capacities.
-  In addition, we applied our automated analysis to the headcam data from a recent paper showing similar effects postre on visual attention (Franchak et al., 2017), finding convergence across methodologies, laboratories, and datasets.
+  The ability to process social information is a critical component of children's early linguistic and cognitive development.
+  However, as children reach their first birthday, they begin to locomote themselves, dramatically changing the way they see the world around them. 
+  How do these postural and locomotor developments affect children's access to the social information relevant for word-learning? 
+  We explored this question by using head-mounted cameras to record the egocentric visual perspective of 36 infants' (at 8, 12, and 16 months of age) during a naturalistic play session with their caregiver.
+To estimate the proportion of faces and hands in the infant view, we used a computer vision algorithm to detect the presence of faces and hands in the entire dataset.
+  We found that infants' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field.
+  We also applied our automated analysis to the egocentric video data from a recent paper documenting how posture changes one-year-olds visual access to faces (Franchak et al., 2017), finding convergence across both methodologies and dataset and confirming previous work that motoric developments play a significant role in the emergence of children’s linguistic and social capacities.
   We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience as they change over development.
@@
 footnotelist      : no
-linenumbers       : yes
+linenumbers       : no
 mask              : no
-draft             : yes
+draft             : no
 
@@
                       fig.path='figs/', echo=FALSE, warning = FALSE, 
-                      cache=TRUE, message = FALSE, sanitize = TRUE)
+                      cache=FALSE, message = FALSE, sanitize = TRUE)
 library(readr)
@@
 library(gridExtra)
-library(ggridges)
 library(assertthat)
@@
 library(papaja)
+
+# base_size in theme_few (changes font size in plots)
+plot_size=16
 ```
@@
 Before the first birthday, children often begin crawling; soon after, they begin to walk independently.
-While not all children crawl [@adolph1998]—or crawl in the same way—children tend to find creative ways of moving (e.g., scooting) or getting others to help them move (e.g cruising) to begin to explore their world on their own [@patrick2012].
-And as independent agents in the world, children are more active in the construction of their own learning environments.
-Yet while toddlers do determine their own input to a much greater degree, they spend much of their time in a world primarily populated by knees
+While not all children crawl [@adolph1998roles]—or crawl in the same way—children tend to find creative ways of moving (e.g., scooting) or getting others to help them move (e.g cruising) to begin to explore their world on their own [@patrick2012developmental].
+As independent agents in the world, children are more active in the construction of their own learning environments--yet spend much of their time in a world primarily populated by knees
 
@@
 In an observational study, @walle2014 also found that children who were able to walk had both higher receptive and productive vocabularies.
-On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access the social information (e.g., faces, gaze) relative to children who are still crawling and sitting.
+On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access the social information (e.g., facial expressions, gaze cues, pointing) relative to children who are still crawling and sitting.
 In other words, the ability of walking infants to access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.
 
-Over the past decade, researchers have started to use egocentric, head-mounted cameras to document the social information that infants and children have access to across early development, and to understand the degree to which there are substantial shifts in their viewpoints that may have downstream developmental consequences [@Yoshida2008].
-Indeed, the infant visual perspective has proved to be both remarkably different than the adult perspective and not easily predicted by our own intuitions [@franchak2011, @yoshida2008; @clerkin2017].
-Over the first two years of life, the infant viewpoint transitions from primarily containing close up view of faces to capturing relatively restricted views of hands paired with the objects they are acting on [@fausey2016]; both computational and empirical work suggest that this restricted viewpoint is more effective for learning about objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017].
+Over the past decade, researchers have started to use egocentric, head-mounted cameras to document the social information that infants and children have access to across early development, and to understand the degree to which there are substantial shifts in their viewpoints that may have downstream developmental consequences [@yoshida2008].
+Indeed, the infant visual perspective has proved to be both remarkably different than the adult perspective and not easily predicted by our own intuitions [@franchak2011, @yoshida2008, @clerkin2017].
+Over the first two years of life, the infant viewpoint seems to transition from primarily containing close up view of faces to capturing relatively restricted views of hands paired with the objects they are acting on [@fausey2016]; both computational and empirical work suggest that this restricted viewpoint may be more effective for learning about objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017].
+
+We hypothesized that children’s postural developments may be partially responsible for some of these changes in the infant perspective.
+During spontaneous play, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011), when they have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).
+More recently, when 12-month-olds [@franchak2017see] participated in a free-play session with their caregivers, their own posture and as well as their caregiver’s posture influenced the proportion of time they spent looking at the faces and bodies of their caregivers [@franchak2017see].
 
 Here, we directly examine the role that postural developments that infants’ experience as they reach their first birthday change the social information in the infant view.
-We hypothesize that children’s postural developments may be partially responsible for some of these changes in the infant perspective.
-During spontaneous play, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011), when they have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).
-More recently, when 12-month-olds [@Franchak2018] participated in a free-play session with their caregivers, their own posture and as well as their caregiver’s posture influenced the proportion of time they spent looking at the faces and bodies of their caregivers [@Franchak2017].
-
-Here, we expand on these previous findings in three key ways.
+We expand on previous findings in three key ways.
 First, using head-mounted cameras during a brief laboratory free-play session, we record the visual experience of three groups of children at 8, 12, and 16 months-of-age, covering a broad range of ages and locomotor abilities around the first birthday; Children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
@@
 Use of this emerging technology allows us to annotate the entirety of the dataset, allowing a more complete picture of the changing infant perspective.
-In addition, we extend this technique to the same dataset used in @Franchak2017, replicating their main pattern of results using this novel methodology and finding convergence between our datasets with different play sessions, populations, and head-mounted cameras.
-Finally, as parents were presented with pairs of novel and familiar objects with their children during the play sessions, we directly examine the availability of social signals during naming events.
+We apply this same automated method to the dataset used in @franchak2017see, validating the generalizability of this technique.
+Third, as parents were presented with pairs of novel and familiar objects with their children during the play sessions, we were able to explore the availability of social signals during naming events.
 
-Thus, our current dataset allows us to analyze changes in the visual access to faces and hands according to children's age, posture, and linguistic input.
-We predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces/hands because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape with greater access to the social information in their environment.
-[language predictions?]
+Thus, the current study allows us to analyze changes in the visual access to faces and hands according to children's age, posture, and linguistic input.
+Broadly, we predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces/hands because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape with greater access to the social information in their environment.
+
+<!-- Second, we predicted that there would be a greater proportion of faces/hands in view around naming events—-broadly construed as any time a parent mentioned a label for one of the objects in the play set -->
 
@@
 ### Participants
-```{r echo=FALSE, include=FALSE}
+```{r echo=FALSE, include=FALSE, demographics}
 all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
@@
 All inclusion decisions were made independent of the results of subsequent analyses.
-Some of these data were also analyzed in @frank2013.
+These data were also analyzed in @frank2013 and @sanchez2018postural.
 
@@
 The second algorithm (based on work by @zhang2016) uses multi-task cascaded convolutional neural networks (MTCNNs) for joint face detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
-We used a Tensorflow implementation of this algorithm avaliable at https://github.com/davidsandberg/facenet.
+We used a Tensorflow implementation of this algorithm available at https://github.com/davidsandberg/facenet.
 
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
+```{r loadDetections, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
 ### READ IN DETECTIONS
@@
 
-```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
+```{r groundTruthSetup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
 ### READ IN GROUND TRUTH and load function
@@
 
-```{r  include=FALSE, echo=FALSE}
+```{r evaluateDetectors, include=FALSE, echo=FALSE}
 ## Evalute face and wrist detectors
@@
 `r round(perfFacesOverall$f[2],2)` OpenPose).
-Although MTCNN and OpenPose performed comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
+MTCNN and OpenPose performed comparably with the random sample, with MTCNN having higher precision on the high density sample, suggesting that OpenPose generated slightly more false positives than MTCNN.
 ViolaJones performed quite poorly relative to the other detectors, especially with respect to the random sample.
-We thus use MTCNN detections in the following analyses.
 For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision but low recall on the randomly sampled frames (see Table 1).
@@
 \hline
-MTCNN-Faces & High density & 0.89 & 0.92 & \textbf{0.90} \\ 
+MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
 MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
 OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
-OpenPose-Faces & Random & 0.72 & 0.80 & \textbf{0.76} \\ 
+OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
 ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
@@
 \caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected).
-P, R, and F denote precision, recall, and F-score, respectively.
-Scores in bold are the highest F-scores for each sample type.}
+P, R, and F denote precision, recall, and F-score, respectively.}
 \vspace{-1em}
@@
 # Results
-First, we report developmental shifts in infants' posture and their orientation relative to their caregiver, consistent with previous literature [@adolph2006; @franchak2017].
+First, we report developmental shifts in infants' posture and their orientation relative to their caregiver, consistent with previous literature [@adolph2006motor; @franchak2017see].
 Then, we examine how these changes influence children's visual access to faces and wrists/hands across this developmental time range, examining the relative contributions of age vs.
@@
 We also explore how these changes impact the accessibility of faces and wrists/hands during labeling events, both when parents were labeling familiar objects (e.g., 'cat') as well as novel objects (e.g., 'zem').
-Finally, we apply the same automated detection method to a egocentric video dataset collected by a different lab [@franch2017] during which children's in-the-moment posture was annotated, replicating their existing results and finding converging support across methods, labs, and paradigms for the impact of postural developments on children's visual access to social information.
+Finally, we apply the same automated detection method to a egocentric video dataset collected by a different lab [@franchak2017see] during which children's in-the-moment posture was annotated.
 
-```{r}
+```{r calcPostAndOrient}
 ## Calculate distributions of infants spent in each category separately for posture / orientation
@@
 
-```{r}
+```{r setupPosOrientPlot}
 ## Make plots of time spent in each orientation / posture
@@
   xlab("Age (months)") + 
-  theme_few(base_size=12) +
+  theme_few(base_size=plot_size) +
   ylim(0,1) +
@@
 
-```{r posture, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
+```{r posture, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
 cowplot::plot_grid(pos_plot, orient_plot, nrow=2)
@@
 
-```{r}
+```{r setupIndivPosOrientPlot}
 ### Calculate individual amounts of time spent in each orientation by subject
@@
   mutate(pos_orient = paste0('Infant ',posture, ', CG ', orientation)) 
-
 posture_by_orientation <- posture_by_orientation %>%
@@
 
-```{r indivPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
+```{r indivPosOrient, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
 
@@
   ylim(0,1) +
-  theme_few(base_size=18) +
+  theme_few(base_size=plot_size) +
   theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
@@
 ## Changes in Access to Faces and Hands
-We first examined the proportion of face and hand detections as a function of children's age without considering their posture (see Figure \ref{fig:detByAge}).
+We first examined the proportion of face and hand detections as a function of children's age without considering their posture (see Figure \ref{fig:detByAge}); here, we report face and wrist detections using OpenPose, although the same pattern of results for face detections was found using the outputs from MTCNN.
 While faces tended to be in the field-of-view overall more often than hands, children's head-mounted cameras were angled slightly upward to capture the presence of faces, and hand detections suffered from somewhat lower recall than face detections.
 Going forward, we thus analyze differences in the relative proportion of faces or hands in view as a function of age, posture, and orientation, rather than comparing them directly.
-Overall, we observed that 12-month-olds appeared to have visual access to slightly fewer faces than 8 or 16-month-olds, creating a slight U-shaped function in face detections; conversely, hand detections were showed a slight increase across this age range, as reported in prior literature [@Fausey2016].
+Overall, we observed that 12-month-olds appeared to have visual access to slightly fewer faces than 8 or 16-month-olds, creating a slight U-shaped function in face detections; conversely, hand detections were showed a slight increase across this age range, as reported in prior literature [@fausey2016].
 
@@
 When caregivers were behind their children, supporting their children's sitting or standing positions, children saw fewer faces and wrists.
-In particular, children who were sitting or in front of their caregiver had a high proportion of faces and hands in their field of view.
-(Figure \ref{fig:detByPosOrient}, lower panel).
+In particular, children who were sitting or in front of their caregiver had a high proportion of faces and hands in their field of view (Figure \ref{fig:detByPosOrient}).
 
@@
 ## Access to Faces and Hands During Labeling Events
-Our play session was designed to provide parents with opporunities to label objects--both familiar and novel--such that we could examine whether children seek out different kinds of social information around  naming events.
+Our play session was designed to provide parents with opporunities to label objects--both familiar and novel--such that we could examine whether children sought out different kinds of social information around naming events.
 We thus explored how face and hand detections changed during object labeling events, analyzing a four-second window around each labeling event (e.g., "Look at the [zem]!").
@@
 
-```{r eval=FALSE, include=FALSE}
+```{r runModels, eval=FALSE, include=FALSE}
 #wrangle data
@@
   group_by(subid, posture, orientation) %>%
-  summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1]) %>%
+  summarize(countFaces=sum(faceOP==TRUE), countNotFaces = sum(faceOP==FALSE), age.at.test=age.at.test[1]) %>%
   ungroup()
@@
 
-## output ---- no interaction
-# cleanedNames_NoInt = c("Intercept","Close","Far","Sit","Stand","Age") 
+## output ---- 
+cleanedNames_Int = c("Intercept","Sit","Stand","Close","Far","Age (Scaled)", "Sit*Close", "Stand*Close","Sit*Far","Stand*Far")
+
 # 
-# faceGLMOut=(summary(faceGLM))
-# rownames(faceGLMOut$coefficients)<-cleanedNames_NoInt
+faceGLMOut=(summary(faceGLM))
+rownames(faceGLMOut$coefficients)<-cleanedNames_Int
 # 
-# wristGLMOut=(summary(wristGLM))
-# rownames(wristGLMOut$coefficients)<-cleanedNames_NoInt
+wristGLMOut=(summary(wristGLM))
+rownames(wristGLMOut$coefficients)<-cleanedNames_Int
 #
-# xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed model predicting the proportion of faces vs. wrists seen by infants.") 
-# xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3))
+
+### 
+xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.")
+
+xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")
 
@@
 
-```{r}
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\
+  \hline
+Intercept & -3.35 & 0.17 & -19.49 & 0.000 \\
+  Sit & 0.31 & 0.20 & 1.58 & 0.114 \\
+  Stand & -0.02 & 0.20 & -0.12 & 0.905 \\
+  Close & 0.03 & 0.19 & 0.15 & 0.882 \\
+  Far & 0.46 & 0.24 & 1.93 & 0.053 \\
+  Age (Scaled) & 0.14 & 0.11 & 1.22 & 0.223 \\
+  Sit*Close & 0.92 & 0.07 & 13.31 & 0.000 \\
+  Stand*Close & 1.30 & 0.08 & 16.02 & 0.000 \\
+  Sit*Far & 0.54 & 0.07 & 7.37 & 0.000 \\
+  Stand*Far & 1.24 & 0.09 & 14.18 & 0.000 \\
+   \hline
+\end{tabular}
+\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants (as detected by OpenPose).}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\
+  \hline
+Intercept & -4.28 & 0.23 & -18.67 & 0.000 \\
+  Sit & 0.54 & 0.19 & 2.90 & 0.004 \\
+  Stand & 0.61 & 0.23 & 2.63 & 0.009 \\
+  Close & 0.67 & 0.22 & 3.05 & 0.002 \\
+  Far & 0.33 & 0.24 & 1.35 & 0.176 \\
+  Age (Scaled) & 0.45 & 0.13 & 3.35 & 0.001 \\
+  Sit*Close & 0.34 & 0.09 & 3.93 & 0.000 \\
+  Stand*Close & 0.73 & 0.09 & 7.92 & 0.000 \\
+  Sit*Far & -0.15 & 0.11 & -1.39 & 0.165 \\
+  Stand*Far & 1.32 & 0.11 & 11.70 & 0.000 \\
+   \hline
+\end{tabular}
+\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants (as detected by OpenPose).}
+\end{table}
+
+```{r setupdetsByAgePlot}
 detectionsByAge <- dlong %>%
@@
 
-```{r detByAge, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=3, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age.
+```{r detByAge, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=3, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age.
 Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
@@
 ggplot(detectionsByAge, aes(x = age.at.test, y = prop_detected, color = detectorType, size=num_detections)) +
-  theme_few(base_size=12) + 
+  theme_few(base_size=plot_size) + 
   geom_point(alpha=.5) +
@@
   scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
-  scale_size_continuous(name = "Number of frames") +
+  scale_size_continuous(name = "Frames") +
   labs(y = "Proportion detected", x = element_blank()) +
-  scale_colour_discrete(name = element_blank()) +
+  # scale_colour_discrete(name = element_blank()) +
   theme(legend.position="none") +
+  ggthemes::scale_color_solarized(name = "") + 
   facet_grid(~detectorType)
@@
 
-```{r}
+```{r setupdetByPosOrientPlot}
 ### Make dataframes for main plot (detections by orientation & posture)
@@
 
-```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation.
+```{r detByPosOrient, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation.
 Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted.
@@
   labs(y = "Proportion detections", x = "Age (months)") +
-  theme_few(base_size=16) +
+  theme_few(base_size=plot_size) +
   theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 12)) +
-  scale_colour_discrete(name = 'Detection type') +
-  scale_size_continuous(name = 'Number of frames') +
+  ggthemes::scale_color_solarized(name = "") + 
+  scale_size_continuous(name = 'Frames') +
   scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
@@
 
-```{r}
+```{r summarizeNamings}
 source("helper.R")
@@
 
-```{r}
+```{r afterNamingsExplore, eval=FALSE, include=FALSE, echo=FALSE}
+afterNamings_Faces <- d %>%
+  group_by(subid) %>%
+  mutate(detections = faceOP) %>%
+  do(summarize.naming((.),window = c(0,4)))
+
+detectionsByGroup <- dlongPlot %>%
+  group_by(age.grp, subid, detectorType) %>%
+  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
+  group_by(age.grp, detectorType) %>%
+  multi_boot_standard(col = 'prop_detected')
+
+afterNamings_Detections_BySub_Faces <- afterNamings_Faces %>%
+  group_by(subid, familiarity, age.grp) %>%
+  summarize(prop_detected = mean(detections, na.rm=TRUE))
+
+afterNamings_Detections_Grouped_Faces <- afterNamings_Detections_BySub_Faces %>%
+  group_by(familiarity, age.grp) %>%
+  multi_boot_standard(col = 'prop_detected')
+
+faces_afternamings_plot <- ggplot(afterNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
+  # plot baseline
+  geom_pointrange(data = (detectionsByGroup %>% filter(detectorType == 'OpenPose-Faces')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=3), alpha=.9) +
+  # plot averages by familiarity 
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
+  # plot familairity detections for each sub
+  geom_point(data=afterNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
+  theme_few() +
+  labs(y = "Prop. face detections", x = "Age (months)") +
+  theme_few(base_size=plot_size) +
+  ylim(0,.6) +
+  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
+  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
+  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
+  scale_size_continuous(name="# of naming events")
+
+```
+
+```{r setUpNamingsPlot}
 ### Analyze detections around naming events (without posture/orientation)
@@
 
-```{r}
+```{r detByNamingPlotPanels}
 ## Plot faces/wrists around namings by age
@@
   labs(y = "Prop. face detections", x = "Age (months)") +
-  theme_few(base_size=16) +
+  theme_few(base_size=plot_size) +
   ylim(0,.6) +
@@
   labs(y = "Prop. wrist detections", x = "Age (months)") +
-  theme_few(base_size=16) +
+  theme_few(base_size=plot_size) +
   ylim(0,.6) +
@@
 
-```{r detByNaming, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\\% bootstrapped confidence intervals."}
+```{r detByNaming, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\\% bootstrapped confidence intervals."}
 cowplot::plot_grid(faces_aroundnamings_plot, wrists_aroundnamings_plot)
@@
 
-# Extension to Franchak et al., 2017
+## Extension to Franchak et al., 2017
 In the present work, we found that infant's in-the-moment posture changed with their age, as did infants' orientation relative their caregiver.
-In a related study with 12-month-olds, @franchak2017 also that children's in-the-moment posture changed the amount of time infants spent looking at faces.
-Here, we sought to replicate the findings from @franchak2017 using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras (shared on Databrary, @Databrary).
+In a related study with 12-month-olds, @franchak2017see also that children's in-the-moment posture changed the amount of time infants spent looking at faces.
+Here, we sought to replicate the findings from @franchak2017see using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras [@simon2015databrary].
 
 This dataset differed from the present in two key ways that could present a challenge for our automated methodology.
-First, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to explore, unlike the present room  dataset which was relatively small (approximately 10' x 10').
+First, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to explore, unlike the  room used for the present dataset which was relatively small (approximately 10' x 10').
 Thus, our automated methodology could fail to generalize to scenes from these more complex environments, where detecting faces and hands could arguably be a much harder task.
-Second, @franchak2017 used a head-mounted eye-tracker, finding that children's posture affected where children were looking within their visual field.
-Nonetheless, if children are often orienting their heads towards where they are allocating their attention  [@Yoshida20008] then we should expect to find the same pattern of results only when analyzing the information in view.
+Second, @franchak2017see used a head-mounted eye-tracker, finding that children's posture affected where children were looking within their visual field.
+Nonetheless, if children are often orienting their heads towards where they are allocating their attention [@yoshida2008] then we should expect to find the same pattern of results only when analyzing the information in view.
 
-```{r}
+```{r loadFranchakData}
 # loads pos_detections used below for franchak dataset
@@
 
-```{r}
+```{r processFranchak}
 faces_by_pos  <- pos_detections %>%
@@
 
-```{r}
+```{r statsFranchak}
 # glmers for franchak dataset
@@
 
-```{r}
+```{r plotPanelsFranchak}
 faces_by_pos_plot <- ggplot(faces_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
@@
   geom_line(data=faces_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label, group=sub_id), alpha=.1) +
-  theme_few(base_size=18) +
+  theme_few(base_size=plot_size) +
   ylab('Prop faces detected') +
@@
   geom_line(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_hands, col=infant_posture_label, group=sub_id), alpha=.1) +
-  theme_few(base_size=18) +
+  theme_few(base_size=plot_size) +
   ylab('Prop hands detected') +
@@
 
-```{r franchak, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture.
+```{r franchak, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture.
 Error bars represent 95 percent bootstrapped confidence intervals."}
@@
 
-Overall, we found convergence between our two methodologies: we replicated @Franchak2017's main results, finding that the proportion of faces in view was greater when infants were sitting or standing vs.
+Overall, we found convergence between our two methodologies: we replicated the main results from @franchak2017see, finding that the proportion of faces in view was greater when infants were sitting or standing vs.
 prone.
-We also found the same patterns of results with hands (i.e., wrist detectinos), which were not originally annotated in @Franch2017 (see Figure \ref{fig:franchak}); results were validated with generalized mixed effect models as in our previous analysis.
-Interestingly, there was a relatively higher proportion of hands in view in @Franchak2017 relative to the present dataset; this is likely mostly due to the fact that there were often multiple people in view: experimenters stayed in the room to film children and assist with technical difficulties, while in our present dataset the experiments left the room for the majority of the play session.
-Nonetheless, we still found that posture modulated the hands in view as well, suggesting that this is a major factor that structures infants' access to visual information.
+We also found the same patterns of results with hands (i.e., wrist detections), though these were not originally annotated in @franchak2017see (see Figure \ref{fig:franchak}); results were validated with generalized mixed effect models as in our previous analysis.
+Interestingly, there was a relatively higher proportion of hands in view in @franchak2017see relative to the present dataset; this is likely mostly due to the fact that there were often multiple people in view: experimenters stayed in the room to film children and assist the caregiver, while in our present dataset the experimenters left the room for the majority of the play session.
+Nonetheless, we still found that posture modulated the proportion of hands in view, suggesting that this is a major factor that structures infants' access to visual information.
 
 # General Discussion
-We used a head-mounted camera to explore how children's postural and locomotor development affects their visual access to social information, here operationalized as the presence of the faces and hands of their caregiver.
-Children's posture and orientation towards their caregiver changed systematically across age, and both of these factors influenced the proportion of faces and hands in the child's visual field.
-Though our cameras were oriented to capture face information, we nonetheless saw an influence of orientation and posture on visual access to hands.
+How do postural and locomotive developments influence the social information that infants see?
+We used a head-mounted camera to explore how these emerging motoric abilities affect children's visual access to social information, here operationalized as the presence of the faces and hands of their caregiver.
+First, we found systematic changes around the first birthday in children's in-the-moment posture and their orientation relative to the caregivers; older children spent more time standing and less time sitting; older children's caregivers spent less time supporting their standing or sitting postures.
+Children's changing posture and orientation to their caregiver jointly shaped the amount of social information that was in their view during one-on-one play sessions with their caregivers: Children saw the most faces/hands when they were sitting or standing and close to their caregiver.
 Motor development appears to modulate how infants experience their visual world and the social information in it.
 
-We created a situation in which the context of interaction was tightly controlled, but as children grow and change, the activities in which they engage with their caregivers also vary, leading to differences in the distribution of contexts they experience.
-Thus, more work is needed to understand how our results relate to children's home experiences [@fausey2016].
-The ability to walk is part of a cascade of changes in children's  experience, and our study captures only a cross-sectional slice of this broader, multifaceted trajectory.
+However, we did not see differences in visual access to social information during naming events, suggesting that children did not change where they were looking when hearing a label for either a novel or familiar object.
+We see several potential explanations for this: first, other work [@yoshida2008; @yu2013joint; @yu2017hand], including @franchak2017see, has found that infants spend much more time looking at the toys vs.
+their caregiver's faces during these play sessions; infants may have thus been primarily interested in exploring these new toys rather than learning their names.
+Furthermore, moving their own neck upwards towards their caregiver still requires quite a physical effort at this age (especially when prone or sitting) and simply may not have been a priority for children in this context.
+A second factor is that this particular play session did not present many opportunities where children would need to use social cues to disambiguate referents.
+Indeed, there were only two possible referents in the room at a time—-and one of them was always a familiar category (i.e., car, kitty).
 
-Understanding these changes has been a persistent challenge for developmental psychology, but the field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools.
-These tools deal better with noisier, more complicated datasets and extract richer information than previous systems.
-We hope that these new tools can be leveraged to understand the changing infant perspective on the visual world and the implications of these changes for linguistic, cognitive, and social development.
+Broadly,these results show promise for the use of automated methods to detect the social information in the egocentric infant viewpoint during more naturalistic parent-child interactons.
+The use of this automated methodology allowed us to easily annotate the entirety of the dataset and analyze all of the frames without the need for hundreds of hours of human coding.
+Importantly, we also applied these same methods to an addiontal dataset that contained more variation in the type and structure of the play session, replicating and extending the effects of children's in-the-moment posture on visual access to social information [@franchak2017see].
+
+Yet more work is needed to understand how these results relate to children's home experiences [@fausey2016].
+Both this play session and that of @franchak2017see were relatively controlled interactions where caregivers are highly aware that they are being montiored by experimenters.
+Furthermore, as children grow and change, the activities in which they engage with their caregivers are likely to also vary, leading to differences in the distribution of contexts they experience that may not be captured in these one-on-one play sessions.
+Finally, the ability to walk is of course only of a cascade of changes in children's abilities and experiences, and this study captures only a cross-sectional slice of this broader, multifaceted trajectory.
+
+Understanding these changes and how they relate to one another has been a persistent challenge for developmental psychology, but the field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools.
+These tools deal better with noisier, more complicated datasets and extract richer information than previous system.
+We hope that these new tools can be leveraged to understand the changing infant perspective on the visual world, both in controlled experimental contexts and in children's home environments.
+By doing so, we believe that we can understand the implications of the changing infant perspective consequences for linguistic, cognitive, and social development.
 
@@
 This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and NSF #1714726 to BLL.
-An earlier version of this work was presented to the Cognitive Science Society in @frank2013 and @sanchez2018.
-
-\vspace{1em} \fbox{\parbox[b][][c]{7.3cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/amsan7/xs-face}}} \vspace{1em}
+An earlier version of this work was presented to the Cognitive Science Society in @frank2013 and @sanchez2018postural.
 
@@
 ```{r create_r-references}
-# r_refs(file = "xsface.bib")
+r_refs(file = "xsface.bib")
 ```

### 2020-04-29 writeup-paper/draft_apa_march2020.tex
clean up writing repository

@@
-\documentclass[english,man,floatsintext,draftall]{apa6}
+\documentclass[english,man,floatsintext]{apa6}
 \usepackage{lmodern}
@@
 \keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: X}
-\usepackage{lineno}
-
-\linenumbers
 \usepackage{csquotes}
@@
 The ability to process social information is a critical component of
-children's early language and cognitive development. However, as
+children's early linguistic and cognitive development. However, as
 children reach their first birthday, they begin to locomote themselves,
-dramatically affecting their visual access to this information. How do
-these postural and locomotor changes affect children's access to the
-social information relevant for word-learning? Here, we explore this
-question by using head-mounted cameras to record 36 infants' (8-16
-months of age) egocentric visual perspective and use computer vision
-algorithms to estimate the proportion of faces and hands in infants'
-environments. We find that infants' posture and orientation to their
-caregiver modulates their access to social information, confirming
-previous work that suggests motoric developments play a significant role
-in the emergence of children's linguistic and social capacities. In
-addition, we applied our automated analysis to the headcam data from a
-recent paper showing similar effects postre on visual attention
-(Franchak et al., 2017), finding convergence across methodologies,
-laboratories, and datasets. We suggest that the combined use of
-head-mounted cameras and the application of new computer vision
-techniques is a promising avenue for understanding the statistics of
-infants' visual and linguistic experience as they change over
-development.
+dramatically changing the way they see the world around them. How do
+these postural and locomotor developments affect children's access to
+the social information relevant for word-learning? We explored this
+question by using head-mounted cameras to record the egocentric visual
+perspective of 36 infants' (at 8, 12, and 16 months of age) during a
+naturalistic play session with their caregiver. To estimate the
+proportion of faces and hands in the infant view, we used a computer
+vision algorithm to detect the presence of faces and hands in the entire
+dataset. We found that infants' posture and orientation to their
+caregiver changed dramatically across this age range and modulated their
+access to this social information; infants who were sitting or standing
+with their caregiver at a close distance tended to have the most
+faces/hands in their visual field. We also applied our automated
+analysis to the egocentric video data from a recent paper documenting
+how posture changes one-year-olds visual access to faces (Franchak et
+al., 2017), finding convergence across both methodologies and dataset
+and confirming previous work that motoric developments play a
+significant role in the emergence of children's linguistic and social
+capacities. We suggest that the combined use of head-mounted cameras and
+the application of new computer vision techniques is a promising avenue
+for understanding the statistics of infants' visual and linguistic
+experience as they change over development.
 }
@@
 experimental context have larger vocabularies at 18 months and
-throughout the second year of life (R. Brooks \& Meltzoff, 2005; Rechele
-Brooks \& Meltzoff, 2008).
+throughout the second year of life (Brooks \& Meltzoff, 2005, 2008).
 
@@
 after, they begin to walk independently. While not all children crawl
-({\textbf{???}})---or crawl in the same way---children tend to find
-creative ways of moving (e.g., scooting) or getting others to help them
-move (e.g cruising) to begin to explore their world on their own
-({\textbf{???}}). And as independent agents in the world, children are
-more active in the construction of their own learning environments. Yet
-while toddlers do determine their own input to a much greater degree,
-they spend much of their time in a world primarily populated by knees
+(Adolph, Vereijken, \& Denny, 1998)---or crawl in the same
+way---children tend to find creative ways of moving (e.g., scooting) or
+getting others to help them move (e.g cruising) to begin to explore
+their world on their own (Patrick, Noah, \& Yang, 2012). As independent
+agents in the world, children are more active in the construction of
+their own learning environments--yet spend much of their time in a world
+primarily populated by knees
 
@@
 independently locomote may fundamentally change their ability to access
-the social information (e.g., faces, gaze) relative to children who are
-still crawling and sitting. In other words, the ability of walking
-infants to access more detailed social information may allow infants to
-learn words quicker and more efficiently, facilitating language growth.
+the social information (e.g., facial expressions, gaze cues, pointing)
+relative to children who are still crawling and sitting. In other words,
+the ability of walking infants to access more detailed social
+information may allow infants to learn words quicker and more
+efficiently, facilitating language growth.
 
@@
 degree to which there are substantial shifts in their viewpoints that
-may have downstream developmental consequences ({\textbf{???}}). Indeed,
-the infant visual perspective has proved to be both remarkably different
-than the adult perspective and not easily predicted by our own
-intuitions (Clerkin, Hart, Rehg, Yu, \& Smith, 2017; Franchak, Kretch,
-Soska, \& Adolph, 2011, Yoshida and Smith (2008)). Over the first two
-years of life, the infant viewpoint transitions from primarily
+may have downstream developmental consequences (Yoshida \& Smith, 2008).
+Indeed, the infant visual perspective has proved to be both remarkably
+different than the adult perspective and not easily predicted by our own
+intuitions (Franchak, Kretch, Soska, \& Adolph, 2011, Yoshida and Smith
+(2008), Clerkin, Hart, Rehg, Yu, and Smith (2017)). Over the first two
+years of life, the infant viewpoint seems to transition from primarily
 containing close up view of faces to capturing relatively restricted
@@
 Jayaraman, \& Smith, 2016); both computational and empirical work
-suggest that this restricted viewpoint is more effective for learning
-about objects and their labels than the comparable adult perspective (D.
-Yurovsky, Smith, \& Yu, in press, Bambach, Crandall, Smith, and Yu
-(2017)).
+suggest that this restricted viewpoint may be more effective for
+learning about objects and their labels than the comparable adult
+perspective (D. Yurovsky, Smith, \& Yu, 2012, Bambach, Crandall, Smith,
+and Yu (2017)).
+
+We hypothesized that children's postural developments may be partially
+responsible for some of these changes in the infant perspective. During
+spontaneous play, toddlers are more likely to look at the floor while
+crawling than while walking (Franchak et al., 2011), when they have full
+visual access to their environment and the people in it (Kretch,
+Franchak, and Adolph, 2014). More recently, when 12-month-olds
+(Franchak, Kretch, \& Adolph, 2017) participated in a free-play session
+with their caregivers, their own posture and as well as their
+caregiver's posture influenced the proportion of time they spent looking
+at the faces and bodies of their caregivers (Franchak et al., 2017).
 
@@
 infants' experience as they reach their first birthday change the social
-information in the infant view. We hypothesize that children's postural
-developments may be partially responsible for some of these changes in
-the infant perspective. During spontaneous play, toddlers are more
-likely to look at the floor while crawling than while walking (Franchak
-et al., 2011), when they have full visual access to their environment
-and the people in it (Kretch, Franchak, and Adolph, 2014). More
-recently, when 12-month-olds ({\textbf{???}}) participated in a
-free-play session with their caregivers, their own posture and as well
-as their caregiver's posture influenced the proportion of time they
-spent looking at the faces and bodies of their caregivers
-({\textbf{???}}).
+information in the infant view. We expand on previous findings in three
+key ways. First, using head-mounted cameras during a brief laboratory
+free-play session, we record the visual experience of three groups of
+children at 8, 12, and 16 months-of-age, covering a broad range of ages
+and locomotor abilities around the first birthday; Children's posture
+and orientation relative to their caregiver were also recorded from a
+third-person perspective and hand-annotated. This cross-sectional design
+thus allows us to directly examine the relative contributions of age
+vs.~postural developments on children's visual access to social
+information. Second, we use a modern computer vision algorithm for the
+automated detection of faces and hands in the egocentric viewpoints of
+infants. In particular, we capitalize on recent improvements in face and
+pose detection algorithms (Cao, Simon, Wei, \& Sheikh, 2017; K. Zhang,
+Zhang, Li, \& Qiao, 2016) to analyze the frequencies of faces and hands
+in the child's visual environment, both overall and relative to naming
+events by their caregivers. Use of this emerging technology allows us to
+annotate the entirety of the dataset, allowing a more complete picture
+of the changing infant perspective. We apply this same automated method
+to the dataset used in Franchak et al. (2017), validating the
+generalizability of this technique. Third, as parents were presented
+with pairs of novel and familiar objects with their children during the
+play sessions, we were able to explore the availability of social
+signals during naming events.
 
-Here, we expand on these previous findings in three key ways. First,
-using head-mounted cameras during a brief laboratory free-play session,
-we record the visual experience of three groups of children at 8, 12,
-and 16 months-of-age, covering a broad range of ages and locomotor
-abilities around the first birthday; Children's posture and orientation
-relative to their caregiver were also recorded from a third-person
-perspective and hand-annotated. This cross-sectional design thus allows
-us to directly examine the relative contributions of age vs.~postural
-developments on children's visual access to social information. Second,
-we use a modern computer vision algorithm for the automated detection of
-faces and hands in the egocentric viewpoints of infants. In particular,
-we capitalize on recent improvements in face and pose detection
-algorithms (Cao, Simon, Wei, \& Sheikh, 2017; K. Zhang, Zhang, Li, \&
-Qiao, 2016) to analyze the frequencies of faces and hands in the child's
-visual environment, both overall and relative to naming events by their
-caregivers. Use of this emerging technology allows us to annotate the
-entirety of the dataset, allowing a more complete picture of the
-changing infant perspective. In addition, we extend this technique to
-the same dataset used in ({\textbf{???}}), replicating their main
-pattern of results using this novel methodology and finding convergence
-between our datasets with different play sessions, populations, and
-head-mounted cameras. Finally, as parents were presented with pairs of
-novel and familiar objects with their children during the play sessions,
-we directly examine the availability of social signals during naming
-events.
-
-Thus, our current dataset allows us to analyze changes in the visual
+Thus, the current study allows us to analyze changes in the visual
 access to faces and hands according to children's age, posture, and
-linguistic input. We predicted that there would be differential access
-to social information based on children's postural developments:
+linguistic input. Broadly, we predicted that there would be differential
+access to social information based on children's postural developments:
 crawling infants would see fewer faces/hands because they would
@@
 access to a richer visual landscape with greater access to the social
-information in their environment. {[}language predictions?{]}
+information in their environment.
 
@@
 inclusion decisions were made independent of the results of subsequent
-analyses. Some of these data were also analyzed in Frank, Simmons,
-Yurovsky, and Pusiol (2013).
+analyses. These data were also analyzed in Frank, Simmons, Yurovsky, and
+Pusiol (2013) and Sanchez, Long, Kraus, and Frank (2018).
 
@@
 real-world environments where varying illuminations and occlusions are
-present. We used a Tensorflow implementation of this algorithm avaliable
+present. We used a Tensorflow implementation of this algorithm available
 at \url{https://github.com/davidsandberg/facenet}.
@@
 For face detection, MTCNN outperformed OpenPose when taking into account
-only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose). Although
-MTCNN and OpenPose performed comparably with the random sample, MTCNN
-performed better on the high density sample (specifically looking at
-precision), suggesting that OpenPose generated more false positives than
-MTCNN. ViolaJones performed quite poorly relative to the other
-detectors, especially with respect to the random sample. We thus use
-MTCNN detections in the following analyses. For wrist detection,
-OpenPose performed moderately well (F = 0.74) with relatively high
-precision but low recall on the randomly sampled frames (see Table 1).
-We thus analyze wrist detections, with the caveat that we are likely
-underestimating the proportion of hands in the dataset.
+only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose). MTCNN and
+OpenPose performed comparably with the random sample, with MTCNN having
+higher precision on the high density sample, suggesting that OpenPose
+generated slightly more false positives than MTCNN. ViolaJones performed
+quite poorly relative to the other detectors, especially with respect to
+the random sample. For wrist detection, OpenPose performed moderately
+well (F = 0.74) with relatively high precision but low recall on the
+randomly sampled frames (see Table 1). We thus analyze wrist detections,
+with the caveat that we are likely underestimating the proportion of
+hands in the dataset.
 
@@
 orientation relative to their caregiver, consistent with previous
-literature ({\textbf{???}}, {\textbf{???}}). Then, we examine how these
-changes influence children's visual access to faces and wrists/hands
-across this developmental time range, examining the relative
-contributions of age vs.~postural developments. We also explore how
-these changes impact the accessibility of faces and wrists/hands during
-labeling events, both when parents were labeling familiar objects (e.g.,
-\enquote{cat}) as well as novel objects (e.g., \enquote{zem}). Finally,
-we apply the same automated detection method to a egocentric video
-dataset collected by a different lab ({\textbf{???}}) during which
-children's in-the-moment posture was annotated, replicating their
-existing results and finding converging support across methods, labs,
-and paradigms for the impact of postural developments on children's
-visual access to social information.
+literature (Adolph \& Berger, 2006; Franchak et al., 2017). Then, we
+examine how these changes influence children's visual access to faces
+and wrists/hands across this developmental time range, examining the
+relative contributions of age vs.~postural developments. We also explore
+how these changes impact the accessibility of faces and wrists/hands
+during labeling events, both when parents were labeling familiar objects
+(e.g., \enquote{cat}) as well as novel objects (e.g., \enquote{zem}).
+Finally, we apply the same automated detection method to a egocentric
+video dataset collected by a different lab (Franchak et al., 2017)
+during which children's in-the-moment posture was annotated.
 
-\begin{figure*}[h]
+\begin{figure*}[!h]
 
-{\centering \includegraphics{figs/posture-1}
+{\centering \includegraphics[width=\textwidth]{figs/posture-1}
 
@@
 
-\begin{figure*}[h]
+\begin{figure*}[!h]
 
-{\centering \includegraphics{figs/indivPosOrient-1}
+{\centering \includegraphics[width=\textwidth]{figs/indivPosOrient-1}
 
@@
 function of children's age without considering their posture (see Figure
-\ref{fig:detByAge}). While faces tended to be in the field-of-view
-overall more often than hands, children's head-mounted cameras were
-angled slightly upward to capture the presence of faces, and hand
-detections suffered from somewhat lower recall than face detections.
-Going forward, we thus analyze differences in the relative proportion of
-faces or hands in view as a function of age, posture, and orientation,
-rather than comparing them directly. Overall, we observed that
-12-month-olds appeared to have visual access to slightly fewer faces
-than 8 or 16-month-olds, creating a slight U-shaped function in face
-detections; conversely, hand detections were showed a slight increase
-across this age range, as reported in prior literature ({\textbf{???}}).
+\ref{fig:detByAge}); here, we report face and wrist detections using
+OpenPose, although the same pattern of results for face detections was
+found using the outputs from MTCNN. While faces tended to be in the
+field-of-view overall more often than hands, children's head-mounted
+cameras were angled slightly upward to capture the presence of faces,
+and hand detections suffered from somewhat lower recall than face
+detections. Going forward, we thus analyze differences in the relative
+proportion of faces or hands in view as a function of age, posture, and
+orientation, rather than comparing them directly. Overall, we observed
+that 12-month-olds appeared to have visual access to slightly fewer
+faces than 8 or 16-month-olds, creating a slight U-shaped function in
+face detections; conversely, hand detections were showed a slight
+increase across this age range, as reported in prior literature (Fausey
+et al., 2016).
 
@@
 particular, children who were sitting or in front of their caregiver had
-a high proportion of faces and hands in their field of view. (Figure
-\ref{fig:detByPosOrient}, lower panel).
+a high proportion of faces and hands in their field of view (Figure
+\ref{fig:detByPosOrient}).
 
@@
 label objects--both familiar and novel--such that we could examine
-whether children seek out different kinds of social information around
+whether children sought out different kinds of social information around
 naming events. We thus explored how face and hand detections changed
@@
 
-\begin{figure*}[h]
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\
+  \hline
+Intercept & -3.35 & 0.17 & -19.49 & 0.000 \\
+  Sit & 0.31 & 0.20 & 1.58 & 0.114 \\
+  Stand & -0.02 & 0.20 & -0.12 & 0.905 \\
+  Close & 0.03 & 0.19 & 0.15 & 0.882 \\
+  Far & 0.46 & 0.24 & 1.93 & 0.053 \\
+  Age (Scaled) & 0.14 & 0.11 & 1.22 & 0.223 \\
+  Sit*Close & 0.92 & 0.07 & 13.31 & 0.000 \\
+  Stand*Close & 1.30 & 0.08 & 16.02 & 0.000 \\
+  Sit*Far & 0.54 & 0.07 & 7.37 & 0.000 \\
+  Stand*Far & 1.24 & 0.09 & 14.18 & 0.000 \\
+   \hline
+\end{tabular}
+\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants (as detected by OpenPose).}
+\end{table}
 
-{\centering \includegraphics{figs/detByAge-1}
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\
+  \hline
+Intercept & -4.28 & 0.23 & -18.67 & 0.000 \\
+  Sit & 0.54 & 0.19 & 2.90 & 0.004 \\
+  Stand & 0.61 & 0.23 & 2.63 & 0.009 \\
+  Close & 0.67 & 0.22 & 3.05 & 0.002 \\
+  Far & 0.33 & 0.24 & 1.35 & 0.176 \\
+  Age (Scaled) & 0.45 & 0.13 & 3.35 & 0.001 \\
+  Sit*Close & 0.34 & 0.09 & 3.93 & 0.000 \\
+  Stand*Close & 0.73 & 0.09 & 7.92 & 0.000 \\
+  Sit*Far & -0.15 & 0.11 & -1.39 & 0.165 \\
+  Stand*Far & 1.32 & 0.11 & 11.70 & 0.000 \\
+   \hline
+\end{tabular}
+\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants (as detected by OpenPose).}
+\end{table}
+
+\begin{figure*}[!h]
+
+{\centering \includegraphics[width=\textwidth]{figs/detByAge-1}
 
@@
 
-\begin{figure*}[h]
+\begin{figure*}[!h]
 
-{\centering \includegraphics{figs/detByPosOrient-1}
+{\centering \includegraphics[width=\textwidth]{figs/detByPosOrient-1}
 
@@
 
-\begin{figure*}[h]
+\begin{figure*}[!h]
 
-{\centering \includegraphics{figs/detByNaming-1}
+{\centering \includegraphics[width=\textwidth]{figs/detByNaming-1}
 
@@
 
-\section{Extension to Franchak et al.,
+\subsection{Extension to Franchak et al.,
 2017}\label{extension-to-franchak-et-al.-2017}
@@
 changed with their age, as did infants' orientation relative their
-caregiver. In a related study with 12-month-olds, ({\textbf{???}}) also
-that children's in-the-moment posture changed the amount of time infants
-spent looking at faces. Here, we sought to replicate the findings from
-({\textbf{???}}) using our automated methodology (OpenPose detections),
-using the footage from their head-mounted cameras (shared on Databrary,
-({\textbf{???}})).
+caregiver. In a related study with 12-month-olds, Franchak et al. (2017)
+also that children's in-the-moment posture changed the amount of time
+infants spent looking at faces. Here, we sought to replicate the
+findings from Franchak et al. (2017) using our automated methodology
+(OpenPose detections), using the footage from their head-mounted cameras
+(D. A. Simon, Gordon, Steiger, \& Gilmore, 2015).
 
@@
 in the present dataset, containing multiple structures and toys in
-different parts of the room for infants to explore, unlike the present
-room dataset which was relatively small (approximately 10' x 10'). Thus,
-our automated methodology could fail to generalize to scenes from these
-more complex environments, where detecting faces and hands could
-arguably be a much harder task. Second, ({\textbf{???}}) used a
-head-mounted eye-tracker, finding that children's posture affected where
-children were looking within their visual field. Nonetheless, if
-children are often orienting their heads towards where they are
-allocating their attention ({\textbf{???}}) then we should expect to
-find the same pattern of results only when analyzing the information in
-view.
+different parts of the room for infants to explore, unlike the room used
+for the present dataset which was relatively small (approximately 10' x
+10'). Thus, our automated methodology could fail to generalize to scenes
+from these more complex environments, where detecting faces and hands
+could arguably be a much harder task. Second, Franchak et al. (2017)
+used a head-mounted eye-tracker, finding that children's posture
+affected where children were looking within their visual field.
+Nonetheless, if children are often orienting their heads towards where
+they are allocating their attention (Yoshida \& Smith, 2008) then we
+should expect to find the same pattern of results only when analyzing
+the information in view.
 
-\begin{figure*}[h]
+\begin{figure*}[!h]
 
-{\centering \includegraphics{figs/franchak-1}
+{\centering \includegraphics[width=\textwidth]{figs/franchak-1}
 
@@
 Overall, we found convergence between our two methodologies: we
-replicated ({\textbf{???}})\enquote{s main results, finding that the
-proportion of faces in view was greater when infants were sitting or
+replicated the main results from Franchak et al. (2017), finding that
+the proportion of faces in view was greater when infants were sitting or
 standing vs.~prone. We also found the same patterns of results with
-hands (i.e., wrist detectinos), which were not originally annotated in
-({\textbf{???}}) (see Figure \ref{fig:franchak}); results were validated
-with generalized mixed effect models as in our previous analysis.
-Interestingly, there was a relatively higher proportion of hands in view
-in ({\textbf{???}}) relative to the present dataset; this is likely
-mostly due to the fact that there were often multiple people in view:
-experimenters stayed in the room to film children and assist with
-technical difficulties, while in our present dataset the experiments
-left the room for the majority of the play session. Nonetheless, we
-still found that posture modulated the hands in view as well, suggesting
-that this is a major factor that structures infants} access to visual
-information.
+hands (i.e., wrist detections), though these were not originally
+annotated in Franchak et al. (2017) (see Figure \ref{fig:franchak});
+results were validated with generalized mixed effect models as in our
+previous analysis. Interestingly, there was a relatively higher
+proportion of hands in view in Franchak et al. (2017) relative to the
+present dataset; this is likely mostly due to the fact that there were
+often multiple people in view: experimenters stayed in the room to film
+children and assist the caregiver, while in our present dataset the
+experimenters left the room for the majority of the play session.
+Nonetheless, we still found that posture modulated the proportion of
+hands in view, suggesting that this is a major factor that structures
+infants' access to visual information.
 
-\section{Discussion}\label{discussion}
+\section{General Discussion}\label{general-discussion}
 
-\section{Discussion}\label{discussion-1}
+How do postural and locomotive developments influence the social
+information that infants see? We used a head-mounted camera to explore
+how these emerging motoric abilities affect children's visual access to
+social information, here operationalized as the presence of the faces
+and hands of their caregiver. First, we found systematic changes around
+the first birthday in children's in-the-moment posture and their
+orientation relative to the caregivers; older children spent more time
+standing and less time sitting; older children's caregivers spent less
+time supporting their standing or sitting postures. Children's changing
+posture and orientation to their caregiver jointly shaped the amount of
+social information that was in their view during one-on-one play
+sessions with their caregivers: Children saw the most faces/hands when
+they were sitting or standing and close to their caregiver. Motor
+development appears to modulate how infants experience their visual
+world and the social information in it.
+
+However, we did not see differences in visual access to social
+information during naming events, suggesting that children did not
+change where they were looking when hearing a label for either a novel
+or familiar object. We see several potential explanations for this:
+first, other work (Yoshida \& Smith, 2008; Yu \& Smith, 2013, 2017),
+including Franchak et al. (2017), has found that infants spend much more
+time looking at the toys vs.~their caregiver's faces during these play
+sessions; infants may have thus been primarily interested in exploring
+these new toys rather than learning their names. Furthermore, moving
+their own neck upwards towards their caregiver still requires quite a
+physical effort at this age (especially when prone or sitting) and
+simply may not have been a priority for children in this context. A
+second factor is that this particular play session did not present many
+opportunities where children would need to use social cues to
+disambiguate referents. Indeed, there were only two possible referents
+in the room at a time----and one of them was always a familiar category
+(i.e., car, kitty).
+
+Broadly,these results show promise for the use of automated methods to
+detect the social information in the egocentric infant viewpoint during
+more naturalistic parent-child interactons. The use of this automated
+methodology allowed us to easily annotate the entirety of the dataset
+and analyze all of the frames without the need for hundreds of hours of
+human coding. Importantly, we also applied these same methods to an
+addiontal dataset that contained more variation in the type and
+structure of the play session, replicating and extending the effects of
+children's in-the-moment posture on visual access to social information
+(Franchak et al., 2017).
+
+Yet more work is needed to understand how these results relate to
+children's home experiences (Fausey et al., 2016). Both this play
+session and that of Franchak et al. (2017) were relatively controlled
+interactions where caregivers are highly aware that they are being
+montiored by experimenters. Furthermore, as children grow and change,
+the activities in which they engage with their caregivers are likely to
+also vary, leading to differences in the distribution of contexts they
+experience that may not be captured in these one-on-one play sessions.
+Finally, the ability to walk is of course only of a cascade of changes
+in children's abilities and experiences, and this study captures only a
+cross-sectional slice of this broader, multifaceted trajectory.
+
+Understanding these changes and how they relate to one another has been
+a persistent challenge for developmental psychology, but the field of
+computer vision has advanced dramatically in recent years, creating a
+new generation of algorithmic tools. These tools deal better with
+noisier, more complicated datasets and extract richer information than
+previous system. We hope that these new tools can be leveraged to
+understand the changing infant perspective on the visual world, both in
+controlled experimental contexts and in children's home environments. By
+doing so, we believe that we can understand the implications of the
+changing infant perspective consequences for linguistic, cognitive, and
+social development.
+
+\section{Acknowledgements}\label{acknowledgements}
+
+Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in
+recruitment, data collection, and annotation. This work was funded by a
+Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF,
+and NSF \#1714726 to BLL. An earlier version of this work was presented
+to the Cognitive Science Society in Frank et al. (2013) and Sanchez et
+al. (2018).
 
@@
 \emph{23}(3), 244--247.
+
+\hypertarget{ref-adolph1998roles}{}
+Adolph, K. E., Vereijken, B., \& Denny, M. (1998). Roles of variability
+and experience in development of crawling. \emph{Child Development},
+\emph{69}(1299), 312.
 
@@
 
+\hypertarget{ref-patrick2012developmental}{}
+Patrick, S. K., Noah, J. A., \& Yang, J. F. (2012). Developmental
+constraints of quadrupedal coordination across crawling styles in human
+infants. \emph{Journal of Neurophysiology}, \emph{107}(11), 3050--3061.
+
+\hypertarget{ref-sanchez2018postural}{}
+Sanchez, A., Long, B., Kraus, A. M., \& Frank, M. C. (2018). Postural
+developments modulate children's visual access to social information.
+
 \hypertarget{ref-scheirer2014perceptual}{}
@@
 Intelligence}, \emph{36}(8), 1679--1686.
+
+\hypertarget{ref-simon2015databrary}{}
+Simon, D. A., Gordon, A. S., Steiger, L., \& Gilmore, R. O. (2015).
+Databrary: Enabling sharing and reuse of research video. In
+\emph{Proceedings of the 15th acm/ieee-cs joint conference on digital
+libraries} (pp. 279--280).
 
@@
 
+\hypertarget{ref-yu2013joint}{}
+Yu, C., \& Smith, L. B. (2013). Joint attention without gaze following:
+Human infants and their parents coordinate visual attention to objects
+through eye-hand coordination. \emph{PloS One}, \emph{8}(11).
+
+\hypertarget{ref-yu2017hand}{}
+Yu, C., \& Smith, L. B. (2017). Hand--eye coordination predicts joint
+attention. \emph{Child Development}, \emph{88}(6), 2060--2078.
+
 \hypertarget{ref-yurovsky2012}{}
-Yurovsky, D., Smith, L., \& Yu, C. (in press). Statistical word learning
-at scale: The baby's view is better. \emph{Developmental Science}.
+Yurovsky, D., Smith, L., \& Yu, C. (2012). Statistical word learning at
+scale: The baby's view is better. \emph{Developmental Science}.
 

### 2020-04-29 writeup-paper/draft_apa_march2020.Rmd
updated xs-face manuscript

@@
 To estimate the proportion of faces and hands in the infant view, we used a computer vision algorithm to detect the presence of faces and hands in the entire dataset.
-  We found that infants' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field.
+  We found that infants\\' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field.
   We also applied our automated analysis to the egocentric video data from a recent paper documenting how posture changes one-year-olds visual access to faces (Franchak et al., 2017), finding convergence across both methodologies and dataset and confirming previous work that motoric developments play a significant role in the emergence of children’s linguistic and social capacities.
-  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience as they change over development.
+  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants\\' visual and linguistic experience as they change over development.
 
@@
 
-floatsintext      : yes
-figsintext        : yes
+figsintext        : no
 figurelist        : no
@@
                       fig.path='figs/', echo=FALSE, warning = FALSE, 
-                      cache=FALSE, message = FALSE, sanitize = TRUE)
+                      cache=TRUE, message = FALSE, sanitize = TRUE)
 library(readr)
@@
 library(papaja)
+library(png)
+library(magick)
 
@@
 These data were also analyzed in @frank2013 and @sanchez2018postural.
+
+### Head-mounted camera
+```{r headcam, fig.env = "figure", fig.height=3, fig.pos = "H", fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}
+
+img <- png::readPNG("images/viewangle_cut.png")
+grid::grid.raster(img)
+```
+
+We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband.
+Videos captured by the headcam were 720x480 pixels with 25 frames per second [^2].
+A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, above).
+
+[^2]: Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
+
+Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988].
+As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field).
 
@@
 
-```{r posture, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
+```{r posture, fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
 cowplot::plot_grid(pos_plot, orient_plot, nrow=2)
@@
 
-```{r indivPosOrient, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
+```{r indivPosOrient,  fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
 
@@
 
-```{r detByAge, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=3, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age.
+```{r detByAge, fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.height=3, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age.
 Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
@@
 
-```{r detByPosOrient, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation.
+```{r detByPosOrient, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation.
 Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted.
@@
 
-```{r detByNaming, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\\% bootstrapped confidence intervals."}
+```{r detByNaming, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\\% bootstrapped confidence intervals."}
 cowplot::plot_grid(faces_aroundnamings_plot, wrists_aroundnamings_plot)
@@
 
-```{r franchak, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture.
+```{r franchak, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture.
 Error bars represent 95 percent bootstrapped confidence intervals."}

### 2020-04-29 writeup-paper/draft_apa_march2020.tex
updated xs-face manuscript

@@
-\documentclass[english,man,floatsintext]{apa6}
+\documentclass[english,man]{apa6}
 \usepackage{lmodern}
@@
 \keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: X}
+\DeclareDelayedFloatFlavor{ThreePartTable}{table}
+\DeclareDelayedFloatFlavor{lltable}{table}
+\DeclareDelayedFloatFlavor*{longtable}{table}
+\makeatletter
+\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
+\makeatother
 \usepackage{csquotes}
@@
 vision algorithm to detect the presence of faces and hands in the entire
-dataset. We found that infants' posture and orientation to their
-caregiver changed dramatically across this age range and modulated their
-access to this social information; infants who were sitting or standing
-with their caregiver at a close distance tended to have the most
-faces/hands in their visual field. We also applied our automated
-analysis to the egocentric video data from a recent paper documenting
-how posture changes one-year-olds visual access to faces (Franchak et
-al., 2017), finding convergence across both methodologies and dataset
-and confirming previous work that motoric developments play a
-significant role in the emergence of children's linguistic and social
+dataset. We found that infants\textbackslash{}`posture and orientation
+to their caregiver changed dramatically across this age range and
+modulated their access to this social information; infants who were
+sitting or standing with their caregiver at a close distance tended to
+have the most faces/hands in their visual field. We also applied our
+automated analysis to the egocentric video data from a recent paper
+documenting how posture changes one-year-olds visual access to faces
+(Franchak et al., 2017), finding convergence across both methodologies
+and dataset and confirming previous work that motoric developments play
+a significant role in the emergence of children's linguistic and social
 capacities. We suggest that the combined use of head-mounted cameras and
 the application of new computer vision techniques is a promising avenue
-for understanding the statistics of infants' visual and linguistic
-experience as they change over development.
+for understanding the statistics of infants\textbackslash{}' visual and
+linguistic experience as they change over development.
 }
@@
 Pusiol (2013) and Sanchez, Long, Kraus, and Frank (2018).
+
+\subsubsection{Head-mounted camera}\label{head-mounted-camera}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics{figs/headcam-1.pdf}
+\caption{\label{fig:headcam}Vertical field of view for two different headcam
+configurations (we used the lower in our current study).}
+\end{figure}
+
+We used a small, head-mounted camera (\enquote{headcam}) that was
+constructed from a MD80 model camera attached to a soft elastic
+headband. Videos captured by the headcam were 720x480 pixels with 25
+frames per second.\footnote{Detailed instructions for creating this
+  headcam can be found at
+  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
+A fisheye lens was attached to the camera to increase the view angle
+from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
+\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
+\ref{fig:headcam}, above).
+
+Even with the fish-eye lens, the vertical field of view of the camera is
+still considerably reduced compared to the child's field of view, which
+spans around 100--120\(^{\circ}\) in the vertical dimension by 6-7
+months of age (Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton,
+1988; Mayer, Fulton, \& Cummings, 1988). As we were primarily interested
+in the presence of faces in the child's field of view, we chose to
+orient the camera upwards to capture the entirety of the child's upper
+visual field where the child is likely to see adult faces, understanding
+that this decision limited our ability to detect hands (especially those
+of the child, which are typically found at the bottom of the visual
+field).
 
@@
 
-\begin{figure*}[!h]
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{figs/posture-1} \caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:posture}
+\end{figure}
 
-{\centering \includegraphics[width=\textwidth]{figs/posture-1}
-
-}
-
-\caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:posture}
-\end{figure*}
-
-\begin{figure*}[!h]
-
-{\centering \includegraphics[width=\textwidth]{figs/indivPosOrient-1}
-
-}
-
-\caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:indivPosOrient}
-\end{figure*}
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{figs/indivPosOrient-1} \caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:indivPosOrient}
+\end{figure}
 
@@
 
-\begin{figure*}[!h]
+\begin{figure}[H]
+\includegraphics[width=\textwidth]{figs/detByAge-1} \caption{Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age.
+Larger dots indicate children who had longer play sessions and thus for whom there was more data.}\label{fig:detByAge}
+\end{figure}
 
-{\centering \includegraphics[width=\textwidth]{figs/detByAge-1}
-
-}
-
-\caption{Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age.
-Larger dots indicate children who had longer play sessions and thus for whom there was more data.}\label{fig:detByAge}
-\end{figure*}
-
-\begin{figure*}[!h]
+\begin{figure}[H]
 
@@
 Error bars represent 95\% bootstrapped confidence intervals}\label{fig:detByPosOrient}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[!h]
+\begin{figure}[H]
 
@@
 \caption{Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\% bootstrapped confidence intervals.}\label{fig:detByNaming}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}[!h]
+\begin{figure}[H]
 
@@
 Error bars represent 95 percent bootstrapped confidence intervals.}\label{fig:franchak}
-\end{figure*}
+\end{figure}
 
@@
 
+\hypertarget{ref-cummings1988}{}
+Cummings, M., Van Hof-Van Duin, J., Mayer, D., Hansen, R., \& Fulton, A.
+(1988). Visual fields of young children. \emph{Behavioural and Brain
+Research}, \emph{29}(1), 7--16.
+
 \hypertarget{ref-farroni2002eye}{}
@@
 \emph{Developmental Science}, \emph{17}(3), 388--395.
+
+\hypertarget{ref-mayer1988}{}
+Mayer, D., Fulton, A., \& Cummings, M. (1988). Visual fields of infants
+assessed with a new perimetric technique. \emph{Investigative
+Ophthalmology \& Visual Science}, \emph{29}(3), 452--459.
 

### 2020-04-30 writeup-paper/draft_apa_march2020.Rmd
edits to paper draft
make figures plausibly work in-line
add marginal r^2 analysis
other typos and edits

@@
   How do these postural and locomotor developments affect children's access to the social information relevant for word-learning? 
-  We explored this question by using head-mounted cameras to record the egocentric visual perspective of 36 infants' (at 8, 12, and 16 months of age) during a naturalistic play session with their caregiver.
+  We explored this question by using head-mounted cameras to record the egocentric visual perspective of 36 infants (at 8, 12, and 16 months of age) during a naturalistic play session with their caregiver.
 To estimate the proportion of faces and hands in the infant view, we used a computer vision algorithm to detect the presence of faces and hands in the entire dataset.
-  We found that infants\\' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field.
+  We found that infants\' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field.
   We also applied our automated analysis to the egocentric video data from a recent paper documenting how posture changes one-year-olds visual access to faces (Franchak et al., 2017), finding convergence across both methodologies and dataset and confirming previous work that motoric developments play a significant role in the emergence of children’s linguistic and social capacities.
-  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants\\' visual and linguistic experience as they change over development.
+  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants\' visual and linguistic experience as they change over development.
 
 keywords          : "Postural developments modulate children's visual access to social information"
-wordcount         : "X"
+wordcount         : "4295"
 
@@
 # Methods
-### Participants
+## Participants
 ```{r echo=FALSE, include=FALSE, demographics}
@@
 
-### Head-mounted camera
+## Head-mounted camera
 ```{r headcam, fig.env = "figure", fig.height=3, fig.pos = "H", fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}
@@
 
-### Procedure
+## Procedure
 All parents signed consent documents while children were fitted with the headcam.
@@
 
-### Data Processing and Annotation
+## Data Processing and Annotation
 Headcam videos were trimmed such that they excluded the instruction phase when the experimenter was in the room and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software.
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
-\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
-The last column features a false positive from OpenPose and a false negative from MTCNN.}
-\end{figure*}
+```{r frames, fig.env = "figure", fig.height=3, fig.pos = "H", fig.label="headcam", fig.cap = "Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
+The last column features a false positive from OpenPose and a false negative from MTCNN."}
+
+img <- png::readPNG("images/detector_samples_banner.png")
+grid::grid.raster(img)
+```
 
@@
 
+```{r runModels, eval=TRUE, include=TRUE}
+library(MuMIn)
+
+#wrangle data
+## Using MTCNN detections for GLM since it was the best
+dGLM_faces <- d %>%
+  mutate(subid = as.factor(subid)) %>%
+  group_by(subid, posture, orientation) %>%
+  summarize(countFaces=sum(faceOP==TRUE), countNotFaces = sum(faceOP==FALSE), age.at.test=age.at.test[1]) %>%
+  ungroup()
+
+dGLM_wrists <- d %>%
+  mutate(subid = as.factor(subid)) %>%
+  group_by(subid, posture, orientation) %>%
+  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
+  ungroup()
+
+### FACES
+
+# Age as fixed effect, full random effects
+faceGLM_age_only <- glmer(cbind(countFaces,countNotFaces) ~  scale(age.at.test) +
+(posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+
+# Age and posture and orientation as fixed effects, full random effects
+faceGLM_no_int <- glmer(cbind(countFaces,countNotFaces) ~ posture + orientation + scale(age.at.test) +  (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+
+# Age and posture * orientation as fixed effects, full random effects
+faceGLM_full <- glmer(cbind(countFaces,countNotFaces) ~ posture * orientation + scale(age.at.test) +  (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+
+### WRISTS
+# Age as fixed effect, full random effects
+wristGLM_age_only <- glmer(cbind(countWrists,countNotWrists) ~ scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+
+wristGLM_no_int <- glmer(cbind(countWrists,countNotWrists) ~ orientation + posture + scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+
+# Age and posture and orientation as fixed effects, full random effects
+wristGLM_full <- glmer(cbind(countWrists,countNotWrists) ~ orientation * posture + scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+```
+
+```{r eval=FALSE}
+## output
+cleanedNames_Int = c("Intercept","Sit","Stand","Close","Far","Age (Scaled)", "Sit*Close", "Stand*Close","Sit*Far","Stand*Far")
+
+#
+faceGLMOut=(summary(faceGLM_full))
+rownames(faceGLMOut$coefficients)<-cleanedNames_Int
+#
+wristGLMOut=(summary(wristGLM_full))
+rownames(wristGLMOut$coefficients)<-cleanedNames_Int
+#
+
+###
+xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.")
+
+xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")
+```
+
+```{r}
+wrists_age_only = r.squaredGLMM(wristGLM_age_only)
+wrists_full= r.squaredGLMM(wristGLM_full)
+
+faces_age_only =  r.squaredGLMM(faceGLM_age_only)
+faces_full =r.squaredGLMM(faceGLM_full)
+
+# wrists_no_int= r.squaredGLMM(wristGLM_no_int)
+# faces_no_int =  r.squaredGLMM(faceGLM_no_int)
+```
+
 These trends were quantified using two generalized linear mixed-effect models estimating the proportion of faces and hands that were in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture.
-A summary of the coefficients of the models can be found in Table 2, confirming that infants who were sitting/standing and in front of their caregivers saw the most faces and hands.
+A summary of the coefficients of the models can be found in Tables 2 and 3, confirming that infants who were sitting/standing and in front of their caregivers saw the most faces and hands.
 When modeling the proportion of faces seen, age was not a significant predictor; however, age remained a significant predictor when modeling the proportion of hands seen by infants.
+
+We directly examined the contributions of posture and orientation by fitting a reduced version of the full model [@nakagawa2013general] without their fixed effects (both models were run with the maximal random effects structure).
+The fixed effects in a model with only the age of the participants accounted for relatively little variance in the proportion of faces (marginal $R^2$ < `r round(faces_age_only[1,1],5)`) or hands in view  (marginal $R^2$ = `r round(wrists_age_only[1,1],3)`).
+However, when adding children's posture and orientation to their caregiver to the model (and their interaction) the marginal $R^2$ were higher for both faces (marginal $R^2$ = `r round(faces_full[1,1],3)`) and wrists (marginal $R^2$ = `r round(wrists_full[1,1],3)`).
+
 Overall, these results suggest that infants' visual access to social information is largely modulated by their posture and orientation to their caregiver, which is in turn a function of their general locomotor development.
@@
 These results suggest that either children did not actively seek out social information during this particular play session around naming events, perhaps because were a limited number of possible referents at any given time.
-
-```{r runModels, eval=FALSE, include=FALSE}
-#wrangle data
-## Using MTCNN detections for GLM since it was the best
-dGLM_faces <- d %>%
-  group_by(subid, posture, orientation) %>%
-  summarize(countFaces=sum(faceOP==TRUE), countNotFaces = sum(faceOP==FALSE), age.at.test=age.at.test[1]) %>%
-  ungroup()
-
-dGLM_wrists <- d %>%
-  group_by(subid, posture, orientation) %>%
-  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
-  ungroup()
-
-## run models 
-faceGLM <- glmer(cbind(countFaces,countNotFaces) ~ posture * orientation + scale(age.at.test) + 
-               (posture + orientation | subid), 
-             dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
-
-wristGLM <- glmer(cbind(countWrists,countNotWrists) ~ orientation * posture + scale(age.at.test) + 
-               (posture + orientation | subid), 
-             dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
-
-## output ---- 
-cleanedNames_Int = c("Intercept","Sit","Stand","Close","Far","Age (Scaled)", "Sit*Close", "Stand*Close","Sit*Far","Stand*Far")
-
-# 
-faceGLMOut=(summary(faceGLM))
-rownames(faceGLMOut$coefficients)<-cleanedNames_Int
-# 
-wristGLMOut=(summary(wristGLM))
-rownames(wristGLMOut$coefficients)<-cleanedNames_Int
-#
-
-### 
-xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.")
-
-xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")
-
-```
 
@@
 Nonetheless, if children are often orienting their heads towards where they are allocating their attention [@yoshida2008] then we should expect to find the same pattern of results only when analyzing the information in view.
-
 ```{r loadFranchakData}

### 2020-04-30 writeup-paper/draft_apa_march2020.tex
edits to paper draft
make figures plausibly work in-line
add marginal r^2 analysis
other typos and edits

@@
 Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
-\keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: X}
+\keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: 4295}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 question by using head-mounted cameras to record the egocentric visual
-perspective of 36 infants' (at 8, 12, and 16 months of age) during a
+perspective of 36 infants (at 8, 12, and 16 months of age) during a
 naturalistic play session with their caregiver. To estimate the
@@
 vision algorithm to detect the presence of faces and hands in the entire
-dataset. We found that infants\textbackslash{}`posture and orientation
-to their caregiver changed dramatically across this age range and
-modulated their access to this social information; infants who were
-sitting or standing with their caregiver at a close distance tended to
-have the most faces/hands in their visual field. We also applied our
-automated analysis to the egocentric video data from a recent paper
-documenting how posture changes one-year-olds visual access to faces
-(Franchak et al., 2017), finding convergence across both methodologies
-and dataset and confirming previous work that motoric developments play
-a significant role in the emergence of children's linguistic and social
+dataset. We found that infants' posture and orientation to their
+caregiver changed dramatically across this age range and modulated their
+access to this social information; infants who were sitting or standing
+with their caregiver at a close distance tended to have the most
+faces/hands in their visual field. We also applied our automated
+analysis to the egocentric video data from a recent paper documenting
+how posture changes one-year-olds visual access to faces (Franchak et
+al., 2017), finding convergence across both methodologies and dataset
+and confirming previous work that motoric developments play a
+significant role in the emergence of children's linguistic and social
 capacities. We suggest that the combined use of head-mounted cameras and
 the application of new computer vision techniques is a promising avenue
-for understanding the statistics of infants\textbackslash{}' visual and
-linguistic experience as they change over development.
+for understanding the statistics of infants' visual and linguistic
+experience as they change over development.
 }
@@
 
-\subsubsection{Participants}\label{participants}
+\subsection{Participants}\label{participants}
 
@@
 
-\subsubsection{Head-mounted camera}\label{head-mounted-camera}
+\subsection{Head-mounted camera}\label{head-mounted-camera}
 
@@
 
-\subsubsection{Procedure}\label{procedure}
+\subsection{Procedure}\label{procedure}
 
@@
 
-\subsubsection{Data Processing and
+\subsection{Data Processing and
 Annotation}\label{data-processing-and-annotation}
@@
 
-\begin{figure*}
+\begin{figure}[htbp]
 \centering
-\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
-\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
-The last column features a false positive from OpenPose and a false negative from MTCNN.}
-\end{figure*}
+\includegraphics{figs/frames-1.pdf}
+\caption{\label{fig:frames}Example face and pose detections made by OpenPose
+(top row) and MTCNN (bottom row) from a child in each age group. The
+last column features a false positive from OpenPose and a false negative
+from MTCNN.}
+\end{figure}
 
@@
 and posture. A summary of the coefficients of the models can be found in
-Table 2, confirming that infants who were sitting/standing and in front
-of their caregivers saw the most faces and hands. When modeling the
-proportion of faces seen, age was not a significant predictor; however,
-age remained a significant predictor when modeling the proportion of
-hands seen by infants. Overall, these results suggest that infants'
-visual access to social information is largely modulated by their
-posture and orientation to their caregiver, which is in turn a function
-of their general locomotor development. Nonetheless, these results also
-suggest that infants may still increase their attention to hands
-throughout this age range as they continue to learn about objects, their
-functions, and their names.
+Tables 2 and 3, confirming that infants who were sitting/standing and in
+front of their caregivers saw the most faces and hands. When modeling
+the proportion of faces seen, age was not a significant predictor;
+however, age remained a significant predictor when modeling the
+proportion of hands seen by infants.
+
+We directly examined the contributions of posture and orientation by
+fitting a reduced version of the full model (Nakagawa \& Schielzeth,
+2013) without their fixed effects (both models were run with the maximal
+random effects structure). The fixed effects in a model with only the
+age of the participants accounted for relatively little variance in the
+proportion of faces (marginal \(R^2\) = 0.00) or hands in view (marginal
+\(R^2\) = 0.15). However, when adding children's posture and orientation
+to their caregiver to the model (and their interaction) the marginal
+\(R^2\) were higher for both faces (marginal \(R^2\) = 0.28) and wrists
+(marginal \(R^2\) = 0.37).
+
+Overall, these results suggest that infants' visual access to social
+information is largely modulated by their posture and orientation to
+their caregiver, which is in turn a function of their general locomotor
+development. Nonetheless, these results also suggest that infants may
+still increase their attention to hands throughout this age range as
+they continue to learn about objects, their functions, and their names.
 
@@
 
+\hypertarget{ref-nakagawa2013general}{}
+Nakagawa, S., \& Schielzeth, H. (2013). A general and simple method for
+obtaining r2 from generalized linear mixed-effects models. \emph{Methods
+in Ecology and Evolution}, \emph{4}(2), 133--142.
+
 \hypertarget{ref-patrick2012developmental}{}

### 2020-08-11 analysis-franchak/franchak_reanalysis.Rmd
Update franchak re-analysis with new pipeline

--- 2020-04-10 analysis-franchak/franchak_reanalysos.Rmd
@@
 library(cowplot)
-
 ```
@@
 
+## How much data do we have for each subject?
+```{r}
+(frames_by_sub <- all_pos_data %>%
+  group_by(sub_id) %>%
+  summarize(num_frames = length(unique(frame))) %>%
+  kable())
+```
+
 ## Read in detections and merge with subid from above
 ```{r}
-detections <- read.csv('../data/franchak_dataset/openpose_detections/franchak_frames_openpose_master_frames.csv') 
+csv_dir = paste0(here::here(),'/data/franchak_dataset/openpose_detections/csvs_compressed_with_conf/')
+files = dir(csv_dir)
+sub_count = 0
+
+for (f in files){
+  print(f)
+  sub_count = sub_count + 1
+  print(sub_count)
+  this_sub <- read.csv(paste0(csv_dir,f)) %>%
+    as.tibble()
+
+  if (sub_count==1){
+     detections <- this_sub
+  } else{
+  detections <- detections %>%
+    full_join(this_sub) 
+  }
+}
+
 dets <- detections %>%
+  mutate(vid_name = str_split_fixed(subid,'.json',2)[,1]) %>%
+  select(-subid) %>%
   left_join(filenames) %>%
-  mutate(sub_id = as.factor(sub_id))
+  mutate(sub_id = as.factor(sub_id)) %>%
+  rename(frame = frame_num)
+
 ```
@@
 pos_detections  <- all_pos_data %>%
-  left_join(dets) 
+  left_join(dets, by=c('frame','sub_id')) 
 ```
@@
 
+pos_detections <- pos_detections %>%
+  mutate(face_openpose = as.logical(face_openpose), hand_openpose = as.logical(hand_openpose))
+
 ```
@@
 ```{r}
-save(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections.RData', pos_detections)
+# save(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections_july2020.RData', detections)
+save(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections_by_posture_july2020.RData', pos_detections)
+
 ```
@@
   group_by(sub_id, infant_posture_label) %>%
-  summarize(prop_faces = mean(face_openpose)) %>%
+  dplyr::summarize(prop_faces = mean(face_openpose)) %>%
   group_by(infant_posture_label) %>%
@@
   theme_few())
+
 ```
@@
 
+```{r}
+faces_by_both_pos  <- pos_detections %>%
+  group_by(sub_id, infant_posture_label, caregiver_posture_label) %>%
+  summarize(prop_faces = mean(face_openpose)) %>%
+  group_by(infant_posture_label,caregiver_posture_label) %>%
+  multi_boot_standard(col = 'prop_faces')
+
+(ggplot(faces_by_both_pos, aes(x=infant_posture_label, y=mean)) +
+  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
+  theme_few() +
+  facet_wrap(~caregiver_posture_label))
+
+```
+
 ### Plot hands by infant posture
@@
   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
-  geom_point(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label), alpha=.2) +
+  geom_point(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_hands, col=infant_posture_label), alpha=.2) +
   theme_few(base_size=18) +
@@
   scale_color_manual(name = "Posture", values=c("#339966","#eb7928","#0099cc")))
-)
 

### 2020-08-11 data/franchak_dataset/openpose_detections/json_to_csv.py
redo franchak preprocessing files

@@ -1,0 +1,20 @@
+import os
+import numpy as np
+import pandas as pd
+
+DATA_PATH = "/Users/brialong/Documents/GitHub/xs-face/data/franchak_dataset/openpose_detections/full_jsons/"
+OUT_PATH = "/Users/brialong/Documents/GitHub/xs-face/data/franchak_dataset/openpose_detections/"
+
+
+files = os.listdir(DATA_PATH)
+full_df = [] 
+
+for f in files:
+	print f
+	if not f.startswith('.'): 
+		file_path = os.path.join(DATA_PATH, f)
+		df = pd.read_json(file_path)
+		df['subid'] = f
+		this_sub=df[['subid','frame_num','face_openpose','hand_openpose','nose_conf','wrist_conf']]
+		this_sub.to_csv(OUT_PATH + f + '_openpose.csv', index=False)
+

### 2020-08-11 writeup-paper/render_examples.Rmd
scripts for naming/posture reliability and rendering examples

---
title: "render_examples"
author: "Bria Long"
date: "7/31/2020"
output: html_document
---

# Load libraries

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(lme4)
library(gridExtra)
library(assertthat)
library(tidyverse)
library(viridis)
library(papaja)
library(png)
library(magick)

# base_size in theme_few (changes font size in plots)
plot_size=16
```

####  Load detections
```{r loadDetections}

d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_july2020.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

## make long form version for plotting both face detections on same graph
dlong <- d %>%
  mutate(id = paste(subid,frame)) %>%
  gather(key = detectorType, value = detection, faceMT, faceOP, faceOPStrict, faceVJ, wristOP, wristOPStrict) %>%
  mutate(detectorType=as.factor(detectorType))  %>%
  mutate(detection = replace_na(detection,FALSE)) ## few frames where OP didn't have detections

# for better plotting names
levels(dlong$detectorType) 
#[1] "faceMT"        "faceOP"        "faceOPStrict"  "faceVJ"       
#[5] "wristOP"       "wristOPStrict"
levels(dlong$detectorType) <- c("MTCNN-Faces","OpenPose-Faces", "OpenPose-Faces-Strict","ViolaJones-Faces","OpenPose-Wrists","OpenPose-Wrists-Strict")

```

## Load ground truth
```{r groundTruthSetup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN GROUND TRUTH and load function
ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv") %>%
  mutate(sample_type=as.factor(sample_type))

ground_truth_wrists <- read_csv("../data/ground_truth/ground_truth_wrists.csv") %>%
  mutate(sample_type=as.factor(sample_type))

# Renaming for table output
levels(ground_truth_faces$sample_type)=c("High density", "Random")
levels(ground_truth_wrists$sample_type)=c("High density", "Random")

# Function to evaluate detectors
evaluate <- function(a, b) {
  if (a == TRUE) {
    if (a == b) return ("TP") # was face/wrist, detected face/wrist
    else return("FN") # was face/wrist, missed face/wrist
  }
  else {
    if (a == b) return("TN") # was not face/wrist, did not detect face/wrist
    else return("FP") # was not face/wrist, detected face/wrist
  }
}
```

## Evaluate face detectors
```{r}
performanceFaces <- ground_truth_faces  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong)   %>%
  filter(!detectorType %in% c('OpenPose-Wrists-Strict','OpenPose-Wrists')) %>%
  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face)) # always have the gronud truth first in evaluate argument
```

## Set up paths for getting detection images
```{r}
example_op_images = paste0(here::here(),'/data/openpose_examples/3sample/')
goldset_image_dir = paste0(here::here(),'/data/ground_truth/goldset_frames/face_images/')
dir.create(paste0(here::here(),'/data','/example_montages/'))
save_dir = paste0(here::here(),'/data','/example_montages/')
no_consent_show_images = c('XS_1224','XS_1622','XS_1628','XS_1631','XS_1634')

performanceFacesToRender <- performanceFaces %>%
  filter(!subid %in% no_consent_show_images) %>% #don't render frames from non-consenting fams
  mutate(frame_string = sprintf("image-%05d", frame)) %>%
  # mutate(image_path = paste0(goldset_image_dir, video, '/',frame_string,'.jpg')) %>%
  mutate(image_path = paste0(example_op_images, video, '_',frame_string,'_rendered.png')) %>% 
  #NOTE: only renders high confidence OP detections
  mutate(file_exists = file.exists(image_path)) %>%
  filter(file_exists==TRUE)
```

## Render out example errors for Viola Jones
```{r}
false_negs_VJ <- performanceFacesToRender %>%
  filter(detectorType=='ViolaJones-Faces') %>%
  filter(result=="FN") %>%
  filter(sample_type=='Random')

false_pos_VJ <- performanceFacesToRender %>%
  filter(detectorType=='ViolaJones-Faces') %>%
  filter(result=="FP") %>%
  filter(sample_type=='Random')

## note these will take a random sample, so set seed if trying to reproduce everything
# set.seed(30)
OpenPose_TP_ViolaJones_FN <- performanceFacesToRender %>%
  filter(image_path %in% false_negs_VJ$image_path) %>%
  filter(detectorType=='OpenPose-Faces') %>%
  filter(result=='TP')  %>%
  sample_n(8)

OpenPose_TN_ViolaJones_FP <- performanceFacesToRender %>%
  filter(image_path %in% false_pos_VJ$image_path) %>%
  filter(detectorType=='OpenPose-Faces') %>%
  filter(result=='TN') %>%
  sample_n(8)
```

### Make montages 
```{r}
image_read(false_negs_VJ$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "false_negs_VJ.png")))

image_read(false_pos_VJ$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "false_pos_VJ.png")))

image_read(OpenPose_TP_ViolaJones_FN$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "OpenPose_TP_ViolaJones_FN.png")))

image_read(OpenPose_TN_ViolaJones_FP$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "OpenPose_TN_ViolaJones_FP.png")))

false_negs_MTCNN <- performanceFacesToRender %>%
  filter(detectorType=='MTCNN-Faces') %>%
  filter(result=="FN")

OpenPose_TP_MTCNN_FN <- performanceFacesToRender %>%
  filter(image_path %in% false_negs_MTCNN$image_path) %>%
  filter(detectorType=='OpenPose-Faces') %>%
  filter(result=='TP') %>%
  sample_n(16)

image_read(OpenPose_TP_MTCNN_FN$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "OpenPose_TP_MTCNN_FN.png")))
```

## Where OpenPose had false alarms that MTCNN correctly rejected 
```{r}
false_pos_OP <- performanceFacesToRender %>%
  filter(detectorType=='OpenPose-Faces') %>%
  filter(result=="FP")

OpenPose_FP_MTCNN_TN <- performanceFacesToRender %>%
  filter(image_path %in% false_pos_OP$image_path) %>%
  filter(detectorType=='MTCNN-Faces') %>%
  filter(result=='TN') %>%
  sample_n(16)

image_read(OpenPose_FP_MTCNN_TN$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "OpenPose_FP_MTCNN_TN.png")))
```

## by posture
```{r}
sit <- performanceFacesToRender %>%
   filter(posture=='sit') %>%
   # filter(is_face==TRUE) %>%
   # filter(detectorType=='OpenPose-Faces-Strict') %>%
   # filter(detection==TRUE) %>%
   sample_n(8)

stand <- performanceFacesToRender %>%
   filter(posture=='stand') %>%
   # filter(is_face==TRUE) %>%
   # filter(detectorType=='OpenPose-Faces-Strict') %>%
   # filter(detection==TRUE) %>%
   sample_n(8)

prone <- performanceFacesToRender %>%
   filter(posture=='prone') %>%
   # filter(is_face==TRUE) %>%
   # filter(detectorType=='OpenPose-Faces-Strict') %>%
   # filter(detection==TRUE) %>%
   sample_n(8)

# carry <- performanceFacesToRender %>%
#    filter(posture=='carry') %>%
#    filter(is_face==TRUE) %>%
#    filter(detectorType=='OpenPose-Faces-Strict') %>%
#    filter(detection==TRUE) %>%
#    sample_n(8)

image_read(prone$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(save_dir, "prone.png")))

image_read(sit$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(save_dir, "sit.png")))

image_read(stand$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(save_dir, "stand.png")))

```

### 2020-08-11 writeup-paper/step1_merge_data_2020.Rmd
add writeup folder for apa draft
add strict faces/hands to main data structure, fix sync timing import error, save out new file

--- 2018-11-06 writeup-paper/step1_merge_data.Rmd
@@
   mutate(faceOP = Nose_conf!=0)  %>%  
-  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   
+  mutate(wristOP = LWrist_conf!=0 | RWrist_conf!=0 )   %>%
+  mutate(faceOPStrict = Nose_conf>.5)  %>%  
+  mutate(wristOPStrict = LWrist_conf>.5 | RWrist_conf>.5 )   
 ```
@@
 alldets=left_join(dets_MTCNN,detsViola[,c("video","frame","faceVJ")])  ## Join MTCNN + Viola Jones
-alldets=left_join(alldets,detsOpenPose[,c("video","frame","faceOP","wristOP")])  ##  ++ OpenPose
+alldets=left_join(alldets,detsOpenPose[,c("video","frame","faceOP","wristOP", "faceOPStrict", "wristOPStrict")])  ##  ++ OpenPose
 ```
@@
 
-## Let's look at which frames these are
+## Let's look at which frames these areall
 missingOP<- alldets %>% 
@@
   group_by(subid) %>%
-  do(add.posture(.)) ## adds both postures and orientatinos
+  do(add.posture(.)) %>% ## adds both postures and orientatinos
+  do(add.posture.second(.)) ## gets reliability coding for postures
+```
+
+```{r}
+# for reliability anlayses
+# all_dets_stripped <- alldets %>%
+#   select(subid, group, frame)
+# 
+# save(all_dets_stripped, file = '../data/consolidated_data/video_by_frame.RData')
+```
+
+### Recheck posture syncing times
+```{r}
+sync_times_check <- read_csv("../data/manual_sync_times/video_sync_times.csv") 
+s <- sync_times_check %>%
+  rename(subid = sid) %>%
+  group_by(subid) %>%
+  mutate(sync_time_string =toString(sync_time_stamp)) %>%
+  mutate(frames = str_split_fixed(sync_time_string,":",3)[,3]) %>%
+  mutate(sec = str_split_fixed(sync_time_string,":",3)[,2]) %>%
+  mutate(min = str_split_fixed(sync_time_string,":",3)[,1]) %>%
+  mutate(sync_time_ms = as.numeric(min)*60000 + as.numeric(sec)*1000 + as.numeric(frames)*33.33)
+
+posture.sync <- function(x) {
+  fname <- paste("../data/posture/",
+                 x$subid[1],
+                 ".csv",
+                 sep="")
+   if (file.exists(fname)) {
+
+    postures <- read_csv(fname)
+    postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order -- the case for some kids
+    x$posture_start_time = postures$start[1]
+    print(paste0("**** ", x$subid[1], " has time",x$posture_start_time))
+   } else {
+     print(paste0("**** ", x$subid[1], " is missing posture ****"))
+   }
+  return(x)
+}
+
+s <- s %>%
+  group_by(subid) %>%
+  do(posture.sync(.))
+
+## problematic kids
+# we want posture coding to start BEFORE the sync time, can zero after
+issues = (s$posture_start_time>s$sync_time_ms) # problematic if greater
+diffs =  s$posture_start_time-s$sync_time_ms
+off_by = diffs[issues]
+seconds_off = round(off_by / 1000*-1,2)
 ```
@@
 ```{r}
-# write_csv(d, "../data/consolidated_data/consolidated_data_4detectors_test.csv") #
+write_csv(d, "../data/consolidated_data/consolidated_data_4detectors_july2020.csv") #
 ```

### 2020-08-11 writeup-paper/helper.R
add functions for getting second naming/posture coding, clean up

@@ -53,6 +53,40 @@ add.posture <- function(x) {
   return(x)
 }
 
+add.posture.second <- function(x) {
+  fname <- paste("../data/posture/second/",
+                 x$subid[1],
+                 ".csv",
+                 sep="")
+  
+  if (file.exists(fname)) {
+    print(fname)
+    postures <- read_csv(fname)
+    postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order
+    
+    ## this reorders the postures so that they are "zeroed" relative to the beginning of the headcam videos
+    postures <- regularize.postures(subset(postures,code=="posture"),x$subid[1]) ## needs subid to get real sync time later 
+    x$posture_second <- factor(NA,levels=levels(postures$posture))
+    x$orientation_second <- factor(NA,levels=levels(postures$orientation))
+    
+    # for each row of p, populate posture and orientation to x
+    for (i in 1:nrow(postures)) {
+      range <- x$time > postures$start[i] & x$time <= postures$end[i]
+      x$posture_second[range] <- postures$posture[i]
+      x$orientation_second[range] <- postures$orientation[i]
+      print(paste0("**** ", x$subid[1], " has  second posture ****"))
+      
+    }
+  } else {
+    x$posture_second <- NA
+    x$orientation_second <- NA
+  }
+  
+  return(x)
+}
+
+
+
 get_sub_sync_time <- function(this_sub_id){
   sync_times_check_file <- read_csv("../data/manual_sync_times/video_sync_times.csv") 
   s <- sync_times_check_file %>%
@@ -61,7 +95,7 @@ get_sub_sync_time <- function(this_sub_id){
     mutate(frames = str_split_fixed(sync_time_string,":",3)[,3]) %>%
     mutate(sec = str_split_fixed(sync_time_string,":",3)[,2]) %>%
     mutate(min = str_split_fixed(sync_time_string,":",3)[,1]) %>%
-    mutate(sync_time_ms = as.numeric(min)*6000 + as.numeric(sec)*1000 + as.numeric(frames)*33.33)
+    mutate(sync_time_ms = as.numeric(min)*60000 + as.numeric(sec)*1000 + as.numeric(frames)*33.33)
 
   sync_time = s$sync_time_ms
   return(sync_time)
@@ -141,6 +175,7 @@ regularize.naming <- function (n) {
   n$familiarity <- factor(n$familiarity)
   
[…]

### 2020-08-11 writeup-paper/old-drafts/backup.Rmd
code for possible extra plots

--- 2020-04-29 writeup-paper/draft_apa_march2020.Rmd
@@
----
-title             : "Postural developments modulate children's visual access to social information"
-shorttitle        : "Posture modulates social access"
-
-author: 
-  - name          : "Bria L. Long"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "450 Serra Mall, Stanford CA 94305"
-    email         : "bria@stanford.edu"
-  - name          : "Alessandro Sanchez"
-    affiliation   : "1"
-  - name          : "Allison M. Kraus"
-    affiliation   : "1"
-  - name          : "Ketan Agrawal"
-    affiliation   : "1"
-  - name          : "Michael C. Frank"
-    affiliation   : "1"
-
-affiliation:
-  - id            : "1"
-    institution   : "Department of Psychology, Stanford University"
-
-# authornote: |
-#   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-# 
-#   Enter author note here.
-
-abstract: |
-  The ability to process social information is a critical component of children's early linguistic and cognitive development.
-  However, as children reach their first birthday, they begin to locomote themselves, dramatically changing the way they see the world around them. 
-  How do these postural and locomotor developments affect children's access to the social information relevant for word-learning? 
-  We explored this question by using head-mounted cameras to record the egocentric visual perspective of 36 infants' (at 8, 12, and 16 months of age) during a naturalistic play session with their caregiver.
-To estimate the proportion of faces and hands in the infant view, we used a computer vision algorithm to detect the presence of faces and hands in the entire dataset.
-  We found that infants' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field.
-  We also applied our automated analysis to the egocentric video data from a recent paper documenting how posture changes one-year-olds visual access to faces (Franchak et al., 2017), finding convergence across both methodologies and dataset and confirming previous work that motoric developments play a significant role in the emergence of children’s linguistic and social capacities.
-  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants' visual and linguistic experience as they change over development.
-
-keywords          : "Postural developments modulate children's visual access to social information"
-wordcount         : "X"
-
-bibliography      : ["xsface.bib"]
-
-floatsintext      : yes
-figsintext        : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : no
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_pdf
----
-
-```{r, libraries}
-library(knitr)
-knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
-                      fig.path='figs/', echo=FALSE, warning = FALSE, 
-                      cache=FALSE, message = FALSE, sanitize = TRUE)
-library(readr)
-library(dplyr)
-library(tidyr)
-library(xtable)
-library(lubridate)
-library(langcog)
-library(ggplot2)
-library(ggthemes)
-library(lme4)
-library(gridExtra)
-library(assertthat)
-library(tidyverse)
-library(viridis)
-library(papaja)
-
-# base_size in theme_few (changes font size in plots)
-plot_size=16
-```
-
-From their earliest months, infants are deeply engaged in learning from others.
-Even newborns tend to prefer to look at faces with direct vs.
-averted gaze [@farroni2002eye] and young infants follow overt gaze shifts.
-And as infants learn to parse and segment their native language(s), the social cues provided by speakers (e.g., eye-gaze) may provide strong scaffolding for early word learning.
-Indeed, empirical work suggests that children’s ability to process social cues is a key factor in early language development: in a longitudinal study, children’s level of joint engagement with their mother at 9-12 months was found to predict both their receptive and productive vocabularies [@carpenter1998].
-More specifically, 10 month-olds who follow an adult’s gaze in an experimental context have larger vocabularies at 18 months and throughout the second year of life [@brooks2005; @brooks2008].
-
-Around their first birthday, however, children’s ability to interact with their world radically changes [@adolph2006motor] as they are no longer constrained to the same spot that their caregivers last placed them in.
-Before the first birthday, children often begin crawling; soon after, they begin to walk independently.
-While not all children crawl [@adolph1998roles]—or crawl in the same way—children tend to find creative ways of moving (e.g., scooting) or getting others to help them move (e.g cruising) to begin to explore their world on their own [@patrick2012developmental].
-As independent agents in the world, children are more active in the construction of their own learning environments--yet spend much of their time in a world primarily populated by knees
-
-One possibility is that these motor improvements have strong developmental cascades, impacting children’s emerging social, cognitive, and linguistic abilities [@iverson2010].
-Indeed, these postural changes change how children interact with their mothers; walking infants make different kinds of object-related bids for attention from their mothers than crawling infants, and tend to hear more action directed statements (e.g., “open it”) [@karasik2014].
-In an observational study, @walle2014 also found that children who were able to walk had both higher receptive and productive vocabularies.
-On their account, children’s ability to stand and independently locomote may fundamentally change their ability to access the social information (e.g., facial expressions, gaze cues, pointing) relative to children who are still crawling and sitting.
-In other words, the ability of walking infants to access more detailed social information may allow infants to learn words quicker and more efficiently, facilitating language growth.
-
-Over the past decade, researchers have started to use egocentric, head-mounted cameras to document the social information that infants and children have access to across early development, and to understand the degree to which there are substantial shifts in their viewpoints that may have downstream developmental consequences [@yoshida2008].
-Indeed, the infant visual perspective has proved to be both remarkably different than the adult perspective and not easily predicted by our own intuitions [@franchak2011, @yoshida2008, @clerkin2017].
-Over the first two years of life, the infant viewpoint seems to transition from primarily containing close up view of faces to capturing relatively restricted views of hands paired with the objects they are acting on [@fausey2016]; both computational and empirical work suggest that this restricted viewpoint may be more effective for learning about objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017].
-
-We hypothesized that children’s postural developments may be partially responsible for some of these changes in the infant perspective.
-During spontaneous play, toddlers are more likely to look at the floor while crawling than while walking (Franchak et al., 2011), when they have full visual access to their environment and the people in it (Kretch, Franchak, and Adolph, 2014).
-More recently, when 12-month-olds [@franchak2017see] participated in a free-play session with their caregivers, their own posture and as well as their caregiver’s posture influenced the proportion of time they spent looking at the faces and bodies of their caregivers [@franchak2017see].
-
-Here, we directly examine the role that postural developments that infants’ experience as they reach their first birthday change the social information in the infant view.
-We expand on previous findings in three key ways.
-First, using head-mounted cameras during a brief laboratory free-play session, we record the visual experience of three groups of children at 8, 12, and 16 months-of-age, covering a broad range of ages and locomotor abilities around the first birthday; Children’s posture and orientation relative to their caregiver were also recorded from a third-person perspective and hand-annotated.
-This cross-sectional design thus allows us to directly examine the relative contributions of age vs.
-postural developments on children’s visual access to social information.
-Second, we use a modern computer vision algorithm for the automated detection of faces and hands in the egocentric viewpoints of infants.
-In particular, we capitalize on recent improvements in face and pose detection algorithms [@zhang2016; @cao2017realtime] to analyze the frequencies of faces and hands in the child’s visual environment, both overall and relative to naming events by their caregivers.
-Use of this emerging technology allows us to annotate the entirety of the dataset, allowing a more complete picture of the changing infant perspective.
-We apply this same automated method to the dataset used in @franchak2017see, validating the generalizability of this technique.
-Third, as parents were presented with pairs of novel and familiar objects with their children during the play sessions, we were able to explore the availability of social signals during naming events.
-
-Thus, the current study allows us to analyze changes in the visual access to faces and hands according to children's age, posture, and linguistic input.
-Broadly, we predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces/hands because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape with greater access to the social information in their environment.
-
-<!-- Second, we predicted that there would be a greater proportion of faces/hands in view around naming events—-broadly construed as any time a parent mentioned a label for one of the objects in the play set -->
-
-# Methods
-### Participants
-```{r echo=FALSE, include=FALSE, demographics}
-all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
-  rename(subid = subID)
-inc_subs <- read.csv("../data/demographics/demographics.csv")
-all_subs$inc <- all_subs$reason=="included"
-
-inc_subs$len.min <- sapply(strsplit(as.character(inc_subs$len),":"),
-                           function(x) {
-                             x <- as.numeric(x)
-                             x[1]+x[2]/60
-                           }
-)
-
-all_subs %>% 
-  group_by(reason) %>%
-  summarise(n = n()) %>% 
-  kable
-
-all_subs %>%
-  left_join(inc_subs) %>%
-  filter(inc==TRUE) %>%
-  group_by(age_group) %>%
-  summarise(n = sum(inc),
-            percent_included = mean(inc),
-            age = mean(age.at.test, na.rm=TRUE),
-            video_length = mean(len.min, na.rm=TRUE),
-            female = sum(gender=="female", na.rm=TRUE)) 
-  # xtable %>%
-  # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
-
-# <!-- \caption{\label{tab:pop} Demographics by age group.} -->
-```
-
-Our final sample consisted of 36 infants and children, with 12 participants in three age groups: 8 months (6 F), 12 months (7 F), and 16 months (6 F).
-Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80 percent English at home.
-Demographics and exclusion rates are  given in the table below.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrr}
-\hline
-Group & N & \% incl. & Avg age & Avg video length (min) \\
-\hline
-8 mo. &   12 & 0.46 & 8.71 & 14.41 \\
-12 mo. &  12 & 0.40 & 12.62 & 12.71 \\
-16 mo. &  12 & 0.31 & 16.29 & 15.10\\
-\hline
-\end{tabular}
-\end{table}
-
-To obtain this final sample, we tested `r length(all_subs$reason)` children, excluding `r sum(all_subs$reason!="included")` children for the following reasons:
-`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
-`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
-`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
-`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
-`r sum(all_subs$reason=="no CDI")` for missing Communicative Development Inventory (CDI) data,
-`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
-`r sum(all_subs$reason=="infant crying")` for fussiness, and one for sample symmetry.
-All inclusion decisions were made independent of the results of subsequent analyses.
-These data were also analyzed in @frank2013 and @sanchez2018postural.
-
 ### Procedure
@@
 
-\begin{figure*}
-\centering
-\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
-\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
-The last column features a false positive from OpenPose and a false negative from MTCNN.}
-\end{figure*}
+<!-- \begin{figure*} -->
+<!-- \centering -->
+<!-- \includegraphics[width=5.5in]{images/detector_samples_banner.pdf} -->
+<!-- \caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group.
+The last column features a false positive from OpenPose and a false negative from MTCNN.} -->
+<!-- \end{figure*} -->
 
@@
 The second algorithm (based on work by @zhang2016) uses multi-task cascaded convolutional neural networks (MTCNNs) for joint face detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
-We used a Tensorflow implementation of this algorithm available at https://github.com/davidsandberg/facenet.
+We used a Tensorflow implementation of this algorithm avaliable at https://github.com/davidsandberg/facenet.
 
@@
 
-```{r loadDetections, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
+```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
 ### READ IN DETECTIONS
@@
 
-```{r groundTruthSetup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
+```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
 ### READ IN GROUND TRUTH and load function
@@
 
-```{r evaluateDetectors, include=FALSE, echo=FALSE}
+```{r  include=FALSE, echo=FALSE}
 ## Evalute face and wrist detectors
@@
 `r round(perfFacesOverall$f[2],2)` OpenPose).
-MTCNN and OpenPose performed comparably with the random sample, with MTCNN having higher precision on the high density sample, suggesting that OpenPose generated slightly more false positives than MTCNN.
+Although MTCNN and OpenPose performed comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN.
 ViolaJones performed quite poorly relative to the other detectors, especially with respect to the random sample.
+We thus use MTCNN detections in the following analyses.
 For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision but low recall on the randomly sampled frames (see Table 1).
@@
 \hline
-MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
+MTCNN-Faces & High density & 0.89 & 0.92 & \textbf{0.90} \\ 
 MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
 OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
-OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
+OpenPose-Faces & Random & 0.72 & 0.80 & \textbf{0.76} \\ 
 ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
@@
 \caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected).
-P, R, and F denote precision, recall, and F-score, respectively.}
+P, R, and F denote precision, recall, and F-score, respectively.
+Scores in bold are the highest F-scores for each sample type.}
 \vspace{-1em}
@@
 # Results
-First, we report developmental shifts in infants' posture and their orientation relative to their caregiver, consistent with previous literature [@adolph2006motor; @franchak2017see].
+First, we report developmental shifts in infants' posture and their orientation relative to their caregiver, consistent with previous literature [@adolph2006; @franchak2017].
 Then, we examine how these changes influence children's visual access to faces and wrists/hands across this developmental time range, examining the relative contributions of age vs.
@@
 We also explore how these changes impact the accessibility of faces and wrists/hands during labeling events, both when parents were labeling familiar objects (e.g., 'cat') as well as novel objects (e.g., 'zem').
-Finally, we apply the same automated detection method to a egocentric video dataset collected by a different lab [@franchak2017see] during which children's in-the-moment posture was annotated.
+Finally, we apply the same automated detection method to a egocentric video dataset collected by a different lab [@franch2017] during which children's in-the-moment posture was annotated, replicating their existing results and finding converging support across methods, labs, and paradigms for the impact of postural developments on children's visual access to social information.
 
-```{r calcPostAndOrient}
+```{r}
 ## Calculate distributions of infants spent in each category separately for posture / orientation
@@
 
-```{r setupPosOrientPlot}
+```{r}
 ## Make plots of time spent in each orientation / posture
@@
   xlab("Age (months)") + 
-  theme_few(base_size=plot_size) +
+  theme_few() +
   ylim(0,1) +
@@
 
-```{r posture, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
+```{r posture, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
 cowplot::plot_grid(pos_plot, orient_plot, nrow=2)
@@
 
-```{r setupIndivPosOrientPlot}
+```{r}
 ### Calculate individual amounts of time spent in each orientation by subject
@@
   mutate(pos_orient = paste0('Infant ',posture, ', CG ', orientation))
+
 posture_by_orientation <- posture_by_orientation %>%
@@
 
-```{r indivPosOrient, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
+```{r individ_pos_orient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=3, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
 
@@
   ylim(0,1) +
-  theme_few(base_size=plot_size) +
+  theme_few(base_size=18) +
   theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
-  scale_fill_manual(name = "Posture and CG Orientation", values=c("#8878cc","#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc", "#98a69f", "#e67300","#ffa64d", "#ffe6cc","#d1b673"))
+  scale_fill_manual(name = "Posture & CG Orientation", values=c("#8878cc","#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc", "#98a69f", "#e67300","#ffa64d", "#ffe6cc","#d1b673"))
 ```
@@
 We also observed changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions than did children at other ages (see Figure \ref{fig:posture}).
-However, we also saw considerable variability across children: some infants spent almost their entire time sitting at a close distance from their caregiver, whereas others showed more considerable variability (see Figure \ref{fig:indivPosOrient}).
+However, we also saw considerable variability across children: some infants spent almost their entire time sitting at a close distance from their caregiver, whereas others showed more considerable variability (see Figure \ref{fig:individ_pos_orient}).
 
 ## Changes in Access to Faces and Hands
-We first examined the proportion of face and hand detections as a function of children's age without considering their posture (see Figure \ref{fig:detByAge}); here, we report face and wrist detections using OpenPose, although the same pattern of results for face detections was found using the outputs from MTCNN.
+We first examined the proportion of face and hand detections as a function of children's age without considering their posture (see Figure \ref{fig:detByAge}).
 While faces tended to be in the field-of-view overall more often than hands, children's head-mounted cameras were angled slightly upward to capture the presence of faces, and hand detections suffered from somewhat lower recall than face detections.
 Going forward, we thus analyze differences in the relative proportion of faces or hands in view as a function of age, posture, and orientation, rather than comparing them directly.
-Overall, we observed that 12-month-olds appeared to have visual access to slightly fewer faces than 8 or 16-month-olds, creating a slight U-shaped function in face detections; conversely, hand detections were showed a slight increase across this age range, as reported in prior literature [@fausey2016].
+Overall, we observed that 12-month-olds appeared to have visual access to slightly fewer faces than 8 or 16-month-olds, creating a slight U-shaped function in face detections; conversely, hand detections were showed a slight increase across this age range, as reported in prior literature [@Fausey2016].
 
@@
 When caregivers were behind their children, supporting their children's sitting or standing positions, children saw fewer faces and wrists.
-In particular, children who were sitting or in front of their caregiver had a high proportion of faces and hands in their field of view (Figure \ref{fig:detByPosOrient}).
+In particular, children who were sitting or in front of their caregiver had a high proportion of faces and hands in their field of view.
+(Figure \ref{fig:detByPosOrient}, lower panel).
 
@@
 ## Access to Faces and Hands During Labeling Events
-Our play session was designed to provide parents with opporunities to label objects--both familiar and novel--such that we could examine whether children sought out different kinds of social information around naming events.
+Our play session was designed to provide parents with opporunities to label objects--both familiar and novel--such that we could examine whether children seek out different kinds of social information around  naming events.
 We thus explored how face and hand detections changed during object labeling events, analyzing a four-second window around each labeling event (e.g., "Look at the [zem]!").
@@
 
-```{r runModels, eval=FALSE, include=FALSE}
+```{r eval=FALSE, include=FALSE}
 #wrangle data
@@
   group_by(subid, posture, orientation) %>%
-  summarize(countFaces=sum(faceOP==TRUE), countNotFaces = sum(faceOP==FALSE), age.at.test=age.at.test[1]) %>%
+  summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1]) %>%
   ungroup()
@@
 
-## output ---- 
-cleanedNames_Int = c("Intercept","Sit","Stand","Close","Far","Age (Scaled)", "Sit*Close", "Stand*Close","Sit*Far","Stand*Far")
-
+## output ---- no interaction
+# cleanedNames_NoInt = c("Intercept","Close","Far","Sit","Stand","Age") 
 # 
-faceGLMOut=(summary(faceGLM))
-rownames(faceGLMOut$coefficients)<-cleanedNames_Int
+# faceGLMOut=(summary(faceGLM))
+# rownames(faceGLMOut$coefficients)<-cleanedNames_NoInt
 # 
-wristGLMOut=(summary(wristGLM))
-rownames(wristGLMOut$coefficients)<-cleanedNames_Int
+# wristGLMOut=(summary(wristGLM))
+# rownames(wristGLMOut$coefficients)<-cleanedNames_NoInt
 # 
-
-### 
-xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.")
-
-xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")
+# xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed model predicting the proportion of faces vs. wrists seen by infants.") 
+# xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3))
 
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\
-  \hline
-Intercept & -3.35 & 0.17 & -19.49 & 0.000 \\
-  Sit & 0.31 & 0.20 & 1.58 & 0.114 \\
-  Stand & -0.02 & 0.20 & -0.12 & 0.905 \\
-  Close & 0.03 & 0.19 & 0.15 & 0.882 \\
-  Far & 0.46 & 0.24 & 1.93 & 0.053 \\
-  Age (Scaled) & 0.14 & 0.11 & 1.22 & 0.223 \\
-  Sit*Close & 0.92 & 0.07 & 13.31 & 0.000 \\
-  Stand*Close & 1.30 & 0.08 & 16.02 & 0.000 \\
-  Sit*Far & 0.54 & 0.07 & 7.37 & 0.000 \\
-  Stand*Far & 1.24 & 0.09 & 14.18 & 0.000 \\
-   \hline
-\end{tabular}
-\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants (as detected by OpenPose).}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\
-  \hline
-Intercept & -4.28 & 0.23 & -18.67 & 0.000 \\
-  Sit & 0.54 & 0.19 & 2.90 & 0.004 \\
-  Stand & 0.61 & 0.23 & 2.63 & 0.009 \\
-  Close & 0.67 & 0.22 & 3.05 & 0.002 \\
-  Far & 0.33 & 0.24 & 1.35 & 0.176 \\
-  Age (Scaled) & 0.45 & 0.13 & 3.35 & 0.001 \\
-  Sit*Close & 0.34 & 0.09 & 3.93 & 0.000 \\
-  Stand*Close & 0.73 & 0.09 & 7.92 & 0.000 \\
-  Sit*Far & -0.15 & 0.11 & -1.39 & 0.165 \\
-  Stand*Far & 1.32 & 0.11 & 11.70 & 0.000 \\
-   \hline
-\end{tabular}
-\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants (as detected by OpenPose).}
-\end{table}
-
-```{r setupdetsByAgePlot}
+```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age.
+Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
 detectionsByAge <- dlong %>%
@@
   mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video
-```
-
-```{r detByAge, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=3, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age.
-Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
 
 ggplot(detectionsByAge, aes(x = age.at.test, y = prop_detected, color = detectorType, size=num_detections)) +
-  theme_few(base_size=plot_size) + 
+  theme_few(base_size=16) + 
   geom_point(alpha=.5) +
@@
   scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
-  scale_size_continuous(name = "Frames") +
+  scale_size_continuous(name = "Number of frames") +
   labs(y = "Proportion detected", x = element_blank()) +
-  # scale_colour_discrete(name = element_blank()) +
+  scale_colour_discrete(name = element_blank()) +
   theme(legend.position="none") +
-  ggthemes::scale_color_solarized(name = "") + 
   facet_grid(~detectorType) 
-
 ```
 
-```{r setupdetByPosOrientPlot}
+```{r}
 ### Make dataframes for main plot (detections by orientation & posture)
@@
 
-```{r detByPosOrient, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation.
+```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation.
 Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted.
-Error bars represent 95\\% bootstrapped confidence intervals"}
+Error bars represent 95/% bootstrapped confidence intervals"}
 
@@
   labs(y = "Proportion detections", x = "Age (months)") +
-  theme_few(base_size=plot_size) +
-  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 12)) +
-  ggthemes::scale_color_solarized(name = "") + 
-  scale_size_continuous(name = 'Frames') +
+  theme_few(base_size=16) +
+  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
+  scale_colour_discrete(name = 'Detection type') +
+  scale_size_continuous(name = 'Number of frames') +
   scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
@@
 
-```{r summarizeNamings}
+```{r}
 source("helper.R")
@@
 
-```{r afterNamingsExplore, eval=FALSE, include=FALSE, echo=FALSE}
-afterNamings_Faces <- d %>%
-  group_by(subid) %>%
-  mutate(detections = faceOP) %>%
-  do(summarize.naming((.),window = c(0,4)))
-
-detectionsByGroup <- dlongPlot %>%
-  group_by(age.grp, subid, detectorType) %>%
-  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
-  group_by(age.grp, detectorType) %>%
-  multi_boot_standard(col = 'prop_detected')
-
-afterNamings_Detections_BySub_Faces <- afterNamings_Faces %>%
-  group_by(subid, familiarity, age.grp) %>%
-  summarize(prop_detected = mean(detections, na.rm=TRUE))
-
-afterNamings_Detections_Grouped_Faces <- afterNamings_Detections_BySub_Faces %>%
-  group_by(familiarity, age.grp) %>%
-  multi_boot_standard(col = 'prop_detected')
-
-faces_afternamings_plot <- ggplot(afterNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
-  # plot baseline
-  geom_pointrange(data = (detectionsByGroup %>% filter(detectorType == 'OpenPose-Faces')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=3), alpha=.9) +
-  # plot averages by familiarity 
-  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
-  # plot familairity detections for each sub
-  geom_point(data=afterNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
-  theme_few() +
-  labs(y = "Prop. face detections", x = "Age (months)") +
-  theme_few(base_size=plot_size) +
-  ylim(0,.6) +
-  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
-  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
-  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
-  scale_size_continuous(name="# of naming events")
-
-```
-
-```{r setUpNamingsPlot}
+```{r}
 ### Analyze detections around naming events (without posture/orientation)
@@
 
-```{r detByNamingPlotPanels}
+```{r}
 ## Plot faces/wrists around namings by age
@@
   labs(y = "Prop. face detections", x = "Age (months)") +
-  theme_few(base_size=plot_size) +
+  theme_few(base_size=16) +
   ylim(0,.6) +
@@
   labs(y = "Prop. wrist detections", x = "Age (months)") +
-  theme_few(base_size=plot_size) +
+  theme_few(base_size=16) +
   ylim(0,.6) +
@@
 
-```{r detByNaming, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\\% bootstrapped confidence intervals."}
+```{r detByNaming, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95 /% bootstrapped confidence intervals."}
 cowplot::plot_grid(faces_aroundnamings_plot, wrists_aroundnamings_plot)
@@
 
-## Extension to Franchak et al., 2017
+# Extension to Franchak et al., 2017
 In the present work, we found that infant's in-the-moment posture changed with their age, as did infants' orientation relative their caregiver.
-In a related study with 12-month-olds, @franchak2017see also that children's in-the-moment posture changed the amount of time infants spent looking at faces.
-Here, we sought to replicate the findings from @franchak2017see using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras [@simon2015databrary].
+In a related study with 12-month-olds, @franchak2017 also that children's in-the-moment posture changed the amount of time infants spent looking at faces.
+Here, we sought to replicate the findings from @franchak2017 using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras (shared on Databrary, @Databrary).
 
 This dataset differed from the present in two key ways that could present a challenge for our automated methodology.
-First, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to explore, unlike the  room used for the present dataset which was relatively small (approximately 10' x 10').
+First, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to explore, unlike the present room  dataset which was relatively small (approximately 10' x 10').
 Thus, our automated methodology could fail to generalize to scenes from these more complex environments, where detecting faces and hands could arguably be a much harder task.
-Second, @franchak2017see used a head-mounted eye-tracker, finding that children's posture affected where children were looking within their visual field.
-Nonetheless, if children are often orienting their heads towards where they are allocating their attention [@yoshida2008] then we should expect to find the same pattern of results only when analyzing the information in view.
+Second, @franchak2017 used a head-mounted eye-tracker, finding that children's posture affected where children were looking within their visual field.
+Nonetheless, if children are often orienting their heads towards where they are allocating their attention  [@Yoshida20008] then we should expect to find the same pattern of results only when analyzing the information in view.
 
-```{r loadFranchakData}
+```{r}
 # loads pos_detections used below for franchak dataset
@@
 
-```{r processFranchak}
+```{r}
 faces_by_pos  <- pos_detections %>%
@@
 
-```{r statsFranchak}
+```{r}
 # glmers for franchak dataset
@@
 
-```{r plotPanelsFranchak}
-faces_by_pos_plot <- ggplot(faces_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
+```{r}
+(faces_by_pos_plot <- ggplot(faces_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
@@
   geom_line(data=faces_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label, group=sub_id), alpha=.1) +
-  theme_few(base_size=plot_size) +
+  theme_few(base_size=18) +
   ylab('Prop faces detected') +
@@
   theme(legend.position = 'none') +
-  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928"))
+  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928")) )
 
-hands_by_pos_plot <- ggplot(hands_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
+(hands_by_pos_plot <- ggplot(hands_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
@@
   geom_line(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_hands, col=infant_posture_label, group=sub_id), alpha=.1) +
-  theme_few(base_size=plot_size) +
+  theme_few(base_size=18) +
   ylab('Prop hands detected') +
@@
   theme(legend.position = 'none') +
-  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928")) 
+  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928")) )
 ```
 
-```{r franchak, fig.env="figure*", out.width = "\\textwidth", fig.pos = "!h", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture.
+```{r franchak, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture.
 Error bars represent 95 percent bootstrapped confidence intervals."}
-cowplot::plot_grid(faces_by_pos_plot,hands_by_pos_plot, nrow=1)
+cowplot::plot_grid(faces_by_pos_plot,hands_by_pos_plot)
 ```
 
-Overall, we found convergence between our two methodologies: we replicated the main results from @franchak2017see, finding that the proportion of faces in view was greater when infants were sitting or standing vs.
+Overall, we found convergence between our two methodologies: we replicated @Franchak2017's main results, finding that the proportion of faces in view was greater when infants were sitting or standing vs.
 prone.
-We also found the same patterns of results with hands (i.e., wrist detections), though these were not originally annotated in @franchak2017see (see Figure \ref{fig:franchak}); results were validated with generalized mixed effect models as in our previous analysis.
-Interestingly, there was a relatively higher proportion of hands in view in @franchak2017see relative to the present dataset; this is likely mostly due to the fact that there were often multiple people in view: experimenters stayed in the room to film children and assist the caregiver, while in our present dataset the experimenters left the room for the majority of the play session.
-Nonetheless, we still found that posture modulated the proportion of hands in view, suggesting that this is a major factor that structures infants' access to visual information.
+We also found the same patterns of results with hands (i.e., wrist detectinos), which were not originally annotated in @Franch2017 (see Figure \ref{fig:franchak}); results were validated with generalized mixed effect models as in our previous analysis.
+Interestingly, there was a relatively higher proportion of hands in view in @Franchak2017 relative to the present dataset; this is likely mostly due to the fact that there were often multiple people in view: experimenters stayed in the room to film children and assist with technical difficulties, while in our present dataset the experiments left the room for the majority of the play session.
+Nonetheless, we still found that posture modulated the hands in view as well, suggesting that this is a major factor that structures infants' access to visual information.
 
-# General Discussion
-How do postural and locomotive developments influence the social information that infants see?
-We used a head-mounted camera to explore how these emerging motoric abilities affect children's visual access to social information, here operationalized as the presence of the faces and hands of their caregiver.
-First, we found systematic changes around the first birthday in children's in-the-moment posture and their orientation relative to the caregivers; older children spent more time standing and less time sitting; older children's caregivers spent less time supporting their standing or sitting postures.
-Children's changing posture and orientation to their caregiver jointly shaped the amount of social information that was in their view during one-on-one play sessions with their caregivers: Children saw the most faces/hands when they were sitting or standing and close to their caregiver.
-Motor development appears to modulate how infants experience their visual world and the social information in it.
-
-However, we did not see differences in visual access to social information during naming events, suggesting that children did not change where they were looking when hearing a label for either a novel or familiar object.
-We see several potential explanations for this: first, other work [@yoshida2008; @yu2013joint; @yu2017hand], including @franchak2017see, has found that infants spend much more time looking at the toys vs.
-their caregiver's faces during these play sessions; infants may have thus been primarily interested in exploring these new toys rather than learning their names.
-Furthermore, moving their own neck upwards towards their caregiver still requires quite a physical effort at this age (especially when prone or sitting) and simply may not have been a priority for children in this context.
-A second factor is that this particular play session did not present many opportunities where children would need to use social cues to disambiguate referents.
-Indeed, there were only two possible referents in the room at a time—-and one of them was always a familiar category (i.e., car, kitty).
-
-Broadly,these results show promise for the use of automated methods to detect the social information in the egocentric infant viewpoint during more naturalistic parent-child interactons.
-The use of this automated methodology allowed us to easily annotate the entirety of the dataset and analyze all of the frames without the need for hundreds of hours of human coding.
-Importantly, we also applied these same methods to an addiontal dataset that contained more variation in the type and structure of the play session, replicating and extending the effects of children's in-the-moment posture on visual access to social information [@franchak2017see].
-
-Yet more work is needed to understand how these results relate to children's home experiences [@fausey2016].
-Both this play session and that of @franchak2017see were relatively controlled interactions where caregivers are highly aware that they are being montiored by experimenters.
-Furthermore, as children grow and change, the activities in which they engage with their caregivers are likely to also vary, leading to differences in the distribution of contexts they experience that may not be captured in these one-on-one play sessions.
-Finally, the ability to walk is of course only of a cascade of changes in children's abilities and experiences, and this study captures only a cross-sectional slice of this broader, multifaceted trajectory.
-
-Understanding these changes and how they relate to one another has been a persistent challenge for developmental psychology, but the field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools.
-These tools deal better with noisier, more complicated datasets and extract richer information than previous system.
-We hope that these new tools can be leveraged to understand the changing infant perspective on the visual world, both in controlled experimental contexts and in children's home environments.
-By doing so, we believe that we can understand the implications of the changing infant perspective consequences for linguistic, cognitive, and social development.
-
-# Acknowledgements
-
-Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation.
-This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and NSF #1714726 to BLL.
-An earlier version of this work was presented to the Cognitive Science Society in @frank2013 and @sanchez2018postural.
-
-\newpage
-
-# References
-```{r create_r-references}
-r_refs(file = "xsface.bib")
-```
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup
+# Discussion

### 2020-08-11 writeup-paper/draft_apa_july2020.Rmd
latest version of writeup

---
title             : "Using automated detections to quantify the effect of postural developments on the social information in the infant view"

shorttitle        : "Automated detection of the social information"

author: 
  - name          : "Bria L. Long"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "450 Serra Mall, Stanford CA 94305"
    email         : "bria@stanford.edu"
  - name          : "Alessandro Sanchez"
    affiliation   : "1"
  - name          : "Allison M. Kraus"
    affiliation   : "1"
  - name          : "Ketan Agrawal"
    affiliation   : "1"
  - name          : "Michael C. Frank"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Stanford University"

# authornote: |
#   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
# 
#   Enter author note here.

abstract: |
Social information plays a key role in infants\' early linguistic and cognitive development. Yet infants locomotor abilities also develop continuously as they learn to sit, crawl, stand, and walk. How do these postural developments affect infants access to the social information relevant for early learning? To address this question, we first created a rich annotated dataset of naturalistic play sessions between infants and their caregivers with sets of novel and familiar toys (N=36, 8-16 months of age), recording both egocentric and third-person video. We then developed an automated analysis method using a pose detection model (Zhang et al., 2017) to detect the faces and hands in the infant view, allowing us to analyze the entirety of this dataset. We found that infants\' posture and orientation to their caregiver changed dramatically across this age range and modulated infants access to social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces and hands in their visual field. We also applied our automated method to a second egocentric video dataset where one-year-olds explored a large play room with their caregivers  (Franchak et al., 2017), andfound convergence across both datasets, confirming the finding that motoric developments play a significant role in shaping the social information that infants have access to over development, and suggesting that they may play a role in emergence of infants\' cognitive and linguistic abiltiies. We propose that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants\' visual and linguistic experience as they change over development and for re-utilizing rich video datasets collected across different labs and diverse contexts.

keywords          : "Postural developments, headcam, social information, face detection, pose detection"
wordcount         : "XXX"

bibliography      : ["xsface.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=FALSE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(lme4)
library(gridExtra)
library(assertthat)
library(tidyverse)
library(viridis)
library(papaja)
library(png)
library(magick)
library(lmerTest)

# base_size in theme_few (changes font size in plots)
plot_size=16
```

From their earliest months, infants are deeply engaged in learning from others. Even newborns tend to prefer to look at faces with direct vs. averted gaze [@farroni2002eye] and young infants follow overt gaze shifts [@gredeback2010development; @bruner1975]. And as infants reach their first birthday, they also tend to follow [@yu2013joint; @yu2017hand] and imitate the gestures of their caregivers (e.g., pointing).

Infants ability to process these social cues may provide strong scaffolding for early word learning. Longitudinal studies provide some evidence for this link: children’s level of joint engagement with their mother at 9-12 months predicts both their receptive and productive vocabularies [@carpenter1998] and 10 month-olds who follow an adult’s gaze (in an experimental context) have larger vocabularies at 18 months and throughout the second year of life [@brooks2005; @brooks2008]. While the relationship between hand-following and language development has been less well characterized, infants who follow their caregivers’ hands tend to be those who spend more time jointly attending to events with their caregivers [@yu2017hand].

Relatively little work, however, has quantified how often infants see and use these kinds of social cues in naturalistic learing environments. By using head-mounted cameras to record what infants see, researchers have begun to document the infant egocentric perspective [@yoshida2008] and to quantify the information -- social and otherwise -- available to infants as they learn. While head-mounted camera data do not provide explicit information on where infants are attending, some work suggests that infants orient their head towards what they are focusing on -- putting those people or objects in view [@yoshida2008]. Initial recordings from in-lab play sessions revealed a different view than many imagined: instead of being dominated by their faces, the infant perspective contained close up views of primarily toys and hands [@yoshida2008; @franchak2011; @yu2017hand]. Subsequent research has revealed that the infant view is far from a unitary construct, undergoing dramatic changes as infants grow. Recordings from home environments suggest that the viewpoints of very young infants -- less than 4 months of age -- do indeed contain persistent and frequent views of faces [@fausey2016; @jayaraman2017faces] but that the infant view tends to contain more and more hands as infants grow older.

These changes in perspective are likely the downstream consequence of a myriad factors, chief of which may be infants' evolving locomotive abilities: an infants ability to sit, crawl, stand, or walk structures the way they interact with the things and people in their world.  These changes can be rapid: the physical body of an infant can lengthen up to 2cm in 24 hours [@lampl1993evidence], leading to changes in their motoric abilities from one day to the next. These motoric developments can be thought of as gateways that open up entirely new phases of development [@iverson2010], and cause a cascade of changes in infants ability to interact with their world and the people in it [@karasik2014].

Thus, infants’ changing locomotor abilities may shape the social cues that infants' see and seek out, in turn impacting their cognitive and linguistic abilities. Supporting this idea, infants experience with sitting predicts their success at 3D object completion tasks [@soska2010systems] as well as their receptive vocabulary [@libertus2016sit]. And as children begin crawling [@adolph1998roles] -- or scooting or cruising [@patrick2012developmental] -- they are no longer constrained to the same spot that their caregivers last placed them in. Yet while crawlers can choose where to go and what they see to a much greater degree, they also appear to spend much of their time in a world populated by floors and knees; during spontaneous play, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011], when they have full visual access to their environment and the people in it [@kretch2014].

On one view, it is children’s ability to stand and walk that fundamentally changes their ability to access social information (e.g., facial expressions, gaze cues, pointing) relative to children who are still crawling and sitting [@walle2016infant]. In turn, this ability to access more detailed social information could allow infants to learn words quicker and more efficiently, facilitating language growth [@walle2016infant]. Supporting this idea, walking infants make different kinds of object-related bids for attention from their caregivers than crawling infants, and tend to hear more action directed statements (e.g., “open it”) [@karasik2014]; further, in an observational study, @walle2014 also found that children who were able to walk had both higher receptive and productive vocabularies. However, not all evidence supports this view: some parental report data suggest a weak relationship between walking and the onset of language [@moore2019point], and head-mounted eye-tracking data have highlighted the role of infants’ in-the-moment posture -- and that of their caregivers’ -- in what infants’ see, finding that these factors are better predictors of the social information in view than their locomotor status per se (i.e., “walker” vs “crawler”) [@franchak2017see].

Understanding this nuanced relationship between the social information in the infant view and their motoric and linguistic development has remained challenging. Indeed, different kinds of play sessions (e.g., exploring novel environments vs. playing with novel objects) might modulate the degree to which infants’ posture changes the social information in view. Furthermore, no work to date has directly examined how the availability of social information changes in naturalistic language learning contexts. For example, it is unclear if and how the social information in view changes during the naming events for different objects (e.g, “Look, a [ball]”).

More broadly, the field is in need of computational tools to reutilize these rich head-mounted camera datasets and to understand the generalizability of findings across different datasets, populations, tasks, and age-ranges. While the field has assembled many head-mounted camera datasets, conducting new analyses on these videos has remained prohibitively time-consuming due to a lack of computational tools for annotations. Instead, hundreds of hours of manual annotations have been required to analyze a fraction of the available frames for a given analysis. Thus, despite containing a wealth of information about the structure of parent-child interactions, these datasets have thus gone dramatically underused.

Recent innovations in computer vision hold promise for automated annotations of the infant view. Over the past decade, deep neural networks have become dramatically better at a wide range of visual tasks, including object classification, @simonyan2014very], scene categorization [@zhou2017places], and pose detection [@zhang2016], arguably facilitating our understanding of visual perception [@peterson2018evaluating, @vanrullen2017perception] and improving computational neuroscience [@kietzmann2018deep]. However, as most models have been trained on photographs or videos taken from the adult perspective, it is still unclear how easily these models can be applied to videos taken from the infant perspective.

In this paper, we thus first validate a model for the automated detection of faces and hands in the infant view, using a open-source model of pose detection -- OpenPose[@zhang2016; @cao2017realtime] -- that provides pose, face, and hand keypoints for every person in an image. We compare the detection accuracy of OpenPose with that of both older and more specialized models of face detection and demonstrate the usability of this off-the-shelf model for quantifying the faces and hands in the infant view.

We then apply this automated method to a head-mounted camera dataset collected in our lab to directly examine the role of postural developments on the social information in the infant view during language learning. Specifically, we constructed a dataset where infants and caregivers played freely using a set of novel and familiar objects; 8, 12, and 16 months-olds (N=26; N=12 in each age group) participated to ensure a wide range of locomotor abilities. Infants’ in-the-moment posture and orientation relative to their caregiver were hand annotated and all play sessions were transcribed. This cross-sectional design thus allowed us to directly examine the relative contributions of age vs. postural developments on children’s visual access to social information, and how the avaliability of social cues changes relative to naming events (e.g., “Yes, you like the [ball!]”).  Broadly, we predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces/hands because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape with greater access to the social information in their environment.

To examine the robustness of our automated method and the generalizability of our findings, we applied OpenPose [@cao2017realtime] to the dataset collected in @franchak2017see, where one-year-olds wore head-mounted eye-tracking cameras during a play session and their in-the-moment posture was hand-annotated. Unlike the present dataset, infants and caregivers roamed a large, open playroom and explored different toys placed throughout. We thus analyzed this dataset with the goal of both (1) validating our automated method on a very different kind of dataset, attempting to replicate their primary findings originally obtained with head-mounted eye-tracking  (2) extending their findings to ask whether whether the proportion of hands in view also varies with infants’ posture in an exploratory play session.

# Methods
## Free-play video dataset  
A primary goal of this project was to provide a rich, open dataset for other researchers to examine the effects of postural developments and labeling behavior during naturalistic parent-child interactions. We thus invited caregivers of 8, 12, and 16-month-olds to participate in play sessions where they were provided with pairs of novel and familiar objects (e.g., a ball and a microfiber duster, called a "zem") in a small playroom in a lab (approximately 10' x 10'). Infants wore head-mounted cameras equipped with a fish-eye lens (see Head-mounted camera), a third-person camera captured a birds-eye view of the play session.  Using these video data, infants' posture and orientation to their caregiver were hand-coded and annotated for the entirety of the play session. All videos were transcribed, and we collected collected MacArthur CDIs for all children who particpated. All materials have been made publicly avaliable on Databrary for whom the parents provided sharing consent (29/36 dyads) via https://nyu.databrary.org/volume/101

### Participants
```{r echo=FALSE, include=FALSE, demographics}
all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
  rename(subid = subID)
inc_subs <- read.csv("../data/demographics/demographics.csv")
all_subs$inc <- all_subs$reason=="included"

inc_subs$len.min <- sapply(strsplit(as.character(inc_subs$len),":"),
                           function(x) {
                             x <- as.numeric(x)
                             x[1]+x[2]/60
                           }
)

all_subs %>% 
  group_by(reason) %>%
  summarise(n = n()) %>% 
  kable

all_subs %>%
  left_join(inc_subs) %>%
  filter(inc==TRUE) %>%
  group_by(age_group) %>%
  summarise(n = sum(inc),
            percent_included = mean(inc),
            age = mean(age.at.test, na.rm=TRUE),
            video_length = mean(len.min, na.rm=TRUE),
            female = sum(gender=="female", na.rm=TRUE)) 
  # xtable %>%
  # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)

# <!-- \caption{\label{tab:pop} Demographics by age group.} -->
```
Our final sample consisted of 36 infants and children, with 12 participants in three age groups: 8 months (6 F), 12 months (7 F), and 16 months (6 F). Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80 percent English at home. Demographics and exclusion rates are given in the table below (see Table 1)/

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
\hline
Group & N & \% incl. & Avg age & Avg video length (min) \\
\hline
8 mo. &   12 & 0.46 & 8.71 & 14.41 \\
12 mo. &  12 & 0.40 & 12.62 & 12.71 \\
16 mo. &  12 & 0.31 & 16.29 & 15.10\\
\hline
\end{tabular}
\caption{Exclusion rates and summary demographics for the infants included in the study.} 
\end{table}

To obtain this final sample, we tested `r length(all_subs$reason)` children, excluding `r sum(all_subs$reason!="included")` children for the following reasons:
`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam (e.g., failure to record, ran out of battery),
`r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
`r sum(all_subs$reason=="less than 4 min HC footage")` for fewer than 4 minutes of headcam footage,
`r sum(all_subs$reason=="multiple people")` for having multiple adults present,
`r sum(all_subs$reason=="no CDI")` for missing Communicative Development Inventory (CDI) data,
`r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
`r sum(all_subs$reason=="infant crying")` for fussiness, and one for sample symmetry. 16-month-olds tolerated the head-mounted camera less well than younger infants, leading to a higher exclusion rate. All inclusion decisions were made independent of the results of subsequent analyses.

### Head-mounted camera
```{r headcam, fig.env = "figure", fig.pos = "H", out.width="40%", fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}
knitr::include_graphics("images/view_angle_redone.png")
```

We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband. Videos captured by the headcam were 720x480 pixels with 25 frames per second [^2]. A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, above).

[^2]: Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.

Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988]. As we were originally primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field).

### Procedure
All parents signed consent documents while children were fitted with the headcam. If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam. When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session. Parents were shown a box containing three pairs of familiar and novel objects. These pairs consisted of a ball paired with a microfiber duster (a "zem"), a toy car paired with a cheese grater (a "manu"), and a brush paired with a back massager (a "tima"). Parents were instructed to play with the object pairs with their child one at a time, "as they typically would." 
[^3]:The first few participants played with a different pair of objects (a toy cat and pedicure foam piece) that was replaced as some infants chewed persistently on the foam.

All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys. The experimenter then left the playroom for approximately 15-20 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.

### Data processing and annotations
Headcam videos were trimmed such that they excluded the instruction phase when the experimenter was in the room and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software. These sessions yielded 
`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a million frames) of video, with an average video length of
`r round(mean(inc_subs$len.min), digits = 2)` minutes
(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).

```{r loadDetections, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN DETECTIONS
d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_july2020.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

## make long form version for plotting both face detections on same graph
dlong <- d %>%
  mutate(id = paste(subid,frame)) %>%
  gather(key = detectorType, value = detection, faceMT, faceOP, faceOPStrict, faceVJ, wristOP, wristOPStrict) %>%
  mutate(detectorType=as.factor(detectorType))  %>%
  mutate(detection = replace_na(detection,FALSE)) ## few frames where OP didn't have detections

# for better plotting names
levels(dlong$detectorType) 
#[1] "faceMT"        "faceOP"        "faceOPStrict"  "faceVJ"       
#[5] "wristOP"       "wristOPStrict"
levels(dlong$detectorType) <- c("MTCNN-Faces","OpenPose-Faces", "OpenPose-Faces-Strict","ViolaJones-Faces","OpenPose-Wrists","OpenPose-Wrists-Strict")

```

```{r}
# calculate NAs for posture/orientation
frame_rate = (1/25)
NA_rates <- d %>%
  group_by(subid) %>%
  summarize(prop_na = sum(is.na(posture)), prop_na_orient= sum(is.na(orientation))) %>%
  ungroup() %>%
  summarize(mean_na_posture = mean(prop_na)*frame_rate,  sd_na_posture = sd(prop_na)*frame_rate, mean_na_orientation = mean(prop_na_orient)*frame_rate, sd_na_orientation = sd(prop_na_orient)*frame_rate)

```

#### Posture and caregiver orientation annotations
We created custom annotations to describe the child's physical posture (i.e. standing) and the orientation of the caregiver relative to the child (e.g. far away). The child's posture was categorized as being carried, prone (crawling or lying), sitting, or standing. The caregiver's orientation was characterized as being close, far, or behind the child (independent of distance). For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child. All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012toward]. Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations; similarily, times when the child was being carried or the caregivers were out of the frame were marked as uncodable for caregiver orientation. On average, posture or orientation was uncodable from 1-2 minutes of data in each child (seconds excluded from analysis for posture;. M =`r NA_rates$mean_na_posture`, SD =`r NA_rates$sd_na_posture`; orientation; M =`r NA_rates$mean_na_posture`, SD =`r NA_rates$sd_na_posture`), and these rates did not vary substantially with the age of the child.  To assess the reliability of these annotations, a second coder coded videos from three different children to calculate Cohen's kappa (posture, irr = XX, orientation, irr = XX).

#### Naming event annotations 
One coder listened to all of the audio from the play sessions and marked the exact timestamps whenever one of the novel or familiar objects was named in any instance (e.g., "Look at the [ball]", "Can you say [zem]?"); a second coder listened to the majority of the play sessions (N=23 sessions) and also annotated all naming events. To asesess reliability, we calculated the proportion of naming events detected by the first coder that were also annotated by a second coder within a sliding window. We found that 82.1% of naming events were detected within a 4 second window (+/- 2s), and 70.9% of namings events were detected within a 2 second window. We also obtained full text transcriptions of the entire play sessions (with time stamps marking 10s intervals). While these full transcriptions are not used in the present analyses, they have been made publically avaliable for future research.

## Face and hand detection
We evaluated three detection systems for the ability to measure infants' access to faces. The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones  [@viola2004robust]. We used this algorithm as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual]. We next tested the performance of two face detectors that both made use of relatively recently developed Convolutional Neural Networks (CNNs) to extract face information. The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract pose information of all the individuals in an image, operationalized as information about the position of 18 different body parts.  For this second algorithm [OpenPose; @cao2017realtime], we used the agent's nose (one of the body keypoints detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.

The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for hands for two reasons. First, as we did not capture children's entire visual field, the presence of a wrist is likely often indicative of the presence of a hand within the field of view. Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.

### Algorithms
Viola Jones, the first face detection system, made use of a series of Haar feature-based cascade classifiers [@viola2004robust] applied to each individual frame. The second algorithm (based on work by @zhang2016) uses multi-task cascaded convolutional neural networks (MTCNNs) for joint face detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present. We used a Tensorflow implementation of this algorithm available at https://github.com/davidsandberg/facenet.

The CNN-based pose detector  [OpenPose; @cao2017realtime; @simon2017hand; @wei2016cpm] provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose. The system uses a convolutional neural network for initial anatomical detection and subsequently applies part affinity fields for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose; here, we only made use of the body parts relevant to the face and hands (nose and wrists), though the entire set of keypoints is publicly avaliable. Each keypoint was accompanied by a confidence score made by the detector.

### Detector evaluation
To evaluate face detector performance, we hand-labeled a "gold set" of frames extracted from the video dataset. To account for the relatively rare appearance of faces in the dataset, we hand-labeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames. Each sample was comprised of an equal number of frames taken from each child's video. For wrist detections, the "gold set" was constructed in the same manner, except frames with a high density of wrists came only from detections made by OpenPose.  Faces were classified as present if at least half of the face was showing; wrists were classified as present if any part of the wrist was showing. Two authors labelled the frames independently and resolved disagreements on a case-by-case basis. Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors.

# Results
First, we report the accuracy of the automated detectors, as assessed by comparison to hand-labelled frames from the free-play video dataset described above. We then apply one of these automated detectors (OpenPose) to the entirety of this video dataset, and use these outputs to examine how postural developments influence children's visual access to faces and hands from 8-16 months of age, as well as how access to these social cues changes during labeling events (e.g., do you see the [zem]?). Finally, we apply this same automated detector to another video dataset [@franchak2017see], replicating and extending their findings on the effects of posture on visual access to social information.

```{r exampledetections, fig.env = "figure", out.width="100%", fig.pos = "H", fig.label="exampledetections", fig.cap = "Example detections made by OpenPose from children in each age group."}

knitr::include_graphics("images/exampe_detections.png")

```

```{r groundTruthSetup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN GROUND TRUTH and load function
ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv") %>%
  mutate(sample_type=as.factor(sample_type))

ground_truth_wrists <- read_csv("../data/ground_truth/ground_truth_wrists.csv") %>%
  mutate(sample_type=as.factor(sample_type))

# Renaming for table output
levels(ground_truth_faces$sample_type)=c("High density", "Random")
levels(ground_truth_wrists$sample_type)=c("High density", "Random")

# Function to evaluate detectors
evaluate <- function(a, b) {
  if (a == TRUE) {
    if (a == b) return ("TP") # was face/wrist, detected face/wrist
    else return("FN") # was face/wrist, missed face/wrist
  }
  else {
    if (a == b) return("TN") # was not face/wrist, did not detect face/wrist
    else return("FP") # was not face/wrist, detected face/wrist
  }
}
```

```{r evaluateDetectors, include=FALSE, echo=FALSE}
## Evalute face and wrist detectors
performanceFaces <- ground_truth_faces  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong)   %>%
  filter(!detectorType %in% c('OpenPose-Wrists-Strict','OpenPose-Wrists')) %>%
  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face)) # always have the gronud truth first in evaluate argument

perfFacesbySampleType <- performanceFaces %>%
  group_by(detectorType, sample_type) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN"))  %>%
  group_by(detectorType, sample_type) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
  # xtable %>%
  # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)

perfFacesOverall <- performanceFaces %>%
  group_by(detectorType) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  group_by(detectorType) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))

## EVALUATE WRIST DETECTOR
performanceWrists <- ground_truth_wrists  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong) %>%
  filter(detectorType %in% c('OpenPose-Wrists-Strict','OpenPose-Wrists')) %>%
  mutate(detector_is_wrist = detection, is_wrist = as.logical(is_wrist)) %>% 
  rowwise() %>%
  mutate(result = evaluate(is_wrist, detector_is_wrist))

perfWristsbySampleType<- performanceWrists %>%
  group_by(detectorType, sample_type) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  group_by(detectorType, sample_type) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
  xtable %>%
  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)

perfWristsOverall<- performanceWrists %>%
  group_by(detectorType) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  group_by(detectorType) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))

## Only keep what we're analyzing
dlong <- dlong %>%
  filter(detectorType %in% c('OpenPose-Faces','OpenPose-Wrists'))

```

## Accuracy of automated detections 
For face detection, we found that both OpenPose and MTCNN dramatically outperformed ViolaJones, our baseline model, especially with respect to the random sample, where ViolaJones generated a relatively high proportion of both false negatives and false positives. When considering only the composite F-score across all frames, MTCNN slightly outperformed OpenPose  (`r round(perfFacesOverall$f[1],2)` MTCNN vs. `r round(perfFacesOverall$f[2],2)` OpenPose), and MTCNN and OpenPose performed comparably with the random sample. Generally, MTCNN exhibited higher precision, whereas OpenPose exhibited higher recall, and these differences were most pronounced on the randomly sample frames. In other words, while OpenPose generated slightly more false positives than MTCNN, MTCNN missed several faces that were accurately detected by OpenPose. When we restricted our analysis to high-confidence detections from OpenPose (>.5 confidence; default threshold for visualization), we found very high precision (P = `r round(perfFacesOverall$p[3],2)`), but much lower recall (R = `r round(perfFacesOverall$r[3],2)`) and thus overall performance (F = `r round(perfFacesOverall$f[3],2)`), indicating that these low-confidence detections often indexed actual faces that were in the infant view.
Figure \ref{fig:exampledetections} shows example successful detections from OpenPose in each age group, and Figure \ref{fig:failures} shows examples of missed faces as well as false positives for context.

We next analyzed the viability of OpenPose as a hand detector. Despite the fact that hand detection is a more computationally challenging problem [@bambach2017], and the fact that we used wrist keypoints as a proxy for hands,  OpenPose performed moderately well as a hand detector (F = `r round(perfWristsOverall$f[1],2)`). OpenPose achieved relatively high precision -- generating relatively few false positives -- but showedlow recall on the randomly sampled frames (see Table 1).  As with face detections, when we restricted our analysis to high-confidence detections, we found much higher precision (P = `r round(perfWristsOverall$p[2],2)`), but much lower recall (R = `r round(perfWristsOverall$r[2],2)`) and thus overall performance (F = `r round(perfWristsOverall$f[2],2)`).

Thus, one major advantage of OpenPose relative to specialized face detectors, such as MTCNN, is that it allows the analysis of both the faces and hands in the infant view with the outputs of only one algorithm, and analyzing the results of all detections (regardless of confidence) yielded reasonably accurate results. Going forward, we analyze face and wrist detections using all detections from OpenPose, with the caveat that we are likely underestimating the proportion of hands in the dataset given the lower recall for hand detections.

```{r failures, fig.env = "figure", out.width="50%", fig.pos = "H", fig.label="headcam", fig.cap = "Example failed detections from OpenPose, showing both false positives and false negatives (i.e. missed detections)."}
knitr::include_graphics("images/example_failures.png")
```

```{r include=FALSE}
## Look at detector performance for precision, recall, and f score for each subjet
perfFacesbySampleTypeByAge <- performanceFaces %>%
  group_by(detectorType, subid, sample_type, age.at.test) %>%
  summarize(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN"))  %>%
  group_by(detectorType, subid, sample_type, age.at.test) %>%
  summarize(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))

ggplot(perfFacesbySampleTypeByAge, aes(x=age.at.test, y=p, col=sample_type)) +
  geom_point() +
  theme_few() +
  geom_smooth() +
  facet_wrap(~detectorType)

ggplot(perfFacesbySampleTypeByAge, aes(x=age.at.test, y=f, col=sample_type)) +
  geom_point() +
  theme_few() +
  geom_smooth() +
  facet_wrap(~detectorType)

ggplot(perfFacesbySampleTypeByAge, aes(x=age.at.test, y=r, col=sample_type)) +
  geom_point() +
  theme_few() +
  geom_smooth() +
  facet_wrap(~detectorType)
```

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
\hline
Algorithm & Sample\ Type & P & R & F \\ 
\hline
MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
OpenPose-Wrists & High density & 0.66 & 1.00 & 0.79 \\ 
OpenPose-Wrists & Random & 0.88 & 0.29 & 0.44 \\ 
\hline
\end{tabular}
\caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected). P, R, and F denote precision, recall, and F-score, respectively.} 
\vspace{-1em}
\end{table}

## Access to social information during postural developments

```{r calcPostAndOrient}
## Calculate distributions of infants spent in each category separately for posture / orientation
ages_inc_subs <- inc_subs %>%
  select(age.at.test, age.grp, subid)

orientation <- d %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, subid, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(subid, age.grp, orientation) %>%
  left_join(ages_inc_subs)

posture <- d %>%
  filter(!is.na(posture)) %>%
  group_by(age.grp, subid, posture) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(subid, age.grp, posture) %>%
  left_join(ages_inc_subs)
```

```{r setupPosOrientPlot}
plot_size=14

## Make plots of time spent in each orientation / posture
orientation_labels <- c(
  behind ='Caregiver behind',
  close = 'Caregiver close',
  far = 'Caregiver far')

posture_labels = c(
  carry = 'Infant carried',
  prone = 'Infant prone',
  sit = 'Infant sitting',
  stand ='Infant standing'
)

pos_plot <- ggplot(posture, aes(x = age.at.test, y = prop.time, col = posture, size=time)) + 
  geom_point(alpha=.6) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  theme_few(base_size=plot_size) +
  ylim(0,1) +
  facet_grid(~posture, labeller = labeller(posture = posture_labels)) +
  geom_smooth(span=5, alpha=.2) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_color_manual(name = "Posture", values=c("#8878cc","#0099cc","#339966","#eb7928")) +
  theme(legend.position = 'none')

orient_plot <- ggplot(orientation, aes(x = age.at.test, y = prop.time, col = orientation, size=time)) + 
  geom_point(alpha=.8) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  theme_few(base_size=plot_size) +
  ylim(0,1) +
  facet_grid(~orientation,  labeller = labeller(orientation= orientation_labels)) +
  geom_smooth(span=10, alpha=.2) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_color_manual(name = "CG", values=c("#171615","#5c5753","#99928d")) +
  theme(legend.position = 'none')
```

```{r posture, fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
cowplot::plot_grid(pos_plot, orient_plot, nrow=2)
```

```{r setupIndivPosOrientPlot}
### Calculate individual amounts of time spent in each orientation by subject
posture_by_orientation <- d %>%
  filter(!is.na(posture)) %>%
  filter(!is.na(orientation)) %>%
  group_by(subid, posture, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  group_by(subid) %>%
  mutate(prop.time = time/sum(time)) %>%
  mutate(pos_orient = paste0('Infant ',posture, ', CG ', orientation))

posture_by_orientation <- posture_by_orientation %>%
  left_join(ages_inc_subs) %>%
  ungroup() %>%
  mutate(subid = fct_reorder(subid, age.at.test))
```

```{r indivPosOrient,  fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times when infant was carried or when posture/orientation were not codable are ommitted for visualization purposes"}

## should be a two column figure
ggplot(posture_by_orientation, aes(x = subid, y = prop.time)) +
  geom_col(aes(fill = pos_orient), position=position_stack()) +
  ylab("Proportion Time") +
  xlab("Individual subjects (ordered by age)") +
  ylim(0,1) +
  theme_few(base_size=plot_size) +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  # scale_fill_manual(name = "Posture and CG Orientation", values=c("#8878cc","#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc", "#98a69f", "#e67300","#ffa64d", "#ffe6cc","#d1b673"))  # including NAs for orientation
scale_fill_manual(name = "Posture and CG Orientation", values=c("#006080", "#0099cc", "#99e6ff", "#339966","#66cc99", "#b3e6cc",  "#e67300","#ffa64d", "#ffe6cc"))

```

### Developmental changes in infant posture and caregiver orientation
We report developmental shifts in infants' posture and their orientation relative to their caregiver, consistent with previous literature [@adolph2006motor; @franchak2017see]. The proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with infants' age. As children got older, their locomotor abilities allowed them to become more independent. Both 8-month-olds and 12-month-olds spent relatively equivalent amounts of time lying/crawling (i.e., 'prone') which was markedly decreased in the 16-month-olds, who spent most of their time sitting or standing (see Figure \ref{fig:posture}). We also observed changes in infants' orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions than did children at other ages (see Figure \ref{fig:posture}). However, we also saw considerable variability across children: some infants spent almost their entire time sitting at a close distance from their caregiver, whereas others showed more considerable variability (see Figure \ref{fig:indivPosOrient}).

### Access to faces and hands
First, we examined the proportion of face and hand detections as a function of infants' age without considering their posture (see Figure \ref{fig:detByAge}).
While faces tended to be in the field-of-view overall more often than hands, infants' head-mounted cameras were angled slightly upward to capture the presence of faces, and hand detections suffered from somewhat lower recall than face detections. We thus only considered differences in the relative proportion of faces or hands in view as a function of age, posture, and orientation, rather than comparing them directly. Overall, we observed that 12-month-olds appeared to have visual access to slightly fewer faces than 8 or 16-month-olds, creating a slight U-shaped function in face detections; conversely, hand detections were showed a slight increase across this age range, as reported in prior work [@fausey2016].

```{r setupdetsByAgePlot}
detectionsByAge <- dlong %>%
  filter(detectorType!="MTCNN-Faces") %>% # for brevity
  filter(detectorType!="ViolaJones-Faces") %>% # low performance
  filter(!is.na(age.at.test)) %>% # filters NAs for combos of orientation/posture that were empty (see merge_data.Rmd)
  group_by(subid, age.at.test, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  left_join(inc_subs) %>%
  mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video
```

```{r detByAge, fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.height=3, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data."}

ggplot(detectionsByAge, aes(x = age.at.test, y = prop_detected, color = detectorType, size=num_detections)) +
  theme_few(base_size=plot_size) + 
  geom_point(alpha=.5) +
  geom_smooth(size=1, alpha=.3, span=10) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
  scale_size_continuous(name = "Frames") +
  labs(y = "Proportion detected", x = element_blank()) +
  # scale_colour_discrete(name = element_blank()) +
  theme(legend.position="none") +
  ggthemes::scale_color_solarized(name = "") + 
  facet_grid(~detectorType)

```

```{r setupdetByPosOrientPlot}
### Make dataframes for main plot (detections by orientation & posture)
# filter out detections w/o posture/orientation codings (wasn't always in view) -- this elimiantes "carry"
dlongPlot <-dlong %>%
  filter(!is.na(posture)) %>%
  filter(!is.na(orientation))

# pop out of actual ages of included subs
ages_inc_subs <- inc_subs %>%
  select(subid, age.at.test, age.grp)

# raw data
detectionsByPostureAndOrient <- dlongPlot %>%
  group_by(age.grp, posture, orientation, subid, detectorType) %>%
  summarise(prop_detected = mean(detection, na.rm=TRUE), num_detections = length(detection)) %>%
  left_join(ages_inc_subs) %>%
  mutate(age.at.test = as.numeric(age.at.test))

# with CIs
detectionsByPostureAndOrientByGroup <- dlongPlot %>%
  group_by(age.grp, orientation, posture, subid, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  group_by(age.grp, orientation, posture, detectorType) %>%
  multi_boot_standard(col = 'prop_detected') %>%
  left_join(ages_inc_subs)

# ggplot(detectionsByPostureAndOrient, aes(x=posture, y=mean, col=detectorType)) +
#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position=position_dodge(width=.8)) +
#   facet_wrap(~orientation) +
#   theme_few(base_size = plot_size)

```

```{r detByPosOrient, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\\% bootstrapped confidence intervals"}

# better labels 
orientation_labels <- c(
  behind ='Caregiver behind',
  close = 'Caregiver close',
  far = 'Caregiver far')

posture_labels = c(
  prone = 'Infant prone',
  sit = 'Infant sitting',
  stand ='Infant standing'
)

 ## plot data all together now, with CIs and individual points from each infant
ggplot(detectionsByPostureAndOrientByGroup, aes(x = age.grp, y = mean, col = detectorType)) + 
  # plot raw data
  geom_point(data=detectionsByPostureAndOrient, alpha=.2, aes(x = age.at.test, y=prop_detected, size=num_detections)) +
  # plot trend lines & SEs of trends based on raw data
  geom_smooth(data=detectionsByPostureAndOrient, aes(x = age.at.test, y=prop_detected, weight = num_detections), alpha=0.1, size=0, span=5) +
  stat_smooth(data=detectionsByPostureAndOrient, aes(x = age.at.test, y=prop_detected, weight = num_detections), geom="line", alpha=0.5, size=1, span=5) +
  # plot means and CIs based on age groups
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=1), alpha=.9) +
  # fix up the plot to look nice
  labs(y = "Proportion detections", x = "Age (months)") +
  theme_few(base_size=plot_size) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 12)) +
  ggthemes::scale_color_solarized(name = "") + 
  scale_size_continuous(name = 'Frames') +
  scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  facet_grid(orientation ~ posture, labeller = labeller(posture = posture_labels, orientation= orientation_labels))

```

```{r runModels, eval=TRUE, include=TRUE}
library(MuMIn)
#wrangle data

dGLM_faces <- d %>%
  mutate(subid = as.factor(subid)) %>%
  group_by(subid, posture, orientation) %>%
  summarize(countFaces=sum(faceOP==TRUE), countNotFaces = sum(faceOP==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

dGLM_wrists <- d %>%
  mutate(subid = as.factor(subid)) %>%
  group_by(subid, posture, orientation) %>%
  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

### FACES

# Age as fixed effect, full random effects
faceGLM_age_only <- glmer(cbind(countFaces,countNotFaces) ~  scale(age.at.test) +
(posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

# Age and posture and orientation as fixed effects, full random effects
faceGLM_no_int <- glmer(cbind(countFaces,countNotFaces) ~ posture + orientation + scale(age.at.test) +  (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

# Age and posture * orientation as fixed effects, full random effects
faceGLM_full <- glmer(cbind(countFaces,countNotFaces) ~ posture * orientation + scale(age.at.test) +  (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

### WRISTS
# Age as fixed effect, full random effects
wristGLM_age_only <- glmer(cbind(countWrists,countNotWrists) ~ scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

wristGLM_no_int <- glmer(cbind(countWrists,countNotWrists) ~ orientation + posture + scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

# Age and posture and orientation as fixed effects, full random effects
wristGLM_full <- glmer(cbind(countWrists,countNotWrists) ~ orientation * posture + scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
```

```{r}
## output
cleanedNames_Int = c("Intercept","Sit","Stand","Close","Far","Age (Scaled)", "Sit*Close", "Stand*Close","Sit*Far","Stand*Far")

#
faceGLMOut=(summary(faceGLM_full))
rownames(faceGLMOut$coefficients)<-cleanedNames_Int
#
wristGLMOut=(summary(wristGLM_full))
rownames(wristGLMOut$coefficients)<-cleanedNames_Int
#
```

```{r eval=FALSE} 
###
xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.")

xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")
```

```{r}
wrists_age_only = r.squaredGLMM(wristGLM_age_only)
wrists_full= r.squaredGLMM(wristGLM_full)

faces_age_only =  r.squaredGLMM(faceGLM_age_only)
faces_full =r.squaredGLMM(faceGLM_full)
```

However, these age-related trends were much smaller than the effect of infant's postural developments on infants' visual access to faces and hands. Infants' in-the-moment posture was a major factor both in how many faces and hands were in view during the play session, as was their orientation relative to their caregiver. This was quantified using two generalized linear mixed-effect models estimating the proportion of faces and hands that were in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture. A summary of the coefficients of the models can be found in Tables 2 and 3.

The interaction between infants' posture and their caregiver's orientation had the most dramatic effect on the social information in view. When caregivers were behind their infants, supporting their infants' sitting or standing positions, infants saw fewer faces. When caregivers were relatively close to their infants, infants who were sitting or standing had more faces in view (Face detections; infant sitting and CG close, b=`r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4],3)`; infant standing and CG close, b=`r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4],3)` than infants who were lying down/crawling (i.e. prone). When caregivers were far away from their infants, face detections were similarity higher (Face detections; infant sitting and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4],4)`; infant standing and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4],4)`.  Overall, we found that age did not remain a significant predictor in accounting for the faces in view when modeling infants posture, their orientation to their caregiver, and the interaction between these two factors (Face detections; Age (scaled), b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4],4)`).

We found a similar pattern of results for wrist detections, even though there were fewer wrist detections overall in the dataset. Infants saw fewer wrists when caregivers were behind their infants, supporting their infants' sitting or standing positions vs when caregivers were relatively closer to their infants (Hand detections; infant sitting and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4],4)`; infant standing and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4],4)` than infants who were lying down/crawling (i.e. prone). Wrist detections were clearly highest when caregivers were far away from their infants (Wrist detections; infant sitting and CG far, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4],4)`; infant standing and CG far, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4],4)`. As with faces,  age did not remain a significant predictor in these models (Wrist detections; Age (scaled), b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4],4)`).

We directly examined the contributions of posture and orientation by fitting a reduced version of the full model [@nakagawa2013general] without their fixed effects (both models were run with the maximal random effects structure). The fixed effects in a model with only the age of the participants accounted for relatively little variance in the proportion of faces (marginal $R^2$ < `r round(faces_age_only[1,1],5)`) or hands in view  (marginal $R^2$ = `r round(wrists_age_only[1,1],3)`). However, when adding infants' posture and orientation to their caregiver to the model (and their interaction) the marginal $R^2$ were higher for both faces (marginal $R^2$ = `r round(faces_full[1,1],3)`) and wrists (marginal $R^2$ = `r round(wrists_full[1,1],3)`).  
Overall, these results suggest that infants' visual access to social information is largely modulated by their posture and orientation to their caregiver, which is in turn a function of their general locomotor development.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & -4.29 & 0.18 & -23.87 & 0.000 \\ 
  Sit & 0.59 & 0.18 & 3.20 & 0.001 \\ 
  Stand & 0.53 & 0.30 & 1.76 & 0.079 \\ 
  Close & 0.53 & 0.19 & 2.75 & 0.006 \\ 
  Far & 0.15 & 0.26 & 0.56 & 0.574 \\ 
  Age (Scaled) & 0.16 & 0.11 & 1.48 & 0.140 \\ 
  Sit*Close & 0.31 & 0.08 & 3.72 & 0.000 \\ 
  Stand*Close & 0.76 & 0.09 & 8.46 & 0.000 \\ 
  Sit*Far & 0.13 & 0.10 & 1.27 & 0.205 \\ 
  Stand*Far & 1.76 & 0.11 & 15.87 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.} 
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & -3.25 & 0.18 & -18.04 & 0.000 \\ 
  Sit & 0.05 & 0.17 & 0.33 & 0.744 \\ 
  Stand & -0.24 & 0.18 & -1.34 & 0.180 \\ 
  Close & 0.01 & 0.19 & 0.07 & 0.942 \\ 
  Far & 0.45 & 0.27 & 1.63 & 0.102 \\ 
  Age (Scaled) & 0.05 & 0.12 & 0.39 & 0.697 \\ 
  Sit*Close & 0.99 & 0.07 & 14.54 & 0.000 \\ 
  Stand*Close & 1.36 & 0.08 & 16.95 & 0.000 \\ 
  Sit*Far & 0.75 & 0.07 & 10.48 & 0.000 \\ 
  Stand*Far & 1.40 & 0.09 & 16.16 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.} 
\end{table}

```{r summarizeNamings}
source("helper.R")
## summarize detections around naming events
aroundNamings_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.))

aroundNamings_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceOP) %>%
  do(summarize.naming(.))
```

```{r basicNamingStatsFaces}
## inferential stats: namings vs baseline 
aroundNamings_Detections_BySub_Faces <- aroundNamings_Faces %>%
  group_by(subid, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
  mutate(detectorType = 'OpenPose-Faces') %>%
  mutate(when = 'all_namings')

baselineDiffs_by_Naming_Faces <- dlong %>%
  filter(detectorType == 'OpenPose-Faces') %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection)) %>%
  mutate(when = 'baseline') %>%
  full_join(aroundNamings_Detections_BySub_Faces) %>%
  left_join(ages_inc_subs)

baselineDiffs_by_Naming_Faces$when<- factor(baselineDiffs_by_Naming_Faces$when, levels=c('baseline','all_namings'))

naming_faces_model = lmer(prop_detected ~  scale(age.at.test) + when +  (1|subid), data = baselineDiffs_by_Naming_Faces)

faces_naming_out = summary(naming_faces_model)
```

```{r}
#### wrists
aroundNamings_Detections_BySub_Wrists <- aroundNamings_Wrists %>%
  group_by(subid, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
  mutate(detectorType = 'OpenPose-Wrists') %>%
  mutate(when = 'all_namings')

baselineDiffs_by_Naming_Wrists <- dlong %>%
  filter(detectorType == 'OpenPose-Wrists') %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection)) %>%
  mutate(when = 'baseline') %>%
  full_join(aroundNamings_Detections_BySub_Wrists)  %>%
  left_join(ages_inc_subs)

baselineDiffs_by_Naming_Wrists$when<- factor(baselineDiffs_by_Naming_Wrists$when, levels=c('baseline','all_namings'))

naming_wrists_model = lmer(prop_detected ~ scale(age.at.test) + when + (1|subid), data = baselineDiffs_by_Naming_Wrists)

wrists_naming_out = summary(naming_wrists_model)

```

```{r include=FALSE}
## explored looking at namings while acconting for familairity
aroundNamings_Detections_BySub_Fam_Faces <- aroundNamings_Faces %>%
  group_by(subid, familiarity, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
  mutate(detectorType = 'OpenPose-Faces')

aroundNamings_Detections_BySub_Fam_Faces_All <- dlong %>%
  filter(detectorType == 'OpenPose-Faces') %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection)) %>%
  mutate(familiarity = 'baseline') %>%
  full_join(aroundNamings_Detections_BySub_Fam_Faces)  %>%
  left_join(ages_inc_subs)

naming_faces_fam = lmer(data = aroundNamings_Detections_BySub_Fam_Faces_All, prop_detected ~ scale(age.at.test) + familiarity +  (1|subid))
summary(naming_faces_fam)
```

```{r}
## explored looking at namings while acconting for familairity
aroundNamings_Detections_BySub_Fam_Wrists <- aroundNamings_Faces %>%
  group_by(subid, familiarity, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
  mutate(detectorType = 'OpenPose-Wrists')

aroundNamings_Detections_BySub_Fam_Wrists_All <- dlong %>%
  filter(detectorType == 'OpenPose-Wrists') %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection)) %>%
  mutate(familiarity = 'baseline') %>%
  full_join(aroundNamings_Detections_BySub_Fam_Wrists)  %>%
  left_join(ages_inc_subs)

naming_wrists_fam = lmer(data = aroundNamings_Detections_BySub_Fam_Wrists_All, prop_detected ~ scale(age.at.test) + familiarity + (1|subid))
naming_wrists_model_out = summary(naming_wrists_fam)
```

```{r include=FALSE}
# difference plots can be confusing, ommitting
# faces_namings_diffs_indiv <- aroundNamings_Detections_BySub_Fam_Faces_All %>%
#   spread(familiarity, prop_detected) %>%
#   mutate(fam_dif = familiar - baseline, nov_diff = novel - baseline) 
#
# ggplot(faces_namings_diffs_indiv, aes(x=age.grp, y=nov_diff, group=age.grp)) +
#   geom_boxplot() +
#   theme_few()

## explored looking at namings while acconting for posture and orientation
# aroundNamings_Detections_BySub_Posture_Faces <- aroundNamings_Faces %>%
#   group_by(subid, familiarity, posture, orientation, age.grp) %>%
#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
#   mutate(detectorType = 'OpenPose-Wrists')
# 
# baseline_diffs_by_posture <- dlong %>%
#   filter(detectorType == 'OpenPose-Wrists') %>%
#   group_by(age.grp, subid, posture, orientation,detectorType) %>%
#   summarise(prop_detected = mean(detection)) %>%
#   mutate(familiarity = 'baseline') %>%
#   full_join(aroundNamings_Detections_BySub_Posture_Faces)
# 
# naming = lmer(data = aroundNamings_Detections_BySub_Posture_Faces, prop_detected ~ familiarity + posture * orientation + (1|subid))
# summary(naming)
```

```{r}
faces_namings_diffs_out <- aroundNamings_Detections_BySub_Fam_Faces_All %>%
  spread(familiarity, prop_detected) %>%
  mutate(fam_dif = familiar - baseline, nov_diff = novel - baseline) %>%
  group_by(age.grp) %>%
  summarize(fam_vs_baseline = mean(fam_dif, na.rm=TRUE), nov_vs_baseline = mean(nov_diff, na.rm=TRUE), fam_vs_baseline_sd = sd(fam_dif, na.rm=TRUE), nov_vs_baseline_sd = sd(nov_diff, na.rm=TRUE))

wrist_namings_diffs_out <- aroundNamings_Detections_BySub_Fam_Wrists_All %>%
  spread(familiarity, prop_detected) %>%
  mutate(fam_dif = familiar - baseline, nov_diff = novel - baseline) %>%
  group_by(age.grp) %>%
  summarize(fam_vs_baseline = mean(fam_dif, na.rm=TRUE), nov_vs_baseline = mean(nov_diff, na.rm=TRUE), fam_vs_baseline_sd = sd(fam_dif, na.rm=TRUE), nov_vs_baseline_sd = sd(nov_diff, na.rm=TRUE))
```

### Social information during naming events 
Our play session was designed to provide parents with opportunities to label objects -- both familiar and novel -- such that we could examine whether children saw different kinds of social information around naming events. In a set of exploratory analyses, we thus analyzed how face and hand detections changed during object labeling events relative to baseline, analyzing a four-second window around each labeling event (e.g., "Look at the [zem]!"). Every utterance of one of the labelled objects (e.g., 'ball') was counted as a 'labeling event'; timestamps of the beginning of each word were hand-annotated and synchronized with the frame-by-frame detections.

To assess whether there were differences in the social information in view during naming events, we first calculated the proportion of detections that were in view during this four second window, and averaged across naming events for each subject as function of whether the named object as a novel or a familiar objet;  this was then then compared to the baseline proportion of faces in view for each subject in linear mixed-effect models, with random effects of subjects and fixed effects of (scaled) age.

Face detections were numerically slightly higher around these novel naming events relative to baseline, with similar effects across age groups
(8 m.o., $M_{fam - baseline}$ = =`r round(faces_namings_diffs_out$nov_vs_baseline[1],3)`,
12 m.o., $M_{nov - baseline}$ =`r round(faces_namings_diffs_out$nov_vs_baseline[2],3)`,
16 m.o. $M_{nov - baseline}$ =`r round(faces_namings_diffs_out$nov_vs_baseline[3],3)`)
however, they were not higher during familiar naming events vs. baseline (8 m.o., $M_{fam - baseline}$ =`r round(faces_namings_diffs_out$fam_vs_baseline[1],3)`,
12 m.o., $M_{fam - baseline}$ =`r round(faces_namings_diffs_out$fam_vs_baseline[2],3)`,
16 m.o. $M_{fam - baseline}$ =`r round(faces_namings_diffs_out$fam_vs_baseline[3],3)`,
Conversely, wrist detections were higher during both familiar (8 m.o., $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$fam_vs_baseline[1],3)`,
12 m.o., $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$fam_vs_baseline[2],3)`,
16 m.o. $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$fam_vs_baseline[3],3)`)
and novel naming events relative to baseline across all age groups (Novel; 8 m.o., (8 m.o.,  $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$nov_vs_baseline[1],3)`,
12 m.o., $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$nov_vs_baseline[2],3)`,
16 m.o.  $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$nov_vs_baseline[3],3)`). These results were confirmed by a linear mixed-effect model with scaled aged as a fixed effect and random intercepts for each subject (Wrist detections; familiar objects vs. baseline, b=`r round(naming_wrists_model_out$coefficients[3,1],2)`, SE  = `r round(naming_wrists_model_out$coefficients[3,2],2)`, t = `r round(naming_wrists_model_out$coefficients[3,4],2)`, P = `r round(naming_wrists_model_out$coefficients[3,5],4)`'; Novel objects vs. baseline, b=`r round(naming_wrists_model_out$coefficients[4,1],2)`, SE  = `r round(naming_wrists_model_out$coefficients[4,2],2)`, t = `r round(naming_wrists_model_out$coefficients[4,4],2)`, P = `r round(naming_wrists_model_out$coefficients[4,5],4)`.

Overall, these exploratory results suggest that children may tend to see more hands around naming events. This finding is consistent with the possibility that caregivers may change how they interact with their infant when presenting them with objects [@gogate2000study, @gogate2006attention].  For example, caregivers may tend to simultaneously name objects when demonstrating their affordances or simply when pointing to them. In turn, infants may be sensitive to these naming events and orient their attention towards their caregiver, consistent with other accounts positing infants' sensitivity to social cues in early word learning environments [@yurovsky2018communicative, @yurovsky2017beyond, @frank2013].

```{r setUpNamingsPlot}
### Analyze detections around naming events (without posture/orientation)

## first get baseline
# baselineFaces <- dlongPlot %>%
#   filter(detectorType == 'OpenPose-Faces') %>%
#   group_by(age.grp, subid, detectorType) %>%
#   summarise(prop_detected = mean(detection)) %>%
#   mutate(familiarity='baseline')
# 
# baselineWrists <- dlongPlot %>%
#   filter(detectorType == 'OpenPose-Wrists') %>%
#   group_by(age.grp, subid, detectorType) %>%
#   summarise(prop_detected = mean(detection)) %>%
#   mutate(familiarity='baseline')
# 
# 
# ### then by faces
# aroundNamings_Detections_BySub_Faces <- aroundNamings_Faces %>%
#   group_by(subid, familiarity, age.grp) %>%
#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
#   mutate(detectorType = 'OpenPose-Faces') %>%
#   full_join(baselineFaces)
# 
# aroundNamings_Detections_Grouped_Faces <- aroundNamings_Detections_BySub_Faces %>%
#   group_by(familiarity, age.grp) %>%
#   multi_boot_standard(col = 'prop_detected')  
# 
# ### now by wrists 
# aroundNamings_Detections_BySub_Wrists <- aroundNamings_Wrists %>%
#   group_by(subid, familiarity, age.grp) %>%
#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
#   mutate(detectorType = 'OpenPose-Wrists') %>%
#   full_join(baselineWrists)
# 
# aroundNamings_Detections_Grouped_Wrists <- aroundNamings_Detections_BySub_Wrists %>%
#   group_by(familiarity, age.grp) %>%
#   multi_boot_standard(col = 'prop_detected')

```

```{r detByNamingPlotPanels}
## Plot faces/wrists around namings by age

# faces_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
#   # plot averages by familiarity 
#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=2), alpha=.9) +
#   # plot familairity detections for each sub
#   geom_point(data=aroundNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=2), alpha=.2) +
#   theme_few() +
#   labs(y = "Prop. face detections", x = "Age (months)") +
#   theme_few(base_size=plot_size) +
#   ylim(0,.4) +
#   theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
#   scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
#   scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18))

## 
# wrists_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Wrists, aes(x=age.grp, y=mean, color=familiarity)) +
#   # plot averages by familiarity 
#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=2), alpha=.9) +
#   # plot familairity detections for each sub
#   geom_point(data=aroundNamings_Detections_BySub_Wrists, aes(x=age.grp, y=prop_detected), position = position_dodge(width=2), alpha=.2) +
#   theme_few() +
#   labs(y = "Prop. wrist detections", x = "Age (months)") +
#   theme_few(base_size=plot_size) +
#   ylim(0,.4) +
#   theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
#   scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
#   scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) 
```

```{r}
age_labels = c('8' = '8 m.o.','12' = '12 m.o.','16' = ' 16 m.o.')
fam_labels = c('baseline' = 'Baseline','familiar' = "Familiar obj named", 'novel' = 'Novel obj named')

wrist_means_naming <- aroundNamings_Detections_BySub_Fam_Wrists_All %>%
  group_by(familiarity, age.grp) %>%
  multi_boot_standard(col = 'prop_detected')

faces_means_naming <- aroundNamings_Detections_BySub_Fam_Faces_All %>%
  group_by(familiarity, age.grp) %>%
  multi_boot_standard(col = 'prop_detected')

wrists_aroundnamings_plot <- ggplot(aroundNamings_Detections_BySub_Fam_Wrists_All, aes(x=familiarity, y=prop_detected, col=familiarity)) +
  geom_point(alpha=.5) +
  geom_line(aes(group=subid), color='grey', alpha=.2) +
  geom_pointrange(data=wrist_means_naming, aes(y=mean, ymin = ci_lower, ymax = ci_upper), alpha=.9) +
  facet_wrap(~age.grp, labeller = labeller(age.grp = age_labels, familiarity = fam_labels))  +
  scale_x_discrete(labels = fam_labels) +
  ylab('Proportion wrists detected') +
  xlab('') +
  theme_few(base_size=plot_size) +
  theme(legend.position='none') +
  scale_color_viridis(discrete=TRUE) +
  scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6))  +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

faces_aroundnamings_plot <- ggplot(aroundNamings_Detections_BySub_Fam_Faces_All, aes(x=familiarity, y=prop_detected, col=familiarity)) +
  geom_point(alpha=.5) +
  geom_line(aes(group=subid), color='grey', alpha=.2) +
  geom_pointrange(data=faces_means_naming, aes(y=mean, ymin = ci_lower, ymax = ci_upper), alpha=.9) +
  facet_wrap(~age.grp, labeller = labeller(age.grp = age_labels))  +
  ylab('Proportion faces detected') +
  xlab('') +
  theme_few(base_size=plot_size) +
  scale_x_discrete(labels = fam_labels) +
  theme(legend.position='none') +
  scale_color_viridis(discrete=TRUE) +
  scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

```{r detByNaming, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label) for familiar and novel objects; these rates are put into context relative to baseline. Error bars represent 95\\% bootstrapped confidence intervals. Grey lines connect points from individual subjects."}
cowplot::plot_grid(faces_aroundnamings_plot, wrists_aroundnamings_plot, rel_widths=c(1,1))
```

## Extension to Franchak et a., 2017
In the present work, we found that infant's in-the-moment posture changed with their age, as did infants' orientation relative their caregiver. In a related study with 12-month-olds, @franchak2017see also found that infants' in-the-moment posture changed the proportion of time infants spent looking at faces. Here, we sought to replicate these findings using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras [@simon2015databrary] for two reasons. First, we sought to validate our novel method, which could fail to generalize to scenes from these more complex environments, where detecting faces and hands could arguably be a much harder task. Second, we sought to replicate the effects of infants' in-the-moment posture on differences in visual access to hands in an independent dataset.

### Comparison between video datasets 
First, despite having different head-mounted camera setups, the view angle of the two head-mounted cameras were only slightly different ($52.2^{\circ}$ horizontal by $42.2^{\circ}$ in @franchak2017see, here $64^{\circ}$ horizontal by $46^{\circ}$ vertical). However, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to climb, explore, and interact with, and infants were unconstrained and allowed to freely wander the room. In contrast, the play room used in Study 1 was relatively small (approximately 10' x 10') and was set-up with focused play with the pairs of novel and familiar objects. Further, multiple people were present during the play session in @franchak2017see-- including their caregiver and two experimenters -- whereas in Study 1 the experimenters left the room during the play session.

```{r loadFranchakData}
# loads pos_detections used below for franchak dataset
load(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections_by_posture_july2020.RData')
```

```{r processFranchak}
faces_by_pos  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_faces = mean(face_openpose)) %>%
  group_by(infant_posture_label) %>%
  multi_boot_standard(col = 'prop_faces')

faces_by_pos_by_sub  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_faces = mean(face_openpose), faces = sum(face_openpose), no_faces = sum(face_openpose==FALSE))

hands_by_pos  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_hands = mean(hand_openpose)) %>%
  group_by(infant_posture_label) %>%
  multi_boot_standard(col = 'prop_hands')

hands_by_pos_by_sub  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_hands = mean(hand_openpose), hands = sum(hand_openpose), no_hands = sum(hand_openpose==FALSE))
```

```{r statsFranchak}
# glmers for franchak dataset
posture_model_faces_franchak <- glmer(cbind(faces,no_faces) ~ infant_posture_label + (1|sub_id), data=faces_by_pos_by_sub, family="binomial")

posture_model_hands_franchak <- glmer(cbind(hands,no_hands) ~ infant_posture_label + (1|sub_id), data=hands_by_pos_by_sub, family="binomial")

franchak_faces_out <- summary(posture_model_faces_franchak)
franchak_wrists_out <- summary(posture_model_hands_franchak)
```

```{r plotPanelsFranchak}
plot_size = 18
faces_by_pos_plot <- ggplot(faces_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
   geom_point(data=faces_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label), alpha=.1) +
  geom_line(data=faces_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label, group=sub_id), alpha=.1) +
  theme_few(base_size=plot_size) +
  ylab('Prop faces detected') +
  xlab('Infant posture') + 
  ylim(c(0,.3)) +
  theme(legend.position = 'none') +
  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928"))

hands_by_pos_plot <- ggplot(hands_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  geom_point(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_hands, col=infant_posture_label), alpha=.1) +
  geom_line(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_hands, col=infant_posture_label, group=sub_id), alpha=.1) +
  theme_few(base_size=plot_size) +
  ylab('Prop hands detected') +
   xlab('Infant posture') + 
  ylim(c(0,.4)) +
  theme(legend.position = 'none') +
  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928")) 
```

```{r franchak, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=4, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals."}
cowplot::plot_grid(faces_by_pos_plot,hands_by_pos_plot, nrow=1)
```

### Replication using automated detections
```{r}
detections_by_conf_by_sub <- pos_detections %>%
  mutate(faces_strict = nose_conf > .5) %>%
  mutate(faces_strict = replace_na(faces_strict, FALSE)) %>%
  group_by(sub_id) %>%
  summarize(avg_faces_by_sub = mean(face_openpose),
            avg_hands_by_sub = mean(hand_openpose), avg_faces_by_sub_strict = mean(faces_strict))

dets_by_conf_by_posture <- pos_detections %>%
  mutate(faces_strict = nose_conf > .5) %>%
  mutate(faces_strict = replace_na(faces_strict, FALSE)) %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(avg_faces_by_sub_strict = mean(faces_strict), avg_faces_by_sub = mean(face_openpose), avg_hands_by_sub = mean(hand_openpose)) %>%
  group_by(infant_posture_label) %>%
  summarize(avg_faces_strict = mean(avg_faces_by_sub_strict),
            avg_faces = mean(avg_faces_by_sub), avg_hands =  mean(avg_hands_by_sub))

```

Overall, we found convergence between our two methodologies, replicating the main results from @franchak2017see, and finding that the proportion of faces detected was greater when infants were sitting or standing vs. prone. We found this result regardless of whether we used all detections (see Figure \ref{fig:franchak}, panel A; (proportion of frames with face detections; Prone: M =`r mean(dets_by_conf_by_posture$avg_faces[1])`%, Sitting: M =`r mean(dets_by_conf_by_posture$avg_faces[2])`%, Upright: M =`r mean(dets_by_conf_by_posture$avg_faces[3])`%) or restricted our analyses to high-confidence detections (percentage of frames with face detections; Prone: M =`r mean(dets_by_conf_by_posture$avg_faces_strict[1])`, Sitting: M =`r mean(dets_by_conf_by_posture$avg_faces_strict[2])`, Upright: M =`r mean(dets_by_conf_by_posture$avg_faces_strict[3])`). These results were confirmed in generalized linear-mixed models, as in the first study (Sitting vs. Prone, b=`r round(franchak_faces_out$coefficients[2,1],3)`, SE  = `r round(franchak_faces_out$coefficients[2,2],3)`, Z = `r round(franchak_faces_out$coefficients[2,3],3)`, P = `r round(franchak_faces_out$coefficients[2,4],4)`. Upright vs. Prone, b=`r round(franchak_faces_out$coefficients[3,1],3)`, SE  = `r round(franchak_faces_out$coefficients[3,2],3)`, Z = `r round(franchak_faces_out$coefficients[3,3],3)`, P = `r round(franchak_faces_out$coefficients[3,3],4)`).

The most noticeable difference between our analysis methods is the overall proportion of frames with faces detected by OpenPose vs. the proportion of frames in which infants were judged to be looking at faces in @franchak2017see. Across the entire session in @franchak2017see, infants looked at faces on 4.7% of frames [@franchak2017see]. When we used all detections from OpenPose, we found a higher proportion of faces:  M = `r mean(detections_by_conf_by_sub$avg_faces_by_sub)`, or `r mean(detections_by_conf_by_sub$avg_faces_by_sub)*100`%.  When we restricted our results to only high-confidence detections, we found M = `r mean(detections_by_conf_by_sub$avg_faces_by_sub_strict)`, closer to the original values reported by @franchak2017see. However, the above analyses on the accuracy of these two methods suggest that high-confidence detections dramatically underestimate the number of faces in view. Thus, we suspect that these differences are due to both (1) the fact head-mounted eye-trackers may sometimes underestimate the proportion of faces attended due to calibration issues, and that (2) OpenPose may sometimes detect faces that are in view for infants but that infants may not be foveating.

We also found that infants' in-the-moment posture also modulated the proportion of hands that were in view (i.e., wrist detections), though these were not originally annotated in @franchak2017see (see Figure \ref{fig:franchak}, (proportion of frames with wrist detections; Prone: M =`r mean(dets_by_conf_by_posture$avg_hands[1])`; Sitting: M =`r mean(dets_by_conf_by_posture$avg_hands[2])`; Upright:  M =`r mean(dets_by_conf_by_posture$avg_hands[3])`). These results were confirmed in generalized linear-mixed models, as in the first study (Sitting vs. Prone, b=`r round(franchak_wrists_out$coefficients[2,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[2,2],3)`, Z = `r round(franchak_wrists_out$coefficients[2,3],4)`, P = `r round(franchak_wrists_out$coefficients[2,4],4)`. Upright vs. prone, b=`r round(franchak_wrists_out$coefficients[3,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[3,2],3)`, Z = `r round(franchak_wrists_out$coefficients[3,3],3)`, P = `r round(franchak_wrists_out$coefficients[3,3],4)`).

Overall, these analyzes extend and validate previous work, replicating the results in Study 1 that infants' in-the-moment posture modulated the proportion of hands in view, and suggesting that posture is a major factor that structures infants' access to visual information broadly construed.

# General Discussion
What social cues do infants see as they learn language, and how does this change as they grow and start to locomote themselves in their environments? We explored this question using video data from head-mounted cameras from two naturalistic datasets of parent-child interactions: a cross-sectional database of naturalistic play sessions from 8-16 month-olds with sets of novel and familiar toys, and database of head-mounted camera videos from one-year-olds who explored a large play area from @franchak2017see. To analyze these datasets, we developed a novel method using a pose detection model to automate the annotation of the social information in the infant view, here operationalized as the presence of the faces and hands of their caregiver; these annotations were then synced with manual annotations of infants in-the-moment posture from third-person videos.

Despite being trained on the adult perspective, the pose detector we used (OpenPose, @cao2017realtime) was able to generalize relatively well to the infant viewpoint, achieving comparable precision and accuracy as a face detector relative to a state-of-the-art model optimized specifically for detecting faces in natural scenes (MTCNN, @zhang2016].  While OpenPose had relatively low recall as a hand detector -- missing some hands that were in the infant view -- it made comparable rates of false alarms. In both cases, we found that overall performance was maximized when all detections were included, regardless of their confidence, suggesting that some low-confidence face and hand detection still index actual faces and hands that were seen by infants.  Thus, while imperfect, we suggest that OpenPose can be applied to infant egocentric videos for the extraction of the social information in the infant viewpoint, reducing the burden of manual annotations and promoting the reusability of rich video datasets for further analyses. The use of this automated methodology allowed us to easily annotate the entirety of our dataset -- additionally analyzing the social information around naming events -- and to re-analyze the data from @franchak2017see to replicate our findings in a very different kind of play session.

Broadly, our results replicate and extend previous work, first by showing systematic changes in infants' in-the-moment posture and their orientation relative to their caregivers  [@adolph2017development]; older children spent more time standing and less time sitting, and older infants' caregivers spent less time supporting their standing or sitting postures. Motor development changes dramatically at the same time that children are breaking into language learning.  Using these automated detections, we found that infants' changing posture and orientation to their caregiver jointly shaped the amount of social information that was in their view during one-on-one play sessions with their caregivers: Children saw the most faces/hands when they were sitting or standing and close to their caregiver vs. crawling or prone. These same findings were recapitulated in a second dataset collected by @franchak2017see with one-year-olds: sitting and upright infants saw more faces--and hands--than infants who were prone. Motor development appears to modulate how infants experience their visual world and the social information in it.

While exploratory, our results also suggest that infants saw a greater proportion of hands during naming events, hinting that children may have been orienting towards their caregiver when they heard labels for objects. While this effect was not present for faces, other work [@yoshida2008; @yu2013joint; @yu2017hand], including @franchak2017see, has found that infants spend much more time looking at the toys vs. their caregiver's faces during these play sessions, and highlighted the importance of hand-following as a component of joint attention [@yu2017hand]. Further, given that there were only two possible referents in the room at a time—-and one of them was always a familiar category (i.e., car, kitty) -- this particular play session did not present many opportunities where children would need to use gaze cues to disambiguate referents.

Importantly, however, all of these findings come from observational, in-lab datasets, posing limits on their generalizability. Future work is needed to relate these the slices of experience captured within in-lab play sessions and infants' everyday experiences [@fausey2016; @clerkin2017]. More broadly, though observational findings allow us to document developmental change and identify potential causal pathways, they cannot confirm them. Locomotive abilities are of course only part of a cascade of changes in infants' abilities and experiences, and these analyzes document only a fraction of this broader, multifaceted trajectory. As children grow and change, the activities in which they engage with their caregivers are likely to also vary, leading to differences in the distribution of social cues that they experience that may not be captured in these in-lab play sessions.

Nonetheless, we think that this approach hold promise for documenting these developmental trajectories and for generating hypotheses about potential causal pathways. Indeed, understanding the relationship between different domains of developmental changes in naturalistic contexts has been a persistent challenge for developmental psychology. However, the field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools that deal better with noisier, more complicated datasets and extract richer information.  We hope that these new tools can now be leveraged to examine the consequences of the changing infant perspective for linguistic, cognitive, and social development.

# Acknowledgements

Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation. This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and NSF #1714726 to BLL. An earlier version of this work was presented to the Cognitive Science Society in @frank2013 and @sanchez2018postural.

\newpage

# References
```{r create_r-references}
r_refs(file = "xsface.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2020-08-11 writeup-paper/draft_apa_july2020.tex
latest version of writeup

\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Using automated detections to quantify how postural developments shape the social information in view for young infants},
            pdfkeywords={Postural developments, headcam, social information, face detection, pose
detection},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Automated detection of the social information}
\author{Bria L. Long\textsuperscript{1}, Alessandro Sanchez\textsuperscript{1}, Allison M. Kraus\textsuperscript{1}, Ketan Agrawal\textsuperscript{1}, \& Michael C. Frank\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\authornote{

Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
\keywords{Postural developments, headcam, social information, face detection, pose detection\newline\indent Word count: XXX}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}

\title{Using automated detections to quantify how postural developments shape
the social information in view for young infants}

\date{}

\abstract{
Social information plays a key role in infants' early linguistic and
cognitive development. Yet infants locomotor abilities also develop
continuously as they learn to sit, crawl, stand, and walk -- influencing
the way that infants see the world around them. How do these postural
developments affect infants access to the social information relevant
for early learning? To address this question, we first created a rich
annotated dataset of naturalistic play sessions between infants and
their caregivers with sets of novel and familiar toys (N=36, 8-16 months
of age), recording both egocentric and third-person video. We then
developed an automated analysis method using a pose detection model
(Zhang et al., 2017) to detect the faces and hands in the infant view,
allowing us to analyze the entirety of this dataset. We also applied
this automated method to second egocentric video dataset where
one-year-olds explored a large play area with their caregivers (Franchak
et al., 2017). We found that infants' posture and orientation to their
caregiver changed dramatically across this age range and modulated
infants access to social information; infants who were sitting or
standing with their caregiver at a close distance tended to have the
most faces and hands in their visual field. We also found convergence
across both datasets, confirming the finding that motoric developments
play a significant role in shaping the social information that infants
have access to over development, and suggesting that they may play a
role in emergence of infants' cognitive and linguistic abiltiies. We
suggest that the combined use of head-mounted cameras and the
application of new computer vision techniques is a promising avenue for
understanding the statistics of infants' visual and linguistic
experience as they change over development and for re-utilizing rich
video datasets collected across different labs and diverse contexts.
}

\begin{document}
\maketitle

From their earliest months, infants are deeply engaged in learning from
others. Even newborns tend to prefer to look at faces with direct
vs.~averted gaze (Farroni, Csibra, Simion, \& Johnson, 2002) and young
infants follow overt gaze shifts (Bruner, 1975; Gredeback, Fikke, \&
Melinder, 2010). And as infants reach their first birthday, they also
tend to follow (Yu \& Smith, 2013, 2017) and imitate the gestures of
their caregivers (e.g., pointing).

Infants ability to process these social cues may provide strong
scaffolding for early word learning. Longitudinal studies provide some
evidence for this link: children's level of joint engagement with their
mother at 9-12 months predicts both their receptive and productive
vocabularies (Carpenter, Nagell, \& Tomasello, 1998) and 10 month-olds
who follow an adult's gaze (in an experimental context) have larger
vocabularies at 18 months and throughout the second year of life (Brooks
\& Meltzoff, 2005, 2008). While the relationship between hand-following
and language development has been less well characterized, infants who
follow their caregivers' hands tend to be those who spend more time
jointly attending to events with their caregivers (Yu \& Smith, 2017).

Relatively little work, however, has quantified how often infants see
and use these kinds of social cues in naturalistic learing environments.
By using head-mounted cameras to record what infants see, researchers
have begun to document the infant egocentric perspective (Yoshida \&
Smith, 2008) and to quantify the information -- social and otherwise --
available to infants as they learn. While head-mounted camera data do
not provide explicit information on where infants are attending, some
work suggests that infants orient their head towards what they are
focusing on -- putting those people or objects in view (Yoshida \&
Smith, 2008). Initial recordings from in-lab play sessions revealed a
different view than many imagined: instead of being dominated by their
faces, the infant perspective contained close up views of primarily toys
and hands (Franchak, Kretch, Soska, \& Adolph, 2011; Yoshida \& Smith,
2008; Yu \& Smith, 2017). Subsequent research has revealed that the
infant view is far from a unitary construct, undergoing dramatic changes
as infants grow. Recordings from home environments suggest that the
viewpoints of very young infants -- less than 4 months of age -- do
indeed contain persistent and frequent views of faces (Fausey,
Jayaraman, \& Smith, 2016; Jayaraman, Fausey, \& Smith, 2017) but that
the infant view tends to contain more and more hands as infants grow
older.

These changes in perspective are likely the downstream consequence of a
myriad factors, chief of which may be infants' evolving locomotive
abilities: an infants ability to sit, crawl, stand, or walk structures
the way they interact with the things and people in their world. These
changes can be rapid: the physical body of an infant can lengthen up to
2cm in 24 hours (Lampl, 1993), leading to changes in their motoric
abilities from one day to the next. These motoric developments can be
thought of as gateways that open up entirely new phases of development
(Iverson, 2010), and cause a cascade of changes in infants ability to
interact with their world and the people in it (Karasik, Tamis-LeMonda,
\& Adolph, 2014).

Thus, infants' changing locomotor abilities may shape the social cues
that infants' see and seek out, in turn impacting their cognitive and
linguistic abilities. Supporting this idea, infants experience with
sitting predicts their success at 3D object completion tasks (Soska,
Adolph, \& Johnson, 2010) as well as their receptive vocabulary
(Libertus \& Violi, 2016). And as children begin crawling (Adolph,
Vereijken, \& Denny, 1998) -- or scooting or cruising (Patrick, Noah, \&
Yang, 2012) -- they are no longer constrained to the same spot that
their caregivers last placed them in. Yet while crawlers can choose
where to go and what they see to a much greater degree, they also appear
to spend much of their time in a world populated by floors and knees;
during spontaneous play, toddlers are more likely to look at the floor
while crawling than while walking (Franchak et al., 2011), when they
have full visual access to their environment and the people in it
(Kretch, Franchak, \& Adolph, 2014).

On one view, it is children's ability to stand and walk that
fundamentally changes their ability to access social information (e.g.,
facial expressions, gaze cues, pointing) relative to children who are
still crawling and sitting (Walle, 2016). In turn, this ability to
access more detailed social information could allow infants to learn
words quicker and more efficiently, facilitating language growth (Walle,
2016). Supporting this idea, walking infants make different kinds of
object-related bids for attention from their caregivers than crawling
infants, and tend to hear more action directed statements (e.g.,
\enquote{open it}) (Karasik et al., 2014); further, in an observational
study, Walle and Campos (2014) also found that children who were able to
walk had both higher receptive and productive vocabularies. However, not
all evidence supports this view: some parental report data suggest a
weak relationship between walking and the onset of language (Moore,
Dailey, Garrison, Amatuni, \& Bergelson, 2019), and head-mounted
eye-tracking data have highlighted the role of infants' in-the-moment
posture -- and that of their caregivers' -- in what infants' see,
finding that these factors are better predictors of the social
information in view than their locomotor status per se (i.e.,
\enquote{walker} vs \enquote{crawler}) (Franchak, Kretch, \& Adolph,
2017).

Understanding this nuanced relationship between the social information
in the infant view and their motoric and linguistic development has
remained challenging. Indeed, different kinds of play sessions (e.g.,
exploring novel environments vs.~playing with novel objects) might
modulate the degree to which infants' posture changes the social
information in view. Furthermore, no work to date has directly examined
how the availability of social information changes in naturalistic
language learning contexts. For example, it is unclear if and how the
social information in view changes during the naming events for
different objects (e.g, \enquote{Look, a {[}ball{]}}).

More broadly, the field is in need of computational tools to reutilize
these rich head-mounted camera datasets and to understand the
generalizability of findings across different datasets, populations,
tasks, and age-ranges. While the field has assembled many head-mounted
camera datasets, conducting new analyses on these videos has remained
prohibitively time-consuming due to a lack of computational tools for
annotations. Instead, hundreds of hours of manual annotations have been
required to analyze a fraction of the available frames for a given
analysis. Thus, despite containing a wealth of information about the
structure of parent-child interactions, these datasets have thus gone
dramatically underused.

Recent innovations in computer vision hold promise for automated
annotations of the infant view. Over the past decade, deep neural
networks have become dramatically better at a wide range of visual
tasks, including object classification, Simonyan and Zisserman
(2014){]}, scene categorization (Zhou, Lapedriza, Khosla, Oliva, \&
Torralba, 2017), and pose detection (K. Zhang, Zhang, Li, \& Qiao,
2016), arguably facilitating our understanding of visual perception
(Peterson, Abbott, \& Griffiths, 2018, VanRullen (2017)) and improving
computational neuroscience (Kietzmann, McClure, \& Kriegeskorte, 2018).
However, as most models have been trained on photographs or videos taken
from the adult perspective, it is still unclear how easily these models
can be applied to videos taken from the infant perspective.

In this paper, we thus first validate a model for the automated
detection of faces and hands in the infant view, using a open-source
model of pose detection -- OpenPose(Cao, Simon, Wei, \& Sheikh, 2017; K.
Zhang et al., 2016) -- that provides pose, face, and hand keypoints for
every person in an image. We compare the detection accuracy of OpenPose
with that of both older and more specialized models of face detection
and demonstrate the usability of this off-the-shelf model for
quantifying the faces and hands in the infant view.

We then apply this automated method to a head-mounted camera dataset
collected in our lab to directly examine the role of postural
developments on the social information in the infant view during
language learning. Specifically, we constructed a dataset where infants
and caregivers played freely using a set of novel and familiar objects;
8, 12, and 16 months-olds (N=26; N=12 in each age group) participated to
ensure a wide range of locomotor abilities. Infants' in-the-moment
posture and orientation relative to their caregiver were hand annotated
and all play sessions were transcribed. This cross-sectional design thus
allowed us to directly examine the relative contributions of age
vs.~postural developments on children's visual access to social
information, and how the avaliability of social cues changes relative to
naming events (e.g., \enquote{Yes, you like the {[}ball!{]}}). Broadly,
we predicted that there would be differential access to social
information based on children's postural developments: crawling infants
would see fewer faces/hands because they would primarily be looking at
the ground, while walking toddlers would have access to a richer visual
landscape with greater access to the social information in their
environment.

To examine the robustness of our automated method and the
generalizability of our findings, we applied OpenPose (Cao et al., 2017)
to the dataset collected in Franchak et al. (2017), where one-year-olds
wore head-mounted eye-tracking cameras during a play session and their
in-the-moment posture was hand-annotated. Unlike the present dataset,
infants and caregivers roamed a large, open playroom and explored
different toys placed throughout. We thus analyzed this dataset with the
goal of both (1) validating our automated method on a very different
kind of dataset, attempting to replicate their primary findings
originally obtained with head-mounted eye-tracking (2) extending their
findings to ask whether whether the proportion of hands in view also
varies with infants' posture in an exploratory play session.

\section{Methods}\label{methods}

\subsection{Free-play video dataset}\label{free-play-video-dataset}

A primary goal of this project was to provide a rich, open dataset for
other researchers to examine the effects of postural developments and
labeling behavior during naturalistic parent-child interactions. We thus
invited caregivers of 8, 12, and 16-month-olds to participate in play
sessions where they were provided with pairs of novel and familiar
objects (e.g., a ball and a microfiber duster, called a \enquote{zem})
in a small playroom in a lab (approximately 10' x 10'). Infants wore
head-mounted cameras equipped with a fish-eye lens (see Head-mounted
camera), a third-person camera captured a birds-eye view of the play
session. Using these video data, infants' posture and orientation to
their caregiver were hand-coded and annotated for the entirety of the
play session. All videos were transcribed, and we collected collected
MacArthur CDIs for all children who particpated. All materials have been
made publicly avaliable on Databrary for whom the parents provided
sharing consent (29/36 dyads) via
\url{https://nyu.databrary.org/volume/101}

\subsubsection{Participants}\label{participants}

Our final sample consisted of 36 infants and children, with 12
participants in three age groups: 8 months (6 F), 12 months (7 F), and
16 months (6 F). Participants were recruited from the surrounding
community via state birth records, had no documented disabilities, and
were reported to hear at least 80 percent English at home. Demographics
and exclusion rates are given in the table below (see Table 1)/

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
\hline
Group & N & \% incl. & Avg age & Avg video length (min) \\
\hline
8 mo. &   12 & 0.46 & 8.71 & 14.41 \\
12 mo. &  12 & 0.40 & 12.62 & 12.71 \\
16 mo. &  12 & 0.31 & 16.29 & 15.10\\
\hline
\end{tabular}
\caption{Exclusion rates and summary demographics for the infants included in the study.} 
\end{table}

To obtain this final sample, we tested 95 children, excluding 59
children for the following reasons: 20 for technical issues related to
the headcam (e.g., failure to record, ran out of battery), 15 for
failing to wear the headcam, 10 for fewer than 4 minutes of headcam
footage, 5 for having multiple adults present, 5 for missing
Communicative Development Inventory (CDI) data, 2 for missing scene
camera footage, 1 for fussiness, and one for sample symmetry.
16-month-olds tolerated the head-mounted camera less well than younger
infants, leading to a higher exclusion rate. All inclusion decisions
were made independent of the results of subsequent analyses.

\subsubsection{Head-mounted camera}\label{head-mounted-camera}

\begin{figure}[H]
\includegraphics[width=0.4\linewidth]{images/view_angle_redone} \caption{Vertical field of view for two different headcam configurations (we used the lower in our current study).}\label{fig:headcam}
\end{figure}

We used a small, head-mounted camera (\enquote{headcam}) that was
constructed from a MD80 model camera attached to a soft elastic
headband. Videos captured by the headcam were 720x480 pixels with 25
frames per second.\footnote{Detailed instructions for creating this
  headcam can be found at
  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
A fisheye lens was attached to the camera to increase the view angle
from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
\ref{fig:headcam}, above).

Even with the fish-eye lens, the vertical field of view of the camera is
still considerably reduced compared to the child's field of view, which
spans around 100--120\(^{\circ}\) in the vertical dimension by 6-7
months of age (Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton,
1988; Mayer, Fulton, \& Cummings, 1988). As we were originally primarily
interested in the presence of faces in the child's field of view, we
chose to orient the camera upwards to capture the entirety of the
child's upper visual field where the child is likely to see adult faces,
understanding that this decision limited our ability to detect hands
(especially those of the child, which are typically found at the bottom
of the visual field).

\subsubsection{Procedure}\label{procedure}

All parents signed consent documents while children were fitted with the
headcam. If the child was uninterested in wearing the headcam or tried
to take it off, the experimenter presented engaging toys to try to draw
the child's focus away from the headcam. When the child was comfortable
wearing the headcam, the child and caregiver were shown to a playroom
for the free-play session. Parents were shown a box containing three
pairs of familiar and novel objects. These pairs consisted of a ball
paired with a microfiber duster (a \enquote{zem}), a toy car paired with
a cheese grater (a \enquote{manu}), and a brush paired with a back
massager (a \enquote{tima}). Parents were instructed to play with the
object pairs with their child one at a time, \enquote{as they typically
would.} {[}\^{}3{]}:The first few participants played with a different
pair of objects (a toy cat and pedicure foam piece) that was replaced as
some infants chewed persistently on the foam.

All parents confirmed that their child had not previously seen the novel
toys and were instructed to use the novel labels to refer to the toys.
The experimenter then left the playroom for approximately 15-20 minutes,
during which a tripod-mounted camera in the corner of the room recorded
the session and the headcam captured video from the child's perspective.

\subsubsection{Data processing and
annotations}\label{data-processing-and-annotations}

Headcam videos were trimmed such that they excluded the instruction
phase when the experimenter was in the room and were automatically
synchronized with the tripod-mounted videos using FinalCut Pro Software.
These sessions yielded 507 minutes (almost a million frames) of video,
with an average video length of 14.07 minutes (min = 4.53, max = 19.35).

\paragraph{Posture and caregiver orientation
annotations}\label{posture-and-caregiver-orientation-annotations}

We created custom annotations to describe the child's physical posture
(i.e.~standing) and the orientation of the caregiver relative to the
child (e.g.~far away). The child's posture was categorized as being
carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations; similarily, times when the child was being carried or the
caregivers were out of the frame were marked as uncodable for caregiver
orientation. On average, posture or orientation was uncodable from 1-2
minutes of data in each child (seconds excluded from analysis for
posture;. M =94.01, SD =225.18; orientation; M =94.01, SD =225.18), and
these rates did not vary substantially with the age of the child. To
assess the reliability of these annotations, a second coder coded videos
from three different children to calculate Cohen's kappa (posture, irr =
XX, orientation, irr = XX).

\paragraph{Naming event annotations}\label{naming-event-annotations}

One coder listened to all of the audio from the play sessions and marked
the exact timestamps whenever one of the novel or familiar objects was
named in any instance (e.g., \enquote{Look at the {[}ball{]}},
\enquote{Can you say {[}zem{]}?}); a second coder listened to the
majority of the play sessions (N=23 sessions) and also annotated all
naming events. To asesess reliability, we calculated the proportion of
naming events detected by the first coder that were also annotated by a
second coder within a sliding window. We found that 82.1\% of naming
events were detected within a 4 second window (+/- 2s), and 70.9\% of
namings events were detected within a 2 second window. We also obtained
full text transcriptions of the entire play sessions (with time stamps
marking 10s intervals). While these full transcriptions are not used in
the present analyses, they have been made publically avaliable for
future research.

\subsection{Face and hand detection}\label{face-and-hand-detection}

We evaluated three detection systems for the ability to measure infants'
access to faces. The first of these is the most commonly-used and widely
available face detection algorithm: Viola-Jones (Viola \& Jones, 2004).
We used this algorithm as a benchmark for performance, as while it can
achieve impressive accuracy in some situations, it is notoriously bad at
dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox, 2014).
We next tested the performance of two face detectors that both made use
of relatively recently developed Convolutional Neural Networks (CNNs) to
extract face information. The first algorithm was specifically optimized
for face detection, and the second algorithm was optimized to extract
pose information of all the individuals in an image, operationalized as
information about the position of 18 different body parts. For this
second algorithm (OpenPose; Cao et al., 2017), we used the agent's nose
(one of the body keypoints detected) to operationalize the presence of
faces, as any half of a face necessarily contains a nose.

The OpenPose detector also provided us with the location of an agent's
wrists, which we used as a proxy for hands for two reasons. First, as we
did not capture children's entire visual field, the presence of a wrist
is likely often indicative of the presence of a hand within the field of
view. Second, hands are often occluded by objects when caregivers are
interacting with children, yet still visually accessible by the child
and part of their joint interaction.

\subsubsection{Algorithms}\label{algorithms}

Viola Jones, the first face detection system, made use of a series of
Haar feature-based cascade classifiers (Viola \& Jones, 2004) applied to
each individual frame. The second algorithm (based on work by K. Zhang
et al. (2016)) uses multi-task cascaded convolutional neural networks
(MTCNNs) for joint face detection and alignment, built to perform well
in real-world environments where varying illuminations and occlusions
are present. We used a Tensorflow implementation of this algorithm
available at \url{https://github.com/davidsandberg/facenet}.

The CNN-based pose detector (OpenPose; Cao et al., 2017; Simon, Joo,
Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016)
provided the locations of 18 body parts (ears, nose, wrists, etc.) and
is available at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a convolutional neural network for initial anatomical
detection and subsequently applies part affinity fields for part
association, producing a series of body part candidates. The candidates
are then matched to a single individual and finally assembled into a
pose; here, we only made use of the body parts relevant to the face and
hands (nose and wrists), though the entire set of keypoints is publicly
avaliable. Each keypoint was accompanied by a confidence score made by
the detector.

\subsubsection{Detector evaluation}\label{detector-evaluation}

To evaluate face detector performance, we hand-labeled a \enquote{gold
set} of frames extracted from the video dataset. To account for the
relatively rare appearance of faces in the dataset, we hand-labeled two
types of samples: a sample containing a high density of faces (half
reported by MTCNN, half by OpenPose) and a random sample from the
remaining frames. Each sample was comprised of an equal number of frames
taken from each child's video. For wrist detections, the \enquote{gold
set} was constructed in the same manner, except frames with a high
density of wrists came only from detections made by OpenPose. Faces were
classified as present if at least half of the face was showing; wrists
were classified as present if any part of the wrist was showing. Two
authors labelled the frames independently and resolved disagreements on
a case-by-case basis. Precision (hits / hits + false alarms), recall
(hits / hits + misses), and F-score (harmonic mean of precision and
recall) were calculated for all detectors.

\section{Results}\label{results}

First, we report the accuracy of the automated detectors, as assessed by
comparison to hand-labelled frames from the free-play video dataset
described above. We then apply one of these automated detectors
(OpenPose) to the entirety of this video dataset, and use these outputs
to examine how postural developments influence children's visual access
to faces and hands from 8-16 months of age, as well as how access to
these social cues changes during labeling events (e.g., do you see the
{[}zem{]}?). Finally, we apply this same automated detector to another
video dataset (Franchak et al., 2017), replicating and extending their
findings on the effects of posture on visual access to social
information.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{images/exampe_detections} \caption{Example detections made by OpenPose from children in each age group.}\label{fig:exampledetections}
\end{figure}

\subsection{Accuracy of automated
detections}\label{accuracy-of-automated-detections}

For face detection, we found that both OpenPose and MTCNN dramatically
outperformed ViolaJones, our baseline model, especially with respect to
the random sample, where ViolaJones generated a relatively high
proportion of both false negatives and false positives. When considering
only the composite F-score across all frames, MTCNN slightly
outperformed OpenPose (0.89 MTCNN vs.~0.83 OpenPose), and MTCNN and
OpenPose performed comparably with the random sample. Generally, MTCNN
exhibited higher precision, whereas OpenPose exhibited higher recall,
and these differences were most pronounced on the randomly sample
frames. In other words, while OpenPose generated slightly more false
positives than MTCNN, MTCNN missed several faces that were accurately
detected by OpenPose. When we restricted our analysis to high-confidence
detections from OpenPose (\textgreater{}.5 confidence; default threshold
for visualization), we found very high precision (P = 0.97), but much
lower recall (R = 0.64) and thus overall performance (F = 0.77),
indicating that these low-confidence detections often indexed actual
faces that were in the infant view. Figure \ref{fig:exampledetections}
shows example successful detections from OpenPose in each age group, and
Figure \ref{fig:failures} shows examples of missed faces as well as
false positives for context.

We next analyzed the viability of OpenPose as a hand detector. Despite
the fact that hand detection is a more computationally challenging
problem (Bambach, Crandall, Smith, \& Yu, 2017), and the fact that we
used wrist keypoints as a proxy for hands, OpenPose performed moderately
well as a hand detector (F = 0.74). OpenPose achieved relatively high
precision -- generating relatively few false positives -- but showedlow
recall on the randomly sampled frames (see Table 1). As with face
detections, when we restricted our analysis to high-confidence
detections, we found much higher precision (P = 0.95), but much lower
recall (R = 0.37) and thus overall performance (F = 0.53).

Thus, one major advantage of OpenPose relative to specialized face
detectors, such as MTCNN, is that it allows the analysis of both the
faces and hands in the infant view with the outputs of only one
algorithm, and analyzing the results of all detections (regardless of
confidence) yielded reasonably accurate results. Going forward, we
analyze face and wrist detections using all detections from OpenPose,
with the caveat that we are likely underestimating the proportion of
hands in the dataset given the lower recall for hand detections.

\begin{figure}[H]
\includegraphics[width=0.5\linewidth]{images/example_failures} \caption{Example failed detections from OpenPose, showing both false positives and false negatives (i.e. missed detections).}\label{fig:failures}
\end{figure}

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
\hline
Algorithm & Sample\ Type & P & R & F \\ 
\hline
MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
OpenPose-Wrists & High density & 0.66 & 1.00 & 0.79 \\ 
OpenPose-Wrists & Random & 0.88 & 0.29 & 0.44 \\ 
\hline
\end{tabular}
\caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected). P, R, and F denote precision, recall, and F-score, respectively.} 
\vspace{-1em}
\end{table}

\subsection{Access to social information during postural
developments}\label{access-to-social-information-during-postural-developments}

\begin{figure}[H]
\includegraphics[width=\textwidth]{figs/posture-1} \caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:posture}
\end{figure}

\begin{figure}[H]
\includegraphics[width=\textwidth]{figs/indivPosOrient-1} \caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times when infant was carried or when posture/orientation were not codable are ommitted for visualization purposes}\label{fig:indivPosOrient}
\end{figure}

\subsubsection{Developmental changes in infant posture and caregiver
orientation}\label{developmental-changes-in-infant-posture-and-caregiver-orientation}

We report developmental shifts in infants' posture and their orientation
relative to their caregiver, consistent with previous literature (Adolph
\& Berger, 2006; Franchak et al., 2017). The proportion of time infants
spent sitting decreased with age, and the proportion of time infants
spent standing increased with infants' age. As children got older, their
locomotor abilities allowed them to become more independent. Both
8-month-olds and 12-month-olds spent relatively equivalent amounts of
time lying/crawling (i.e., \enquote{prone}) which was markedly decreased
in the 16-month-olds, who spent most of their time sitting or standing
(see Figure \ref{fig:posture}). We also observed changes in infants'
orientation relative to their caregivers: the 8-month-olds spent more
time with their caregiver behind them supporting their sitting positions
than did children at other ages (see Figure \ref{fig:posture}). However,
we also saw considerable variability across children: some infants spent
almost their entire time sitting at a close distance from their
caregiver, whereas others showed more considerable variability (see
Figure \ref{fig:indivPosOrient}).

\subsubsection{Access to faces and
hands}\label{access-to-faces-and-hands}

First, we examined the proportion of face and hand detections as a
function of infants' age without considering their posture (see Figure
\ref{fig:detByAge}). While faces tended to be in the field-of-view
overall more often than hands, infants' head-mounted cameras were angled
slightly upward to capture the presence of faces, and hand detections
suffered from somewhat lower recall than face detections. We thus only
considered differences in the relative proportion of faces or hands in
view as a function of age, posture, and orientation, rather than
comparing them directly. Overall, we observed that 12-month-olds
appeared to have visual access to slightly fewer faces than 8 or
16-month-olds, creating a slight U-shaped function in face detections;
conversely, hand detections were showed a slight increase across this
age range, as reported in prior work (Fausey et al., 2016).

\begin{figure}[H]
\includegraphics[width=\textwidth]{figs/detByAge-1} \caption{Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data.}\label{fig:detByAge}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=\textwidth]{figs/detByPosOrient-1}

}

\caption{Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\% bootstrapped confidence intervals}\label{fig:detByPosOrient}
\end{figure}

However, these age-related trends were much smaller than the effect of
infant's postural developments on infants' visual access to faces and
hands. Infants' in-the-moment posture was a major factor both in how
many faces and hands were in view during the play session, as was their
orientation relative to their caregiver. This was quantified using two
generalized linear mixed-effect models estimating the proportion of
faces and hands that were in view, with orientation, posture, their
interaction, and scaled participant's age as fixed effects, and with
random slopes for infants' orientation and posture. A summary of the
coefficients of the models can be found in Tables 2 and 3.

The interaction between infants' posture and their caregiver's
orientation had the most dramatic effect on the social information in
view. When caregivers were behind their infants, supporting their
infants' sitting or standing positions, infants saw fewer faces. When
caregivers were relatively close to their infants, infants who were
sitting or standing had more faces in view (Face detections; infant
sitting and CG close, b=0.99, SE = 0.07, Z = 14.54, P = 0; infant
standing and CG close, b=1.36, SE = 0.08, Z = 16.95, P = 0 than infants
who were lying down/crawling (i.e.~prone). When caregivers were far away
from their infants, face detections were similarity higher (Face
detections; infant sitting and CG far, b = 0.75, SE = 0.07, Z = 10.48, P
= 0; infant standing and CG far, b = 1.40, SE = 0.09, Z = 16.16, P = 0.
Overall, we found that age did not remain a significant predictor in
accounting for the faces in view when modeling infants posture, their
orientation to their caregiver, and the interaction between these two
factors (Face detections; Age (scaled), b = 0.05, SE = 0.12, Z = 0.39, P
= 0.70).

We found a similar pattern of results for wrist detections, even though
there were fewer wrist detections overall in the dataset. Infants saw
fewer wrists when caregivers were behind their infants, supporting their
infants' sitting or standing positions vs when caregivers were
relatively closer to their infants (Hand detections; infant sitting and
CG close, b=0.31, SE = 0.08, Z = 3.72, P = 0.00; infant standing and CG
close, b=0.76, SE = 0.09, Z = 8.46, P = 0 than infants who were lying
down/crawling (i.e.~prone). Wrist detections were clearly highest when
caregivers were far away from their infants (Wrist detections; infant
sitting and CG far, b=0.13, SE = 0.10, Z = 1.27, P = 0.21; infant
standing and CG far, b=1.76, SE = 0.11, Z = 15.87, P = 0. As with faces,
age did not remain a significant predictor in these models (Wrist
detections; Age (scaled), b = 0.16, SE = 0.11, Z = 1.48, P = 0.14).

We directly examined the contributions of posture and orientation by
fitting a reduced version of the full model (Nakagawa \& Schielzeth,
2013) without their fixed effects (both models were run with the maximal
random effects structure). The fixed effects in a model with only the
age of the participants accounted for relatively little variance in the
proportion of faces (marginal \(R^2\) \textless{} 0.01) or hands in view
(marginal \(R^2\) = 0.02). However, when adding infants' posture and
orientation to their caregiver to the model (and their interaction) the
marginal \(R^2\) were higher for both faces (marginal \(R^2\) = 0.28)
and wrists (marginal \(R^2\) = 0.31).\\
Overall, these results suggest that infants' visual access to social
information is largely modulated by their posture and orientation to
their caregiver, which is in turn a function of their general locomotor
development.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & -4.29 & 0.18 & -23.87 & 0.000 \\ 
  Sit & 0.59 & 0.18 & 3.20 & 0.001 \\ 
  Stand & 0.53 & 0.30 & 1.76 & 0.079 \\ 
  Close & 0.53 & 0.19 & 2.75 & 0.006 \\ 
  Far & 0.15 & 0.26 & 0.56 & 0.574 \\ 
  Age (Scaled) & 0.16 & 0.11 & 1.48 & 0.140 \\ 
  Sit*Close & 0.31 & 0.08 & 3.72 & 0.000 \\ 
  Stand*Close & 0.76 & 0.09 & 8.46 & 0.000 \\ 
  Sit*Far & 0.13 & 0.10 & 1.27 & 0.205 \\ 
  Stand*Far & 1.76 & 0.11 & 15.87 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.} 
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & -3.25 & 0.18 & -18.04 & 0.000 \\ 
  Sit & 0.05 & 0.17 & 0.33 & 0.744 \\ 
  Stand & -0.24 & 0.18 & -1.34 & 0.180 \\ 
  Close & 0.01 & 0.19 & 0.07 & 0.942 \\ 
  Far & 0.45 & 0.27 & 1.63 & 0.102 \\ 
  Age (Scaled) & 0.05 & 0.12 & 0.39 & 0.697 \\ 
  Sit*Close & 0.99 & 0.07 & 14.54 & 0.000 \\ 
  Stand*Close & 1.36 & 0.08 & 16.95 & 0.000 \\ 
  Sit*Far & 0.75 & 0.07 & 10.48 & 0.000 \\ 
  Stand*Far & 1.40 & 0.09 & 16.16 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.} 
\end{table}

\subsubsection{Social information during naming
events}\label{social-information-during-naming-events}

Our play session was designed to provide parents with opportunities to
label objects -- both familiar and novel -- such that we could examine
whether children saw different kinds of social information around naming
events. In a set of exploratory analyses, we thus analyzed how face and
hand detections changed during object labeling events relative to
baseline, analyzing a four-second window around each labeling event
(e.g., \enquote{Look at the {[}zem{]}!}). Every utterance of one of the
labelled objects (e.g., \enquote{ball}) was counted as a
\enquote{labeling event}; timestamps of the beginning of each word were
hand-annotated and synchronized with the frame-by-frame detections.

To assess whether there were differences in the social information in
view during naming events, we first calculated the proportion of
detections that were in view during this four second window, and
averaged across naming events for each subject as function of whether
the named object as a novel or a familiar objet; this was then then
compared to the baseline proportion of faces in view for each subject in
linear mixed-effect models, with random effects of subjects and fixed
effects of (scaled) age.

Face detections were numerically slightly higher around these novel
naming events relative to baseline, with similar effects across age
groups (8 m.o., \(M_{fam - baseline}\) = =0.02, 12 m.o.,
\(M_{nov - baseline}\) =0.05, 16 m.o. \(M_{nov - baseline}\) =0.02)
however, they were not higher during familiar naming events vs.~baseline
(8 m.o., \(M_{fam - baseline}\) =0.00, 12 m.o., \(M_{fam - baseline}\)
=0.00, 16 m.o. \(M_{fam - baseline}\) =0.02, Conversely, wrist
detections were higher during both familiar (8 m.o.,
\(M_{fam - baseline}\) =0.07, 12 m.o., \(M_{fam - baseline}\) =0.04, 16
m.o. \(M_{fam - baseline}\) =0.08) and novel naming events relative to
baseline across all age groups (Novel; 8 m.o., (8 m.o.,
\(M_{fam - baseline}\) =0.09, 12 m.o., \(M_{fam - baseline}\) =0.08, 16
m.o. \(M_{fam - baseline}\) =0.08). These results were confirmed by a
linear mixed-effect model with scaled aged as a fixed effect and random
intercepts for each subject (Wrist detections; familiar objects
vs.~baseline, b=0.06, SE = 0.01, t = 4.33, P = 0.00'; Novel objects
vs.~baseline, b=0.08, SE = 0.02, t = 5.21, P = 0.

Overall, these exploratory results suggest that children may tend to see
more hands around naming events. This finding is consistent with the
possibility that caregivers may change how they interact with their
infant when presenting them with objects (Gogate, Bahrick, \& Watson,
2000, Gogate, Bolzani, and Betancourt (2006)). For example, caregivers
may tend to simultaneously name objects when demonstrating their
affordances or simply when pointing to them. In turn, infants may be
sensitive to these naming events and orient their attention towards
their caregiver, consistent with other accounts positing infants'
sensitivity to social cues in early word learning environments
(Yurovsky, 2018, Yurovsky and Frank (2017), Frank, Simmons, Yurovsky,
and Pusiol (2013)).

\begin{figure}[H]

{\centering \includegraphics[width=\textwidth]{figs/detByNaming-1}

}

\caption{Proportion of face / wrist detections during naming events (+/- 2 seconds around label) for familiar and novel objects; these rates are put into context relative to baseline. Error bars represent 95\% bootstrapped confidence intervals. Grey lines connect points from individual subjects.}\label{fig:detByNaming}
\end{figure}

\subsection{Extension to Franchak et a.,
2017}\label{extension-to-franchak-et-a.-2017}

In the present work, we found that infant's in-the-moment posture
changed with their age, as did infants' orientation relative their
caregiver. In a related study with 12-month-olds, Franchak et al. (2017)
also found that infants' in-the-moment posture changed the proportion of
time infants spent looking at faces. Here, we sought to replicate these
findings using our automated methodology (OpenPose detections), using
the footage from their head-mounted cameras (D. A. Simon, Gordon,
Steiger, \& Gilmore, 2015) for two reasons. First, we sought to validate
our novel method, which could fail to generalize to scenes from these
more complex environments, where detecting faces and hands could
arguably be a much harder task. Second, we sought to replicate the
effects of infants' in-the-moment posture on differences in visual
access to hands in an independent dataset.

\subsubsection{Comparison between video
datasets}\label{comparison-between-video-datasets}

First, despite having different head-mounted camera setups, the view
angle of the two head-mounted cameras were only slightly different
(\(52.2^{\circ}\) horizontal by \(42.2^{\circ}\) in Franchak et al.
(2017), here \(64^{\circ}\) horizontal by \(46^{\circ}\) vertical).
However, the environment that infants were immersed in with their
caregivers (and experimenters) was much larger and more varied than the
play room used in the present dataset, containing multiple structures
and toys in different parts of the room for infants to climb, explore,
and interact with, and infants were unconstrained and allowed to freely
wander the room. In contrast, the play room used in Study 1 was
relatively small (approximately 10' x 10') and was set-up with focused
play with the pairs of novel and familiar objects. Further, multiple
people were present during the play session in Franchak et al. (2017)--
including their caregiver and two experimenters -- whereas in Study 1
the experimenters left the room during the play session.

\begin{figure}[H]

{\centering \includegraphics[width=\textwidth]{figs/franchak-1}

}

\caption{Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals.}\label{fig:franchak}
\end{figure}

\subsubsection{Replication using automated
detections}\label{replication-using-automated-detections}

Overall, we found convergence between our two methodologies, replicating
the main results from Franchak et al. (2017), and finding that the
proportion of faces detected was greater when infants were sitting or
standing vs.~prone. We found this result regardless of whether we used
all detections (see Figure \ref{fig:franchak}, panel A; (proportion of
frames with face detections; Prone: M =0.13\%, Sitting: M =0.20\%,
Upright: M =0.22\%) or restricted our analyses to high-confidence
detections (percentage of frames with face detections; Prone: M =0.03,
Sitting: M =0.06, Upright: M =0.05). These results were confirmed in
generalized linear-mixed models, as in the first study (Sitting
vs.~Prone, b=0.34, SE = 0.02, Z = 23.37, P = 0. Upright vs.~Prone,
b=0.38, SE = 0.02, Z = 24.01, P = 24.01).

The most noticeable difference between our analysis methods is the
overall proportion of frames with faces detected by OpenPose vs.~the
proportion of frames in which infants were judged to be looking at faces
in Franchak et al. (2017). Across the entire session in Franchak et al.
(2017), infants looked at faces on 4.7\% of frames (Franchak et al.,
2017). When we used all detections from OpenPose, we found a higher
proportion of faces: M = 0.22, or 21.80\%. When we restricted our
results to only high-confidence detections, we found M = 0.06, closer to
the original values reported by Franchak et al. (2017). However, the
above analyses on the accuracy of these two methods suggest that
high-confidence detections dramatically underestimate the number of
faces in view. Thus, we suspect that these differences are due to both
(1) the fact head-mounted eye-trackers may sometimes underestimate the
proportion of faces attended due to calibration issues, and that (2)
OpenPose may sometimes detect faces that are in view for infants but
that infants may not be foveating.

We also found that infants' in-the-moment posture also modulated the
proportion of hands that were in view (i.e., wrist detections), though
these were not originally annotated in Franchak et al. (2017) (see
Figure \ref{fig:franchak}, (proportion of frames with wrist detections;
Prone: M =0.15; Sitting: M =0.21; Upright: M =0.22). These results were
confirmed in generalized linear-mixed models, as in the first study
(Sitting vs.~Prone, b=0.27, SE = 0.01, Z = 19.09, P = 0. Upright
vs.~prone, b=0.30, SE = 0.02, Z = 19.73, P = 19.73).

Overall, these analyzes extend and validate previous work, replicating
the results in Study 1 that infants' in-the-moment posture modulated the
proportion of hands in view, and suggesting that posture is a major
factor that structures infants' access to visual information broadly
construed.

\section{General Discussion}\label{general-discussion}

What social cues do infants see as they learn language, and how does
this change as they grow and are start to locomote themselves in their
environments? We explored this question using video data from
head-mounted cameras from two naturalistic datasets of parent-child
interactions: a cross-sectional database of naturalistic play sessions
from 8-16 month-olds with sets of novel and familiar toys, and database
of head-mounted camera videos from one-year-olds who explored a large
play area from Franchak et al. (2017). To analyze these datasets, we
developed a novel method using a pose detection model to automate the
annotation of the social information in the infant view, here
operationalized as the presence of the faces and hands of their
caregiver; these annotations were then synced with manual annotations of
infants in-the-moment posture from third-person videos.

Despite being trained on the adult perspective, the pose detector we
used (OpenPose, Cao et al. (2017)) was able to generalize relatively
well to the infant viewpoint, achieving comparable precision and
accuracy as a face detector relative to a state-of-the-art model
optimized specifically for detecting faces in natural scenes (MTCNN, K.
Zhang et al. (2016){]}. While OpenPose had relatively low recall as a
hand detector -- missing some hands that were in the infant view -- it
made comparable rates of false alarms. In both cases, we found that
overall performance was maximized when all detections were included,
regardless of their confidence, suggesting that some low-confidence face
and hand detection still index actual faces and hands that were seen by
infants. Thus, while imperfect, we suggest that OpenPose can be applied
to infant egocentric videos for the extraction of the social information
in the infant viewpoint, reducing the burden of manual annotations and
promoting the reusability of rich video datasets for further analyses.
The use of this automated methodology allowed us to easily annotate the
entirety of our dataset -- additionally analyzing the social information
around naming events -- and to re-analyze the data from Franchak et al.
(2017) to replicate our findings in a very different kind of play
session.

Broadly, our results replicate and extend previous work, first by
showing systematic changes in infants' in-the-moment posture and their
orientation relative to their caregivers (Adolph \& Franchak, 2017);
older children spent more time standing and less time sitting, and older
infants' caregivers spent less time supporting their standing or sitting
postures. Motor development changes dramatically at the same time that
children are breaking into language learning. Using these automated
detections, we found that infants' changing posture and orientation to
their caregiver jointly shaped the amount of social information that was
in their view during one-on-one play sessions with their caregivers:
Children saw the most faces/hands when they were sitting or standing and
close to their caregiver vs.~crawling or prone. These same findings were
recapitulated in a second dataset collected by Franchak et al. (2017)
with one-year-olds: sitting and upright infants saw more faces--and
hands--than infants who were prone. Motor development appears to
modulate how infants experience their visual world and the social
information in it.

While exploratory, our results also suggest that infants saw a greater
proportion of hands during naming events, hinting that children may have
been orienting towards their caregiver when they heard labels for
objects. While this effect was not present for faces, other work
(Yoshida \& Smith, 2008; Yu \& Smith, 2013, 2017), including Franchak et
al. (2017), has found that infants spend much more time looking at the
toys vs.~their caregiver's faces during these play sessions, and
highlighted the importance of hand-following as a component of joint
attention (Yu \& Smith, 2017). Further, given that there were only two
possible referents in the room at a time----and one of them was always a
familiar category (i.e., car, kitty) -- this particular play session did
not present many opportunities where children would need to use gaze
cues to disambiguate referents.

Importantly, however, all of these findings come from observational,
in-lab datasets, posing limits on their generalizability. Future work is
needed to relate these the slices of experience captured within in-lab
play sessions and infants' everyday experiences (Clerkin, Hart, Rehg,
Yu, \& Smith, 2017; Fausey et al., 2016). More broadly, though
observational findings allow us to document developmental change and
identify potential causal pathways, they cannot confirm them. Locomotive
abilities are of course only part of a cascade of changes in infants'
abilities and experiences, and these analyzes document only a fraction
of this broader, multifaceted trajectory. As children grow and change,
the activities in which they engage with their caregivers are likely to
also vary, leading to differences in the distribution of social cues
that they experience that may not be captured in these in-lab play
sessions.

Nonetheless, we think that this approach hold promise for documenting
these developmental trajectories and for generating hypotheses about
potential causal pathways. Indeed, understanding the relationship
between different domains of developmental changes in naturalistic
contexts has been a persistent challenge for developmental psychology.
However, the field of computer vision has advanced dramatically in
recent years, creating a new generation of algorithmic tools that deal
better with noisier, more complicated datasets and extract richer
information. We hope that these new tools can now be leveraged to
examine the consequences of the changing infant perspective for
linguistic, cognitive, and social development.

\section{Acknowledgements}\label{acknowledgements}

Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in
recruitment, data collection, and annotation. This work was funded by a
Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF,
and NSF \#1714726 to BLL. An earlier version of this work was presented
to the Cognitive Science Society in Frank et al. (2013) and Sanchez,
Long, Kraus, and Frank (2018).

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-adolph2006motor}{}
Adolph, K. E., \& Berger, S. E. (2006). Motor development.
\emph{Handbook of Child Psychology}.

\hypertarget{ref-adolph2017development}{}
Adolph, K. E., \& Franchak, J. M. (2017). The development of motor
behavior. \emph{Wiley Interdisciplinary Reviews: Cognitive Science},
\emph{8}(1-2), e1430.

\hypertarget{ref-adolph2012toward}{}
Adolph, K. E., Gilmore, R. O., Freeman, C., Sanderson, P., \& Millman,
D. (2012). Toward open behavioral science. \emph{Psychological Inquiry},
\emph{23}(3), 244--247.

\hypertarget{ref-adolph1998roles}{}
Adolph, K. E., Vereijken, B., \& Denny, M. (1998). Roles of variability
and experience in development of crawling. \emph{Child Development},
\emph{69}(1299), 312.

\hypertarget{ref-bambach2017}{}
Bambach, S., Crandall, D. J., Smith, L. B., \& Yu, C. (2017). An
egocentric perspective on active vision and visual object learning in
toddlers. In \emph{Proceedings of the seventh joint ieee conference on
development and learning and on epigenetic robotics}.

\hypertarget{ref-brooks2005}{}
Brooks, R., \& Meltzoff, A. (2005). The development of gaze following
and its relation to language. \emph{Developmental Science}, \emph{8}(6),
535--543.

\hypertarget{ref-brooks2008}{}
Brooks, R., \& Meltzoff, A. N. (2008). Infant gaze following and
pointing predict accelerated vocabulary growth through two years of age:
A longitudinal, growth curve modeling study. \emph{Journal of Child
Language}, \emph{35}(1), 207--220.

\hypertarget{ref-bruner1975}{}
Bruner, J. (1975). From communication to language: A psychological
perspective. \emph{Cognition}, \emph{3}(3), 255--287.

\hypertarget{ref-cao2017realtime}{}
Cao, Z., Simon, T., Wei, S.-E., \& Sheikh, Y. (2017). Realtime
multi-person 2D pose estimation using part affinity fields. In
\emph{CVPR}.

\hypertarget{ref-carpenter1998}{}
Carpenter, M., Nagell, K., \& Tomasello, M. (1998). Social cognition,
joint attention, and communicative competence from 9 to 15 months of
age. \emph{Monographs of the Society for Research in Child Development},
\emph{63}(4).

\hypertarget{ref-clerkin2017}{}
Clerkin, E. M., Hart, E., Rehg, J. M., Yu, C., \& Smith, L. B. (2017).
Real-world visual statistics and infants' first-learned object names.
\emph{Phil. Trans. R. Soc. B}, \emph{372}(1711), 20160055.

\hypertarget{ref-cummings1988}{}
Cummings, M., Van Hof-Van Duin, J., Mayer, D., Hansen, R., \& Fulton, A.
(1988). Visual fields of young children. \emph{Behavioural and Brain
Research}, \emph{29}(1), 7--16.

\hypertarget{ref-farroni2002eye}{}
Farroni, T., Csibra, G., Simion, F., \& Johnson, M. H. (2002). Eye
contact detection in humans from birth. \emph{Proceedings of the
National Academy of Sciences}, \emph{99}(14), 9602--9605.

\hypertarget{ref-fausey2016}{}
Fausey, C. M., Jayaraman, S., \& Smith, L. B. (2016). From faces to
hands: Changing visual input in the first two years. \emph{Cognition},
\emph{152}, 101--107.

\hypertarget{ref-franchak2017see}{}
Franchak, J. M., Kretch, K. S., \& Adolph, K. E. (2017). See and be
seen: Infant--caregiver social looking during locomotor free play.
\emph{Developmental Science}.

\hypertarget{ref-franchak2011}{}
Franchak, J. M., Kretch, K. S., Soska, K. C., \& Adolph, K. E. (2011).
Head-mounted eye tracking: A new method to describe infant looking.
\emph{Child Development}, \emph{82}(6), 1738--1750.

\hypertarget{ref-frank2013}{}
Frank, M. C., Simmons, K., Yurovsky, D., \& Pusiol, G. (2013).
Developmental and postural changes in children's visual access to faces.
In \emph{Proceedings of the 35th annual meeting of the cognitive science
society} (pp. 454--459).

\hypertarget{ref-gogate2000study}{}
Gogate, L. J., Bahrick, L. E., \& Watson, J. D. (2000). A study of
multimodal motherese: The role of temporal synchrony between verbal
labels and gestures. \emph{Child Development}, \emph{71}(4), 878--894.

\hypertarget{ref-gogate2006attention}{}
Gogate, L. J., Bolzani, L. H., \& Betancourt, E. A. (2006). Attention to
maternal multimodal naming by 6-to 8-month-old infants and learning of
word--object relations. \emph{Infancy}, \emph{9}(3), 259--288.

\hypertarget{ref-gredeback2010development}{}
Gredeback, G., Fikke, L., \& Melinder, A. (2010). The development of
joint visual attention: A longitudinal study of gaze following during
interactions with mothers and strangers. \emph{Developmental Science},
\emph{13}(6), 839--848.

\hypertarget{ref-iverson2010}{}
Iverson, J. M. (2010). Developing language in a developing body: The
relationship between motor development and language development.
\emph{Journal of Child Language}, \emph{37}(2), 229--261.

\hypertarget{ref-jayaraman2017faces}{}
Jayaraman, S., Fausey, C. M., \& Smith, L. B. (2017). Why are faces
denser in the visual experiences of younger than older infants?
\emph{Developmental Psychology}, \emph{53}(1), 38.

\hypertarget{ref-karasik2014}{}
Karasik, L. B., Tamis-LeMonda, C. S., \& Adolph, K. E. (2014). Crawling
and walking infants elicit different verbal responses from mothers.
\emph{Developmental Science}, \emph{17}(3), 388--395.

\hypertarget{ref-kietzmann2018deep}{}
Kietzmann, T. C., McClure, P., \& Kriegeskorte, N. (2018). Deep neural
networks in computational neuroscience. \emph{BioRxiv}, 133504.

\hypertarget{ref-kretch2014}{}
Kretch, K. S., Franchak, J. M., \& Adolph, K. E. (2014). Crawling and
walking infants see the world differently. \emph{Child Development},
\emph{85}(4), 1503--1518.

\hypertarget{ref-lampl1993evidence}{}
Lampl, M. (1993). Evidence of saltatory growth in infancy.
\emph{American Journal of Human Biology}, \emph{5}(6), 641--652.

\hypertarget{ref-libertus2016sit}{}
Libertus, K., \& Violi, D. A. (2016). Sit to talk: Relation between
motor skills and language development in infancy. \emph{Frontiers in
Psychology}, \emph{7}, 475.

\hypertarget{ref-mayer1988}{}
Mayer, D., Fulton, A., \& Cummings, M. (1988). Visual fields of infants
assessed with a new perimetric technique. \emph{Investigative
Ophthalmology \& Visual Science}, \emph{29}(3), 452--459.

\hypertarget{ref-moore2019point}{}
Moore, C., Dailey, S., Garrison, H., Amatuni, A., \& Bergelson, E.
(2019). Point, walk, talk: Links between three early milestones, from
observation and parental report. \emph{Developmental Psychology}.

\hypertarget{ref-nakagawa2013general}{}
Nakagawa, S., \& Schielzeth, H. (2013). A general and simple method for
obtaining r2 from generalized linear mixed-effects models. \emph{Methods
in Ecology and Evolution}, \emph{4}(2), 133--142.

\hypertarget{ref-patrick2012developmental}{}
Patrick, S. K., Noah, J. A., \& Yang, J. F. (2012). Developmental
constraints of quadrupedal coordination across crawling styles in human
infants. \emph{Journal of Neurophysiology}, \emph{107}(11), 3050--3061.

\hypertarget{ref-peterson2018evaluating}{}
Peterson, J. C., Abbott, J. T., \& Griffiths, T. L. (2018). Evaluating
(and improving) the correspondence between deep neural networks and
human representations. \emph{Cognitive Science}, \emph{42}(8),
2648--2669.

\hypertarget{ref-sanchez2018postural}{}
Sanchez, A., Long, B., Kraus, A. M., \& Frank, M. C. (2018). Postural
developments modulate children's visual access to social information.

\hypertarget{ref-scheirer2014perceptual}{}
Scheirer, W. J., Anthony, S. E., Nakayama, K., \& Cox, D. D. (2014).
Perceptual annotation: Measuring human vision to improve computer
vision. \emph{IEEE Transactions on Pattern Analysis and Machine
Intelligence}, \emph{36}(8), 1679--1686.

\hypertarget{ref-simon2015databrary}{}
Simon, D. A., Gordon, A. S., Steiger, L., \& Gilmore, R. O. (2015).
Databrary: Enabling sharing and reuse of research video. In
\emph{Proceedings of the 15th acm/ieee-cs joint conference on digital
libraries} (pp. 279--280).

\hypertarget{ref-simon2017hand}{}
Simon, T., Joo, H., Matthews, I., \& Sheikh, Y. (2017). Hand keypoint
detection in single images using multiview bootstrapping. In
\emph{CVPR}.

\hypertarget{ref-simonyan2014very}{}
Simonyan, K., \& Zisserman, A. (2014). Very deep convolutional networks
for large-scale image recognition. \emph{ArXiv Preprint
ArXiv:1409.1556}.

\hypertarget{ref-soska2010systems}{}
Soska, K. C., Adolph, K. E., \& Johnson, S. P. (2010). Systems in
development: Motor skill acquisition facilitates three-dimensional
object completion. \emph{Developmental Psychology}, \emph{46}(1), 129.

\hypertarget{ref-vanrullen2017perception}{}
VanRullen, R. (2017). Perception science in the age of deep neural
networks. \emph{Frontiers in Psychology}, \emph{8}, 142.

\hypertarget{ref-viola2004robust}{}
Viola, P., \& Jones, M. J. (2004). Robust real-time face detection.
\emph{International Journal of Computer Vision}, \emph{57}(2), 137--154.

\hypertarget{ref-walle2016infant}{}
Walle, E. A. (2016). Infant social development across the transition
from crawling to walking. \emph{Frontiers in Psychology}, \emph{7}, 960.

\hypertarget{ref-walle2014}{}
Walle, E. A., \& Campos, J. J. (2014). Infant language development is
related to the acquisition of walking. \emph{Developmental Psychology},
\emph{50}(2), 336.

\hypertarget{ref-wei2016cpm}{}
Wei, S.-E., Ramakrishna, V., Kanade, T., \& Sheikh, Y. (2016).
Convolutional pose machines. In \emph{CVPR}.

\hypertarget{ref-yoshida2008}{}
Yoshida, H., \& Smith, L. (2008). What's in view for toddlers? Using a
head camera to study visual experience. \emph{Infancy}, \emph{13},
229--248.

\hypertarget{ref-yu2013joint}{}
Yu, C., \& Smith, L. B. (2013). Joint attention without gaze following:
Human infants and their parents coordinate visual attention to objects
through eye-hand coordination. \emph{PloS One}, \emph{8}(11).

\hypertarget{ref-yu2017hand}{}
Yu, C., \& Smith, L. B. (2017). Hand--eye coordination predicts joint
attention. \emph{Child Development}, \emph{88}(6), 2060--2078.

\hypertarget{ref-yurovsky2018communicative}{}
Yurovsky, D. (2018). A communicative approach to early word learning.
\emph{New Ideas in Psychology}, \emph{50}, 73--79.

\hypertarget{ref-yurovsky2017beyond}{}
Yurovsky, D., \& Frank, M. C. (2017). Beyond naïve cue combination:
Salience and social cues in early word learning. \emph{Developmental
Science}, \emph{20}(2), e12349.

\hypertarget{ref-zhang2016}{}
Zhang, K., Zhang, Z., Li, Z., \& Qiao, Y. (2016). Joint face detection
and alignment using multitask cascaded convolutional networks.
\emph{IEEE Signal Processing Letters}, \emph{23}(10), 1499--1503.

\hypertarget{ref-zhou2017places}{}
Zhou, B., Lapedriza, A., Khosla, A., Oliva, A., \& Torralba, A. (2017).
Places: A 10 million image database for scene recognition. \emph{IEEE
Transactions on Pattern Analysis and Machine Intelligence},
\emph{40}(6), 1452--1464.

\endgroup

\end{document}

### 2020-08-11 writeup-paper/draft_apa_march2020.Rmd
latest version of writeup

@@
 Indeed, the infant visual perspective has proved to be both remarkably different than the adult perspective and not easily predicted by our own intuitions [@franchak2011, @yoshida2008, @clerkin2017].
-Over the first two years of life, the infant viewpoint seems to transition from primarily containing close up view of faces to capturing relatively restricted views of hands paired with the objects they are acting on [@fausey2016]; both computational and empirical work suggest that this restricted viewpoint may be more effective for learning about objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017].
+
+Over the first two years of life, the infant viewpoint seems to transition from primarily containing close up view of faces to capturing relatively restricted views of hands paired with the objects they are acting on [@fausey2016].
+(FAUSEY REVIEW HERE) both computational and empirical work suggest that this restricted viewpoint may be more effective for learning about objects and their labels than the comparable adult perspective [@yurovsky2012, @bambach2017].
 
@@
 More recently, when 12-month-olds [@franchak2017see] participated in a free-play session with their caregivers, their own posture and as well as their caregiver’s posture influenced the proportion of time they spent looking at the faces and bodies of their caregivers [@franchak2017see].
+
+TWO GOALS
 
@@
 # Methods
+A secondary goal of this project was to provide a rich, open dataset for other researchers to examine labeling behavior during naturalistic parent-child interactions.
+We thus invited parents of 8, 12, and 16-month-olds to participate in play sessions where they were provided with pairs of novel and familiar objects (e.g., a ball and a microfiber duster, called a "zem").
+Third-person cameras show a birds-eye videw of the play session, and all videos were transcribed.
+Finally, we collected collected MacArthur CDIs for all children who particpated.
+
 ## Participants
@@
 To obtain this final sample, we tested `r length(all_subs$reason)` children, excluding `r sum(all_subs$reason!="included")` children for the following reasons:
-`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam,
+`r sum(all_subs$reason=="HC technical error")` for technical issues related to the headcam (e.g., failure to record),
 `r sum(all_subs$reason=="would not wear camera")` for failing to wear the headcam,
@@
 `r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
-`r sum(all_subs$reason=="infant crying")` for fussiness, and one for sample symmetry.
-All inclusion decisions were made independent of the results of subsequent analyses.
-These data were also analyzed in @frank2013 and @sanchez2018postural.
+`r sum(all_subs$reason=="infant crying")` for fussiness, and one for sample symmetry. All inclusion decisions were made independent of the results of subsequent analyses.
 
@@
 When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session.
-Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys.
+Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would."  Objects were provided in a box with written labels so that parents did not need to remember the novel labels.
+All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys.
 The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.
@@
 (min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).
+
+### Transcriptions
+Audio from headcam video was transcribed...
 

### 2020-08-11 analysis/0_sanity_checks/check_posture_frames.Rmd
port analysis scripts to analysis folder; rename

---
title: "check_posture_times"
author: "Bria Long"
date: "2/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Step 0: Libraries and helper functions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) # remove all objects from current workspace
library(dplyr)
library(readr)
library(magrittr)
library(tidyr)
library(assertthat)
source("helper.R") # Contains critical helper functions for merging annotations
```

## Step 3. Merge demographics and manual annotations
```{r}
# load demographics and pose / orientation / timing
demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
  select(-ra, -assist, -len) ## -dot -dib fields did not exist, deleted bll
```

```{r}
######## ADD POSTURE CODING #########
add.posture <- function(x) {
  fname <- paste("../data/posture/",
                 x$subid[1],
                 ".csv",
                 sep="")

  if (file.exists(fname)) {
    print(fname)
    postures <- read_csv(fname)
    postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order
    postures <- regularize.postures(subset(postures,code=="posture"))
    x$posture <- factor(NA,levels=levels(postures$posture))
    x$orientation <- factor(NA,levels=levels(postures$orientation))

    # for each row of p, populate posture and orientation to x
    for (i in 1:nrow(postures)) {
      range <- x$time > postures$start[i] & x$time <= postures$end[i]
      x$posture[range] <- postures$posture[i]
      x$orientation[range] <- postures$orientation[i]
    }
  } else {
    print(paste0("**** ", x$subid[1], " is missing posture ****"))
    x$posture <- NA
    x$orientation <- NA
  }

  return(x)
}
```

### 2020-08-11 analysis/0_sanity_checks/fps_sanity_checks.Rmd
port analysis scripts to analysis folder; rename

--- 2018-03-13 analysis-validation/fps_sanity_checks.Rmd
@@
 new <- read_csv("../data/video_stats/new_movie_lengths.csv") 
-old <- read_csv("../data/video_stats/old_movie_lengths.csv")
+old <- read_csv("../data/video_stats/old_movie_lengths.csv") ## from previous iteration
 ```

### 2020-08-11 analysis/0_sanity_checks/length_sanity_check.Rmd
port analysis scripts to analysis folder; rename

--- 2018-03-13 analysis-validation/length_sanity_check.Rmd
@@
 movies <- read_csv("../data/video_stats/new_movie_lengths.csv") 
-MTCNN <- read_csv("../data/final_output/mtcnn3.csv")
-openPose <- read_csv("../data/final_output/openpose_results_truncated_2.csv")
+MTCNN <- read_csv("../data/final_detector_output/mtcnn3.csv")
+openPose <- read_csv("../data/final_detector_output/openpose_results_truncated_2.csv")
 frameTimes <- read_csv("../data/video_stats/frametimechecks_showframes.csv")
@@
 
-XS_1229 has been trimmed (headcam came off) and therefore there are different movie lenghts.
-
 =====
 
-Mike's notes (Explanations of old discrepancies :)
+<!-- XS_1229 has been trimmed (headcam came off) and therefore there are different movie lenghts. -->
 
-+ 0825 - unknown, ~500 frames off (15s), spot checks seem to match decently well. 
-+ 1205, 1211, and 1225 were post-trimed, hypothesis is that the _QT files are wrappers to source files that trim length to appropriate? 
-+ 1628 and 1639 both off by a lot of frames, unknown.
+<!-- Old notes pre-cleaning of dataset (irrelevant) -->
 
-to check in the videos
-+ 1229 is reported to be much longer in the demographics file, perhaps this is post-trimming? 
-+ 1612 is reported to be shorter than it is, check this. 
-+ 1622 is reported to be longer than it is, check this.
+<!-- + 0825 - unknown, ~500 frames off (15s), spot checks seem to match decently well.  -->
+<!-- + 1205, 1211, and 1225 were post-trimed, hypothesis is that the _QT files are wrappers to source files that trim length to appropriate?  -->
+<!-- + 1628 and 1639 both off by a lot of frames, unknown. -->
+
+<!-- to check in the videos -->
+<!-- + 1229 is reported to be much longer in the demographics file, perhaps this is post-trimming?  -->
+<!-- + 1612 is reported to be shorter than it is, check this.  -->
+<!-- + 1622 is reported to be longer than it is, check this.  -->

### 2020-08-11 analysis/0_sanity_checks/z-old/main_analyses_bll.Rmd
port analysis scripts to analysis folder; rename

--- 2018-03-13 analysis-main/main_analyses_bll.Rmd
@@
 ```{r}
-d <- read_csv("../data/consolidated_data_4detectors.feather") %>%
+d <- read_csv("../data/consolidated_data_4detectors_march7.csv") %>%
   mutate(posture = factor(posture), 

### 2020-08-11 analysis/1_merge_data/merge_all_data.Rmd
port analysis scripts to analysis folder; rename

--- 2020-08-11 writeup-paper/step1_merge_data_2020.Rmd
@@
 source("helper.R") # Contains critical helper functions for merging annotations
+library(here)
 ```
@@
 ```{r}
-dets_MTCNN <- read_csv("../data/final_detector_output/mtcnn3.csv") %>%
+dets_MTCNN <- read_csv("../../data/final_detector_output/mtcnn3.csv") %>%
   mutate(subid = video) %>%
@@
 ```{r}
-detsViola <- read_csv("../data/final_detector_output/viola.csv") 
+detsViola <- read_csv("../../data/final_detector_output/viola.csv") 
 detsViola <- detsViola %>%
@@
 ```{r}
-detsOpenPose <- read_csv("../data/final_detector_output/openpose_results_truncated_2.csv") 
+detsOpenPose <- read_csv("../../data/final_detector_output/openpose_results_truncated_2.csv") 
 detsOpenPose <- detsOpenPose %>%
@@
 # load demographics and pose / orientation / timing
-demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
+demo.data <- read_csv("../../data/demographics/demographics.csv") %>% 
   select(-ra, -assist, -len) ## -dot -dib fields did not exist, deleted bll
@@
 ```{r}
-sync_times_check <- read_csv("../data/manual_sync_times/video_sync_times.csv") 
+sync_times_check <- read_csv("../../data/manual_sync_times/video_sync_times.csv") 
 s <- sync_times_check %>%
@@
 posture.sync <- function(x) {
-  fname <- paste("../data/posture/",
+  fname <- paste("../../data/posture/",
                  x$subid[1],

### 2020-08-11 analysis/2_coding_reliability/reliability_naming_posture.Rmd
port analysis scripts to analysis folder; rename

---
title: "reliability_analyses"
author: "Bria Long"
date: "7/27/2020"
output: html_document
---

# Step 0: Libraries and helper functions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) # remove all objects from current workspace
library(dplyr)
library(readr)
library(magrittr)
library(tidyr)
library(lubridate)
library(irr)
library(forcats)
library(assertthat)
source("helper.R") # Contains critical helper functions for merging annotations
```

## Load basics
```{r}
# frame by video structure
load(file = '../data/consolidated_data/video_by_frame.RData')
```

## Naming reliabilty
```{r}
# paths to datafiles
file_path = '../data/naming/second/'
second_codes = dir(file_path) %>%
  tibble() %>%
  mutate(subid = str_split_fixed(., '-',2)[,1])

##
all_namings_second_coder <- all_dets_stripped %>%
  filter(subid %in% second_codes$subid) %>%
  do(add.times(.)) %>%  # calls helper functions to add times
  group_by(subid) %>%
  do(get.namings.second(.)) %>%
  rename(time_second_coder = time)  %>%
  select(-naming.instance) 
##
all_namings_first_coder <- all_dets_stripped %>%
  filter(subid %in% second_codes$subid) %>%
  do(add.times(.)) %>%  # calls helper functions to add times
  group_by(subid) %>%
  do(get.namings(.))  %>%
  rename(time_first_coder = time) %>%
  select(-naming.instance, -X)

joined_namings_4s <- all_namings_first_coder %>%
  full_join(all_namings_second_coder) %>%
  mutate(time_lower = time_first_coder-2, time_upper = time_first_coder+2) %>%
  mutate(match = time_second_coder>=time_lower & time_second_coder<=time_upper) %>%
  group_by(subid, name, time_first_coder) %>%
  summarize(found_both = sum(match)>0) %>% 
  mutate(found_both = replace_na(found_both, FALSE))

joined_namings_2s <- all_namings_first_coder %>%
  full_join(all_namings_second_coder) %>%
  mutate(time_lower = time_first_coder-1, time_upper = time_first_coder+1) %>%
  mutate(match = time_second_coder>=time_lower & time_second_coder<=time_upper) %>%
  group_by(subid, name, time_first_coder) %>%
  summarize(found_both = sum(match)>0) %>% 
  mutate(found_both = replace_na(found_both, FALSE))

mean(joined_namings_4s$found_both)
mean(joined_namings_2s$found_both)

# if we want to examine second -> first coder reliability
#
# joined_namings_2 <- all_namings_second_coder %>%
#   full_join(all_namings_first_coder) %>%
#   mutate(time_lower = time_second_coder-2, time_upper = time_second_coder+2) %>%
#   mutate(match = time_first_coder>=time_lower & time_first_coder<=time_upper) %>%
#   group_by(subid, name, time_second_coder) %>%
#   summarize(found_both = sum(match)>0) %>%
#   mutate(found_both = replace_na(found_both, FALSE))

# mean(joined_namings_2$found_both, na.rm=TRUE)

```
We found that `r round(mean(joined_namings_4s$found_both),2)*100`% of naming events were detected within a 4 second window around the naming event, and `r round(mean(joined_namings_2s$found_both),2)*100`% of namings events were detected within a 2 second window.

```{r}
for_binning <- all_dets_stripped %>%
  filter(subid %in% second_codes$subid) %>%
  do(add.times(.))

bin_size = 3 # seconds
min_time = min(for_binning$time, na.rm=TRUE)
max_time = max(for_binning$time, na.rm=TRUE)
bin_starts = seq(min_time, max_time+1,bin_size)
bins = c(bin_starts, max_time)

test <- for_binning %>%
  select(subid, time) %>%
  mutate(time_bin = cut(time, bins, bin_starts))

```

## Posture reliability
```{r}

d <- all_dets_stripped %>%
  group_by(subid) %>% # get subid
  do(add.times(.)) %>%  # calls helper functions to add times
  group_by(subid) %>%
  do(add.posture(.)) %>% ## adds both postures and orientatinos
  do(add.posture.second(.)) ## and from a second coder...
```

```{r}
posture_reliability <- d %>%
  select(subid, frame, posture, posture_second, orientation, orientation_second) %>%
  mutate(posture = fct_explicit_na(posture,'NIF')) %>%
  mutate(posture_second = fct_explicit_na(posture_second,'NIF')) %>%
  filter(!posture=='NIF') %>%
  filter(!posture_second=='NIF')

col_names = c('rater1','rater2')
posture_eval = data.frame(posture_reliability$posture, posture_reliability$posture_second)
irr::kappa2(posture_eval)

orient_eval = data.frame(posture_reliability$orientation, posture_reliability$orientation_second)
irr::kappa2(orient_eval)

```

```{r}
coder_1 <- ggplot(posture_reliability, aes(x=frame, y=posture_second, fill=posture_second)) +
  geom_tile() +
  facet_wrap(~subid) + 
  theme(legend.position='none')

coder_2 <- ggplot(posture_reliability, aes(x=frame, y=posture, fill=posture)) +
  geom_tile() +
  facet_wrap(~subid) + 
  theme(legend.position='none')

cowplot::plot_grid(coder_1,coder_2, nrow=2)
```

```{r}
# posture_every_second <- d %>%
#   mutate(time_seconds = round(time,0)) %>%
#   group_by(time_seconds) %>%
#   mutate(common_posture = names(sort(table(posture)))[5],
#   common_posture_second = names(sort(table(posture_second)))[5])
# 
# posture_eval_second = data.frame(posture_every_second$common_posture, posture_every_second$common_posture_second)
# irr::kappa2(posture_eval_second)
```

### 2020-08-11 analysis/4_extra_plots/SI_extraplots.Rmd
port analysis scripts to analysis folder; rename

---
title: "SI_ExtraPlots"
author: "Bria Long"
date: "4/8/2020"
output: pdf_document
---

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(lme4)
library(gridExtra)
library(ggridges)
library(assertthat)
library(tidyverse)
library(viridis)
library(papaja)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN DETECTIONS
d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_march2020.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

## make long form version for plotting both face detections on same graph
dlong <- d %>%
  mutate(id = paste(subid,frame)) %>%
  gather(key = detectorType, value = detection, faceMT, faceOP, faceVJ, wristOP) %>%
  mutate(detectorType=as.factor(detectorType))

# for better plotting names
levels(dlong$detectorType) 
# [1] "faceMT"  "faceOP"  "faceVJ"  "wristOP"
levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces","OpenPose-Wrists")

```

```{r}
source("helper.R")
## summarize detections around naming events
aroundNamings_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.))

aroundNamings_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceOP) %>%
  do(summarize.naming(.)) 
```

```{r eval=FALSE,include=FALSE}

### SUMMARIES FOR SI PLOTS
### Faces/Wrists by naming by posture/orientation

## summarize faces/wrists detected by orientation/posture/naming events
aroundNamings_Faces_PosOrient_Group <- aroundNamings_Faces %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(propFacesDetected = mean(detections, na.rm=TRUE)) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  group_by(orientation, posture, familiarity, age.grp) %>%
  multi_boot_standard(col = "propFacesDetected")

aroundNamings_Wrists_PosOrient_Group <- aroundNamings_Wrists %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(propWristsDetected = mean(detections, na.rm=TRUE)) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  group_by(orientation, posture, familiarity, age.grp) %>%
  multi_boot_standard(col = "propWristsDetected")

## for plotting
aroundNamings_Faces_PosOrient_Raw <- aroundNamings_Faces %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE), num_namings= length(detections)) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  left_join(ages_inc_subs)

aroundNamings_Wrists_PosOrient_Raw <- aroundNamings_Wrists %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE), num_namings=length(detections)) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  left_join(ages_inc_subs)

```

```{r eval=FALSE,include=FALSE}

### SI PLOT 1
### Faces by naming by posture/orientation

# plots means by familairity and posture
ggplot(aroundNamings_Faces_PosOrient_Group, aes(x=age.grp, y=mean, color=familiarity)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper, color=familiarity), position = position_dodge(width=3), alpha=.9) +
    # plot indiv data points
  geom_point(data=aroundNamings_Faces_PosOrient_Raw, aes(x=age.grp, y=prop_detected, size=num_namings, color=familiarity), position = position_dodge(width=3), alpha=.2) +
  # # plot baseline
    geom_pointrange(data = (detectionsByPostureAndOrientByGroup %>% filter(detectorType == 'OpenPose-Faces')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=0), alpha=.9) +
  theme_few() +
  ylim(0,.6) +
  facet_grid(orientation ~ posture) +
  labs(y = "Prop. face detections around naming events", x = "Age (months)") +
  theme_few(base_size=16) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")
```
```{r eval=FALSE,include=FALSE}

### SI PLOT 2
### Wrists by naming by posture/orientation

# plots means by familairity and posture
ggplot(aroundNamings_Wrists_PosOrient_Group, aes(x=age.grp, y=mean, color=familiarity)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper, color=familiarity), position = position_dodge(width=3), alpha=.9) +
    # plot indiv data points
  geom_point(data=aroundNamings_Wrists_PosOrient_Raw, aes(x=age.grp, y=prop_detected, size=num_namings, color=familiarity), position = position_dodge(width=3), alpha=.2) +
  # # plot baseline
    geom_pointrange(data = (detectionsByPostureAndOrientByGroup %>% filter(detectorType == 'OpenPose-Wrists')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=0), alpha=.9) +
  theme_few() +
  ylim(0,.6) +
  facet_grid(orientation ~ posture) +
  labs(y = "Prop. wrist detections around naming events", x = "Age (months)") +
  theme_few(base_size=16) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")
```

```{r eval=FALSE}
### Parallel plot to detections -- not included
# better labels 
orientation_labels <- c(
  behind ='Caregiver behind',
  close = 'Caregiver close',
  far = 'Caregiver far')

posture_labels = c(
  prone = 'Infant prone',
  sit = 'Infant sitting',
  stand ='Infant standing'
)
posture_by_orientation_subset <- d %>%
  filter(!is.na(posture)) %>%
  filter(!is.na(orientation)) %>%
  group_by(subid, posture, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  group_by(subid) %>%
  mutate(prop.time = time/sum(time)) %>%
  left_join(ages_inc_subs) %>%
  mutate(pos_orient = paste0('Infant ',posture, ', CG ', orientation)) # for colors

ggplot(posture_by_orientation_subset, aes(x = age.at.test, y = prop.time, col=pos_orient, size=time)) +
  geom_point(alpha=.9) +
  ylab("Proportion Time") +
  xlab("Age (months)") +
  theme_few(base_size=18) +
  ylim(0,1) +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position='none') +
  scale_color_manual(name = "Posture & CG Orientation", values=c("#006080", "#0099cc", "#99e6ff","#339966","#66cc99", "#b3e6cc",  "#e67300","#ffa64d", "#ffe6cc")) +
  facet_grid(orientation ~ posture, labeller = labeller(posture = posture_labels, orientation= orientation_labels)) +
  geom_smooth(span=10, alpha=.2)

```

```{r afterNamingsExplore, eval=FALSE, include=FALSE, echo=FALSE}
afterNamings_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceOP) %>%
  do(summarize.naming((.),window = c(0,4)))

detectionsByGroup <- dlongPlot %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  group_by(age.grp, detectorType) %>%
  multi_boot_standard(col = 'prop_detected')

afterNamings_Detections_BySub_Faces <- afterNamings_Faces %>%
  group_by(subid, familiarity, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE))

afterNamings_Detections_Grouped_Faces <- afterNamings_Detections_BySub_Faces %>%
  group_by(familiarity, age.grp) %>%
  multi_boot_standard(col = 'prop_detected')

faces_afternamings_plot <- ggplot(afterNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
  # plot baseline
  geom_pointrange(data = (detectionsByGroup %>% filter(detectorType == 'OpenPose-Faces')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=3), alpha=.9) +
  # plot averages by familiarity 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
  # plot familairity detections for each sub
  geom_point(data=afterNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
  theme_few() +
  labs(y = "Prop. face detections", x = "Age (months)") +
  theme_few(base_size=plot_size) +
  ylim(0,.6) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")

```

```{r}
### get individual points
baseline_diffs_faces <- dlong %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection)) %>%
  mutate(familiarity = 'baseline') %>%
  full_join(aroundNamings_Detections_BySub_Faces) %>%
  spread(familiarity,  prop_detected) %>%
  summarize(fam_vs_baseline = familiar - baseline, novel_vs_baseline = novel - baseline)

baseline_by_group <-  baseline_diffs_faces %>%
  group_by(age.grp) %>%
  filter(!is.na(novel_vs_baseline)) %>%
  multi_boot_standard(col='novel_vs_baseline')

ggplot(baseline_by_group, aes(x=age.grp, y=mean)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
  # plot familairity detections for each sub
  geom_point(data=afterNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
  theme_few() +
```

### 2020-08-11 analysis/4_extra_plots/extra_plotting_code.Rmd
port analysis scripts to analysis folder; rename

---
title: "extra_plotting_code"
author: "Bria Long"
date: "3/18/2020"
output: html_document
---

```{r}
# ggplot(posture, aes(y = posture, x = prop.time, fill=posture, col=posture)) + 
#   ggridges::geom_density_ridges(alpha=.7) +
#   theme_few() +
#   facet_grid(~age.grp) 
# coord_flip()

# p1=ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
#   geom_violin() + 
#   ylab("Proportion Time") + 
#   xlab("") + 
#   theme_few() +
#   scale_x_discrete( labels=c("8 m.", "12 m.", "16 m."))
# 
# p2=ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
#   geom_boxplot() + 
#   ylab("Proportion Time") + 
#   xlab("") + 
#   theme_few() +
#   scale_x_discrete( labels=c("8 m.", "12 m.", "16 m.")) +
#   theme(axis.ticks = element_blank())
# 
# ggarrange(p1,p2)
```

```{r detByPosOrientOld, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=6, fig.height=3, fig.cap = "Proportion of face / wrist detections by children's age, posture (top panel), and orientation (bottom panel)."}
# ages_inc_subs <- inc_subs %>%
#   select(subid, age.at.test, age.grp)
# 
# detectionsByOrientation <- dlongPlot %>%
#   filter(!is.na(orientation)) %>%
#   group_by(age.grp, orientation, subid, detectorType) %>%
#   summarise(prop_detected = mean(detection), num_detections = length(detection), detection_size = mean(w*h, na.rm=TRUE)) %>%
#   left_join(ages_inc_subs) %>%
#   mutate(age.at.test = as.numeric(age.at.test))
# 
# ggplot(detectionsByOrientation, aes(x = age.at.test, y = detection_size, col = detectorType, size=num_detections)) + 
#   geom_point(alpha=.5) + 
#   labs(y = "Prop. detected", x = "Age (months)") +
#   facet_grid(~detectorType) +
#   theme_few() +
#   theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
#   # scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6)) +
#   facet_grid(~orientation) +
#   geom_smooth(span=10, alpha=.2, aes(weight = num_detections), show.legend = FALSE)

```

```{r}
detectionsByPosture <- dlongPlot %>%
  filter(!is.na(posture)) %>%
  group_by(age.grp, posture, subid, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  left_join(ages_inc_subs) %>%
  mutate(age.at.test = as.numeric(age.at.test))

ggplot(detectionsByPosture, aes(x = age.at.test, y = prop_detected, col = detectorType, size=num_detections)) +
  geom_point(alpha=.5) +
  labs(y = "Prop. detected", x = "Age (months)") +
  facet_grid(~detectorType) +
  theme_few() +
  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
  scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6)) +
  facet_grid(~posture) +
  geom_smooth(span=10, alpha=.2, show.legend = FALSE)

```

```{r eval=FALSE}
p1=ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
  geom_boxplot(outlier.size = 0.5) + 
  labs(y = "Prop. detected", x = element_blank()) +
  facet_grid(~detectorType) +
  theme_few() +
  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
  scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6))

detectionsByOrientation <- dlongPlot %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, orientation, subid, detectorType) %>%
  summarise(detections = mean(detection)) %>%
  group_by(age.grp)

p2=ggplot(detectionsByOrientation, aes(x = factor(age.grp), y = detections, col = orientation)) + 
  geom_boxplot(outlier.size = 0.5) + 
  labs(y = "Prop. detected", x = element_blank()) +
  facet_grid(~detectorType) +
  theme_few() +
  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))  +
  scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6))

ggarrange(p1, p2)

```

```{r}
# naming_rate_posorient <- aroundNamings_Faces %>%
#   group_by(subid, age.grp, orientation, posture) %>%
#   summarize(count_namings= length(naming.instance)) %>%
#   left_join(posture_by_orientation) %>%
#   filter(!is.na(posture)) %>%
#   filter(!is.na(orientation)) %>%
#   mutate(naming_rate = count_namings / time) %>%
#   left_join(ages_inc_subs) %>%
#   ungroup() %>%
#   mutate(subid = fct_reorder(subid, age.at.test))

# ggplot(naming_rate_posorient, aes(x = subid, y = naming_rate)) + 
#   geom_col(aes(fill = pos_orient), position=position_stack()) +
#   ylab("Objects named  / second ") + 
#   xlab("Individual subjects (ordered by age)") +
#   theme_few() +
#   theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
#   scale_fill_manual(name = "Posture & CG Orientation", values=c("#8878cc",  "#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc",  "#e67300","#ffa64d", "#ffe6cc","#d1b673")) 
# 
# ggplot(naming_rate_posorient, aes(x = age.grp, y = naming_rate, col=posture,size=prop.time)) + 
#   geom_boxplot(alpha=.4) +
#   ylab("Objects named  / second ") + 
#   # xlab("Age (months)") +
#   theme_few() +
#   theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
#   facet_grid(~age.grp)
#   scale_color_manual(name = "Posture & CG Orientation", values=c("#8878cc",  "#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc",  "#e67300","#ffa64d", "#ffe6cc","#d1b673"))

```

```{r}
namings_prop <- aroundNamings_Faces %>%
  group_by(subid, age.grp, orientation, posture) %>%
  summarize(count_namings=length(naming.instance)) %>%
  group_by(subid) %>%
  mutate(total_namings=sum(count_namings)) %>%
  group_by(subid, age.grp, orientation, posture) %>%
  mutate(prop_namings = count_namings / total_namings) %>%
  # filter(!is.na(posture)) %>%
  # filter(!is.na(orientation)) %>%
  left_join(ages_inc_subs) %>%
  ungroup() %>%
  mutate(subid = fct_reorder(subid, age.at.test)) %>%
  mutate(pos_orient = as.factor(paste0(posture,'_',orientation)))

ggplot(namings_prop, aes(x = subid, y = prop_namings)) + 
  geom_col(aes(fill = pos_orient), position=position_stack()) +
  ylab("Proportion naming events") + 
  xlab("Individual subjects (ordered by age)") +
  theme_few() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  scale_fill_manual(name = "Posture & CG Orientation", values=c("#8878cc","#ebebeb","#ebebeb", "#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc", "#98a69f", "#e67300","#ffa64d", "#ffe6cc","#d1b673")) 
```

```{r}
## does naming rate differ by posture?

## number of namimgs in each posture/orientation combo by sub
posture_orientation_naming_count <- aroundNamings_Faces %>%
  group_by(subid, posture, orientation) %>%
  summarize(num_namings = length(naming.instance))

## time spent in each posture/orient combo for each sub
posture_orientation_time <- d %>%
  group_by(subid, posture, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE))

naming_rate_by_posture <- posture_orientation_naming_count %>%
  left_join(posture_orientation_time) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  mutate(naming_rate = num_namings / time) %>%
  group_by(posture, orientation) %>%
  multi_boot_standard(col = 'naming_rate')

## nope, not really
ggplot(naming_rate_by_posture, aes(x=posture, y=mean, col=orientation)) +
  geom_pointrange(aes(ymin=ci_lower, ymax=ci_upper), position=position_dodge(width=1)) +
  theme_few()
```

```{r}
countNamings_Raw <- aroundNamings_Faces %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE), num_namings=length(detections)) %>%
  left_join(ages_inc_subs)

countNamings_2 <- countNamings_Raw %>%
  mutate(pos_orient = as.factor(paste0('Infant ',posture, ', CG ', orientation))) %>%
  left_join(ages_inc_subs) %>%
  ungroup() %>%
  mutate(subid = fct_reorder(subid, age.at.test))

ggplot(countNamings_2, aes(x = subid, y = num_namings)) + 
  geom_col(aes(fill = pos_orient), position=position_stack()) +
  ylab("Number of naming events") + 
  xlab("Individual subjects (ordered by age)") +
  theme_few() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  scale_fill_manual(name = "Posture & CG Orientation", values=c("#8878cc", "#ebebeb", "#ebebeb", "#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc",  "#e67300","#ffa64d", "#ffe6cc","#d1b673"))

## more naming events wtih age

```

### 2020-08-11 writeup-paper/draft_apa_july2020.Rmd
port analysis scripts to analysis folder; rename

@@
 
-shorttitle        : "Automated detection of the social information"
+shorttitle        : "Automated detection of social information"
 
@@
 We found that infants\' posture and orientation to their caregiver changed dramatically across this age range and modulated infants access to social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces and hands in their visual field.
-We also applied our automated method to a second egocentric video dataset where one-year-olds explored a large play room with their caregivers  (Franchak et al., 2017), andfound convergence across both datasets, confirming the finding that motoric developments play a significant role in shaping the social information that infants have access to over development, and suggesting that they may play a role in emergence of infants\' cognitive and linguistic abiltiies.
+We also applied our automated method to a second egocentric video dataset where one-year-olds explored a large play room with their caregivers  (Franchak et al., 2017), and found convergence across both datasets, confirming the finding that motoric developments play a significant role in shaping the social information that infants have access to over development, and suggesting that they may play a role in emergence of infants\' cognitive and linguistic abiltiies.
 We propose that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants\' visual and linguistic experience as they change over development and for re-utilizing rich video datasets collected across different labs and diverse contexts.
@@
 When caregivers were behind their infants, supporting their infants' sitting or standing positions, infants saw fewer faces.
-When caregivers were relatively close to their infants, infants who were sitting or standing had more faces in view (Face detections; infant sitting and CG close, b=`r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4],3)`; infant standing and CG close, b=`r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4],3)` than infants who were lying down/crawling (i.e.
+When caregivers were relatively close to their infants, infants who were sitting or standing had more faces in view (Face detections; infant sitting and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4],3)`; infant standing and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4],4)`) than infants who were lying down/crawling (i.e.
 prone).
@@
 We found a similar pattern of results for wrist detections, even though there were fewer wrist detections overall in the dataset.
-Infants saw fewer wrists when caregivers were behind their infants, supporting their infants' sitting or standing positions vs when caregivers were relatively closer to their infants (Hand detections; infant sitting and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4],4)`; infant standing and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4],4)` than infants who were lying down/crawling (i.e.
+Infants saw fewer wrists when caregivers were behind their infants, supporting their infants' sitting or standing positions vs when caregivers were relatively closer to their infants (Hand detections; infant sitting and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4],4)`; infant standing and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4],4)`) than infants who were lying down/crawling (i.e.
 prone).
-Wrist detections were clearly highest when caregivers were far away from their infants (Wrist detections; infant sitting and CG far, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4],4)`; infant standing and CG far, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4],4)`.
+Wrist detections were  highest when caregivers were far away from their infants and those infants were standing (Wrist detections;infant standing and CG far, b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4],4)`.
 As with faces, age did not remain a significant predictor in these models (Wrist detections; Age (scaled), b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4],4)`).

### 2020-08-11 writeup-paper/draft_apa_july2020.tex
port analysis scripts to analysis folder; rename

@@
 \hypersetup{
-            pdftitle={Using automated detections to quantify how postural developments shape the social information in view for young infants},
+            pdftitle={Using automated detections to quantify the effect of postural developments on the social information in the infant view},
             pdfkeywords={Postural developments, headcam, social information, face detection, pose
@@
 
-\title{Using automated detections to quantify how postural developments shape
-the social information in view for young infants}
+\title{Using automated detections to quantify the effect of postural
+developments on the social information in the infant view}
 
@@
 cognitive development. Yet infants locomotor abilities also develop
-continuously as they learn to sit, crawl, stand, and walk -- influencing
-the way that infants see the world around them. How do these postural
-developments affect infants access to the social information relevant
-for early learning? To address this question, we first created a rich
-annotated dataset of naturalistic play sessions between infants and
-their caregivers with sets of novel and familiar toys (N=36, 8-16 months
-of age), recording both egocentric and third-person video. We then
-developed an automated analysis method using a pose detection model
+continuously as they learn to sit, crawl, stand, and walk. How do these
+postural developments affect infants access to the social information
+relevant for early learning? To address this question, we first created
+a rich annotated dataset of naturalistic play sessions between infants
+and their caregivers with sets of novel and familiar toys (N=36, 8-16
+months of age), recording both egocentric and third-person video. We
+then developed an automated analysis method using a pose detection model
 (Zhang et al., 2017) to detect the faces and hands in the infant view,
-allowing us to analyze the entirety of this dataset. We also applied
-this automated method to second egocentric video dataset where
-one-year-olds explored a large play area with their caregivers (Franchak
-et al., 2017). We found that infants' posture and orientation to their
-caregiver changed dramatically across this age range and modulated
-infants access to social information; infants who were sitting or
-standing with their caregiver at a close distance tended to have the
-most faces and hands in their visual field. We also found convergence
+allowing us to analyze the entirety of this dataset. We found that
+infants' posture and orientation to their caregiver changed dramatically
+across this age range and modulated infants access to social
+information; infants who were sitting or standing with their caregiver
+at a close distance tended to have the most faces and hands in their
+visual field. We also applied our automated method to a second
+egocentric video dataset where one-year-olds explored a large play room
+with their caregivers (Franchak et al., 2017), and found convergence
 across both datasets, confirming the finding that motoric developments
@@
 role in emergence of infants' cognitive and linguistic abiltiies. We
-suggest that the combined use of head-mounted cameras and the
+propose that the combined use of head-mounted cameras and the
 application of new computer vision techniques is a promising avenue for
@@
 What social cues do infants see as they learn language, and how does
-this change as they grow and are start to locomote themselves in their
+this change as they grow and start to locomote themselves in their
 environments? We explored this question using video data from

### 2020-08-11 analysis/1_merge_data/helper.R

@@ -1,0 +1,259 @@
+library(stringr)
+
+######## MERGE IN EXACT TIMES FOR FRAMES #########
+add.times <- function(x) {
+  print(x$subid[1])
+  ## from initial submission
+  fname <- paste0("../../data/frame_times_redone/",
+                 str_sub(as.character(x$subid[1]),start=4,end=8),
+                 ".csv")
+  
+  if (file.exists(fname)) {
+    times <- read_csv(fname)
+    times$time <- times$best_time ## comes from show_frames python script
+    x <- left_join(x, times)
+    x$dt <- c(diff(x$time),.032) ## adds in 33ms for the last frame but otherwise goes on actual frame times, making an fps rate differences not a problem
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing frames ****"))
+    x$time <- NA
+    x$dt <- NA
+  }
+  return(x)
+}
+
+######## ADD POSTURE CODING #########
+add.posture <- function(x) {
+  fname <- paste("../../data/posture/",
+                 x$subid[1],
+                 ".csv",
+                 sep="")
+  
+  if (file.exists(fname)) {
+    print(fname)
+    postures <- read_csv(fname)
+    postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order
+   
+    ## this reorders the postures so that they are "zeroed" relative to the beginning of the headcam videos
+     postures <- regularize.postures(subset(postures,code=="posture"),x$subid[1]) ## needs subid to get real sync time later 
+    x$posture <- factor(NA,levels=levels(postures$posture))
+    x$orientation <- factor(NA,levels=levels(postures$orientation))
+    
+    # for each row of p, populate posture and orientation to x
+    for (i in 1:nrow(postures)) {
+      range <- x$time > postures$start[i] & x$time <= postures$end[i]
+      x$posture[range] <- postures$posture[i]
+      x$orientation[range] <- postures$orientation[i]
+    }
+  } else {
+    print(paste0("**** ", x$subid[1], " is missing posture ****"))
+    x$posture <- NA
+    x$orientation <- NA
+  }
+  
+  return(x)
+}
+
+add.posture.second <- function(x) {
+  fname <- paste("../../data/posture/second/",
+                 x$subid[1],
+                 ".csv",
+                 sep="")
+  
+  if (file.exists(fname)) {
[…]

### 2020-08-11 writeup-paper/helper.R

@@ -1,259 +1,0 @@
-library(stringr)
-
-######## MERGE IN EXACT TIMES FOR FRAMES #########
-add.times <- function(x) {
-  print(x$subid[1])
-  ## from initial submission
-  fname <- paste0("../data/frame_times_redone/",
-                 str_sub(as.character(x$subid[1]),start=4,end=8),
-                 ".csv")
-  
-  if (file.exists(fname)) {
-    times <- read_csv(fname)
-    times$time <- times$best_time ## comes from show_frames python script
-    x <- left_join(x, times)
-    x$dt <- c(diff(x$time),.032) ## adds in 33ms for the last frame but otherwise goes on actual frame times, making an fps rate differences not a problem
-  } else {
-    print(paste0("**** ", x$subid[1], " is missing frames ****"))
-    x$time <- NA
-    x$dt <- NA
-  }
-  return(x)
-}
-
-######## ADD POSTURE CODING #########
-add.posture <- function(x) {
-  fname <- paste("../data/posture/",
-                 x$subid[1],
-                 ".csv",
-                 sep="")
-  
-  if (file.exists(fname)) {
-    print(fname)
-    postures <- read_csv(fname)
-    postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order
-   
-    ## this reorders the postures so that they are "zeroed" relative to the beginning of the headcam videos
-     postures <- regularize.postures(subset(postures,code=="posture"),x$subid[1]) ## needs subid to get real sync time later 
-    x$posture <- factor(NA,levels=levels(postures$posture))
-    x$orientation <- factor(NA,levels=levels(postures$orientation))
-    
-    # for each row of p, populate posture and orientation to x
-    for (i in 1:nrow(postures)) {
-      range <- x$time > postures$start[i] & x$time <= postures$end[i]
-      x$posture[range] <- postures$posture[i]
-      x$orientation[range] <- postures$orientation[i]
-    }
-  } else {
-    print(paste0("**** ", x$subid[1], " is missing posture ****"))
-    x$posture <- NA
-    x$orientation <- NA
-  }
-  
-  return(x)
-}
-
-add.posture.second <- function(x) {
-  fname <- paste("../data/posture/second/",
-                 x$subid[1],
-                 ".csv",
-                 sep="")
-  
-  if (file.exists(fname)) {
[…]

### 2020-08-12 analysis/1_merge_data/helper.R
fix paths

@@ -1,10 +1,11 @@
 library(stringr)
+library(here)
 
 ######## MERGE IN EXACT TIMES FOR FRAMES #########
 add.times <- function(x) {
   print(x$subid[1])
   ## from initial submission
-  fname <- paste0("../../data/frame_times_redone/",
+  fname <- paste0(here::here(), "/data/frame_times_redone/",
                  str_sub(as.character(x$subid[1]),start=4,end=8),
                  ".csv")
   
@@ -23,7 +24,7 @@ add.times <- function(x) {
 
 ######## ADD POSTURE CODING #########
 add.posture <- function(x) {
-  fname <- paste("../../data/posture/",
+  fname <- paste(here::here(), "/data/posture/",
                  x$subid[1],
                  ".csv",
                  sep="")
@@ -54,7 +55,7 @@ add.posture <- function(x) {
 }
 
 add.posture.second <- function(x) {
-  fname <- paste("../../data/posture/second/",
+  fname <- paste(here::here(),"/data/posture/second/",
                  x$subid[1],
                  ".csv",
                  sep="")
@@ -88,7 +89,7 @@ add.posture.second <- function(x) {
 
 
 get_sub_sync_time <- function(this_sub_id){
-  sync_times_check_file <- read_csv("../../data/manual_sync_times/video_sync_times.csv") 
+  sync_times_check_file <- read_csv(paste0(here::here(),"/data/manual_sync_times/video_sync_times.csv")) 
   s <- sync_times_check_file %>%
     filter(sid == this_sub_id) %>%
     mutate(sync_time_string =toString(sync_time_stamp)) %>%
@@ -206,7 +207,7 @@ naming.instance <- function(ns) {
 
 ######## GET SUMMARY MEASURES OVER NAMINGS #########
 
-get.namings <- function (x, window = c(-2,2), file_dir = '../../data/naming/') {  
+get.namings <- function (x, window = c(-2,2), file_dir = paste0(here::here(),'/data/naming/')) {  
   # read in naming times
   namings <- read.csv(paste(file_dir,
                             x$subid[1],
@@ -217,7 +218,7 @@ get.namings <- function (x, window = c(-2,2), file_dir = '../../data/naming/') {
   namings <- regularize.naming(namings)
 }
 
-get.namings.second <- function (x, window = c(-2,2), file_dir = '../../data/naming/second/') {
[…]

### 2020-08-12 analysis/4_extra_plots/SI_extraplots.Rmd
port old plotting code to backup

@@
 ```
+
+### Analyze detections around naming events (without posture/orientation)
+
+```{r include=FALSE}
+# difference plots can be confusing, ommitting
+# faces_namings_diffs_indiv <- aroundNamings_Detections_BySub_Fam_Faces_All %>%
+#   spread(familiarity, prop_detected) %>%
+#   mutate(fam_dif = familiar - baseline, nov_diff = novel - baseline) 
+#
+# ggplot(faces_namings_diffs_indiv, aes(x=age.grp, y=nov_diff, group=age.grp)) +
+#   geom_boxplot() +
+#   theme_few()
+
+## explored looking at namings while acconting for posture and orientation
+# aroundNamings_Detections_BySub_Posture_Faces <- aroundNamings_Faces %>%
+#   group_by(subid, familiarity, posture, orientation, age.grp) %>%
+#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
+#   mutate(detectorType = 'OpenPose-Wrists')
+# 
+# baseline_diffs_by_posture <- dlong %>%
+#   filter(detectorType == 'OpenPose-Wrists') %>%
+#   group_by(age.grp, subid, posture, orientation,detectorType) %>%
+#   summarise(prop_detected = mean(detection)) %>%
+#   mutate(familiarity = 'baseline') %>%
+#   full_join(aroundNamings_Detections_BySub_Posture_Faces)
+# 
+# naming = lmer(data = aroundNamings_Detections_BySub_Posture_Faces, prop_detected ~ familiarity + posture * orientation + (1|subid))
+# summary(naming)
+```
+
+```{r setUpNamingsPlot}
+
+## first get baseline
+# baselineFaces <- dlongPlot %>%
+#   filter(detectorType == 'OpenPose-Faces') %>%
+#   group_by(age.grp, subid, detectorType) %>%
+#   summarise(prop_detected = mean(detection)) %>%
+#   mutate(familiarity='baseline')
+# 
+# baselineWrists <- dlongPlot %>%
+#   filter(detectorType == 'OpenPose-Wrists') %>%
+#   group_by(age.grp, subid, detectorType) %>%
+#   summarise(prop_detected = mean(detection)) %>%
+#   mutate(familiarity='baseline')
+# 
+# 
+# ### then by faces
+# aroundNamings_Detections_BySub_Faces <- aroundNamings_Faces %>%
+#   group_by(subid, familiarity, age.grp) %>%
+#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
+#   mutate(detectorType = 'OpenPose-Faces') %>%
+#   full_join(baselineFaces)
+# 
+# aroundNamings_Detections_Grouped_Faces <- aroundNamings_Detections_BySub_Faces %>%
+#   group_by(familiarity, age.grp) %>%
+#   multi_boot_standard(col = 'prop_detected')  
+# 
+# ### now by wrists 
+# aroundNamings_Detections_BySub_Wrists <- aroundNamings_Wrists %>%
+#   group_by(subid, familiarity, age.grp) %>%
+#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
+#   mutate(detectorType = 'OpenPose-Wrists') %>%
+#   full_join(baselineWrists)
+# 
+# aroundNamings_Detections_Grouped_Wrists <- aroundNamings_Detections_BySub_Wrists %>%
+#   group_by(familiarity, age.grp) %>%
+#   multi_boot_standard(col = 'prop_detected')
+
+```
+
+```{r detByNamingPlotPanels}
+## Plot faces/wrists around namings by age
+
+# faces_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
+#   # plot averages by familiarity 
+#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=2), alpha=.9) +
+#   # plot familairity detections for each sub
+#   geom_point(data=aroundNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=2), alpha=.2) +
+#   theme_few() +
+#   labs(y = "Prop. face detections", x = "Age (months)") +
+#   theme_few(base_size=plot_size) +
+#   ylim(0,.4) +
+#   theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
+#   scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
+#   scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18))
+
+## 
+# wrists_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Wrists, aes(x=age.grp, y=mean, color=familiarity)) +
+#   # plot averages by familiarity 
+#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=2), alpha=.9) +
+#   # plot familairity detections for each sub
+#   geom_point(data=aroundNamings_Detections_BySub_Wrists, aes(x=age.grp, y=prop_detected), position = position_dodge(width=2), alpha=.2) +
+#   theme_few() +
+#   labs(y = "Prop. wrist detections", x = "Age (months)") +
+#   theme_few(base_size=plot_size) +
+#   ylim(0,.4) +
+#   theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
+#   scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
+#   scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) 
+```

### 2020-08-12 writeup-paper/draft_apa_july2020.Rmd
latest version of ms

@@
 We found that infants\' posture and orientation to their caregiver changed dramatically across this age range and modulated infants access to social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces and hands in their visual field.
-We also applied our automated method to a second egocentric video dataset where one-year-olds explored a large play room with their caregivers  (Franchak et al., 2017), and found convergence across both datasets, confirming the finding that motoric developments play a significant role in shaping the social information that infants have access to over development, and suggesting that they may play a role in emergence of infants\' cognitive and linguistic abiltiies.
-We propose that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants\' visual and linguistic experience as they change over development and for re-utilizing rich video datasets collected across different labs and diverse contexts.
+We also applied our automated method to a second egocentric video dataset where one-year-olds explored a large play room with their caregivers  (Franchak et al., 2017), replicating and extending their findings.
+Together, these results establish the validity of a new method for detecting social information, confirm that motoric developments play a significant role in shaping the social information that infants have access to.
+We propose that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for re-utilizing rich video datasets collected across different labs and diverse contexts and understanding the statistics of infants\' visual and linguistic experience as they change over development.
 
@@
 library(lmerTest)
-
+```
+
+```{r} 
+## Sources critical functions for merging data files amd timestamps
+source("../analysis/1_merge_data/helper.R")
+```
+
+```{r}
+## Cosmetic rendering functions/settings
 # base_size in theme_few (changes font size in plots)
 plot_size=16
+
+reportP <- function(pValue){
+  if (pValue < 0.001){
+    result <- "P < 0.001"
+  } else {
+    result <- sprintf("P = %.3f", pValue) # inserts a float into a string and simultaneously do rounding
+  }
+  return(result)
+}
 ```
@@
 We then apply this automated method to a head-mounted camera dataset collected in our lab to directly examine the role of postural developments on the social information in the infant view during language learning.
-Specifically, we constructed a dataset where infants and caregivers played freely using a set of novel and familiar objects; 8, 12, and 16 months-olds (N=26; N=12 in each age group) participated to ensure a wide range of locomotor abilities.
+Specifically, we constructed a dataset where infants and caregivers played freely using a set of novel and familiar objects; 8, 12, and 16 months-olds (N=36; N=12 in each age group) participated to ensure a wide range of locomotor abilities.
 Infants’ in-the-moment posture and orientation relative to their caregiver were hand annotated and all play sessions were transcribed.
@@
 ```{r echo=FALSE, include=FALSE, demographics}
-all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
+all_subs <- read.csv(paste0(here::here(),"/data/demographics/all_participants.csv")) %>%
   rename(subid = subID)
-inc_subs <- read.csv("../data/demographics/demographics.csv")
+inc_subs <- read.csv(paste0(here::here(),"/data/demographics/demographics.csv"))
 all_subs$inc <- all_subs$reason=="included"
@@
 ### READ IN DETECTIONS
-d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_july2020.csv") %>%
+d <- read_csv(paste0(here::here(),"/data/consolidated_data/consolidated_data_4detectors_july2020.csv")) %>%
   mutate(posture = factor(posture), orientation = factor(orientation))
@@
 Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations; similarily, times when the child was being carried or the caregivers were out of the frame were marked as uncodable for caregiver orientation.
-On average, posture or orientation was uncodable from 1-2 minutes of data in each child (seconds excluded from analysis for posture;.
-M =`r NA_rates$mean_na_posture`, SD =`r NA_rates$sd_na_posture`; orientation; M =`r NA_rates$mean_na_posture`, SD =`r NA_rates$sd_na_posture`), and these rates did not vary substantially with the age of the child.
+On average, posture or orientation was uncodable from 1-2 minutes of data in each child (seconds excluded from analysis for posture, M = `r round(NA_rates$mean_na_posture,2)`s, SD = `r round(NA_rates$sd_na_posture,2)`s; orientation; M = `r round(NA_rates$mean_na_posture,2)`s, SD = `r round(NA_rates$sd_na_posture,2)`s), and these rates did not vary substantially with the age of the child.
 To assess the reliability of these annotations, a second coder coded videos from three different children to calculate Cohen's kappa (posture, irr = XX, orientation, irr = XX).
@@
 ### READ IN GROUND TRUTH and load function
-ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv") %>%
+ground_truth_faces <- read_csv(paste0(here::here(),"/data/ground_truth/ground_truth_faces.csv")) %>%
   mutate(sample_type=as.factor(sample_type))
 
-ground_truth_wrists <- read_csv("../data/ground_truth/ground_truth_wrists.csv") %>%
+ground_truth_wrists <- read_csv(paste0(here::here(),"/data/ground_truth/ground_truth_wrists.csv")) %>%
   mutate(sample_type=as.factor(sample_type))
@@
 Despite the fact that hand detection is a more computationally challenging problem [@bambach2017], and the fact that we used wrist keypoints as a proxy for hands,  OpenPose performed moderately well as a hand detector (F = `r round(perfWristsOverall$f[1],2)`).
-OpenPose achieved relatively high precision -- generating relatively few false positives -- but showedlow recall on the randomly sampled frames (see Table 1).
+OpenPose achieved relatively high precision -- generating relatively few false positives -- but showed low recall on the randomly sampled frames (see Table 1).
 As with face detections, when we restricted our analysis to high-confidence detections, we found much higher precision (P = `r round(perfWristsOverall$p[2],2)`), but much lower recall (R = `r round(perfWristsOverall$r[2],2)`) and thus overall performance (F = `r round(perfWristsOverall$f[2],2)`).
@@
 
-## Access to social information during postural developments
+## Study 1: Access to social information during postural developments
 
@@
 When caregivers were behind their infants, supporting their infants' sitting or standing positions, infants saw fewer faces.
-When caregivers were relatively close to their infants, infants who were sitting or standing had more faces in view (Face detections; infant sitting and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4],3)`; infant standing and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4],4)`) than infants who were lying down/crawling (i.e.
+When caregivers were relatively close to their infants, infants who were sitting or standing had more faces in view (Face detections; infant sitting and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4])`; infant standing and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4])`) than infants who were lying down/crawling (i.e.
 prone).
-When caregivers were far away from their infants, face detections were similarity higher (Face detections; infant sitting and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4],4)`; infant standing and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4],4)`.
-Overall, we found that age did not remain a significant predictor in accounting for the faces in view when modeling infants posture, their orientation to their caregiver, and the interaction between these two factors (Face detections; Age (scaled), b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, P = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4],4)`).
+When caregivers were far away from their infants, face detections were similarity higher (Face detections; infant sitting and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, P = `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4])`; infant standing and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, P = `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`.
+Overall, we found that age did not remain a significant predictor in accounting for the faces in view when modeling infants posture, their orientation to their caregiver, and the interaction between these two factors (Face detections; Age (scaled), b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, P = `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
 
 We found a similar pattern of results for wrist detections, even though there were fewer wrist detections overall in the dataset.
-Infants saw fewer wrists when caregivers were behind their infants, supporting their infants' sitting or standing positions vs when caregivers were relatively closer to their infants (Hand detections; infant sitting and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4],4)`; infant standing and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4],4)`) than infants who were lying down/crawling (i.e.
+Infants saw fewer wrists when caregivers were behind their infants, supporting their infants' sitting or standing positions vs when caregivers were relatively closer to their infants (Hand detections; infant sitting and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`,  `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4])`; infant standing and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`,  `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4])`) than infants who were lying down/crawling (i.e.
 prone).
-Wrist detections were  highest when caregivers were far away from their infants and those infants were standing (Wrist detections;infant standing and CG far, b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4],4)`.
-As with faces, age did not remain a significant predictor in these models (Wrist detections; Age (scaled), b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, P = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4],4)`).
+Wrist detections were  highest when caregivers were far away from their infants and those infants were standing (Wrist detections;infant standing and CG far, b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
+As with faces, age did not remain a significant predictor in these models (Wrist detections; Age (scaled), b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
 
@@
 ```{r summarizeNamings}
-source("helper.R")
 ## summarize detections around naming events
@@
 
-```{r basicNamingStatsFaces}
-## inferential stats: namings vs baseline 
+```{r basicNamingStatsFaces, include=FALSE}
+## Face detedctinos around namings sans familiarity 
 aroundNamings_Detections_BySub_Faces <- aroundNamings_Faces %>%
@@
 
-naming_faces_model = lmer(prop_detected ~  scale(age.at.test) + when +  (1|subid), data = baselineDiffs_by_Naming_Faces)
-
-faces_naming_out = summary(naming_faces_model)
-```
-
-```{r}
-#### wrists
+lmer(prop_detected ~  scale(age.at.test) + when +  (1|subid), data = baselineDiffs_by_Naming_Faces)
+```
+
+```{r include=FALSE}
+#### Wrists namings without familiarity 
 aroundNamings_Detections_BySub_Wrists <- aroundNamings_Wrists %>%
@@
 
-naming_wrists_model = lmer(prop_detected ~ scale(age.at.test) + when + (1|subid), data = baselineDiffs_by_Naming_Wrists)
-
-wrists_naming_out = summary(naming_wrists_model)
-
+lmer(prop_detected ~ scale(age.at.test) + when + (1|subid), data = baselineDiffs_by_Naming_Wrists)
 ```
@@
 ```{r include=FALSE}
-## explored looking at namings while acconting for familairity
+## Face detections during namings, coefficeints actually not outputted
 aroundNamings_Detections_BySub_Fam_Faces <- aroundNamings_Faces %>%
@@
 
-naming_faces_fam = lmer(data = aroundNamings_Detections_BySub_Fam_Faces_All, prop_detected ~ scale(age.at.test) + familiarity +  (1|subid))
-summary(naming_faces_fam)
+lmer(data = aroundNamings_Detections_BySub_Fam_Faces_All, prop_detected ~ scale(age.at.test) + familiarity +  (1|subid))
+
 ```
@@
 ```{r}
-## explored looking at namings while acconting for familairity
+## Wrist detections during namings 
 aroundNamings_Detections_BySub_Fam_Wrists <- aroundNamings_Faces %>%
@@
 
-```{r include=FALSE}
-# difference plots can be confusing, ommitting
-# faces_namings_diffs_indiv <- aroundNamings_Detections_BySub_Fam_Faces_All %>%
-#   spread(familiarity, prop_detected) %>%
-#   mutate(fam_dif = familiar - baseline, nov_diff = novel - baseline) 
-#
-# ggplot(faces_namings_diffs_indiv, aes(x=age.grp, y=nov_diff, group=age.grp)) +
-#   geom_boxplot() +
-#   theme_few()
-
-## explored looking at namings while acconting for posture and orientation
-# aroundNamings_Detections_BySub_Posture_Faces <- aroundNamings_Faces %>%
-#   group_by(subid, familiarity, posture, orientation, age.grp) %>%
-#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
-#   mutate(detectorType = 'OpenPose-Wrists')
-# 
-# baseline_diffs_by_posture <- dlong %>%
-#   filter(detectorType == 'OpenPose-Wrists') %>%
-#   group_by(age.grp, subid, posture, orientation,detectorType) %>%
-#   summarise(prop_detected = mean(detection)) %>%
-#   mutate(familiarity = 'baseline') %>%
-#   full_join(aroundNamings_Detections_BySub_Posture_Faces)
-# 
-# naming = lmer(data = aroundNamings_Detections_BySub_Posture_Faces, prop_detected ~ familiarity + posture * orientation + (1|subid))
-# summary(naming)
-```
-
 ```{r}
+## output means for in-line text rendering
 faces_namings_diffs_out <- aroundNamings_Detections_BySub_Fam_Faces_All %>%
@@
 Face detections were numerically slightly higher around these novel naming events relative to baseline, with similar effects across age groups
-(8 m.o., $M_{fam - baseline}$ = =`r round(faces_namings_diffs_out$nov_vs_baseline[1],3)`,
-12 m.o., $M_{nov - baseline}$ =`r round(faces_namings_diffs_out$nov_vs_baseline[2],3)`,
-16 m.o. $M_{nov - baseline}$ =`r round(faces_namings_diffs_out$nov_vs_baseline[3],3)`)
+(8 m.o., $M_{fam - baseline}$ = `r round(faces_namings_diffs_out$nov_vs_baseline[1],3)`,
+12 m.o., $M_{nov - baseline}$ = `r round(faces_namings_diffs_out$nov_vs_baseline[2],3)`,
+16 m.o. $M_{nov - baseline}$ = `r round(faces_namings_diffs_out$nov_vs_baseline[3],3)`)
 however, they were not higher during familiar naming events vs. baseline (8 m.o., $M_{fam - baseline}$ =`r round(faces_namings_diffs_out$fam_vs_baseline[1],3)`,
-12 m.o., $M_{fam - baseline}$ =`r round(faces_namings_diffs_out$fam_vs_baseline[2],3)`,
-16 m.o. $M_{fam - baseline}$ =`r round(faces_namings_diffs_out$fam_vs_baseline[3],3)`,
+12 m.o., $M_{fam - baseline}$ = `r round(faces_namings_diffs_out$fam_vs_baseline[2],3)`,
+16 m.o. $M_{fam - baseline}$ = `r round(faces_namings_diffs_out$fam_vs_baseline[3],3)`).
 Conversely, wrist detections were higher during both familiar (8 m.o., $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$fam_vs_baseline[1],3)`,
-12 m.o., $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$fam_vs_baseline[2],3)`,
-16 m.o. $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$fam_vs_baseline[3],3)`)
-and novel naming events relative to baseline across all age groups (Novel; 8 m.o., (8 m.o.,  $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$nov_vs_baseline[1],3)`,
-12 m.o., $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$nov_vs_baseline[2],3)`,
+12 m.o., $M_{fam - baseline}$ = `r round(wrist_namings_diffs_out$fam_vs_baseline[2],3)`,
+16 m.o. $M_{fam - baseline}$ = `r round(wrist_namings_diffs_out$fam_vs_baseline[3],3)`)
+and novel naming events relative to baseline across all age groups (8 m.o.,  $M_{nov - baseline}$ = `r round(wrist_namings_diffs_out$nov_vs_baseline[1],3)`,
+12 m.o., $M_{nov - baseline}$ = `r round(wrist_namings_diffs_out$nov_vs_baseline[2],3)`,
 16 m.o.
-$M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$nov_vs_baseline[3],3)`).
+$M_{nov - baseline}$ = `r round(wrist_namings_diffs_out$nov_vs_baseline[3],3)`).
 These results were confirmed by a linear mixed-effect model with scaled aged as a fixed effect and random intercepts for each subject (Wrist detections; familiar objects vs.
-baseline, b=`r round(naming_wrists_model_out$coefficients[3,1],2)`, SE  = `r round(naming_wrists_model_out$coefficients[3,2],2)`, t = `r round(naming_wrists_model_out$coefficients[3,4],2)`, P = `r round(naming_wrists_model_out$coefficients[3,5],4)`'; Novel objects vs.
-baseline, b=`r round(naming_wrists_model_out$coefficients[4,1],2)`, SE  = `r round(naming_wrists_model_out$coefficients[4,2],2)`, t = `r round(naming_wrists_model_out$coefficients[4,4],2)`, P = `r round(naming_wrists_model_out$coefficients[4,5],4)`.
+baseline, b = `r round(naming_wrists_model_out$coefficients[3,1],2)`, SE  = `r round(naming_wrists_model_out$coefficients[3,2],2)`, t = `r round(naming_wrists_model_out$coefficients[3,4],2)`, `r reportP(naming_wrists_model_out$coefficients[3,5])`; Novel objects vs.
+baseline, b = `r round(naming_wrists_model_out$coefficients[4,1],2)`, SE  = `r round(naming_wrists_model_out$coefficients[4,2],2)`, t = `r round(naming_wrists_model_out$coefficients[4,4],2)`, `r reportP(naming_wrists_model_out$coefficients[4,5])`).
 
 Overall, these exploratory results suggest that children may tend to see more hands around naming events.
-This finding is consistent with the possibility that caregivers may change how they interact with their infant when presenting them with objects [@gogate2000study, @gogate2006attention].
+This finding is consistent with the possibility that caregivers may change how they interact with their infant when presenting them with objects [@gogate2000study; @gogate2006attention].
 For example, caregivers may tend to simultaneously name objects when demonstrating their affordances or simply when pointing to them.
-In turn, infants may be sensitive to these naming events and orient their attention towards their caregiver, consistent with other accounts positing infants' sensitivity to social cues in early word learning environments [@yurovsky2018communicative, @yurovsky2017beyond, @frank2013].
-
-```{r setUpNamingsPlot}
-### Analyze detections around naming events (without posture/orientation)
-
-## first get baseline
-# baselineFaces <- dlongPlot %>%
-#   filter(detectorType == 'OpenPose-Faces') %>%
-#   group_by(age.grp, subid, detectorType) %>%
-#   summarise(prop_detected = mean(detection)) %>%
-#   mutate(familiarity='baseline')
-# 
-# baselineWrists <- dlongPlot %>%
-#   filter(detectorType == 'OpenPose-Wrists') %>%
-#   group_by(age.grp, subid, detectorType) %>%
-#   summarise(prop_detected = mean(detection)) %>%
-#   mutate(familiarity='baseline')
-# 
-# 
-# ### then by faces
-# aroundNamings_Detections_BySub_Faces <- aroundNamings_Faces %>%
-#   group_by(subid, familiarity, age.grp) %>%
-#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
-#   mutate(detectorType = 'OpenPose-Faces') %>%
-#   full_join(baselineFaces)
-# 
-# aroundNamings_Detections_Grouped_Faces <- aroundNamings_Detections_BySub_Faces %>%
-#   group_by(familiarity, age.grp) %>%
-#   multi_boot_standard(col = 'prop_detected')  
-# 
-# ### now by wrists 
-# aroundNamings_Detections_BySub_Wrists <- aroundNamings_Wrists %>%
-#   group_by(subid, familiarity, age.grp) %>%
-#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
-#   mutate(detectorType = 'OpenPose-Wrists') %>%
-#   full_join(baselineWrists)
-# 
-# aroundNamings_Detections_Grouped_Wrists <- aroundNamings_Detections_BySub_Wrists %>%
-#   group_by(familiarity, age.grp) %>%
-#   multi_boot_standard(col = 'prop_detected')
-
-```
-
-```{r detByNamingPlotPanels}
-## Plot faces/wrists around namings by age
-
-# faces_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
-#   # plot averages by familiarity 
-#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=2), alpha=.9) +
-#   # plot familairity detections for each sub
-#   geom_point(data=aroundNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=2), alpha=.2) +
-#   theme_few() +
-#   labs(y = "Prop. face detections", x = "Age (months)") +
-#   theme_few(base_size=plot_size) +
-#   ylim(0,.4) +
-#   theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
-#   scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
-#   scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18))
-
-## 
-# wrists_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Wrists, aes(x=age.grp, y=mean, color=familiarity)) +
-#   # plot averages by familiarity 
-#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=2), alpha=.9) +
-#   # plot familairity detections for each sub
-#   geom_point(data=aroundNamings_Detections_BySub_Wrists, aes(x=age.grp, y=prop_detected), position = position_dodge(width=2), alpha=.2) +
-#   theme_few() +
-#   labs(y = "Prop. wrist detections", x = "Age (months)") +
-#   theme_few(base_size=plot_size) +
-#   ylim(0,.4) +
-#   theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
-#   scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
-#   scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) 
-```
+In turn, infants may be sensitive to these naming events and orient their attention towards their caregiver, consistent with other accounts positing infants' sensitivity to social cues in early word learning environments [@yurovsky2018communicative; @yurovsky2017beyond; @frank2013].
 
 ```{r}
+## set up naming plots
 age_labels = c('8' = '8 m.o.','12' = '12 m.o.','16' = ' 16 m.o.')
@@
 
-## Extension to Franchak et a., 2017
+## Study 2: Extension of Franchak et al., 2017
 In the present work, we found that infant's in-the-moment posture changed with their age, as did infants' orientation relative their caregiver.
 In a related study with 12-month-olds, @franchak2017see also found that infants' in-the-moment posture changed the proportion of time infants spent looking at faces.
-Here, we sought to replicate these findings using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras [@simon2015databrary] for two reasons.
+Here, we sought to replicate these findings using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras [hosted on Databrary; @simon2015databrary] for two reasons.
 First, we sought to validate our novel method, which could fail to generalize to scenes from these more complex environments, where detecting faces and hands could arguably be a much harder task.
@@
 ### Comparison between video datasets 
-First, despite having different head-mounted camera setups, the view angle of the two head-mounted cameras were only slightly different ($52.2^{\circ}$ horizontal by $42.2^{\circ}$ in @franchak2017see, here $64^{\circ}$ horizontal by $46^{\circ}$ vertical).
-However, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to climb, explore, and interact with, and infants were unconstrained and allowed to freely wander the room.
+First, despite having different head-mounted camera setups, the view angle of the two head-mounted cameras used in Study 1 and @franchak2017see were only slightly different ($52.2^{\circ}$ horizontal by $42.2^{\circ}$ in @franchak2017see, here $64^{\circ}$ horizontal by $46^{\circ}$ vertical).
+However, the play environment in @franchak2017see that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to climb, explore, and interact with, and infants were unconstrained and allowed to freely wander the room.
 In contrast, the play room used in Study 1 was relatively small (approximately 10' x 10') and was set-up with focused play with the pairs of novel and familiar objects.
@@
 # loads pos_detections used below for franchak dataset
-load(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections_by_posture_july2020.RData')
+load(file = paste0(here::here(),'/data/franchak_dataset/preprocessed/franchak_dataset_detections_by_posture_july2020.RData'))
 ```
@@
 ```{r processFranchak}
+# preprocess franchak data
 faces_by_pos  <- pos_detections %>%
@@
 
-### Replication using automated detections
 ```{r}
+## means for outputting in the text
 detections_by_conf_by_sub <- pos_detections %>%
@@
 
+### Replication using automated detections
 Overall, we found convergence between our two methodologies, replicating the main results from @franchak2017see, and finding that the proportion of faces detected was greater when infants were sitting or standing vs.
 prone.
-We found this result regardless of whether we used all detections (see Figure \ref{fig:franchak}, panel A; (proportion of frames with face detections; Prone: M =`r mean(dets_by_conf_by_posture$avg_faces[1])`%, Sitting: M =`r mean(dets_by_conf_by_posture$avg_faces[2])`%, Upright: M =`r mean(dets_by_conf_by_posture$avg_faces[3])`%) or restricted our analyses to high-confidence detections (percentage of frames with face detections; Prone: M =`r mean(dets_by_conf_by_posture$avg_faces_strict[1])`, Sitting: M =`r mean(dets_by_conf_by_posture$avg_faces_strict[2])`, Upright: M =`r mean(dets_by_conf_by_posture$avg_faces_strict[3])`).
+We found this result regardless of whether we used all detections (see Figure \ref{fig:franchak}, panel A; (proportion of frames with face detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_faces[1])`, Sitting: M = `r mean(dets_by_conf_by_posture$avg_faces[2])`, Upright: M = `r mean(dets_by_conf_by_posture$avg_faces[3])`) or restricted our analyses to high-confidence detections (percentage of frames with face detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[1])`, Sitting: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[2])`, Upright: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[3])`).
 These results were confirmed in generalized linear-mixed models, as in the first study (Sitting vs.
-Prone, b=`r round(franchak_faces_out$coefficients[2,1],3)`, SE  = `r round(franchak_faces_out$coefficients[2,2],3)`, Z = `r round(franchak_faces_out$coefficients[2,3],3)`, P = `r round(franchak_faces_out$coefficients[2,4],4)`.
+Prone, b = `r round(franchak_faces_out$coefficients[2,1],3)`, SE  = `r round(franchak_faces_out$coefficients[2,2],3)`, Z = `r round(franchak_faces_out$coefficients[2,3],3)`, `r reportP(franchak_faces_out$coefficients[2,4])`.
 Upright vs.
-Prone, b=`r round(franchak_faces_out$coefficients[3,1],3)`, SE  = `r round(franchak_faces_out$coefficients[3,2],3)`, Z = `r round(franchak_faces_out$coefficients[3,3],3)`, P = `r round(franchak_faces_out$coefficients[3,3],4)`).
+Prone, b = `r round(franchak_faces_out$coefficients[3,1],3)`, SE  = `r round(franchak_faces_out$coefficients[3,2],3)`, Z = `r round(franchak_faces_out$coefficients[3,3],3)`, `r reportP(franchak_faces_out$coefficients[3,4])`).
 
@@
 Across the entire session in @franchak2017see, infants looked at faces on 4.7% of frames [@franchak2017see].
-When we used all detections from OpenPose, we found a higher proportion of faces:  M = `r mean(detections_by_conf_by_sub$avg_faces_by_sub)`, or `r mean(detections_by_conf_by_sub$avg_faces_by_sub)*100`%.
-When we restricted our results to only high-confidence detections, we found M = `r mean(detections_by_conf_by_sub$avg_faces_by_sub_strict)`, closer to the original values reported by @franchak2017see.
-However, the above analyses on the accuracy of these two methods suggest that high-confidence detections dramatically underestimate the number of faces in view.
-Thus, we suspect that these differences are due to both (1) the fact head-mounted eye-trackers may sometimes underestimate the proportion of faces attended due to calibration issues, and that (2) OpenPose may sometimes detect faces that are in view for infants but that infants may not be foveating.
-
-We also found that infants' in-the-moment posture also modulated the proportion of hands that were in view (i.e., wrist detections), though these were not originally annotated in @franchak2017see (see Figure \ref{fig:franchak}, (proportion of frames with wrist detections; Prone: M =`r mean(dets_by_conf_by_posture$avg_hands[1])`; Sitting: M =`r mean(dets_by_conf_by_posture$avg_hands[2])`; Upright:  M =`r mean(dets_by_conf_by_posture$avg_hands[3])`).
+When we used all detections from OpenPose, we found a higher proportion of faces --  `r mean(detections_by_conf_by_sub$avg_faces_by_sub)*100`%.
+When we restricted our results to only high-confidence detections, we found `r mean(detections_by_conf_by_sub$avg_faces_by_sub_strict)*100`% of frames with faces, closer to the original values reported by @franchak2017see.
+However, the above analyses on the accuracy of this method suggest that high-confidence detections dramatically underestimate the number of faces in view.
+Thus, we suspect that these differences between our findings are due to both (1) the fact head-mounted eye-trackers may sometimes underestimate the proportion of faces attended due to calibration issues, and that (2) OpenPose may sometimes detect faces that are in view for infants but that infants may not be foveating.
+
+We also found that infants' in-the-moment posture also modulated the proportion of hands that were in view (i.e., wrist detections), though these were not originally annotated in @franchak2017see (see Figure \ref{fig:franchak}, proportion of frames with wrist detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_hands[1])`; Sitting: M = `r mean(dets_by_conf_by_posture$avg_hands[2])`; Upright:  M = `r mean(dets_by_conf_by_posture$avg_hands[3])`).
 These results were confirmed in generalized linear-mixed models, as in the first study (Sitting vs.
-Prone, b=`r round(franchak_wrists_out$coefficients[2,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[2,2],3)`, Z = `r round(franchak_wrists_out$coefficients[2,3],4)`, P = `r round(franchak_wrists_out$coefficients[2,4],4)`.
+Prone, b = `r round(franchak_wrists_out$coefficients[2,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[2,2],3)`, Z = `r round(franchak_wrists_out$coefficients[2,3],4)`, `r reportP(franchak_wrists_out$coefficients[2,4])`.
 Upright vs.
-prone, b=`r round(franchak_wrists_out$coefficients[3,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[3,2],3)`, Z = `r round(franchak_wrists_out$coefficients[3,3],3)`, P = `r round(franchak_wrists_out$coefficients[3,3],4)`).
+prone, b = `r round(franchak_wrists_out$coefficients[3,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[3,2],3)`, Z = `r round(franchak_wrists_out$coefficients[3,3],3)`, `r reportP(franchak_wrists_out$coefficients[3,4])`).
 
@@
 # General Discussion
-What social cues do infants see as they learn language, and how does this change as they grow and start to locomote themselves in their environments?
-We explored this question using video data from head-mounted cameras from two naturalistic datasets of parent-child interactions: a cross-sectional database of naturalistic play sessions from 8-16 month-olds with sets of novel and familiar toys, and database of head-mounted camera videos from one-year-olds who explored a large play area from @franchak2017see.
+What social cues do infants see as they learn language, and how does this change as they grow and start to locomote themselves?
+We explored this question using video data from head-mounted cameras from two naturalistic datasets of parent-child interactions: a cross-sectional database of play sessions from 8-16 month-olds with sets of novel and familiar toys, and database of head-mounted camera videos from one-year-olds who explored a large play area from @franchak2017see.
 To analyze these datasets, we developed a novel method using a pose detection model to automate the annotation of the social information in the infant view, here operationalized as the presence of the faces and hands of their caregiver; these annotations were then synced with manual annotations of infants in-the-moment posture from third-person videos.
 
-Despite being trained on the adult perspective, the pose detector we used (OpenPose, @cao2017realtime) was able to generalize relatively well to the infant viewpoint, achieving comparable precision and accuracy as a face detector relative to a state-of-the-art model optimized specifically for detecting faces in natural scenes (MTCNN, @zhang2016].
+Despite being trained on the adult perspective, the pose detector we used (OpenPose, @cao2017realtime) was able to generalize relatively well to the infant viewpoint, achieving comparable precision and accuracy as a face detector relative to a state-of-the-art model optimized specifically for detecting faces in natural scenes (MTCNN, @zhang2016).
 While OpenPose had relatively low recall as a hand detector -- missing some hands that were in the infant view -- it made comparable rates of false alarms.
@@
 Motor development changes dramatically at the same time that children are breaking into language learning.
-Using these automated detections, we found that infants' changing posture and orientation to their caregiver jointly shaped the amount of social information that was in their view during one-on-one play sessions with their caregivers: Children saw the most faces/hands when they were sitting or standing and close to their caregiver vs.
+Using these automated detections, we found that infants' changing posture and orientation to their caregiver jointly shaped the amount of social information that was in their view during one-on-one play sessions with their caregivers.
+Children saw the most faces/hands when they were sitting or standing and close to their caregiver vs.
 crawling or prone.
-These same findings were recapitulated in a second dataset collected by @franchak2017see with one-year-olds: sitting and upright infants saw more faces--and hands--than infants who were prone.
+These same findings were recapitulated in a second dataset collected by @franchak2017see with one-year-olds: sitting and upright infants saw more faces -- and hands -- than infants who were prone.
 Motor development appears to modulate how infants experience their visual world and the social information in it.
 
-While exploratory, our results also suggest that infants saw a greater proportion of hands during naming events, hinting that children may have been orienting towards their caregiver when they heard labels for objects.
+While exploratory, our results also suggest that infants saw a greater proportion of hands around naming events, hinting that children may have been orienting towards their caregiver when they heard labels for objects.
 While this effect was not present for faces, other work [@yoshida2008; @yu2013joint; @yu2017hand], including @franchak2017see, has found that infants spend much more time looking at the toys vs.
 their caregiver's faces during these play sessions, and highlighted the importance of hand-following as a component of joint attention [@yu2017hand].
-Further, given that there were only two possible referents in the room at a time—-and one of them was always a familiar category (i.e., car, kitty) -- this particular play session did not present many opportunities where children would need to use gaze cues to disambiguate referents.
-
-Importantly, however, all of these findings come from observational, in-lab datasets, posing limits on their generalizability.
-Future work is needed to relate these the slices of experience captured within in-lab play sessions and infants' everyday experiences [@fausey2016; @clerkin2017].
+Further, given that there were only two possible referents in the room at a time -- and one of them was always a familiar category -- this particular play session did not present many opportunities where children would need to use gaze cues to disambiguate referents.
+
+Importantly, however, all of these findings come from observational, in-lab datasets, posing important limits on their generalizability.
+Future work is needed to relate these  slices of experience captured within in-lab play sessions with infants' everyday experiences [@fausey2016; @clerkin2017].
 More broadly, though observational findings allow us to document developmental change and identify potential causal pathways, they cannot confirm them.
-Locomotive abilities are of course only part of a cascade of changes in infants' abilities and experiences, and these analyzes document only a fraction of this broader, multifaceted trajectory.
+Locomotive abilities are of course only part of a cascade of changes in infants' abilities and experiences, and these analyses document only a fraction of this broader, multifaceted trajectory.
 As children grow and change, the activities in which they engage with their caregivers are likely to also vary, leading to differences in the distribution of social cues that they experience that may not be captured in these in-lab play sessions.
 
-Nonetheless, we think that this approach hold promise for documenting these developmental trajectories and for generating hypotheses about potential causal pathways.
-Indeed, understanding the relationship between different domains of developmental changes in naturalistic contexts has been a persistent challenge for developmental psychology.
-However, the field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools that deal better with noisier, more complicated datasets and extract richer information.
+Understanding the relationship between different domains of developmental changes in naturalistic contexts has been a persistent challenge for developmental psychology.
+Nonetheless, we think that this approach hold promise for documenting these developmental trajectories and for generating new hypotheses.The field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools that deal better with noisier, more complicated datasets and extract richer information.
 We hope that these new tools can now be leveraged to examine the consequences of the changing infant perspective for linguistic, cognitive, and social development.
@@
 # Acknowledgements
-
 Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation.

### 2020-08-12 writeup-paper/draft_apa_july2020.tex
latest version of ms

@@
 \makeatother
-\shorttitle{Automated detection of the social information}
+\shorttitle{Automated detection of social information}
 \author{Bria L. Long\textsuperscript{1}, Alessandro Sanchez\textsuperscript{1}, Allison M. Kraus\textsuperscript{1}, Ketan Agrawal\textsuperscript{1}, \& Michael C. Frank\textsuperscript{1}}
@@
 egocentric video dataset where one-year-olds explored a large play room
-with their caregivers (Franchak et al., 2017), and found convergence
-across both datasets, confirming the finding that motoric developments
-play a significant role in shaping the social information that infants
-have access to over development, and suggesting that they may play a
-role in emergence of infants' cognitive and linguistic abiltiies. We
-propose that the combined use of head-mounted cameras and the
-application of new computer vision techniques is a promising avenue for
-understanding the statistics of infants' visual and linguistic
-experience as they change over development and for re-utilizing rich
-video datasets collected across different labs and diverse contexts.
+with their caregivers (Franchak et al., 2017), replicating and extending
+their findings. Together, these results establish the validity of a new
+method for detecting social information, confirm that motoric
+developments play a significant role in shaping the social information
+that infants have access to. We propose that the combined use of
+head-mounted cameras and the application of new computer vision
+techniques is a promising avenue for re-utilizing rich video datasets
+collected across different labs and diverse contexts and understanding
+the statistics of infants' visual and linguistic experience as they
+change over development.
 }
@@
 and caregivers played freely using a set of novel and familiar objects;
-8, 12, and 16 months-olds (N=26; N=12 in each age group) participated to
+8, 12, and 16 months-olds (N=36; N=12 in each age group) participated to
 ensure a wide range of locomotor abilities. Infants' in-the-moment
@@
 minutes of data in each child (seconds excluded from analysis for
-posture;. M =94.01, SD =225.18; orientation; M =94.01, SD =225.18), and
-these rates did not vary substantially with the age of the child. To
-assess the reliability of these annotations, a second coder coded videos
-from three different children to calculate Cohen's kappa (posture, irr =
-XX, orientation, irr = XX).
+posture, M = 94.01s, SD = 225.18s; orientation; M = 94.01s, SD =
+225.18s), and these rates did not vary substantially with the age of the
+child. To assess the reliability of these annotations, a second coder
+coded videos from three different children to calculate Cohen's kappa
+(posture, irr = XX, orientation, irr = XX).
 
@@
 well as a hand detector (F = 0.74). OpenPose achieved relatively high
-precision -- generating relatively few false positives -- but showedlow
+precision -- generating relatively few false positives -- but showed low
 recall on the randomly sampled frames (see Table 1). As with face
@@
 
-\subsection{Access to social information during postural
-developments}\label{access-to-social-information-during-postural-developments}
+\subsection{Study 1: Access to social information during postural
+developments}\label{study-1-access-to-social-information-during-postural-developments}
 
@@
 sitting or standing had more faces in view (Face detections; infant
-sitting and CG close, b=0.99, SE = 0.07, Z = 14.54, P = 0; infant
-standing and CG close, b=1.36, SE = 0.08, Z = 16.95, P = 0 than infants
-who were lying down/crawling (i.e.~prone). When caregivers were far away
-from their infants, face detections were similarity higher (Face
-detections; infant sitting and CG far, b = 0.75, SE = 0.07, Z = 10.48, P
-= 0; infant standing and CG far, b = 1.40, SE = 0.09, Z = 16.16, P = 0.
-Overall, we found that age did not remain a significant predictor in
-accounting for the faces in view when modeling infants posture, their
+sitting and CG close, b = 0.99, SE = 0.07, Z = 14.54, P \textless{}
+0.001; infant standing and CG close, b = 1.36, SE = 0.08, Z = 16.95, P
+\textless{} 0.001) than infants who were lying down/crawling
+(i.e.~prone). When caregivers were far away from their infants, face
+detections were similarity higher (Face detections; infant sitting and
+CG far, b = 0.75, SE = 0.07, Z = 10.48, P = P \textless{} 0.001; infant
+standing and CG far, b = 1.40, SE = 0.09, Z = 16.16, P = P \textless{}
+0.001. Overall, we found that age did not remain a significant predictor
+in accounting for the faces in view when modeling infants posture, their
 orientation to their caregiver, and the interaction between these two
 factors (Face detections; Age (scaled), b = 0.05, SE = 0.12, Z = 0.39, P
-= 0.70).
+= P = 0.697).
 
@@
 relatively closer to their infants (Hand detections; infant sitting and
-CG close, b=0.31, SE = 0.08, Z = 3.72, P = 0.00; infant standing and CG
-close, b=0.76, SE = 0.09, Z = 8.46, P = 0 than infants who were lying
-down/crawling (i.e.~prone). Wrist detections were clearly highest when
-caregivers were far away from their infants (Wrist detections; infant
-sitting and CG far, b=0.13, SE = 0.10, Z = 1.27, P = 0.21; infant
-standing and CG far, b=1.76, SE = 0.11, Z = 15.87, P = 0. As with faces,
-age did not remain a significant predictor in these models (Wrist
-detections; Age (scaled), b = 0.16, SE = 0.11, Z = 1.48, P = 0.14).
+CG close, b=0.31, SE = 0.08, Z = 3.72, P \textless{} 0.001; infant
+standing and CG close, b=0.76, SE = 0.09, Z = 8.46, P \textless{} 0.001)
+than infants who were lying down/crawling (i.e.~prone). Wrist detections
+were highest when caregivers were far away from their infants and those
+infants were standing (Wrist detections;infant standing and CG far, b =
+1.76, SE = 0.11, Z = 15.87, P \textless{} 0.001). As with faces, age did
+not remain a significant predictor in these models (Wrist detections;
+Age (scaled), b = 0.16, SE = 0.11, Z = 1.48, P = 0.140).
 
@@
 marginal \(R^2\) were higher for both faces (marginal \(R^2\) = 0.28)
-and wrists (marginal \(R^2\) = 0.31).\\
-Overall, these results suggest that infants' visual access to social
-information is largely modulated by their posture and orientation to
-their caregiver, which is in turn a function of their general locomotor
-development.
+and wrists (marginal \(R^2\) = 0.31). Overall, these results suggest
+that infants' visual access to social information is largely modulated
+by their posture and orientation to their caregiver, which is in turn a
+function of their general locomotor development.
 
@@
 naming events relative to baseline, with similar effects across age
-groups (8 m.o., \(M_{fam - baseline}\) = =0.02, 12 m.o.,
-\(M_{nov - baseline}\) =0.05, 16 m.o. \(M_{nov - baseline}\) =0.02)
+groups (8 m.o., \(M_{fam - baseline}\) = 0.02, 12 m.o.,
+\(M_{nov - baseline}\) = 0.05, 16 m.o. \(M_{nov - baseline}\) = 0.02)
 however, they were not higher during familiar naming events vs.~baseline
-(8 m.o., \(M_{fam - baseline}\) =0.00, 12 m.o., \(M_{fam - baseline}\)
-=0.00, 16 m.o. \(M_{fam - baseline}\) =0.02, Conversely, wrist
+(8 m.o., \(M_{fam - baseline}\) =0.00, 12 m.o., \(M_{fam - baseline}\) =
+0.00, 16 m.o. \(M_{fam - baseline}\) = 0.02). Conversely, wrist
 detections were higher during both familiar (8 m.o.,
-\(M_{fam - baseline}\) =0.07, 12 m.o., \(M_{fam - baseline}\) =0.04, 16
-m.o. \(M_{fam - baseline}\) =0.08) and novel naming events relative to
-baseline across all age groups (Novel; 8 m.o., (8 m.o.,
-\(M_{fam - baseline}\) =0.09, 12 m.o., \(M_{fam - baseline}\) =0.08, 16
-m.o. \(M_{fam - baseline}\) =0.08). These results were confirmed by a
-linear mixed-effect model with scaled aged as a fixed effect and random
-intercepts for each subject (Wrist detections; familiar objects
-vs.~baseline, b=0.06, SE = 0.01, t = 4.33, P = 0.00'; Novel objects
-vs.~baseline, b=0.08, SE = 0.02, t = 5.21, P = 0.
+\(M_{fam - baseline}\) =0.07, 12 m.o., \(M_{fam - baseline}\) = 0.04, 16
+m.o. \(M_{fam - baseline}\) = 0.08) and novel naming events relative to
+baseline across all age groups (8 m.o., \(M_{nov - baseline}\) = 0.09,
+12 m.o., \(M_{nov - baseline}\) = 0.08, 16 m.o. \(M_{nov - baseline}\) =
+0.08). These results were confirmed by a linear mixed-effect model with
+scaled aged as a fixed effect and random intercepts for each subject
+(Wrist detections; familiar objects vs.~baseline, b = 0.06, SE = 0.01, t
+= 4.33, P \textless{} 0.001; Novel objects vs.~baseline, b = 0.08, SE =
+0.02, t = 5.21, P \textless{} 0.001).
 
@@
 infant when presenting them with objects (Gogate, Bahrick, \& Watson,
-2000, Gogate, Bolzani, and Betancourt (2006)). For example, caregivers
-may tend to simultaneously name objects when demonstrating their
-affordances or simply when pointing to them. In turn, infants may be
-sensitive to these naming events and orient their attention towards
-their caregiver, consistent with other accounts positing infants'
-sensitivity to social cues in early word learning environments
-(Yurovsky, 2018, Yurovsky and Frank (2017), Frank, Simmons, Yurovsky,
-and Pusiol (2013)).
+2000; Gogate, Bolzani, \& Betancourt, 2006). For example, caregivers may
+tend to simultaneously name objects when demonstrating their affordances
+or simply when pointing to them. In turn, infants may be sensitive to
+these naming events and orient their attention towards their caregiver,
+consistent with other accounts positing infants' sensitivity to social
+cues in early word learning environments (Frank, Simmons, Yurovsky, \&
+Pusiol, 2013; Yurovsky, 2018; Yurovsky \& Frank, 2017).
 
@@
 
-\subsection{Extension to Franchak et a.,
-2017}\label{extension-to-franchak-et-a.-2017}
+\subsection{Study 2: Extension of Franchak et al.,
+2017}\label{study-2-extension-of-franchak-et-al.-2017}
 
@@
 findings using our automated methodology (OpenPose detections), using
-the footage from their head-mounted cameras (D. A. Simon, Gordon,
-Steiger, \& Gilmore, 2015) for two reasons. First, we sought to validate
-our novel method, which could fail to generalize to scenes from these
-more complex environments, where detecting faces and hands could
-arguably be a much harder task. Second, we sought to replicate the
-effects of infants' in-the-moment posture on differences in visual
-access to hands in an independent dataset.
+the footage from their head-mounted cameras (hosted on Databrary; D. A.
+Simon, Gordon, Steiger, \& Gilmore, 2015) for two reasons. First, we
+sought to validate our novel method, which could fail to generalize to
+scenes from these more complex environments, where detecting faces and
+hands could arguably be a much harder task. Second, we sought to
+replicate the effects of infants' in-the-moment posture on differences
+in visual access to hands in an independent dataset.
 
@@
 First, despite having different head-mounted camera setups, the view
-angle of the two head-mounted cameras were only slightly different
-(\(52.2^{\circ}\) horizontal by \(42.2^{\circ}\) in Franchak et al.
-(2017), here \(64^{\circ}\) horizontal by \(46^{\circ}\) vertical).
-However, the environment that infants were immersed in with their
+angle of the two head-mounted cameras used in Study 1 and Franchak et
+al. (2017) were only slightly different (\(52.2^{\circ}\) horizontal by
+\(42.2^{\circ}\) in Franchak et al. (2017), here \(64^{\circ}\)
+horizontal by \(46^{\circ}\) vertical). However, the play environment in
+Franchak et al. (2017) that infants were immersed in with their
 caregivers (and experimenters) was much larger and more varied than the
@@
 all detections (see Figure \ref{fig:franchak}, panel A; (proportion of
-frames with face detections; Prone: M =0.13\%, Sitting: M =0.20\%,
-Upright: M =0.22\%) or restricted our analyses to high-confidence
-detections (percentage of frames with face detections; Prone: M =0.03,
-Sitting: M =0.06, Upright: M =0.05). These results were confirmed in
+frames with face detections; Prone: M = 0.13, Sitting: M = 0.20,
+Upright: M = 0.22) or restricted our analyses to high-confidence
+detections (percentage of frames with face detections; Prone: M = 0.03,
+Sitting: M = 0.06, Upright: M = 0.05). These results were confirmed in
 generalized linear-mixed models, as in the first study (Sitting
-vs.~Prone, b=0.34, SE = 0.02, Z = 23.37, P = 0. Upright vs.~Prone,
-b=0.38, SE = 0.02, Z = 24.01, P = 24.01).
+vs.~Prone, b = 0.34, SE = 0.02, Z = 23.37, P \textless{} 0.001. Upright
+vs.~Prone, b = 0.38, SE = 0.02, Z = 24.01, P \textless{} 0.001).
 
@@
 2017). When we used all detections from OpenPose, we found a higher
-proportion of faces: M = 0.22, or 21.80\%. When we restricted our
-results to only high-confidence detections, we found M = 0.06, closer to
-the original values reported by Franchak et al. (2017). However, the
-above analyses on the accuracy of these two methods suggest that
+proportion of faces -- 21.80\%. When we restricted our results to only
+high-confidence detections, we found 6.11\% of frames with faces, closer
+to the original values reported by Franchak et al. (2017). However, the
+above analyses on the accuracy of this method suggest that
 high-confidence detections dramatically underestimate the number of
-faces in view. Thus, we suspect that these differences are due to both
-(1) the fact head-mounted eye-trackers may sometimes underestimate the
-proportion of faces attended due to calibration issues, and that (2)
-OpenPose may sometimes detect faces that are in view for infants but
-that infants may not be foveating.
+faces in view. Thus, we suspect that these differences between our
+findings are due to both (1) the fact head-mounted eye-trackers may
+sometimes underestimate the proportion of faces attended due to
+calibration issues, and that (2) OpenPose may sometimes detect faces
+that are in view for infants but that infants may not be foveating.
 
@@
 these were not originally annotated in Franchak et al. (2017) (see
-Figure \ref{fig:franchak}, (proportion of frames with wrist detections;
-Prone: M =0.15; Sitting: M =0.21; Upright: M =0.22). These results were
-confirmed in generalized linear-mixed models, as in the first study
-(Sitting vs.~Prone, b=0.27, SE = 0.01, Z = 19.09, P = 0. Upright
-vs.~prone, b=0.30, SE = 0.02, Z = 19.73, P = 19.73).
+Figure \ref{fig:franchak}, proportion of frames with wrist detections;
+Prone: M = 0.15; Sitting: M = 0.21; Upright: M = 0.22). These results
+were confirmed in generalized linear-mixed models, as in the first study
+(Sitting vs.~Prone, b = 0.27, SE = 0.01, Z = 19.09, P \textless{} 0.001.
+Upright vs.~prone, b = 0.30, SE = 0.02, Z = 19.73, P \textless{} 0.001).
 
@@
 What social cues do infants see as they learn language, and how does
-this change as they grow and start to locomote themselves in their
-environments? We explored this question using video data from
-head-mounted cameras from two naturalistic datasets of parent-child
-interactions: a cross-sectional database of naturalistic play sessions
-from 8-16 month-olds with sets of novel and familiar toys, and database
-of head-mounted camera videos from one-year-olds who explored a large
-play area from Franchak et al. (2017). To analyze these datasets, we
-developed a novel method using a pose detection model to automate the
-annotation of the social information in the infant view, here
-operationalized as the presence of the faces and hands of their
-caregiver; these annotations were then synced with manual annotations of
-infants in-the-moment posture from third-person videos.
+this change as they grow and start to locomote themselves? We explored
+this question using video data from head-mounted cameras from two
+naturalistic datasets of parent-child interactions: a cross-sectional
+database of play sessions from 8-16 month-olds with sets of novel and
+familiar toys, and database of head-mounted camera videos from
+one-year-olds who explored a large play area from Franchak et al.
+(2017). To analyze these datasets, we developed a novel method using a
+pose detection model to automate the annotation of the social
+information in the infant view, here operationalized as the presence of
+the faces and hands of their caregiver; these annotations were then
+synced with manual annotations of infants in-the-moment posture from
+third-person videos.
 
@@
 optimized specifically for detecting faces in natural scenes (MTCNN, K.
-Zhang et al. (2016){]}. While OpenPose had relatively low recall as a
-hand detector -- missing some hands that were in the infant view -- it
-made comparable rates of false alarms. In both cases, we found that
-overall performance was maximized when all detections were included,
-regardless of their confidence, suggesting that some low-confidence face
-and hand detection still index actual faces and hands that were seen by
-infants. Thus, while imperfect, we suggest that OpenPose can be applied
-to infant egocentric videos for the extraction of the social information
-in the infant viewpoint, reducing the burden of manual annotations and
+Zhang et al. (2016)). While OpenPose had relatively low recall as a hand
+detector -- missing some hands that were in the infant view -- it made
+comparable rates of false alarms. In both cases, we found that overall
+performance was maximized when all detections were included, regardless
+of their confidence, suggesting that some low-confidence face and hand
+detection still index actual faces and hands that were seen by infants.
+Thus, while imperfect, we suggest that OpenPose can be applied to infant
+egocentric videos for the extraction of the social information in the
+infant viewpoint, reducing the burden of manual annotations and
 promoting the reusability of rich video datasets for further analyses.
@@
 their caregiver jointly shaped the amount of social information that was
-in their view during one-on-one play sessions with their caregivers:
+in their view during one-on-one play sessions with their caregivers.
 Children saw the most faces/hands when they were sitting or standing and
@@
 recapitulated in a second dataset collected by Franchak et al. (2017)
-with one-year-olds: sitting and upright infants saw more faces--and
-hands--than infants who were prone. Motor development appears to
+with one-year-olds: sitting and upright infants saw more faces -- and
+hands -- than infants who were prone. Motor development appears to
 modulate how infants experience their visual world and the social
@@
 While exploratory, our results also suggest that infants saw a greater
-proportion of hands during naming events, hinting that children may have
+proportion of hands around naming events, hinting that children may have
 been orienting towards their caregiver when they heard labels for
@@
 attention (Yu \& Smith, 2017). Further, given that there were only two
-possible referents in the room at a time----and one of them was always a
-familiar category (i.e., car, kitty) -- this particular play session did
-not present many opportunities where children would need to use gaze
-cues to disambiguate referents.
+possible referents in the room at a time -- and one of them was always a
+familiar category -- this particular play session did not present many
+opportunities where children would need to use gaze cues to disambiguate
+referents.
 
 Importantly, however, all of these findings come from observational,
-in-lab datasets, posing limits on their generalizability. Future work is
-needed to relate these the slices of experience captured within in-lab
-play sessions and infants' everyday experiences (Clerkin, Hart, Rehg,
-Yu, \& Smith, 2017; Fausey et al., 2016). More broadly, though
-observational findings allow us to document developmental change and
-identify potential causal pathways, they cannot confirm them. Locomotive
-abilities are of course only part of a cascade of changes in infants'
-abilities and experiences, and these analyzes document only a fraction
-of this broader, multifaceted trajectory. As children grow and change,
-the activities in which they engage with their caregivers are likely to
-also vary, leading to differences in the distribution of social cues
-that they experience that may not be captured in these in-lab play
-sessions.
-
-Nonetheless, we think that this approach hold promise for documenting
-these developmental trajectories and for generating hypotheses about
-potential causal pathways. Indeed, understanding the relationship
-between different domains of developmental changes in naturalistic
-contexts has been a persistent challenge for developmental psychology.
-However, the field of computer vision has advanced dramatically in
-recent years, creating a new generation of algorithmic tools that deal
-better with noisier, more complicated datasets and extract richer
-information. We hope that these new tools can now be leveraged to
-examine the consequences of the changing infant perspective for
-linguistic, cognitive, and social development.
+in-lab datasets, posing important limits on their generalizability.
+Future work is needed to relate these slices of experience captured
+within in-lab play sessions with infants' everyday experiences (Clerkin,
+Hart, Rehg, Yu, \& Smith, 2017; Fausey et al., 2016). More broadly,
+though observational findings allow us to document developmental change
+and identify potential causal pathways, they cannot confirm them.
+Locomotive abilities are of course only part of a cascade of changes in
+infants' abilities and experiences, and these analyses document only a
+fraction of this broader, multifaceted trajectory. As children grow and
+change, the activities in which they engage with their caregivers are
+likely to also vary, leading to differences in the distribution of
+social cues that they experience that may not be captured in these
+in-lab play sessions.
+
+Understanding the relationship between different domains of
+developmental changes in naturalistic contexts has been a persistent
+challenge for developmental psychology. Nonetheless, we think that this
+approach hold promise for documenting these developmental trajectories
+and for generating new hypotheses.The field of computer vision has
+advanced dramatically in recent years, creating a new generation of
+algorithmic tools that deal better with noisier, more complicated
+datasets and extract richer information. We hope that these new tools
+can now be leveraged to examine the consequences of the changing infant
+perspective for linguistic, cognitive, and social development.
 

### 2020-08-14 analysis/1_merge_data/merge_all_data.Rmd
add cropping by fisheye to preprocessing, redo paths

@@
 ## Step 1: Load detectors
+
+### Load demographics
+```{r}
+# All of the videos were filmed with an MD-80 headcam.
+Some videos had a fisheye lens attached to them that increased their FOV by approximately 50% in both directions (see manuscript; in demographics column "fisheye").
+This was determined by looking at the third-person videos.
+
+# The original resolutoin fo ALL of the videos is 480 x 720.
+But, then, OpenPose resizes them (not sure why?) and adds some blank space so that it is the right space -- this blank space is on the RIGHT.
+The final resolution becomes 1280 x 720; all vertical space has meaningful detections, but all horizontal space does not
+
+# In order to make the detections comparable, we need to restrict detections to center FOV for fisheye participants.
+### Specify cropping for participants with fish-eye lens
+
+## pop out fisheye versus none
+demo <- read.csv(paste0(here::here(),"/data/demographics/demographics_updated.csv"))
+
+fisheye_subs <- demo %>%
+  filter(Fisheye=='yes')
+
+no_fisheye_subs <- demo %>%
+  filter(Fisheye=='no')
+
+# based on difference in view angle between veho/muvi-80
+vertical_diff = (1 - (47/64))/2
+horizontal_diff = (1 - (36/46))/2
+
+# vid params
+# x_max = 1280 # actual max -- black space between 1080 and 1280
+x_max = 1080 # max with content
+y_max = 720
+
+# cropping params -- c
+limit_lower_y = vertical_diff*y_max
+limit_upper_y = (1 - vertical_diff)*y_max
+
+limit_lower_x = horizontal_diff*x_max
+limit_upper_x = (1 - horizontal_diff)*x_max
+
+```
+
 ### Load detectors from MTCNN
@@
 
-### Load OpenPose detections (face and wrists)
+### Load & Preprocess OpenPose detections (face and wrists)
 Wrists: if right or left wrist was ever detected
@@
 ```{r}
-detsOpenPose <- read_csv("../../data/final_detector_output/openpose_results_truncated_2.csv") 
-detsOpenPose <- detsOpenPose %>%
-  mutate(video = name) %>%
-  distinct(video,frame,.keep_all=TRUE)  %>%
-  mutate(frame = as.numeric(frame) + 1) %>% ## Frame output is off by one since we extracted from videos vs. frames.
+detsOpenPoseRaw <- read_csv(paste0(here::here(),"/data/final_detector_output/openpose_results_truncated_2.csv")) %>%
+  mutate(subid = name, video = name) %>%
+  distinct(subid,frame,.keep_all=TRUE) %>%
+  mutate(frame = as.numeric(frame) + 1)  ## Frame output is off by one since we extracted from videos vs. frames.
+```
+
+## break out detetions for fisheye vs non
+```{r}
+# dets for subs w/o fisheye lens 
+detsOpenPose_no_fisheye <- no_fisheye_subs %>%
+  left_join(detsOpenPoseRaw) %>%
   mutate(faceOP = Nose_conf!=0)  %>%  
@@
   mutate(faceOPStrict = Nose_conf>.5)  %>%  
-  mutate(wristOPStrict = LWrist_conf>.5 | RWrist_conf>.5 )   
+  mutate(wristOPStrict = LWrist_conf>.5 | RWrist_conf>.5 )   %>%
+  mutate(faceOPAll = faceOP,wristOPAll = wristOP, faceOPStrictAll = faceOPStrict, wristOPStrictAll = wristOPStrict)
+
+# dets for subs w fisheye lens 
+detsOpenPose_fisheye <- fisheye_subs %>%
+  left_join(detsOpenPoseRaw) %>%
+  mutate(faceOPAll = Nose_conf!=0)  %>%  
+  mutate(wristOPAll = LWrist_conf!=0 | RWrist_conf!=0 )   %>%
+  mutate(faceOPStrictAll = Nose_conf>.5)  %>%  
+  mutate(wristOPStrictAll = LWrist_conf>.5 | RWrist_conf>.5 ) %>%
+  mutate(noseOP_center = (Nose_y<(limit_upper_y) & Nose_x<(limit_upper_x) & Nose_y>(limit_lower_y) & Nose_x>(limit_lower_x))) %>%
+  mutate(wristOPLeft_center = (LWrist_y<(limit_upper_y) & LWrist_x<(limit_upper_x) & LWrist_y>(limit_lower_y) & LWrist_x>(limit_lower_x)))  %>%
+  mutate(wristOPRight_center = (RWrist_y<(limit_upper_y) & RWrist_x<(limit_upper_x) & RWrist_y>(limit_lower_y) & RWrist_x>(limit_lower_x)))  %>%
+  mutate(faceOP = Nose_conf!=0 & noseOP_center) %>%  
+  mutate(wristOP = (LWrist_conf!=0 & wristOPLeft_center) | (wristOPRight_center & RWrist_conf!=0 )) %>%
+  mutate(faceOPStrict = Nose_conf>.5 & noseOP_center) %>%  
+  mutate(wristOPStrict = (LWrist_conf>.5 & wristOPLeft_center) | (wristOPRight_center & RWrist_conf>.5 )) %>%
+  select(-noseOP_center, -wristOPLeft_center, -wristOPRight_center)
+
+detsOpenPose <- detsOpenPose_no_fisheye %>%
+  full_join(detsOpenPose_fisheye)
 ```
@@
 alldets=left_join(dets_MTCNN,detsViola[,c("video","frame","faceVJ")])  ## Join MTCNN + Viola Jones
-alldets=left_join(alldets,detsOpenPose[,c("video","frame","faceOP","wristOP", "faceOPStrict", "wristOPStrict")])  ##  ++ OpenPose
+
+alldets=left_join(alldets,detsOpenPose[,c("video","frame","faceOP","wristOP", "faceOPStrict", "wristOPStrict", "faceOPAll","wristOPAll", "faceOPStrictAll", "wristOPStrictAll")])  ##  ++ OpenPose
 ```
@@
   group_by(subid) %>%
-  do(add.posture(.)) %>% ## adds both postures and orientatinos
-  do(add.posture.second(.)) ## gets reliability coding for postures
+  do(add.posture(.)) 
 ```
@@
 ```{r}
-write_csv(d, "../data/consolidated_data/consolidated_data_4detectors_july2020.csv") #
+write_csv(d, paste0(here::here(), "/data/consolidated_data/consolidated_data_4detectors_aug2020.csv")) 
 ```

### 2020-08-14 analysis/1_merge_data/helper.R
redo path typo

@@ -218,7 +218,7 @@ get.namings <- function (x, window = c(-2,2), file_dir = paste0(here::here(),'/d
   namings <- regularize.naming(namings)
 }
 
-get.namings.second <- function (x, window = c(-2,2), file_dir = paste0(here::here(),'/data/naming/second')) {
+get.namings.second <- function (x, window = c(-2,2), file_dir = paste0(here::here(),'/data/naming/second/')) {
   # try for second file
   second_file = Sys.glob(paste(file_dir,x$subid[1], '*',".csv", sep=""))
   naming_second = read.csv(second_file)

### 2020-08-14 analysis/2_coding_reliability/reliability_naming_posture.Rmd
Update reliability_naming_posture.Rmd

@@
 library(assertthat)
-source("helper.R") # Contains critical helper functions for merging annotations
+library(here)
+library(ggplot2)
+# Contains critical helper functions for merging annotations
+source(paste0(here::here(), "/analysis/1_merge_data/helper.R")) 
 ```
@@
 # frame by video structure
-load(file = '../data/consolidated_data/video_by_frame.RData')
+load(file = paste0(here::here(),'/data/consolidated_data/video_by_frame.RData'))
 ```
 
-## Naming reliabilty
+## Load naming data for first and second coders
 ```{r}
 # paths to datafiles
-file_path = '../data/naming/second/'
+file_path = paste0(here::here(), '/data/naming/second/')
 second_codes = dir(file_path) %>%
@@
   select(-naming.instance, -X) 
+```
 
+## Calculate namings that occur within the same range
+```{r}
 joined_namings_4s <- all_namings_first_coder %>%
@@
 
-## Posture reliability
+# Posture & oreintation coding reliability
+
+## Load data, frame times, coding 
+### Call functions sourced in helper.R
 ```{r}
-
 d <- all_dets_stripped %>%
@@
 
+### Create dataframe
 ```{r}
@@
 posture_eval = data.frame(posture_reliability$posture, posture_reliability$posture_second)
-irr::kappa2(posture_eval)
-
-orient_eval = data.frame(posture_reliability$orientation, posture_reliability$orientation_second)
-irr::kappa2(orient_eval)
-
 ```
 
+## Posture reliability plots and calculation
 ```{r}
@@
 ```{r}
-# posture_every_second <- d %>%
-#   mutate(time_seconds = round(time,0)) %>%
-#   group_by(time_seconds) %>%
-#   mutate(common_posture = names(sort(table(posture)))[5],
-#   common_posture_second = names(sort(table(posture_second)))[5])
-# 
-# posture_eval_second = data.frame(posture_every_second$common_posture, posture_every_second$common_posture_second)
-# irr::kappa2(posture_eval_second)
+posture_irr =irr::kappa2(posture_eval)
 ```
+Cohen's kappa for posture is `r round(posture_irr$value,2)`
+
+## Orientation reliability plots and calculation
+```{R}
+coder_1 <- ggplot(posture_reliability, aes(x=frame, y=orientation_second, fill=orientation_second)) +
+  geom_tile() +
+  facet_wrap(~subid) + 
+  theme(legend.position='none')
+
+coder_2 <- ggplot(posture_reliability, aes(x=frame, y=orientation, fill=orientation)) +
+  geom_tile() +
+  facet_wrap(~subid) + 
+  theme(legend.position='none')
+
+cowplot::plot_grid(coder_1,coder_2, nrow=2)
+```
+
+```{R}
+orient_eval = data.frame(posture_reliability$orientation, posture_reliability$orientation_second)
+orient_irr = irr::kappa2(orient_eval)
+```
+Cohen's kappa for orientation is `r round(orient_irr$value,2)`

### 2020-08-14 writeup-paper/draft_apa_july2020.Rmd
latest paper draft

@@
 ---
-title             : "Using automated detections to quantify the effect of postural developments on the social information in the infant view"
-
+title             : "Automated detections reveal the social information in the changing infant view"
 shorttitle        : "Automated detection of social information"
@@
 
-# authornote: |
-#   Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-# 
-#   Enter author note here.
+authornote: |
 
@@
 We also applied our automated method to a second egocentric video dataset where one-year-olds explored a large play room with their caregivers  (Franchak et al., 2017), replicating and extending their findings.
-Together, these results establish the validity of a new method for detecting social information, confirm that motoric developments play a significant role in shaping the social information that infants have access to.
-We propose that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for re-utilizing rich video datasets collected across different labs and diverse contexts and understanding the statistics of infants\' visual and linguistic experience as they change over development.
+Together, these results establish the validity of a new method for detecting social information and confirm that motoric developments play a significant role in shaping the social information that infants have access to.
+We propose that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for re-utilizing rich video datasets and understanding how the statistics of infants\' visual and linguistic experience change over development.
 
@@
 While head-mounted camera data do not provide explicit information on where infants are attending, some work suggests that infants orient their head towards what they are focusing on -- putting those people or objects in view [@yoshida2008].
-Initial recordings from in-lab play sessions revealed a different view than many imagined: instead of being dominated by their faces, the infant perspective contained close up views of primarily toys and hands [@yoshida2008; @franchak2011; @yu2017hand].
+Initial recordings using this technique during in-lab play sessions revealed a different view than many imagined: instead of being dominated by faces, the infant perspective contained close up views of primarily toys and hands [@yoshida2008; @franchak2011; @yu2017hand].
 Subsequent research has revealed that the infant view is far from a unitary construct, undergoing dramatic changes as infants grow.
@@
 
-These changes in perspective are likely the downstream consequence of a myriad factors, chief of which may be infants' evolving locomotive abilities: an infants ability to sit, crawl, stand, or walk structures the way they interact with the things and people in their world.
-These changes can be rapid: the physical body of an infant can lengthen up to 2cm in 24 hours [@lampl1993evidence], leading to changes in their motoric abilities from one day to the next.
+These changes in perspective are likely the downstream consequence of a myriad factors, chief of which may be infants' evolving locomotive abilities: an infant's ability to sit, crawl, stand, or walk structures the way they interact with the things and people in their world.
+Infants' locomotive abilities can change rapidly: the physical body of an infant can lengthen up to 2cm in 24 hours [@lampl1993evidence], leading to changes in their physical abilities from one day to the next.
 These motoric developments can be thought of as gateways that open up entirely new phases of development [@iverson2010], and cause a cascade of changes in infants ability to interact with their world and the people in it [@karasik2014].
@@
 
-More broadly, the field is in need of computational tools to reutilize these rich head-mounted camera datasets and to understand the generalizability of findings across different datasets, populations, tasks, and age-ranges.
+More broadly, the field is in need of computational tools to reutilize rich video datasets and to understand the generalizability of findings across different populations, tasks, and age-ranges.
 While the field has assembled many head-mounted camera datasets, conducting new analyses on these videos has remained prohibitively time-consuming due to a lack of computational tools for annotations.
@@
 Unlike the present dataset, infants and caregivers roamed a large, open playroom and explored different toys placed throughout.
-We thus analyzed this dataset with the goal of both (1) validating our automated method on a very different kind of dataset, attempting to replicate their primary findings originally obtained with head-mounted eye-tracking  (2) extending their findings to ask whether whether the proportion of hands in view also varies with infants’ posture in an exploratory play session.
+We thus analyzed this dataset with the goal of both (1) validating our automated method on a very different kind of dataset, attempting to replicate their primary findings originally obtained with a head-mounted eye-tracker  (2) examining whether the proportion of hands in view also varies with infants’ posture in an exploratory play session.
 
@@
 ## Free-play video dataset  
-A primary goal of this project was to provide a rich, open dataset for other researchers to examine the effects of postural developments and labeling behavior during naturalistic parent-child interactions.
+A primary goal of this project was to provide a annotated, open dataset for other researchers to examine the effects of postural developments and labeling behavior during naturalistic parent-child interactions.
 We thus invited caregivers of 8, 12, and 16-month-olds to participate in play sessions where they were provided with pairs of novel and familiar objects (e.g., a ball and a microfiber duster, called a "zem") in a small playroom in a lab (approximately 10' x 10').
-Infants wore head-mounted cameras equipped with a fish-eye lens (see Head-mounted camera), a third-person camera captured a birds-eye view of the play session.
+Infants wore head-mounted cameras equipped with a fish-eye lens (see Head-mounted camera), a tripod-mounted camera captured a third-person view of the play session.
 Using these video data, infants' posture and orientation to their caregiver were hand-coded and annotated for the entirety of the play session.
-All videos were transcribed, and we collected collected MacArthur CDIs for all children who particpated.
+All videos were transcribed, and MacArthur CDIs were collected for all children who particpated.
 All materials have been made publicly avaliable on Databrary for whom the parents provided sharing consent (29/36 dyads) via https://nyu.databrary.org/volume/101
@@
   rename(subid = subID)
-inc_subs <- read.csv(paste0(here::here(),"/data/demographics/demographics.csv"))
+inc_subs <- read.csv(paste0(here::here(),"/data/demographics/demographics_updated.csv"))
 all_subs$inc <- all_subs$reason=="included"
@@
 Participants were recruited from the surrounding community via state birth records, had no documented disabilities, and were reported to hear at least 80 percent English at home.
-Demographics and exclusion rates are given in the table below (see Table 1)/
+Demographics and exclusion rates are given in the table below (see Table 1).
 
@@
 `r sum(all_subs$reason=="no DV cam footage")` for missing scene camera footage,
-`r sum(all_subs$reason=="infant crying")` for fussiness, and one for sample symmetry.
+`r sum(all_subs$reason=="infant crying")` for fussiness, and 1 for sample symmetry.
 16-month-olds tolerated the head-mounted camera less well than younger infants, leading to a higher exclusion rate.
@@
 ### Head-mounted camera
-```{r headcam, fig.env = "figure", fig.pos = "H", out.width="40%", fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}
-knitr::include_graphics("images/view_angle_redone.png")
-```
-
-We used a small, head-mounted camera ("headcam") that was constructed from a MD80 model camera attached to a soft elastic headband.
-Videos captured by the headcam were 720x480 pixels with 25 frames per second [^2].
-A fisheye lens was attached to the camera to increase the view angle from $32^{\circ}$ horizontal by $24^{\circ}$ vertical to $64^{\circ}$ horizontal by $46^{\circ}$ vertical (see Figure \ref{fig:headcam}, above).
+
+We used a head-mounted camera ("headcam") that was constructed from a small camera attached to a soft elastic headband.[^2]  Initial participants wore an MD-80 camera, which was then replaced by a Veho pro camera which had better battery life and a larger view angle.
+The view angle of the MD-80 camera was $32^{\circ}$ horizontal by $24^{\circ}$ vertical, and we attached a fish-eye lens to the camera to increase the view angle to $64^{\circ}$ horizontal by $46^{\circ}$ vertical.
+The view angle of the Veho pro camera was wider, $47^{\circ}$ horizontal by $36^{\circ}$ vertical.
+Videos captured by MD-80/Veho cameras  were 640x480/720x480 pixels, respectively, and both cameras had a frame rate of ~30 frames per second.
+To ensure taht detections across these different cameras were comparable, we excluded detections from the outer edges of the videos taken with the MD-80 cameras based on these view angle differences in our main analyses (i.e.
+excluding top/bottom 13% of the frames, and left/right 10% of the frames).
 
@@
 
-Even with the fish-eye lens, the vertical field of view of the camera is still considerably reduced compared to the child's field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988].
-As we were originally primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field).
+However, the vertical field of view of the cameras were still considerably reduced compared to the infants' field of view, which spans around 100--120$^{\circ}$ in the vertical dimension by 6-7 months of age [@mayer1988;@cummings1988].
+As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field).
 
@@
 Parents were instructed to play with the object pairs with their child one at a time, "as they typically would."
+
 [^3]:The first few participants played with a different pair of objects (a toy cat and pedicure foam piece) that was replaced as some infants chewed persistently on the foam.
@@
 ### READ IN DETECTIONS
-d <- read_csv(paste0(here::here(),"/data/consolidated_data/consolidated_data_4detectors_july2020.csv")) %>%
+d <- read_csv(paste0(here::here(),"/data/consolidated_data/consolidated_data_4detectors_aug2020.csv")) %>%
   mutate(posture = factor(posture), orientation = factor(orientation))
+
+## OpenPose detections key
+## faceOP, wristOP == faces and wrists in equalized FOV across cameras
+## faceOPAll, wristOPAll == faces and wrists of all confidences across all areas of the frames
+## "faceOPStrict", "wristOPStrict" == faces and wrists in equalized FOV across cameras with high confidence
+## "faceOPStrictAll", "wristOPStrictAll" == faces and wrists across all areas of the frames  with high confidence
 
@@
   mutate(id = paste(subid,frame)) %>%
-  gather(key = detectorType, value = detection, faceMT, faceOP, faceOPStrict, faceVJ, wristOP, wristOPStrict) %>%
+  gather(key = detectorType, value = detection, faceMT, faceOP, faceOPAll, faceOPStrict, faceOPStrictAll, faceVJ, wristOP, wristOPAll, wristOPStrict, wristOPStrictAll) %>%
   mutate(detectorType=as.factor(detectorType))  %>%
@@
 # for better plotting names
-levels(dlong$detectorType) 
-#[1] "faceMT"        "faceOP"        "faceOPStrict"  "faceVJ"       
-#[5] "wristOP"       "wristOPStrict"
-levels(dlong$detectorType) <- c("MTCNN-Faces","OpenPose-Faces", "OpenPose-Faces-Strict","ViolaJones-Faces","OpenPose-Wrists","OpenPose-Wrists-Strict")
+# levels(dlong$detectorType)
+
+levels(dlong$detectorType) <- c("MTCNN-Faces","OpenPose-Faces","OpenPose-Faces-All", "OpenPose-Faces-Strict","OpenPose-Faces-Strict-All","ViolaJones-Faces","OpenPose-Wrists","OpenPose-Wrists-All", "OpenPose-Wrists-Strict",  "OpenPose-Wrists-Strict-All")
 
@@
 On average, posture or orientation was uncodable from 1-2 minutes of data in each child (seconds excluded from analysis for posture, M = `r round(NA_rates$mean_na_posture,2)`s, SD = `r round(NA_rates$sd_na_posture,2)`s; orientation; M = `r round(NA_rates$mean_na_posture,2)`s, SD = `r round(NA_rates$sd_na_posture,2)`s), and these rates did not vary substantially with the age of the child.
-To assess the reliability of these annotations, a second coder coded videos from three different children to calculate Cohen's kappa (posture, irr = XX, orientation, irr = XX).
+To assess the reliability of these annotations, a second coder annotated videos from three different children to calculate Cohen's kappa (posture, irr = .82; caregiver orientation, irr = .73).
 
 #### Naming event annotations 
-One coder listened to all of the audio from the play sessions and marked the exact timestamps whenever one of the novel or familiar objects was named in any instance (e.g., "Look at the [ball]", "Can you say [zem]?"); a second coder listened to the majority of the play sessions (N=23 sessions) and also annotated all naming events.
-To asesess reliability, we calculated the proportion of naming events detected by the first coder that were also annotated by a second coder within a sliding window.
-We found that 82.1% of naming events were detected within a 4 second window (+/- 2s), and 70.9% of namings events were detected within a 2 second window.
+One coder listened to all of the audio from the play sessions and marked the exact timestamps whenever one of the novel or familiar objects was named in any instance (e.g., "Look at the [ball]", "Can you say [zem]?"); a second coder listened to the majority of the play sessions (N = 23 sessions) and also annotated all naming events.
+To asesess reliability, we calculated the proportion of naming events detected by the first coder that were also annotated by the second coder within a sliding window.
+We found that 82.1% of naming events were detected within a 4 second window (+/- 2s), and 70.9% of namings events were detected within a 2 second window (+/- 1s).
 We also obtained full text transcriptions of the entire play sessions (with time stamps marking 10s intervals).
@@
 ## Face and hand detection
-We evaluated three detection systems for the ability to measure infants' access to faces.
+We evaluated three detection systems for the ability to measure infants' visual access to faces.
 The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones  [@viola2004robust].
@@
 ```{r evaluateDetectors, include=FALSE, echo=FALSE}
+## Humans saw the entire frames -- do these analyses on uncropped detections (cropped detections useful for the rest of analyses)
+
+# c("MTCNN-Faces","OpenPose-Faces","OpenPose-Faces-All", "OpenPose-Faces-Strict","OpenPose-Faces-Strict-All","ViolaJones-Faces","OpenPose-Wrists","OpenPose-Wrists-All", "OpenPose-Wrists-Strict",  "OpenPose-Wrists-Strict-All")
+
 ## Evalute face and wrist detectors
@@
   left_join(dlong)   %>%
-  filter(!detectorType %in% c('OpenPose-Wrists-Strict','OpenPose-Wrists')) %>%
+  filter(detectorType %in% c('ViolaJones-Faces','MTCNN-Faces','OpenPose-Faces-All','OpenPose-Faces-Strict-All')) %>%
   mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
@@
 
-perfFacesbySampleType <- performanceFaces %>%
+perfFacessbySampleType <- performanceFaces %>%
   group_by(detectorType, sample_type) %>%
@@
   summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
-  xtable %>%
-  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
+  # xtable %>%
+  # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
 
@@
 ## Accuracy of automated detections 
-For face detection, we found that both OpenPose and MTCNN dramatically outperformed ViolaJones, our baseline model, especially with respect to the random sample, where ViolaJones generated a relatively high proportion of both false negatives and false positives.
+For face detection, we found that both OpenPose and MTCNN dramatically outperformed ViolaJones, our baseline model, especially with respect to the random sample, where ViolaJones missed many face that were in view.
 When considering only the composite F-score across all frames, MTCNN slightly outperformed OpenPose  (`r round(perfFacesOverall$f[1],2)` MTCNN vs.
@@
 
-We next analyzed the viability of OpenPose as a hand detector.
+We next assessed the viability of OpenPose as a hand detector.
 Despite the fact that hand detection is a more computationally challenging problem [@bambach2017], and the fact that we used wrist keypoints as a proxy for hands,  OpenPose performed moderately well as a hand detector (F = `r round(perfWristsOverall$f[1],2)`).
@@
 \centering
-\begin{tabular}{rllrrr}
+\begin{tabular}{llrrr}
   \hline
@@
   OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
+  OpenPose-Faces-Strict & High density & 0.97 & 0.65 & 0.78 \\ 
+  OpenPose-Faces-Strict & Random & 0.97 & 0.62 & 0.76 \\ 
   ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
   ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
-OpenPose-Wrists & High density & 0.66 & 1.00 & 0.79 \\ 
-OpenPose-Wrists & Random & 0.88 & 0.29 & 0.44 \\ 
+  \hline
+  \hline
+  OpenPose-Wrists & High density & 0.66 & 0.96 & 0.78 \\ 
+  OpenPose-Wrists & Random & 0.88 & 0.29 & 0.43 \\ 
+  OpenPose-Wrists-Strict & High density & 0.95 & 0.44 & 0.60 \\ 
+  OpenPose-Wrists-Strict & Random & 1.00 & 0.10 & 0.18 \\
    \hline
@@
 \caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected).
-P, R, and F denote precision, recall, and F-score, respectively.}
+P, R, and F denote precision, recall, and F-score, respectively.
+"Strict" denotes when we only considered high confidence detections.}
 \vspace{-1em}
@@
 
-### Access to faces and hands
+### Access to faces and hands across age
 First, we examined the proportion of face and hand detections as a function of infants' age without considering their posture (see Figure \ref{fig:detByAge}).
@@
 We thus only considered differences in the relative proportion of faces or hands in view as a function of age, posture, and orientation, rather than comparing them directly.
-Overall, we observed that 12-month-olds appeared to have visual access to slightly fewer faces than 8 or 16-month-olds, creating a slight U-shaped function in face detections; conversely, hand detections were showed a slight increase across this age range, as reported in prior work [@fausey2016].
+Overall, we did not observe strong age related trends from 8-16 months of age; if anything, face detections showed a slight u-shaped pattern, with 12-month-olds having slightly fewer faces in their visual field that 8- or 16-month-olds.
 
@@
 
-However, these age-related trends were much smaller than the effect of infant's postural developments on infants' visual access to faces and hands.
-Infants' in-the-moment posture was a major factor both in how many faces and hands were in view during the play session, as was their orientation relative to their caregiver.
+In contrast, infant's locomotor developments had a major effect on the faces and hands that were in the field of view.
+Infants' in-the-moment posture was a major factor both in how many faces and hands were in view during the play session, as was their orientation relative to their caregiver (see Figure \ref{fig:detByPosOrient}).
 This was quantified using two generalized linear mixed-effect models estimating the proportion of faces and hands that were in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture.
@@
 prone).
-When caregivers were far away from their infants, face detections were similarity higher (Face detections; infant sitting and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, P = `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4])`; infant standing and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, P = `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`.
-Overall, we found that age did not remain a significant predictor in accounting for the faces in view when modeling infants posture, their orientation to their caregiver, and the interaction between these two factors (Face detections; Age (scaled), b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, P = `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
+When caregivers were far away from their infants, face detections were similarity higher (Face detections; infant sitting and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4])`; infant standing and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`,`r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
+Infants' age was not significant predictor in accounting for the faces in view (Face detections; Age (scaled), b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
 
@@
 prone).
-Wrist detections were  highest when caregivers were far away from their infants and those infants were standing (Wrist detections;infant standing and CG far, b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
-As with faces, age did not remain a significant predictor in these models (Wrist detections; Age (scaled), b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
-
-We directly examined the contributions of posture and orientation by fitting a reduced version of the full model [@nakagawa2013general] without their fixed effects (both models were run with the maximal random effects structure).
+Wrist detections were  highest when caregivers were far away from their infants and those infants were standing (Wrist detections; infant standing and CG far, b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
+As with faces, age was not a significant predictor in these models (Wrist detections; Age (scaled), b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
+
+We directly examined the contributions of posture and orientation vs.
+age by fitting a reduced version of the full model [@nakagawa2013general] without their fixed effects (both models were run with the maximal random effects structure).
 The fixed effects in a model with only the age of the participants accounted for relatively little variance in the proportion of faces (marginal $R^2$ < `r round(faces_age_only[1,1],5)`) or hands in view  (marginal $R^2$ = `r round(wrists_age_only[1,1],3)`).
@@
   \hline
-Intercept & -4.29 & 0.18 & -23.87 & 0.000 \\ 
-  Sit & 0.59 & 0.18 & 3.20 & 0.001 \\ 
-  Stand & 0.53 & 0.30 & 1.76 & 0.079 \\ 
-  Close & 0.53 & 0.19 & 2.75 & 0.006 \\ 
-  Far & 0.15 & 0.26 & 0.56 & 0.574 \\ 
-  Age (Scaled) & 0.16 & 0.11 & 1.48 & 0.140 \\ 
-  Sit*Close & 0.31 & 0.08 & 3.72 & 0.000 \\ 
-  Stand*Close & 0.76 & 0.09 & 8.46 & 0.000 \\ 
-  Sit*Far & 0.13 & 0.10 & 1.27 & 0.205 \\ 
-  Stand*Far & 1.76 & 0.11 & 15.87 & 0.000 \\ 
+Intercept & -3.24 & 0.20 & -16.37 & 0.000 \\ 
+  Sit & -0.00 & 0.19 & -0.02 & 0.988 \\ 
+  Stand & -0.29 & 0.20 & -1.44 & 0.151 \\ 
+  Close & 0.02 & 0.18 & 0.10 & 0.917 \\ 
+  Far & 0.51 & 0.25 & 2.04 & 0.041 \\ 
+  Age (Scaled) & 0.04 & 0.11 & 0.42 & 0.671 \\ 
+  Sit*Close & 0.90 & 0.07 & 13.80 & 0.000 \\ 
+  Stand*Close & 1.23 & 0.08 & 15.44 & 0.000 \\ 
+  Sit*Far & 0.63 & 0.07 & 9.16 & 0.000 \\ 
+  Stand*Far & 1.23 & 0.09 & 14.43 & 0.000 \\ 
    \hline
 \end{tabular}
-\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.} 
+\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.} 
 \end{table}
@@
   \hline
-Intercept & -3.25 & 0.18 & -18.04 & 0.000 \\ 
-  Sit & 0.05 & 0.17 & 0.33 & 0.744 \\ 
-  Stand & -0.24 & 0.18 & -1.34 & 0.180 \\ 
-  Close & 0.01 & 0.19 & 0.07 & 0.942 \\ 
-  Far & 0.45 & 0.27 & 1.63 & 0.102 \\ 
-  Age (Scaled) & 0.05 & 0.12 & 0.39 & 0.697 \\ 
-  Sit*Close & 0.99 & 0.07 & 14.54 & 0.000 \\ 
-  Stand*Close & 1.36 & 0.08 & 16.95 & 0.000 \\ 
-  Sit*Far & 0.75 & 0.07 & 10.48 & 0.000 \\ 
-  Stand*Far & 1.40 & 0.09 & 16.16 & 0.000 \\ 
+Intercept & -4.20 & 0.20 & -21.48 & 0.000 \\ 
+  Sit & 0.61 & 0.18 & 3.34 & 0.001 \\ 
+  Stand & 0.80 & 0.31 & 2.56 & 0.011 \\ 
+  Close & 0.37 & 0.20 & 1.87 & 0.061 \\ 
+  Far & 0.01 & 0.27 & 0.05 & 0.961 \\ 
+  Age (Scaled) & 0.09 & 0.11 & 0.86 & 0.390 \\ 
+  Sit*Close & 0.27 & 0.08 & 3.35 & 0.001 \\ 
+  Stand*Close & 0.53 & 0.09 & 6.07 & 0.000 \\ 
+  Sit*Far & 0.09 & 0.11 & 0.86 & 0.387 \\ 
+  Stand*Far & 1.57 & 0.11 & 14.16 & 0.000 \\ 
    \hline
 \end{tabular}
-\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.} 
+\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.} 
 \end{table}
@@
 
-lmer(data = aroundNamings_Detections_BySub_Fam_Faces_All, prop_detected ~ scale(age.at.test) + familiarity +  (1|subid))
+naming_faces<- lmer(data = aroundNamings_Detections_BySub_Fam_Faces_All, prop_detected ~ scale(age.at.test) + familiarity +  (1|subid))
+summary(naming_faces)
 
@@
 
-Face detections were numerically slightly higher around these novel naming events relative to baseline, with similar effects across age groups
+Face detections were not higher around these novel naming events relative to baseline, with similar effects across age groups
 (8 m.o., $M_{fam - baseline}$ = `r round(faces_namings_diffs_out$nov_vs_baseline[1],3)`,
@@
 16 m.o. $M_{nov - baseline}$ = `r round(faces_namings_diffs_out$nov_vs_baseline[3],3)`)
-however, they were not higher during familiar naming events vs. baseline (8 m.o., $M_{fam - baseline}$ =`r round(faces_namings_diffs_out$fam_vs_baseline[1],3)`,
+nor during familiar naming events vs. baseline (8 m.o., $M_{fam - baseline}$ =`r round(faces_namings_diffs_out$fam_vs_baseline[1],3)`,
 12 m.o., $M_{fam - baseline}$ = `r round(faces_namings_diffs_out$fam_vs_baseline[2],3)`,
 16 m.o. $M_{fam - baseline}$ = `r round(faces_namings_diffs_out$fam_vs_baseline[3],3)`).
-Conversely, wrist detections were higher during both familiar (8 m.o., $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$fam_vs_baseline[1],3)`,
+Conversely, wrist detections were higher during both familiar naming events (8 m.o., $M_{fam - baseline}$ =`r round(wrist_namings_diffs_out$fam_vs_baseline[1],3)`,
 12 m.o., $M_{fam - baseline}$ = `r round(wrist_namings_diffs_out$fam_vs_baseline[2],3)`,
@@
 ### Comparison between video datasets 
-First, despite having different head-mounted camera setups, the view angle of the two head-mounted cameras used in Study 1 and @franchak2017see were only slightly different ($52.2^{\circ}$ horizontal by $42.2^{\circ}$ in @franchak2017see, here $64^{\circ}$ horizontal by $46^{\circ}$ vertical).
-However, the play environment in @franchak2017see that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to climb, explore, and interact with, and infants were unconstrained and allowed to freely wander the room.
+First, the play environment in @franchak2017see that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to climb, explore, and interact with, and infants were unconstrained and allowed to freely wander the room.
 In contrast, the play room used in Study 1 was relatively small (approximately 10' x 10') and was set-up with focused play with the pairs of novel and familiar objects.
 Further, multiple people were present during the play session in @franchak2017see-- including their caregiver and two experimenters -- whereas in Study 1 the experimenters left the room during the play session.
+The view angle of the two head-mounted cameras used in Study 1 and @franchak2017see were relatively similar ($52.2^{\circ}$ horizontal by $42.2^{\circ}$ in @franchak2017see, $47^{\circ}$ horizontal by $36^{\circ}$ vertical in Study 1).
 
@@
 Overall, we found convergence between our two methodologies, replicating the main results from @franchak2017see, and finding that the proportion of faces detected was greater when infants were sitting or standing vs.
-prone.
-We found this result regardless of whether we used all detections (see Figure \ref{fig:franchak}, panel A; (proportion of frames with face detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_faces[1])`, Sitting: M = `r mean(dets_by_conf_by_posture$avg_faces[2])`, Upright: M = `r mean(dets_by_conf_by_posture$avg_faces[3])`) or restricted our analyses to high-confidence detections (percentage of frames with face detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[1])`, Sitting: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[2])`, Upright: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[3])`).
+prone (see Figure \ref{fig:franchak}).
+We found this result regardless of whether we used all detections (proportion of frames with face detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_faces[1])`, Sitting: M = `r mean(dets_by_conf_by_posture$avg_faces[2])`, Upright: M = `r mean(dets_by_conf_by_posture$avg_faces[3])`) or restricted our analyses to high-confidence detections (percentage of frames with face detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[1])`, Sitting: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[2])`, Upright: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[3])`).
 These results were confirmed in generalized linear-mixed models, as in the first study (Sitting vs.
@@
 
-The most noticeable difference between our analysis methods is the overall proportion of frames with faces detected by OpenPose vs.
+The most noticeable difference between our results is the overall proportion of frames with faces detected by OpenPose vs.
 the proportion of frames in which infants were judged to be looking at faces in @franchak2017see.
@@
 
-Overall, these analyzes extend and validate previous work, replicating the results in Study 1 that infants' in-the-moment posture modulated the proportion of hands in view, and suggesting that posture is a major factor that structures infants' access to visual information broadly construed.
+Overall, these analyses extend and validate previous work, replicating the results in Study 1 that infants' in-the-moment posture modulated the proportion of hands in view, and suggesting that posture is a major factor that structures infants' access to visual information broadly construed.
 
@@
 More broadly, though observational findings allow us to document developmental change and identify potential causal pathways, they cannot confirm them.
-Locomotive abilities are of course only part of a cascade of changes in infants' abilities and experiences, and these analyses document only a fraction of this broader, multifaceted trajectory.
-As children grow and change, the activities in which they engage with their caregivers are likely to also vary, leading to differences in the distribution of social cues that they experience that may not be captured in these in-lab play sessions.
+As children grow and change, the activities in which they engage with their caregivers are likely to also vary, leading to differences in the distribution of social cues that they experience that may not be captured.
+Finally, locomotive abilities are of course only part of a cascade of changes in infants' abilities and experiences, and these analyses document only a fraction of this broader, multifaceted trajectory in a population of children from primarily WEIRD contexts -- white, educated, industrialized, rich, and democratic [@jones2010weird].
 
 Understanding the relationship between different domains of developmental changes in naturalistic contexts has been a persistent challenge for developmental psychology.
-Nonetheless, we think that this approach hold promise for documenting these developmental trajectories and for generating new hypotheses.The field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools that deal better with noisier, more complicated datasets and extract richer information.
+Nonetheless, we think that this approach hold promise for documenting these developmental trajectories and for generating new hypotheses.
+The field of computer vision has advanced dramatically in recent years, creating a new generation of algorithmic tools that deal better with noisier, more complicated datasets and extract richer information.
 We hope that these new tools can now be leveraged to examine the consequences of the changing infant perspective for linguistic, cognitive, and social development.

### 2020-08-14 writeup-paper/draft_apa_july2020.tex
latest paper draft

@@
 \hypersetup{
-            pdftitle={Using automated detections to quantify the effect of postural developments on the social information in the infant view},
+            pdftitle={Automated detections reveal the social information in the changing infant view},
             pdfkeywords={Postural developments, headcam, social information, face detection, pose
@@
 
-\title{Using automated detections to quantify the effect of postural
-developments on the social information in the infant view}
+\title{Automated detections reveal the social information in the changing
+infant view}
 
@@
 their findings. Together, these results establish the validity of a new
-method for detecting social information, confirm that motoric
+method for detecting social information and confirm that motoric
 developments play a significant role in shaping the social information
@@
 techniques is a promising avenue for re-utilizing rich video datasets
-collected across different labs and diverse contexts and understanding
-the statistics of infants' visual and linguistic experience as they
-change over development.
+and understanding how the statistics of infants' visual and linguistic
+experience change over development.
 }
@@
 focusing on -- putting those people or objects in view (Yoshida \&
-Smith, 2008). Initial recordings from in-lab play sessions revealed a
-different view than many imagined: instead of being dominated by their
-faces, the infant perspective contained close up views of primarily toys
-and hands (Franchak, Kretch, Soska, \& Adolph, 2011; Yoshida \& Smith,
-2008; Yu \& Smith, 2017). Subsequent research has revealed that the
-infant view is far from a unitary construct, undergoing dramatic changes
-as infants grow. Recordings from home environments suggest that the
-viewpoints of very young infants -- less than 4 months of age -- do
-indeed contain persistent and frequent views of faces (Fausey,
-Jayaraman, \& Smith, 2016; Jayaraman, Fausey, \& Smith, 2017) but that
-the infant view tends to contain more and more hands as infants grow
-older.
+Smith, 2008). Initial recordings using this technique during in-lab play
+sessions revealed a different view than many imagined: instead of being
+dominated by faces, the infant perspective contained close up views of
+primarily toys and hands (Franchak, Kretch, Soska, \& Adolph, 2011;
+Yoshida \& Smith, 2008; Yu \& Smith, 2017). Subsequent research has
+revealed that the infant view is far from a unitary construct,
+undergoing dramatic changes as infants grow. Recordings from home
+environments suggest that the viewpoints of very young infants -- less
+than 4 months of age -- do indeed contain persistent and frequent views
+of faces (Fausey, Jayaraman, \& Smith, 2016; Jayaraman, Fausey, \&
+Smith, 2017) but that the infant view tends to contain more and more
+hands as infants grow older.
 
@@
 myriad factors, chief of which may be infants' evolving locomotive
-abilities: an infants ability to sit, crawl, stand, or walk structures
-the way they interact with the things and people in their world. These
-changes can be rapid: the physical body of an infant can lengthen up to
-2cm in 24 hours (Lampl, 1993), leading to changes in their motoric
-abilities from one day to the next. These motoric developments can be
-thought of as gateways that open up entirely new phases of development
-(Iverson, 2010), and cause a cascade of changes in infants ability to
-interact with their world and the people in it (Karasik, Tamis-LeMonda,
-\& Adolph, 2014).
+abilities: an infant's ability to sit, crawl, stand, or walk structures
+the way they interact with the things and people in their world.
+Infants' locomotive abilities can change rapidly: the physical body of
+an infant can lengthen up to 2cm in 24 hours (Lampl, 1993), leading to
+changes in their physical abilities from one day to the next. These
+motoric developments can be thought of as gateways that open up entirely
+new phases of development (Iverson, 2010), and cause a cascade of
+changes in infants ability to interact with their world and the people
+in it (Karasik, Tamis-LeMonda, \& Adolph, 2014).
 
@@
 More broadly, the field is in need of computational tools to reutilize
-these rich head-mounted camera datasets and to understand the
-generalizability of findings across different datasets, populations,
-tasks, and age-ranges. While the field has assembled many head-mounted
-camera datasets, conducting new analyses on these videos has remained
-prohibitively time-consuming due to a lack of computational tools for
-annotations. Instead, hundreds of hours of manual annotations have been
-required to analyze a fraction of the available frames for a given
-analysis. Thus, despite containing a wealth of information about the
-structure of parent-child interactions, these datasets have thus gone
-dramatically underused.
+rich video datasets and to understand the generalizability of findings
+across different populations, tasks, and age-ranges. While the field has
+assembled many head-mounted camera datasets, conducting new analyses on
+these videos has remained prohibitively time-consuming due to a lack of
+computational tools for annotations. Instead, hundreds of hours of
+manual annotations have been required to analyze a fraction of the
+available frames for a given analysis. Thus, despite containing a wealth
+of information about the structure of parent-child interactions, these
+datasets have thus gone dramatically underused.
 
@@
 kind of dataset, attempting to replicate their primary findings
-originally obtained with head-mounted eye-tracking (2) extending their
-findings to ask whether whether the proportion of hands in view also
-varies with infants' posture in an exploratory play session.
+originally obtained with a head-mounted eye-tracker (2) examining
+whether the proportion of hands in view also varies with infants'
+posture in an exploratory play session.
 
@@
 
-A primary goal of this project was to provide a rich, open dataset for
-other researchers to examine the effects of postural developments and
-labeling behavior during naturalistic parent-child interactions. We thus
-invited caregivers of 8, 12, and 16-month-olds to participate in play
-sessions where they were provided with pairs of novel and familiar
+A primary goal of this project was to provide a annotated, open dataset
+for other researchers to examine the effects of postural developments
+and labeling behavior during naturalistic parent-child interactions. We
+thus invited caregivers of 8, 12, and 16-month-olds to participate in
+play sessions where they were provided with pairs of novel and familiar
 objects (e.g., a ball and a microfiber duster, called a \enquote{zem})
@@
 head-mounted cameras equipped with a fish-eye lens (see Head-mounted
-camera), a third-person camera captured a birds-eye view of the play
-session. Using these video data, infants' posture and orientation to
-their caregiver were hand-coded and annotated for the entirety of the
-play session. All videos were transcribed, and we collected collected
-MacArthur CDIs for all children who particpated. All materials have been
-made publicly avaliable on Databrary for whom the parents provided
-sharing consent (29/36 dyads) via
-\url{https://nyu.databrary.org/volume/101}
+camera), a tripod-mounted camera captured a third-person view of the
+play session. Using these video data, infants' posture and orientation
+to their caregiver were hand-coded and annotated for the entirety of the
+play session. All videos were transcribed, and MacArthur CDIs were
+collected for all children who particpated. All materials have been made
+publicly avaliable on Databrary for whom the parents provided sharing
+consent (29/36 dyads) via \url{https://nyu.databrary.org/volume/101}
 
@@
 were reported to hear at least 80 percent English at home. Demographics
-and exclusion rates are given in the table below (see Table 1)/
+and exclusion rates are given in the table below (see Table 1).
 
@@
 Communicative Development Inventory (CDI) data, 2 for missing scene
-camera footage, 1 for fussiness, and one for sample symmetry.
+camera footage, 1 for fussiness, and 1 for sample symmetry.
 16-month-olds tolerated the head-mounted camera less well than younger
@@
 
-\begin{figure}[H]
-\includegraphics[width=0.4\linewidth]{images/view_angle_redone} \caption{Vertical field of view for two different headcam configurations (we used the lower in our current study).}\label{fig:headcam}
-\end{figure}
-
-We used a small, head-mounted camera (\enquote{headcam}) that was
-constructed from a MD80 model camera attached to a soft elastic
-headband. Videos captured by the headcam were 720x480 pixels with 25
-frames per second.\footnote{Detailed instructions for creating this
-  headcam can be found at
+We used a head-mounted camera (\enquote{headcam}) that was constructed
+from a small camera attached to a soft elastic headband.\footnote{Detailed
+  instructions for creating this headcam can be found at
   \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
-A fisheye lens was attached to the camera to increase the view angle
-from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
-\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
-\ref{fig:headcam}, above).
-
-Even with the fish-eye lens, the vertical field of view of the camera is
-still considerably reduced compared to the child's field of view, which
-spans around 100--120\(^{\circ}\) in the vertical dimension by 6-7
-months of age (Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton,
-1988; Mayer, Fulton, \& Cummings, 1988). As we were originally primarily
-interested in the presence of faces in the child's field of view, we
-chose to orient the camera upwards to capture the entirety of the
-child's upper visual field where the child is likely to see adult faces,
-understanding that this decision limited our ability to detect hands
-(especially those of the child, which are typically found at the bottom
-of the visual field).
+Initial participants wore an MD-80 camera, which was then replaced by a
+Veho pro camera which had better battery life and a larger view angle.
+The view angle of the MD-80 camera was \(32^{\circ}\) horizontal by
+\(24^{\circ}\) vertical, and we attached a fish-eye lens to the camera
+to increase the view angle to \(64^{\circ}\) horizontal by
+\(46^{\circ}\) vertical. The view angle of the Veho pro camera was
+wider, \(47^{\circ}\) horizontal by \(36^{\circ}\) vertical. Videos
+captured by MD-80/Veho cameras were 640x480/720x480 pixels,
+respectively, and both cameras had a frame rate of \textasciitilde{}30
+frames per second. To ensure taht detections across these different
+cameras were comparable, we excluded detections from the outer edges of
+the videos taken with the MD-80 cameras based on these view angle
+differences in our main analyses (i.e.~excluding top/bottom 13\% of the
+frames, and left/right 10\% of the frames).
+
+However, the vertical field of view of the cameras were still
+considerably reduced compared to the infants' field of view, which spans
+around 100--120\(^{\circ}\) in the vertical dimension by 6-7 months of
+age (Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton, 1988; Mayer,
+Fulton, \& Cummings, 1988). As we were primarily interested in the
+presence of faces in the child's field of view, we chose to orient the
+camera upwards to capture the entirety of the child's upper visual field
+where the child is likely to see adult faces, understanding that this
+decision limited our ability to detect hands (especially those of the
+child, which are typically found at the bottom of the visual field).
 
@@
 object pairs with their child one at a time, \enquote{as they typically
-would.} {[}\^{}3{]}:The first few participants played with a different
-pair of objects (a toy cat and pedicure foam piece) that was replaced as
-some infants chewed persistently on the foam.
+would.}
 
@@
 minutes of data in each child (seconds excluded from analysis for
-posture, M = 94.01s, SD = 225.18s; orientation; M = 94.01s, SD =
-225.18s), and these rates did not vary substantially with the age of the
+posture, M = 104.99s, SD = 233.86s; orientation; M = 104.99s, SD =
+233.86s), and these rates did not vary substantially with the age of the
 child. To assess the reliability of these annotations, a second coder
-coded videos from three different children to calculate Cohen's kappa
-(posture, irr = XX, orientation, irr = XX).
+annotated videos from three different children to calculate Cohen's
+kappa (posture, irr = .82; caregiver orientation, irr = .73).
 
@@
 \enquote{Can you say {[}zem{]}?}); a second coder listened to the
-majority of the play sessions (N=23 sessions) and also annotated all
+majority of the play sessions (N = 23 sessions) and also annotated all
 naming events. To asesess reliability, we calculated the proportion of
-naming events detected by the first coder that were also annotated by a
-second coder within a sliding window. We found that 82.1\% of naming
+naming events detected by the first coder that were also annotated by
+the second coder within a sliding window. We found that 82.1\% of naming
 events were detected within a 4 second window (+/- 2s), and 70.9\% of
-namings events were detected within a 2 second window. We also obtained
-full text transcriptions of the entire play sessions (with time stamps
-marking 10s intervals). While these full transcriptions are not used in
-the present analyses, they have been made publically avaliable for
-future research.
+namings events were detected within a 2 second window (+/- 1s). We also
+obtained full text transcriptions of the entire play sessions (with time
+stamps marking 10s intervals). While these full transcriptions are not
+used in the present analyses, they have been made publically avaliable
+for future research.
 
@@
 We evaluated three detection systems for the ability to measure infants'
-access to faces. The first of these is the most commonly-used and widely
-available face detection algorithm: Viola-Jones (Viola \& Jones, 2004).
-We used this algorithm as a benchmark for performance, as while it can
-achieve impressive accuracy in some situations, it is notoriously bad at
-dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox, 2014).
-We next tested the performance of two face detectors that both made use
-of relatively recently developed Convolutional Neural Networks (CNNs) to
-extract face information. The first algorithm was specifically optimized
-for face detection, and the second algorithm was optimized to extract
-pose information of all the individuals in an image, operationalized as
-information about the position of 18 different body parts. For this
-second algorithm (OpenPose; Cao et al., 2017), we used the agent's nose
-(one of the body keypoints detected) to operationalize the presence of
-faces, as any half of a face necessarily contains a nose.
+visual access to faces. The first of these is the most commonly-used and
+widely available face detection algorithm: Viola-Jones (Viola \& Jones,
+2004). We used this algorithm as a benchmark for performance, as while
+it can achieve impressive accuracy in some situations, it is notoriously
+bad at dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox,
+2014). We next tested the performance of two face detectors that both
+made use of relatively recently developed Convolutional Neural Networks
+(CNNs) to extract face information. The first algorithm was specifically
+optimized for face detection, and the second algorithm was optimized to
+extract pose information of all the individuals in an image,
+operationalized as information about the position of 18 different body
+parts. For this second algorithm (OpenPose; Cao et al., 2017), we used
+the agent's nose (one of the body keypoints detected) to operationalize
+the presence of faces, as any half of a face necessarily contains a
+nose.
 
@@
 outperformed ViolaJones, our baseline model, especially with respect to
-the random sample, where ViolaJones generated a relatively high
-proportion of both false negatives and false positives. When considering
-only the composite F-score across all frames, MTCNN slightly
-outperformed OpenPose (0.89 MTCNN vs.~0.83 OpenPose), and MTCNN and
-OpenPose performed comparably with the random sample. Generally, MTCNN
-exhibited higher precision, whereas OpenPose exhibited higher recall,
-and these differences were most pronounced on the randomly sample
-frames. In other words, while OpenPose generated slightly more false
-positives than MTCNN, MTCNN missed several faces that were accurately
-detected by OpenPose. When we restricted our analysis to high-confidence
-detections from OpenPose (\textgreater{}.5 confidence; default threshold
-for visualization), we found very high precision (P = 0.97), but much
-lower recall (R = 0.64) and thus overall performance (F = 0.77),
-indicating that these low-confidence detections often indexed actual
-faces that were in the infant view. Figure \ref{fig:exampledetections}
-shows example successful detections from OpenPose in each age group, and
-Figure \ref{fig:failures} shows examples of missed faces as well as
-false positives for context.
-
-We next analyzed the viability of OpenPose as a hand detector. Despite
+the random sample, where ViolaJones missed many face that were in view.
+When considering only the composite F-score across all frames, MTCNN
+slightly outperformed OpenPose (0.89 MTCNN vs.~0.83 OpenPose), and MTCNN
+and OpenPose performed comparably with the random sample. Generally,
+MTCNN exhibited higher precision, whereas OpenPose exhibited higher
+recall, and these differences were most pronounced on the randomly
+sample frames. In other words, while OpenPose generated slightly more
+false positives than MTCNN, MTCNN missed several faces that were
+accurately detected by OpenPose. When we restricted our analysis to
+high-confidence detections from OpenPose (\textgreater{}.5 confidence;
+default threshold for visualization), we found very high precision (P =
+0.97), but much lower recall (R = 0.64) and thus overall performance (F
+= 0.77), indicating that these low-confidence detections often indexed
+actual faces that were in the infant view. Figure
+\ref{fig:exampledetections} shows example successful detections from
+OpenPose in each age group, and Figure \ref{fig:failures} shows examples
+of missed faces as well as false positives for context.
+
+We next assessed the viability of OpenPose as a hand detector. Despite
 the fact that hand detection is a more computationally challenging
@@
 used wrist keypoints as a proxy for hands, OpenPose performed moderately
-well as a hand detector (F = 0.74). OpenPose achieved relatively high
+well as a hand detector (F = 0.73). OpenPose achieved relatively high
 precision -- generating relatively few false positives -- but showed low
@@
 detections, we found much higher precision (P = 0.95), but much lower
-recall (R = 0.37) and thus overall performance (F = 0.53).
+recall (R = 0.36) and thus overall performance (F = 0.52).
 
@@
 
-\subsubsection{Access to faces and
-hands}\label{access-to-faces-and-hands}
+\subsubsection{Access to faces and hands across
+age}\label{access-to-faces-and-hands-across-age}
 
@@
 view as a function of age, posture, and orientation, rather than
-comparing them directly. Overall, we observed that 12-month-olds
-appeared to have visual access to slightly fewer faces than 8 or
-16-month-olds, creating a slight U-shaped function in face detections;
-conversely, hand detections were showed a slight increase across this
-age range, as reported in prior work (Fausey et al., 2016).
+comparing them directly. Overall, we did not observe strong age related
+trends from 8-16 months of age; if anything, face detections showed a
+slight u-shaped pattern, with 12-month-olds having slightly fewer faces
+in their visual field that 8- or 16-month-olds.
 
@@
 
-However, these age-related trends were much smaller than the effect of
-infant's postural developments on infants' visual access to faces and
-hands. Infants' in-the-moment posture was a major factor both in how
-many faces and hands were in view during the play session, as was their
-orientation relative to their caregiver. This was quantified using two
-generalized linear mixed-effect models estimating the proportion of
-faces and hands that were in view, with orientation, posture, their
-interaction, and scaled participant's age as fixed effects, and with
-random slopes for infants' orientation and posture. A summary of the
-coefficients of the models can be found in Tables 2 and 3.
+In contrast, infant's locomotor developments had a major effect on the
+faces and hands that were in the field of view. Infants' in-the-moment
+posture was a major factor both in how many faces and hands were in view
+during the play session, as was their orientation relative to their
+caregiver (see Figure \ref{fig:detByPosOrient}). This was quantified
+using two generalized linear mixed-effect models estimating the
+proportion of faces and hands that were in view, with orientation,
+posture, their interaction, and scaled participant's age as fixed
+effects, and with random slopes for infants' orientation and posture. A
+summary of the coefficients of the models can be found in Tables 2 and
+3.
 
@@
 sitting or standing had more faces in view (Face detections; infant
-sitting and CG close, b = 0.99, SE = 0.07, Z = 14.54, P \textless{}
-0.001; infant standing and CG close, b = 1.36, SE = 0.08, Z = 16.95, P
+sitting and CG close, b = 0.90, SE = 0.07, Z = 13.80, P \textless{}
+0.001; infant standing and CG close, b = 1.23, SE = 0.08, Z = 15.44, P
 \textless{} 0.001) than infants who were lying down/crawling
@@
 detections were similarity higher (Face detections; infant sitting and
-CG far, b = 0.75, SE = 0.07, Z = 10.48, P = P \textless{} 0.001; infant
-standing and CG far, b = 1.40, SE = 0.09, Z = 16.16, P = P \textless{}
-0.001. Overall, we found that age did not remain a significant predictor
-in accounting for the faces in view when modeling infants posture, their
-orientation to their caregiver, and the interaction between these two
-factors (Face detections; Age (scaled), b = 0.05, SE = 0.12, Z = 0.39, P
-= P = 0.697).
+CG far, b = 0.63, SE = 0.07, Z = 9.16, P \textless{} 0.001; infant
+standing and CG far, b = 1.23, SE = 0.09, Z = 14.43,P \textless{}
+0.001). Infants' age was not significant predictor in accounting for the
+faces in view (Face detections; Age (scaled), b = 0.04, SE = 0.11, Z =
+0.42, P = 0.671).
 
@@
 relatively closer to their infants (Hand detections; infant sitting and
-CG close, b=0.31, SE = 0.08, Z = 3.72, P \textless{} 0.001; infant
-standing and CG close, b=0.76, SE = 0.09, Z = 8.46, P \textless{} 0.001)
+CG close, b=0.27, SE = 0.08, Z = 3.35, P \textless{} 0.001; infant
+standing and CG close, b=0.53, SE = 0.09, Z = 6.07, P \textless{} 0.001)
 than infants who were lying down/crawling (i.e.~prone). Wrist detections
 were highest when caregivers were far away from their infants and those
-infants were standing (Wrist detections;infant standing and CG far, b =
-1.76, SE = 0.11, Z = 15.87, P \textless{} 0.001). As with faces, age did
-not remain a significant predictor in these models (Wrist detections;
-Age (scaled), b = 0.16, SE = 0.11, Z = 1.48, P = 0.140).
-
-We directly examined the contributions of posture and orientation by
-fitting a reduced version of the full model (Nakagawa \& Schielzeth,
-2013) without their fixed effects (both models were run with the maximal
-random effects structure). The fixed effects in a model with only the
-age of the participants accounted for relatively little variance in the
-proportion of faces (marginal \(R^2\) \textless{} 0.01) or hands in view
-(marginal \(R^2\) = 0.02). However, when adding infants' posture and
-orientation to their caregiver to the model (and their interaction) the
-marginal \(R^2\) were higher for both faces (marginal \(R^2\) = 0.28)
-and wrists (marginal \(R^2\) = 0.31). Overall, these results suggest
-that infants' visual access to social information is largely modulated
-by their posture and orientation to their caregiver, which is in turn a
-function of their general locomotor development.
+infants were standing (Wrist detections; infant standing and CG far, b =
+1.57, SE = 0.11, Z = 14.16, P \textless{} 0.001). As with faces, age was
+not a significant predictor in these models (Wrist detections; Age
+(scaled), b = 0.09, SE = 0.11, Z = 0.86, P = 0.390).
+
+We directly examined the contributions of posture and orientation
+vs.~age by fitting a reduced version of the full model (Nakagawa \&
+Schielzeth, 2013) without their fixed effects (both models were run with
+the maximal random effects structure). The fixed effects in a model with
+only the age of the participants accounted for relatively little
+variance in the proportion of faces (marginal \(R^2\) \textless{} 0.01)
+or hands in view (marginal \(R^2\) = 0.02). However, when adding
+infants' posture and orientation to their caregiver to the model (and
+their interaction) the marginal \(R^2\) were higher for both faces
+(marginal \(R^2\) = 0.25) and wrists (marginal \(R^2\) = 0.26). Overall,
+these results suggest that infants' visual access to social information
+is largely modulated by their posture and orientation to their
+caregiver, which is in turn a function of their general locomotor
+development.
 
@@
 
-Face detections were numerically slightly higher around these novel
-naming events relative to baseline, with similar effects across age
-groups (8 m.o., \(M_{fam - baseline}\) = 0.02, 12 m.o.,
-\(M_{nov - baseline}\) = 0.05, 16 m.o. \(M_{nov - baseline}\) = 0.02)
-however, they were not higher during familiar naming events vs.~baseline
-(8 m.o., \(M_{fam - baseline}\) =0.00, 12 m.o., \(M_{fam - baseline}\) =
-0.00, 16 m.o. \(M_{fam - baseline}\) = 0.02). Conversely, wrist
-detections were higher during both familiar (8 m.o.,
-\(M_{fam - baseline}\) =0.07, 12 m.o., \(M_{fam - baseline}\) = 0.04, 16
-m.o. \(M_{fam - baseline}\) = 0.08) and novel naming events relative to
-baseline across all age groups (8 m.o., \(M_{nov - baseline}\) = 0.09,
-12 m.o., \(M_{nov - baseline}\) = 0.08, 16 m.o. \(M_{nov - baseline}\) =
-0.08). These results were confirmed by a linear mixed-effect model with
-scaled aged as a fixed effect and random intercepts for each subject
-(Wrist detections; familiar objects vs.~baseline, b = 0.06, SE = 0.01, t
-= 4.33, P \textless{} 0.001; Novel objects vs.~baseline, b = 0.08, SE =
-0.02, t = 5.21, P \textless{} 0.001).
+Face detections were not higher around these novel naming events
+relative to baseline, with similar effects across age groups (8 m.o.,
+\(M_{fam - baseline}\) = 0.01, 12 m.o., \(M_{nov - baseline}\) = 0.04,
+16 m.o. \(M_{nov - baseline}\) = 0.02) nor during familiar naming events
+vs.~baseline (8 m.o., \(M_{fam - baseline}\) =0.00, 12 m.o.,
+\(M_{fam - baseline}\) = 0.00, 16 m.o. \(M_{fam - baseline}\) = 0.01).
+Conversely, wrist detections were higher during both familiar naming
+events (8 m.o., \(M_{fam - baseline}\) =0.03, 12 m.o.,
+\(M_{fam - baseline}\) = 0.04, 16 m.o. \(M_{fam - baseline}\) = 0.06)
+and novel naming events relative to baseline across all age groups (8
+m.o., \(M_{nov - baseline}\) = 0.06, 12 m.o., \(M_{nov - baseline}\) =
+0.07, 16 m.o. \(M_{nov - baseline}\) = 0.07). These results were
+confirmed by a linear mixed-effect model with scaled aged as a fixed
+effect and random intercepts for each subject (Wrist detections;
+familiar objects vs.~baseline, b = 0.05, SE = 0.01, t = 3.76, P
+\textless{} 0.001; Novel objects vs.~baseline, b = 0.06, SE = 0.01, t =
+5.01, P \textless{} 0.001).
 
@@
 
-First, despite having different head-mounted camera setups, the view
-angle of the two head-mounted cameras used in Study 1 and Franchak et
-al. (2017) were only slightly different (\(52.2^{\circ}\) horizontal by
-\(42.2^{\circ}\) in Franchak et al. (2017), here \(64^{\circ}\)
-horizontal by \(46^{\circ}\) vertical). However, the play environment in
-Franchak et al. (2017) that infants were immersed in with their
-caregivers (and experimenters) was much larger and more varied than the
-play room used in the present dataset, containing multiple structures
-and toys in different parts of the room for infants to climb, explore,
-and interact with, and infants were unconstrained and allowed to freely
-wander the room. In contrast, the play room used in Study 1 was
-relatively small (approximately 10' x 10') and was set-up with focused
-play with the pairs of novel and familiar objects. Further, multiple
-people were present during the play session in Franchak et al. (2017)--
-including their caregiver and two experimenters -- whereas in Study 1
-the experimenters left the room during the play session.
+First, the play environment in Franchak et al. (2017) that infants were
+immersed in with their caregivers (and experimenters) was much larger
+and more varied than the play room used in the present dataset,
+containing multiple structures and toys in different parts of the room
+for infants to climb, explore, and interact with, and infants were
+unconstrained and allowed to freely wander the room. In contrast, the
+play room used in Study 1 was relatively small (approximately 10' x 10')
+and was set-up with focused play with the pairs of novel and familiar
+objects. Further, multiple people were present during the play session
+in Franchak et al. (2017)-- including their caregiver and two
+experimenters -- whereas in Study 1 the experimenters left the room
+during the play session. The view angle of the two head-mounted cameras
+used in Study 1 and Franchak et al. (2017) were relatively similar
+(\(52.2^{\circ}\) horizontal by \(42.2^{\circ}\) in Franchak et al.
+(2017), \(47^{\circ}\) horizontal by \(36^{\circ}\) vertical in Study
+1).
 
@@
 proportion of faces detected was greater when infants were sitting or
-standing vs.~prone. We found this result regardless of whether we used
-all detections (see Figure \ref{fig:franchak}, panel A; (proportion of
-frames with face detections; Prone: M = 0.13, Sitting: M = 0.20,
-Upright: M = 0.22) or restricted our analyses to high-confidence
-detections (percentage of frames with face detections; Prone: M = 0.03,
-Sitting: M = 0.06, Upright: M = 0.05). These results were confirmed in
-generalized linear-mixed models, as in the first study (Sitting
-vs.~Prone, b = 0.34, SE = 0.02, Z = 23.37, P \textless{} 0.001. Upright
-vs.~Prone, b = 0.38, SE = 0.02, Z = 24.01, P \textless{} 0.001).
-
-The most noticeable difference between our analysis methods is the
-overall proportion of frames with faces detected by OpenPose vs.~the
-proportion of frames in which infants were judged to be looking at faces
-in Franchak et al. (2017). Across the entire session in Franchak et al.
+standing vs.~prone (see Figure \ref{fig:franchak}). We found this result
+regardless of whether we used all detections (proportion of frames with
+face detections; Prone: M = 0.13, Sitting: M = 0.20, Upright: M = 0.22)
+or restricted our analyses to high-confidence detections (percentage of
+frames with face detections; Prone: M = 0.03, Sitting: M = 0.06,
+Upright: M = 0.05). These results were confirmed in generalized
+linear-mixed models, as in the first study (Sitting vs.~Prone, b = 0.34,
+SE = 0.02, Z = 23.37, P \textless{} 0.001. Upright vs.~Prone, b = 0.38,
+SE = 0.02, Z = 24.01, P \textless{} 0.001).
+
+The most noticeable difference between our results is the overall
+proportion of frames with faces detected by OpenPose vs.~the proportion
+of frames in which infants were judged to be looking at faces in
+Franchak et al. (2017). Across the entire session in Franchak et al.
 (2017), infants looked at faces on 4.7\% of frames (Franchak et al.,
@@
 
-Overall, these analyzes extend and validate previous work, replicating
+Overall, these analyses extend and validate previous work, replicating
 the results in Study 1 that infants' in-the-moment posture modulated the
@@
 though observational findings allow us to document developmental change
-and identify potential causal pathways, they cannot confirm them.
-Locomotive abilities are of course only part of a cascade of changes in
-infants' abilities and experiences, and these analyses document only a
-fraction of this broader, multifaceted trajectory. As children grow and
-change, the activities in which they engage with their caregivers are
-likely to also vary, leading to differences in the distribution of
-social cues that they experience that may not be captured in these
-in-lab play sessions.
+and identify potential causal pathways, they cannot confirm them. As
+children grow and change, the activities in which they engage with their
+caregivers are likely to also vary, leading to differences in the
+distribution of social cues that they experience that may not be
+captured. Finally, locomotive abilities are of course only part of a
+cascade of changes in infants' abilities and experiences, and these
+analyses document only a fraction of this broader, multifaceted
+trajectory in a population of children from primarily WEIRD contexts --
+white, educated, industrialized, rich, and democratic (D. Jones, 2010).
 
@@
 approach hold promise for documenting these developmental trajectories
-and for generating new hypotheses.The field of computer vision has
+and for generating new hypotheses. The field of computer vision has
 advanced dramatically in recent years, creating a new generation of
@@
 
+\hypertarget{ref-jones2010weird}{}
+Jones, D. (2010). A weird view of human nature skews psychologists'
+studies. American Association for the Advancement of Science.
+
 \hypertarget{ref-karasik2014}{}

### 2020-08-24 writeup-paper/draft_apa_blinded.Rmd
latest paper draft

--- 2020-08-14 writeup-paper/draft_apa_july2020.Rmd
@@
 abstract: |
+•	We provided an open, annotated dataset of naturalistic parent-child interactions during a play session with novel and familiar toys, recording egocentric and third-person views.
+•	We develop a novel method using a pose detection model to automatically quantify the faces and hands in the infant view.
+•	We apply our novel method to this and another egoecentric video dataset, confirming that locomotive developments shape the faces and hands that infants see.
+•	The combined use of head-mounted cameras and automated analyses have the potential to reveal the changing statistics of the infant view.
+
   Social information plays a key role in infants\' early linguistic and cognitive development.
@@
 How do these postural developments affect infants access to the social information relevant for early learning?
-To address this question, we first created a rich annotated dataset of naturalistic play sessions between infants and their caregivers with sets of novel and familiar toys (N=36, 8-16 months of age), recording both egocentric and third-person video.
-We then developed an automated analysis method using a pose detection model (Zhang et al., 2017) to detect the faces and hands in the infant view, allowing us to analyze the entirety of this dataset.
+To address this question, we first created a rich annotated dataset of naturalistic play sessions between infants and their caregivers with sets of novel and familiar toys (N=36, 8-16 month-olds), recording both egocentric and third-person video.
+We then developed an automated analysis method using a pose detection model to detect the faces and hands in the infant view.
 We found that infants\' posture and orientation to their caregiver changed dramatically across this age range and modulated infants access to social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces and hands in their visual field.
-We also applied our automated method to a second egocentric video dataset where one-year-olds explored a large play room with their caregivers  (Franchak et al., 2017), replicating and extending their findings.
-Together, these results establish the validity of a new method for detecting social information and confirm that motoric developments play a significant role in shaping the social information that infants have access to.
-We propose that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for re-utilizing rich video datasets and understanding how the statistics of infants\' visual and linguistic experience change over development.
+We applied this automated method to a second egocentric video dataset from a previous study, replicating and extending their findings on the influence of infants' posture on the social information in view.
+Together, these results establish the validity of a new method for detecting social information in egocentric videos and confirm that motoric developments play a significant role in shaping the social information that infants have access to.
+The combined use of head-mounted cameras and new computer vision techniques is a promising avenue for understanding how the statistics of infants\' visual and linguistic experience change over development.
 
 keywords          : "Postural developments, headcam, social information, face detection, pose detection"
-wordcount         : "XXX"
+wordcount         : "7324"
 
@@
 
-figsintext        : no
-figurelist        : no
+figsintext        : yes
+figurelist        : yes
 tablelist         : no
@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_docx
 ---
@@
 library(knitr)
-knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
+knitr::opts_chunk$set(fig.crop = FALSE, 
                       fig.path='figs/', echo=FALSE, warning = FALSE, 
@@
 And as infants reach their first birthday, they also tend to follow [@yu2013joint; @yu2017hand] and imitate the gestures of their caregivers (e.g., pointing).
-
 Infants ability to process these social cues may provide strong scaffolding for early word learning.
@@
 
-Relatively little work, however, has quantified how often infants see and use these kinds of social cues in naturalistic learing environments.
+Relatively little work, however, has quantified how often infants see and use these kinds of social cues in naturalistic learning environments.
 By using head-mounted cameras to record what infants see, researchers have begun to document the infant egocentric perspective [@yoshida2008] and to quantify the information -- social and otherwise -- available to infants as they learn.
@@
 Initial recordings using this technique during in-lab play sessions revealed a different view than many imagined: instead of being dominated by faces, the infant perspective contained close up views of primarily toys and hands [@yoshida2008; @franchak2011; @yu2017hand].
-Subsequent research has revealed that the infant view is far from a unitary construct, undergoing dramatic changes as infants grow.
+Subsequent research has revealed that the infant view undergoes dramatic changes as infants grow.
 Recordings from home environments suggest that the viewpoints of very young infants -- less than 4 months of age -- do indeed contain persistent and frequent views of faces [@fausey2016; @jayaraman2017faces] but that the infant view tends to contain more and more hands as infants grow older.
 
-These changes in perspective are likely the downstream consequence of a myriad factors, chief of which may be infants' evolving locomotive abilities: an infant's ability to sit, crawl, stand, or walk structures the way they interact with the things and people in their world.
-Infants' locomotive abilities can change rapidly: the physical body of an infant can lengthen up to 2cm in 24 hours [@lampl1993evidence], leading to changes in their physical abilities from one day to the next.
-These motoric developments can be thought of as gateways that open up entirely new phases of development [@iverson2010], and cause a cascade of changes in infants ability to interact with their world and the people in it [@karasik2014].
-
+These changes in perspective are likely the downstream consequence of myriad factors, chief of which may be infants' evolving locomotive abilities: an infant's ability to sit, crawl, stand, or walk structures the way they interact with the things and people in their world.
+Infants' locomotive abilities can change rapidly: the physical body of an infant can lengthen up to 2cm in 24 hours [@lampl1993evidence], leading to changes in their abilities from one day to the next.
+These motoric developments can be thought of as gateways that open up entirely new phases of development [@iverson2010], causing a cascade of changes in an infant's ability to interact with their world and the people in it [@karasik2014].
 Thus, infants’ changing locomotor abilities may shape the social cues that infants' see and seek out, in turn impacting their cognitive and linguistic abilities.
-Supporting this idea, infants experience with sitting predicts their success at 3D object completion tasks [@soska2010systems] as well as their receptive vocabulary [@libertus2016sit].
-And as children begin crawling [@adolph1998roles] -- or scooting or cruising [@patrick2012developmental] -- they are no longer constrained to the same spot that their caregivers last placed them in.
+
+However, understanding the nuanced relationship between the social information in the infant view and their motoric and linguistic development has remained challenging.
+For example, as children begin crawling [@adolph1998roles] -- or scooting or cruising [@patrick2012developmental] -- they are no longer constrained to the same spot that their caregivers last placed them in.
 Yet while crawlers can choose where to go and what they see to a much greater degree, they also appear to spend much of their time in a world populated by floors and knees; during spontaneous play, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011], when they have full visual access to their environment and the people in it [@kretch2014].
-
-On one view, it is children’s ability to stand and walk that fundamentally changes their ability to access social information (e.g., facial expressions, gaze cues, pointing) relative to children who are still crawling and sitting [@walle2016infant].
-In turn, this ability to access more detailed social information could allow infants to learn words quicker and more efficiently, facilitating language growth [@walle2016infant].
-Supporting this idea, walking infants make different kinds of object-related bids for attention from their caregivers than crawling infants, and tend to hear more action directed statements (e.g., “open it”) [@karasik2014]; further, in an observational study, @walle2014 also found that children who were able to walk had both higher receptive and productive vocabularies.
-However, not all evidence supports this view: some parental report data suggest a weak relationship between walking and the onset of language [@moore2019point], and head-mounted eye-tracking data have highlighted the role of infants’ in-the-moment posture -- and that of their caregivers’ -- in what infants’ see, finding that these factors are better predictors of the social information in view than their locomotor status per se (i.e., “walker” vs “crawler”) [@franchak2017see].
-
-Understanding this nuanced relationship between the social information in the infant view and their motoric and linguistic development has remained challenging.
-Indeed, different kinds of play sessions (e.g., exploring novel environments vs.
+Further, infants' experience with sitting predicts their success at 3D object completion tasks [@soska2010systems] as well as their receptive vocabulary [@libertus2016sit], suggesting the importance of focused play sessions for language development.
+
+Others have posited that it is children’s ability to stand and walk that fundamentally changes their ability to access social information (e.g., facial expressions, gaze cues, pointing) relative to children who are still crawling and sitting [@walle2016infant], which in turn allow infants to learn words quicker and more efficiently [@walle2016infant].
+Supporting this idea, walking vs.
+crawling infants tend to make different kinds of object-related bids for attention from their caregivers [@karasik2014], hear more action directed statements (e.g., “open it”) [@karasik2014], and have higher receptive and productive vocabularies [@walle2014].
+However, not all evidence supports this view: parental report data suggest a weak relationship between walking and the onset of language [@moore2019point], and head-mounted eye-tracking data from one-year-olds highlights the role of infants’ in-the-moment posture -- and that of their caregivers’ -- in what infants’ see, finding that these factors are better predictors of the social information in view than their locomotor status per se (i.e., “walker” vs “crawler”) [@franchak2017see].
+
+Broadly, the field is in need of computational tools to reuse these rich video datasets and understand the generalizability of findings across different populations, tasks, and age-ranges.
+For example, different kinds of play sessions (e.g., exploring novel environments vs.
 playing with novel objects) might modulate the degree to which infants’ posture changes the social information in view.
-Furthermore, no work to date has directly examined how the availability of social information changes in naturalistic language learning contexts.
-For example, it is unclear if and how the social information in view changes during the naming events for different objects (e.g, “Look, a [ball]”).
-
-More broadly, the field is in need of computational tools to reutilize rich video datasets and to understand the generalizability of findings across different populations, tasks, and age-ranges.
-While the field has assembled many head-mounted camera datasets, conducting new analyses on these videos has remained prohibitively time-consuming due to a lack of computational tools for annotations.
+Yet while the field has assembled many head-mounted camera datasets, conducting new analyses on these videos has remained prohibitively time-consuming due to a lack of computational tools for annotations.
 Instead, hundreds of hours of manual annotations have been required to analyze a fraction of the available frames for a given analysis.
-Thus, despite containing a wealth of information about the structure of parent-child interactions, these datasets have thus gone dramatically underused.
+Thus despite containing a wealth of information about the structure of parent-child interactions, these datasets have thus gone dramatically underused.
 
 Recent innovations in computer vision hold promise for automated annotations of the infant view.
-Over the past decade, deep neural networks have become dramatically better at a wide range of visual tasks, including object classification, @simonyan2014very], scene categorization [@zhou2017places], and pose detection [@zhang2016], arguably facilitating our understanding of visual perception [@peterson2018evaluating, @vanrullen2017perception] and improving computational neuroscience [@kietzmann2018deep].
+Over the past decade, deep neural networks have become dramatically better at a wide range of visual tasks, including object classification [@simonyan2014very] scene categorization [@zhou2017places], and pose detection [@zhang2016], arguably facilitating our understanding of visual perception [@peterson2018evaluating; @vanrullen2017perception] and improving computational neuroscience [@kietzmann2018deep].
 However, as most models have been trained on photographs or videos taken from the adult perspective, it is still unclear how easily these models can be applied to videos taken from the infant perspective.
 
-In this paper, we thus first validate a model for the automated detection of faces and hands in the infant view, using a open-source model of pose detection -- OpenPose[@zhang2016; @cao2017realtime] -- that provides pose, face, and hand keypoints for every person in an image.
-We compare the detection accuracy of OpenPose with that of both older and more specialized models of face detection and demonstrate the usability of this off-the-shelf model for quantifying the faces and hands in the infant view.
-
-We then apply this automated method to a head-mounted camera dataset collected in our lab to directly examine the role of postural developments on the social information in the infant view during language learning.
-Specifically, we constructed a dataset where infants and caregivers played freely using a set of novel and familiar objects; 8, 12, and 16 months-olds (N=36; N=12 in each age group) participated to ensure a wide range of locomotor abilities.
-Infants’ in-the-moment posture and orientation relative to their caregiver were hand annotated and all play sessions were transcribed.
-This cross-sectional design thus allowed us to directly examine the relative contributions of age vs.
-postural developments on children’s visual access to social information, and how the avaliability of social cues changes relative to naming events (e.g., “Yes, you like the [ball!]”).
+In the present paper, we make progress on understanding the relationship  between the infant visual perspective and motoric and linguistic development in three ways.
+First, we validate a model for the automated detection of faces and hands in the infant view.
+We use an open-source model of pose detection -- OpenPose [@zhang2016; @cao2017realtime] -- that provides pose, face, and hand keypoints for every person in an image.
+We compare the detection accuracy of this method with that of both older and more specialized models of face detection, and demonstrate the usability of this off-the-shelf model for quantifying the faces and hands in the infant view.
+
+Second, we provide an annotated, open dataset for researchers to examine the effects of postural developments and naming behavior during naturalistic parent-child interactions.
+To do so, we invited caregivers of 8, 12, and 16-month-olds to participate in play sessions where they were provided with pairs of novel and familiar objects (e.g., a ball and a microfiber duster, called a "zem").
+We chose this age-range to span the months when infants typically transition from sitting to crawling to standing, and documented the infant egocentric view as well as a third-person view of the play session.
+We then annotated infants' in-the-moment posture and orientation relative to their caregiver, and obtained full transcripts of the play sessions as well as MacArthur-Bates Communicative Development Inventories for each child.
+
+Third, we apply our automated method to this and one other openly available head-mounted camera dataset, examining the generality of the effects of motor development on the social information in view.
+In Study 1, we use the cross-sectional design of our dataset to directly examine the relative contributions of age vs.
+in-the-moment posture on infants' visual access to social information, and use the transcriptions to examine how the availability of social cues changes relative to naming events (e.g., “Yes, you like the [ball!]”).
+Indeed, despite positing links between the social information in view and language development, no work to date has directly examined how the availability of social information changes around naming events in naturalistic contexts.
+In Study 2, we apply this automated method to @franchak2017see, where one-year-olds wore head-mounted eye-tracking cameras during a play session and their in-the-moment posture was hand-annotated.
+Unlike in Study 1, infants and caregivers roamed a large, open playroom and explored different toys placed throughout.
+We thus analyze this second dataset with a first goal of  validating our automated method on a very different kind of video dataset, attempting to replicate their primary findings originally obtained with a head-mounted eye-tracker, and then further examine whether the proportion of hands in view also varies with infants’ posture.
 Broadly, we predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces/hands because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape with greater access to the social information in their environment.
 
-To examine the robustness of our automated method and the generalizability of our findings, we applied OpenPose [@cao2017realtime] to the dataset collected in @franchak2017see, where one-year-olds wore head-mounted eye-tracking cameras during a play session and their in-the-moment posture was hand-annotated.
-Unlike the present dataset, infants and caregivers roamed a large, open playroom and explored different toys placed throughout.
-We thus analyzed this dataset with the goal of both (1) validating our automated method on a very different kind of dataset, attempting to replicate their primary findings originally obtained with a head-mounted eye-tracker  (2) examining whether the proportion of hands in view also varies with infants’ posture in an exploratory play session.
-
-# Methods
-## Free-play video dataset  
-A primary goal of this project was to provide a annotated, open dataset for other researchers to examine the effects of postural developments and labeling behavior during naturalistic parent-child interactions.
-We thus invited caregivers of 8, 12, and 16-month-olds to participate in play sessions where they were provided with pairs of novel and familiar objects (e.g., a ball and a microfiber duster, called a "zem") in a small playroom in a lab (approximately 10' x 10').
-Infants wore head-mounted cameras equipped with a fish-eye lens (see Head-mounted camera), a tripod-mounted camera captured a third-person view of the play session.
+# Study 1
+## Methods
+Caregivers of 8, 12, and 16-month-olds were invited to participate in play sessions where they were provided with pairs of novel and familiar objects (e.g., a ball and a microfiber duster, called a "zem") in a small playroom in a lab (approximately 10' x 10').
+Infants wore head-mounted cameras equipped with a fish-eye lens (see Head-mounted camera); a tripod-mounted camera captured a third-person view of the play session.
 Using these video data, infants' posture and orientation to their caregiver were hand-coded and annotated for the entirety of the play session.
-All videos were transcribed, and MacArthur CDIs were collected for all children who particpated.
-All materials have been made publicly avaliable on Databrary for whom the parents provided sharing consent (29/36 dyads) via https://nyu.databrary.org/volume/101
+All videos were transcribed, and MacArthur CDIs were collected for all children who participated.
+All materials have been made publicly available on Databrary for whom the parents provided sharing consent (29/36 dyads) via [BLINDED].
 
@@
 
-all_subs %>%
+sub_summary <- all_subs %>%
   left_join(inc_subs) %>%
-  filter(inc==TRUE) %>%
   group_by(age_group) %>%
@@
             female = sum(gender=="female", na.rm=TRUE))
-  # xtable %>%
-  # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)
-
-# <!-- \caption{\label{tab:pop} Demographics by age group.} -->
+
+colnames(sub_summary) <- c('Group','N','% incl.','Avg. age', 'Avg video length', 'Num female')
+
 ```
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrr}
-\hline
-Group & N & \% incl. & Avg age & Avg video length (min) \\
-\hline
-8 mo. &   12 & 0.46 & 8.71 & 14.41 \\
-12 mo. &  12 & 0.40 & 12.62 & 12.71 \\
-16 mo. &  12 & 0.31 & 16.29 & 15.10\\
-\hline
-\end{tabular}
-\caption{Exclusion rates and summary demographics for the infants included in the study.} 
-\end{table}
+```{r, echo=F}
+apa_table(sub_summary, caption = "Exclusion rates and summary demographics for the infants included in the study.")
+```
 
@@
 Videos captured by MD-80/Veho cameras  were 640x480/720x480 pixels, respectively, and both cameras had a frame rate of ~30 frames per second.
-To ensure taht detections across these different cameras were comparable, we excluded detections from the outer edges of the videos taken with the MD-80 cameras based on these view angle differences in our main analyses (i.e.
+To ensure that detections across these different cameras were comparable, we excluded detections from the outer edges of the videos taken with the MD-80 cameras based on these view angle differences in our main analyses (i.e.
 excluding top/bottom 13% of the frames, and left/right 10% of the frames).
 
-[^2]: Detailed instructions for creating this headcam can be found  at \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
+[^2]: Detailed instructions for creating this headcam can be found at [BLINDED].
 
@@
 As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field).
+We note that these limitations regarding field of view and camera angle affect all studies using this method, not only our own.
 
@@
 ## "faceOPStrictAll", "wristOPStrictAll" == faces and wrists across all areas of the frames  with high confidence
-
+```
+
+```{r}
 ## make long form version for plotting both face detections on same graph
@@
 Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations; similarily, times when the child was being carried or the caregivers were out of the frame were marked as uncodable for caregiver orientation.
-On average, posture or orientation was uncodable from 1-2 minutes of data in each child (seconds excluded from analysis for posture, M = `r round(NA_rates$mean_na_posture,2)`s, SD = `r round(NA_rates$sd_na_posture,2)`s; orientation; M = `r round(NA_rates$mean_na_posture,2)`s, SD = `r round(NA_rates$sd_na_posture,2)`s), and these rates did not vary substantially with the age of the child.
-To assess the reliability of these annotations, a second coder annotated videos from three different children to calculate Cohen's kappa (posture, irr = .82; caregiver orientation, irr = .73).
+On average, posture or orientation was uncodable from 1-2 minutes of data in each child (seconds excluded from analysis for posture, M = `r round(NA_rates$mean_na_posture,0)`s, SD = `r round(NA_rates$sd_na_posture,0)`s; orientation; M = `r round(NA_rates$mean_na_posture,0)`s, SD = `r round(NA_rates$sd_na_posture,0)`s), and these rates did not vary substantially with the age of the child.
+To assess the reliability of these annotations, a second coder annotated videos from three different children to calculate Cohen's kappa (posture, $\kappa$ = .82; caregiver orientation, $\kappa$ = .73).
 
@@
 ## Face and hand detection
-We evaluated three detection systems for the ability to measure infants' visual access to faces.
+We evaluated three automated detection systems for the ability to measure infants' visual access to faces.
 The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones  [@viola2004robust].
@@
 Viola Jones, the first face detection system, made use of a series of Haar feature-based cascade classifiers [@viola2004robust] applied to each individual frame.
-The second algorithm (based on work by @zhang2016) uses multi-task cascaded convolutional neural networks (MTCNNs) for joint face detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
+The second algorithm [based on work by @zhang2016] uses multi-task cascaded convolutional neural networks (MTCNNs) for joint face detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present.
 We used a Tensorflow implementation of this algorithm available at https://github.com/davidsandberg/facenet.
@@
 
-# Results
+```{r exampledetections, fig.env = "figure", out.width="100%", fig.pos = "H", fig.label="exampledetections", fig.cap = "Example detections made by OpenPose from children in each age group."}
+knitr::include_graphics("images/exampe_detections.png")
+```
+
+## Results
 First, we report the accuracy of the automated detectors, as assessed by comparison to hand-labelled frames from the free-play video dataset described above.
-We then apply one of these automated detectors (OpenPose) to the entirety of this video dataset, and use these outputs to examine how postural developments influence children's visual access to faces and hands from 8-16 months of age, as well as how access to these social cues changes during labeling events (e.g., do you see the [zem]?).
-Finally, we apply this same automated detector to another video dataset [@franchak2017see], replicating and extending their findings on the effects of posture on visual access to social information.
-
-```{r exampledetections, fig.env = "figure", out.width="100%", fig.pos = "H", fig.label="exampledetections", fig.cap = "Example detections made by OpenPose from children in each age group."}
-
-knitr::include_graphics("images/exampe_detections.png")
-
-```
+We then apply one of these automated detectors (OpenPose) to the entirety of this video dataset, and use these outputs to examine how postural developments influence children's visual access to faces and hands from 8-16 months of age.
+We further use the detections to examine how access to these social cues changes during naming events (e.g., do you see the [zem]?).
 
@@
 
+```
+
+```{r}
 ## Only keep what we're analyzing
@@
   filter(detectorType %in% c('OpenPose-Faces','OpenPose-Wrists'))
-
-```
-
-## Accuracy of automated detections 
-For face detection, we found that both OpenPose and MTCNN dramatically outperformed ViolaJones, our baseline model, especially with respect to the random sample, where ViolaJones missed many face that were in view.
+```
+
+```{r}
+## Clean up and output tables
+colnames(perfWristsbySampleType) <- c('Algorithm','Sample Type','P','R','F')
+
+colnames(perfFacessbySampleType) <- c('Algorithm','Sample Type','P','R','F')
+
+## rename variables for interpretability in table
+levels(perfFacessbySampleType$Algorithm) <- c("MTCNN-Faces","OpenPose-Faces-NoFilter","OpenPose-Faces", "OpenPose-Faces-Strict-NoFilter","OpenPose-Faces-Strict","ViolaJones-Faces","OpenPose-Wrists","OpenPose-Wrists-All", "OpenPose-Wrists-Strict",  "OpenPose-Wrists-Strict-All")
+
+apa_table(perfFacessbySampleType,  caption = "Detector performance for faces in high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected).
+P, R, and F denote precision, recall, and F-score, respectively.
+\'Strict\' denotes when only  high confidence detections are considered.")
+```
+
+```{r}
+apa_table(perfWristsbySampleType,  caption = "Detector performance for wrists in both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected).
+P, R, and F denote precision, recall, and F-score, respectively.
+\'Strict\' denotes when only  high confidence detections are considered.")
+```
+
+```{r failures, fig.env = "figure", fig.pos = "h", fig.label="headcam", fig.width = 2, fig.cap = "Example failed detections from OpenPose, showing both false positives and false negatives (i.e.
+missed detections)."}
+knitr::include_graphics("images/example_failures.png")
+```
+
+### Accuracy of automated detections 
+For face detection, we found that both OpenPose and MTCNN dramatically outperformed ViolaJones (our baseline model) especially with respect to the random sample, where ViolaJones missed many face that were in view (see Table 2).
 When considering only the composite F-score across all frames, MTCNN slightly outperformed OpenPose  (`r round(perfFacesOverall$f[1],2)` MTCNN vs.
@@
 In other words, while OpenPose generated slightly more false positives than MTCNN, MTCNN missed several faces that were accurately detected by OpenPose.
-When we restricted our analysis to high-confidence detections from OpenPose (>.5 confidence; default threshold for visualization), we found very high precision (P = `r round(perfFacesOverall$p[3],2)`), but much lower recall (R = `r round(perfFacesOverall$r[3],2)`) and thus overall performance (F = `r round(perfFacesOverall$f[3],2)`), indicating that these low-confidence detections often indexed actual faces that were in the infant view.
+When we restricted our analysis to high-confidence detections from OpenPose (>.5 confidence; default threshold for visualization), we found very high precision (P = `r round(perfFacesOverall$p[3],2)`), but much lower recall (R = `r round(perfFacesOverall$r[3],2)`) and thus overall lower performance (F = `r round(perfFacesOverall$f[3],2)`), indicating that these low-confidence detections often indexed actual faces that were in the infant view.
 Figure \ref{fig:exampledetections} shows example successful detections from OpenPose in each age group, and Figure \ref{fig:failures} shows examples of missed faces as well as false positives for context.
@@
 Despite the fact that hand detection is a more computationally challenging problem [@bambach2017], and the fact that we used wrist keypoints as a proxy for hands,  OpenPose performed moderately well as a hand detector (F = `r round(perfWristsOverall$f[1],2)`).
-OpenPose achieved relatively high precision -- generating relatively few false positives -- but showed low recall on the randomly sampled frames (see Table 1).
-As with face detections, when we restricted our analysis to high-confidence detections, we found much higher precision (P = `r round(perfWristsOverall$p[2],2)`), but much lower recall (R = `r round(perfWristsOverall$r[2],2)`) and thus overall performance (F = `r round(perfWristsOverall$f[2],2)`).
+OpenPose achieved relatively high precision -- generating relatively few false positives -- but showed low recall on the randomly sampled frames (see Table 3).
+As with face detections, when we restricted our analysis to high-confidence detections, we found much higher precision (P = `r round(perfWristsOverall$p[2],2)`), but much lower recall (R = `r round(perfWristsOverall$r[2],2)`) and thus lower overall performance (F = `r round(perfWristsOverall$f[2],2)`).
 
@@
 Going forward, we analyze face and wrist detections using all detections from OpenPose, with the caveat that we are likely underestimating the proportion of hands in the dataset given the lower recall for hand detections.
-
-```{r failures, fig.env = "figure", out.width="50%", fig.pos = "H", fig.label="headcam", fig.cap = "Example failed detections from OpenPose, showing both false positives and false negatives (i.e.
-missed detections)."}
-knitr::include_graphics("images/example_failures.png")
-```
 
@@
 ```
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{llrrr}
-  \hline
-Algorithm & Sample\ Type & P & R & F \\ 
-  \hline
-MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
-  MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
-  OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
-  OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
-  OpenPose-Faces-Strict & High density & 0.97 & 0.65 & 0.78 \\ 
-  OpenPose-Faces-Strict & Random & 0.97 & 0.62 & 0.76 \\ 
-  ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
-  ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
-  \hline
-  \hline
-  OpenPose-Wrists & High density & 0.66 & 0.96 & 0.78 \\ 
-  OpenPose-Wrists & Random & 0.88 & 0.29 & 0.43 \\ 
-  OpenPose-Wrists-Strict & High density & 0.95 & 0.44 & 0.60 \\ 
-  OpenPose-Wrists-Strict & Random & 1.00 & 0.10 & 0.18 \\
-   \hline
-\end{tabular}
-\caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected).
-P, R, and F denote precision, recall, and F-score, respectively.
-"Strict" denotes when we only considered high confidence detections.}
-\vspace{-1em}
-\end{table}
-
-## Study 1: Access to social information during postural developments
 
@@
 
-```{r posture, fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
+```{r posture, fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are omitted for visualization purposes."}
 cowplot::plot_grid(pos_plot, orient_plot, nrow=2)
@@
 
-```{r indivPosOrient,  fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times when infant was carried or when posture/orientation were not codable are ommitted for visualization purposes"}
+```{r indivPosOrient,  fig.env = "figure", fig.pos = "H", echo=FALSE, out.width = "\\textwidth", fig.width=8, fig.height=4, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times when infant was carried or when posture/orientation were not codable are omitted for visualization purposes."}
 
@@
 ### Developmental changes in infant posture and caregiver orientation
-We report developmental shifts in infants' posture and their orientation relative to their caregiver, consistent with previous literature [@adolph2006motor; @franchak2017see].
-The proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with infants' age.
-As children got older, their locomotor abilities allowed them to become more independent.
+Consistent with previous literature [@adolph2006motor], the proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with infants' age.
+As infants got older, their locomotor abilities allowed them to become more independent.
 Both 8-month-olds and 12-month-olds spent relatively equivalent amounts of time lying/crawling (i.e., 'prone') which was markedly decreased in the 16-month-olds, who spent most of their time sitting or standing (see Figure \ref{fig:posture}).
-We also observed changes in infants' orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions than did children at other ages (see Figure \ref{fig:posture}).
+We also observed changes in infants' orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions than did infants at other ages (see Figure \ref{fig:posture}).
 However, we also saw considerable variability across children: some infants spent almost their entire time sitting at a close distance from their caregiver, whereas others showed more considerable variability (see Figure \ref{fig:indivPosOrient}).
 
-### Access to faces and hands across age
+### Changes in access to faces and hands 
 First, we examined the proportion of face and hand detections as a function of infants' age without considering their posture (see Figure \ref{fig:detByAge}).
@@
 rownames(wristGLMOut$coefficients)<-cleanedNames_Int
-#
-```
-
-```{r eval=FALSE} 
 ###
-xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.")
-
-xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")
+# xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.")
+
+# xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")
+```
+
+```{r}
+apa_table(faceGLMOut$coefficients, "Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")
+```
+
+```{r}
+apa_table(wristGLMOut$coefficients, "Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.")
 ```
@@
 
-In contrast, infant's locomotor developments had a major effect on the faces and hands that were in the field of view.
-Infants' in-the-moment posture was a major factor both in how many faces and hands were in view during the play session, as was their orientation relative to their caregiver (see Figure \ref{fig:detByPosOrient}).
-This was quantified using two generalized linear mixed-effect models estimating the proportion of faces and hands that were in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture.
-A summary of the coefficients of the models can be found in Tables 2 and 3.
-
-The interaction between infants' posture and their caregiver's orientation had the most dramatic effect on the social information in view.
+In contrast, infant's locomotor developments had a major effect on the faces and hands that were in the field of view (see Figure \ref{fig:detByPosOrient}).
+Two generalized linear mixed-effect models were used to predict the proportion of faces and hands in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture (see all coefficients in Tables 4 and 5).
+In particular, the interaction between infants' posture and their caregiver's orientation had the most dramatic effect on the social information in view.
 When caregivers were behind their infants, supporting their infants' sitting or standing positions, infants saw fewer faces.
-When caregivers were relatively close to their infants, infants who were sitting or standing had more faces in view (Face detections; infant sitting and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4])`; infant standing and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4])`) than infants who were lying down/crawling (i.e.
+When caregivers were relatively close to their infants, infants who were sitting or standing had more faces in view (Face detections; infant sitting and caregiver (CG) close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4])`; infant standing and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4])`) than infants who were lying down/crawling (i.e.
 prone).
@@
 We directly examined the contributions of posture and orientation vs.
-age by fitting a reduced version of the full model [@nakagawa2013general] without their fixed effects (both models were run with the maximal random effects structure).
+age by fitting a reduced version of the full model [@nakagawa2013general] without their fixed effects (both models were run with the maximal random effects structure), and comparing model fits for each of these.
 The fixed effects in a model with only the age of the participants accounted for relatively little variance in the proportion of faces (marginal $R^2$ < `r round(faces_age_only[1,1],5)`) or hands in view  (marginal $R^2$ = `r round(wrists_age_only[1,1],3)`).
@@
 Our play session was designed to provide parents with opportunities to label objects -- both familiar and novel -- such that we could examine whether children saw different kinds of social information around naming events.
-In a set of exploratory analyses, we thus analyzed how face and hand detections changed during object labeling events relative to baseline, analyzing a four-second window around each labeling event (e.g., "Look at the [zem]!").
-Every utterance of one of the labelled objects (e.g., 'ball') was counted as a 'labeling event'; timestamps of the beginning of each word were hand-annotated and synchronized with the frame-by-frame detections.
+In a set of exploratory analyses, we thus analyzed how face and hand detections changed during object naming events relative to baseline.
+We analyzed a four-second window (+/- 2 seconds) each time a caregiver uttered a name for one of the objects (e.g., "Look at the [zem]!"); this time window was chosen in keeping with previous work suggesting that parents attend to their target referents during this time range [@trueswell2016perceiving].
+Every utterance of one of the objects (e.g., 'ball') was counted as a 'naming event'; timestamps of the beginning of each word were hand-annotated and synchronized with the frame-by-frame detections.
 
@@
 Overall, these exploratory results suggest that children may tend to see more hands around naming events.
-This finding is consistent with the possibility that caregivers may change how they interact with their infant when presenting them with objects [@gogate2000study; @gogate2006attention].
+This finding is consistent with the possibility that caregivers may change how they interact with their infant when presenting them with objects [@gogate2000study; @gogate2006attention] and that hands could play a key role in guiding infants' attention during dyadic interactions [@yu2017hand].
 For example, caregivers may tend to simultaneously name objects when demonstrating their affordances or simply when pointing to them.
-In turn, infants may be sensitive to these naming events and orient their attention towards their caregiver, consistent with other accounts positing infants' sensitivity to social cues in early word learning environments [@yurovsky2018communicative; @yurovsky2017beyond; @frank2013].
+In turn, infants may be sensitive to these naming events and orient their attention towards their caregiver, consistent with other accounts positing infants' sensitivity to social cues in early word learning environments [@yurovsky2018communicative; @yurovsky2017beyond].
 
@@
 
-## Study 2: Extension of Franchak et al., 2017
-In the present work, we found that infant's in-the-moment posture changed with their age, as did infants' orientation relative their caregiver.
-In a related study with 12-month-olds, @franchak2017see also found that infants' in-the-moment posture changed the proportion of time infants spent looking at faces.
+# Study 2
+In Study 1, we found that infant's in-the-moment posture changed with their age, as did infants' orientation relative their caregiver.
+In a related study with one-year-olds, @franchak2017see also found that infants' in-the-moment posture changed the proportion of time infants spent looking at faces.
 Here, we sought to replicate these findings using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras [hosted on Databrary; @simon2015databrary] for two reasons.
@@
 
-### Comparison between video datasets 
-First, the play environment in @franchak2017see that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to climb, explore, and interact with, and infants were unconstrained and allowed to freely wander the room.
-In contrast, the play room used in Study 1 was relatively small (approximately 10' x 10') and was set-up with focused play with the pairs of novel and familiar objects.
-Further, multiple people were present during the play session in @franchak2017see-- including their caregiver and two experimenters -- whereas in Study 1 the experimenters left the room during the play session.
-The view angle of the two head-mounted cameras used in Study 1 and @franchak2017see were relatively similar ($52.2^{\circ}$ horizontal by $42.2^{\circ}$ in @franchak2017see, $47^{\circ}$ horizontal by $36^{\circ}$ vertical in Study 1).
+## Methods
+
+### Participants
+With the aid of @franchak2017see, we obtained the scene camera footage from the head-mounted eye-trackers for the 17 one-year-old infants (range 11.8–12.4 months) who participated.
+As noted in @franchak2017see, families were recruited from maternity wards of local hospitals in the New York City metropolitan area
+and were predominantly white and middle class.
+
+### Head-mounted camera
+The view angle of the two head-mounted cameras used in these two studies were relatively similar ($52.2^{\circ}$ horizontal by $42.2^{\circ}$ in @franchak2017see, $47^{\circ}$ horizontal by $36^{\circ}$ vertical in Study 1).
+However, in Study 2 the camera was mounted on the side of the infants' head in Study 2, closer to infants temples, whereas in Study 1 the camera was situated in the middle of infants' forehead and oriented slightly upwards.
+
+### Procedure
+The play environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in Study 1, containing multiple structures and toys in different parts of the room for infants to climb, explore, and interact with, and infants were allowed to freely wander the room.
+In contrast, the play room used in Study 1 was relatively small (approximately 10' x 10') and was set-up for focused play on a mat with the pairs of novel and familiar objects.
+In addition, multiple people were present during the play session -- including their caregiver and two experimenters -- whereas in Study 1 the experimenters left the room during the play session.
+
+### Video annotations
+The first five minutes of each of the videos were coded for the infants' posture (upright, prone, or sitting) by trained coders in @franchak2017see.
+These frame-by-frame posture annotations were synced with the outputs of the same automated annotation used in Study 1.
 
@@
 
-### Replication using automated detections
-Overall, we found convergence between our two methodologies, replicating the main results from @franchak2017see, and finding that the proportion of faces detected was greater when infants were sitting or standing vs.
+## Results
+### Differences between eye-tracking vs. automated detections
+First, we compared the overall proportion of frames in which infants foveated faces -- as assessed by the head-mounted eye-tracker --  in @franchak2017see vs.
+the proportion of frames with faces detected by OpenPose.
+We expected some differences, as (1) head-mounted eye-trackers may underestimate the proportion of faces attended due to calibration issues, and (2) OpenPose may detect faces that are in view for infants but that infants may not be foveating.
+Across the entire session in @franchak2017see, infants looked at faces on 4.7% of frames.
+When we used all detections from OpenPose, we found a higher proportion of faces --  `r mean(detections_by_conf_by_sub$avg_faces_by_sub)*100`%.
+When we restricted our results to only high-confidence detections, we found `r mean(detections_by_conf_by_sub$avg_faces_by_sub_strict)*100`% of frames with faces, closer to the original values reported by @franchak2017see.
+However, the above analyses on the accuracy of this method suggest that high-confidence detections dramatically underestimate the number of faces in view.
+Thus, OpenPose may somewhat overestimate the proportion of faces that are actually foveated, while head-mounted eye-trackers may underestimate the proportion of faces that infants' could be attending to.
+
+### Replication & extension using automated detections
+Despite these differences, we found convergence between our two methodologies, replicating the main results from @franchak2017see, and finding that the proportion of faces detected was greater when infants were sitting or standing vs.
 prone (see Figure \ref{fig:franchak}).
 We found this result regardless of whether we used all detections (proportion of frames with face detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_faces[1])`, Sitting: M = `r mean(dets_by_conf_by_posture$avg_faces[2])`, Upright: M = `r mean(dets_by_conf_by_posture$avg_faces[3])`) or restricted our analyses to high-confidence detections (percentage of frames with face detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[1])`, Sitting: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[2])`, Upright: M = `r mean(dets_by_conf_by_posture$avg_faces_strict[3])`).
-These results were confirmed in generalized linear-mixed models, as in the first study (Sitting vs.
+These results were confirmed in generalized linear-mixed models, with random intercepts for each subjects and infants' posture as a fixed effect (Sitting vs.
 Prone, b = `r round(franchak_faces_out$coefficients[2,1],3)`, SE  = `r round(franchak_faces_out$coefficients[2,2],3)`, Z = `r round(franchak_faces_out$coefficients[2,3],3)`, `r reportP(franchak_faces_out$coefficients[2,4])`.
@@
 
-The most noticeable difference between our results is the overall proportion of frames with faces detected by OpenPose vs.
-the proportion of frames in which infants were judged to be looking at faces in @franchak2017see.
-Across the entire session in @franchak2017see, infants looked at faces on 4.7% of frames [@franchak2017see].
-When we used all detections from OpenPose, we found a higher proportion of faces --  `r mean(detections_by_conf_by_sub$avg_faces_by_sub)*100`%.
-When we restricted our results to only high-confidence detections, we found `r mean(detections_by_conf_by_sub$avg_faces_by_sub_strict)*100`% of frames with faces, closer to the original values reported by @franchak2017see.
-However, the above analyses on the accuracy of this method suggest that high-confidence detections dramatically underestimate the number of faces in view.
-Thus, we suspect that these differences between our findings are due to both (1) the fact head-mounted eye-trackers may sometimes underestimate the proportion of faces attended due to calibration issues, and that (2) OpenPose may sometimes detect faces that are in view for infants but that infants may not be foveating.
-
-We also found that infants' in-the-moment posture also modulated the proportion of hands that were in view (i.e., wrist detections), though these were not originally annotated in @franchak2017see (see Figure \ref{fig:franchak}, proportion of frames with wrist detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_hands[1])`; Sitting: M = `r mean(dets_by_conf_by_posture$avg_hands[2])`; Upright:  M = `r mean(dets_by_conf_by_posture$avg_hands[3])`).
-These results were confirmed in generalized linear-mixed models, as in the first study (Sitting vs.
+We also found that infants' in-the-moment posture modulated the proportion of hands that were in view (i.e., wrist detections), though these were not originally annotated in @franchak2017see (see Figure \ref{fig:franchak}, proportion of frames with wrist detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_hands[1])`; Sitting: M = `r mean(dets_by_conf_by_posture$avg_hands[2])`; Upright:  M = `r mean(dets_by_conf_by_posture$avg_hands[3])`).
+These results were confirmed in generalized linear-mixed models with the same model structure as with faces (Sitting vs.
 Prone, b = `r round(franchak_wrists_out$coefficients[2,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[2,2],3)`, Z = `r round(franchak_wrists_out$coefficients[2,3],4)`, `r reportP(franchak_wrists_out$coefficients[2,4])`.
@@
 # General Discussion
-What social cues do infants see as they learn language, and how does this change as they grow and start to locomote themselves?
-We explored this question using video data from head-mounted cameras from two naturalistic datasets of parent-child interactions: a cross-sectional database of play sessions from 8-16 month-olds with sets of novel and familiar toys, and database of head-mounted camera videos from one-year-olds who explored a large play area from @franchak2017see.
+What social cues do infants see as they learn language, and how does infants' access to these cues change as they grow and start to locomote themselves?
+We explored this question using video data from head-mounted cameras from two naturalistic datasets of parent-child interactions: a cross-sectional database of play sessions from 8-16 month-olds with sets of novel and familiar toys, and database of head-mounted camera videos from one-year-olds who explored a large play area [@franchak2017see].
 To analyze these datasets, we developed a novel method using a pose detection model to automate the annotation of the social information in the infant view, here operationalized as the presence of the faces and hands of their caregiver; these annotations were then synced with manual annotations of infants in-the-moment posture from third-person videos.
 
-Despite being trained on the adult perspective, the pose detector we used (OpenPose, @cao2017realtime) was able to generalize relatively well to the infant viewpoint, achieving comparable precision and accuracy as a face detector relative to a state-of-the-art model optimized specifically for detecting faces in natural scenes (MTCNN, @zhang2016).
+Despite being trained on the adult perspective, the pose detector we used [OpenPose, @cao2017realtime] was able to generalize relatively well to the infant viewpoint, achieving comparable precision and accuracy as a face detector relative to a state-of-the-art model optimized specifically for detecting faces in natural scenes [@zhang2016].
 While OpenPose had relatively low recall as a hand detector -- missing some hands that were in the infant view -- it made comparable rates of false alarms.
@@
 Thus, while imperfect, we suggest that OpenPose can be applied to infant egocentric videos for the extraction of the social information in the infant viewpoint, reducing the burden of manual annotations and promoting the reusability of rich video datasets for further analyses.
-The use of this automated methodology allowed us to easily annotate the entirety of our dataset -- additionally analyzing the social information around naming events -- and to re-analyze the data from @franchak2017see to replicate our findings in a very different kind of play session.
+The use of this automated methodology allowed us to easily annotate the entirety of our dataset -- additionally analyzing the social information around naming events -- and to re-analyze the data from @franchak2017see, replicating our findings in a very different kind of play session.
+Furthermore, future work may be able to fine-tune pose detectors for even better accuracy, leveraging human annotations of the faces and hands that infants see to adapt models for the infant view.
 
@@
 As children grow and change, the activities in which they engage with their caregivers are likely to also vary, leading to differences in the distribution of social cues that they experience that may not be captured.
-Finally, locomotive abilities are of course only part of a cascade of changes in infants' abilities and experiences, and these analyses document only a fraction of this broader, multifaceted trajectory in a population of children from primarily WEIRD contexts -- white, educated, industrialized, rich, and democratic [@jones2010weird].
+Finally, locomotive abilities are of course only part of a cascade of changes in infants' abilities and experiences, and these analyses document only a fraction of this broader, multifaceted trajectory in a population of children from primarily WEIRD contexts -- white, educated, industrialized, rich, and democratic [@henrich2010most].
+Parenting practices with respect to motor development can and do vary widely across cultures [@karasik2018ties] -- and these choices likely influence the social cues that children see and how they use them.
 
@@
 # Acknowledgements
-Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in recruitment, data collection, and annotation.
-This work was funded by a Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF, and NSF #1714726 to BLL.
-An earlier version of this work was presented to the Cognitive Science Society in @frank2013 and @sanchez2018postural.
+[BLINDED]
 

### 2020-08-24 writeup-paper/draft_apa_blinded.tex
latest paper draft

--- 2020-08-14 writeup-paper/draft_apa_july2020.tex
@@
 Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
-\keywords{Postural developments, headcam, social information, face detection, pose detection\newline\indent Word count: XXX}
+\keywords{Postural developments, headcam, social information, face detection, pose detection\newline\indent Word count: 7324}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \abstract{
+•We provided a open, annotated dataset of naturalistic parent-child
+interactions during a play session with with novel and familiar toys,
+recording the egocentric and third-person views.
+
+•We develop a novel method using a pose detection model to automatically
+quantify the faces and hands in the infant view.
+
+•We apply our novel method to this and another egoecentric video
+dataset, and confirm that locomotive developments around the first
+birthday shape the faces and hands that infants see in naturalistic
+interactions.
+
+•The combined use of head-mounted cameras and automated analyses has the
+potential to reveal the changing statistics of the information in the
+infant view.
+
 Social information plays a key role in infants' early linguistic and
@@
 and their caregivers with sets of novel and familiar toys (N=36, 8-16
-months of age), recording both egocentric and third-person video. We
-then developed an automated analysis method using a pose detection model
-(Zhang et al., 2017) to detect the faces and hands in the infant view,
-allowing us to analyze the entirety of this dataset. We found that
-infants' posture and orientation to their caregiver changed dramatically
-across this age range and modulated infants access to social
-information; infants who were sitting or standing with their caregiver
-at a close distance tended to have the most faces and hands in their
-visual field. We also applied our automated method to a second
-egocentric video dataset where one-year-olds explored a large play room
-with their caregivers (Franchak et al., 2017), replicating and extending
-their findings. Together, these results establish the validity of a new
-method for detecting social information and confirm that motoric
-developments play a significant role in shaping the social information
-that infants have access to. We propose that the combined use of
-head-mounted cameras and the application of new computer vision
-techniques is a promising avenue for re-utilizing rich video datasets
-and understanding how the statistics of infants' visual and linguistic
-experience change over development.
+month-olds), recording both egocentric and third-person video. We then
+developed an automated analysis method using a pose detection model to
+detect the faces and hands in the infant view. We found that infants'
+posture and orientation to their caregiver changed dramatically across
+this age range and modulated infants access to social information;
+infants who were sitting or standing with their caregiver at a close
+distance tended to have the most faces and hands in their visual field.
+We applied this automated method to a second egocentric video dataset
+from a previous study, replicating and extending their findings on the
+influence of infants' posture on the social information in view.
+Together, these results establish the validity of a new method for
+detecting social information in egocentric videos and confirm that
+motoric developments play a significant role in shaping the social
+information that infants have access to. The combined use of
+head-mounted cameras and new computer vision techniques is a promising
+avenue for understanding how the statistics of infants' visual and
+linguistic experience change over development.
 }
@@
 tend to follow (Yu \& Smith, 2013, 2017) and imitate the gestures of
-their caregivers (e.g., pointing).
-
-Infants ability to process these social cues may provide strong
-scaffolding for early word learning. Longitudinal studies provide some
-evidence for this link: children's level of joint engagement with their
-mother at 9-12 months predicts both their receptive and productive
-vocabularies (Carpenter, Nagell, \& Tomasello, 1998) and 10 month-olds
-who follow an adult's gaze (in an experimental context) have larger
-vocabularies at 18 months and throughout the second year of life (Brooks
-\& Meltzoff, 2005, 2008). While the relationship between hand-following
-and language development has been less well characterized, infants who
-follow their caregivers' hands tend to be those who spend more time
-jointly attending to events with their caregivers (Yu \& Smith, 2017).
+their caregivers (e.g., pointing). Infants ability to process these
+social cues may provide strong scaffolding for early word learning.
+Longitudinal studies provide some evidence for this link: children's
+level of joint engagement with their mother at 9-12 months predicts both
+their receptive and productive vocabularies (Carpenter, Nagell, \&
+Tomasello, 1998) and 10 month-olds who follow an adult's gaze (in an
+experimental context) have larger vocabularies at 18 months and
+throughout the second year of life (Brooks \& Meltzoff, 2005, 2008).
+While the relationship between hand-following and language development
+has been less well characterized, infants who follow their caregivers'
+hands tend to be those who spend more time jointly attending to events
+with their caregivers (Yu \& Smith, 2017).
 
 Relatively little work, however, has quantified how often infants see
-and use these kinds of social cues in naturalistic learing environments.
-By using head-mounted cameras to record what infants see, researchers
-have begun to document the infant egocentric perspective (Yoshida \&
-Smith, 2008) and to quantify the information -- social and otherwise --
-available to infants as they learn. While head-mounted camera data do
-not provide explicit information on where infants are attending, some
-work suggests that infants orient their head towards what they are
-focusing on -- putting those people or objects in view (Yoshida \&
-Smith, 2008). Initial recordings using this technique during in-lab play
-sessions revealed a different view than many imagined: instead of being
-dominated by faces, the infant perspective contained close up views of
-primarily toys and hands (Franchak, Kretch, Soska, \& Adolph, 2011;
-Yoshida \& Smith, 2008; Yu \& Smith, 2017). Subsequent research has
-revealed that the infant view is far from a unitary construct,
-undergoing dramatic changes as infants grow. Recordings from home
-environments suggest that the viewpoints of very young infants -- less
-than 4 months of age -- do indeed contain persistent and frequent views
-of faces (Fausey, Jayaraman, \& Smith, 2016; Jayaraman, Fausey, \&
-Smith, 2017) but that the infant view tends to contain more and more
-hands as infants grow older.
-
-These changes in perspective are likely the downstream consequence of a
+and use these kinds of social cues in naturalistic learning
+environments. By using head-mounted cameras to record what infants see,
+researchers have begun to document the infant egocentric perspective
+(Yoshida \& Smith, 2008) and to quantify the information -- social and
+otherwise -- available to infants as they learn. While head-mounted
+camera data do not provide explicit information on where infants are
+attending, some work suggests that infants orient their head towards
+what they are focusing on -- putting those people or objects in view
+(Yoshida \& Smith, 2008). Initial recordings using this technique during
+in-lab play sessions revealed a different view than many imagined:
+instead of being dominated by faces, the infant perspective contained
+close up views of primarily toys and hands (Franchak, Kretch, Soska, \&
+Adolph, 2011; Yoshida \& Smith, 2008; Yu \& Smith, 2017). Subsequent
+research has revealed that the infant view undergoes dramatic changes as
+infants grow. Recordings from home environments suggest that the
+viewpoints of very young infants -- less than 4 months of age -- do
+indeed contain persistent and frequent views of faces (Fausey,
+Jayaraman, \& Smith, 2016; Jayaraman, Fausey, \& Smith, 2017) but that
+the infant view tends to contain more and more hands as infants grow
+older.
+
+These changes in perspective are likely the downstream consequence of
 myriad factors, chief of which may be infants' evolving locomotive
@@
 an infant can lengthen up to 2cm in 24 hours (Lampl, 1993), leading to
-changes in their physical abilities from one day to the next. These
-motoric developments can be thought of as gateways that open up entirely
-new phases of development (Iverson, 2010), and cause a cascade of
-changes in infants ability to interact with their world and the people
-in it (Karasik, Tamis-LeMonda, \& Adolph, 2014).
-
-Thus, infants' changing locomotor abilities may shape the social cues
-that infants' see and seek out, in turn impacting their cognitive and
-linguistic abilities. Supporting this idea, infants experience with
-sitting predicts their success at 3D object completion tasks (Soska,
-Adolph, \& Johnson, 2010) as well as their receptive vocabulary
-(Libertus \& Violi, 2016). And as children begin crawling (Adolph,
-Vereijken, \& Denny, 1998) -- or scooting or cruising (Patrick, Noah, \&
-Yang, 2012) -- they are no longer constrained to the same spot that
-their caregivers last placed them in. Yet while crawlers can choose
-where to go and what they see to a much greater degree, they also appear
-to spend much of their time in a world populated by floors and knees;
-during spontaneous play, toddlers are more likely to look at the floor
-while crawling than while walking (Franchak et al., 2011), when they
-have full visual access to their environment and the people in it
-(Kretch, Franchak, \& Adolph, 2014).
-
-On one view, it is children's ability to stand and walk that
+changes in their abilities from one day to the next. These motoric
+developments can be thought of as gateways that open up entirely new
+phases of development (Iverson, 2010), causing a cascade of changes in
+an infant's ability to interact with their world and the people in it
+(Karasik, Tamis-LeMonda, \& Adolph, 2014). Thus, infants' changing
+locomotor abilities may shape the social cues that infants' see and seek
+out, in turn impacting their cognitive and linguistic abilities.
+
+However, understanding the nuanced relationship between the social
+information in the infant view and their motoric and linguistic
+development has remained challenging. For example, as children begin
+crawling (Adolph, Vereijken, \& Denny, 1998) -- or scooting or cruising
+(Patrick, Noah, \& Yang, 2012) -- they are no longer constrained to the
+same spot that their caregivers last placed them in. Yet while crawlers
+can choose where to go and what they see to a much greater degree, they
+also appear to spend much of their time in a world populated by floors
+and knees; during spontaneous play, toddlers are more likely to look at
+the floor while crawling than while walking (Franchak et al., 2011),
+when they have full visual access to their environment and the people in
+it (Kretch, Franchak, \& Adolph, 2014). Further, infants' experience
+with sitting predicts their success at 3D object completion tasks
+(Soska, Adolph, \& Johnson, 2010) as well as their receptive vocabulary
+(Libertus \& Violi, 2016), suggesting the importance of focused play
+sessions for language development.
+
+Others have posited that it is children's ability to stand and walk that
 fundamentally changes their ability to access social information (e.g.,
 facial expressions, gaze cues, pointing) relative to children who are
-still crawling and sitting (Walle, 2016). In turn, this ability to
-access more detailed social information could allow infants to learn
-words quicker and more efficiently, facilitating language growth (Walle,
-2016). Supporting this idea, walking infants make different kinds of
-object-related bids for attention from their caregivers than crawling
-infants, and tend to hear more action directed statements (e.g.,
-\enquote{open it}) (Karasik et al., 2014); further, in an observational
-study, Walle and Campos (2014) also found that children who were able to
-walk had both higher receptive and productive vocabularies. However, not
-all evidence supports this view: some parental report data suggest a
-weak relationship between walking and the onset of language (Moore,
-Dailey, Garrison, Amatuni, \& Bergelson, 2019), and head-mounted
-eye-tracking data have highlighted the role of infants' in-the-moment
-posture -- and that of their caregivers' -- in what infants' see,
-finding that these factors are better predictors of the social
-information in view than their locomotor status per se (i.e.,
-\enquote{walker} vs \enquote{crawler}) (Franchak, Kretch, \& Adolph,
-2017).
-
-Understanding this nuanced relationship between the social information
-in the infant view and their motoric and linguistic development has
-remained challenging. Indeed, different kinds of play sessions (e.g.,
-exploring novel environments vs.~playing with novel objects) might
-modulate the degree to which infants' posture changes the social
-information in view. Furthermore, no work to date has directly examined
-how the availability of social information changes in naturalistic
-language learning contexts. For example, it is unclear if and how the
-social information in view changes during the naming events for
-different objects (e.g, \enquote{Look, a {[}ball{]}}).
-
-More broadly, the field is in need of computational tools to reutilize
-rich video datasets and to understand the generalizability of findings
-across different populations, tasks, and age-ranges. While the field has
+still crawling and sitting (Walle, 2016), which in turn allow infants to
+learn words quicker and more efficiently (Walle, 2016). Supporting this
+idea, walking vs.~crawling infants tend to make different kinds of
+object-related bids for attention from their caregivers (Karasik et al.,
+2014), hear more action directed statements (e.g., \enquote{open it})
+(Karasik et al., 2014), and have higher receptive and productive
+vocabularies (Walle \& Campos, 2014). However, not all evidence supports
+this view: parental report data suggest a weak relationship between
+walking and the onset of language (Moore, Dailey, Garrison, Amatuni, \&
+Bergelson, 2019), and head-mounted eye-tracking data from one-year-olds
+highlights the role of infants' in-the-moment posture -- and that of
+their caregivers' -- in what infants' see, finding that these factors
+are better predictors of the social information in view than their
+locomotor status per se (i.e., \enquote{walker} vs \enquote{crawler})
+(Franchak, Kretch, \& Adolph, 2017).
+
+Broadly, the field is in need of computational tools to reuse these rich
+video datasets and understand the generalizability of findings across
+different populations, tasks, and age-ranges. For example, different
+kinds of play sessions (e.g., exploring novel environments vs.~playing
+with novel objects) might modulate the degree to which infants' posture
+changes the social information in view. Yet while the field has
 assembled many head-mounted camera datasets, conducting new analyses on
@@
 manual annotations have been required to analyze a fraction of the
-available frames for a given analysis. Thus, despite containing a wealth
+available frames for a given analysis. Thus despite containing a wealth
 of information about the structure of parent-child interactions, these
@@
 networks have become dramatically better at a wide range of visual
-tasks, including object classification, Simonyan and Zisserman
-(2014){]}, scene categorization (Zhou, Lapedriza, Khosla, Oliva, \&
-Torralba, 2017), and pose detection (K. Zhang, Zhang, Li, \& Qiao,
-2016), arguably facilitating our understanding of visual perception
-(Peterson, Abbott, \& Griffiths, 2018, VanRullen (2017)) and improving
-computational neuroscience (Kietzmann, McClure, \& Kriegeskorte, 2018).
-However, as most models have been trained on photographs or videos taken
-from the adult perspective, it is still unclear how easily these models
-can be applied to videos taken from the infant perspective.
-
-In this paper, we thus first validate a model for the automated
-detection of faces and hands in the infant view, using a open-source
-model of pose detection -- OpenPose(Cao, Simon, Wei, \& Sheikh, 2017; K.
-Zhang et al., 2016) -- that provides pose, face, and hand keypoints for
-every person in an image. We compare the detection accuracy of OpenPose
-with that of both older and more specialized models of face detection
-and demonstrate the usability of this off-the-shelf model for
+tasks, including object classification, Simonyan and Zisserman (2014),
+scene categorization (Zhou, Lapedriza, Khosla, Oliva, \& Torralba,
+2017), and pose detection (K. Zhang, Zhang, Li, \& Qiao, 2016), arguably
+facilitating our understanding of visual perception (Peterson, Abbott,
+\& Griffiths, 2018; VanRullen, 2017) and improving computational
+neuroscience (Kietzmann, McClure, \& Kriegeskorte, 2018). However, as
+most models have been trained on photographs or videos taken from the
+adult perspective, it is still unclear how easily these models can be
+applied to videos taken from the infant perspective.
+
+In the present paper, we make progress on understanding the relationship
+the between the infant visual perspective and motoric and linguistic
+development in three ways. First, we validate a model for the automated
+detection of faces and hands in the infant view. We use an open-source
+model of pose detection -- OpenPose (Cao, Simon, Wei, \& Sheikh, 2017;
+K. Zhang et al., 2016) -- that provides pose, face, and hand keypoints
+for every person in an image. We compare the detection accuracy of this
+method with that of both older and more specialized models of face
+detection, and demonstrate the usability of this off-the-shelf model for
 quantifying the faces and hands in the infant view.
 
-We then apply this automated method to a head-mounted camera dataset
-collected in our lab to directly examine the role of postural
-developments on the social information in the infant view during
-language learning. Specifically, we constructed a dataset where infants
-and caregivers played freely using a set of novel and familiar objects;
-8, 12, and 16 months-olds (N=36; N=12 in each age group) participated to
-ensure a wide range of locomotor abilities. Infants' in-the-moment
-posture and orientation relative to their caregiver were hand annotated
-and all play sessions were transcribed. This cross-sectional design thus
-allowed us to directly examine the relative contributions of age
-vs.~postural developments on children's visual access to social
-information, and how the avaliability of social cues changes relative to
-naming events (e.g., \enquote{Yes, you like the {[}ball!{]}}). Broadly,
-we predicted that there would be differential access to social
-information based on children's postural developments: crawling infants
-would see fewer faces/hands because they would primarily be looking at
-the ground, while walking toddlers would have access to a richer visual
-landscape with greater access to the social information in their
-environment.
-
-To examine the robustness of our automated method and the
-generalizability of our findings, we applied OpenPose (Cao et al., 2017)
-to the dataset collected in Franchak et al. (2017), where one-year-olds
-wore head-mounted eye-tracking cameras during a play session and their
-in-the-moment posture was hand-annotated. Unlike the present dataset,
-infants and caregivers roamed a large, open playroom and explored
-different toys placed throughout. We thus analyzed this dataset with the
-goal of both (1) validating our automated method on a very different
-kind of dataset, attempting to replicate their primary findings
-originally obtained with a head-mounted eye-tracker (2) examining
+Second, we provide an annotated, open dataset for researchers to examine
+the effects of postural developments and naming behavior during
+naturalistic parent-child interactions. To do so, we invited caregivers
+of 8, 12, and 16-month-olds to participate in play sessions where they
+were provided with pairs of novel and familiar objects (e.g., a ball and
+a microfiber duster, called a \enquote{zem}). We chose this age-range to
+span the months when infants typically transition from sitting to
+crawling to standing, and documented the infant egocentric view as well
+as a third-person view of the play session. We then annotated infants'
+in-the-moment posture and orientation relative to their caregiver, and
+obtained full transcripts of the play sessions as well as
+MacArthur-Bates Communicative Development Inventories for each child.
+
+Third, we apply our automated method to this and one other openly
+available head-mounted camera dataset, examining the generality of the
+effects of motor development on the social information in view. In Study
+1, we use the cross-sectional design of our dataset to directly examine
+the relative contributions of age vs.~in-the-moment posture on infants'
+visual access to social information, and use the transcriptions to
+examine how the availability of social cues changes relative to naming
+events (e.g., \enquote{Yes, you like the {[}ball!{]}}). Indeed, despite
+positing links between the social information in view and language
+development, no work to date has directly examined how the availability
+of social information changes around naming events in naturalistic
+contexts. In Study 2, we apply this automated method to Franchak et al.
+(2017), where one-year-olds wore head-mounted eye-tracking cameras
+during a play session and their in-the-moment posture was
+hand-annotated. Unlike the present dataset, infants and caregivers
+roamed a large, open playroom and explored different toys placed
+throughout. We thus analyzed this second dataset with a first goal of
+validating our automated method on a very different kind of video
+dataset, attempting to replicate their primary findings originally
+obtained with a head-mounted eye-tracker, and then further examine
 whether the proportion of hands in view also varies with infants'
-posture in an exploratory play session.
-
-\section{Methods}\label{methods}
-
-\subsection{Free-play video dataset}\label{free-play-video-dataset}
-
-A primary goal of this project was to provide a annotated, open dataset
-for other researchers to examine the effects of postural developments
-and labeling behavior during naturalistic parent-child interactions. We
-thus invited caregivers of 8, 12, and 16-month-olds to participate in
+posture. Broadly, we predicted that there would be differential access
+to social information based on children's postural developments:
+crawling infants would see fewer faces/hands because they would
+primarily be looking at the ground, while walking toddlers would have
+access to a richer visual landscape with greater access to the social
+information in their environment.
+
+\section{Study 1}\label{study-1}
+
+\subsection{Methods}\label{methods}
+
+Caregivers of 8, 12, and 16-month-olds were invited to participate in
 play sessions where they were provided with pairs of novel and familiar
@@
 play session. All videos were transcribed, and MacArthur CDIs were
-collected for all children who particpated. All materials have been made
-publicly avaliable on Databrary for whom the parents provided sharing
-consent (29/36 dyads) via \url{https://nyu.databrary.org/volume/101}
+collected for all children who participated. All materials have been
+made publicly available on Databrary for whom the parents provided
+sharing consent (29/36 dyads) via {[}BLINDED{]}.
 
@@
 from a small camera attached to a soft elastic headband.\footnote{Detailed
-  instructions for creating this headcam can be found at
-  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
+  instructions for creating this headcam can be found at {[}BLINDED{]}.}
 Initial participants wore an MD-80 camera, which was then replaced by a
@@
 respectively, and both cameras had a frame rate of \textasciitilde{}30
-frames per second. To ensure taht detections across these different
+frames per second. To ensure that detections across these different
 cameras were comparable, we excluded detections from the outer edges of
@@
 decision limited our ability to detect hands (especially those of the
-child, which are typically found at the bottom of the visual field).
+child, which are typically found at the bottom of the visual field). We
+note that these limitations regarding field of view and camera angle
+affect all studies using this method, not only our own.
 
@@
 minutes of data in each child (seconds excluded from analysis for
-posture, M = 104.99s, SD = 233.86s; orientation; M = 104.99s, SD =
-233.86s), and these rates did not vary substantially with the age of the
-child. To assess the reliability of these annotations, a second coder
-annotated videos from three different children to calculate Cohen's
-kappa (posture, irr = .82; caregiver orientation, irr = .73).
+posture, M = 105s, SD = 234s; orientation; M = 105s, SD = 234s), and
+these rates did not vary substantially with the age of the child. To
+assess the reliability of these annotations, a second coder annotated
+videos from three different children to calculate Cohen's kappa
+(posture, \(\kappa\) = .82; caregiver orientation, \(\kappa\) = .73).
 
@@
 each individual frame. The second algorithm (based on work by K. Zhang
-et al. (2016)) uses multi-task cascaded convolutional neural networks
+et al., 2016) uses multi-task cascaded convolutional neural networks
 (MTCNNs) for joint face detection and alignment, built to perform well
@@
 
-\section{Results}\label{results}
+\subsection{Results}\label{results}
 
@@
 to examine how postural developments influence children's visual access
-to faces and hands from 8-16 months of age, as well as how access to
-these social cues changes during labeling events (e.g., do you see the
-{[}zem{]}?). Finally, we apply this same automated detector to another
-video dataset (Franchak et al., 2017), replicating and extending their
-findings on the effects of posture on visual access to social
-information.
+to faces and hands from 8-16 months of age. We further use the
+detections to examine how access to these social cues changes during
+naming events (e.g., do you see the {[}zem{]}?). Finally, we apply this
+same automated detector to another video dataset (Franchak et al.,
+2017), replicating and extending their findings on the effects of
+posture on visual access to social information.
 
@@
 
-\subsection{Accuracy of automated
+\subsubsection{Accuracy of automated
 detections}\label{accuracy-of-automated-detections}
@@
 For face detection, we found that both OpenPose and MTCNN dramatically
-outperformed ViolaJones, our baseline model, especially with respect to
-the random sample, where ViolaJones missed many face that were in view.
-When considering only the composite F-score across all frames, MTCNN
-slightly outperformed OpenPose (0.89 MTCNN vs.~0.83 OpenPose), and MTCNN
-and OpenPose performed comparably with the random sample. Generally,
-MTCNN exhibited higher precision, whereas OpenPose exhibited higher
-recall, and these differences were most pronounced on the randomly
-sample frames. In other words, while OpenPose generated slightly more
-false positives than MTCNN, MTCNN missed several faces that were
-accurately detected by OpenPose. When we restricted our analysis to
-high-confidence detections from OpenPose (\textgreater{}.5 confidence;
-default threshold for visualization), we found very high precision (P =
-0.97), but much lower recall (R = 0.64) and thus overall performance (F
-= 0.77), indicating that these low-confidence detections often indexed
-actual faces that were in the infant view. Figure
-\ref{fig:exampledetections} shows example successful detections from
-OpenPose in each age group, and Figure \ref{fig:failures} shows examples
-of missed faces as well as false positives for context.
+outperformed ViolaJones (our baseline model) especially with respect to
+the random sample, where ViolaJones missed many face that were in view
+(see Table 1). When considering only the composite F-score across all
+frames, MTCNN slightly outperformed OpenPose (0.89 MTCNN vs.~0.83
+OpenPose), and MTCNN and OpenPose performed comparably with the random
+sample. Generally, MTCNN exhibited higher precision, whereas OpenPose
+exhibited higher recall, and these differences were most pronounced on
+the randomly sample frames. In other words, while OpenPose generated
+slightly more false positives than MTCNN, MTCNN missed several faces
+that were accurately detected by OpenPose. When we restricted our
+analysis to high-confidence detections from OpenPose (\textgreater{}.5
+confidence; default threshold for visualization), we found very high
+precision (P = 0.97), but much lower recall (R = 0.64) and thus overall
+lower performance (F = 0.77), indicating that these low-confidence
+detections often indexed actual faces that were in the infant view.
+Figure \ref{fig:exampledetections} shows example successful detections
+from OpenPose in each age group, and Figure \ref{fig:failures} shows
+examples of missed faces as well as false positives for context.
 
@@
 detections, we found much higher precision (P = 0.95), but much lower
-recall (R = 0.36) and thus overall performance (F = 0.52).
+recall (R = 0.36) and thus lower overall performance (F = 0.52).
 
@@
 
-\subsection{Study 1: Access to social information during postural
-developments}\label{study-1-access-to-social-information-during-postural-developments}
-
 \begin{figure}[H]
@@
 
-We report developmental shifts in infants' posture and their orientation
-relative to their caregiver, consistent with previous literature (Adolph
-\& Berger, 2006; Franchak et al., 2017). The proportion of time infants
-spent sitting decreased with age, and the proportion of time infants
-spent standing increased with infants' age. As children got older, their
-locomotor abilities allowed them to become more independent. Both
-8-month-olds and 12-month-olds spent relatively equivalent amounts of
-time lying/crawling (i.e., \enquote{prone}) which was markedly decreased
-in the 16-month-olds, who spent most of their time sitting or standing
-(see Figure \ref{fig:posture}). We also observed changes in infants'
-orientation relative to their caregivers: the 8-month-olds spent more
-time with their caregiver behind them supporting their sitting positions
-than did children at other ages (see Figure \ref{fig:posture}). However,
-we also saw considerable variability across children: some infants spent
-almost their entire time sitting at a close distance from their
-caregiver, whereas others showed more considerable variability (see
-Figure \ref{fig:indivPosOrient}).
-
-\subsubsection{Access to faces and hands across
-age}\label{access-to-faces-and-hands-across-age}
+Consistent with previous literature (Adolph \& Berger, 2006), the
+proportion of time infants spent sitting decreased with age, and the
+proportion of time infants spent standing increased with infants' age.
+As children got older, their locomotor abilities allowed them to become
+more independent. Both 8-month-olds and 12-month-olds spent relatively
+equivalent amounts of time lying/crawling (i.e., \enquote{prone}) which
+was markedly decreased in the 16-month-olds, who spent most of their
+time sitting or standing (see Figure \ref{fig:posture}). We also
+observed changes in infants' orientation relative to their caregivers:
+the 8-month-olds spent more time with their caregiver behind them
+supporting their sitting positions than did children at other ages (see
+Figure \ref{fig:posture}). However, we also saw considerable variability
+across children: some infants spent almost their entire time sitting at
+a close distance from their caregiver, whereas others showed more
+considerable variability (see Figure \ref{fig:indivPosOrient}).
+
+\subsubsection{Changes in access to faces and
+hands}\label{changes-in-access-to-faces-and-hands}
 
@@
 In contrast, infant's locomotor developments had a major effect on the
-faces and hands that were in the field of view. Infants' in-the-moment
-posture was a major factor both in how many faces and hands were in view
-during the play session, as was their orientation relative to their
-caregiver (see Figure \ref{fig:detByPosOrient}). This was quantified
-using two generalized linear mixed-effect models estimating the
-proportion of faces and hands that were in view, with orientation,
-posture, their interaction, and scaled participant's age as fixed
-effects, and with random slopes for infants' orientation and posture. A
-summary of the coefficients of the models can be found in Tables 2 and
-3.
-
-The interaction between infants' posture and their caregiver's
-orientation had the most dramatic effect on the social information in
-view. When caregivers were behind their infants, supporting their
-infants' sitting or standing positions, infants saw fewer faces. When
-caregivers were relatively close to their infants, infants who were
-sitting or standing had more faces in view (Face detections; infant
-sitting and CG close, b = 0.90, SE = 0.07, Z = 13.80, P \textless{}
-0.001; infant standing and CG close, b = 1.23, SE = 0.08, Z = 15.44, P
-\textless{} 0.001) than infants who were lying down/crawling
-(i.e.~prone). When caregivers were far away from their infants, face
-detections were similarity higher (Face detections; infant sitting and
-CG far, b = 0.63, SE = 0.07, Z = 9.16, P \textless{} 0.001; infant
-standing and CG far, b = 1.23, SE = 0.09, Z = 14.43,P \textless{}
-0.001). Infants' age was not significant predictor in accounting for the
-faces in view (Face detections; Age (scaled), b = 0.04, SE = 0.11, Z =
-0.42, P = 0.671).
+faces and hands that were in the field of view (see Figure
+\ref{fig:detByPosOrient}). Two generalized linear mixed-effect models
+were used to predict the proportion of faces and hands in view, with
+orientation, posture, their interaction, and scaled participant's age as
+fixed effects, and with random slopes for infants' orientation and
+posture (see all coefficients in Tables 3 and 4). In particular, the
+interaction between infants' posture and their caregiver's orientation
+had the most dramatic effect on the social information in view. When
+caregivers were behind their infants, supporting their infants' sitting
+or standing positions, infants saw fewer faces. When caregivers were
+relatively close to their infants, infants who were sitting or standing
+had more faces in view (Face detections; infant sitting and caregiver
+(CG) close, b = 0.90, SE = 0.07, Z = 13.80, P \textless{} 0.001; infant
+standing and CG close, b = 1.23, SE = 0.08, Z = 15.44, P \textless{}
+0.001) than infants who were lying down/crawling (i.e.~prone). When
+caregivers were far away from their infants, face detections were
+similarity higher (Face detections; infant sitting and CG far, b = 0.63,
+SE = 0.07, Z = 9.16, P \textless{} 0.001; infant standing and CG far, b
+= 1.23, SE = 0.09, Z = 14.43,P \textless{} 0.001). Infants' age was not
+significant predictor in accounting for the faces in view (Face
+detections; Age (scaled), b = 0.04, SE = 0.11, Z = 0.42, P = 0.671).
 
@@
 Schielzeth, 2013) without their fixed effects (both models were run with
-the maximal random effects structure). The fixed effects in a model with
-only the age of the participants accounted for relatively little
-variance in the proportion of faces (marginal \(R^2\) \textless{} 0.01)
-or hands in view (marginal \(R^2\) = 0.02). However, when adding
-infants' posture and orientation to their caregiver to the model (and
-their interaction) the marginal \(R^2\) were higher for both faces
-(marginal \(R^2\) = 0.25) and wrists (marginal \(R^2\) = 0.26). Overall,
-these results suggest that infants' visual access to social information
-is largely modulated by their posture and orientation to their
-caregiver, which is in turn a function of their general locomotor
-development.
+the maximal random effects structure), and comparing model fits for each
+of these. The fixed effects in a model with only the age of the
+participants accounted for relatively little variance in the proportion
+of faces (marginal \(R^2\) \textless{} 0.01) or hands in view (marginal
+\(R^2\) = 0.02). However, when adding infants' posture and orientation
+to their caregiver to the model (and their interaction) the marginal
+\(R^2\) were higher for both faces (marginal \(R^2\) = 0.25) and wrists
+(marginal \(R^2\) = 0.26). Overall, these results suggest that infants'
+visual access to social information is largely modulated by their
+posture and orientation to their caregiver, which is in turn a function
+of their general locomotor development.
 
@@
 events. In a set of exploratory analyses, we thus analyzed how face and
-hand detections changed during object labeling events relative to
-baseline, analyzing a four-second window around each labeling event
-(e.g., \enquote{Look at the {[}zem{]}!}). Every utterance of one of the
-labelled objects (e.g., \enquote{ball}) was counted as a
-\enquote{labeling event}; timestamps of the beginning of each word were
-hand-annotated and synchronized with the frame-by-frame detections.
+hand detections changed during object naming events relative to
+baseline. We analyzed a four-second window (+/- 2 seconds) each time a
+caregiver uttered a name for one of the objects (e.g., \enquote{Look at
+the {[}zem{]}!}); this time window was chosen in keeping with previous
+work suggesting that parents attend to their target referents during
+this time range (Trueswell et al., 2016). Every utterance of one of the
+objects (e.g., \enquote{ball}) was counted as a \enquote{naming event};
+timestamps of the beginning of each word were hand-annotated and
+synchronized with the frame-by-frame detections.
 
@@
 infant when presenting them with objects (Gogate, Bahrick, \& Watson,
-2000; Gogate, Bolzani, \& Betancourt, 2006). For example, caregivers may
-tend to simultaneously name objects when demonstrating their affordances
-or simply when pointing to them. In turn, infants may be sensitive to
-these naming events and orient their attention towards their caregiver,
-consistent with other accounts positing infants' sensitivity to social
-cues in early word learning environments (Frank, Simmons, Yurovsky, \&
-Pusiol, 2013; Yurovsky, 2018; Yurovsky \& Frank, 2017).
+2000; Gogate, Bolzani, \& Betancourt, 2006) and that hands could play a
+key role in guiding infants' attention during dyadic interactions (Yu \&
+Smith, 2017). For example, caregivers may tend to simultaneously name
+objects when demonstrating their affordances or simply when pointing to
+them. In turn, infants may be sensitive to these naming events and
+orient their attention towards their caregiver, consistent with other
+accounts positing infants' sensitivity to social cues in early word
+learning environments (Yurovsky, 2018; Yurovsky \& Frank, 2017).
 
@@
 
-\subsection{Study 2: Extension of Franchak et al.,
-2017}\label{study-2-extension-of-franchak-et-al.-2017}
-
-In the present work, we found that infant's in-the-moment posture
-changed with their age, as did infants' orientation relative their
-caregiver. In a related study with 12-month-olds, Franchak et al. (2017)
-also found that infants' in-the-moment posture changed the proportion of
-time infants spent looking at faces. Here, we sought to replicate these
-findings using our automated methodology (OpenPose detections), using
-the footage from their head-mounted cameras (hosted on Databrary; D. A.
-Simon, Gordon, Steiger, \& Gilmore, 2015) for two reasons. First, we
-sought to validate our novel method, which could fail to generalize to
-scenes from these more complex environments, where detecting faces and
-hands could arguably be a much harder task. Second, we sought to
-replicate the effects of infants' in-the-moment posture on differences
-in visual access to hands in an independent dataset.
-
-\subsubsection{Comparison between video
-datasets}\label{comparison-between-video-datasets}
-
-First, the play environment in Franchak et al. (2017) that infants were
-immersed in with their caregivers (and experimenters) was much larger
-and more varied than the play room used in the present dataset,
-containing multiple structures and toys in different parts of the room
-for infants to climb, explore, and interact with, and infants were
-unconstrained and allowed to freely wander the room. In contrast, the
-play room used in Study 1 was relatively small (approximately 10' x 10')
-and was set-up with focused play with the pairs of novel and familiar
-objects. Further, multiple people were present during the play session
-in Franchak et al. (2017)-- including their caregiver and two
-experimenters -- whereas in Study 1 the experimenters left the room
-during the play session. The view angle of the two head-mounted cameras
-used in Study 1 and Franchak et al. (2017) were relatively similar
-(\(52.2^{\circ}\) horizontal by \(42.2^{\circ}\) in Franchak et al.
-(2017), \(47^{\circ}\) horizontal by \(36^{\circ}\) vertical in Study
-1).
+\section{Study 2}\label{study-2}
+
+In Study 1, we found that infant's in-the-moment posture changed with
+their age, as did infants' orientation relative their caregiver. In a
+related study with 12-month-olds, Franchak et al. (2017) also found that
+infants' in-the-moment posture changed the proportion of time infants
+spent looking at faces. Here, we sought to replicate these findings
+using our automated methodology (OpenPose detections), using the footage
+from their head-mounted cameras (hosted on Databrary; D. A. Simon,
+Gordon, Steiger, \& Gilmore, 2015) for two reasons. First, we sought to
+validate our novel method, which could fail to generalize to scenes from
+these more complex environments, where detecting faces and hands could
+arguably be a much harder task. Second, we sought to replicate the
+effects of infants' in-the-moment posture on differences in visual
+access to hands in an independent dataset.
+
+\subsection{Methods}\label{methods-1}
+
+\subsubsection{Participants}\label{participants-1}
+
+With the aid of Franchak et al. (2017), we obtained the scene camera
+footage from the head-mounted eye-trackers for the 17 one-year-old
+infants (range 11.8--12.4 months) who participated. As noted in Franchak
+et al. (2017), families were recruited from maternity wards of local
+hospitals in the New York City metropolitan area and were predominantly
+white and middle class.
+
+\subsubsection{Head-mounted camera}\label{head-mounted-camera-1}
+
+The view angle of the two head-mounted cameras used in these two studies
+were relatively similar (\(52.2^{\circ}\) horizontal by \(42.2^{\circ}\)
+in Franchak et al. (2017), \(47^{\circ}\) horizontal by \(36^{\circ}\)
+vertical in Study 1). However, in Study 2 the camera was mounted on the
+side of the infants' head in Study 2, closer to infants temples, whereas
+in Study 1 the camera was situated in the middle of infants' forehead
+and oriented slightly upwards.
+
+\subsubsection{Procedure}\label{procedure-1}
+
+The play environment that infants were immersed in with their caregivers
+(and experimenters) was much larger and more varied than the play room
+used in Study 1, containing multiple structures and toys in different
+parts of the room for infants to climb, explore, and interact with, and
+infants were unconstrained and allowed to freely wander the room. In
+contrast, the play room used in Study 1 was relatively small
+(approximately 10' x 10') and was set-up for focused play on a mat with
+the pairs of novel and familiar objects. In addition, multiple people
+were present during the play session -- including their caregiver and
+two experimenters -- whereas in Study 1 the experimenters left the room
+during the play session.
+
+\subsubsection{Video annotations}\label{video-annotations}
+
+The first five minutes of each of the videos were coded for the infants'
+posture (upright, prone, or sitting) by trained coders in Franchak et
+al. (2017). These frame-by-frame posture annotations were synced with
+the outputs of the same automated annotation used in Study 1.
 
@@
 
-\subsubsection{Replication using automated
-detections}\label{replication-using-automated-detections}
-
-Overall, we found convergence between our two methodologies, replicating
-the main results from Franchak et al. (2017), and finding that the
-proportion of faces detected was greater when infants were sitting or
-standing vs.~prone (see Figure \ref{fig:franchak}). We found this result
-regardless of whether we used all detections (proportion of frames with
-face detections; Prone: M = 0.13, Sitting: M = 0.20, Upright: M = 0.22)
-or restricted our analyses to high-confidence detections (percentage of
-frames with face detections; Prone: M = 0.03, Sitting: M = 0.06,
-Upright: M = 0.05). These results were confirmed in generalized
-linear-mixed models, as in the first study (Sitting vs.~Prone, b = 0.34,
-SE = 0.02, Z = 23.37, P \textless{} 0.001. Upright vs.~Prone, b = 0.38,
-SE = 0.02, Z = 24.01, P \textless{} 0.001).
-
-The most noticeable difference between our results is the overall
-proportion of frames with faces detected by OpenPose vs.~the proportion
-of frames in which infants were judged to be looking at faces in
-Franchak et al. (2017). Across the entire session in Franchak et al.
-(2017), infants looked at faces on 4.7\% of frames (Franchak et al.,
-2017). When we used all detections from OpenPose, we found a higher
+\subsection{Results}\label{results-1}
+
+\subsubsection{Differences between eye-tracking vs.~automated
+detections}\label{differences-between-eye-tracking-vs.automated-detections}
+
+First, we compared the overall proportion of frames in which infants
+foveated faces -- as assessed by the head-mounted eye-tracker -- in
+Franchak et al. (2017) vs.~the proportion of frames with faces detected
+by OpenPose. We expected some differences, as (1) head-mounted
+eye-trackers may underestimate the proportion of faces attended due to
+calibration issues, and (2) OpenPose may detect faces that are in view
+for infants but that infants may not be foveating. Across the entire
+session in Franchak et al. (2017), infants looked at faces on 4.7\% of
+frames. When we used all detections from OpenPose, we found a higher
 proportion of faces -- 21.80\%. When we restricted our results to only
@@
 high-confidence detections dramatically underestimate the number of
-faces in view. Thus, we suspect that these differences between our
-findings are due to both (1) the fact head-mounted eye-trackers may
-sometimes underestimate the proportion of faces attended due to
-calibration issues, and that (2) OpenPose may sometimes detect faces
-that are in view for infants but that infants may not be foveating.
-
-We also found that infants' in-the-moment posture also modulated the
+faces in view. Thus, OpenPose may somewhat overestimate the proportion
+of faces that are actually foveated, while head-mounted eye-trackers may
+underestimate the proportion of faces that infants' could be attending
+to.
+
+\subsubsection{Replication \& extension using automated
+detections}\label{replication-extension-using-automated-detections}
+
+Despite these differences, we found convergence between our two
+methodologies, replicating the main results from Franchak et al. (2017),
+and finding that the proportion of faces detected was greater when
+infants were sitting or standing vs.~prone (see Figure
+\ref{fig:franchak}). We found this result regardless of whether we used
+all detections (proportion of frames with face detections; Prone: M =
+0.13, Sitting: M = 0.20, Upright: M = 0.22) or restricted our analyses
+to high-confidence detections (percentage of frames with face
+detections; Prone: M = 0.03, Sitting: M = 0.06, Upright: M = 0.05).
+These results were confirmed in generalized linear-mixed models, with
+random intercepts for each subjects and infants' posture as a fixed
+effect (Sitting vs.~Prone, b = 0.34, SE = 0.02, Z = 23.37, P \textless{}
+0.001. Upright vs.~Prone, b = 0.38, SE = 0.02, Z = 24.01, P \textless{}
+0.001).
+
+We also found that infants' in-the-moment posture modulated the
 proportion of hands that were in view (i.e., wrist detections), though
@@
 Prone: M = 0.15; Sitting: M = 0.21; Upright: M = 0.22). These results
-were confirmed in generalized linear-mixed models, as in the first study
-(Sitting vs.~Prone, b = 0.27, SE = 0.01, Z = 19.09, P \textless{} 0.001.
-Upright vs.~prone, b = 0.30, SE = 0.02, Z = 19.73, P \textless{} 0.001).
+were confirmed in generalized linear-mixed models with the same model
+structure as with faces (Sitting vs.~Prone, b = 0.27, SE = 0.01, Z =
+19.09, P \textless{} 0.001. Upright vs.~prone, b = 0.30, SE = 0.02, Z =
+19.73, P \textless{} 0.001).
 
@@
 What social cues do infants see as they learn language, and how does
-this change as they grow and start to locomote themselves? We explored
-this question using video data from head-mounted cameras from two
-naturalistic datasets of parent-child interactions: a cross-sectional
-database of play sessions from 8-16 month-olds with sets of novel and
-familiar toys, and database of head-mounted camera videos from
-one-year-olds who explored a large play area from Franchak et al.
-(2017). To analyze these datasets, we developed a novel method using a
+infants' access to these cues change as they grow and start to locomote
+themselves? We explored this question using video data from head-mounted
+cameras from two naturalistic datasets of parent-child interactions: a
+cross-sectional database of play sessions from 8-16 month-olds with sets
+of novel and familiar toys, and database of head-mounted camera videos
+from one-year-olds who explored a large play area (Franchak et al.,
+2017). To analyze these datasets, we developed a novel method using a
 pose detection model to automate the annotation of the social
@@
 Despite being trained on the adult perspective, the pose detector we
-used (OpenPose, Cao et al. (2017)) was able to generalize relatively
-well to the infant viewpoint, achieving comparable precision and
-accuracy as a face detector relative to a state-of-the-art model
-optimized specifically for detecting faces in natural scenes (MTCNN, K.
-Zhang et al. (2016)). While OpenPose had relatively low recall as a hand
-detector -- missing some hands that were in the infant view -- it made
-comparable rates of false alarms. In both cases, we found that overall
-performance was maximized when all detections were included, regardless
-of their confidence, suggesting that some low-confidence face and hand
-detection still index actual faces and hands that were seen by infants.
-Thus, while imperfect, we suggest that OpenPose can be applied to infant
+used (OpenPose, Cao et al., 2017) was able to generalize relatively well
+to the infant viewpoint, achieving comparable precision and accuracy as
+a face detector relative to a state-of-the-art model optimized
+specifically for detecting faces in natural scenes (K. Zhang et al.,
+2016). While OpenPose had relatively low recall as a hand detector --
+missing some hands that were in the infant view -- it made comparable
+rates of false alarms. In both cases, we found that overall performance
+was maximized when all detections were included, regardless of their
+confidence, suggesting that some low-confidence face and hand detection
+still index actual faces and hands that were seen by infants. Thus,
+while imperfect, we suggest that OpenPose can be applied to infant
 egocentric videos for the extraction of the social information in the
@@
 around naming events -- and to re-analyze the data from Franchak et al.
-(2017) to replicate our findings in a very different kind of play
-session.
+(2017), replicating our findings in a very different kind of play
+session. Furthermore, future work may be able to fine-tune pose
+detectors for even better accuracy, leveraging human annotations of the
+faces and hands that infants see to adapt models for the infant view.
 
@@
 trajectory in a population of children from primarily WEIRD contexts --
-white, educated, industrialized, rich, and democratic (D. Jones, 2010).
+white, educated, industrialized, rich, and democratic (Henrich, Heine,
+\& Norenzayan, 2010). Parenting practices with respect to motor
+development can and do vary widely across cultures (Karasik,
+Tamis-LeMonda, Ossmy, \& Adolph, 2018) -- and these choices likely
+influence the social cues that children see and how they use them.
 
@@
 
-Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in
-recruitment, data collection, and annotation. This work was funded by a
-Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF,
-and NSF \#1714726 to BLL. An earlier version of this work was presented
-to the Cognitive Science Society in Frank et al. (2013) and Sanchez,
-Long, Kraus, and Frank (2018).
+{[}BLINDED{]}
 
@@
 
-\hypertarget{ref-frank2013}{}
-Frank, M. C., Simmons, K., Yurovsky, D., \& Pusiol, G. (2013).
-Developmental and postural changes in children's visual access to faces.
-In \emph{Proceedings of the 35th annual meeting of the cognitive science
-society} (pp. 454--459).
-
 \hypertarget{ref-gogate2000study}{}
@@
 
+\hypertarget{ref-henrich2010most}{}
+Henrich, J., Heine, S. J., \& Norenzayan, A. (2010). Most people are not
+weird. \emph{Nature}, \emph{466}(7302), 29--29.
+
 \hypertarget{ref-iverson2010}{}
@@
 
-\hypertarget{ref-jones2010weird}{}
-Jones, D. (2010). A weird view of human nature skews psychologists'
-studies. American Association for the Advancement of Science.
-
 \hypertarget{ref-karasik2014}{}
@@
 \emph{Developmental Science}, \emph{17}(3), 388--395.
+
+\hypertarget{ref-karasik2018ties}{}
+Karasik, L. B., Tamis-LeMonda, C. S., Ossmy, O., \& Adolph, K. E.
+(2018). The ties that bind: Cradling in tajikistan. \emph{PloS One},
+\emph{13}(10), e0204428.
 
@@
 
-\hypertarget{ref-sanchez2018postural}{}
-Sanchez, A., Long, B., Kraus, A. M., \& Frank, M. C. (2018). Postural
-developments modulate children's visual access to social information.
-
 \hypertarget{ref-scheirer2014perceptual}{}
@@
 
+\hypertarget{ref-trueswell2016perceiving}{}
+Trueswell, J. C., Lin, Y., Armstrong III, B., Cartmill, E. A.,
+Goldin-Meadow, S., \& Gleitman, L. R. (2016). Perceiving referential
+intent: Dynamics of reference in natural parent--child interactions.
+\emph{Cognition}, \emph{148}, 117--135.
+
 \hypertarget{ref-vanrullen2017perception}{}

### 2020-08-26 writeup-paper/draft_apa_blinded.Rmd
fixing last typos; add to gitignore

@@
 And as infants reach their first birthday, they also tend to follow [@yu2013joint; @yu2017hand] and imitate the gestures of their caregivers (e.g., pointing).
-Infants ability to process these social cues may provide strong scaffolding for early word learning.
+Infants' ability to process these social cues may provide strong scaffolding for early word learning.
 Longitudinal studies provide some evidence for this link: children’s level of joint engagement with their mother at 9-12 months predicts both their receptive and productive vocabularies [@carpenter1998] and 10 month-olds who follow an adult’s gaze (in an experimental context) have larger vocabularies at 18 months and throughout the second year of life [@brooks2005; @brooks2008].
@@
 These motoric developments can be thought of as gateways that open up entirely new phases of development [@iverson2010], causing a cascade of changes in an infant's ability to interact with their world and the people in it [@karasik2014].
-Thus, infants’ changing locomotor abilities may shape the social cues that infants' see and seek out, in turn impacting their cognitive and linguistic abilities.
+Thus, infants’ changing locomotor abilities may shape the social cues that infants see and seek out, in turn impacting their cognitive and linguistic abilities.
 
@@
 ### Accuracy of automated detections 
-For face detection, we found that both OpenPose and MTCNN dramatically outperformed ViolaJones (our baseline model) especially with respect to the random sample, where ViolaJones missed many face that were in view (see Table 2).
+For face detection, we found that both OpenPose and MTCNN dramatically outperformed ViolaJones (our baseline model) especially with respect to the random sample, where ViolaJones missed many faces that were in view (see Table 2).
 When considering only the composite F-score across all frames, MTCNN slightly outperformed OpenPose  (`r round(perfFacesOverall$f[1],2)` MTCNN vs.
 `r round(perfFacesOverall$f[2],2)` OpenPose), and MTCNN and OpenPose performed comparably with the random sample.
-Generally, MTCNN exhibited higher precision, whereas OpenPose exhibited higher recall, and these differences were most pronounced on the randomly sample frames.
+Generally, MTCNN exhibited higher precision, whereas OpenPose exhibited higher recall, and these differences were most pronounced on the randomly sampled frames.
 In other words, while OpenPose generated slightly more false positives than MTCNN, MTCNN missed several faces that were accurately detected by OpenPose.
@@
 We thus only considered differences in the relative proportion of faces or hands in view as a function of age, posture, and orientation, rather than comparing them directly.
-Overall, we did not observe strong age related trends from 8-16 months of age; if anything, face detections showed a slight u-shaped pattern, with 12-month-olds having slightly fewer faces in their visual field that 8- or 16-month-olds.
+Overall, we did not observe strong age related trends from 8-16 months of age; if anything, face detections showed a slight U-shaped pattern, with 12-month-olds having slightly fewer faces in their visual field than 8- or 16-month-olds.
 
@@
 
-```{r detByPosOrient, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation.
+```{r detByPosOrient, fig.env="figure", out.width = "\\textwidth", fig.pos = "H", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregiver's orientation.
 Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted.
@@
 
-In contrast, infant's locomotor developments had a major effect on the faces and hands that were in the field of view (see Figure \ref{fig:detByPosOrient}).
+In contrast, infants' locomotor developments had a major effect on the faces and hands that were in the field of view (see Figure \ref{fig:detByPosOrient}).
 Two generalized linear mixed-effect models were used to predict the proportion of faces and hands in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture (see all coefficients in Tables 4 and 5).
@@
 prone).
-When caregivers were far away from their infants, face detections were similarity higher (Face detections; infant sitting and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4])`; infant standing and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`,`r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
+When caregivers were far away from their infants, face detections were similarily higher (Face detections; infant sitting and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4])`; infant standing and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`,`r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
 Infants' age was not significant predictor in accounting for the faces in view (Face detections; Age (scaled), b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).

### 2021-03-15 analysis/1_merge_data/merge_all_data.Rmd
amend details to cameras in preprocessing

@@
 ```{r}
-# All of the videos were filmed with an MD-80 headcam.
-Some videos had a fisheye lens attached to them that increased their FOV by approximately 50% in both directions (see manuscript; in demographics column "fisheye").
+# Some of the videos were filmed with an MD-80 headcam where a fisheye lens attached to them that increased their FOV by approximately 50% in both directions (see manuscript; in demographics column "fisheye").
 This was determined by looking at the third-person videos.
@@
 ### Sanity checks for merging detectors
-We know that openPose works on videos and so might notget all frames. Let's check out the frames it doesn't have
+We know that openPose works on videos and so might not get all frames. Let's check out the frames it doesn't have
 

### 2021-03-25 writeup-paper/SI_extraplots.Rmd
appendix figure 1, si plotting code

--- 2020-08-11 analysis/4_extra_plots/SI_extraplots.Rmd
@@
 ```
-
-```{r afterNamingsExplore, eval=FALSE, include=FALSE, echo=FALSE}
-afterNamings_Faces <- d %>%
-  group_by(subid) %>%
-  mutate(detections = faceOP) %>%
-  do(summarize.naming((.),window = c(0,4)))
-
-detectionsByGroup <- dlongPlot %>%
-  group_by(age.grp, subid, detectorType) %>%
-  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
-  group_by(age.grp, detectorType) %>%
-  multi_boot_standard(col = 'prop_detected')
-
-afterNamings_Detections_BySub_Faces <- afterNamings_Faces %>%
-  group_by(subid, familiarity, age.grp) %>%
-  summarize(prop_detected = mean(detections, na.rm=TRUE))
-
-afterNamings_Detections_Grouped_Faces <- afterNamings_Detections_BySub_Faces %>%
-  group_by(familiarity, age.grp) %>%
-  multi_boot_standard(col = 'prop_detected')
-
-faces_afternamings_plot <- ggplot(afterNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
-  # plot baseline
-  geom_pointrange(data = (detectionsByGroup %>% filter(detectorType == 'OpenPose-Faces')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=3), alpha=.9) +
-  # plot averages by familiarity 
-  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
-  # plot familairity detections for each sub
-  geom_point(data=afterNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
-  theme_few() +
-  labs(y = "Prop. face detections", x = "Age (months)") +
-  theme_few(base_size=plot_size) +
-  ylim(0,.6) +
-  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
-  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
-  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
-  scale_size_continuous(name="# of naming events")
-
-```
-
-```{r}
-### get individual points
-baseline_diffs_faces <- dlong %>%
-  group_by(age.grp, subid, detectorType) %>%
-  summarise(prop_detected = mean(detection)) %>%
-  mutate(familiarity = 'baseline') %>%
-  full_join(aroundNamings_Detections_BySub_Faces) %>%
-  spread(familiarity,  prop_detected) %>%
-  summarize(fam_vs_baseline = familiar - baseline, novel_vs_baseline = novel - baseline)
-
-baseline_by_group <-  baseline_diffs_faces %>%
-  group_by(age.grp) %>%
-  filter(!is.na(novel_vs_baseline)) %>%
-  multi_boot_standard(col='novel_vs_baseline')
-
-ggplot(baseline_by_group, aes(x=age.grp, y=mean)) +
-  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
-  # plot familairity detections for each sub
-  geom_point(data=afterNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
-  theme_few() +
-```

### 2021-04-03 writeup-paper/draft_apa_blinded_revised.Rmd
April revisions to .rmd

--- 2020-08-26 writeup-paper/draft_apa_blinded.Rmd
@@
 
-  Social information plays a key role in infants\' early linguistic and cognitive development.
-Yet infants locomotor abilities also develop continuously as they learn to sit, crawl, stand, and walk.
-How do these postural developments affect infants access to the social information relevant for early learning?
-To address this question, we first created a rich annotated dataset of naturalistic play sessions between infants and their caregivers with sets of novel and familiar toys (N=36, 8-16 month-olds), recording both egocentric and third-person video.
-We then developed an automated analysis method using a pose detection model to detect the faces and hands in the infant view.
-We found that infants\' posture and orientation to their caregiver changed dramatically across this age range and modulated infants access to social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces and hands in their visual field.
-We applied this automated method to a second egocentric video dataset from a previous study, replicating and extending their findings on the influence of infants' posture on the social information in view.
-Together, these results establish the validity of a new method for detecting social information in egocentric videos and confirm that motoric developments play a significant role in shaping the social information that infants have access to.
-The combined use of head-mounted cameras and new computer vision techniques is a promising avenue for understanding how the statistics of infants\' visual and linguistic experience change over development.
+  How do postural developments affect infants’ access to social information?
+We recorded egocentric and third-person video while infants and their caregivers (N=36, 8–16-month-olds) participated in naturalistic play sessions.
+We then validated the use of a neural network pose detection model to detect faces and hands in the infant view.
+We used this automated method to analyze our data and a prior egocentric video dataset (N=17, 12-month-olds).
+Infants’ average posture and orientation with respect to their caregiver changed dramatically across this age range; both posture and orientation modulated access to social information.
+Together, these results confirm that infant’s ability to move and act on the world plays a significant role in shaping the social information in their view.
 
 keywords          : "Postural developments, headcam, social information, face detection, pose detection"
-wordcount         : "7324"
+wordcount         : "8148"
 
@@
 
-figsintext        : yes
+figsintext        : no
 figurelist        : yes
@@
 
-```{r, libraries}
+```{r, libraries include=FALSE}
 library(knitr)
@@
                       cache=FALSE, message = FALSE, sanitize = TRUE)
+
+library(here)
 library(readr)
@@
 Longitudinal studies provide some evidence for this link: children’s level of joint engagement with their mother at 9-12 months predicts both their receptive and productive vocabularies [@carpenter1998] and 10 month-olds who follow an adult’s gaze (in an experimental context) have larger vocabularies at 18 months and throughout the second year of life [@brooks2005; @brooks2008].
-While the relationship between hand-following and language development has been less well characterized, infants who follow their caregivers’ hands tend to be those who spend more time jointly attending to events with their caregivers [@yu2017hand].
+While the relationship between hand-following and language development has been less well characterized, infants who follow their caregivers’ hands tend to be those who spend more time jointly attending to events with their caregivers [@yu2017hand] and caregivers tend to create referential clarity by holding objects for their infants to see [@suanda2019signal].
 
@@
 By using head-mounted cameras to record what infants see, researchers have begun to document the infant egocentric perspective [@yoshida2008] and to quantify the information -- social and otherwise -- available to infants as they learn.
-While head-mounted camera data do not provide explicit information on where infants are attending, some work suggests that infants orient their head towards what they are focusing on -- putting those people or objects in view [@yoshida2008].
+While head-mounted camera data do not provide explicit information about what infants are looking at (unlike head-mounted eye-trackers), some work suggests that infants orient their head towards what they are focusing on -- putting those people or objects in view [@yoshida2008].
 Initial recordings using this technique during in-lab play sessions revealed a different view than many imagined: instead of being dominated by faces, the infant perspective contained close up views of primarily toys and hands [@yoshida2008; @franchak2011; @yu2017hand].
@@
 
-These changes in perspective are likely the downstream consequence of myriad factors, chief of which may be infants' evolving locomotive abilities: an infant's ability to sit, crawl, stand, or walk structures the way they interact with the things and people in their world.
-Infants' locomotive abilities can change rapidly: the physical body of an infant can lengthen up to 2cm in 24 hours [@lampl1993evidence], leading to changes in their abilities from one day to the next.
-These motoric developments can be thought of as gateways that open up entirely new phases of development [@iverson2010], causing a cascade of changes in an infant's ability to interact with their world and the people in it [@karasik2014].
-Thus, infants’ changing locomotor abilities may shape the social cues that infants see and seek out, in turn impacting their cognitive and linguistic abilities.
-
-However, understanding the nuanced relationship between the social information in the infant view and their motoric and linguistic development has remained challenging.
-For example, as children begin crawling [@adolph1998roles] -- or scooting or cruising [@patrick2012developmental] -- they are no longer constrained to the same spot that their caregivers last placed them in.
-Yet while crawlers can choose where to go and what they see to a much greater degree, they also appear to spend much of their time in a world populated by floors and knees; during spontaneous play, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011], when they have full visual access to their environment and the people in it [@kretch2014].
-Further, infants' experience with sitting predicts their success at 3D object completion tasks [@soska2010systems] as well as their receptive vocabulary [@libertus2016sit], suggesting the importance of focused play sessions for language development.
-
-Others have posited that it is children’s ability to stand and walk that fundamentally changes their ability to access social information (e.g., facial expressions, gaze cues, pointing) relative to children who are still crawling and sitting [@walle2016infant], which in turn allow infants to learn words quicker and more efficiently [@walle2016infant].
+Broadly, the field is in need of computational tools to reuse these rich video datasets and understand the generalizability of findings across populations, tasks, and age-ranges and to understand how changes in the infant view are influenced by other aspects of development.
+For example, children’s evolving motoric abilities likely change how they participate with their caregivers in different kinds of play sessions (e.g., exploring novel environments vs.
+playing with novel objects) which in turn may shape the social cues that children see and use during learning.
+Yet while the field has assembled many head-mounted camera datasets, conducting new analyses on these videos has remained prohibitively time-consuming due to a lack of computational tools for annotations.
+Instead, hundreds of hours of manual annotations have been required to analyze a fraction of the available frames for a given analysis.
+Thus, despite containing a wealth of information about the structure of parent-child interactions, these datasets have thus gone dramatically underused.
+As a result, an understanding of the nuanced relationship between the social information in the infant view and children’s motoric and linguistic development has remained challenging.
+
+Indeed, developmental changes in the infant view are likely the downstream consequence of myriad factors, including infants’ evolving locomotive abilities: an infant’s ability to sit, crawl, stand, or walk structures the way they interact with the things and people in their world.
+These motoric developments have been thought of as gateways that open up entirely new phases of development [@iverson2010] causing a cascade of changes in an infant’s ability to interact with their world and the people in it [@karasik2014].
+
+Thus, one idea is that infants’ changing locomotor abilities could shape the social cues that infants see and seek out, in turn impacting their cognitive and linguistic abilities.
+Some evidence supports this view: for example, infants’ experience with sitting predicts their success at 3D object completion tasks [@soska2010systems] as well as their receptive vocabulary [@libertus2016sit], suggesting the importance of focused play sessions for language development.
+Later, as children begin crawling [@adolph1998learning] – or scooting or cruising [@patrick2012developmental] – their view of the world changes as they are no longer constrained to the same spot that their caregivers last placed them in.
+Yet while crawlers can choose where to go and what they see to a much greater degree, they also appear to spend much of their time in a world populated by floors and knees; during spontaneous play, toddlers are more likely to look at the floor while crawling than while walking [@franchak2011],, when they have full visual access to their environment and the people in it [@kretch2014].
+
+On one theoretical view, it is primarily children’s ability to stand and walk that fundamentally changes their ability to access social information (e.g., facial expressions, gaze cues, pointing) relative to children who are still crawling and sitting, which could in turn allow infants to learn words quicker and more efficiently [@walle2016infant].
 Supporting this idea, walking vs.
 crawling infants tend to make different kinds of object-related bids for attention from their caregivers [@karasik2014], hear more action directed statements (e.g., “open it”) [@karasik2014], and have higher receptive and productive vocabularies [@walle2014].
-However, not all evidence supports this view: parental report data suggest a weak relationship between walking and the onset of language [@moore2019point], and head-mounted eye-tracking data from one-year-olds highlights the role of infants’ in-the-moment posture -- and that of their caregivers’ -- in what infants’ see, finding that these factors are better predictors of the social information in view than their locomotor status per se (i.e., “walker” vs “crawler”) [@franchak2017see].
-
-Broadly, the field is in need of computational tools to reuse these rich video datasets and understand the generalizability of findings across different populations, tasks, and age-ranges.
-For example, different kinds of play sessions (e.g., exploring novel environments vs.
-playing with novel objects) might modulate the degree to which infants’ posture changes the social information in view.
-Yet while the field has assembled many head-mounted camera datasets, conducting new analyses on these videos has remained prohibitively time-consuming due to a lack of computational tools for annotations.
-Instead, hundreds of hours of manual annotations have been required to analyze a fraction of the available frames for a given analysis.
-Thus despite containing a wealth of information about the structure of parent-child interactions, these datasets have thus gone dramatically underused.
-
-Recent innovations in computer vision hold promise for automated annotations of the infant view.
+However, not all evidence supports this view: parental report data suggest no relationship between walking and the onset of language [@moore2019point].
+Further, using head-mounted eye-tracking data from one-year-olds, Franchak et al., 2018 found that infants’ in-the-moment posture also interacts with their caregivers’ posture to shape the social information in view [@franchak2018see], highlighting the need to consider not only children’s motoric abilities but how caregivers adapt to them.
+
+Recent innovations in computer vision hold promise for understanding the generality of these findings.
+By automating annotations of the infant view, we can go beyond limited sets of manual annotations to characterize the consistency and variability in the social information that children see during early learning.
 Over the past decade, deep neural networks have become dramatically better at a wide range of visual tasks, including object classification [@simonyan2014very] scene categorization [@zhou2017places], and pose detection [@zhang2016], arguably facilitating our understanding of visual perception [@peterson2018evaluating; @vanrullen2017perception] and improving computational neuroscience [@kietzmann2018deep].
-However, as most models have been trained on photographs or videos taken from the adult perspective, it is still unclear how easily these models can be applied to videos taken from the infant perspective.
-
-In the present paper, we make progress on understanding the relationship  between the infant visual perspective and motoric and linguistic development in three ways.
-First, we validate a model for the automated detection of faces and hands in the infant view.
-We use an open-source model of pose detection -- OpenPose [@zhang2016; @cao2017realtime] -- that provides pose, face, and hand keypoints for every person in an image.
-We compare the detection accuracy of this method with that of both older and more specialized models of face detection, and demonstrate the usability of this off-the-shelf model for quantifying the faces and hands in the infant view.
-
-Second, we provide an annotated, open dataset for researchers to examine the effects of postural developments and naming behavior during naturalistic parent-child interactions.
-To do so, we invited caregivers of 8, 12, and 16-month-olds to participate in play sessions where they were provided with pairs of novel and familiar objects (e.g., a ball and a microfiber duster, called a "zem").
-We chose this age-range to span the months when infants typically transition from sitting to crawling to standing, and documented the infant egocentric view as well as a third-person view of the play session.
-We then annotated infants' in-the-moment posture and orientation relative to their caregiver, and obtained full transcripts of the play sessions as well as MacArthur-Bates Communicative Development Inventories for each child.
-
-Third, we apply our automated method to this and one other openly available head-mounted camera dataset, examining the generality of the effects of motor development on the social information in view.
-In Study 1, we use the cross-sectional design of our dataset to directly examine the relative contributions of age vs.
-in-the-moment posture on infants' visual access to social information, and use the transcriptions to examine how the availability of social cues changes relative to naming events (e.g., “Yes, you like the [ball!]”).
+Yet as most models have been trained on photographs or videos taken from the adult perspective, it is unclear how easily these models can be applied to videos taken from the infant perspective.
+While some computer vision algorithms have indeed been adapted for egocentric vision (e.g., gaze predictions, [@zhang2018coarse]), very few have been adapted for egocentric video data from infants [@bambach2015lending].
+
+Here, we make progress on understanding the social information in the changing infant view by adapting novel computational methods.
+We use a publicly available pose detection model [@cao2017realtime; @zhang2016]) for the detection of faces and hands in infant egocentric video data, a similar approach to that used by @long2020detecting.
+We compare the detection accuracy of this method with that of both older and more specialized models of face detection, demonstrating the usability of this off-the-shelf model for quantifying the faces and hands in the infant view in egocentric video datasets.
+
+We then use these automated detections to examine how the social information in view changes with respect to infant’s age and real-time posture in two different egocentric video datasets.
+In Study 1, we use the cross-sectional design of our dataset to examine the relative contributions of children’s age vs.
+real-time posture on infants’ visual access to social information, and use the transcriptions to explore how the availability of social cues changes relative to naming events (e.g., “Yes, you like the [ball!]”).
 Indeed, despite positing links between the social information in view and language development, no work to date has directly examined how the availability of social information changes around naming events in naturalistic contexts.
-In Study 2, we apply this automated method to @franchak2017see, where one-year-olds wore head-mounted eye-tracking cameras during a play session and their in-the-moment posture was hand-annotated.
+In Study 2, we apply this same automated method to @franchak2018see, where one-year-olds wore head-mounted eye-tracking cameras during a play session and their in-the-moment posture was hand-annotated (https://nyu.databrary.org/volume/135).
 Unlike in Study 1, infants and caregivers roamed a large, open playroom and explored different toys placed throughout.
-We thus analyze this second dataset with a first goal of  validating our automated method on a very different kind of video dataset, attempting to replicate their primary findings originally obtained with a head-mounted eye-tracker, and then further examine whether the proportion of hands in view also varies with infants’ posture.
-Broadly, we predicted that there would be differential access to social information based on children's postural developments: crawling infants would see fewer faces/hands because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape with greater access to the social information in their environment.
+We analyze this second dataset with the goal of validating our automated method on a very different kind of video dataset, extending their primary findings originally obtained with a head-mounted eye-tracker.
+Across both datasets, we predicted that there would be differential access to social information based on children’s postural developments in both datasets: crawling infants would see fewer faces/hands because they would primarily be looking at the ground, while walking toddlers would have access to a richer visual landscape with greater access to the social information in their environment.
+
+To preview our results, we find that infants’ changing locomotor abilities are a major factor that shape the statistics of the social visual environment, confirming and extending previous work.
+Thus, children’s social learning environments appears to change dramatically as children change in their ability to move on their own and interact with the world, in line with recent proposals emphasizing the child as an active learner [@xu2019towards]
 
@@
 ## Methods
-Caregivers of 8, 12, and 16-month-olds were invited to participate in play sessions where they were provided with pairs of novel and familiar objects (e.g., a ball and a microfiber duster, called a "zem") in a small playroom in a lab (approximately 10' x 10').
-Infants wore head-mounted cameras equipped with a fish-eye lens (see Head-mounted camera); a tripod-mounted camera captured a third-person view of the play session.
-Using these video data, infants' posture and orientation to their caregiver were hand-coded and annotated for the entirety of the play session.
-All videos were transcribed, and MacArthur CDIs were collected for all children who participated.
+We provide an annotated, open dataset for researchers to examine the effects of postural developments and naming behavior during naturalistic parent-child interactions.
+Caregivers of 8, 12, and 16-month-olds were invited to participate in play sessions where they were provided with pairs of novel and familiar objects (e.g., a ball and a microfiber duster, called a “zem”) in a small playroom in a lab (approximately 10 x 10 feet).
+Infants wore head-mounted cameras (see Head-mounted camera) and a tripod-mounted camera captured a third-person view of the play session.
+Using these video data, infants’ posture and orientation to their caregiver were hand-coded and annotated for the entirety of the play session; this age-range spans the months when infants typically transition from sitting to crawling to standing All videos were transcribed, and MacArthur CDIs were collected for all children who participated.
 All materials have been made publicly available on Databrary for whom the parents provided sharing consent (29/36 dyads) via [BLINDED].
@@
   rename(subid = subID)
-inc_subs <- read.csv(paste0(here::here(),"/data/demographics/demographics_updated.csv"))
+
+inc_subs <- read.csv(paste0(here::here(),"/data/demographics/demographics_updated.csv")) %>%
+  mutate(camera = factor(Fisheye, levels = c('no','yes'), labels = c('veho','md80')))
+
 all_subs$inc <- all_subs$reason=="included"
@@
             age = mean(age.at.test, na.rm=TRUE),
+            age_min = min(age.at.test, na.rm=TRUE),
+            age_max = max(age.at.test, na.rm=TRUE),
             video_length = mean(len.min, na.rm=TRUE),
@@
 
-colnames(sub_summary) <- c('Group','N','% incl.','Avg. age', 'Avg video length', 'Num female')
+colnames(sub_summary) <- c('Group','N','% incl.','Avg. age', 'Min age','Max age','Avg video length', 'Num female')
+
+ages_inc_subs <- inc_subs %>%
+  select(subid, age.at.test, age.grp, camera)
 
@@
 `r sum(all_subs$reason=="infant crying")` for fussiness, and 1 for sample symmetry.
+Technical issues related to the initial headcam model (MD-80) led us to switch to a different head-mounted camera during data collection (see Head-mounted camera).
 16-month-olds tolerated the head-mounted camera less well than younger infants, leading to a higher exclusion rate.
@@
 To ensure that detections across these different cameras were comparable, we excluded detections from the outer edges of the videos taken with the MD-80 cameras based on these view angle differences in our main analyses (i.e.
-excluding top/bottom 13% of the frames, and left/right 10% of the frames).
+excluding top/bottom 13% of the frames, and left/right 10% of the frames), and the camera that was used was included as a covariate in our generalized linear mixed effect models.
 
@@
 As we were primarily interested in the presence of faces in the child's field of view, we chose to orient the camera upwards to capture the entirety of the child's upper visual field where the child is likely to see adult faces, understanding that this decision limited our ability to detect hands (especially those of the child, which are typically found at the bottom of the visual field).
-We note that these limitations regarding field of view and camera angle affect all studies using this method, not only our own.
+We note that these limitations regarding field of view and camera angle affect all studies to date using this method, not only our own; future innovations in lightweight, wearable cameras may alleviate these field-of-view limitations.
 
@@
 The caregiver's orientation was characterized as being close, far, or behind the child (independent of distance).
-For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child.
+“Close” to the caregiver was defined as being within the caregiver’s reach in any direction; for the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child.
+When children were sitting in their caregiver’s lap, this was characterized as the caregiver being “behind” with the child sitting (instead of the child being carried);coding instructions accompany the repository for this dataset.
 All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012toward].
-Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations; similarily, times when the child was being carried or the caregivers were out of the frame were marked as uncodable for caregiver orientation.
-On average, posture or orientation was uncodable from 1-2 minutes of data in each child (seconds excluded from analysis for posture, M = `r round(NA_rates$mean_na_posture,0)`s, SD = `r round(NA_rates$sd_na_posture,0)`s; orientation; M = `r round(NA_rates$mean_na_posture,0)`s, SD = `r round(NA_rates$sd_na_posture,0)`s), and these rates did not vary substantially with the age of the child.
-To assess the reliability of these annotations, a second coder annotated videos from three different children to calculate Cohen's kappa (posture, $\kappa$ = .82; caregiver orientation, $\kappa$ = .73).
+Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations; similarly, times when the child was being carried or the caregivers were out of the frame were marked as uncodable for caregiver orientation.
+On average, posture or orientation was uncodable from 1-2 minutes of data in each child (seconds excluded from analysis for posture, M = `r round(NA_rates$mean_na_posture,0)`s, SD = `r round(NA_rates$sd_na_posture,0)`s; orientation; M = `r round(NA_rates$mean_na_orientation,0)`s, SD = `r round(NA_rates$sd_na_orientation,0)`s), and these rates did not vary substantially with the age of the child.
+To assess the reliability of these annotations, a second coder annotated videos from five different children to calculate Cohen's kappa (posture, $\kappa$ = .76; caregiver orientation, $\kappa$ = .65).
 
@@
 We also obtained full text transcriptions of the entire play sessions (with time stamps marking 10s intervals).
-While these full transcriptions are not used in the present analyses, they have been made publically avaliable for future research.
+While these full transcriptions are not used in the present analyses, they have been made publicly available for future research.
 
@@
 We further use the detections to examine how access to these social cues changes during naming events (e.g., do you see the [zem]?).
+Our main analyses were not pre-registered though they were driven by the hypotheses and findings in @frank2012measuring and @franchak2018see.
+In contrast, we did not have strong predictions regarding how access to social cues would change during naming, thus we consider these analyses completely exploratory.
+All data and code for all analyses are available at https://osf.io/bje2u/?view_only=3e2bdfbe2294422a9decca74313856f3.
 
@@
   summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
+```
+
+```{r include=FALSE, eval=FALSE}
+# perfWristsByCamera <- performanceWrists %>%
+#   left_join(ages_inc_subs, by = c("video" = "subid")) %>%
+#   group_by(detectorType, camera) %>%
+#   summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
+#   group_by(detectorType, camera) %>%
+#   summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
+# 
+# perfFacesByCamera <- performanceFaces %>%
+#   left_join(ages_inc_subs, by = c("video" = "subid")) %>%
+#   group_by(detectorType, camera) %>%
+#   summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
+#   group_by(detectorType, camera) %>%
+#   summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))
 
@@
 ages_inc_subs <- inc_subs %>%
-  select(age.at.test, age.grp, subid)
+  select(age.at.test, age.grp, subid, camera)
 
@@
   summarise(time = sum(dt, na.rm=TRUE)) %>%
-  mutate(prop.time = time/sum(time)) %>%
+  mutate(prop.time = time/sum(time), sum_time = sum(time)) %>%
   group_by(subid, age.grp, orientation) %>%
@@
   summarise(time = sum(dt, na.rm=TRUE)) %>%
-  mutate(prop.time = time/sum(time)) %>%
+  mutate(prop.time = time/sum(time), sum_time = sum(time)) %>%
   group_by(subid, age.grp, posture) %>%
@@
 )
+```
+
+```{r}
 
@@
   facet_grid(~posture, labeller = labeller(posture = posture_labels)) +
-  geom_smooth(span=5, alpha=.2) +
+  geom_smooth(method = "glm", method.args = list(family = "binomial"), formula = y ~ x, mapping = aes(weight = time)) +
   scale_x_continuous(breaks=c(8,12,16)) +
@@
   facet_grid(~orientation,  labeller = labeller(orientation= orientation_labels)) +
-  geom_smooth(span=10, alpha=.2) +
+    geom_smooth(method = "glm", method.args = list(family = "binomial"), formula = y ~ x, mapping = aes(weight = time)) +
   scale_x_continuous(breaks=c(8,12,16)) +
@@
 ```{r setupdetsByAgePlot}
+
 detectionsByAge <- dlong %>%
@@
   mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video
+
 ```
@@
 
+```{r include-FALSE}
+goldFaces <- ground_truth_faces %>%
+  filter(sample_type == 'Random') %>%
+  rename(subid = video) %>%
+  group_by(subid) %>%
+  summarise(prop_detected = mean(is_face), num_detections = length(is_face)) %>%
+  mutate(detectorType='Goldset-Faces') %>%
+  left_join(inc_subs %>% select(subid, age.at.test))
+
+goldWrists <- ground_truth_wrists %>%
+  filter(sample_type == 'Random') %>%
+  rename(subid = video) %>%
+  group_by(subid) %>%
+  summarise(prop_detected = mean(is_wrist), num_detections = length(is_wrist)) %>%
+  mutate(detectorType='Goldset-Wrists') %>%
+  left_join(inc_subs %>% select(subid, age.at.test))
+
+detectionsByAgeWithGold <- detectionsByAge %>%
+  full_join(goldFaces) %>%
+  full_join(goldWrists) %>%
+  mutate(detectorType = factor(detectorType, levels = c('OpenPose-Faces','Goldset-Faces','OpenPose-Wrists','Goldset-Wrists')))
+
+# ggplot(detectionsByAgeWithGold, aes(x = age.at.test, y = prop_detected, color = detectorType, size=num_detections)) +
+#   theme_few(base_size=plot_size) + 
+#   geom_point(alpha=.5) +
+#   geom_smooth(size=1, alpha=.3, span=10) +
+#   scale_x_continuous(breaks=c(8,12,16)) +
+#   scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
+#   scale_size_continuous(name = "Frames") +
+#   labs(y = "Proportion detected", x = element_blank()) +
+#   scale_colour_manual(name = element_blank(), values = c('navy','blue','dark red','red')) +
+#   theme(legend.position="none") +
+#   # ggthemes::scale_color_solarized(name = "") + 
+#   facet_grid(~detectorType)
+
+# ggsave('figure_gen/appendix_figure_1.png', height=3, width=6, units='in')
+```
+
 ```{r setupdetByPosOrientPlot}
@@
 ages_inc_subs <- inc_subs %>%
-  select(subid, age.at.test, age.grp)
+  select(subid, age.at.test, age.grp, camera)
 
@@
   mutate(subid = as.factor(subid)) %>%
+  left_join(ages_inc_subs) %>% # get camera back
   group_by(subid, posture, orientation) %>%
-  summarize(countFaces=sum(faceOP==TRUE), countNotFaces = sum(faceOP==FALSE), age.at.test=age.at.test[1]) %>%
+  summarize(countFaces=sum(faceOP==TRUE), countNotFaces = sum(faceOP==FALSE), age.at.test=age.at.test[1], camera=camera[1]) %>%
   ungroup()
@@
   mutate(subid = as.factor(subid)) %>%
+  left_join(ages_inc_subs) %>% # get camera back
   group_by(subid, posture, orientation) %>%
-  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
+  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1], camera=camera[1]) %>%
   ungroup()
@@
 # Age as fixed effect, full random effects
-faceGLM_age_only <- glmer(cbind(countFaces,countNotFaces) ~  scale(age.at.test) +
+faceGLM_age_only <- glmer(cbind(countFaces,countNotFaces) ~  scale(age.at.test) + camera +
 (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
@@
 # Age and posture and orientation as fixed effects, full random effects
-faceGLM_no_int <- glmer(cbind(countFaces,countNotFaces) ~ posture + orientation + scale(age.at.test) +  (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+faceGLM_no_int <- glmer(cbind(countFaces,countNotFaces) ~ posture + orientation + camera + scale(age.at.test) +  (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
 
 # Age and posture * orientation as fixed effects, full random effects
-faceGLM_full <- glmer(cbind(countFaces,countNotFaces) ~ posture * orientation + scale(age.at.test) +  (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+faceGLM_full <- glmer(cbind(countFaces,countNotFaces) ~ posture * orientation +   scale(age.at.test) + camera + (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
 
@@
 # Age as fixed effect, full random effects
-wristGLM_age_only <- glmer(cbind(countWrists,countNotWrists) ~ scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
-
-wristGLM_no_int <- glmer(cbind(countWrists,countNotWrists) ~ orientation + posture + scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+wristGLM_age_only <- glmer(cbind(countWrists,countNotWrists) ~ scale(age.at.test) + camera + (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+
+wristGLM_no_int <- glmer(cbind(countWrists,countNotWrists) ~ orientation + posture + scale(age.at.test) + camera + (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
 
 # Age and posture and orientation as fixed effects, full random effects
-wristGLM_full <- glmer(cbind(countWrists,countNotWrists) ~ orientation * posture + scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
+wristGLM_full <- glmer(cbind(countWrists,countNotWrists) ~ posture * orientation + scale(age.at.test) + camera + (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
 ```
@@
 ## output
-cleanedNames_Int = c("Intercept","Sit","Stand","Close","Far","Age (Scaled)", "Sit*Close", "Stand*Close","Sit*Far","Stand*Far")
+cleanedNames_Int = c("Intercept","Sit","Stand","Close","Far","Age (Scaled)", "Camera Model", "Sit*Close", "Stand*Close","Sit*Far","Stand*Far")
 
@@
 When caregivers were behind their infants, supporting their infants' sitting or standing positions, infants saw fewer faces.
-When caregivers were relatively close to their infants, infants who were sitting or standing had more faces in view (Face detections; infant sitting and caregiver (CG) close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4])`; infant standing and CG close, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4])`) than infants who were lying down/crawling (i.e.
+When caregivers were relatively close to their infants, infants who were sitting or standing had more faces in view (Face detections; infant sitting and caregiver (CG) close, _b_ =  `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, _SE_  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, _Z_ = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4])`; infant standing and CG close, _b_ =  `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, _SE_  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, _Z_ = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4])`) than infants who were lying down/crawling (i.e.
 prone).
-When caregivers were far away from their infants, face detections were similarily higher (Face detections; infant sitting and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4])`; infant standing and CG far, b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`,`r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
-Infants' age was not significant predictor in accounting for the faces in view (Face detections; Age (scaled), b = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
+When caregivers were far away from their infants, face detections were similarily higher (Face detections; infant sitting and CG far, _b_ =  `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),1],2)`, _SE_  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),2],2)`, _Z_ = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Sit*Far'),4])`; infant standing and CG far, _b_ =  `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, _SE_  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, _Z_ = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`,`r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
+Infants' age was not significant predictor in accounting for the faces in view (Face detections; Age (scaled), _b_ =  `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, _SE_  = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, _Z_ = `r round(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, `r reportP(faceGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
 
 We found a similar pattern of results for wrist detections, even though there were fewer wrist detections overall in the dataset.
-Infants saw fewer wrists when caregivers were behind their infants, supporting their infants' sitting or standing positions vs when caregivers were relatively closer to their infants (Hand detections; infant sitting and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`,  `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4])`; infant standing and CG close, b=`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`,  `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4])`) than infants who were lying down/crawling (i.e.
+Infants saw fewer wrists when caregivers were behind their infants, supporting their infants' sitting or standing positions vs when caregivers were relatively closer to their infants (Hand detections; infant sitting and CG close, _b_ =`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),1],2)`, _SE_  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),3],2)`,  `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Sit*Close'),4])`; infant standing and CG close, _b_ =`r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),1],2)`, _SE_  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),2],2)`, _Z_ = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),3],2)`,  `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Close'),4])`) than infants who were lying down/crawling (i.e.
 prone).
-Wrist detections were  highest when caregivers were far away from their infants and those infants were standing (Wrist detections; infant standing and CG far, b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
-As with faces, age was not a significant predictor in these models (Wrist detections; Age (scaled), b = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, SE  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, Z = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
+Wrist detections were  highest when caregivers were far away from their infants and those infants were standing (Wrist detections; infant standing and CG far, _b_ =  `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),1],2)`, _SE_  = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),2],2)`, _Z_ = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Stand*Far'),4])`).
+As with faces, age was not a significant predictor in these models (Wrist detections; Age (scaled), _b_ =  `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),1],2)`, _SE_ =  `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),2],2)`, _Z_ = `r round(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),3],2)`, `r reportP(wristGLMOut$coefficients[which(cleanedNames_Int=='Age (Scaled)'),4])`).
 
@@
 Overall, these results suggest that infants' visual access to social information is largely modulated by their posture and orientation to their caregiver, which is in turn a function of their general locomotor development.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-Intercept & -3.24 & 0.20 & -16.37 & 0.000 \\ 
-  Sit & -0.00 & 0.19 & -0.02 & 0.988 \\ 
-  Stand & -0.29 & 0.20 & -1.44 & 0.151 \\ 
-  Close & 0.02 & 0.18 & 0.10 & 0.917 \\ 
-  Far & 0.51 & 0.25 & 2.04 & 0.041 \\ 
-  Age (Scaled) & 0.04 & 0.11 & 0.42 & 0.671 \\ 
-  Sit*Close & 0.90 & 0.07 & 13.80 & 0.000 \\ 
-  Stand*Close & 1.23 & 0.08 & 15.44 & 0.000 \\ 
-  Sit*Far & 0.63 & 0.07 & 9.16 & 0.000 \\ 
-  Stand*Far & 1.23 & 0.09 & 14.43 & 0.000 \\ 
-   \hline
-\end{tabular}
-\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.} 
-\end{table}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrr}
-  \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
-  \hline
-Intercept & -4.20 & 0.20 & -21.48 & 0.000 \\ 
-  Sit & 0.61 & 0.18 & 3.34 & 0.001 \\ 
-  Stand & 0.80 & 0.31 & 2.56 & 0.011 \\ 
-  Close & 0.37 & 0.20 & 1.87 & 0.061 \\ 
-  Far & 0.01 & 0.27 & 0.05 & 0.961 \\ 
-  Age (Scaled) & 0.09 & 0.11 & 0.86 & 0.390 \\ 
-  Sit*Close & 0.27 & 0.08 & 3.35 & 0.001 \\ 
-  Stand*Close & 0.53 & 0.09 & 6.07 & 0.000 \\ 
-  Sit*Far & 0.09 & 0.11 & 0.86 & 0.387 \\ 
-  Stand*Far & 1.57 & 0.11 & 14.16 & 0.000 \\ 
-   \hline
-\end{tabular}
-\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.} 
-\end{table}
 
@@
 These results were confirmed by a linear mixed-effect model with scaled aged as a fixed effect and random intercepts for each subject (Wrist detections; familiar objects vs.
-baseline, b = `r round(naming_wrists_model_out$coefficients[3,1],2)`, SE  = `r round(naming_wrists_model_out$coefficients[3,2],2)`, t = `r round(naming_wrists_model_out$coefficients[3,4],2)`, `r reportP(naming_wrists_model_out$coefficients[3,5])`; Novel objects vs.
-baseline, b = `r round(naming_wrists_model_out$coefficients[4,1],2)`, SE  = `r round(naming_wrists_model_out$coefficients[4,2],2)`, t = `r round(naming_wrists_model_out$coefficients[4,4],2)`, `r reportP(naming_wrists_model_out$coefficients[4,5])`).
+baseline, _b_ =  `r round(naming_wrists_model_out$coefficients[3,1],2)`, _SE_  = `r round(naming_wrists_model_out$coefficients[3,2],2)`, _t_ = `r round(naming_wrists_model_out$coefficients[3,4],2)`, `r reportP(naming_wrists_model_out$coefficients[3,5])`; Novel objects vs.
+baseline, _b_ =  `r round(naming_wrists_model_out$coefficients[4,1],2)`, _SE_  = `r round(naming_wrists_model_out$coefficients[4,2],2)`, _t_ = `r round(naming_wrists_model_out$coefficients[4,4],2)`, `r reportP(naming_wrists_model_out$coefficients[4,5])`).
 
@@
 In Study 1, we found that infant's in-the-moment posture changed with their age, as did infants' orientation relative their caregiver.
-In a related study with one-year-olds, @franchak2017see also found that infants' in-the-moment posture changed the proportion of time infants spent looking at faces.
+In a related study with one-year-olds, @franchak2018see also found that infants' in-the-moment posture changed the proportion of time infants spent looking at faces.
 Here, we sought to replicate these findings using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras [hosted on Databrary; @simon2015databrary] for two reasons.
@@
 ### Participants
-With the aid of @franchak2017see, we obtained the scene camera footage from the head-mounted eye-trackers for the 17 one-year-old infants (range 11.8–12.4 months) who participated.
-As noted in @franchak2017see, families were recruited from maternity wards of local hospitals in the New York City metropolitan area
-and were predominantly white and middle class.
+With the aid of @franchak2018see, we obtained the scene camera footage from the head-mounted eye-trackers for the 17 one-year-old infants (range 11.8–12.4 months) who participated.
+As noted in @franchak2018see, families were recruited from maternity wards of local hospitals in the New York City metropolitan area and were predominantly white and middle class.
 
 ### Head-mounted camera
-The view angle of the two head-mounted cameras used in these two studies were relatively similar ($52.2^{\circ}$ horizontal by $42.2^{\circ}$ in @franchak2017see, $47^{\circ}$ horizontal by $36^{\circ}$ vertical in Study 1).
+The view angle of the two head-mounted cameras used in these two studies were relatively similar ($52.2^{\circ}$ horizontal by $42.2^{\circ}$ in @franchak2018see, $47^{\circ}$ horizontal by $36^{\circ}$ vertical in Study 1).
 However, in Study 2 the camera was mounted on the side of the infants' head in Study 2, closer to infants temples, whereas in Study 1 the camera was situated in the middle of infants' forehead and oriented slightly upwards.
@@
 ### Video annotations
-The first five minutes of each of the videos were coded for the infants' posture (upright, prone, or sitting) by trained coders in @franchak2017see.
+The first five minutes of each of the videos were coded for the infants' posture (upright, prone, or sitting) by trained coders in @franchak2018see.
 These frame-by-frame posture annotations were synced with the outputs of the same automated annotation used in Study 1.
@@
 ### Differences between eye-tracking vs. automated detections
-First, we compared the overall proportion of frames in which infants foveated faces -- as assessed by the head-mounted eye-tracker --  in @franchak2017see vs.
+First, we compared the overall proportion of frames in which infants foveated faces -- as assessed by the head-mounted eye-tracker --  in @franchak2018see vs.
 the proportion of frames with faces detected by OpenPose.
 We expected some differences, as (1) head-mounted eye-trackers may underestimate the proportion of faces attended due to calibration issues, and (2) OpenPose may detect faces that are in view for infants but that infants may not be foveating.
-Across the entire session in @franchak2017see, infants looked at faces on 4.7% of frames.
+Across the entire session in @franchak2018see, infants looked at faces on 4.7% of frames.
 When we used all detections from OpenPose, we found a higher proportion of faces --  `r mean(detections_by_conf_by_sub$avg_faces_by_sub)*100`%.
-When we restricted our results to only high-confidence detections, we found `r mean(detections_by_conf_by_sub$avg_faces_by_sub_strict)*100`% of frames with faces, closer to the original values reported by @franchak2017see.
+When we restricted our results to only high-confidence detections, we found `r mean(detections_by_conf_by_sub$avg_faces_by_sub_strict)*100`% of frames with faces, closer to the original values reported by @franchak2018see.
 However, the above analyses on the accuracy of this method suggest that high-confidence detections dramatically underestimate the number of faces in view.
@@
 ### Replication & extension using automated detections
-Despite these differences, we found convergence between our two methodologies, replicating the main results from @franchak2017see, and finding that the proportion of faces detected was greater when infants were sitting or standing vs.
+Despite these differences, we found convergence between our two methodologies, replicating the main results from @franchak2018see, and finding that the proportion of faces detected was greater when infants were sitting or standing vs.
 prone (see Figure \ref{fig:franchak}).
@@
 These results were confirmed in generalized linear-mixed models, with random intercepts for each subjects and infants' posture as a fixed effect (Sitting vs.
-Prone, b = `r round(franchak_faces_out$coefficients[2,1],3)`, SE  = `r round(franchak_faces_out$coefficients[2,2],3)`, Z = `r round(franchak_faces_out$coefficients[2,3],3)`, `r reportP(franchak_faces_out$coefficients[2,4])`.
+Prone, _b_ =  `r round(franchak_faces_out$coefficients[2,1],3)`, SE  = `r round(franchak_faces_out$coefficients[2,2],3)`, Z = `r round(franchak_faces_out$coefficients[2,3],3)`, `r reportP(franchak_faces_out$coefficients[2,4])`.
 Upright vs.
-Prone, b = `r round(franchak_faces_out$coefficients[3,1],3)`, SE  = `r round(franchak_faces_out$coefficients[3,2],3)`, Z = `r round(franchak_faces_out$coefficients[3,3],3)`, `r reportP(franchak_faces_out$coefficients[3,4])`).
-
-We also found that infants' in-the-moment posture modulated the proportion of hands that were in view (i.e., wrist detections), though these were not originally annotated in @franchak2017see (see Figure \ref{fig:franchak}, proportion of frames with wrist detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_hands[1])`; Sitting: M = `r mean(dets_by_conf_by_posture$avg_hands[2])`; Upright:  M = `r mean(dets_by_conf_by_posture$avg_hands[3])`).
+Prone, _b_ =  `r round(franchak_faces_out$coefficients[3,1],3)`, SE  = `r round(franchak_faces_out$coefficients[3,2],3)`, Z = `r round(franchak_faces_out$coefficients[3,3],3)`, `r reportP(franchak_faces_out$coefficients[3,4])`).
+
+We also found that infants' in-the-moment posture modulated the proportion of hands that were in view (i.e., wrist detections), though these were not originally annotated in @franchak2018see (see Figure \ref{fig:franchak}, proportion of frames with wrist detections; Prone: M = `r mean(dets_by_conf_by_posture$avg_hands[1])`; Sitting: M = `r mean(dets_by_conf_by_posture$avg_hands[2])`; Upright:  M = `r mean(dets_by_conf_by_posture$avg_hands[3])`).
 These results were confirmed in generalized linear-mixed models with the same model structure as with faces (Sitting vs.
-Prone, b = `r round(franchak_wrists_out$coefficients[2,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[2,2],3)`, Z = `r round(franchak_wrists_out$coefficients[2,3],4)`, `r reportP(franchak_wrists_out$coefficients[2,4])`.
+Prone, _b_ =  `r round(franchak_wrists_out$coefficients[2,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[2,2],3)`, Z = `r round(franchak_wrists_out$coefficients[2,3],4)`, `r reportP(franchak_wrists_out$coefficients[2,4])`.
 Upright vs.
-prone, b = `r round(franchak_wrists_out$coefficients[3,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[3,2],3)`, Z = `r round(franchak_wrists_out$coefficients[3,3],3)`, `r reportP(franchak_wrists_out$coefficients[3,4])`).
+prone, _b_ =  `r round(franchak_wrists_out$coefficients[3,1],3)`, SE  = `r round(franchak_wrists_out$coefficients[3,2],3)`, Z = `r round(franchak_wrists_out$coefficients[3,3],3)`, `r reportP(franchak_wrists_out$coefficients[3,4])`).
 
@@
 What social cues do infants see as they learn language, and how does infants' access to these cues change as they grow and start to locomote themselves?
-We explored this question using video data from head-mounted cameras from two naturalistic datasets of parent-child interactions: a cross-sectional database of play sessions from 8-16 month-olds with sets of novel and familiar toys, and database of head-mounted camera videos from one-year-olds who explored a large play area [@franchak2017see].
+We explored this question using video data from head-mounted cameras from two naturalistic datasets of parent-child interactions: a cross-sectional database of play sessions from 8-16 month-olds with sets of novel and familiar toys, and database of head-mounted camera videos from one-year-olds who explored a large play area [@franchak2018see].
 To analyze these datasets, we developed a novel method using a pose detection model to automate the annotation of the social information in the infant view, here operationalized as the presence of the faces and hands of their caregiver; these annotations were then synced with manual annotations of infants in-the-moment posture from third-person videos.
@@
 Thus, while imperfect, we suggest that OpenPose can be applied to infant egocentric videos for the extraction of the social information in the infant viewpoint, reducing the burden of manual annotations and promoting the reusability of rich video datasets for further analyses.
-The use of this automated methodology allowed us to easily annotate the entirety of our dataset -- additionally analyzing the social information around naming events -- and to re-analyze the data from @franchak2017see, replicating our findings in a very different kind of play session.
+The use of this automated methodology allowed us to easily annotate the entirety of our dataset -- additionally analyzing the social information around naming events -- and to re-analyze the data from @franchak2018see, replicating our findings in a very different kind of play session.
 Furthermore, future work may be able to fine-tune pose detectors for even better accuracy, leveraging human annotations of the faces and hands that infants see to adapt models for the infant view.
+
+Nonetheless, it is important to note that the use of automated methods does come with some drawbacks.
+While using OpenPose does allow the analysis of the much more data, it is possible that these coarser detections may obscure fine-grained differences that may only be seen with careful, manual annotations or the use of head-mounted eye-trackers.
+In addition, researchers often gain insight into the phenomena that they are studying by carefully annotating these kinds of data, leading to new intuitions or ideas for future research.
+For these reasons, going forward we suggest that these automated methods continue to be complemented and validated by manual annotations on a portion of the relevant datasets they are applied to.
 
@@
 crawling or prone.
-These same findings were recapitulated in a second dataset collected by @franchak2017see with one-year-olds: sitting and upright infants saw more faces -- and hands -- than infants who were prone.
+These same findings were recapitulated in a second dataset collected by @franchak2018see with one-year-olds: sitting and upright infants saw more faces -- and hands -- than infants who were prone.
 Motor development appears to modulate how infants experience their visual world and the social information in it.
@@
 While exploratory, our results also suggest that infants saw a greater proportion of hands around naming events, hinting that children may have been orienting towards their caregiver when they heard labels for objects.
-While this effect was not present for faces, other work [@yoshida2008; @yu2013joint; @yu2017hand], including @franchak2017see, has found that infants spend much more time looking at the toys vs.
+While this effect was not present for faces, other work [@yoshida2008; @yu2013joint; @yu2017hand], including @franchak2018see, has found that infants spend much more time looking at the toys vs.
 their caregiver's faces during these play sessions, and highlighted the importance of hand-following as a component of joint attention [@yu2017hand].
@@
 
+Overall, our results suggest children changing locomotor abilities substantially change the social information that children have access to as they are learning.
+These results join an literature highlighting children as active learners whose own abilities to act on the world are major factors in the social information they see (Xu, 2019).
+From this perspective, children are far from sponges that soak up combinations of statistical regularities and social cues in their environment: rather, children’s changing cognitive, linguistic, and motoric abilities modulates the kinds of social information that they experience.
+
 Importantly, however, all of these findings come from observational, in-lab datasets, posing important limits on their generalizability.
+Further, while the quantity of video data analyzed in these studies goes far beyond that of prior work, the data come from a relatively small number of children, and this sample was selected based on those who tolerated wearing the camera during the experimental session.
 Future work is needed to relate these  slices of experience captured within in-lab play sessions with infants' everyday experiences [@fausey2016; @clerkin2017].
+
 More broadly, though observational findings allow us to document developmental change and identify potential causal pathways, they cannot confirm them.
